### 2018-11-21 paper/ms.tex
Initial commit

\documentclass[twocolumn]{aastex62}

% \submitjournal{ApJ}

\shortauthors{Beane et al.}

\begin{document}

\title{Actions are Bad, Mmkay}

\correspondingauthor{Angus Beane}
\email{abeane@sas.upenn.edu}

\author{Angus Beane}
\affil{Center for Computational Astrophysics, Flatiron Institute,
162 5th Ave., New York, NY, 10010, USA}
\affil{Department of Physics \& Astronomy, University of Pennsylvania,
209 South 33rd St., Philadelphia, PA 19104, USA}

\author{others}

\begin{abstract}

In an axisymmetric potential the 6D phase space of stars' positions and velocities can be reduced to a 3D space of invariant actions. For this reason actions are commonly used to help interpret phase space. However, it is not expected that actions are invariant in detail for the Milky Way --- it's spiral arms, bar(s), and gas clouds are all non-axisymmetric features that should induce long-term evolution in action space. However, these same features will cause a star's true orbit to deviate slightly from its orbit in an axisymmetric potential. If an observer assumes an axisymmetric potential when computing actions, this will induce an artificial short-term phase dependence of the actions, which we interpret as systematic error. Using the potential of cosmological zoom-in simulations of Milky Way Mass galaxies from the FIRE collaboration (the Latte suite), we investigate the magnitude of this systematic error. We find XXX. Something something lower limit.

\end{abstract}

\keywords{actions}

\section{Introduction} \label{sec:intro}

% \bibliography{references}

\end{document}

### 2018-11-21 paper/ms.tex
Update paper

@@
 \shortauthors{Beane et al.}
+
+\newcommand{\Gus}[1]{\textcolor{red}{#1}}
+
+\newcommand{\Msun}{\text{M}_\odot}
+\newcommand{\pc}{\text{pc}}
+\newcommand{\kpc}{\text{kpc}}
+
+\newcommand{\mi}{\texttt{m12i}}
+\newcommand{\mf}{\texttt{m12f}}
+\newcommand{\mm}{\texttt{m12m}}
 
@@
 \affil{Center for Computational Astrophysics, Flatiron Institute,
-162 5th Ave., New York, NY, 10010, USA}
+162 5th Ave., New York, NY 10010, USA}
 \affil{Department of Physics \& Astronomy, University of Pennsylvania,
@@
 
-\author{others}
+\author{Many Helpful People}
+\noaffiliation
 
@@
 If an observer assumes an axisymmetric potential when computing actions, this will induce an artificial short-term phase dependence of the actions, which we interpret as systematic error.
-Using the potential of cosmological zoom-in simulations of Milky Way Mass galaxies from the FIRE collaboration (the Latte suite), we investigate the magnitude of this systematic error.
+Using the potential of cosmological zoom-in simulations of Milky Way mass galaxies from the FIRE collaboration (the Latte suite), we investigate the magnitude of this systematic error.
 We find XXX.
@@
 
-% \bibliography{references}
+We could have proceeded by taking an axisymmetric potential and adding analytic spiral arms, a bar, and gas clouds.
+However, in preliminary work with the FIRE galaxies we found that defining a consistent reference frame between snapshots was very difficult.
+An inaccuracy in the direction of the $z$-axis by a single degree induces a $140\,\pc$ change in the height of the midplane at $8.2\,\kpc$.
+In a cosmological context it is not apparent that there exists a consistent definition of the galactocentric frame, and we therefore proceed in a cosmological context to include this effect.
+
+We perform a set of three different simulations repeated for two of the Milky Way zoom-ins, \mi and \mf.\footnote{The number corresponds to the mass of the dark matter halo (here, $\sim 10^{12} \Msun$).
+The first letter stands for mass and the last letter is a unique identifier for the {\em initial conditions}.} The first assumes an axisymmetric potential
+
+\section{Methods} \label{sec:methods}
+We begin by
+
+\subsection{Simulated Galaxies} \label{ssec:fire}
+We wish to perform N-body integration within a realistic Galactic potential.
+In order to do this, we use the {\em Latte} suite of Milky Way-mass zoom-ins from the Feedback in Realistic Environments (FIRE) collaboration\footnote{\url{http://fire.northwestern.edu}} \citep{2016ApJ...827L..23W,2018MNRAS.480..800H}.
+In a zoom-in cosmological simulation a high resolution region of interest is embedded within a low resolution cosmological context (\Gus{cites}).
+In the case of {\em Latte}, a region that will evolve into a Milky Way-mass galaxy, identified in a preliminary dark matter only run, is increased in resolution using MUSIC.
+This allows one to achieve ultra-high resolution simulations of individual galaxies than possible in a full cosmological run.
+
+\section{Simulations}
+
+\bibliography{references}
 

### 2018-11-24 runs/axisymmetric/run_m12i.sh
Update oceanic and first run

@@ -5,21 +5,13 @@
 #SBATCH -p gpu
 #SBATCH --gres=gpu:1 -n 8 
 #SBATCH --ntasks-per-node=8
-#SBATCH --mem=200g
+#SBATCH --mem=150g
 #SBATCH --time=7-00:00:00
 #SBATCH --output=log/asmw_%j.log
 #SBATCH --constraint=p100
 pwd; hostname; date
 
 source /mnt/home/abeane/.bash_profile
-
-module load gcc
-module load cuda
-module load cudnn
-module load openmpi
-
-source /mnt/home/abeane/amuse_env/bin/activate
-export PYTHONPATH=
 
 echo $CUDA_VISIBLE_DEVICES
 
@@ -30,7 +22,6 @@ cd $dir
 
 export OMP_NUM_THREADS=1 #this prevents conflicts in parallel libraries
 
-##mpirun -np 4 python axisymmetric_milkyway.py 
-python -u oc_nbody.py options 
+python -u oc_nbody.py options_m12i 
 
 date

### 2018-11-24 runs/init.sh

@@ -1,0 +1,4 @@
+module load gcc openmpi lib/x264/stable-aaa9aa8 ffmpeg
+module load cuda cudnn
+source ~/amuse_env/bin/activate
+export PYTHONPATH=`pwd`/include:`pwd`

### 2018-11-24 runs/oceanic/gizmo_interface.py

@@ -33,7 +33,12 @@ class gizmo_interface(object):
         self._read_snapshots_()
 
         # find starting star
-        self._init_starting_star_()
+        if self.axisymmetric:
+            if self.axi_Rinit is None or self.axi_vcircfrac is None or self.axi_zinit is None:
+                self._init_starting_star_()
+            self._gen_axisymmetric_()
+        else:
+            self._init_starting_star_()
 
         self._init_kdtree_()
 
@@ -82,6 +87,38 @@ class gizmo_interface(object):
             self.first_snapshot.snapshot['time'] * 1000.0
 
         self._clean_Rmag_(self.first_snapshot)
+
+    def _gen_axisymmetric_(self):
+        import agama
+        potential_cache_file = self.cache_directory + '/potential_id'+str(self.snap_index)
+        potential_cache_file += '_' + self.sim_name + '_pot'
+        try:
+            self.potential = agama.Potential(file=potential_cache_file)
+        except:
+            star_position = self.first_snapshot['star'].prop('host.distance.principal')
+            gas_position = self.first_snapshot['gas'].prop('host.distance.principal')
+            dark_position = self.first_snapshot['dark'].prop('host.distance.principal')
+
+            star_mass = self.first_snapshot['star']['mass']
+            gas_mass = self.first_snapshot['gas']['mass']
+            dark_mass = self.first_snapshot['dark']['mass']
+
+            position = np.concatenate((star_position, gas_position))
+            mass = np.concatenate((star_mass, gas_mass))
+
+            #TODO make these user-controllable
+            self.pdark = agama.Potential(type="Multipole",
+                                        particles=(dark_position, dark_mass),
+                                        symmetry='a', gridsizeR=20, lmax=2)
+            self.pbar = agama.Potential(type="CylSpline",
+                                        particles=(position, mass),
+                                        symmetry='a', gridsizer=20, gridsizez=20,
+                                        mmax=0, Rmin=0.2,
+                                        Rmax=50, Zmin=0.02, Zmax=10)
+            self.potential = agama.Potential(self.pdark, self.pbar)
+            self.potential.export(potential_cache_file)
+        return None
+
 
     def _clean_Rmag_(self, snap):
         # cleans out all particles greater than Rmag from galactic center

### 2018-11-24 runs/oceanic/oc_nbody.py

@@ -32,16 +32,29 @@ def evolve_cluster_in_galaxy(options_file):
         import gala.dynamics as gd
         import gala.potential as gp
 
-        pos = [opt.options['axi_Rinit'], 0, 0] * u.kpc
-        vel = [0, 0, 0] * u.km/u.s
-        mw = gp.MilkyWayPotential()
-        phase = gd.PhaseSpacePosition(pos, vel)
-        vc = mw.circular_velocity(phase).to_value(u.km/u.s) | units.kms
+        if opt.options['axi_Rinit'] is None \
+            or opt.options['axi_Rinit'] is None \
+            or opt.options['axi_zinit'] is None:
+            
+            pos = galaxy_code.chosen_position_z0
+            vel = galaxy_code.chosen_velocity_z0
+            stars.x += pos[0] | units.kpc
+            stars.y += pos[1] | units.kpc
+            stars.z += pos[2] | units.kpc
+            stars.vx += vel[0] | units.kms
+            stars.vy += vel[1] | units.kms
+            stars.vz += vel[2] | units.kms
+        else:
+            pos = [opt.options['axi_Rinit'], 0, 0] * u.kpc
+            vel = [0, 0, 0] * u.km/u.s
+            mw = gp.MilkyWayPotential()
+            phase = gd.PhaseSpacePosition(pos, vel)
+            vc = mw.circular_velocity(phase).to_value(u.km/u.s) | units.kms
 
-        stars = cluster_code.particles.copy()
-        stars.x += opt.options['axi_Rinit'] | units.kpc
-        stars.vy += vc
-        stars.z += opt.options['axi_zinit'] | units.kpc
+            stars = cluster_code.particles.copy()
+            stars.x += opt.options['axi_Rinit'] | units.kpc
+            stars.vy += vc
+            stars.z += opt.options['axi_zinit'] | units.kpc
 
     else:
         pos = galaxy_code.chosen_position_z0

### 2018-11-24 runs/oceanic/options.py

@@ -41,12 +41,14 @@ class options_reader(object):
         self._read_optional_option_('force_calculation', 'axisymmetric', 'false')
         self.options['axisymmetric'] = self._convert_bool_(self.options['axisymmetric'])
         if self.options['axisymmetric']:
-            self._read_required_option_('force_calculation', 'axi_Rinit')
-            self._read_optional_option_('force_calculation', 'axi_vcircfrac', 1.0)
-            self._read_optional_option_('force_calculation', 'axi_zinit', 0)
-            self.options['axi_Rinit'] = float(self.options['axi_Rinit'])
-            self.options['axi_vcircfrac'] = float(self.options['axi_vcircfrac'])
-            self.options['axi_zinit'] = float(self.options['axi_zinit'])
+            self._read_optional_option_('force_calculation', 'axi_Rinit', None)
+            self._read_optional_option_('force_calculation', 'axi_vcircfrac', None)
+            self._read_optional_option_('force_calculation', 'axi_zinit', None)
+            if self.options['axi_Rinit'] is not None and self.options['axi_vcircfrac'] is not None \
+                and self.options['axi_zinit'] is not None:
+                self.options['axi_Rinit'] = float(self.options['axi_Rinit'])
+                self.options['axi_vcircfrac'] = float(self.options['axi_vcircfrac'])
+                self.options['axi_zinit'] = float(self.options['axi_zinit'])
 
         # read in cluster parameters
         for opt in ['N', 'W0', 'Rcluster', 'softening',
@@ -113,7 +115,6 @@ class options_reader(object):
             if opt in self.options.keys():
                 self.options[opt] = self._convert_bool_(self.options[opt])
 
-        self._convert_rbf_basis_(self.options['basis'])
         self._convert_nbodycode_(self.options['nbodycode'])
 
     def set_options(self, object):

### 2018-11-24 paper/ms.tex
Update paper

@@
 We perform a set of three different simulations repeated for two of the Milky Way zoom-ins, \mi and \mf.\footnote{The number corresponds to the mass of the dark matter halo (here, $\sim 10^{12} \Msun$).
-The first letter stands for mass and the last letter is a unique identifier for the {\em initial conditions}.} The first assumes an axisymmetric potential
+The first ``m'' stands for mass and the last letter is a unique identifier for the {\em initial conditions}.} The first assumes an axisymmetric potential and also implicitly a well-defined galactocentric frame.
+This first simulation shows that our integration technique respects the fact that in an axisymmetric potential actions are invariant and that our calculation of the actions does not imprint some artificial phase dependence in the actions.
 
@@
 We begin by
+
+Rockstar import in gizmo_analysis is commented out
 

### 2018-11-26 runs/oceanic/oceanic_io.py
Create out file option

@@ -29,11 +29,11 @@ class snapshot_reader(object):
         self.frames = meta_array(np.append(self.frames, f), **self.frames.meta)
         # self.frames.append(self._grab_frame_(system, galaxy_code, time))
         if np.mod(i, self.write_frequency) == 0:
-            fout = self.output_directory+'/cluster_snapshots.p'
+            fout = self.output_directory+'/' + self.out_name
             dill.dump(self.frames, open(fout, 'wb'))
 
     def finish_sim(self):
-        fout = self.output_directory+'/cluster_snapshots.p'
+        fout = self.output_directory+'/' + self.out_name
         dill.dump(self.frames, open(fout, 'wb'))
 
     def _grab_frame_(self, system, galaxy_code, com, time):

### 2018-11-26 runs/oceanic/options.py

@@ -20,6 +20,7 @@ class options_reader(object):
         self._read_optional_option_('general', 'gpu_enabled', 'false')
         self._read_optional_option_('general', 'ngpu', '1')
         self._read_optional_option_('general', 'write_frequency', '10')
+        self._read_optional_option_('general', 'out_file', 'cluster_snapshots.p')
 
         # read in simulation parameters
         for opt in ['simulation_directory', 'cache_directory', 'snap_index',

### 2018-11-26 pattern_speed/dump_histogram.py
Update histogram dump to use ananke pa

@@ -1,23 +1,32 @@
 import numpy as np
 import gizmo_analysis as gizmo
+from tqdm import tqdm
 
-start = 591
-end = 600
+start = 590
+end = 600 
 snap_ids = list(range(start, end+1))
+
+pa = np.array([[-0.116813976049, 0.981662058456, -0.150645603947],
+               [-0.860269343337, -0.024217138031, 0.509264358796],
+               [-0.496277293369, -0.189084989395, -0.847322674588]])
 
 rmin = 7.2
 rmax = 9.2
 zmax = 1.0
 
-for gal in ['m12i', 'm12f']:
+#for gal in ['m12i', 'm12f']:
+for gal in ['m12i']:
     sim_directory = '/mnt/ceph/users/firesims/fire2/metaldiff/' + gal + '_res7100/'
     snaps = gizmo.io.Read.read_snapshots(['star', 'gas', 'dark'], 'index', snap_ids,
                                          properties=['position', 'mass', 'velocity',
                                                      'form.scalefactor'],
                                          assign_principal_axes=True,
                                          simulation_directory=sim_directory)
-    for idx, snap in zip(snap_ids, snaps):
+    for idx, snap in tqdm(zip(snap_ids, snaps)):
+        #snap.principal_axes_vectors = snaps[0].principal_axes_vectors
+        snap.principal_axes_vectors = pa 
         for k in snap.keys():
+            snap[k].principal_axes_vectors = pa
             fout = k + '_' + gal + '_id' + str(idx) + '.npy'
             cyl_pos = snap[k].prop('host.distance.principal.cylindrical')
 

### 2018-11-26 paper/ms.tex
Update line wrapping

@@
 \author{Angus Beane}
-\affil{Center for Computational Astrophysics, Flatiron Institute,
-162 5th Ave., New York, NY 10010, USA}
-\affil{Department of Physics \& Astronomy, University of Pennsylvania,
-209 South 33rd St., Philadelphia, PA 19104, USA}
+\affil{Center for Computational Astrophysics, Flatiron Institute, 162 5th
+Ave., New York, NY 10010, USA}
+\affil{Department of Physics \& Astronomy, University of Pennsylvania, 209
+South 33rd St., Philadelphia, PA 19104, USA}
 
@@
 
-In an axisymmetric potential the 6D phase space of stars' positions and velocities can be reduced to a 3D space of invariant actions.
-For this reason actions are commonly used to help interpret phase space.
-However, it is not expected that actions are invariant in detail for the Milky Way --- it's spiral arms, bar(s), and gas clouds are all non-axisymmetric features that should induce long-term evolution in action space.
-However, these same features will cause a star's true orbit to deviate slightly from its orbit in an axisymmetric potential.
-If an observer assumes an axisymmetric potential when computing actions, this will induce an artificial short-term phase dependence of the actions, which we interpret as systematic error.
-Using the potential of cosmological zoom-in simulations of Milky Way mass galaxies from the FIRE collaboration (the Latte suite), we investigate the magnitude of this systematic error.
-We find XXX.
-Something something lower limit.
+In an axisymmetric potential the 6D phase space of stars' positions and
+velocities can be reduced to a 3D space of invariant actions. For this reason
+actions are commonly used to help interpret phase space. However, it is not
+expected that actions are invariant in detail for the Milky Way --- it's
+spiral arms, bar(s), and gas clouds are all non-axisymmetric features that
+should induce long-term evolution in action space. However, these same
+features will cause a star's true orbit to deviate slightly from its orbit in
+an axisymmetric potential. If an observer assumes an axisymmetric potential
+when computing actions, this will induce an artificial short-term phase
+dependence of the actions, which we interpret as systematic error. Using the
+potential of cosmological zoom-in simulations of Milky Way mass galaxies from
+the FIRE collaboration (the Latte suite), we investigate the magnitude of
+this systematic error. We find XXX. Something something lower limit.
 
@@
 
-We could have proceeded by taking an axisymmetric potential and adding analytic spiral arms, a bar, and gas clouds.
-However, in preliminary work with the FIRE galaxies we found that defining a consistent reference frame between snapshots was very difficult.
-An inaccuracy in the direction of the $z$-axis by a single degree induces a $140\,\pc$ change in the height of the midplane at $8.2\,\kpc$.
-In a cosmological context it is not apparent that there exists a consistent definition of the galactocentric frame, and we therefore proceed in a cosmological context to include this effect.
+We could have proceeded by taking an axisymmetric potential and adding
+analytic spiral arms, a bar, and gas clouds. However, in preliminary work
+with the FIRE galaxies we found that defining a consistent reference frame
+between snapshots was very difficult. An inaccuracy in the direction of the
+$z$-axis by a single degree induces a $140\,\pc$ change in the height of the
+midplane at $8.2\,\kpc$. In a cosmological context it is not apparent that
+there exists a consistent definition of the galactocentric frame, and we
+therefore proceed in a cosmological context to include this effect.
 
-We perform a set of three different simulations repeated for two of the Milky Way zoom-ins, \mi and \mf.\footnote{The number corresponds to the mass of the dark matter halo (here, $\sim 10^{12} \Msun$).
-The first ``m'' stands for mass and the last letter is a unique identifier for the {\em initial conditions}.} The first assumes an axisymmetric potential and also implicitly a well-defined galactocentric frame.
-This first simulation shows that our integration technique respects the fact that in an axisymmetric potential actions are invariant and that our calculation of the actions does not imprint some artificial phase dependence in the actions.
+We perform a set of three different simulations repeated for two of the Milky
+Way zoom-ins, \mi and \mf.\footnote{The number corresponds to the mass of the
+dark matter halo (here, $\sim 10^{12} \Msun$). The first ``m'' stands for
+mass and the last letter is a unique identifier for the {\em initial
+conditions}.} The first assumes an axisymmetric potential and also implicitly
+a well-defined galactocentric frame. This first simulation shows that our
+integration technique respects the fact that in an axisymmetric potential
+actions are invariant and that our calculation of the actions does not
+imprint some artificial phase dependence in the actions.
 
@@
 We wish to perform N-body integration within a realistic Galactic potential.
-In order to do this, we use the {\em Latte} suite of Milky Way-mass zoom-ins from the Feedback in Realistic Environments (FIRE) collaboration\footnote{\url{http://fire.northwestern.edu}} \citep{2016ApJ...827L..23W,2018MNRAS.480..800H}.
-In a zoom-in cosmological simulation a high resolution region of interest is embedded within a low resolution cosmological context (\Gus{cites}).
-In the case of {\em Latte}, a region that will evolve into a Milky Way-mass galaxy, identified in a preliminary dark matter only run, is increased in resolution using MUSIC.
-This allows one to achieve ultra-high resolution simulations of individual galaxies than possible in a full cosmological run.
+In order to do this, we use the {\em Latte} suite of Milky Way-mass zoom-ins
+from the Feedback in Realistic Environments (FIRE)
+collaboration\footnote{\url{http://fire.northwestern.edu}}
+\citep{2016ApJ...827L..23W,2018MNRAS.480..800H}. In a zoom-in cosmological
+simulation a high resolution region of interest is embedded within a low
+resolution cosmological context (\Gus{cites}). In the case of {\em Latte}, a
+region that will evolve into a Milky Way-mass galaxy, identified in a
+preliminary dark matter only run, is increased in resolution using MUSIC.
+This allows one to achieve ultra-high resolution simulations of individual
+galaxies than possible in a full cosmological run.
+
+Our centering differs from Sanderson et al. by $\sim3.2\,\pc$ for \mi.
 

### 2018-11-26 paper/ms.tex
Add pattern speed updates

@@
 
+\appendix \section{Pattern Speed} Our estimation of the pattern speed is very
+crude, but throughout the text we have adopted three different speeds in
+order to demonstrate that qualitatively our results do not depend on the
+precise numerical value of the pattern speed. We identify the spiral arms by
+first plotting a $\phi$-$R$ 2D histogram, shown in
+Figure~\ref{fig:pattern_speed}. In order to compute $\phi$ and $R$, we use the principal axes from \citep{2018arXiv180610564S}.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{{pattern_speed_by_eye}.png}
+\caption{Pattern speed.}
+\label{fig:pattern_speed}
+\end{figure}
+
 \bibliography{references}

### 2018-12-01 runs/schmactions/actions/compute_actions.py
Add schmactions action computations

@@ -1,0 +1,15 @@
+from oceanic.analysis import snapshot_action_calculator
+from oceanic.options import options_reader
+import dill
+
+opt = options_reader('../options_m12i')
+ac = snapshot_action_calculator(opt, snapshot_file='cluster_snapshots_m12i.p')
+ac.all_actions(fileout='cluster_snapshots_m12i_actions.p')
+
+opt = options_reader('../options_m12i_slow')
+ac = snapshot_action_calculator(opt, snapshot_file='cluster_snapshots_m12i_slow.p')
+ac.all_actions(fileout='cluster_snapshots_m12i_slow_actions.p')
+
+opt = options_reader('../options_m12i_fast')
+ac = snapshot_action_calculator(opt, snapshot_file='cluster_snapshots_m12i_fast.p')
+ac.all_actions(fileout='cluster_snapshots_m12i_fast_actions.p')

### 2018-12-01 runs/schmactions/pa_rotate/init.sh
Prelim pa rotation script

@@ -1,0 +1,1 @@
+../runs/init.sh

### 2018-12-01 runs/schmactions/pa_rotate/rotate_pa.py

@@ -1,0 +1,81 @@
+import numpy as np
+from scipy.optimize import minimize
+from tqdm import tqdm
+
+from oceanic.analysis import snapshot_action_calculator
+from oceanic.options import options_reader
+
+import dill
+import utilities as ut
+
+from scipy.optimize import minimize
+
+gal_info = np.genfromtxt('../m12i_info.txt', comments='#', delimiter=',')
+
+cen = gal_info[0]
+vel = gal_info[1]
+pa = gal_info[2:]
+
+def euler_matrix(theta) :
+    theta = np.array(theta)
+    if len(theta)==2:
+        theta = np.hstack((theta, 0))
+    R_x = np.array([[1,         0,                  0  ] ,                
+                    [0,         np.cos(theta[0]), -np.sin(theta[0]) ],
+                    [0,         np.sin(theta[0]), np.cos(theta[0])  ]
+                    ])
+    R_y = np.array([[np.cos(theta[1]),    0,      np.sin(theta[1])  ],
+                    [0,                     1,      0 ] ,                
+                    [-np.sin(theta[1]),   0,      np.cos(theta[1])  ]
+                    ])
+    R_z = np.array([[np.cos(theta[2]),    -np.sin(theta[2]),    0],
+                    [np.sin(theta[2]),    np.cos(theta[2]),     0],
+                    [0,                     0,                      1]
+                    ])
+    R = np.dot(R_x, np.dot(R_y, R_z))
+    
+    # this is the real one
+    # R = np.dot(R_z, np.dot( R_y, R_x ))
+    return R
+
+def euler_rotate(theta, vec):
+    mat = euler_matrix(theta)
+    return np.transpose(np.tensordot(mat, np.transpose(vec), axes=1))
+
+opt = options_reader('../options_m12i')
+ac = snapshot_action_calculator(opt, snapshot_file='cluster_snapshots_m12i.p')
+
+cluster = dill.load(open('cluster_snapshots_m12i.p', 'rb'))
+npart = len(cluster[0]['position'])
+nsnap = len(cluster)
+
+pos = np.array([cl['position'] for cl in cluster])
+vel = np.array([cl['velocity'] for cl in cluster])
+time = np.array([cl['time'] for cl in cluster])
+
+pos_flat = np.reshape(pos, (npart*nsnap, 3))
+vel_flat = np.reshape(vel, (npart*nsnap, 3))
[…]

### 2018-12-01 runs/axisymmetric/actions/compute_actions.py
Axisymmetric sim, maybe not useful

@@ -2,7 +2,7 @@ from oceanic.analysis import snapshot_action_calculator
 from oceanic.options import options_reader
 import dill
 
-opt = options_reader('../options')
+opt = options_reader('../options_m12i')
 #ac = snapshot_action_calculator(opt)
 ac = snapshot_action_calculator(opt, snapshot_file='cluster_snapshots.p')
 

### 2018-12-01 runs/axisymmetric/run_m12i.sh

@@ -8,7 +8,7 @@
 #SBATCH --mem=150g
 #SBATCH --time=7-00:00:00
 #SBATCH --output=log/asmw_%j.log
-#SBATCH --constraint=p100
+#SBATCH --constraint=v100
 pwd; hostname; date
 
 source /mnt/home/abeane/.bash_profile

### 2018-12-01 paper/ms.tex
Update README, fix paper
Test CI
Update pattern speed image
Test pattern speed image
Try bibtex
Update paper

@@
 \shortauthors{Beane et al.}
+
+\usepackage{graphicx}
 
@@
 
-\title{Actions are Bad, Mmkay}
+\title{Actions Computed in the Milky Way are Not Invariant}
 
@@
 velocities can be reduced to a 3D space of invariant actions. For this reason
-actions are commonly used to help interpret phase space. However, it is not
-expected that actions are invariant in detail for the Milky Way --- it's
-spiral arms, bar(s), and gas clouds are all non-axisymmetric features that
-should induce long-term evolution in action space. However, these same
-features will cause a star's true orbit to deviate slightly from its orbit in
-an axisymmetric potential. If an observer assumes an axisymmetric potential
-when computing actions, this will induce an artificial short-term phase
-dependence of the actions, which we interpret as systematic error. Using the
-potential of cosmological zoom-in simulations of Milky Way mass galaxies from
-the FIRE collaboration (the Latte suite), we investigate the magnitude of
-this systematic error. We find XXX. Something something lower limit.
+actions are commonly used to help interpret phase space. We show that if one
+assumes the wrong galactocentric reference frame this can induce artificial
+non-axisymmetries which manifest as time dependence in the computed actions.
+However, the Milky Way is not axisymmetric in detail and so there are no
+truly axisymmetric axes. We instead define the  ``optimally axisymmetric''
+axes as the reference frame which minimizes the time dependence of the
+actions. Using cosmological zoom-in simulations from FIRE-2, we show that the
+optimally axisymmetric reference frame can differ from standard methods of
+calculating the galactocentric reference frame by (some numbers). We further
+show that if one uses the conventional method to define the Galactic midplane
+in this simulated galaxy, the midplane can vary by nearly a scale height
+across the galaxy. This work calls into question the use of actions as
+dynamical invariants unless these reference frame issues can be resolved.
+
+% However, it is not expected that actions are invariant in detail for the
+% Milky Way. This in-invariance can come about from assuming the wrong
+% Galactic potential, but it can also come about due to true
+% non-axisymmetries. We show that actions can gain artificial time-dependence
+% also by assuming the wrong reference frame. We argue that standard methods
+% for determining the galactocentric reference frame are not gu
+
+% it's spiral arms, bar(s), and gas clouds are all non-axisymmetric features
+% that should induce long-term evolution in action space. However, these same
+% features will cause a star's true orbit to deviate slightly from its orbit
+% in an axisymmetric potential. If an observer assumes an axisymmetric
+% potential when computing actions, this will induce an artificial short-term
+% phase dependence of the actions, which we interpret as systematic error.
+% Using the potential of cosmological zoom-in simulations of Milky Way mass
+% galaxies from the FIRE collaboration (the Latte suite), we investigate the
+% magnitude of this systematic error. We find XXX. Something something lower
+% limit.
 
@@
 \section{Introduction} \label{sec:intro}
+
+Midplane resolved with $\sim 370$ to $570$ star particles.
 
@@
 
-Rockstar import in gizmo_analysis is commented out
+Rockstar import in gizmo\_analysis is commented out
 
@@
 first plotting a $\phi$-$R$ 2D histogram, shown in
-Figure~\ref{fig:pattern_speed}. In order to compute $\phi$ and $R$, we use the principal axes from \citep{2018arXiv180610564S}.
+Figure~\ref{fig:pattern_speed}. In order to compute $\phi$ and $R$, we use the principal axes from \citep{2018arXiv180610564S}. Push push
 
 \begin{figure}
-\includegraphics[width=\columnwidth]{{pattern_speed_by_eye}.png}
+\centering
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{{pattern_speed_by_eye}.png}
 \caption{Pattern speed.}

### 2018-12-03 paper/ms.tex
Update paper

@@
 
-\title{Actions Computed in the Milky Way are Not Invariant}
+\title{The Actions We Compute are Not Invariant}
 
@@
 
-\section{Simulations}
+\section{Optimally Axisymmetric Frame}
+An analytic, axisymmetric potential with a known reference frame is a far too
+simple description of the Milky Way. We thus turn to cosmological zoom-in
+simulations as approximations to a ``true'' Milky Way potential. It is important to remember that
+
+\subsection{Cosmological Zoom-in} \label{ssec:cosmozoom}
+Cosmological zoom-ins allow one to simulate a selected region at high resolution embedded in a low-resolution cosmological background \citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}
+
+\subsection{Cluster Simulations} \label{ssec:cluster_sim}
 
@@
 first plotting a $\phi$-$R$ 2D histogram, shown in
-Figure~\ref{fig:pattern_speed}. In order to compute $\phi$ and $R$, we use the principal axes from \citep{2018arXiv180610564S}. Push push
+Figure~\ref{fig:pattern_speed}. In order to compute $\phi$ and $R$, we use
+the principal axes from \citep{2018arXiv180610564S}. Push push
 

### 2018-12-03 runs/oceanic_schmactions/rotate_pa.py
Add minimize to rotate pa

@@ -1,5 +1,9 @@
 import gizmo_analysis as gizmo
 import numpy as np
+import agama
+import sys
+import utilities as ut
+from scipy.optimize import minimize
 
 np.random.seed(162)
 agama.setUnits(mass=1, length=1, velocity=1)
@@ -82,23 +86,25 @@ vel = np.load(velfile)
 nframes, npart, ndim = np.shape(pos)
 
 def actions(pos, vel):
-	flat_pos = np.reshape(pos, (nframes*npart, ndim))
-	flat_vel = np.reshape(vel, (nframes*npart, ndim))
-	phase = np.c_[pos, vel]
-	act = af(phase)
-	return np.reshape(act, (nframes, npart, ndim))
+    flat_pos = np.reshape(pos, (nframes*npart, ndim))
+    flat_vel = np.reshape(vel, (nframes*npart, ndim))
+    phase = np.c_[flat_pos, flat_vel]
+    act = af(phase)
+    return np.reshape(act, (nframes, npart, ndim))
 
 def chisq(x):
-	offset = x[:3]
-	theta = x[3:]
+    offset = x[:3]
+    theta = x[3:]
 
-	rotated_pa = euler_rotate(theta, fiducial_pa)
+    rotated_pa = euler_rotate(theta, fiducial_pa)
 
-	new_pos = get_position(snap, pos, offset, rotated_pa)
-	new_vel = get_position(snap, vel, np.array([0,0,0]), rotated_pa)
+    new_pos = get_position(snap, pos, offset, rotated_pa)
+    new_vel = get_position(snap, vel, np.array([0,0,0]), rotated_pa)
 
-	act = actions(new_pos, new_vel)
-	act_collapse = np.median(act, axis=1)
-	act_med = np.median(act_collapse, axis=0)
-	act_collapse = np.subtract(act_collapse, act_med)
-	return np.sum(np.square(act_collapse))
+    act = actions(new_pos, new_vel)
+    act_collapse = np.median(act, axis=1)
+    act_med = np.median(act_collapse, axis=0)
+    act_collapse = np.subtract(act_collapse, act_med)
+    return np.sum(np.square(act_collapse))
+
+res = minimize(chisq, np.array([0, 0, 0, 0, 0, 0]))

### 2018-12-03 paper/ms.tex
Update paper

@@
 \newcommand{\kpc}{\text{kpc}}
+
+\newcommand{\abs}[1]{\left| #1 \right|}
+\newcommand{\z}{z}
 
@@
 simple description of the Milky Way. We thus turn to cosmological zoom-in
-simulations as approximations to a ``true'' Milky Way potential. It is important to remember that
+simulations as approximations to a ``true'' Milky Way potential. We will perform a few simulations of small clusters in order to
 
 \subsection{Cosmological Zoom-in} \label{ssec:cosmozoom}
-Cosmological zoom-ins allow one to simulate a selected region at high resolution embedded in a low-resolution cosmological background \citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}
+Cosmological zoom-ins allow one to simulate a selected region at high
+resolution embedded in a low-resolution cosmological background
+\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}. Here we use a
+zoom-in of a Milky Way-mass galaxy from the FIRE-2 collaboration.
 
-\subsection{Cluster Simulations} \label{ssec:cluster_sim}
+First, a dark matter only simulation was run with $\Lambda$CDM cosmology
+parameterized by $(\Omega_m, \Omega_{\Lambda}, \Omega_b, h, \sigma_8, n_s) =
+(0.272, 0.728, 0.0455, 0.702, 0.807, 0.961)$, consistent with current
+constraints \citep{2018arXiv180706209P}. Halos are selected at $z=0$ based
+solely on their mass ($\sim 1\text{--}2 \times 10^{12} \Msun$). Particles
+within $5 R_{200\text{m}}$ of each halo are traced back to $\z=99$ and the
+initial conditions are regenerated at higher resolution using MUSIC
+\citep{2011MNRAS.415.2101H}. A total of $\sim 10$ halos were simulated at
+higher resolution \citep{2018MNRAS.481.4133G}.
 
-\appendix \section{Pattern Speed} Our estimation of the pattern speed is very
-crude, but throughout the text we have adopted three different speeds in
-order to demonstrate that qualitatively our results do not depend on the
-precise numerical value of the pattern speed. We identify the spiral arms by
-first plotting a $\phi$-$R$ 2D histogram, shown in
-Figure~\ref{fig:pattern_speed}. In order to compute $\phi$ and $R$, we use
-the principal axes from \citep{2018arXiv180610564S}. Push push
+These halos are uncontaminated and contain gas particles of mass $\sim 7000
+\text{--} 20000 \Msun$ and star particles of mass $\sim 5000 \text{--} 7000
+\Msun$, with the lower end coming from stellar evolution
+\citep{2018arXiv180610564S}. Softening lengths for dark matter and star
+particles are fixed at $112\,\pc$ and $11.2\,\pc$,
+respectively.\footnote{This is $2.8$ times the often-quoted
+Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
+median softening length for gas particles around roughly solar positions
+(with cylindrical radii within $500\,\pc$ of $8.2\,\kpc$ and
+$\abs{z}<1\,\kpc$) is $98\,\pc$.
 
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{{pattern_speed_by_eye}.png}
-\caption{Pattern speed.}
-\label{fig:pattern_speed}
-\end{figure}
+We restrict ourselves to the simulation \texttt{m12i} \citep[first introduced
+in][]{2016ApJ...827L..23W}. For this work, this galaxy is the best-case
+scenario. It does not contain a bar (like \texttt{m12m}) or companion galaxy,
+and by eye is the closest to axisymmetric of the three galaxies presented in
+\citet{2018arXiv180610564S} \citep{2018MNRAS.481.4133G}.
+
+It is important to remember that simulations of Milky Way-mass galaxies are
+not perfect representations of the true Milky Way, as discussed in
+\citet{2018arXiv180610564S}. For instance, the velocity structure of
+\texttt{m12i} is closer to M31 than the Milky Way. However, in this work we
+are most interested in the global properties of the \texttt{m12i} potential,
+and specifically in deviations from axisymmetry. From this perspective,
+\texttt{m12i} is actually far more axisymmetric than we might expect of the
+Milky Way. While \texttt{m12i} has prominent spiral arms, it lacks a bar and
+a companion like the Large Magellanic Cloud. In future work, these additional
+complications also need to be addressed.
+
+\subsection{Cluster Simulations} \label{ssec:cluster_sim} We simulate a small
+star cluster of $256$ stars in the frozen $\z=0$ potential of
+$\texttt{m12i}$.
+
+We use the simple tree code
+\texttt{pykdgrav}\footnote{\url{https://github.com/omgspace/pykdgrav},
+commit:\texttt{fbd1ddc}}.
+
+We use the initial position and velocity of two star particles in
+$\texttt{m12i}$
+
+\appendix \section{Appendix 1}
+Gonna write an appendix probably
 

### 2018-12-04 paper/ms.tex
Update paper

@@
 
-\subsection{Cluster Simulations} \label{ssec:cluster_sim} We simulate a small
-star cluster of $256$ stars in the frozen $\z=0$ potential of
-$\texttt{m12i}$.
+\subsection{Cluster Simulations} \label{ssec:cluster_sim}
+We simulate a small star cluster of $256$ stars in the frozen $\z=0$
+potential of $\texttt{m12i}$. We use the code \texttt{ph4} within the
+\texttt{AMUSE} framework \citep{2011ascl.soft07007P, 2013CoPhC.184..456P,
+2013A&A...557A..84P} for the cluster integration. We use a softening length
+of $0.01\,\pc$ which means we do not correctly solve the internal dynamics of
+the cluster, but this is not important for our purposes. To couple the
+gravitational forces of the frozen snapshot to the cluster stars, we use the
+\texttt{BRIDGE} \citep{2007PASJ...59.1095F}, also implemented in
+\texttt{AMUSE}. Our simulation runs for $2\,\Gyr$, allowing us to solve for
+$\sim 10$ orbits.
 
-We use the simple tree code
+To compute the forces from the frozen potential, we use the simple tree code
 \texttt{pykdgrav}\footnote{\url{https://github.com/omgspace/pykdgrav},

### 2018-12-04 runs/oceanic_schmactions/rotate_pa.py
Dump new oa frame

@@ -12,6 +12,8 @@ snap_idx=600
 sim_dir = '/mnt/ceph/users/firesims/fire2/metaldiff/m12i_res7100/'
 gal_info = 'm12i_info.txt'
 
+cadence_list = [50, 20]
+
 snap = gizmo.io.Read.read_snapshots(['star', 'gas', 'dark'],
                                      'index', snap_idx,
                                      simulation_directory=sim_dir,
@@ -26,6 +28,8 @@ for k in snap.keys():
                  'principal_axes_vectors']:
         setattr(snap[k], attr, getattr(snap, attr))
 
+fiducial_center = ref[0]
+fiducial_velocity = ref[1]
 fiducial_pa = ref[2:]
 
 def get_position(snap, pos, center, pa=None):
@@ -80,10 +84,15 @@ af = agama.ActionFinder(potential, interp=False)
 # read in simulated positions and velocities
 posfile = 'pos_' + sys.argv[1] + '.npy'
 velfile = 'vel_' + sys.argv[1] + '.npy'
-pos = np.load(posfile)
-vel = np.load(velfile)
+pos_full = np.load(posfile)
+vel_full = np.load(velfile)
 
-nframes, npart, ndim = np.shape(pos)
+def init_minimizer(cadence):
+    global nframes, npart, ndim
+    pos = pos_full[::cadence]
+    vel = vel_full[::cadence]
+
+    nframes, npart, ndim = np.shape(pos)
 
 def actions(pos, vel):
     flat_pos = np.reshape(pos, (nframes*npart, ndim))
@@ -107,4 +116,18 @@ def chisq(x):
     act_collapse = np.subtract(act_collapse, act_med)
     return np.sum(np.square(act_collapse))
 
-res = minimize(chisq, np.array([0, 0, 0, 0, 0, 0]))
+xinit = np.array([0, 0, 0, 0, 0, 0])
+for cd in cadence_list:
+    init_minimizer(cd)
+    res = minimize(chisq, xinit, method='Nelder-Mead')
+    xinit = res.x
+
+offset = xinit[:3]
+theta = xinit[3:]
+
+oa_center_position = fiducial_center + offset
+oa_pa = euler_rotate(theta, fiducial_pa)
+oa_center_velocity = get_position(snap, fiducial_velocity, np.array([0, 0, 0]), oa_pa)
+
+new_frame = np.vstack((oa_center_position, oa_center_velocity, oa_pa))
+np.savetxt('oa_frame.txt', new_frame)

### 2018-12-04 paper/ms.tex
Fix paper
Update paper

@@
 \newcommand{\kpc}{\text{kpc}}
+\newcommand{\Myr}{\text{Myr}}
+\newcommand{\Gyr}{\text{Gyr}}
+\newcommand{\kms}{\text{km}/\text{s}}
 
@@
 \newcommand{\mm}{\texttt{m12m}}
+\newcommand{\gdr}{{\em Gaia} DR2}
 
@@
 
-\title{The Actions We Compute are Not Invariant}
+\title{Our Galactocentric Reference Frame Is Probably Wrong, and Not by a
+Little}
 
@@
 In an axisymmetric potential the 6D phase space of stars' positions and
-velocities can be reduced to a 3D space of invariant actions. For this reason
-actions are commonly used to help interpret phase space. We show that if one
-assumes the wrong galactocentric reference frame this can induce artificial
-non-axisymmetries which manifest as time dependence in the computed actions.
-However, the Milky Way is not axisymmetric in detail and so there are no
-truly axisymmetric axes. We instead define the  ``optimally axisymmetric''
-axes as the reference frame which minimizes the time dependence of the
-actions. Using cosmological zoom-in simulations from FIRE-2, we show that the
-optimally axisymmetric reference frame can differ from standard methods of
-calculating the galactocentric reference frame by (some numbers). We further
-show that if one uses the conventional method to define the Galactic midplane
-in this simulated galaxy, the midplane can vary by nearly a scale height
-across the galaxy. This work calls into question the use of actions as
-dynamical invariants unless these reference frame issues can be resolved.
+velocities can be reduced to a 3D space of invariant actions. We show that if
+one assumes the wrong galactocentric reference frame this can induce
+artificial non-axisymmetries which manifest as time dependence in the
+computed actions. However, the Milky Way is not axisymmetric in detail and so
+there are no truly axisymmetric axes. We instead define the  ``optimally
+axisymmetric'' axes as the reference frame which minimizes the time
+dependence of the computed actions. Standard methods for determining the
+galactocentric reference frame typically involve finding the median height of
+stars in the disk. Using cosmological zoom-in simulations from FIRE-2, we
+show that the optimally axisymmetric reference frame can differ from standard
+methods of calculating the galactocentric reference frame by nearly a scale
+height. This work calls into question the use of actions as dynamical
+invariants unless these reference frame issues can be resolved.
 
@@
 \section{Introduction} \label{sec:intro}
+Our understanding of the Milky Way is currently undergoing a revolution in the era of \gdr.
+Recent major discoveries include the remnants of a major merger \citep{2018Natur.563...85H}, a phase-space spiral (\Gus{cites}) indicating local substructure infall (\Gus{cites}), a gap indicative of a dark matter subtructure in GD1 \citep{2018arXiv181103631B}, and \Gus{others}.
 
-Midplane resolved with $\sim 370$ to $570$ star particles.
+\section{Reference Frame Gone Wrong} \label{sec:ref_frame}
+We begin by quantifying how
 
-We could have proceeded by taking an axisymmetric potential and adding
-analytic spiral arms, a bar, and gas clouds. However, in preliminary work
-with the FIRE galaxies we found that defining a consistent reference frame
-between snapshots was very difficult. An inaccuracy in the direction of the
-$z$-axis by a single degree induces a $140\,\pc$ change in the height of the
-midplane at $8.2\,\kpc$. In a cosmological context it is not apparent that
-there exists a consistent definition of the galactocentric frame, and we
-therefore proceed in a cosmological context to include this effect.
-
-We perform a set of three different simulations repeated for two of the Milky
-Way zoom-ins, \mi and \mf.\footnote{The number corresponds to the mass of the
-dark matter halo (here, $\sim 10^{12} \Msun$). The first ``m'' stands for
-mass and the last letter is a unique identifier for the {\em initial
-conditions}.} The first assumes an axisymmetric potential and also implicitly
-a well-defined galactocentric frame. This first simulation shows that our
-integration technique respects the fact that in an axisymmetric potential
-actions are invariant and that our calculation of the actions does not
-imprint some artificial phase dependence in the actions.
-
-\section{Methods} \label{sec:methods}
-We begin by
-
-Rockstar import in gizmo\_analysis is commented out
-
-\subsection{Simulated Galaxies} \label{ssec:fire}
-We wish to perform N-body integration within a realistic Galactic potential.
-In order to do this, we use the {\em Latte} suite of Milky Way-mass zoom-ins
-from the Feedback in Realistic Environments (FIRE)
-collaboration\footnote{\url{http://fire.northwestern.edu}}
-\citep{2016ApJ...827L..23W,2018MNRAS.480..800H}. In a zoom-in cosmological
-simulation a high resolution region of interest is embedded within a low
-resolution cosmological context (\Gus{cites}). In the case of {\em Latte}, a
-region that will evolve into a Milky Way-mass galaxy, identified in a
-preliminary dark matter only run, is increased in resolution using MUSIC.
-This allows one to achieve ultra-high resolution simulations of individual
-galaxies than possible in a full cosmological run.
-
-Our centering differs from Sanderson et al. by $\sim3.2\,\pc$ for \mi.
-
-\section{Optimally Axisymmetric Frame}
+\section{Optimally Axisymmetric Frame} \label{sec:oa_frame}
 An analytic, axisymmetric potential with a known reference frame is a far too
 simple description of the Milky Way. We thus turn to cosmological zoom-in
-simulations as approximations to a ``true'' Milky Way potential. We will perform a few simulations of small clusters in order to
+simulations as approximations to a ``true'' Milky Way potential. We will
+perform a few simulations of small clusters embedded in the potential of a
+frozen $\z=0$ snapshot of one such zoom-in. We take our fiducial reference
+frame from \citet{2018arXiv180610564S}.
+
+Freezing the potential means that during the course of our simulation the
+cluster will rapidly pass by much of the substructure in the potential. To
+illustrate that this has little effect on our results, we resimulate each
+cluster with the softening lengths of the embedded potential increased by a
+factor of \Gus{something}.
+
+Once we perform this simulation, we calculate the actions of each star in the
+cluster assuming the best-fit smooth, axisymmetric potential, as we do
+observationally for the Milky Way.
+We refer to this space as the observational action space.
+We will find that there is significant time evolution in observational action space.
+Attributing some of this evolution to our fiducial reference frame being incorrect, we minimize the time-dispersion in action space
 
@@
 First, a dark matter only simulation was run with $\Lambda$CDM cosmology
-parameterized by $(\Omega_m, \Omega_{\Lambda}, \Omega_b, h, \sigma_8, n_s) =
-(0.272, 0.728, 0.0455, 0.702, 0.807, 0.961)$, consistent with current
-constraints \citep{2018arXiv180706209P}. Halos are selected at $z=0$ based
-solely on their mass ($\sim 1\text{--}2 \times 10^{12} \Msun$). Particles
-within $5 R_{200\text{m}}$ of each halo are traced back to $\z=99$ and the
-initial conditions are regenerated at higher resolution using MUSIC
+parameterized by $\Omega_m = 0.272$, $\Omega_{\Lambda} = 0.728$, $\Omega_b =
+0.0455$, $h = 0.702$, $\sigma_8 = 0.807$, and $n_s = 0.961$, consistent with
+current constraints \citep{2018arXiv180706209P}. Halos are selected at $z=0$
+based solely on their mass ($\sim 1\text{--}2 \times 10^{12} \Msun$).
+Particles within $5 R_{200\text{m}}$ of each halo are traced back to $\z=99$
+and the initial conditions are regenerated at higher resolution using MUSIC
 \citep{2011MNRAS.415.2101H}. A total of $\sim 10$ halos were simulated at
@@
 These halos are uncontaminated and contain gas particles of mass $\sim 7000
-\text{--} 20000 \Msun$ and star particles of mass $\sim 5000 \text{--} 7000
+\text{--} 20,000\,\Msun$ and star particles of mass $\sim 5000 \text{--} 7000\,
 \Msun$, with the lower end coming from stellar evolution
@@
 
+Even though we change the center, we do not recalculate the center velocity.
+This is largely irrelevant, as the recalculated center velocity differs from
+the fiducial center velocity by $<0.1\,\kms$.
+
 \appendix \section{Appendix 1}

### 2018-12-04 paper/ms.tex
Update paper

@@
 \newcommand{\mm}{\texttt{m12m}}
-\newcommand{\gdr}{{\em Gaia} DR2}
 
@@
 one assumes the wrong galactocentric reference frame this can induce
-artificial non-axisymmetries which manifest as time dependence in the
-computed actions. However, the Milky Way is not axisymmetric in detail and so
-there are no truly axisymmetric axes. We instead define the  ``optimally
-axisymmetric'' axes as the reference frame which minimizes the time
-dependence of the computed actions. Standard methods for determining the
-galactocentric reference frame typically involve finding the median height of
-stars in the disk. Using cosmological zoom-in simulations from FIRE-2, we
-show that the optimally axisymmetric reference frame can differ from standard
-methods of calculating the galactocentric reference frame by nearly a scale
-height. This work calls into question the use of actions as dynamical
-invariants unless these reference frame issues can be resolved.
+artificial non-axisymmetries which manifest as time dependence in the computed
+actions. However, the Milky Way is not axisymmetric in detail and so there are
+no truly axisymmetric axes. We instead define the  ``optimally axisymmetric''
+axes as the reference frame which minimizes the time dependence of the
+computed actions. Standard methods for determining the galactocentric
+reference frame typically involve finding the median height of stars in the
+disk. Using cosmological zoom-in simulations from FIRE-2, we show that the
+optimally axisymmetric reference frame can differ from standard methods of
+calculating the galactocentric reference frame by nearly a scale height. This
+work calls into question the use of actions as dynamical invariants unless
+these reference frame issues can be resolved.
 
@@
 \section{Introduction} \label{sec:intro}
-Our understanding of the Milky Way is currently undergoing a revolution in the era of \gdr.
-Recent major discoveries include the remnants of a major merger \citep{2018Natur.563...85H}, a phase-space spiral (\Gus{cites}) indicating local substructure infall (\Gus{cites}), a gap indicative of a dark matter subtructure in GD1 \citep{2018arXiv181103631B}, and \Gus{others}.
+Our understanding of the Milky Way is currently undergoing a revolution in the
+era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
+major merger \citep{2018Natur.563...85H}, a phase-space spiral (\Gus{cites})
+indicating local substructure infall (\Gus{cites}), a gap indicative of a dark
+matter subtructure in GD1 \citep{2018arXiv181103631B}, and \Gus{others}.
+
+Underlying all of these discoveries is the assumption of a galactocentric
+reference frame \citep{2008gady.book.....B}. In angular
+coordinates, the center of the galaxy is taken to be the location of
+Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From stellar
+motions near Sgr~A\textsuperscript{*}, the distance to the center is taken to
+be $\sim 8.1\text{--}8.3\,\kpc$
+\citep{2009ApJ...692.1075G,2018A&A...615L..15G}. The vertical height is found
+by identifying where the stellar positions and velocities reach a maximum
+(effectively the median height of all disk stars), and is usually taken to be
+$\sim 25\,\pc$ \citep{2001ApJ...553..184C}, with more recent measurements from
+{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}.
+
+These measurements of the galactocentric reference frame rely on the
+assumption of an axisymmetric Milky Way. Under this assumption, these
+estimates would provide the correct parameters for the galactocentric
+reference frame up to measurement error. However, the Milky Way is not
+axisymmetric in detail. Spiral arms, bar(s), and infalling satellite galaxies
+such as Sagittarius and the Large Magellanic Cloud all induce
+non-axisymmetries.
+
+Since the Milky Way is not truly axisymmetric, it is not clear the best way to
+define the galactocentric reference frame. In this work, we consider one
+possible physically motivated approach to defining the galactocentric
+reference frame, and show that it can differ from conventional methods of
+frame measurements.
+
+In a slowly-evolving axisymmetric potential, the actions of stars' orbits are
+conserved quantities \citep{2008gady.book.....B,2014RvMP...86....1S}. In other
+words, if we could measure the actions of a star in such a potential, we would
+measure the same actions over the course of the orbit. The non-axisymmetries
+of a more realistic potential induce time evolution in the {\em measured}
+actions --- the true orbital invariants (if they exist) are invariant by
+definition. Through simple tests, we will see that even in a truly
+axisymmetric potential, if the wrong reference frame is assumed then there
+will be additional time evolution in the measured actions.
+
+To provide one such realization of a realistic Galactic potential, we turn to
+cosmological zoom-ins of Milky Way-mass galaxies from the Feedback in
+Realistic Environments (FIRE)
+collaboration\footnote{\url{https://fire.northwestern.edu}} (the Latte suite
+of simulations).
 
 \section{Reference Frame Gone Wrong} \label{sec:ref_frame}
-We begin by quantifying how
+We show and quantify the time evolution of measured actions that comes about
+from inaccuracies in the galactocentric reference frame.
+
+\subsection{Cartoon} \label{ssec:cartoon}
+We begin with a cartoon illustrating what we mean in Figure~\ref{fig:cartoon}.
+The solid gray line indicates a hypothetical ``true'' orbit. The y-axis
+corresponds to the vertical height of the orbit and the x-axis to either time
+or orbital phase. The dashed gray line is an incorrect midplane which an
+observer uses to compute orbits.
+
+Now suppose this observer makes a measurement of the true orbit at the orange
+point or the red point. Then the orange and red lines correspond to the orbits
+that the observer would compute for each point. In action space, this would
+correspond to a different value of $J_z$ for the orange and red points. In
+this way, assuming the wrong reference frame induces time dependence in the
+actions {\em computed by an observer}.
+
+\begin{figure*}
+\plotone{cartoon.pdf}
+\caption{Caption.}
+\label{fig:cartoon}
+\end{figure*}
+
+\subsection{Quantification} \label{ssec:quant}
+We now turn to classical orbital integration in a commonly assumed Milky Way potential to quantify the time variation of computed actions for different reference frame errors.
 
@@
 cluster assuming the best-fit smooth, axisymmetric potential, as we do
-observationally for the Milky Way.
-We refer to this space as the observational action space.
-We will find that there is significant time evolution in observational action space.
-Attributing some of this evolution to our fiducial reference frame being incorrect, we minimize the time-dispersion in action space
+observationally for the Milky Way. We refer to this space as the observational
+action space. We will find that there is significant time evolution in
+observational action space. Attributing some of this evolution to our fiducial
+reference frame being incorrect, we minimize the time-dispersion in action
+space
 
@@
 This is largely irrelevant, as the recalculated center velocity differs from
-the fiducial center velocity by $<0.1\,\kms$.
+the fiducial center velocity by $<0.1\,\kms$. In recentering the galaxy, the
+package \texttt{gizmo\_analysis} will subtract a slightly different value for
+the Hubble flow from each particle's velocity to convert to a physical
+reference frame. The error this introduces is $\sim 0.02\,\kms$, which we also
+ignore.
 

### 2018-12-04 schmactions/plot_one_orbit.py
Add gala orbit integration

@@ -1,0 +1,135 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+import astropy.units as u
+
+import gala.dynamics as gd
+import gala.integrate as gi
+import gala.potential as gp
+
+from scipy.stats import sigmaclip
+
+import warnings
+
+from tqdm import tqdm
+
+import matplotlib as mpl
+from matplotlib import rc
+rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+rc('text', usetex=True)
+mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
+
+tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
+        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
+
+dt = 1 * u.Myr
+t1 = 0.0 * u.Gyr
+t2 = 5.0 * u.Gyr
+
+wrong_max = 1000 # times dt, so 1 Gyr
+
+mw = gp.MilkyWayPotential()
+
+def compute_actions_wrong_ref_frame(init_pos, init_vel, offset, cadence=25, wrong_max=None):
+    q = gd.PhaseSpacePosition(pos=init_pos, vel=init_vel)
+    with warnings.catch_warnings(record=True):
+        warnings.simplefilter("ignore")
+        orbit = mw.integrate_orbit(q, dt=dt, t1=t1, t2=t2, Integrator=gi.DOPRI853Integrator)
+        res = gd.find_actions(orbit, N_max=8)
+        ans = res['actions'].to(u.kpc * u.km / u.s).value
+        print(ans)
+
+    pos = np.transpose(orbit.pos.xyz.to_value(u.kpc))
+    offset = offset.to_value(u.kpc)
+    pos = np.subtract(pos, offset)
+    vel = orbit.vel.d_xyz
+    pos = np.transpose(pos) * u.kpc
+    off_q = gd.PhaseSpacePosition(pos=pos, vel=vel)
+
+    if wrong_max is None:
+        wrong_max = len(orbit)
+
+    out_action = []
+    time = orbit.t[:wrong_max][::cadence]
+    with warnings.catch_warnings(record=True):
+        warnings.simplefilter("ignore")
[…]

### 2018-12-04 paper/ms.tex
Fix previous commit
Update paper

@@
 \newcommand{\kms}{\text{km}/\text{s}}
+\newcommand{\actunit}{\text{kpc}\,\kms}
 
@@
 \newcommand{\mm}{\texttt{m12m}}
+
+% affiliations
+\newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute, 162 5th
+Ave., New York, NY 10010, USA}
+\newcommand{\penn}{Department of Physics \& Astronomy, University of Pennsylvania, 209
+South 33rd St., Philadelphia, PA 19104, USA}
+\newcommand{\amnh}{Department of Astrophysics, American Museum of Natural History, Central Park West at 79th St., New York, NY 10024, USA}
+\newcommand{\columbia}{Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 10027, USA}
 
@@
 \author{Angus Beane}
-\affil{Center for Computational Astrophysics, Flatiron Institute, 162 5th
-Ave., New York, NY 10010, USA}
-\affil{Department of Physics \& Astronomy, University of Pennsylvania, 209
-South 33rd St., Philadelphia, PA 19104, USA}
+\affil{\cca}
+\affil{\penn}
 
-\author{Many Helpful People}
-\noaffiliation
+\author{Robyn Sanderson}
+\affil{\cca}
+\affil{\penn}
+
+\author{Mordecai-Mark Mac Low}
+\affil{\cca}
+\affil{\amnh}
+
+\author{Melissa K. Ness}
+\affil{\cca}
+\affil{\columbia}
+
+\author{Daniel Angl\'es-Alc\'azar}
+\affil{\cca}
+
+\author{Megan Bedell}
+\affil{\cca}
 
@@
 \begin{figure*}
-\plotone{cartoon.pdf}
+\plotone{fig/cartoon.pdf}
 \caption{Caption.}
@@
 \subsection{Quantification} \label{ssec:quant}
-We now turn to classical orbital integration in a commonly assumed Milky Way potential to quantify the time variation of computed actions for different reference frame errors.
+We now turn to classical orbital integration in a commonly assumed Milky Way
+potential to quantify the time variation of computed actions for different
+reference frame errors.
+
+We use the code \texttt{gala} v0.3 to perform orbit integrations and
+conversion to action space \citep{2017JOSS....2..388P,Price-Whelan:2018},
+which uses an action estimator technique presented by
+\citet{2014MNRAS.441.3284S}. We use the default \texttt{MWPotential} as our
+potential, which is based on the Milky Way potential available in
+\texttt{galpy} \citep{2015ApJS..216...29B}. This potential includes a
+Hernquist bulge and nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
+\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
+is fit to empirically match some observations. We use the Dormand-Prince
+8(5,3) integration scheme \citep{Dormand80:integrator}. We use a timestep of
+$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for
+a Sun-like star.
+
+First, we illustrate the time-dependence of one possible inaccurate reference
+frame. We begin by taking an orbit with initial position $(8, 0, 0)\,\kpc$ and
+initial velocity $(0, -190, 50)\,\kms$ --- this corresponds to an orbit with
+actions $(J_r, L_z, J_z) = (32.5, -1520, 23.0)\,\actunit$. The orbit is
+plotted in \Gus{Appendix x}. We then suppose that an observer makes
+observations of this orbit in a reference frame that is incorrect in radius by
+$200\,\pc$ and height by $100\,\pc$ --- we subtract the vector $(200, 0,
+100)\,\pc$ from each position in the orbit. This corresponds to an observer
+physically located at the position $(8, 0, 0)\,\kpc$ but erroneously thinking
+they are located at $(8.2, 0, 0.1)\,\kpc$.
+
+Using the observations of the orbit that this erroneous observer makes every
+$\Myr$, we perform the same orbit integration procedure as before to compute
+the actions our observer would make in the incorrect reference frame over the
+duration of the orbit. The erroneous computed actions are shown for the first
+$\Gyr$ of the orbit in Figure~\ref{fig:one_orbit_wrong_ref}. Occasionally the
+numerical scheme fails and very large actions are reported by \texttt{gala}
+--- we perform a $5\sigma$ clip on each action to exclude such orbits, but
+this only excludes a total of $5$ orbits out of the $1000$ considered for
+Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
+vast majority of orbits are computed properly.
+
+\begin{figure*}
+\plotone{fig/one_orbit.pdf}
+\caption{Caption.}
+\label{fig:one_orbit_wrong_ref}
+\end{figure*}
 
@@
 \texttt{pykdgrav}\footnote{\url{https://github.com/omgspace/pykdgrav},
-commit:\texttt{fbd1ddc}}.
+commit:\texttt{fbd1ddc}}. We use an opening angle of $0.5$ and exclude stars
+at distances $R>50\,\kpc$ in the fiducial reference frame.
 
 We use the initial position and velocity of two star particles in
-$\texttt{m12i}$
+$\texttt{m12i}$. These two star particles are randomly chosen from all star
+particles in the simulation with the following properties:
 

### 2018-12-05 runs/oceanic_schmactions/rotate_pa.py
Just minimize Jz

@@ -13,6 +13,7 @@ sim_dir = '/mnt/ceph/users/firesims/fire2/metaldiff/m12i_res7100/'
 gal_info = 'm12i_info.txt'
 
 cadence_list = [50, 20]
+#cadence_list = [50]
 
 tend = 2000 # Myr
 dt = 0.1 # Myr
@@ -122,7 +123,7 @@ def chisq(x, return_actions=False):
     if return_actions:
         return act_collapse
     act_collapse = np.subtract(act_collapse, act_med)
-    return np.sum(np.square(act_collapse/act_med))
+    return np.sum(np.square(act_collapse/act_med)[:,2])
 
 xinit = np.array([0, 0, 0, 0, 0, 0])
 for cd in cadence_list:

### 2018-12-05 runs/oceanic_schmactions/plot_actions.py
update plots

@@ -17,7 +17,7 @@ tlist = np.arange(0 , tend, dt)
 cadence = 20
 tlist = tlist[::cadence]
 
-tmax = 1000
+tmax = 2000
 
 #idx_list = ['0', '1']
 idx_list = ['0']

### 2018-12-05 schmactions/plot_one_orbit.py
Add variable zoff

@@ -13,6 +13,9 @@ import warnings
 
 from tqdm import tqdm
 
+from joblib import Parallel, delayed
+import multiprocessing
+
 import matplotlib as mpl
 from matplotlib import rc
 rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
@@ -26,10 +29,11 @@ dt = 1 * u.Myr
 t1 = 0.0 * u.Gyr
 t2 = 5.0 * u.Gyr
 
+ncpu = 8
+
 wrong_max = 1000 # times dt, so 1 Gyr
 
 mw = gp.MilkyWayPotential()
-
 def compute_actions_wrong_ref_frame(init_pos, init_vel, offset, cadence=25, wrong_max=None):
     q = gd.PhaseSpacePosition(pos=init_pos, vel=init_vel)
     with warnings.catch_warnings(record=True):
@@ -51,13 +55,15 @@ def compute_actions_wrong_ref_frame(init_pos, init_vel, offset, cadence=25, wron
 
     out_action = []
     time = orbit.t[:wrong_max][::cadence]
-    with warnings.catch_warnings(record=True):
-        warnings.simplefilter("ignore")
-        for this_q in tqdm(off_q[:wrong_max][::cadence]):
+    
+    for this_q in tqdm(off_q[:wrong_max][::cadence]):
+        with warnings.catch_warnings(record=True):
+            warnings.simplefilter("ignore")
             this_orbit = mw.integrate_orbit(this_q, dt=dt, t1=t1, t2=t2, Integrator=gi.DOPRI853Integrator)
             res = gd.actionangle.find_actions(this_orbit, N_max=8)
             ans = res['actions'].to(u.kpc * u.km / u.s).value
             out_action.append(ans)
+    #out_action = Parallel(n_jobs=ncpu) (delayed(loop)(this_q) for this_q in tqdm(orbit[:wrong_max][::cadence]))
     return time, np.array(out_action), orbit
 
 def plot_wrong_act(t, act, rel_error=False):
@@ -123,13 +129,30 @@ init_pos = [8, 0, 0] * u.kpc
 init_vel = [0, -190, 50] * u.km/u.s
 offset = [100, 0, 200] * u.pc
 
+sys.exit(0)
+
 t, act, orbit = compute_actions_wrong_ref_frame(init_pos, init_vel, offset, cadence=1, wrong_max=wrong_max)
[…]

### 2018-12-05 paper/ms.tex
Update paper

@@
 
+Figure~\ref{fig:one_orbit_wrong_ref} shows that the erroneous observer would
+infer significant time evolution in their computed actions. The
+95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $8.7$,
+$47.1$, and $19.3\,\actunit$ for $J_r$, $L_z$, and $J_z$, respectively. These
+are fractional errors of $26.5\%$, $3.1\%$, and $81.4\%$, respectively.
+
 \begin{figure*}
@@
 \end{figure*}
+
+We now
 
@@
 \texttt{pykdgrav}\footnote{\url{https://github.com/omgspace/pykdgrav},
-commit:\texttt{fbd1ddc}}. We use an opening angle of $0.5$ and exclude stars
+commit: \texttt{fbd1ddc}}. We use an opening angle of $0.5$ and exclude stars
 at distances $R>50\,\kpc$ in the fiducial reference frame.
@@
 $\texttt{m12i}$. These two star particles are randomly chosen from all star
-particles in the simulation with the following properties:
+particles in the simulation with the following properties (in the fiducial
+reference frame):
+\begin{enumerate}
+    \item age in the range $(0.25, 0.75)\,\Gyr$,
+    \item cylindrical radius in the range $(7.7, 8.7)\,\kpc$,
+    \item height (absolute value) less than $0.5\,\kpc$,
+    \item $J_r$ and $J_z$ in the range:
+        \subitem $(10, 20)\,\actunit$ for the first star,
+        \subitem $(40, 60)\,\actunit$ for the second star.
+\end{enumerate}
+The ids we choose are \texttt{23693026} for the first star and
+\texttt{17012804} for the second star. We refrain from calling these ``low''
+or ``high'' action stars since the actions are computed in a presumably
+incorrect reference frame.
+
+To compute actions, both in the choosing of each of these star particles and
+in what follows, we use the package
+\texttt{agama}\footnote{\url{https://github.com/GalacticDynamics-Oxford/Agama},
+commit: \texttt{0e48336}}. Using only the high-resolution particles, we
+compute the symmetrized potential in two components. First, a multipole
+expansion with $20$ radial grid points and a max order of $2$ for the
+spherical-harmonic expansion is used for the dark matter. Second, a
+cylindrical spline expansion with $20$ radial and $z$ grid points, a $0$ order
+max index azimuthal-harmonic expansion, $0.2$ and $50\,\kpc$ for the minimum
+and maximum radial grid node, $0.02$ and $10\,\kpc$ for the minimum and
+maximum $z$ grid node is used for the baryonic matter (stars and gas). In both
+cases, axisymmetry is explicitly enforced. Other \texttt{agama} parameters are
+left as default.
+
+\subsection{Frame Optimization} \label{ssec:frame_opt}
+Figure~\Gus{something} shows the action computed at each time step in the
+fiducial reference frame. Clearly there is significant time evolution.
 

### 2018-12-05 schmactions/plot_one_orbit.py
Add sigma clip

@@ -150,7 +150,7 @@ def z_off(z):
     this_offset = [0, 0, z] * u.pc
     print('my offset is!:', z)
     t, act, orbit = compute_actions_wrong_ref_frame(init_pos, init_vel, this_offset, cadence=10, wrong_max=wrong_max)
-    return np.percentile(act, 95, axis=0) - np.percentile(act, 5, axis=0), t, act, orbit
+    return np.percentile(act, 95, axis=0) - np.percentile(act, 5, axis=0)
 
 zofflist = np.linspace(0, 500, 50) # u.pc
 #perc_list = np.array([z_off(z) for z in tqdm(zofflist)])

### 2018-12-06 paper/ms.tex
Update paper

@@
 South 33rd St., Philadelphia, PA 19104, USA}
-\newcommand{\amnh}{Department of Astrophysics, American Museum of Natural History, Central Park West at 79th St., New York, NY 10024, USA}
-\newcommand{\columbia}{Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 10027, USA}
+\newcommand{\amnh}{Department of Astrophysics, American Museum of Natural
+History, Central Park West at 79th St., New York, NY 10024, USA}
+\newcommand{\columbia}{Department of Astronomy, Columbia University, 550 W
+120th St., New York, NY 10027, USA}
 
@@
 
-\title{Our Galactocentric Reference Frame Is Probably Wrong, and Not by a
+\title{Our Galactocentric Coordinate System Is Probably Wrong, and Not by a
 Little}
@@
 
-In an axisymmetric potential the 6D phase space of stars' positions and
-velocities can be reduced to a 3D space of invariant actions. We show that if
-one assumes the wrong galactocentric reference frame this can induce
-artificial non-axisymmetries which manifest as time dependence in the computed
-actions. However, the Milky Way is not axisymmetric in detail and so there are
-no truly axisymmetric axes. We instead define the  ``optimally axisymmetric''
-axes as the reference frame which minimizes the time dependence of the
-computed actions. Standard methods for determining the galactocentric
-reference frame typically involve finding the median height of stars in the
-disk. Using cosmological zoom-in simulations from FIRE-2, we show that the
-optimally axisymmetric reference frame can differ from standard methods of
-calculating the galactocentric reference frame by nearly a scale height. This
-work calls into question the use of actions as dynamical invariants unless
-these reference frame issues can be resolved.
+In a completely axisymmetric potential, the notion of a global Galactic
+midplane is well defined. However, since the Milky Way is not completely
+axisymmetric, the global Galactic midplane is to some level of precision not
+well defined. The local Galactic midplane has been found by determining the
+height at which the number counts of stars reaches a maximum. This local
+midplane is then used as the global midplane, an assumption which would be
+justified in a completely axisymmetric potential. Assuming a global Galactic
+midplane exists, we show in three cosmological zoom-in simulations from the
+FIRE collaboration that local determinations of the midplane can vary by
+nearly a thin disk scale height. Using standard orbit integration techniques,
+we show that such an inaccuracy in the midplane can induce artifical time
+evolution in the vertical action that makes it virtually unusable for
+individual stars in the disk. The situation is less dire for angular momentum
 
-% However, it is not expected that actions are invariant in detail for the
-% Milky Way. This in-invariance can come about from assuming the wrong
-% Galactic potential, but it can also come about due to true
-% non-axisymmetries. We show that actions can gain artificial time-dependence
-% also by assuming the wrong reference frame. We argue that standard methods
-% for determining the galactocentric reference frame are not gu
+% Local determinations of the Galactic midplane rely on finding the height at
+% which the number count of stars reaches a maximum. This is not guaranteed to
+% be
 
-% it's spiral arms, bar(s), and gas clouds are all non-axisymmetric features
-% that should induce long-term evolution in action space. However, these same
-% features will cause a star's true orbit to deviate slightly from its orbit
-% in an axisymmetric potential. If an observer assumes an axisymmetric
-% potential when computing actions, this will induce an artificial short-term
-% phase dependence of the actions, which we interpret as systematic error.
-% Using the potential of cosmological zoom-in simulations of Milky Way mass
-% galaxies from the FIRE collaboration (the Latte suite), we investigate the
-% magnitude of this systematic error. We find XXX. Something something lower
-% limit.
+% In an axisymmetric, adiabatic potential, the 6D phase space of stars' positions and velocities can be reduced to a 3D space of invariant actions.
+We show that if one assumes the wrong galactocentric coordinate system this can induce significant artificial time dependence in the computed actions.
+
+% In an axisymmetric potential the 6D phase space of stars' positions and
+% velocities can be reduced to a 3D space of invariant actions. We show that if
+% one assumes the wrong galactocentric coordinate system this can induce
+% artificial non-axisymmetries which manifest as time dependence in the computed
+% actions. However, the Milky Way is not axisymmetric in detail and so there are
+% no truly axisymmetric axes. We instead define the  ``optimally axisymmetric''
+% axes as the coordinate system which minimizes the time dependence of the
+% computed actions. Standard methods for determining the galactocentric
+% coordinate system typically involve finding the median height of stars in the
+% disk. Using cosmological zoom-in simulations from FIRE-2, we show that the
+% optimally axisymmetric coordinate system can differ from standard methods of
+% calculating the galactocentric coordinate system by nearly a scale height.
+% This work calls into question the use of actions as dynamical invariants
+% unless these coordinate system issues can be resolved.
 
@@
 Underlying all of these discoveries is the assumption of a galactocentric
-reference frame \citep{2008gady.book.....B}. In angular
+coordinate system \citep{2008gady.book.....B}. In angular
 coordinates, the center of the galaxy is taken to be the location of
@@
 
-These measurements of the galactocentric reference frame rely on the
+These measurements of the galactocentric coordinate system rely on the
 assumption of an axisymmetric Milky Way. Under this assumption, these
 estimates would provide the correct parameters for the galactocentric
-reference frame up to measurement error. However, the Milky Way is not
+coordinate system up to measurement error. However, the Milky Way is not
 axisymmetric in detail. Spiral arms, bar(s), and infalling satellite galaxies
@@
 Since the Milky Way is not truly axisymmetric, it is not clear the best way to
-define the galactocentric reference frame. In this work, we consider one
+define the galactocentric coordinate system. In this work, we consider one
 possible physically motivated approach to defining the galactocentric
-reference frame, and show that it can differ from conventional methods of
+coordinate system, and show that it can differ from conventional methods of
 frame measurements.
@@
 definition. Through simple tests, we will see that even in a truly
-axisymmetric potential, if the wrong reference frame is assumed then there
+axisymmetric potential, if the wrong coordinate system is assumed then there
 will be additional time evolution in the measured actions.
@@
 
-\section{Reference Frame Gone Wrong} \label{sec:ref_frame}
+\section{Local Midplane in FIRE} \label{sec:local_fire}
+\subsection{Description of Simulation} \label{ssec:cosmozoom}
+Here we briefly describe the cosmological zoom-ins used in this work.
+Cosmological zoom-ins allow one to simulate a selected region at high
+resolution embedded in a low-resolution cosmological background
+\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}. The FIRE
+collaboration has simulated a number of Milky Way-mass zoom-ins as part of the
+{\em Latte} suite of FIRE-2 simulations
+\citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. We use the three zoom-ins
+considered in \citet{2018arXiv180610564S} --- \texttt{m12i}, \texttt{m12f},
+\texttt{m12m}, since they have associated mock {\em Gaia} DR2 catalogues and
+the $z=0$ snapshots are publicly available.
+
+First, a dark matter only simulation was run with $\Lambda$CDM cosmology
+parameterized by $\Omega_m = 0.272$, $\Omega_{\Lambda} = 0.728$, $\Omega_b =
+0.0455$, $h = 0.702$, $\sigma_8 = 0.807$, and $n_s = 0.961$, consistent with
+current constraints \citep{2018arXiv180706209P}. Halos are selected at $z=0$
+based solely on their mass ($\sim 1\text{--}2 \times 10^{12} \Msun$).
+Particles within $5 R_{200\text{m}}$ of each halo are traced back to $\z=99$
+and the initial conditions are regenerated at higher resolution using MUSIC
+\citep{2011MNRAS.415.2101H}.
+
+Simulations were run using
+\texttt{GIZMO}\footnote{\url{http://www.tapir.caltech.edu/~phopkins/Site/GIZMO.html}}
+\citep{2015MNRAS.450...53H}. Hydrodynamics is solved using the meshless
+finite-mass method and gravity is solved using a modified version of the
+Tree-PM solver of \texttt{GADGET-3}, using fully adaptive and fully
+conservative gravitational force softening for gas.
+
+These halos are uncontaminated and contain gas particles of mass $\sim 7000
+\text{--} 20,000\,\Msun$ and star particles of mass $\sim 5000 \text{--} 7000\,
+\Msun$, with the lower end coming from stellar evolution
+\citep{2018arXiv180610564S}. Softening lengths for dark matter and star
+particles are fixed at $112\,\pc$ and $11.2\,\pc$,
+respectively.\footnote{This is $2.8$ times the often-quoted
+Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
+median softening length for gas particles around roughly solar positions
+(with cylindrical radii within $500\,\pc$ of $8.2\,\kpc$ and
+$\abs{z}<1\,\kpc$) is $98\,\pc$.
+
+% We restrict ourselves to the simulation \texttt{m12i} \citep[first introduced
+% in][]{2016ApJ...827L..23W}. For this work, this galaxy is the best-case
+% scenario. It does not contain a bar (like \texttt{m12m}) or companion galaxy,
+% and by eye is the closest to axisymmetric of the three galaxies presented in
+% \citet{2018arXiv180610564S} \citep{2018MNRAS.481.4133G}.
+
+It is important to remember that simulations of Milky Way-mass galaxies are
+not perfect representations of the true Milky Way, as discussed in
+\citet{2018arXiv180610564S}. For instance, the velocity structure of
+\texttt{m12i} is closer to M31 than the Milky Way. However, in this work we
+are most interested in the global properties of the \texttt{m12i} potential,
+and specifically in deviations from axisymmetry. From this perspective,
+\texttt{m12i} is actually far more axisymmetric than we might expect of the
+Milky Way. While \texttt{m12i} has prominent spiral arms, it lacks a bar and
+a companion like the Large Magellanic Cloud. In future work, these additional
+complications also need to be addressed.
+
+\subsection{Local Midplane} \label{sec:local_midplane}
+We
+
+\section{Coordinate System Gone Wrong} \label{sec:ref_frame}
 We show and quantify the time evolution of measured actions that comes about
-from inaccuracies in the galactocentric reference frame.
+from inaccuracies in the galactocentric coordinate system.
 
@@
 correspond to a different value of $J_z$ for the orange and red points. In
-this way, assuming the wrong reference frame induces time dependence in the
+this way, assuming the wrong coordinate system induces time dependence in the
 actions {\em computed by an observer}.
@@
 potential to quantify the time variation of computed actions for different
-reference frame errors.
+coordinate system errors.
 
@@
 plotted in \Gus{Appendix x}. We then suppose that an observer makes
-observations of this orbit in a reference frame that is incorrect in radius by
+observations of this orbit in a coordinate system that is incorrect in radius by
 $200\,\pc$ and height by $100\,\pc$ --- we subtract the vector $(200, 0,
@@
 $\Myr$, we perform the same orbit integration procedure as before to compute
-the actions our observer would make in the incorrect reference frame over the
+the actions our observer would make in the incorrect coordinate system over the
 duration of the orbit. The erroneous computed actions are shown for the first
@@
 \section{Optimally Axisymmetric Frame} \label{sec:oa_frame}
-An analytic, axisymmetric potential with a known reference frame is a far too
-simple description of the Milky Way. We thus turn to cosmological zoom-in
+An analytic, axisymmetric potential with a known coordinate system is a far
+too simple description of the Milky Way. We thus turn to cosmological zoom-in
 simulations as approximations to a ``true'' Milky Way potential. We will
@@
 observational action space. Attributing some of this evolution to our fiducial
-reference frame being incorrect, we minimize the time-dispersion in action
+coordinate system being incorrect, we minimize the time-dispersion in action
 space
-
-\subsection{Cosmological Zoom-in} \label{ssec:cosmozoom}
-Cosmological zoom-ins allow one to simulate a selected region at high
-resolution embedded in a low-resolution cosmological background
-\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}. Here we use a
-zoom-in of a Milky Way-mass galaxy from the FIRE-2 collaboration.
-
-First, a dark matter only simulation was run with $\Lambda$CDM cosmology
-parameterized by $\Omega_m = 0.272$, $\Omega_{\Lambda} = 0.728$, $\Omega_b =
-0.0455$, $h = 0.702$, $\sigma_8 = 0.807$, and $n_s = 0.961$, consistent with
-current constraints \citep{2018arXiv180706209P}. Halos are selected at $z=0$
-based solely on their mass ($\sim 1\text{--}2 \times 10^{12} \Msun$).
-Particles within $5 R_{200\text{m}}$ of each halo are traced back to $\z=99$
-and the initial conditions are regenerated at higher resolution using MUSIC
-\citep{2011MNRAS.415.2101H}. A total of $\sim 10$ halos were simulated at
-higher resolution \citep{2018MNRAS.481.4133G}.
-
-These halos are uncontaminated and contain gas particles of mass $\sim 7000
-\text{--} 20,000\,\Msun$ and star particles of mass $\sim 5000 \text{--} 7000\,
-\Msun$, with the lower end coming from stellar evolution
-\citep{2018arXiv180610564S}. Softening lengths for dark matter and star
-particles are fixed at $112\,\pc$ and $11.2\,\pc$,
-respectively.\footnote{This is $2.8$ times the often-quoted
-Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
-median softening length for gas particles around roughly solar positions
-(with cylindrical radii within $500\,\pc$ of $8.2\,\kpc$ and
-$\abs{z}<1\,\kpc$) is $98\,\pc$.
-
-We restrict ourselves to the simulation \texttt{m12i} \citep[first introduced
-in][]{2016ApJ...827L..23W}. For this work, this galaxy is the best-case
-scenario. It does not contain a bar (like \texttt{m12m}) or companion galaxy,
-and by eye is the closest to axisymmetric of the three galaxies presented in
-\citet{2018arXiv180610564S} \citep{2018MNRAS.481.4133G}.
-
-It is important to remember that simulations of Milky Way-mass galaxies are
-not perfect representations of the true Milky Way, as discussed in
-\citet{2018arXiv180610564S}. For instance, the velocity structure of
-\texttt{m12i} is closer to M31 than the Milky Way. However, in this work we
-are most interested in the global properties of the \texttt{m12i} potential,
-and specifically in deviations from axisymmetry. From this perspective,
-\texttt{m12i} is actually far more axisymmetric than we might expect of the
-Milky Way. While \texttt{m12i} has prominent spiral arms, it lacks a bar and
-a companion like the Large Magellanic Cloud. In future work, these additional
-complications also need to be addressed.
 
@@
 commit: \texttt{fbd1ddc}}. We use an opening angle of $0.5$ and exclude stars
-at distances $R>50\,\kpc$ in the fiducial reference frame.
+at distances $R>50\,\kpc$ in the fiducial coordinate system.
 
@@
 particles in the simulation with the following properties (in the fiducial
-reference frame):
+coordinate system):
 \begin{enumerate}
@@
 or ``high'' action stars since the actions are computed in a presumably
-incorrect reference frame.
+incorrect coordinate system.
 
@@
 Figure~\Gus{something} shows the action computed at each time step in the
-fiducial reference frame. Clearly there is significant time evolution.
+fiducial coordinate system. Clearly there is significant time evolution.
 
@@
 the Hubble flow from each particle's velocity to convert to a physical
-reference frame. The error this introduces is $\sim 0.02\,\kms$, which we also
-ignore.
+coordinate system. The error this introduces is $\sim 0.02\,\kms$, which we
+also ignore.
 

### 2018-12-08 paper/ms.tex
Update paper

@@
 \newcommand{\z}{z}
-
-\newcommand{\mi}{\texttt{m12i}}
-\newcommand{\mf}{\texttt{m12f}}
-\newcommand{\mm}{\texttt{m12m}}
 
@@
 midplane exists, we show in three cosmological zoom-in simulations from the
-FIRE collaboration that local determinations of the midplane can vary by
-nearly a thin disk scale height. Using standard orbit integration techniques,
-we show that such an inaccuracy in the midplane can induce artifical time
-evolution in the vertical action that makes it virtually unusable for
-individual stars in the disk. The situation is less dire for angular momentum
-
-% Local determinations of the Galactic midplane rely on finding the height at
-% which the number count of stars reaches a maximum. This is not guaranteed to
-% be
-
-% In an axisymmetric, adiabatic potential, the 6D phase space of stars' positions and velocities can be reduced to a 3D space of invariant actions.
-We show that if one assumes the wrong galactocentric coordinate system this can induce significant artificial time dependence in the computed actions.
-
-% In an axisymmetric potential the 6D phase space of stars' positions and
-% velocities can be reduced to a 3D space of invariant actions. We show that if
-% one assumes the wrong galactocentric coordinate system this can induce
-% artificial non-axisymmetries which manifest as time dependence in the computed
-% actions. However, the Milky Way is not axisymmetric in detail and so there are
-% no truly axisymmetric axes. We instead define the  ``optimally axisymmetric''
-% axes as the coordinate system which minimizes the time dependence of the
-% computed actions. Standard methods for determining the galactocentric
-% coordinate system typically involve finding the median height of stars in the
-% disk. Using cosmological zoom-in simulations from FIRE-2, we show that the
-% optimally axisymmetric coordinate system can differ from standard methods of
-% calculating the galactocentric coordinate system by nearly a scale height.
-% This work calls into question the use of actions as dynamical invariants
-% unless these coordinate system issues can be resolved.
+FIRE collaboration that local determinations of the midplane can vary by up to
+a thin disk scale height. The effect is more pronounced in galaxies with less
+quiescent histories (e.g. Sagittarius-like encounters). We explore one
+consequence of such an inaccuracy of the midplane in the context of orbital
+actions. Using standard orbit integration techniques, we show that midplane
+inaccuracy induces artifical time evolution in the vertical action that makes
+it virtually unusable for individual stars in the disk. The situation is less
+dire for angular momentum and radial action determinations, though we also
+show that inaccuracies in the solar radius can have a similar effect on the
+latter. As a result, we question the use of actions as dynamical invariants in
+the Milky Way disk.
 
@@
 \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. We use the three zoom-ins
-considered in \citet{2018arXiv180610564S} --- \texttt{m12i}, \texttt{m12f},
-\texttt{m12m}, since they have associated mock {\em Gaia} DR2 catalogues and
-the $z=0$ snapshots are publicly available.
+considered in \citet{2018arXiv180610564S} --- m12i, m12f, m12m, since they
+have associated mock {\em Gaia} DR2 catalogues and the $z=0$ snapshots are
+publicly available.
 
@@
 \text{--} 20,000\,\Msun$ and star particles of mass $\sim 5000 \text{--} 7000\,
-\Msun$, with the lower end coming from stellar evolution
-\citep{2018arXiv180610564S}. Softening lengths for dark matter and star
-particles are fixed at $112\,\pc$ and $11.2\,\pc$,
+\Msun$\footnote{In fact, the simulation m12i contained a gas
+splitting bug which causes a small number of gas and star particles to have
+larger than typical masses. These constitute only $\sim0.2\%$ of the star
+particles in the $z=0$ snapshots, and are only a factor of a few more massive
+than typical star particles. Since our results are broadly consistent between
+the three simulations, we ignore this minor complication.}, with the lower end
+coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
+for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
 respectively.\footnote{This is $2.8$ times the often-quoted
 Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
-median softening length for gas particles around roughly solar positions
-(with cylindrical radii within $500\,\pc$ of $8.2\,\kpc$ and
-$\abs{z}<1\,\kpc$) is $98\,\pc$.
+median softening length for gas particles around roughly solar positions (with
+cylindrical radii within $500\,\pc$ of $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is
+$98\,\pc$.
 
-% We restrict ourselves to the simulation \texttt{m12i} \citep[first introduced
+In this work, we take the coordinate systems used in
+\citet{2018arXiv180610564S} as our fiducial coordinate systems for each
+galaxy. The center of the galaxy is found through an iterative ``shrinking
+spheres'' method, where the center is calculated for a given radius which is
+then reduced by $50\%$ until converging on the galaxy's center. The center of
+mass velocity is then determined by all star particles within $15\,\kpc$ of
+this center. The galaxy is then rotated onto the principal axis frame as
+determined by stars younger than $1\,\Gyr$ inside of the fiducial solar radius
+$R_{\odot} = 8.2\,\kpc$.
+
+% We restrict ourselves to the simulation m12i \citep[first introduced
 % in][]{2016ApJ...827L..23W}. For this work, this galaxy is the best-case
@@
 not perfect representations of the true Milky Way, as discussed in
-\citet{2018arXiv180610564S}. For instance, the velocity structure of
-\texttt{m12i} is closer to M31 than the Milky Way. However, in this work we
-are most interested in the global properties of the \texttt{m12i} potential,
-and specifically in deviations from axisymmetry. From this perspective,
-\texttt{m12i} is actually far more axisymmetric than we might expect of the
-Milky Way. While \texttt{m12i} has prominent spiral arms, it lacks a bar and
-a companion like the Large Magellanic Cloud. In future work, these additional
-complications also need to be addressed.
+\citet{2018arXiv180610564S}. For instance, the velocity structure of m12i is
+closer to M31 than the Milky Way. However, in this work we are most interested
+in the global properties of the m12i potential, and specifically in deviations
+from axisymmetry. From this perspective, m12i is actually far more
+axisymmetric than we might expect of the Milky Way. While m12i has prominent
+spiral arms, it lacks a bar and a companion like the Large Magellanic Cloud.
+In future work, these additional complications also need to be addressed.
 
 \subsection{Local Midplane} \label{sec:local_midplane}
-We
+We now turn to defining the local midplane that an observer might measure if
+they were situated in each of these galaxies. We place our imaginary observer
+at the solar radius of $8.2\,\kpc$ and azimuth $\phi$, and consider stars
+within a cylinder perpendicular to the fiducial disk
+
+For a given azimuth $\phi$ at the solar radius of $8.2\,\kpc$, we take the
+median height of all star particles within a cylinder of radius $0.5\,\kpc$
+and total height $2\,\kpc$. This median height is taken to be what our
+imaginary observer would measure as the local Galactic midplane, and is
+reported for each galaxy in Figure~\ref{fig:midplane}. We find that only $10$
+iterations of this procedure are necessary for convergence. We bootstrap
+resample $1000$ times and report the $95^{\text{th}}$ minus $5^{\text{th}}$
+percentile as the dashed-line error bars.
+
+To allow for potential small inaccuracies in the fiducial coordinate systems,
+we also subtract the best fit $\text{A} \sin{\phi + \text{B}} + C$ from the
+midplane as a function of azimuth. For $\text{A}$, the values are $-165$,
+$45$, $9\,\pc$ and for $\text{C}$ the values are $-69$, $19$, and $-18\,\pc$
+for m12i, m12f, and m12m, respectively. For the assumed solar radius of
+$8.2\,\kpc$, we can approximate the angle offset for the $z$-axis from the
+values of $\text{A}$ --- we compute $1.15$, $0.31$, and $0.06\,\deg$ for m12i,
+m12f, and m12m. These angle offsets are consistent with the values given in
+\citet{2018arXiv180610564S} for the difference between the $z$-axis as defined
+by the gas and stars.
+
+Figure~\ref{fig:midplane} shows the inferred midplane our imaginary observer
+would make as a function of azimuth for each galaxy. The $90\%$ interquartile
+range for each galaxy is $185$, $162$, $84\,\pc$ for m12i, m12f, and m12m.
+
+\begin{figure*}
+\plotone{fig/midplane_fit.pdf}
+\caption{Caption.}
+\label{fig:midplane}
+\end{figure*}
+
+Our analysis makes the assumption that a global midplane is well defined.
+However, our results are also consistent with the notion that the global
+midplane is not a well-defined concept. Furthermore, none of the galaxies chosen in this work have
 
@@
 \subsection{Cartoon} \label{ssec:cartoon}
-We begin with a cartoon illustrating what we mean in Figure~\ref{fig:cartoon}.
+We begin with a cartoon in Figure~\ref{fig:cartoon} to show that an inaccurate determination of 
 The solid gray line indicates a hypothetical ``true'' orbit. The y-axis
@@
 We simulate a small star cluster of $256$ stars in the frozen $\z=0$
-potential of $\texttt{m12i}$. We use the code \texttt{ph4} within the
+potential of m12i. We use the code \texttt{ph4} within the
 \texttt{AMUSE} framework \citep{2011ascl.soft07007P, 2013CoPhC.184..456P,
@@
 We use the initial position and velocity of two star particles in
-$\texttt{m12i}$. These two star particles are randomly chosen from all star
+m12i. These two star particles are randomly chosen from all star
 particles in the simulation with the following properties (in the fiducial

### 2018-12-11 paper/ms.tex
Update paper

@@
 \usepackage{graphicx}
+\usepackage{gensymb}
+\usepackage{bm}
 
@@
 \newcommand{\z}{z}
+
+\newcommand{\beq}{\begin{equation}}
+\newcommand{\eeq}{\end{equation}}
 
@@
 However, our results are also consistent with the notion that the global
-midplane is not a well-defined concept. Furthermore, none of the galaxies chosen in this work have
+midplane is not a well-defined concept.
+
+\subsection{Dynamical Midplane}
+We now seek to define the midplane in a dynamically consistent manner, which
+we refer to as the ``dynamical midplane''. By this, we mean we would like to
+define the midplane in a manner consistent with some sort of dynamical
+constraint. Our method, which we describe now, is just one of many possible
+ways the dynamical midplane can be defined.
+
+We define the dynamical midplane based on the principle that perfectly
+circular, planar orbits will trace the midplane. Once the circular, planar
+orbits have been identified, the procedure for finding the dynamical midplane
+is relatively simple. From the measurement of each star's orbit, the midplane
+of the galaxy can be computed at $z=0$ easily. Furthermore, we will find that
+each orbit also gives a measure of the: dynamical center of the galaxy,
+velocity of the galaxy, and acceleration of the galaxy.
+
+We pause briefly to consider general difficulties in measuring stellar orbits
+in a cosmological simulation. It may seem reasonable to place the galaxy in an
+inertial reference frame and then to simply measure the positions of each star
+particle over time. However, it is not generally the case that the center of
+mass of the stellar disk experiences a vanishing acceleration. Indeed, in a
+re-run of the simulations presented here\footnote{The re-run contained a
+cosmic ray bug fix which is unimportant here, but the re-run also output the
+acceleration of each particle, which is not the case for the simulations in
+this work.}, we measure a total acceleration of $\sim 0.4\,\kms/\text{Myr}$
+for stars with cylindrical radii within $0.5\,\kpc$ of $8.2\,\kpc$ and
+vertical heights less than $0.4\,\kpc$. Assuming a constant acceleration, over
+$22\,\Myr$ --- the time spacing of snapshots considered in this work --- this
+results in a $\sim90\,\pc$ shift of the disk. Since the offset scales with
+$t^2$, this presents a significant problem for measuring stellar orbits over
+longer timescales.
+
+Another option is to recenter and reorient the galaxy at each snapshot.
+However, the centering and orientation procedures are not without
+stochasticity themselves. This stochasticity can easily dominate the magnitude
+of the orbit. For instance, a shift in the direction of the $z$-axis by a mere
+$0.7\degree$ results in a shift in the midplane of $100\,\pc$ at a radius of
+$8.2\,\kpc$. We attempted this procedure in various ways (i.e. exploring
+different centering and orientation procedures), but were not able to cleanly
+measure the orbits of disk stars.
+
+Here we choose to simplify the problem by considering only perfectly circular,
+planar orbits. Since these orbits have no phase-dependent structure, we can
+work on shorter timescales, i.e. for orbits with significant radial or
+vertical action a large fraction of their orbit would need to be sampled in
+order to measure the orbit accurately. Because we know the analytic form of
+these orbits, we can forward model the trajectory of each star particle in the
+cosmological simulation. Working in comoving coordinates, we can write down
+the trajectory of a star particle as
+
+\beq
+\bm{r}(t) = R(\theta, \phi, \psi) \left( (A\cos{\omega t}, B\sin{\omega t}, C t) + \bm{x_0} \right)
+\eeq
 
@@
 \subsection{Cartoon} \label{ssec:cartoon}
-We begin with a cartoon in Figure~\ref{fig:cartoon} to show that an inaccurate determination of 
-The solid gray line indicates a hypothetical ``true'' orbit. The y-axis
-corresponds to the vertical height of the orbit and the x-axis to either time
-or orbital phase. The dashed gray line is an incorrect midplane which an
-observer uses to compute orbits.
+We begin with a cartoon in Figure~\ref{fig:cartoon} to show that an inaccurate
+determination of The solid gray line indicates a hypothetical ``true'' orbit.
+The y-axis corresponds to the vertical height of the orbit and the x-axis to
+either time or orbital phase. The dashed gray line is an incorrect midplane
+which an observer uses to compute orbits.
 
@@
 
-\section{Optimally Axisymmetric Frame} \label{sec:oa_frame}
-An analytic, axisymmetric potential with a known coordinate system is a far
-too simple description of the Milky Way. We thus turn to cosmological zoom-in
-simulations as approximations to a ``true'' Milky Way potential. We will
-perform a few simulations of small clusters embedded in the potential of a
-frozen $\z=0$ snapshot of one such zoom-in. We take our fiducial reference
-frame from \citet{2018arXiv180610564S}.
+% \section{Optimally Axisymmetric Frame} \label{sec:oa_frame}
+% An analytic, axisymmetric potential with a known coordinate system is a far
+% too simple description of the Milky Way. We thus turn to cosmological zoom-in
+% simulations as approximations to a ``true'' Milky Way potential. We will
+% perform a few simulations of small clusters embedded in the potential of a
+% frozen $\z=0$ snapshot of one such zoom-in. We take our fiducial reference
+% frame from \citet{2018arXiv180610564S}.
 
-Freezing the potential means that during the course of our simulation the
-cluster will rapidly pass by much of the substructure in the potential. To
-illustrate that this has little effect on our results, we resimulate each
-cluster with the softening lengths of the embedded potential increased by a
-factor of \Gus{something}.
+% Freezing the potential means that during the course of our simulation the
+% cluster will rapidly pass by much of the substructure in the potential. To
+% illustrate that this has little effect on our results, we resimulate each
+% cluster with the softening lengths of the embedded potential increased by a
+% factor of \Gus{something}.
 
-Once we perform this simulation, we calculate the actions of each star in the
-cluster assuming the best-fit smooth, axisymmetric potential, as we do
-observationally for the Milky Way. We refer to this space as the observational
-action space. We will find that there is significant time evolution in
-observational action space. Attributing some of this evolution to our fiducial
-coordinate system being incorrect, we minimize the time-dispersion in action
-space
+% Once we perform this simulation, we calculate the actions of each star in the
+% cluster assuming the best-fit smooth, axisymmetric potential, as we do
+% observationally for the Milky Way. We refer to this space as the observational
+% action space. We will find that there is significant time evolution in
+% observational action space. Attributing some of this evolution to our fiducial
+% coordinate system being incorrect, we minimize the time-dispersion in action
+% space
 
-\subsection{Cluster Simulations} \label{ssec:cluster_sim}
-We simulate a small star cluster of $256$ stars in the frozen $\z=0$
-potential of m12i. We use the code \texttt{ph4} within the
-\texttt{AMUSE} framework \citep{2011ascl.soft07007P, 2013CoPhC.184..456P,
-2013A&A...557A..84P} for the cluster integration. We use a softening length
-of $0.01\,\pc$ which means we do not correctly solve the internal dynamics of
-the cluster, but this is not important for our purposes. To couple the
-gravitational forces of the frozen snapshot to the cluster stars, we use the
-\texttt{BRIDGE} \citep{2007PASJ...59.1095F}, also implemented in
-\texttt{AMUSE}. Our simulation runs for $2\,\Gyr$, allowing us to solve for
-$\sim 10$ orbits.
+% \subsection{Cluster Simulations} \label{ssec:cluster_sim}
+% We simulate a small star cluster of $256$ stars in the frozen $\z=0$
+% potential of m12i. We use the code \texttt{ph4} within the
+% \texttt{AMUSE} framework \citep{2011ascl.soft07007P, 2013CoPhC.184..456P,
+% 2013A&A...557A..84P} for the cluster integration. We use a softening length
+% of $0.01\,\pc$ which means we do not correctly solve the internal dynamics of
+% the cluster, but this is not important for our purposes. To couple the
+% gravitational forces of the frozen snapshot to the cluster stars, we use the
+% \texttt{BRIDGE} \citep{2007PASJ...59.1095F}, also implemented in
+% \texttt{AMUSE}. Our simulation runs for $2\,\Gyr$, allowing us to solve for
+% $\sim 10$ orbits.
 
-To compute the forces from the frozen potential, we use the simple tree code
-\texttt{pykdgrav}\footnote{\url{https://github.com/omgspace/pykdgrav},
-commit: \texttt{fbd1ddc}}. We use an opening angle of $0.5$ and exclude stars
-at distances $R>50\,\kpc$ in the fiducial coordinate system.
+% To compute the forces from the frozen potential, we use the simple tree code
+% \texttt{pykdgrav}\footnote{\url{https://github.com/omgspace/pykdgrav},
+% commit: \texttt{fbd1ddc}}. We use an opening angle of $0.5$ and exclude stars
+% at distances $R>50\,\kpc$ in the fiducial coordinate system.
 
-We use the initial position and velocity of two star particles in
-m12i. These two star particles are randomly chosen from all star
-particles in the simulation with the following properties (in the fiducial
-coordinate system):
-\begin{enumerate}
-    \item age in the range $(0.25, 0.75)\,\Gyr$,
-    \item cylindrical radius in the range $(7.7, 8.7)\,\kpc$,
-    \item height (absolute value) less than $0.5\,\kpc$,
-    \item $J_r$ and $J_z$ in the range:
-        \subitem $(10, 20)\,\actunit$ for the first star,
-        \subitem $(40, 60)\,\actunit$ for the second star.
-\end{enumerate}
-The ids we choose are \texttt{23693026} for the first star and
-\texttt{17012804} for the second star. We refrain from calling these ``low''
-or ``high'' action stars since the actions are computed in a presumably
-incorrect coordinate system.
+% We use the initial position and velocity of two star particles in
+% m12i. These two star particles are randomly chosen from all star
+% particles in the simulation with the following properties (in the fiducial
+% coordinate system):
+% \begin{enumerate}
+%     \item age in the range $(0.25, 0.75)\,\Gyr$,
+%     \item cylindrical radius in the range $(7.7, 8.7)\,\kpc$,
+%     \item height (absolute value) less than $0.5\,\kpc$,
+%     \item $J_r$ and $J_z$ in the range:
+%         \subitem $(10, 20)\,\actunit$ for the first star,
+%         \subitem $(40, 60)\,\actunit$ for the second star.
+% \end{enumerate}
+% The ids we choose are \texttt{23693026} for the first star and
+% \texttt{17012804} for the second star. We refrain from calling these ``low''
+% or ``high'' action stars since the actions are computed in a presumably
+% incorrect coordinate system.
 
-To compute actions, both in the choosing of each of these star particles and
-in what follows, we use the package
-\texttt{agama}\footnote{\url{https://github.com/GalacticDynamics-Oxford/Agama},
-commit: \texttt{0e48336}}. Using only the high-resolution particles, we
-compute the symmetrized potential in two components. First, a multipole
-expansion with $20$ radial grid points and a max order of $2$ for the
-spherical-harmonic expansion is used for the dark matter. Second, a
-cylindrical spline expansion with $20$ radial and $z$ grid points, a $0$ order
-max index azimuthal-harmonic expansion, $0.2$ and $50\,\kpc$ for the minimum
-and maximum radial grid node, $0.02$ and $10\,\kpc$ for the minimum and
-maximum $z$ grid node is used for the baryonic matter (stars and gas). In both
-cases, axisymmetry is explicitly enforced. Other \texttt{agama} parameters are
-left as default.
+% To compute actions, both in the choosing of each of these star particles and
+% in what follows, we use the package
+% \texttt{agama}\footnote{\url{https://github.com/GalacticDynamics-Oxford/Agama},
+% commit: \texttt{0e48336}}. Using only the high-resolution particles, we
+% compute the symmetrized potential in two components. First, a multipole
+% expansion with $20$ radial grid points and a max order of $2$ for the
+% spherical-harmonic expansion is used for the dark matter. Second, a
+% cylindrical spline expansion with $20$ radial and $z$ grid points, a $0$ order
+% max index azimuthal-harmonic expansion, $0.2$ and $50\,\kpc$ for the minimum
+% and maximum radial grid node, $0.02$ and $10\,\kpc$ for the minimum and
+% maximum $z$ grid node is used for the baryonic matter (stars and gas). In both
+% cases, axisymmetry is explicitly enforced. Other \texttt{agama} parameters are
+% left as default.
 
-\subsection{Frame Optimization} \label{ssec:frame_opt}
-Figure~\Gus{something} shows the action computed at each time step in the
-fiducial coordinate system. Clearly there is significant time evolution.
+% \subsection{Frame Optimization} \label{ssec:frame_opt}
+% Figure~\Gus{something} shows the action computed at each time step in the
+% fiducial coordinate system. Clearly there is significant time evolution.
 
-Even though we change the center, we do not recalculate the center velocity.
-This is largely irrelevant, as the recalculated center velocity differs from
-the fiducial center velocity by $<0.1\,\kms$. In recentering the galaxy, the
-package \texttt{gizmo\_analysis} will subtract a slightly different value for
-the Hubble flow from each particle's velocity to convert to a physical
-coordinate system. The error this introduces is $\sim 0.02\,\kms$, which we
-also ignore.
+% Even though we change the center, we do not recalculate the center velocity.
+% This is largely irrelevant, as the recalculated center velocity differs from
+% the fiducial center velocity by $<0.1\,\kms$. In recentering the galaxy, the
+% package \texttt{gizmo\_analysis} will subtract a slightly different value for
+% the Hubble flow from each particle's velocity to convert to a physical
+% coordinate system. The error this introduces is $\sim 0.02\,\kms$, which we
+% also ignore.
 

### 2018-12-11 paper/ms.tex
Update paper

@@
 the trajectory of a star particle as
+\beq\label{eq:helix}
+\bm{r}(t) = R(\theta, \phi, \psi) \left[ (A\cos{\omega t}, B\sin{\omega t}, C
+t) + \bm{x_0} \right]\text{,}
+\eeq
+where $R(\theta, \phi, \psi)$ is an Euler angle transformation, $A$, $B$, $C$,
+and $\omega$ are arbitrary constants, and $\bm{x_0}$ is an arbitrary 3-vector.
+This equation is a helix with an arbitrary constant and rotation. It describes
+the
 
-\beq
-\bm{r}(t) = R(\theta, \phi, \psi) \left( (A\cos{\omega t}, B\sin{\omega t}, C t) + \bm{x_0} \right)
-\eeq
+We minimize the summed squared difference between Equation~\ref{eq:helix} and
+the true trajectory of each star with $R$ within $1\,\kpc$ of $R_{\odot} =
+8.2\,\kpc$ in each simulation. We consider minimum rms error (i.e. the typical
+error in the Equation~\ref{eq:helix} model in describing the true trajectory).
+This rms error characterizes how well the helix model we consider, which would
+be exactly correct if the overall motion of the galaxy is well-described by
+constant speed motion and the orbit of a particular store is exactly circular
+and planar.
+
+We plot a histogram of the rms error for each galaxy in
+Figure~\ref{fig:something}. As we estimated before, the total effect of the
+galaxy's acceleration is roughly $90\,\pc$ (\Gus{Check for m12f, m12m}). We
+find that this is small compared to the typical error, even for the stars
+whose trajectories are best modeled by Equation~\ref{eq:helix}.
+
+There are many reasons for the failure of this model. First, there could be no
+stars on circular orbits. This seems unlikely considering that typical
+distribution functions peak on perfectly circular, planar orbits
+\citep[e.g.][]{2013MNRAS.434..652T}. Second, the trajectory of the galaxy may
+not be well-described by constant-speed motion. This also seems unlikely,
+considering that the bulk acceleration of the stellar disk only results in a
+$\sim90\,\pc$ (\Gus{m12f, m12m}) deviation from constant-speed motion over the
+$22\,\Myr$ time we considered.
+
+Furthermore, in Figure~\ref{fig:something} we plot the center of mass of the
+galaxy (in comoving coordinates) for the $22\,\Myr$ timeframe described. The
+center of mass at each snapshot is computed using the iterative shrinking
+spheres method described in Section~\ref{ssec:cosmozoom}. The rms error on the
+constant-speed motion is $\sim\text{something}$, which we interpret as
+\Gus{something}.
+
+One potential explanation is that {\em no star can remain on a circular,
+planar orbit}. The order unity density variations may be strong enough such
+that each star's orbit is perturbed. If there are truly stars in the
+simulation on perfectly circular, planar orbits, then the magnitude of their
+acceleration over the last $22\,\Myr$ must be constant. Furthermore, their
+trajectory in acceleration space must be well-described by a rotation. In the re-runs,
 
@@
 
-We now
+We now repeat the procedure but for
 

### 2018-12-14 paper/ms.tex
Update paper

--- 2018-12-11 paper/ms.tex
@@
 \shortauthors{Beane et al.}
+\shorttitle{Where the Hell Are We?}
 
@@
 \newcommand{\z}{z}
+\newcommand{\uth}{\textsuperscript{th}} % why is this so long in the first place??
 
@@
 
-\title{Our Galactocentric Coordinate System Is Probably Wrong, and Not by a
-Little}
+\title{Our Galactic Midplane Is Probably Wrong, and Not by a Little}
 
@@
 Since the Milky Way is not truly axisymmetric, it is not clear the best way to
-define the galactocentric coordinate system. In this work, we consider one
-possible physically motivated approach to defining the galactocentric
-coordinate system, and show that it can differ from conventional methods of
-frame measurements.
+define the Galactocentric coordinate system. In this work, we focus on the midplane and, equivalently, the solar height.
 
@@
 of simulations).
+
+\section{Impact on Actions} \label{sec:ref_frame}
+We first demonstrate the importance of measuring the Galactic midplane to very
+high accuracy. We explore the consequence of an incorrect midplane in the
+context of computing orbital actions of disk stars. An inaccuracy in the
+midplane is only one potential systematic error in the way actions are
+presently computed. The assumed axisymmetric model of the Galaxy may be a poor
+description of the Galactic potential, and the parameters in this axisymmetric
+model may be estimated incorrectly.
+
+In this work, we assume both that an axisymmetric model of the Galaxy is a good description of the potential and that we have correctly determined the parameters of that model.
+We are only concerned with an inaccuracy
+
+\subsection{Summary of Action Computation} \label{ssec:action_comp}
+We begin by summarizing one way in which actions are computed in the
+literature \Gus{cite}. One first assumes an axisymmetric model of the Galaxy.
+This model may itself be a poor description of the Galactic potential,
+
+We use the code \texttt{gala} v0.3 to perform orbit integrations and
+conversion to action space \citep{2017JOSS....2..388P,Price-Whelan:2018},
+which uses an action estimator technique presented by
+\citet{2014MNRAS.441.3284S}. Another common alternative is \texttt{galpy}
+\citep{2015ApJS..216...29B} using either direct orbit integration or the
+\Gus{Stackel Fudge} approximation. We use the default \texttt{MWPotential} as
+our potential, which is based on the Milky Way potential available in
+\texttt{galpy} \citep{2015ApJS..216...29B}. This potential includes a
+Hernquist bulge and nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
+\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
+is fit to empirically match some observations. We use the Dormand-Prince
+8(5,3) integration scheme \citep{Dormand80:integrator}. We use a timestep of
+$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
+Sun-like star.
+
+\subsection{Cartoon} \label{ssec:cartoon}
+We present a cartoon in Figure~\ref{fig:cartoon} to show that an inaccurate
+determination of the midplane can lead to a phase-dependence of the {\em
+observed} actions. The solid gray line indicates a hypothetical ``true''
+orbit. The y-axis corresponds to the vertical height of the orbit and the
+x-axis orbital phase. The dashed gray line is an incorrect midplane which an
+observer uses to compute orbits.
+
+Now suppose this observer makes a measurement of the true orbit at the orange
+point or the red point. Then the orange and red lines correspond to the orbits
+that the observer would compute for each point. In action space, this would
+correspond to a different value of $J_z$ for the orange and red points. In
+this way, assuming the wrong coordinate system induces time dependence in the
+actions {\em computed by an observer}.
+
+\begin{figure*}
+\plotone{fig/cartoon.pdf}
+\caption{Caption.}
+\label{fig:cartoon}
+\end{figure*}
+
+\subsection{Quantification} \label{ssec:quant}
+We now turn to classical orbital integration in a commonly assumed Milky Way
+potential to quantify the time variation of computed actions for different
+coordinate system errors. Our action computation procedure is described in
+Section~\ref{ssec:action_comp}.
+
+First, we illustrate the time-dependence of one possible inaccurate reference
+frame. We begin by taking an orbit with initial position $(8, 0, 0)\,\kpc$ and
+initial velocity $(0, -190, 50)\,\kms$ --- this corresponds to an orbit with
+actions $(J_r, L_z, J_z) = (32.5, -1520, 23.0)\,\actunit$. The orbit is
+plotted in \Gus{Appendix x}. We then suppose that an observer makes
+observations of this orbit in a coordinate system that is incorrect in radius by
+$200\,\pc$ and height by $100\,\pc$ --- we subtract the vector $(200, 0,
+100)\,\pc$ from each position in the orbit. This corresponds to an observer
+physically located at the position $(8, 0, 0)\,\kpc$ but erroneously thinking
+they are located at $(8.2, 0, 0.1)\,\kpc$.
+
+Using the observations of the orbit that this erroneous observer makes every
+$\Myr$\footnote{In addition to being erroneous, we also assume our fictitious
+observer is immortal.}, we perform the same orbit integration procedure as
+before to compute the actions our observer would make in the incorrect
+coordinate system over the duration of the orbit. The erroneous computed
+actions are shown for the first $\Gyr$ of the orbit in
+Figure~\ref{fig:one_orbit_wrong_ref}. Occasionally the numerical scheme fails
+and very large actions are reported by \texttt{gala}
+--- we perform a $5\sigma$ clip on each action to exclude such orbits, but
+this only excludes a total of $5$ orbits out of the $1000$ considered for
+Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
+vast majority of orbits are computed properly.
+
+Figure~\ref{fig:one_orbit_wrong_ref} shows that the erroneous observer would
+infer significant time evolution in their computed actions. The
+95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $8.7$,
+$47.1$, and $19.3\,\actunit$ for $J_r$, $L_z$, and $J_z$, respectively. These
+are fractional errors of $26.5\%$, $3.1\%$, and $81.4\%$, respectively.
+
+\begin{figure*}
+\plotone{fig/one_orbit.pdf}
+\caption{Caption.}
+\label{fig:one_orbit_wrong_ref}
+\end{figure*}
+
+We now repeat the procedure but continuously changing the error in the assumed
+midplane. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth minus
+5\uth
+
+\begin{figure*}
+% \plotone{fig/something.pdf}
+\caption{Caption.}
+\label{fig:many_orbit_wrong_ref}
+\end{figure*}
 
@@
 at the solar radius of $8.2\,\kpc$ and azimuth $\phi$, and consider stars
-within a cylinder perpendicular to the fiducial disk
+within a cylinder perpendicular to the fiducial disk.
 
@@
 To allow for potential small inaccuracies in the fiducial coordinate systems,
-we also subtract the best fit $\text{A} \sin{\phi + \text{B}} + C$ from the
-midplane as a function of azimuth. For $\text{A}$, the values are $-165$,
+we also subtract the best fit $\text{A} \sin{\phi + \text{B}} + C$ curve from
+the midplane as a function of azimuth. For $\text{A}$, the values are $-165$,
 $45$, $9\,\pc$ and for $\text{C}$ the values are $-69$, $19$, and $-18\,\pc$
@@
 
-\subsection{Dynamical Midplane}
-We now seek to define the midplane in a dynamically consistent manner, which
-we refer to as the ``dynamical midplane''. By this, we mean we would like to
-define the midplane in a manner consistent with some sort of dynamical
-constraint. Our method, which we describe now, is just one of many possible
-ways the dynamical midplane can be defined.
+% \subsection{Dynamical Midplane}
+% We now seek to define the midplane in a dynamically consistent manner, which
+% we refer to as the ``dynamical midplane''. By this, we mean we would like to
+% define the midplane in a manner consistent with some sort of dynamical
+% constraint. Our method, which we describe now, is just one of many possible
+% ways the dynamical midplane can be defined.
 
-We define the dynamical midplane based on the principle that perfectly
-circular, planar orbits will trace the midplane. Once the circular, planar
-orbits have been identified, the procedure for finding the dynamical midplane
-is relatively simple. From the measurement of each star's orbit, the midplane
-of the galaxy can be computed at $z=0$ easily. Furthermore, we will find that
-each orbit also gives a measure of the: dynamical center of the galaxy,
-velocity of the galaxy, and acceleration of the galaxy.
+% We define the dynamical midplane based on the principle that perfectly
+% circular, planar orbits will trace the midplane. Once the circular, planar
+% orbits have been identified, the procedure for finding the dynamical midplane
+% is relatively simple. From the measurement of each star's orbit, the midplane
+% of the galaxy can be computed at $z=0$ easily. Furthermore, we will find that
+% each orbit also gives a measure of the: dynamical center of the galaxy,
+% velocity of the galaxy, and acceleration of the galaxy.
 
-We pause briefly to consider general difficulties in measuring stellar orbits
-in a cosmological simulation. It may seem reasonable to place the galaxy in an
-inertial reference frame and then to simply measure the positions of each star
-particle over time. However, it is not generally the case that the center of
-mass of the stellar disk experiences a vanishing acceleration. Indeed, in a
-re-run of the simulations presented here\footnote{The re-run contained a
-cosmic ray bug fix which is unimportant here, but the re-run also output the
-acceleration of each particle, which is not the case for the simulations in
-this work.}, we measure a total acceleration of $\sim 0.4\,\kms/\text{Myr}$
-for stars with cylindrical radii within $0.5\,\kpc$ of $8.2\,\kpc$ and
-vertical heights less than $0.4\,\kpc$. Assuming a constant acceleration, over
-$22\,\Myr$ --- the time spacing of snapshots considered in this work --- this
-results in a $\sim90\,\pc$ shift of the disk. Since the offset scales with
-$t^2$, this presents a significant problem for measuring stellar orbits over
-longer timescales.
+% We pause briefly to consider general difficulties in measuring stellar orbits
+% in a cosmological simulation. It may seem reasonable to place the galaxy in an
+% inertial reference frame and then to simply measure the positions of each star
+% particle over time. However, it is not generally the case that the center of
+% mass of the stellar disk experiences a vanishing acceleration. Indeed, in a
+% re-run of the simulations presented here\footnote{The re-run contained a
+% cosmic ray bug fix which is unimportant here, but the re-run also output the
+% acceleration of each particle, which is not the case for the simulations in
+% this work.}, we measure a total acceleration of $\sim 0.4\,\kms/\text{Myr}$
+% for stars with cylindrical radii within $0.5\,\kpc$ of $8.2\,\kpc$ and
+% vertical heights less than $0.4\,\kpc$. Assuming a constant acceleration, over
+% $22\,\Myr$ --- the time spacing of snapshots considered in this work --- this
+% results in a $\sim90\,\pc$ shift of the disk. Since the offset scales with
+% $t^2$, this presents a significant problem for measuring stellar orbits over
+% longer timescales.
 
-Another option is to recenter and reorient the galaxy at each snapshot.
-However, the centering and orientation procedures are not without
-stochasticity themselves. This stochasticity can easily dominate the magnitude
-of the orbit. For instance, a shift in the direction of the $z$-axis by a mere
-$0.7\degree$ results in a shift in the midplane of $100\,\pc$ at a radius of
-$8.2\,\kpc$. We attempted this procedure in various ways (i.e. exploring
-different centering and orientation procedures), but were not able to cleanly
-measure the orbits of disk stars.
+% Another option is to recenter and reorient the galaxy at each snapshot.
+% However, the centering and orientation procedures are not without
+% stochasticity themselves. This stochasticity can easily dominate the magnitude
+% of the orbit. For instance, a shift in the direction of the $z$-axis by a mere
+% $0.7\degree$ results in a shift in the midplane of $100\,\pc$ at a radius of
+% $8.2\,\kpc$. We attempted this procedure in various ways (i.e. exploring
+% different centering and orientation procedures), but were not able to cleanly
+% measure the orbits of disk stars.
 
-Here we choose to simplify the problem by considering only perfectly circular,
-planar orbits. Since these orbits have no phase-dependent structure, we can
-work on shorter timescales, i.e. for orbits with significant radial or
-vertical action a large fraction of their orbit would need to be sampled in
-order to measure the orbit accurately. Because we know the analytic form of
-these orbits, we can forward model the trajectory of each star particle in the
-cosmological simulation. Working in comoving coordinates, we can write down
-the trajectory of a star particle as
+% Here we choose to simplify the problem by considering only perfectly circular,
+% planar orbits. Since these orbits have no phase-dependent structure, we can
+% work on shorter timescales, i.e. for orbits with significant radial or
+% vertical action a large fraction of their orbit would need to be sampled in
+% order to measure the orbit accurately. Because we know the analytic form of
+% these orbits, we can forward model the trajectory of each star particle in the
+% cosmological simulation. Working in comoving coordinates, we can write down
+% the trajectory of a star particle as
+% \beq\label{eq:helix}
+% \bm{r}(t) = R(\theta, \phi, \psi) \left[ (A\cos{\omega t}, B\sin{\omega t}, C
+% t) + \bm{x_0} \right]\text{,}
+% \eeq
+% where $R(\theta, \phi, \psi)$ is an Euler angle transformation, $A$, $B$, $C$,
+% and $\omega$ are arbitrary constants, and $\bm{x_0}$ is an arbitrary 3-vector.
+% This equation is a helix with an arbitrary constant and rotation. It describes
+% the
 
-\beq
-\bm{r}(t) = R(\theta, \phi, \psi) \left( (A\cos{\omega t}, B\sin{\omega t}, C t) + \bm{x_0} \right)
-\eeq
+% We minimize the summed squared difference between Equation~\ref{eq:helix} and
+% the true trajectory of each star with $R$ within $1\,\kpc$ of $R_{\odot} =
+% 8.2\,\kpc$ in each simulation. We consider minimum rms error (i.e. the typical
+% error in the Equation~\ref{eq:helix} model in describing the true trajectory).
+% This rms error characterizes how well the helix model we consider, which would
+% be exactly correct if the overall motion of the galaxy is well-described by
+% constant speed motion and the orbit of a particular store is exactly circular
+% and planar.
 
-\section{Coordinate System Gone Wrong} \label{sec:ref_frame}
-We show and quantify the time evolution of measured actions that comes about
-from inaccuracies in the galactocentric coordinate system.
+% We plot a histogram of the rms error for each galaxy in
+% Figure~\ref{fig:something}. As we estimated before, the total effect of the
+% galaxy's acceleration is roughly $90\,\pc$ (\Gus{Check for m12f, m12m}). We
+% find that this is small compared to the typical error, even for the stars
+% whose trajectories are best modeled by Equation~\ref{eq:helix}.
 
-\subsection{Cartoon} \label{ssec:cartoon}
-We begin with a cartoon in Figure~\ref{fig:cartoon} to show that an inaccurate
-determination of The solid gray line indicates a hypothetical ``true'' orbit.
-The y-axis corresponds to the vertical height of the orbit and the x-axis to
-either time or orbital phase. The dashed gray line is an incorrect midplane
-which an observer uses to compute orbits.
+% There are many reasons for the failure of this model. First, there could be no
+% stars on circular orbits. This seems unlikely considering that typical
+% distribution functions peak on perfectly circular, planar orbits
+% \citep[e.g.][]{2013MNRAS.434..652T}. Second, the trajectory of the galaxy may
+% not be well-described by constant-speed motion. This also seems unlikely,
+% considering that the bulk acceleration of the stellar disk only results in a
+% $\sim90\,\pc$ (\Gus{m12f, m12m}) deviation from constant-speed motion over the
+% $22\,\Myr$ time we considered.
 
-Now suppose this observer makes a measurement of the true orbit at the orange
-point or the red point. Then the orange and red lines correspond to the orbits
-that the observer would compute for each point. In action space, this would
-correspond to a different value of $J_z$ for the orange and red points. In
-this way, assuming the wrong coordinate system induces time dependence in the
-actions {\em computed by an observer}.
+% Furthermore, in Figure~\ref{fig:something} we plot the center of mass of the
+% galaxy (in comoving coordinates) for the $22\,\Myr$ timeframe described. The
+% center of mass at each snapshot is computed using the iterative shrinking
+% spheres method described in Section~\ref{ssec:cosmozoom}. The rms error on the
+% constant-speed motion is $\sim\text{something}$, which we interpret as
+% \Gus{something}.
 
-\begin{figure*}
-\plotone{fig/cartoon.pdf}
-\caption{Caption.}
-\label{fig:cartoon}
-\end{figure*}
-
-\subsection{Quantification} \label{ssec:quant}
-We now turn to classical orbital integration in a commonly assumed Milky Way
-potential to quantify the time variation of computed actions for different
-coordinate system errors.
-
-We use the code \texttt{gala} v0.3 to perform orbit integrations and
-conversion to action space \citep{2017JOSS....2..388P,Price-Whelan:2018},
-which uses an action estimator technique presented by
-\citet{2014MNRAS.441.3284S}. We use the default \texttt{MWPotential} as our
-potential, which is based on the Milky Way potential available in
-\texttt{galpy} \citep{2015ApJS..216...29B}. This potential includes a
-Hernquist bulge and nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
-\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
-is fit to empirically match some observations. We use the Dormand-Prince
-8(5,3) integration scheme \citep{Dormand80:integrator}. We use a timestep of
-$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for
-a Sun-like star.
-
-First, we illustrate the time-dependence of one possible inaccurate reference
-frame. We begin by taking an orbit with initial position $(8, 0, 0)\,\kpc$ and
-initial velocity $(0, -190, 50)\,\kms$ --- this corresponds to an orbit with
-actions $(J_r, L_z, J_z) = (32.5, -1520, 23.0)\,\actunit$. The orbit is
-plotted in \Gus{Appendix x}. We then suppose that an observer makes
-observations of this orbit in a coordinate system that is incorrect in radius by
-$200\,\pc$ and height by $100\,\pc$ --- we subtract the vector $(200, 0,
-100)\,\pc$ from each position in the orbit. This corresponds to an observer
-physically located at the position $(8, 0, 0)\,\kpc$ but erroneously thinking
-they are located at $(8.2, 0, 0.1)\,\kpc$.
-
-Using the observations of the orbit that this erroneous observer makes every
-$\Myr$, we perform the same orbit integration procedure as before to compute
-the actions our observer would make in the incorrect coordinate system over the
-duration of the orbit. The erroneous computed actions are shown for the first
-$\Gyr$ of the orbit in Figure~\ref{fig:one_orbit_wrong_ref}. Occasionally the
-numerical scheme fails and very large actions are reported by \texttt{gala}
---- we perform a $5\sigma$ clip on each action to exclude such orbits, but
-this only excludes a total of $5$ orbits out of the $1000$ considered for
-Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
-vast majority of orbits are computed properly.
-
-Figure~\ref{fig:one_orbit_wrong_ref} shows that the erroneous observer would
-infer significant time evolution in their computed actions. The
-95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $8.7$,
-$47.1$, and $19.3\,\actunit$ for $J_r$, $L_z$, and $J_z$, respectively. These
-are fractional errors of $26.5\%$, $3.1\%$, and $81.4\%$, respectively.
-
-\begin{figure*}
-\plotone{fig/one_orbit.pdf}
-\caption{Caption.}
-\label{fig:one_orbit_wrong_ref}
-\end{figure*}
-
-We now
+% One potential explanation is that {\em no star can remain on a circular,
+% planar orbit}. The order unity density variations may be strong enough such
+% that each star's orbit is perturbed. If there are truly stars in the
+% simulation on perfectly circular, planar orbits, then the magnitude of their
+% acceleration over the last $22\,\Myr$ must be constant. Furthermore, their
+% trajectory in acceleration space must be well-described by a rotation. In the re-runs,
 

### 2018-12-15 paper/ms.tex
Update paper

@@
 they were situated in each of these galaxies. We place our imaginary observer
-at the solar radius of $8.2\,\kpc$ and azimuth $\phi$, and consider stars
-within a cylinder perpendicular to the fiducial disk.
+at the solar radius of $8.2\,\kpc$ and azimuth $\phi$\footnote{Our erroneous,
+immortal observer is also warp-drive capable.}, and consider stars within a
+cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to the
+fiducial disk.
 
-For a given azimuth $\phi$ at the solar radius of $8.2\,\kpc$, we take the
-median height of all star particles within a cylinder of radius $0.5\,\kpc$
-and total height $2\,\kpc$. This median height is taken to be what our
-imaginary observer would measure as the local Galactic midplane, and is
-reported for each galaxy in Figure~\ref{fig:midplane}. We find that only $10$
-iterations of this procedure are necessary for convergence. We bootstrap
-resample $1000$ times and report the $95^{\text{th}}$ minus $5^{\text{th}}$
-percentile as the dashed-line error bars.
+This median height is taken to be what our observer would measure as the local
+Galactic midplane at each $\phi$, and is reported for each galaxy in
+Figure~\ref{fig:midplane}. We find that only $10$ iterations of this procedure
+are necessary for convergence. We bootstrap resample $1000$ times and report
+the $95^{\text{th}}$ minus $5^{\text{th}}$ percentile as the dashed-line error
+bars. This bootstrap resampling is performed by first selecting all stars
+within a height of $2\,\kpc$, resampling that selection, and then repeating
+the $10$ iterations.
 
 To allow for potential small inaccuracies in the fiducial coordinate systems,
-we also subtract the best fit $\text{A} \sin{\phi + \text{B}} + C$ curve from
-the midplane as a function of azimuth. For $\text{A}$, the values are $-165$,
-$45$, $9\,\pc$ and for $\text{C}$ the values are $-69$, $19$, and $-18\,\pc$
-for m12i, m12f, and m12m, respectively. For the assumed solar radius of
-$8.2\,\kpc$, we can approximate the angle offset for the $z$-axis from the
-values of $\text{A}$ --- we compute $1.15$, $0.31$, and $0.06\,\deg$ for m12i,
-m12f, and m12m. These angle offsets are consistent with the values given in
-\citet{2018arXiv180610564S} for the difference between the $z$-axis as defined
-by the gas and stars.
+we also subtract the best fit $\text{A} \sin{\left(\phi + \text{B}\right)} +
+\text{C}$ curve from the midplane as a function of azimuth. For $\text{A}$,
+the values are $-165$, $45$, $9\,\pc$ and for $\text{C}$ the values are $-69$,
+$19$, and $-18\,\pc$ for m12i, m12f, and m12m, respectively \Gus{Double check
+numbers}. For the assumed solar radius of $8.2\,\kpc$, we can approximate the
+angle offset for the $z$-axis from the values of $\text{A}$ --- we compute
+\Gus{$1.15$, $0.31$, and $0.06\,\deg$} for m12i, m12f, and m12m. These angle
+offsets are consistent with the values given in \citet{2018arXiv180610564S}
+for the difference between the $z$-axis as defined by the gas and stars.
 
@@
 \end{figure*}
-
-Our analysis makes the assumption that a global midplane is well defined.
-However, our results are also consistent with the notion that the global
-midplane is not a well-defined concept.
 

### 2018-12-19 paper/ms.tex

@@
 \newcommand{\z}{z}
-\newcommand{\uth}{\textsuperscript{th}} % why is this so long in the first place??
+\newcommand{\uth}{\textsuperscript{th}}
 
@@
 % affiliations
-\newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute, 162 5th
-Ave., New York, NY 10010, USA}
-\newcommand{\penn}{Department of Physics \& Astronomy, University of Pennsylvania, 209
-South 33rd St., Philadelphia, PA 19104, USA}
+\newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute,
+162 5th Ave., New York, NY 10010, USA}
+\newcommand{\penn}{Department of Physics \& Astronomy, University of
+Pennsylvania, 209 South 33rd St., Philadelphia, PA 19104, USA}
 \newcommand{\amnh}{Department of Astrophysics, American Museum of Natural
@@
 
-Underlying all of these discoveries is the assumption of a galactocentric
+Underlying all of these discoveries is the assumption of a Galactocentric
 coordinate system \citep{2008gady.book.....B}. In angular
@@
 assumption of an axisymmetric Milky Way. Under this assumption, these
-estimates would provide the correct parameters for the galactocentric
+estimates would provide the correct parameters for the Galactocentric
 coordinate system up to measurement error. However, the Milky Way is not
@@
 Since the Milky Way is not truly axisymmetric, it is not clear the best way to
-define the Galactocentric coordinate system. In this work, we focus on the midplane and, equivalently, the solar height.
+define the Galactocentric coordinate system --- or, perhaps, the best way to
+modify the current implicit definition of the Galactocentric coordinate system
+in order to account for non-axisymmetries.
 
@@
 will be additional time evolution in the measured actions.
+
+The transformation to action space is not trivial, and many of the assumptions
+of the transformation are potentially incorrect. It is therefore natural to
+consider the reason to transform to action space. One reason is dimensionality
+reduction --- in the action space paradigm, a stellar orbit is completely
+described by three actions, as opposed to six dimensions of phase space. This
+means that the relationship between {\em orbital} properties of individual
+stars and stellar properties such as age or metallicity can be investigated.
 
@@
 
-In this work, we assume both that an axisymmetric model of the Galaxy is a good description of the potential and that we have correctly determined the parameters of that model.
-We are only concerned with an inaccuracy
+In this work, we assume both that an axisymmetric model of the Galaxy is a
+good description of the potential and that we have correctly determined the
+parameters of that model. We are only concerned with an inaccuracy
 
@@
 
+\section{Discussion} \label{sec:discussion}
+{\em Gaia} DR2 has ushered in a new era of Galactic dynamics. High-precision
+phase space measurements have been made for an unprecedented number of stars.
+In the face of such large data, it is worth considering the goal of such
+measurements --- a goal which should inform the usage of such data.
+
+In this work we considered one such question: given the current position in
+phase space of a single star, what can we infer about that star's past phase
+space trajectory? Current methodology involves using a smoothed, analytic form
+for the Galactic potential (with perhaps crude implementations of spiral arms
+or sattelites), fitting such form to some selection of observations, and
+integrating from the currently measured position in phase space. This orbit
+integration is also commonly accompanied with a transformation into
+action-angle space. Other than the integration, every step of that process
+should be viewed with significant scrutiny.
+
+The more general problem of measuring the orbits of stars in cosmological
+simulations is essentially unexplored in this work and the literature, and we
+delay discussion to future work.
+
 % \subsection{Dynamical Midplane}

### 2018-12-19 paper/ms.tex
Add many orbits plot to paper

@@
 midplane. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth minus
-5\uth
+5\uth percentile divided by the true action value for three different orbits.
+The red orbit is exactly the same as in Figure~\ref{fig:one_orbit_wrong_ref},
+and we also consider two more planar orbits with initial vertical velocities
+of $30\,\kms$ (\Gus{orange}) and $10\,\kms$ (\Gus{blue}). For these two new
+orbits, the true actions are $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$
+and $(40.4, -1520, 0.7)\,\actunit$ for the \Gus{orange} and \Gus{blue} orbits,
+respectively. \Gus{Easy way to explain change in $J_r$?}
 
 \begin{figure*}
-% \plotone{fig/something.pdf}
+\plotone{fig/many_orbits_schmactions.pdf}
 \caption{Caption.}

### 2018-12-19 schmactions/plot_many_orbit.py
Make Lz bounds more intepretable

@@ -38,7 +38,7 @@ max_offset = 500
 d_offset = 50
 
 JrIQR_max = 100
-LzIQR_max = 10
+LzIQR_max = 100
 JzIQR_max = 200
 
 mw = gp.MilkyWayPotential()

### 2018-12-19 schmactions/plot_many_orbit.py
Setup orbit list for ipython inspection

@@ -53,6 +53,11 @@ def compute_actions(pos=None, vel=None, phase=None):
     ans = res['actions'].to(u.kpc * u.km / u.s).value
     return ans
 
+def compute_orbit(pos, vel):
+    q = gd.PhaseSpacePosition(pos=pos, vel=vel)
+    orbit = mw.integrate_orbit(q, dt=dt, t1=t1, t2=t2, Integrator=gi.DOPRI853Integrator)
+    return orbit
+
 
 def compute_actions_wrong_ref_frame(init_pos, init_vel, offset, cadence=25, wrong_max=None):
     q = gd.PhaseSpacePosition(pos=init_pos, vel=init_vel)
@@ -217,4 +222,10 @@ if __name__ == '__main__':
 
     save_fig(fig, ax, 'many_orbits_schmactions.pdf', true_act=True)
 
+    # examine orbits, I know this is hacky, just want orbits loaded in ipython
+    init_pos = [8, 0, 0] * u.kpc
+    init_vel_list = [[0, -190, 10] * u.km/u.s,
+                    [0, -190, 30] * u.km/u.s,
+                    [0, -190, 50] * u.km/u.s]
 
+    orbit_list = [compute_orbit(init_pos, init_vel) for init_vel in init_vel_list]

### 2018-12-19 paper/ms.tex
Add zmax to text
Add text to describe many orbit fig

@@
 actions $(J_r, L_z, J_z) = (32.5, -1520, 23.0)\,\actunit$. The orbit is
-plotted in \Gus{Appendix x}. We then suppose that an observer makes
-observations of this orbit in a coordinate system that is incorrect in radius by
-$200\,\pc$ and height by $100\,\pc$ --- we subtract the vector $(200, 0,
-100)\,\pc$ from each position in the orbit. This corresponds to an observer
-physically located at the position $(8, 0, 0)\,\kpc$ but erroneously thinking
-they are located at $(8.2, 0, 0.1)\,\kpc$.
+plotted in \Gus{Appendix x}. For reference, the $z_{\text{max}}$ for this
+orbit is $854\,\pc$. We then suppose that an observer makes observations of
+this orbit in a coordinate system that is incorrect in radius by $200\,\pc$
+and height by $100\,\pc$ --- we subtract the vector $(200, 0, 100)\,\pc$ from
+each position in the orbit. This corresponds to an observer physically located
+at the position $(8, 0, 0)\,\kpc$ but erroneously thinking they are located at
+$(8.2, 0, 0.1)\,\kpc$.
 
@@
 and $(40.4, -1520, 0.7)\,\actunit$ for the \Gus{orange} and \Gus{blue} orbits,
-respectively. \Gus{Easy way to explain change in $J_r$?}
+respectively. The value of $z_{\text{max}}$ for these orbits are $122\,\pc$
+and $425\,\pc$, respectively (with the blue orbit having $z_{\text{max}}$ of
+$854\,\pc$.) \Gus{Easy way to explain change in $J_r$?}
+
+The top row of Figure~\ref{Fig:many_orbit_wrong_ref} shows the error in each
+action assuming an offset in the midplane. In the {\em bottom row} we also
+consider offsets in the Galactocentric radius $R$. The {\em left}, {\em
+middle}, and {\em right} columns show the 95\uth minus 5\uth percentile of the
+measured action over the course of the incorrect orbit divided by the true
+action value as a percentage.
+
+In the {\em upper middle} panel, there is essentially no error in the
+determination of $L_z$. This is expected since, in principle, $L_z$ only
+depends on the $x$- and $y$-components of the position and velocity of the
+stars\footnote{Though this is not how we compute $L_z$ in practice.}, and is
+thus unaffected by offsets in $z$.
+
+The {\em upper right} panel shows that the fractional error in $J_z$ is more
+exaggerated for more planar orbits. For the nearly planar (blue) orbit,
+deviations that are $\sim30\,\pc$ already give $100\%$ deviations in the
+actions. The required offset for $100\%$ deviation is $\sim120$ and
+$\sim225\,\pc$ for the orange and red orbits, respectively. \Gus{Make these
+numbers exact.}
+
+For completeness, we also consider the effect of offsets in the Galactocentric
+radius $R$ in the {\em bottom panels}. The error is largest for $J_r$, with
+some deviations resulting in $L_z$ and relatively small deviations in $J_z$.
+In the {\em bottom left} panel the orange and blue lines overlap, and in the
+{\em bottom middle} and {\em bottom right} panels all three lines overlap.
 

### 2018-12-20 paper/ms.tex
Add to discussion section

@@
 \section{Discussion} \label{sec:discussion}
-{\em Gaia} DR2 has ushered in a new era of Galactic dynamics. High-precision
-phase space measurements have been made for an unprecedented number of stars.
-In the face of such large data, it is worth considering the goal of such
-measurements --- a goal which should inform the usage of such data.
+We have shown that, in simulation, the ``local midplane'' can have strong
+azimuthal dependence. The implications of this finding are not clear, but
+clearly demonstrate that assuming our local midplane is the same as the global
+midplane is insufficient for extracting the most dynamical information from
+the {\em Gaia} mission.
 
-In this work we considered one such question: given the current position in
-phase space of a single star, what can we infer about that star's past phase
-space trajectory? Current methodology involves using a smoothed, analytic form
-for the Galactic potential (with perhaps crude implementations of spiral arms
-or sattelites), fitting such form to some selection of observations, and
-integrating from the currently measured position in phase space. This orbit
-integration is also commonly accompanied with a transformation into
-action-angle space. Other than the integration, every step of that process
-should be viewed with significant scrutiny.
+A possible interpretation of our finding is that one should use a star's local
+midplane in order to compute dynamical quantities like the vertical action.
+Under the assumption that this interpretation is correct, we now consider the
+implications for dynamical studies of disrupting open clusters (the original
+motivation of this study).
 
-The more general problem of measuring the orbits of stars in cosmological
-simulations is essentially unexplored in this work and the literature, and we
-delay discussion to future work.
+Suppose an open cluster of mass $M_c$ on an orbit with actions $\bm{J}_c$ is
+disrupted, and that the cluster is currently at an orbit phase of
+$\bm{\theta}_0$. We would like to determine all stars in the {\em Gaia}
+dataset\footnote{This is likely to be more successful with DR3/4 than DR2,
+since upcoming data releases will contain a far greater number of stars with
+radial velocities --- though targeted, follow-up radial velocity surveys may
+allow this program to be performed sooner.} that are likely to be ejected
+members from that cluster. One could proceed by integrating the cluster's
+orbit and selecting all stars within a certain distance of that orbit,
+accounting for the fact that the stream will not exactly follow the orbit of
+the cluster \citep[e.g.][]{2011MNRAS.413.1852E}.
+
+One could then make a further selection by requiring that the action of each
+field star $i$ be close to the cluster's actions to within some bound, i.e.
+$\bm{J}_i - \bm{J}_c \in V_e$ where $V_e$ is some volume in action space
+enclosing a region of acceptable error. A natural choice for $V_e$ is to
+center it at the origin and allow its extent to be the intrinsic action-space
+dispersion of a disrupted cluster. This dispersion can be estimated as
+\citep[\S~8.3.3][]{2008gady.book.....B}
+\beq \label{eq:action_disp}
+\frac{\Delta J_i}{J_i} \sim \left(\frac{m}{M}\right)^{1/3}\text{,}
+\eeq
+where $m$ is the mass of the cluster and $M$ is the mass enclosed by the
+cluster's orbit. A more sophisticated treatment of the action-space dispersion
+of disrupted clusters can be obtained by following
+\citet{2011MNRAS.413.1852E}, but such a treatment is premature for this work.
+
+The act of comparing a field star's action to a cluster's action (i.e. in
+computing $\bm{J}_i - \bm{J}_c$) is subject to significant error if the
+cluster and field star have significantly different local midplanes and an
+observer assumed a single global midplane.
+
+To explore the magnitude of this effect, we consider the following program.
+First, we assume a cluster mass $m$ and estimate the expected action-space
+dispersion from Equation~\ref{eq:action_disp}. We then consult
+Figure~\ref{fig:many_orbit_wrong_ref} to determine the $z$ offset necessary in
+order for the induced error in the vertical action to be the same as the
+expected action-space dispersion. If this were the case, an observer would
+have to account for the different midplane of the field star and the cluster.
+We repeat this analysis for each of the three orbits considered in
+Section~\ref{ssec:quant}.
+
+With a characteristic $z$ offset in hand, we consult Figure~\ref{fig:deltaphi}
+to convert this $z$ offset into a $\Delta \phi$, which we convert to an
+arclength or characteristic distance (the difference between the two will be
+negligible for the small $\Delta \phi$'s we will compute). Therefore, if an
+observer is attempting to reconstruct a cluster with a given mass on a given
+orbit, we will give a characteristic distance between a field star and the
+cluster within which an observer might not need to consider differences in
+their respective local midplanes.
+
+From the simulation we have the best handle on the $z$ offset
+(Figure~\ref{fig:midplane}), and this $z$ offset has the greatest impact on
+the vertical action $J_z$ (Figure~\ref{fig:many_orbit_wrong_ref}). Therefore,
+we only consider the impact of a $z$ offset on the vertical action $J_z$.
+
+% {\em Gaia} DR2 has ushered in a new era of Galactic dynamics. High-precision
+% phase space measurements have been made for an unprecedented number of stars.
+% In the face of such large data, it is worth considering the goal of such
+% measurements --- a goal which should inform the usage of such data.
+
+% In this work we considered one such question: given the current position in
+% phase space of a single star, what can we infer about that star's past phase
+% space trajectory? Current methodology involves using a smoothed, analytic form
+% for the Galactic potential (with perhaps crude implementations of spiral arms
+% or sattelites), fitting such form to some selection of observations, and
+% integrating from the currently measured position in phase space. This orbit
+% integration is also commonly accompanied with a transformation into
+% action-angle space. Other than the integration, every step of that process
+% should be viewed with significant scrutiny.
+
+% The more general problem of measuring the orbits of stars in cosmological
+% simulations is essentially unexplored in this work and the literature, and we
+% delay discussion to future work.
 

### 2018-12-21 paper/ms.tex
Add cluster offset to paper
Add captions to figures

@@
 
-\correspondingauthor{Angus Beane}
+% \correspondingauthor{Angus Beane}
 \email{abeane@sas.upenn.edu}
@@
 \plotone{fig/cartoon.pdf}
-\caption{Caption.}
+\caption{Cartoon showing the effect an incorrect midplane can have on orbit
+integrations and as an extension on action computations. The x-axis shows the
+orbital phase and the y-axis the vertical height. The top gray curve depicts a
+hypothetical true orbit oscillating about the true midplane, the horizontal
+solid gray line. Consider an immortal observer who erroneously assumes the
+midplane is located at the horizontal dashed line. Suppose this immortal
+observer measures the orbit at the blue and pink points. If the immortal
+observer were to perform orbit integrations at each observation, they would
+assume the blue and pink curves for the respective orbits. In this way, an
+incorrect midplane will induce phase-dependence in the {\em observed}
+actions.}
 \label{fig:cartoon}
@@
 \plotone{fig/one_orbit.pdf}
-\caption{Caption.}
+\caption{The artificial phase-dependence in the observed actions induced by an
+error in the Galactocentric coordinate system. We consider here an orbit with
+initial position $(8,0,0)\,\kpc$ and initial velocity $(0,-190,50)\,\kms$.
+This corresponds to an orbit with actions $(J_r, L_z, J_z) = (32.5, -1520,
+23.0)\,\actunit$ and $z_{\text{max}}=854\,\pc$. Considering the first $\Gyr$
+of this orbit, we assume an immortal observer }
 \label{fig:one_orbit_wrong_ref}
@@
 
+\begin{figure}
+\plotone{fig/cluster_offset.pdf}
+\caption{Caption.}
+\label{fig:cluster_offset}
+\end{figure}
+
+In Figure~\ref{fig:cluster_offset}, we show the $z$ offset necessary to induce
+the action space dispersion computed from Equation~\ref{eq:action_disp}, using
+the \texttt{gala} computed mass enclosed within a radius of $8.2\,\kpc$ ($\sim
+10^{11}\,[\,M_{\odot}\,]$).
+
 From the simulation we have the best handle on the $z$ offset

### 2018-12-21 paper/ms.tex
Make orbit par. tighter

@@
 initial velocity $(0, -190, 50)\,\kms$ --- this corresponds to an orbit with
-actions $(J_r, L_z, J_z) = (32.5, -1520, 23.0)\,\actunit$. The orbit is
-plotted in \Gus{Appendix x}. For reference, the $z_{\text{max}}$ for this
-orbit is $854\,\pc$. We then suppose that an observer makes observations of
-this orbit in a coordinate system that is incorrect in radius by $200\,\pc$
-and height by $100\,\pc$ --- we subtract the vector $(200, 0, 100)\,\pc$ from
-each position in the orbit. This corresponds to an observer physically located
-at the position $(8, 0, 0)\,\kpc$ but erroneously thinking they are located at
-$(8.2, 0, 0.1)\,\kpc$.
+actions $(J_r, L_z, J_z) = (32.5, -1520, 23.0)\,\actunit$ and
+$z_{\text{max}}=854\,\pc$. The orbit is plotted in \Gus{Appendix x}. We then
+suppose that an observer makes observations of this orbit in a coordinate
+system that is incorrect in radius by $200\,\pc$ and height by $100\,\pc$ ---
+we subtract the vector $(200, 0, 100)\,\pc$ from each position in the orbit.
+This corresponds to an observer physically located at the position $(8, 0,
+0)\,\kpc$ but erroneously thinking they are located at $(8.2, 0, 0.1)\,\kpc$.
 

### 2018-12-21 real_cluster/compute_real_cluster_actions.py
Add real cluster action calc

@@ -1,0 +1,59 @@
+import numpy as np 
+
+import gala.dynamics as gd
+import gala.potential as gp
+import gala.integrate as gi
+
+from pyia import GaiaData
+from astropy.table import Table 
+import astropy.units as u
+import astropy.coordinates as coord
+
+import warnings
+
+dt = 1 * u.Myr
+mw = gp.MilkyWayPotential()
+
+t = Table.read('J_A+A_616_A10_tablea3.dat.fits', format='fits')
+
+# rename stuff
+t['ra'] = t['RAdeg']
+t['dec'] = t['DEdeg']
+t['parallax'] = t['plx']
+t['pmra'] = t['pmRA']
+t['pmdec'] = t['pmDE']
+t['radial_velocity'] = t['RV']
+
+g = GaiaData(t)
+
+dist = 1/(g.parallax.to_value(u.mas)) * 1000 # pc
+name = t['Cluster'].tolist()
+m = np.c_[name, dist]
+
+rsun = 8.2 * u.kpc
+zsun = 0.025 * u.kpc
+vsun = [11.1, 232.24, 7.25] * u.km/u.s
+gc_frame = coord.Galactocentric(galcen_distance=rsun, galcen_v_sun=coord.CartesianDifferential(*vsun), z_sun=zsun)
+
+sc = g.skycoord
+scdyn = gd.PhaseSpacePosition(sc.transform_to(gc_frame).cartesian)
+
+def actions(star):
+    with warnings.catch_warnings(record=True):
+        warnings.simplefilter("ignore")
+        orbit = mw.integrate_orbit(star, dt=dt, t1=0*u.Gyr, t2=5*u.Gyr, Integrator=gi.DOPRI853Integrator)
+        res = gd.actionangle.find_actions(orbit, N_max=8)
+        ans = res['actions'].to(u.kpc * u.km / u.s).value
+        return np.append(ans,orbit.zmax(approximate=True).to(u.pc).value)
+
+result = [actions(s) for s in scdyn]
+result2 = np.append((m, result), axis=1)
+r = Table(result2, names=('cluster', 'distance', 'Jr', 'Lz', 'Jz', 'zmax'))
+r['distance'].unit = u.pc
+r['Jr'].unit = u.kpc * u.km/u.s
+r['Lz'].unit = u.kpc * u.km/u.s
+r['Jz'].unit = u.kpc * u.km/u.s
+r['zmax'].unit = u.pc
+
[…]

### 2018-12-22 real_cluster/compute_real_cluster_actions.py
Update real cluster actions output

@@ -47,7 +47,7 @@ def actions(star):
         return np.append(ans,orbit.zmax(approximate=True).to(u.pc).value)
 
 result = [actions(s) for s in scdyn]
-result2 = np.append((m, result), axis=1)
+result2 = np.concatenate((m, result), axis=1)
 r = Table(result2, names=('cluster', 'distance', 'Jr', 'Lz', 'Jz', 'zmax'))
 r['distance'].unit = u.pc
 r['Jr'].unit = u.kpc * u.km/u.s

### 2018-12-23 midplane_test/plot_deltaphi.py
Output range vs phi data

@@ -84,6 +84,8 @@ for gal,ax_col in zip(glist, ax.transpose()):
     ax_col.plot(dphi_mean/np.pi, r_up*1000, alpha=0.75, ls='dashed', c=tb_c[0])
     ax_col.plot(dphi_mean/np.pi, r_low*1000, alpha=0.75, ls='dashed', c=tb_c[0])
 
+    out = np.hstack((dphi_mean, r_mean, r_sigma))
+    np.save('r_vs_dphi_'+gal+'.npy', out)
 
     # chord length
     def tick_function(l):

### 2018-12-23 paper/ms.tex
Swearing is bad
Update paper

@@
 \shortauthors{Beane et al.}
-\shorttitle{Where the Hell Are We?}
+\shorttitle{Our Galactic Midplane}
 
@@
 \newcommand{\abs}[1]{\left| #1 \right|}
-\newcommand{\z}{z}
+\newcommand{\z}{z_r}
 \newcommand{\uth}{\textsuperscript{th}}
+\newcommand{\n}{\text{n}}
 
@@
 0.0455$, $h = 0.702$, $\sigma_8 = 0.807$, and $n_s = 0.961$, consistent with
-current constraints \citep{2018arXiv180706209P}. Halos are selected at $z=0$
-based solely on their mass ($\sim 1\text{--}2 \times 10^{12} \Msun$).
-Particles within $5 R_{200\text{m}}$ of each halo are traced back to $\z=99$
-and the initial conditions are regenerated at higher resolution using MUSIC
-\citep{2011MNRAS.415.2101H}.
+current constraints \citep{2018arXiv180706209P}. Halos are selected at $\z=0$
+based solely on their mass ($\sim 1\text{--}2 \times 10^{12} \Msun$). In this
+work, to avoid confusion with the vertical height $z$, we refer to
+cosmological redshift as $\z$. Particles within $5 R_{200\text{m}}$ of each
+halo are traced back to $\z=99$ and the initial conditions are regenerated at
+higher resolution using MUSIC \citep{2011MNRAS.415.2101H}.
 
@@
 
+To gain further handle on the midplane behavior as a function of azimuth, we
+compute the range of midplane values for a given $\Delta \phi$ segment of the
+midplane in Figure~\ref{fig:range_deltaphi}. With the observer's azimuth set
+to zero, we consider a region extending from $-\Delta \phi/2$ to
+$\Delta\phi/2$. We also compute the chord length for the chord extending from
+the observer to $\pm \Delta\phi/2$. We repeat this for each of the $50$ bins
+in azimuth. Each bin is plotted as a gray line, with the mean value in solid
+blue and the $1\,\sigma$ dispersion as the dashed blue lines.
+
+\begin{figure*}
+\plotone{fig/range_dphi.pdf}
+\caption{Caption.}
+\label{fig:rand_deltaphi}
+\end{figure*}
+
 A possible interpretation of our finding is that one should use a star's local
@@
 \beq \label{eq:action_disp}
-\frac{\Delta J_i}{J_i} \sim \left(\frac{m}{M}\right)^{1/3}\text{,}
+\frac{\Delta J_i}{J_i} \sim \left(\frac{m_c}{M}\right)^{1/3}\text{,}
 \eeq
-where $m$ is the mass of the cluster and $M$ is the mass enclosed by the
+where $m_c$ is the mass of the cluster and $M$ is the mass enclosed by the
 cluster's orbit. A more sophisticated treatment of the action-space dispersion
@@
 the \texttt{gala} computed mass enclosed within a radius of $8.2\,\kpc$ ($\sim
-10^{11}\,[\,M_{\odot}\,]$).
+10^{11}\,M_{\odot}$).
 
-From the simulation we have the best handle on the $z$ offset
-(Figure~\ref{fig:midplane}), and this $z$ offset has the greatest impact on
-the vertical action $J_z$ (Figure~\ref{fig:many_orbit_wrong_ref}). Therefore,
-we only consider the impact of a $z$ offset on the vertical action $J_z$.
+We now convert the $z$~offset-$m_c$ relation to an $R_{\n}$-$m_c$ relation,
+where $R_{\n}$ indicates a sort of ``solar neighborhood'' radius. An observer
+attempting to reconstruct an open cluster of mass $m_c$ with perfect knowledge
+of the midplane at the location of the cluster would expect to find the
+ejected members using a dynamical cut for field stars within a distance
+$R_{\n}$ of the cluster. If such an observer wants to find ejected stars that
+are at a distance greater than $R_{\n}$, a more sophisticated analysis would
+be required, the details of which are not apparent to the authors of this
+work.
+
+We use the relation in Figure~\ref{fig:deltaphi}, which shows the typical $z$
+offset at each $\Delta \phi$, to convert a given $z$ offset into a $\Delta
+\phi$, which we then convert into a chord length which we interpret as
+$R_{\n}$.
 

### 2018-12-26 paper/ms.tex
Update paper

@@
 \caption{Caption.}
-\label{fig:rand_deltaphi}
+\label{fig:range_deltaphi}
 \end{figure*}
@@
 
-With a characteristic $z$ offset in hand, we consult Figure~\ref{fig:deltaphi}
-to convert this $z$ offset into a $\Delta \phi$, which we convert to an
-arclength or characteristic distance (the difference between the two will be
-negligible for the small $\Delta \phi$'s we will compute). Therefore, if an
-observer is attempting to reconstruct a cluster with a given mass on a given
-orbit, we will give a characteristic distance between a field star and the
-cluster within which an observer might not need to consider differences in
-their respective local midplanes.
+With a characteristic $z$ offset in hand, we consult
+Figure~\ref{fig:range_deltaphi} to convert this $z$ offset into a $\Delta
+\phi$, which we convert to an arclength or characteristic distance (the
+difference between the two will be negligible for the small $\Delta \phi$'s we
+will compute). Therefore, if an observer is attempting to reconstruct a
+cluster with a given mass on a given orbit, we will give a characteristic
+distance between a field star and the cluster within which an observer might
+not need to consider differences in their respective local midplanes.
 
@@
 
-We use the relation in Figure~\ref{fig:deltaphi}, which shows the typical $z$
-offset at each $\Delta \phi$, to convert a given $z$ offset into a $\Delta
-\phi$, which we then convert into a chord length which we interpret as
-$R_{\n}$.
+We use the relation in Figure~\ref{fig:range_deltaphi}, which shows the
+typical $z$ offset at each $\Delta \phi$, to convert a given $z$ offset into a
+$\Delta\phi$, which we then convert into a chord length which we interpret as
+$R_{\n}$. The result of this procedure is shown in Figure~\ref{fig:Rn_mc}.
+
+\begin{figure*}
+\plotone{fig/}
+\end{figure*}
 

### 2018-12-26 midplane_test/midplane.py
Fix concatenate bug

@@ -133,7 +133,7 @@ def main(gal):
     print('A=', A, 'B=', B, 'C=', C)
     fit = A*np.cos(theta + B) + C
 
-    out = np.concatenate((theta, result, fit), axis=1)
+    out = np.concatenate((theta.reshape(nspoke, 1), result, fit), axis=1)
 
     np.save('output/out_'+gal+'.npy', out)
 

### 2019-01-02 schmactions/cluster_eval.py
Change Rn to kpc

@@ -10,6 +10,7 @@ from scipy.optimize import minimize
 import matplotlib as mpl
 from matplotlib import rc
 import matplotlib.pyplot as plt
+import matplotlib.ticker as mticker
 rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
@@ -135,9 +136,9 @@ for ax, gal in zip(ax_list, glist):
         Rn_low_target = np.abs(Rn_low_target)
         Rn_high_target = np.abs(Rn_high_target)
 
-        ax.plot(mc_list, Rn_target*1000, c=c, label="{0:0.1f}".format(Jz))
-        ax.plot(mc_list, Rn_low_target*1000, c=c, ls='dashed')
-        ax.plot(mc_list, Rn_high_target*1000, c=c, ls='dashed')
+        ax.plot(mc_list, Rn_target, c=c, label="{0:0.1f}".format(Jz))
+        ax.plot(mc_list, Rn_low_target, c=c, ls='dashed')
+        ax.plot(mc_list, Rn_high_target, c=c, ls='dashed')
 
         if gal=='m12f':
             l = 0.05
@@ -153,11 +154,12 @@ for ax, gal in zip(ax_list, glist):
 for ax in ax_list:
     ax.set_xlabel(r'$m_c\,[\,M_{\odot}\,]$')
     ax.set_xscale('log')
-    ax.set_ylim((200, 20000))
+    ax.set_ylim((0.2, 20))
     ax.set_xlim((mc_min, mc_max))
     ax.set_yscale('log')
+    ax.yaxis.set_major_formatter(mticker.ScalarFormatter())
 
-ax_list[0].set_ylabel(r'$R_n[\,\text{pc}\,]$')
+ax_list[0].set_ylabel(r'$R_n[\,\text{kpc}\,]$')
 
 ax.legend(frameon=False, title=r'$J_{z,\text{true}}$')
 

### 2019-01-02 paper/ms.tex
Fix paper temporarily
Add Rn plot to paper
Add some text to describe Rn plot

@@
 $\Delta\phi$, which we then convert into a chord length which we interpret as
-$R_{\n}$. The result of this procedure is shown in Figure~\ref{fig:Rn_mc}.
+$R_{\n}$, the neighborhood around the cluster. If the $z$~offset is greater
+than the maximum range value of the galaxy, we simply report the maximum chord
+length ($2\times8.2\,\kpc$). The result of this procedure is shown in
+Figure~\ref{fig:Rn_mc}. The solid lines indicate the neighborhood values
+assuming the mean relation between $z$~offset and $m_c$, and the dashed lines
+show the $+1\,\sigma$ and $-1\,\sigma$ relations, indicative of the strong
+$\phi$-dependence of the midplane effect.
+
+Figure~\ref{fig:Rn_mc} is most informative for the low $J_z$ orbit (blue). The
+neighborhood around $\sim100\Msun$ clusters can be as small as $\sim300\,\pc$
+but up to $\sim1.2\,\kpc$ for m12i and m12f, though the situation is
+marginally better for m12m with values ranging from $\sim0.8\,\kpc$ to
+$\sim3\,\kpc$. The neighborhood values steadily increase with $m_c$, topping
+out at $\sim5\,\kpc$ for $\sim10^4\,\Msun$ clusters in m12i and m12f. In m12m,
+the neighborhood value reaches the maximum allowed of $16.4\,\kpc$ for
+clusters of mass $\sim800\,\Msun$. The interpretation of this is that the
+expected action space dispersion is so high for such high mass clusters that
+the measured $z$~offset is not large enough throughout the entire galaxy to be
+important.
+
+For the medium $J_z$ orbit (orange), the situation is better, although such an
+orbit is not a good fit to any of the nearby open clusters \Gus{Table x}. For
+m12i and m12f, the neighborhood is $\sim \text{few}\,\kpc$ for
+$m_c\sim10^2\,\Msun$. At $\sim10^3\,\Msun$ the curves reach their maximum
+value. For m12m, the neighborhood starts at $\sim10\,\kpc$ and quickly rises
+to the maximum value. For the high $J_z$ orbit (red), the neighborhood is at
+the maximum value for nearly all masses, except at the very low-mass end in
+m12i.
 
 \begin{figure*}
-\plotone{fig/}
+\plotone{fig/Rn_vs_mc.pdf}
+\caption{Caption.}
+\label{fig:Rn_mc}
 \end{figure*}

### 2019-01-03 paper/ms.tex
Remove commented text
Update info on action comp

@@
 \subsection{Summary of Action Computation} \label{ssec:action_comp}
-We begin by summarizing one way in which actions are computed in the
-literature \Gus{cite}. One first assumes an axisymmetric model of the Galaxy.
-This model may itself be a poor description of the Galactic potential,
-
+We follow our previous work in computing actions \citep{2018ApJ...867...31B}.
 We use the code \texttt{gala} v0.3 to perform orbit integrations and
@@
 which uses an action estimator technique presented by
-\citet{2014MNRAS.441.3284S}. Another common alternative is \texttt{galpy}
-\citep{2015ApJS..216...29B} using either direct orbit integration or the
-\Gus{Stackel Fudge} approximation. We use the default \texttt{MWPotential} as
+\citet{2014MNRAS.441.3284S}. We use the default \texttt{MWPotential} as
 our potential, which is based on the Milky Way potential available in
@@
 Sun-like star.
+
+Other methods for computing actions are used in the literature. For example,
+the St\"ackel Fudge method \citep{2016MNRAS.457.2107S} which uses the
+St\"ackel potential to approximate the Galactic potential
+\citep{1985MNRAS.216..273D,2012MNRAS.426.1324B} is commonly used
+\citep{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}.
 
@@
 
-% {\em Gaia} DR2 has ushered in a new era of Galactic dynamics. High-precision
-% phase space measurements have been made for an unprecedented number of stars.
-% In the face of such large data, it is worth considering the goal of such
-% measurements --- a goal which should inform the usage of such data.
-
-% In this work we considered one such question: given the current position in
-% phase space of a single star, what can we infer about that star's past phase
-% space trajectory? Current methodology involves using a smoothed, analytic form
-% for the Galactic potential (with perhaps crude implementations of spiral arms
-% or sattelites), fitting such form to some selection of observations, and
-% integrating from the currently measured position in phase space. This orbit
-% integration is also commonly accompanied with a transformation into
-% action-angle space. Other than the integration, every step of that process
-% should be viewed with significant scrutiny.
-
-% The more general problem of measuring the orbits of stars in cosmological
-% simulations is essentially unexplored in this work and the literature, and we
-% delay discussion to future work.
-
-% \subsection{Dynamical Midplane}
-% We now seek to define the midplane in a dynamically consistent manner, which
-% we refer to as the ``dynamical midplane''. By this, we mean we would like to
-% define the midplane in a manner consistent with some sort of dynamical
-% constraint. Our method, which we describe now, is just one of many possible
-% ways the dynamical midplane can be defined.
-
-% We define the dynamical midplane based on the principle that perfectly
-% circular, planar orbits will trace the midplane. Once the circular, planar
-% orbits have been identified, the procedure for finding the dynamical midplane
-% is relatively simple. From the measurement of each star's orbit, the midplane
-% of the galaxy can be computed at $z=0$ easily. Furthermore, we will find that
-% each orbit also gives a measure of the: dynamical center of the galaxy,
-% velocity of the galaxy, and acceleration of the galaxy.
-
-% We pause briefly to consider general difficulties in measuring stellar orbits
-% in a cosmological simulation. It may seem reasonable to place the galaxy in an
-% inertial reference frame and then to simply measure the positions of each star
-% particle over time. However, it is not generally the case that the center of
-% mass of the stellar disk experiences a vanishing acceleration. Indeed, in a
-% re-run of the simulations presented here\footnote{The re-run contained a
-% cosmic ray bug fix which is unimportant here, but the re-run also output the
-% acceleration of each particle, which is not the case for the simulations in
-% this work.}, we measure a total acceleration of $\sim 0.4\,\kms/\text{Myr}$
-% for stars with cylindrical radii within $0.5\,\kpc$ of $8.2\,\kpc$ and
-% vertical heights less than $0.4\,\kpc$. Assuming a constant acceleration, over
-% $22\,\Myr$ --- the time spacing of snapshots considered in this work --- this
-% results in a $\sim90\,\pc$ shift of the disk. Since the offset scales with
-% $t^2$, this presents a significant problem for measuring stellar orbits over
-% longer timescales.
-
-% Another option is to recenter and reorient the galaxy at each snapshot.
-% However, the centering and orientation procedures are not without
-% stochasticity themselves. This stochasticity can easily dominate the magnitude
-% of the orbit. For instance, a shift in the direction of the $z$-axis by a mere
-% $0.7\degree$ results in a shift in the midplane of $100\,\pc$ at a radius of
-% $8.2\,\kpc$. We attempted this procedure in various ways (i.e. exploring
-% different centering and orientation procedures), but were not able to cleanly
-% measure the orbits of disk stars.
-
-% Here we choose to simplify the problem by considering only perfectly circular,
-% planar orbits. Since these orbits have no phase-dependent structure, we can
-% work on shorter timescales, i.e. for orbits with significant radial or
-% vertical action a large fraction of their orbit would need to be sampled in
-% order to measure the orbit accurately. Because we know the analytic form of
-% these orbits, we can forward model the trajectory of each star particle in the
-% cosmological simulation. Working in comoving coordinates, we can write down
-% the trajectory of a star particle as
-% \beq\label{eq:helix}
-% \bm{r}(t) = R(\theta, \phi, \psi) \left[ (A\cos{\omega t}, B\sin{\omega t}, C
-% t) + \bm{x_0} \right]\text{,}
-% \eeq
-% where $R(\theta, \phi, \psi)$ is an Euler angle transformation, $A$, $B$, $C$,
-% and $\omega$ are arbitrary constants, and $\bm{x_0}$ is an arbitrary 3-vector.
-% This equation is a helix with an arbitrary constant and rotation. It describes
-% the
-
-% We minimize the summed squared difference between Equation~\ref{eq:helix} and
-% the true trajectory of each star with $R$ within $1\,\kpc$ of $R_{\odot} =
-% 8.2\,\kpc$ in each simulation. We consider minimum rms error (i.e. the typical
-% error in the Equation~\ref{eq:helix} model in describing the true trajectory).
-% This rms error characterizes how well the helix model we consider, which would
-% be exactly correct if the overall motion of the galaxy is well-described by
-% constant speed motion and the orbit of a particular store is exactly circular
-% and planar.
-
-% We plot a histogram of the rms error for each galaxy in
-% Figure~\ref{fig:something}. As we estimated before, the total effect of the
-% galaxy's acceleration is roughly $90\,\pc$ (\Gus{Check for m12f, m12m}). We
-% find that this is small compared to the typical error, even for the stars
-% whose trajectories are best modeled by Equation~\ref{eq:helix}.
-
-% There are many reasons for the failure of this model. First, there could be no
-% stars on circular orbits. This seems unlikely considering that typical
-% distribution functions peak on perfectly circular, planar orbits
-% \citep[e.g.][]{2013MNRAS.434..652T}. Second, the trajectory of the galaxy may
-% not be well-described by constant-speed motion. This also seems unlikely,
-% considering that the bulk acceleration of the stellar disk only results in a
-% $\sim90\,\pc$ (\Gus{m12f, m12m}) deviation from constant-speed motion over the
-% $22\,\Myr$ time we considered.
-
-% Furthermore, in Figure~\ref{fig:something} we plot the center of mass of the
-% galaxy (in comoving coordinates) for the $22\,\Myr$ timeframe described. The
-% center of mass at each snapshot is computed using the iterative shrinking
-% spheres method described in Section~\ref{ssec:cosmozoom}. The rms error on the
-% constant-speed motion is $\sim\text{something}$, which we interpret as
-% \Gus{something}.
-
-% One potential explanation is that {\em no star can remain on a circular,
-% planar orbit}. The order unity density variations may be strong enough such
-% that each star's orbit is perturbed. If there are truly stars in the
-% simulation on perfectly circular, planar orbits, then the magnitude of their
-% acceleration over the last $22\,\Myr$ must be constant. Furthermore, their
-% trajectory in acceleration space must be well-described by a rotation. In the re-runs,
-
-% \section{Optimally Axisymmetric Frame} \label{sec:oa_frame}
-% An analytic, axisymmetric potential with a known coordinate system is a far
-% too simple description of the Milky Way. We thus turn to cosmological zoom-in
-% simulations as approximations to a ``true'' Milky Way potential. We will
-% perform a few simulations of small clusters embedded in the potential of a
-% frozen $\z=0$ snapshot of one such zoom-in. We take our fiducial reference
-% frame from \citet{2018arXiv180610564S}.
-
-% Freezing the potential means that during the course of our simulation the
-% cluster will rapidly pass by much of the substructure in the potential. To
-% illustrate that this has little effect on our results, we resimulate each
-% cluster with the softening lengths of the embedded potential increased by a
-% factor of \Gus{something}.
-
-% Once we perform this simulation, we calculate the actions of each star in the
-% cluster assuming the best-fit smooth, axisymmetric potential, as we do
-% observationally for the Milky Way. We refer to this space as the observational
-% action space. We will find that there is significant time evolution in
-% observational action space. Attributing some of this evolution to our fiducial
-% coordinate system being incorrect, we minimize the time-dispersion in action
-% space
-
-% \subsection{Cluster Simulations} \label{ssec:cluster_sim}
-% We simulate a small star cluster of $256$ stars in the frozen $\z=0$
-% potential of m12i. We use the code \texttt{ph4} within the
-% \texttt{AMUSE} framework \citep{2011ascl.soft07007P, 2013CoPhC.184..456P,
-% 2013A&A...557A..84P} for the cluster integration. We use a softening length
-% of $0.01\,\pc$ which means we do not correctly solve the internal dynamics of
-% the cluster, but this is not important for our purposes. To couple the
-% gravitational forces of the frozen snapshot to the cluster stars, we use the
-% \texttt{BRIDGE} \citep{2007PASJ...59.1095F}, also implemented in
-% \texttt{AMUSE}. Our simulation runs for $2\,\Gyr$, allowing us to solve for
-% $\sim 10$ orbits.
-
-% To compute the forces from the frozen potential, we use the simple tree code
-% \texttt{pykdgrav}\footnote{\url{https://github.com/omgspace/pykdgrav},
-% commit: \texttt{fbd1ddc}}. We use an opening angle of $0.5$ and exclude stars
-% at distances $R>50\,\kpc$ in the fiducial coordinate system.
-
-% We use the initial position and velocity of two star particles in
-% m12i. These two star particles are randomly chosen from all star
-% particles in the simulation with the following properties (in the fiducial
-% coordinate system):
-% \begin{enumerate}
-%     \item age in the range $(0.25, 0.75)\,\Gyr$,
-%     \item cylindrical radius in the range $(7.7, 8.7)\,\kpc$,
-%     \item height (absolute value) less than $0.5\,\kpc$,
-%     \item $J_r$ and $J_z$ in the range:
-%         \subitem $(10, 20)\,\actunit$ for the first star,
-%         \subitem $(40, 60)\,\actunit$ for the second star.
-% \end{enumerate}
-% The ids we choose are \texttt{23693026} for the first star and
-% \texttt{17012804} for the second star. We refrain from calling these ``low''
-% or ``high'' action stars since the actions are computed in a presumably
-% incorrect coordinate system.
-
-% To compute actions, both in the choosing of each of these star particles and
-% in what follows, we use the package
-% \texttt{agama}\footnote{\url{https://github.com/GalacticDynamics-Oxford/Agama},
-% commit: \texttt{0e48336}}. Using only the high-resolution particles, we
-% compute the symmetrized potential in two components. First, a multipole
-% expansion with $20$ radial grid points and a max order of $2$ for the
-% spherical-harmonic expansion is used for the dark matter. Second, a
-% cylindrical spline expansion with $20$ radial and $z$ grid points, a $0$ order
-% max index azimuthal-harmonic expansion, $0.2$ and $50\,\kpc$ for the minimum
-% and maximum radial grid node, $0.02$ and $10\,\kpc$ for the minimum and
-% maximum $z$ grid node is used for the baryonic matter (stars and gas). In both
-% cases, axisymmetry is explicitly enforced. Other \texttt{agama} parameters are
-% left as default.
-
-% \subsection{Frame Optimization} \label{ssec:frame_opt}
-% Figure~\Gus{something} shows the action computed at each time step in the
-% fiducial coordinate system. Clearly there is significant time evolution.
-
-% Even though we change the center, we do not recalculate the center velocity.
-% This is largely irrelevant, as the recalculated center velocity differs from
-% the fiducial center velocity by $<0.1\,\kms$. In recentering the galaxy, the
-% package \texttt{gizmo\_analysis} will subtract a slightly different value for
-% the Hubble flow from each particle's velocity to convert to a physical
-% coordinate system. The error this introduces is $\sim 0.02\,\kms$, which we
-% also ignore.
-
 \appendix \section{Appendix 1}

### 2019-01-03 paper/ms.tex
Finish caption in one orbit fig

@@
 error in the Galactocentric coordinate system. We consider here an orbit with
-initial position $(8,0,0)\,\kpc$ and initial velocity $(0,-190,50)\,\kms$.
-This corresponds to an orbit with actions $(J_r, L_z, J_z) = (32.5, -1520,
-23.0)\,\actunit$ and $z_{\text{max}}=854\,\pc$. Considering the first $\Gyr$
-of this orbit, we assume an immortal observer }
+initial position $(8,0,0)\,\kpc$ and initial velocity $(0,-190,30)\,\kms$.
+This corresponds to an orbit with actions $(J_r, L_z, J_z) = (37.9, -1520,
+7.0)\,\actunit$ and $z_{\text{max}}=425\,\pc$. We integrate the orbit
+according to the procedure laid out in Section~\ref{ssec:action_comp}. Then,
+we subtract $100\,\pc$ from the $z$ value of each position in the orbit,
+corresponding to an erroneous observer assuming a midplane that is off by
+$100\,\pc$. We then allow our (immortal) observer to observe the orbit over
+$1\,\Gyr$ and perform the same orbit integration procedure at each timestep,
+and report the values of the actions. The computation of $L_z$ is pristine to
+errors in $z$, with only numerical artifacts remaining. Only small errors are
+induced in $J_r$, with the middle $90\%$ of values over the $\Gyr$ being
+$\sim6\%$ of the true $J_r$. As expected, large errors are induced in $J_z$,
+with the middle $90\%$ of values being $\sim85\%$ of the true $J_z$.}
 \label{fig:one_orbit_wrong_ref}

### 2019-01-03 paper/ms.tex
Update fractional errors for one orbit

@@
 infer significant time evolution in their computed actions. The
-95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $8.7$,
-$47.1$, and $19.3\,\actunit$ for $J_r$, $L_z$, and $J_z$, respectively. These
-are fractional errors of $26.5\%$, $3.1\%$, and $81.4\%$, respectively.
+95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
+$6.2 \actunit$ for $J_r$ and $J_z$, respectively. These are fractional errors
+of $5.7\%$ and $85.7\%$, respectively. The error induced in $L_z$ is
+negligible, as expected since $L_z$ does not depend on the value of $z$. It is
+worth pointing out that a $100\,\pc$ error in an orbit with
+$z_{\text{max}}=425\,\pc$ --- a $24\%$ error --- induced an $85.7\%$ error in
+the computation of $J_z$.
 

### 2019-01-03 paper/ms.tex
Minor change
Update numbers in text

@@
 which uses an action estimator technique presented by
-\citet{2014MNRAS.441.3284S}. We use the default \texttt{MWPotential} as
-our potential, which is based on the Milky Way potential available in
+\citet{2014MNRAS.441.3284S}. We use the default \texttt{MWPotential} as our
+potential, which is based on the Milky Way potential available in
 \texttt{galpy} \citep{2015ApJS..216...29B}. This potential includes a
@@
 frame. We begin by taking an orbit with initial position $(8, 0, 0)\,\kpc$ and
-initial velocity $(0, -190, 50)\,\kms$ --- this corresponds to an orbit with
-actions $(J_r, L_z, J_z) = (32.5, -1520, 23.0)\,\actunit$ and
-$z_{\text{max}}=854\,\pc$. The orbit is plotted in \Gus{Appendix x}. We then
+initial velocity $(0, -190, 30)\,\kms$ --- this corresponds to an orbit with
+actions $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$ and
+$z_{\text{max}}=425\,\pc$. The orbit is plotted in \Gus{Appendix x}. We then
 suppose that an observer makes observations of this orbit in a coordinate

### 2019-01-03 paper/ms.tex
Update intro to schmaction section

@@
 good description of the potential and that we have correctly determined the
-parameters of that model. We are only concerned with an inaccuracy
+parameters of that model. We are only concerned with an inaccuracy in the
+Galactocentric coordinate system and, in particular, the measurement of the
+Galactic midplane. Here we determine how an error in either the Galactic
+midplane or Galactocentric radius propagates into an error in the various
+actions.
 

### 2019-01-03 paper/ms.tex
Rewrite section 2 intro

--- 2019-01-03 paper/ms.tex
@@
 \section{Impact on Actions} \label{sec:ref_frame}
-We first demonstrate the importance of measuring the Galactic midplane to very
-high accuracy. We explore the consequence of an incorrect midplane in the
-context of computing orbital actions of disk stars. An inaccuracy in the
-midplane is only one potential systematic error in the way actions are
-presently computed. The assumed axisymmetric model of the Galaxy may be a poor
-description of the Galactic potential, and the parameters in this axisymmetric
-model may be estimated incorrectly.
-
-In this work, we assume both that an axisymmetric model of the Galaxy is a
-good description of the potential and that we have correctly determined the
-parameters of that model. We are only concerned with an inaccuracy
+We first demonstrate the importance of measuring the Galactic midplane and
+Solar radius to very high accuracy. To be clear, the consequences we explore
+here may also arise from various other observational errors. For instance, the
+symmetrized Galactic potential used by observers may not be a good description
+of the true potential --- or the parameters used may be incorrect. In this
+Section, we assume these approximations are valid and simply explore the
+consequences of an inaccurate Galactocentric coordinate system.
 

### 2019-01-03 paper/ms.tex
Move action comp summary around

@@
 
-\subsection{Summary of Action Computation} \label{ssec:action_comp}
-We follow our previous work in computing actions \citep{2018ApJ...867...31B}.
-We use the code \texttt{gala} v0.3 to perform orbit integrations and
-conversion to action space \citep{2017JOSS....2..388P,Price-Whelan:2018},
-which uses an action estimator technique presented by
-\citet{2014MNRAS.441.3284S}. We use the default \texttt{MWPotential} as our
-potential, which is based on the Milky Way potential available in
-\texttt{galpy} \citep{2015ApJS..216...29B}. This potential includes a
-Hernquist bulge and nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
-\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
-is fit to empirically match some observations. We use the Dormand-Prince
-8(5,3) integration scheme \citep{Dormand80:integrator}. We use a timestep of
-$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
-Sun-like star.
-
-Other methods for computing actions are used in the literature. For example,
-the St\"ackel Fudge method \citep{2016MNRAS.457.2107S} which uses the
-St\"ackel potential to approximate the Galactic potential
-\citep{1985MNRAS.216..273D,2012MNRAS.426.1324B} is commonly used
-\citep{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}.
-
 \subsection{Cartoon} \label{ssec:cartoon}
@@
 
+\subsection{Summary of Action Computation} \label{ssec:action_comp}
+We follow our previous work in computing actions \citep{2018ApJ...867...31B}.
+We use the code \texttt{gala} v0.3 to perform orbit integrations and
+conversion to action space \citep{2017JOSS....2..388P,Price-Whelan:2018},
+which uses an action estimator technique presented by
+\citet{2014MNRAS.441.3284S}. We use the default \texttt{MWPotential} as our
+potential, which is based on the Milky Way potential available in
+\texttt{galpy} \citep{2015ApJS..216...29B}. This potential includes a
+Hernquist bulge and nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
+\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
+is fit to empirically match some observations. We use the Dormand-Prince
+8(5,3) integration scheme \citep{Dormand80:integrator}. We use a timestep of
+$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
+Sun-like star.
+
+Other methods for computing actions are used in the literature. For example,
+the St\"ackel Fudge method \citep{2016MNRAS.457.2107S} which uses the
+St\"ackel potential to approximate the Galactic potential
+\citep{1985MNRAS.216..273D,2012MNRAS.426.1324B} is commonly used
+\citep{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}.
+
 \subsection{Quantification} \label{ssec:quant}
-We now turn to classical orbital integration in a commonly assumed Milky Way
-potential to quantify the time variation of computed actions for different
-coordinate system errors. Our action computation procedure is described in
-Section~\ref{ssec:action_comp}.
-
 First, we illustrate the time-dependence of one possible inaccurate reference

### 2019-01-03 paper/ms.tex
Update color

@@
 
-Now suppose this observer makes a measurement of the true orbit at the orange
-point or the red point. Then the orange and red lines correspond to the orbits
+Now suppose this observer makes a measurement of the true orbit at the blue
+point or the pink point. Then the blue and pink lines correspond to the orbits
 that the observer would compute for each point. In action space, this would
-correspond to a different value of $J_z$ for the orange and red points. In
-this way, assuming the wrong coordinate system induces time dependence in the
+correspond to a different value of $J_z$ for the blue and pink points. In this
+way, assuming the wrong coordinate system induces time dependence in the
 actions {\em computed by an observer}.

### 2019-01-03 paper/ms.tex
Add some placeholders to Section 2, small improvements

@@
 suppose that an observer makes observations of this orbit in a coordinate
-system that is incorrect in radius by $200\,\pc$ and height by $100\,\pc$ ---
-we subtract the vector $(200, 0, 100)\,\pc$ from each position in the orbit.
-This corresponds to an observer physically located at the position $(8, 0,
-0)\,\kpc$ but erroneously thinking they are located at $(8.2, 0, 0.1)\,\kpc$.
+system that is incorrect in height by $100\,\pc$ --- i.e. we subtract the
+vector $(0, 0, 100)\,\pc$ from each position in the orbit. This corresponds to
+an observer physically located at the position $(8, 0, 0)\,\kpc$ but
+erroneously thinking they are located at $(8, 0, 0.1)\,\kpc$.
 
-Using the observations of the orbit that this erroneous observer makes every
-$\Myr$\footnote{In addition to being erroneous, we also assume our fictitious
-observer is immortal.}, we perform the same orbit integration procedure as
-before to compute the actions our observer would make in the incorrect
+Using the observations of the orbit that this erroneous (and immortal)
+observer makes every $\Myr$, we perform the same orbit integration procedure
+as before to compute the actions our observer would observe in the incorrect
 coordinate system over the duration of the orbit. The erroneous computed
-actions are shown for the first $\Gyr$ of the orbit in
+actions are shown for the first $\Gyr$ of the orbit in the {\em upper} panels
 Figure~\ref{fig:one_orbit_wrong_ref}. Occasionally the numerical scheme fails
@@
     this only excludes a total of $5$ orbits out of the $1000$ considered for
-Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
-vast majority of orbits are computed properly.
+    Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but
+    the vast majority of orbits are computed properly. We perform the same
+    procedure but assume a $100\,\pc$ offset in the $x$ component (i.e.
+    subtracting the vector $(100, 0, 0)\,\pc$) in the {\em lower} panels. This
+    is equivalent to an observer making an incorrect measurement of the Solar
+    radius.
 
 Figure~\ref{fig:one_orbit_wrong_ref} shows that the erroneous observer would
-infer significant time evolution in their computed actions. The
-95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
-$6.2 \actunit$ for $J_r$ and $J_z$, respectively. These are fractional errors
-of $5.7\%$ and $85.7\%$, respectively. The error induced in $L_z$ is
-negligible, as expected since $L_z$ does not depend on the value of $z$. It is
-worth pointing out that a $100\,\pc$ error in an orbit with
-$z_{\text{max}}=425\,\pc$ --- a $24\%$ error --- induced an $85.7\%$ error in
-the computation of $J_z$.
+infer significant time evolution in their computed actions. For an error in
+$z$ ({\em upper} panels), the 95\textsuperscript{th} minus
+5\textsuperscript{th} percentiles are $2.2$ and $6.2 \actunit$ for $J_r$ and
+$J_z$, respectively. These are fractional errors of $5.7\%$ and $85.7\%$,
+respectively. The error induced in $L_z$ is negligible, as expected since
+$L_z$ does not depend on the value of $z$. It is worth pointing out that a
+$100\,\pc$ error in an orbit with $z_{\text{max}}=425\,\pc$ --- a $24\%$ error
+--- induced an $85.7\%$ error in the computation of $J_z$.
+
+For an error in the $x$ component (or Solar radius), the 95\textsuperscript{th} minus
+5\textsuperscript{th} percentiles are $x$, $y$, and $z \actunit$ for $J_r$, $L_z$ and
+$J_z$, respectively. These are fractional errors of $x\%$, $y\%$ and $z\%$,
+respectively.
 

### 2019-01-03 paper/ms.tex
Update multi-orbit text, update terminology and numbers

@@
 
-We now repeat the procedure but continuously changing the error in the assumed
-midplane. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth minus
-5\uth percentile divided by the true action value for three different orbits.
-The red orbit is exactly the same as in Figure~\ref{fig:one_orbit_wrong_ref},
-and we also consider two more planar orbits with initial vertical velocities
-of $30\,\kms$ (\Gus{orange}) and $10\,\kms$ (\Gus{blue}). For these two new
-orbits, the true actions are $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$
-and $(40.4, -1520, 0.7)\,\actunit$ for the \Gus{orange} and \Gus{blue} orbits,
-respectively. The value of $z_{\text{max}}$ for these orbits are $122\,\pc$
-and $425\,\pc$, respectively (with the blue orbit having $z_{\text{max}}$ of
-$854\,\pc$.) \Gus{Easy way to explain change in $J_r$?}
+We now repeat the procedure but continuously changing the error in the $z$ and
+$x$ components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth
+minus 5\uth percentile divided by the true action value for three different
+orbits. The \Gus{orange} orbit is exactly the same as in
+Figure~\ref{fig:one_orbit_wrong_ref}, and we also consider two more planar
+orbits with initial vertical velocities of $50\,\kms$ (\Gus{red}) and
+$10\,\kms$ (\Gus{blue}). For these two new orbits, the true actions are $(J_r,
+L_z, J_z) = (32.6, -1520, 22.9)\,\actunit$ and $(40.4, -1520, 0.7)\,\actunit$
+for the \Gus{red} and \Gus{blue} orbits, respectively. The value of
+$z_{\text{max}}$ for these orbits are $854\,\pc$ and $122\,\pc$, respectively
+(with the orange orbit having $z_{\text{max}}$ of $425\,\pc$.) \Gus{Easy way
+to explain change in $J_r$?}
 
 The top row of Figure~\ref{Fig:many_orbit_wrong_ref} shows the error in each
-action assuming an offset in the midplane. In the {\em bottom row} we also
-consider offsets in the Galactocentric radius $R$. The {\em left}, {\em
-middle}, and {\em right} columns show the 95\uth minus 5\uth percentile of the
-measured action over the course of the incorrect orbit divided by the true
-action value as a percentage.
+action assuming an offset in the $z$ component (i.e. the Galactic midplane).
+In the {\em bottom row} we also consider offsets in the $x$ component (i.e.
+the Solar radius). The {\em left}, {\em middle}, and {\em right} columns show
+the relative error in $J_z$, $L_z$, and $J_r$, respectively. We compute the
+relative error as the  95\uth minus 5\uth percentile of the measured action
+($\Delta J_i$) over the course of the first $\Gyr$ of the incorrectly observed
+orbit divided by the true action value as a percentage.
 
@@
 stars\footnote{Though this is not how we compute $L_z$ in practice.}, and is
-thus unaffected by offsets in $z$.
+thus unaffected by offsets in $z$. This also agrees with the result we found
+in Figure~\ref{fig:one_orbit_wrong_ref}.
 
@@
 
-For completeness, we also consider the effect of offsets in the Galactocentric
-radius $R$ in the {\em bottom panels}. The error is largest for $J_r$, with
-some deviations resulting in $L_z$ and relatively small deviations in $J_z$.
-In the {\em bottom left} panel the orange and blue lines overlap, and in the
-{\em bottom middle} and {\em bottom right} panels all three lines overlap.
+For the offset in the Solar radius {\em bottom panels}, the error is largest
+for $J_r$, with some deviations resulting in $L_z$ and relatively small
+deviations in $J_z$. In the {\em bottom left} panel the orange and blue lines
+overlap, and in the {\em bottom middle} and {\em bottom right} panels all
+three lines overlap.
 

### 2019-01-03 paper/ms.tex
Remove commented text
Small variable name updates

@@
 
-% We restrict ourselves to the simulation m12i \citep[first introduced
-% in][]{2016ApJ...827L..23W}. For this work, this galaxy is the best-case
-% scenario. It does not contain a bar (like \texttt{m12m}) or companion galaxy,
-% and by eye is the closest to axisymmetric of the three galaxies presented in
-% \citet{2018arXiv180610564S} \citep{2018MNRAS.481.4133G}.
-
 It is important to remember that simulations of Milky Way-mass galaxies are
@@
 
-Suppose an open cluster of mass $M_c$ on an orbit with actions $\bm{J}_c$ is
+Suppose an open cluster of mass $m_c$ on an orbit with actions $\bm{J}_c$ is
 disrupted, and that the cluster is currently at an orbit phase of
@@
 To explore the magnitude of this effect, we consider the following program.
-First, we assume a cluster mass $m$ and estimate the expected action-space
+First, we assume a cluster mass $m_c$ and estimate the expected action-space
 dispersion from Equation~\ref{eq:action_disp}. We then consult

### 2019-01-04 paper/ms.tex
Add caveat about solar radius
Fix typo

@@
 
-The top row of Figure~\ref{Fig:many_orbit_wrong_ref} shows the error in each
+The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the error in each
 action assuming an offset in the $z$ component (i.e. the Galactic midplane).
@@
 midplane is insufficient for extracting the most dynamical information from
-the {\em Gaia} mission.
+the {\em Gaia} mission. From now on, we set aside the discussion of
+inaccuracies in the Solar radius. While all that we state about the midplane
+applies to the Solar radius as well, it is not clear the best way to determine
+the dynamical center of the galaxy in the FIRE simulations.
 

### 2019-01-04 paper/ms.tex
Add caption to many orbits

@@
 \plotone{fig/many_orbits_schmactions.pdf}
-\caption{Caption.}
+\caption{\Gus{update figure} We report the 95\uth minus 5\uth percentile of the error in the
+measured action ($\Delta J_i$) from coordinate system errors for three orbits
+with low-, medium-, and high-$J_z$ values (blue, orange, and red,
+respectively). The {\em left}, {\em center}, and {\em right} panels show the
+result for $J_r$, $L_z$, and $J_z$, respectively. The {\em upper} panels
+consider an offset in $z$ and the {\em lower} panels consider an offset in $x$
+(equivalently, an offset in the Solar radius).}
 \label{fig:many_orbit_wrong_ref}

### 2019-01-04 paper/ms.tex
Add caption to midplane figure

@@
 \plotone{fig/midplane_fit.pdf}
-\caption{Caption.}
+\caption{The local midplane determined at the fiducial Solar radius
+($8.2\,\kpc$) for the three FIRE galaxies m12i, m21f, and m12m ({\em left},
+{\em center}, and {\em right} panels). The local midplane is determined at a
+position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
+and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
+using the new height $10$ times to converge on the local midplane height. In
+order to allow for the possibility that the fiducial Galactocentric coordinate
+system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
+each panel --- this figure is reproduced with the original midplane
+determination (i.e. before subtracting the best fit sine curve) in
+Appendix~\Gus{x}. We then bootstrap resample $1000$ times on all stars within
+a $2\,\kpc$ height of the fiducial midplane to determine error bars (95\uth
+and 5\uth percentiles), which we report as dashed lines.}
 \label{fig:midplane}

### 2019-01-04 paper/ms.tex
Make discussion section into subsections, add applicability to Milky Way

@@
 
+\subsection{Applicability to the Milky Way} \label{ssec:is_it_real}
+As we briefly discussed in Section~\ref{ssec:cosmozoom}, cosmological zoom-in
+simulations of Milky Way-mass galaxies are not perfect representations of the
+Milky Way, and do not reproduce every available observation of the Milky Way.
+For instance, the velocity dispersion-age relation in m12i, m12f, and m12m is
+closer to that of M31 \citep{2018arXiv180610564S}.
+
+In order to assess the applicability of our work to the real Milky Way, we
+consider a subset of Table~3 from \citet{2018arXiv180610564S} --- our
+Table~\ref{tab:scale_height} shows the stellar scale heights of m12i, m12f,
+and m12m and the current best estimates for the Milky Way. One noticeable
+feature is that the FIRE galaxies are more ``puffy'' than the Milky Way.
+\Gus{Maybe FIRE collaboration has some insight into the mechanism for this?}.
+This would seem to suggest that the midplane estimates from
+Figure~\ref{fig:midplane} over-estimate the true variation in the Milky Way.
+
+\begin{deluxetable}{ccc}
+\tablecaption{Stellar Disk Scale Heights}
+\tablehead{\colhead{galaxy} & \colhead{thin disk} & \colhead{thick disk} \\ 
+\colhead{} & \colhead{(pc)} & \colhead{(pc)} } 
+\startdata
+MW\tablenotemark{a} & 300 & 900 \\
+m12i\tablenotemark{b} & 480 & 2000 \\
+m12f\tablenotemark{b} & 440 & 1280 \\
+m12m\tablenotemark{b} & 290 & 1030 \\
+\enddata
+% \tablerefs{\citet{2016ARAA..54..529B,2018arXiv180610564S,2008ApJ...673..864J}}
+
+\tablenotetext{a}{\citet{2008ApJ...673..864J,2016ARAA..54..529B}}
+\tablenotetext{b}{\citet{2018arXiv180610564S}}
+
+\label{tab:scale_height}
+\end{deluxetable}
+
+Another important aspect to consider when comparing the simulated galaxies to
+the Milky Way is the environment/merger history. For instance, m12i and m12m
+both lack a massive neighbor analogous to the LMC. On the other hand, m12f has
+had a recenter merger with \Gus{something}. \Gus{Expand more here - how to
+incorporate merger trees?}
+
+\subsection{Solar Neighborhood} \label{ssec:neighborhood}
 To gain further handle on the midplane behavior as a function of azimuth, we

### 2019-01-04 paper/ms.tex
Add caption to cluster offset fig

@@
 \plotone{fig/cluster_offset.pdf}
-\caption{Caption.}
+\caption{The $z$ offset necessary in order for the action error induced by our
+midplane effect to be comparable to the intrinsic action dispersion from a
+cluster of mass $m_c$. We compute the intrinsic dispersion from
+Equation~\ref{eq:action_disp} and convert the dispersion to a $z$ offset using
+the result from Figure~\ref{fig:many_orbit_wrong_ref}. We consider the result
+for three different orbits with varying vertical actions of $J_z = 0.7, 7.0,
+22.9\,\actunit$ ($z_{\text{max}} = 122,425,854\,\pc$) --- colored by
+\Gus{blue, orange, red}, respectively. Nearby open clusters have actions
+closest to the $J_z=0.7\,\actunit$ (blue) orbit, see Table~\Gus{x}.}
 \label{fig:cluster_offset}

### 2019-01-04 paper/ms.tex
Add caption to the range-deltaphi plot

@@
 \plotone{fig/range_dphi.pdf}
-\caption{Caption.}
+\caption{The range of midplane heights encountered as a function of angular
+width. At each angle $\phi$ from Figure~\ref{fig:midplane} we consider an
+angular width of $\Delta \phi$ centered on $\phi$ and report the range of
+midplane heights within that width. We repeat the procedure for each $\phi$
+and plot the result as translucent gray lines. We also plot the mean range as
+a solid blue line and the $\pm1\sigma$ lines as dashed blue lines. The upper
+$x$-axis shows the chord length from the position $\phi$ to
+$\pm\Delta\phi/2$.}
 \label{fig:range_deltaphi}

### 2019-01-04 paper/ms.tex
Add caption to neighborhood-mass relation fig

@@
 \plotone{fig/Rn_vs_mc.pdf}
-\caption{Caption.}
+\caption{The neighborhood around an open cluster of mass $m_c$ for our three
+orbits (blue, orange, and red) and for each of the FIRE galaxies considered
+here --- m12i, m12f, and m12m ({\em left}, {\em center}, and {\em right}
+panels, respectively). Given the $z$ offset for a cluster of mass $m_c$
+computed in Figure~\ref{fig:cluster_offset}, we convert this to a chord length
+using the result from Figure~\ref{fig:range_deltaphi} and plot as a solid blue
+line. We interpret this chord length as the ``neighborhood'' around the
+cluster, i.e. the distance from the cluster one could go before the $J_z$
+error induced by our midplane effect is comparable to the intrinsic action
+dispersion of the cluster. When the $z$ offset is larger than the maximum
+range in Figure~\ref{fig:range_deltaphi}, we report the maximum chord length
+($2\times8.2\,\kpc$). We aso report the same procedure but using the
+$\pm1\sigma$ lines from Figure~\ref{fig:range_deltaphi}, which we plot here as
+dashed lines. For the medium- and high-$J_z$ orbits (orange and red), the
+neighborhood is quite large and thus the effect is negligible. However, for
+the low-$J_z$ orbit (blue), which is closest to the orbits of nearby open
+clusters (see Table~\Gus{x}), the neighborhood is only a few hundred $\pc$ for
+low-mass clusters.}
 \label{fig:Rn_mc}

### 2019-01-14 paper/ms.tex
Expand intro a bit

@@
 
-Since the Milky Way is not truly axisymmetric, it is not clear the best way to
-define the Galactocentric coordinate system --- or, perhaps, the best way to
-modify the current implicit definition of the Galactocentric coordinate system
-in order to account for non-axisymmetries.
+We define the ``global Galactic midplane'' to be the plane of $z$ symmetry in
+the axisymmetric model of the Milky Way. We define the ``local Galactic
+midplane'' to be the plane of $z$ symmetry inferred by an observer at a
+particular location. The true global Galactic midplane is the plane of $z$
+symmetry in the best-fit axisymmetric model. It is not guaranteed that the
+local Galactic midplane coincide with the global Galactic midplane. Even
+without a high-precision definition and measurement of the global Galactic
+midplane, we can still assess how different the two should be by measuring the
+azimuthal dependence of the local Galactic midplane. This procedure may be
+readily tractable with current datasets. However, we turn to simulation in
+this work.
 
-In a slowly-evolving axisymmetric potential, the actions of stars' orbits are
+It's important for us to emphasize that when we state our local Galactic
+midplane may be erroneous, we do not mean that past measurements/estimates of
+it are wrong. Rather, what we mean is that it may vary from the local midplane
+at other points in the Galaxy. In the face of {\em Gaia} DR2 (and soon DR3/4),
+this may be an important effect to consider when performing dynamical
+modelling across large distances.
+
+Since an azimuthal-dependent midplane is effectively a departure from a perfectly axisymmetric potential, we In a slowly-evolving axisymmetric potential, the actions of stars' orbits are
 conserved quantities \citep{2008gady.book.....B,2014RvMP...86....1S}. In other

### 2019-01-14 paper/ms.tex
Fix line breaks

@@
 
-Since an azimuthal-dependent midplane is effectively a departure from a perfectly axisymmetric potential, we In a slowly-evolving axisymmetric potential, the actions of stars' orbits are
-conserved quantities \citep{2008gady.book.....B,2014RvMP...86....1S}. In other
-words, if we could measure the actions of a star in such a potential, we would
-measure the same actions over the course of the orbit. The non-axisymmetries
-of a more realistic potential induce time evolution in the {\em measured}
-actions --- the true orbital invariants (if they exist) are invariant by
-definition. Through simple tests, we will see that even in a truly
-axisymmetric potential, if the wrong coordinate system is assumed then there
-will be additional time evolution in the measured actions.
+Since an azimuthal-dependent midplane is effectively a departure from a
+perfectly axisymmetric potential, we In a slowly-evolving axisymmetric
+potential, the actions of stars' orbits are conserved quantities
+\citep{2008gady.book.....B,2014RvMP...86....1S}. In other words, if we could
+measure the actions of a star in such a potential, we would measure the same
+actions over the course of the orbit. The non-axisymmetries of a more
+realistic potential induce time evolution in the {\em measured} actions ---
+the true orbital invariants (if they exist) are invariant by definition.
+Through simple tests, we will see that even in a truly axisymmetric potential,
+if the wrong coordinate system is assumed then there will be additional time
+evolution in the measured actions.
 
@@
 
-For an error in the $x$ component (or Solar radius), the 95\textsuperscript{th} minus
-5\textsuperscript{th} percentiles are $x$, $y$, and $z \actunit$ for $J_r$, $L_z$ and
-$J_z$, respectively. These are fractional errors of $x\%$, $y\%$ and $z\%$,
-respectively.
+For an error in the $x$ component (or Solar radius), the
+95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $x$, $y$,
+and $z \actunit$ for $J_r$, $L_z$ and $J_z$, respectively. These are
+fractional errors of $x\%$, $y\%$ and $z\%$, respectively.
 
@@
 \plotone{fig/many_orbits_schmactions.pdf}
-\caption{\Gus{update figure} We report the 95\uth minus 5\uth percentile of the error in the
-measured action ($\Delta J_i$) from coordinate system errors for three orbits
-with low-, medium-, and high-$J_z$ values (blue, orange, and red,
-respectively). The {\em left}, {\em center}, and {\em right} panels show the
-result for $J_r$, $L_z$, and $J_z$, respectively. The {\em upper} panels
+\caption{\Gus{update figure} We report the 95\uth minus 5\uth percentile of
+the error in the measured action ($\Delta J_i$) from coordinate system errors
+for three orbits with low-, medium-, and high-$J_z$ values (blue, orange, and
+red, respectively). The {\em left}, {\em center}, and {\em right} panels show
+the result for $J_r$, $L_z$, and $J_z$, respectively. The {\em upper} panels
 consider an offset in $z$ and the {\em lower} panels consider an offset in $x$

### 2019-01-14 paper/ms.tex
Add real clusters to paper

@@
 stars and stellar properties such as age or metallicity can be investigated.
+
+We explore one science goal that relies on the conservation of actions --- the
+reconstruction of open clusters across large distances using dynamical
+tagging. The basic program is to match stars with roughly the same actions. To
+set the scene, we consider the orbits of the nine open clusters within
+$250\,\pc$ as reported by the {\em Gaia} collaboration
+\citep{2018A&A...616A..10G}. In Table~\ref{tab:real_clusters}, we report the
+distance to each cluster, the three actions, and the maximum vertical extent
+of their orbits ($z_{\text{max}}$). The actions are computed using the
+technique described in Section~\ref{ssec:action_comp}.
+
+\begin{deluxetable*}{cccccc}
+\tablehead{\colhead{cluster} & \colhead{distance} & \colhead{$J_r$} &
+\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$}\\ \colhead{ } &
+\colhead{$\mathrm{pc}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
+\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
+\colhead{$\mathrm{pc}$}}
+\startdata
+alphaPer & 174.9 & 14.0 & -1739 & 0.0045 & 10.5 \\
+Blanco1 & 237.2 & 1.6 & -1841 & 1.53 & 207.0 \\
+ComaBer & 85.9 & 1.6 & -1860 & 0.71 & 139.7 \\
+Hyades & 47.5 & 20.3 & -1757 & 0.26 & 81.5 \\
+IC2391 & 151.6 & 5.5 & -1790 & 0.018 & 21.2 \\
+IC2602 & 152.2 & 11.5 & -1718 & 0.32 & 88.6 \\
+NGC2451 & 193.7 & 5.8 & -1805 & 0.21 & 73.4 \\
+Pleiades & 135.8 & 19.3 & -1696 & 0.38 & 96.3 \\
+Praesepe & 186.2 & 21.2 & -1766 & 0.58 & 122.9
+\enddata
+\tablerefs{\citet{2018AA...616A..10G}}
+\label{tab:real_clusters}
+\end{deluxetable*}
 

### 2019-01-14 paper/ms.tex
Add more about what an action is

@@
 Since an azimuthal-dependent midplane is effectively a departure from a
-perfectly axisymmetric potential, we In a slowly-evolving axisymmetric
-potential, the actions of stars' orbits are conserved quantities
-\citep{2008gady.book.....B,2014RvMP...86....1S}. In other words, if we could
-measure the actions of a star in such a potential, we would measure the same
-actions over the course of the orbit. The non-axisymmetries of a more
-realistic potential induce time evolution in the {\em measured} actions ---
-the true orbital invariants (if they exist) are invariant by definition.
+perfectly axisymmetric potential, we consider the effect that an erroneous
+midplane has on the conservations of actions. The actions of an orbit are
+typically computed in a cylindrical coordinate system as,
+\beq\label{eq:actions}
+J_i \equiv
+\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
+\eeq
+where $i=R,z,\phi$. Note that $J_{\phi} \equiv L_z$. In a slowly-evolving axisymmetric potential, the actions
+of stars' orbits are conserved quantities
+\citep{2008gady.book.....B,2014RvMP...86....1S}. The non-axisymmetries of a
+more realistic potential induce time evolution in the {\em measured} actions
+--- the true orbital invariants (if they exist) are invariant by definition.
 Through simple tests, we will see that even in a truly axisymmetric potential,

### 2019-01-14 paper/ms.tex
Expand on limitations of actions

@@
 The transformation to action space is not trivial, and many of the assumptions
-of the transformation are potentially incorrect. It is therefore natural to
-consider the reason to transform to action space. One reason is dimensionality
-reduction --- in the action space paradigm, a stellar orbit is completely
-described by three actions, as opposed to six dimensions of phase space. This
-means that the relationship between {\em orbital} properties of individual
-stars and stellar properties such as age or metallicity can be investigated.
+of the transformation are potentially incorrect. For instance, the assumption
+that the Galactic potential is axisymmetric may be inadequate with current
+measurement errors. Even if this assumption is valid, the parameters used in
+our axisymmetric models may be inadequately tuned to observations. In
+addition, the necessity of 6D phase space measurements of a star limits the
+Galactic volume observable in action space.
+
+With these limitations in mind, it is therefore natural to consider the reason
+to transform to action space. One reason is dimensionality reduction --- in
+the action space paradigm, a stellar orbit is completely described by three
+actions, as opposed to six dimensions of phase space. This means that the
+relationship between {\em orbital} properties of individual stars and stellar
+properties such as age or metallicity can be investigated.
 
@@
 reconstruction of open clusters across large distances using dynamical
-tagging. The basic program is to match stars with roughly the same actions. To
-set the scene, we consider the orbits of the nine open clusters within
-$250\,\pc$ as reported by the {\em Gaia} collaboration
-\citep{2018A&A...616A..10G}. In Table~\ref{tab:real_clusters}, we report the
-distance to each cluster, the three actions, and the maximum vertical extent
-of their orbits ($z_{\text{max}}$). The actions are computed using the
-technique described in Section~\ref{ssec:action_comp}.
+tagging. The basic program is to match stars with roughly the same actions as
+their supposed natal cluster. To set the scene, we consider the orbits of the
+nine open clusters within $250\,\pc$ as reported by the {\em Gaia}
+collaboration \citep{2018A&A...616A..10G}. In Table~\ref{tab:real_clusters},
+we report the distance to each cluster, the three actions, and the maximum
+vertical extent of their orbits ($z_{\text{max}}$). The actions are computed
+using the technique described in Section~\ref{ssec:action_comp}.
 
@@
 
-To provide one such realization of a realistic Galactic potential, we turn to
+To provide a realization of a realistic Galactic potential, we turn to
 cosmological zoom-ins of Milky Way-mass galaxies from the Feedback in
 Realistic Environments (FIRE)
-collaboration\footnote{\url{https://fire.northwestern.edu}} (the Latte suite
+collaboration\footnote{\url{https://fire.northwestern.edu}} (the {\em Latte} suite
 of simulations).

### 2019-01-14 ticker_tape/ticker_tape.py
Clean up old files

@@ -1,59 +1,0 @@
-import gizmo_analysis as gizmo
-import numpy as np
-from tqdm import tqdm
-import matplotlib.pyplot as plt
-
-from matplotlib import rc
-import matplotlib as mpl
-rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
-rc('text', usetex=True)
-mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
-
-tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
-        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
-
-#gal_info = 'oa_frame.txt'
-gal_list = ['m12i', 'm12f', 'm12m']
-
-for gal in gal_list:
-    gal_info = 'fiducial_coord/' + gal + '_res7100_center.txt'
-
-    sim_directory = '/mnt/ceph/users/firesims/fire2/metaldiff/' + gal + '_res7100'
-    snap = gizmo.io.Read.read_snapshots(['star', 'gas', 'dark'], 'index', 600,
-                                   properties=['position', 'id',
-                                               'mass', 'velocity',
-                                               'form.scalefactor'],
-                                     assign_center=False,
-                                     simulation_directory=sim_directory)
-
-    ref = np.genfromtxt(gal_info, comments='#', delimiter=',')
-    snap.center_position = ref[0]
-    snap.center_velocity = ref[1]
-    snap.principal_axes_vectors = ref[2:]
-    for k in snap.keys():
-        for attr in ['center_position', 'center_velocity', 
-                     'principal_axes_vectors']:
-            setattr(snap[k], attr, getattr(snap, attr))
-
-    star_pos = snap['star'].prop('host.distance.principal.cylindrical')
-
-    R = star_pos[:,0]
-    z = star_pos[:,1]
-    phi = star_pos[:,2]
-
-    Rmid = 8.2
-    dR = 0.5
-
-    Rbool = np.logical_and(R > Rmid - dR, R < Rmid + dR)
-    zbool = np.abs(z) < 1
-
-    keys = np.where(np.logical_and(Rbool, zbool))[0]
-
-    phibins = np.linspace(0, 2.*np.pi, 51)
-    zbins = np.arange(-1, 1, 0.01)
-    plt.figure(figsize=(7,2))
[…]

### 2019-01-15 paper/ms.tex

@@
 of simulations).
+
+In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
+system errors have on the measured actions. In Section~\ref{sec:local_fire},
+we measure the local midplane in cosmological zoom-in simulations from the
+FIRE collaboration. In Section~\ref{sec:discussion} we discuss and give some
+context for our results, and implications for the definition of the Solar
+neighborhood. We conclude in Section~\ref{sec:conclusion}.
 

### 2019-01-23 schmactions/many_orbits/plot_many_orbits.py
Small updates to many_orbits

@@ -46,7 +46,7 @@ def sclip(a, s=4):
 
 name_list = ['thin', 'thick', 'halo']
 
-fig, ax = plt.subplots(2, 3, figsize=(7, 3.5))
+fig, ax = plt.subplots(2, 3, figsize=(7, 4))
 init_pos = [8, 0, 0] * u.kpc
 init_vel_list = [[0, -190, 10] * u.km/u.s,
                  [0, -190, 50] * u.km/u.s,
@@ -112,11 +112,11 @@ for x in ax[1]:
     x.set_xticks(np.arange(xmin,xmax,100), minor=True)
 
 for x in ax[:,0]:
-    x.set_ylabel(r'$\Delta J_r/J_r\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+    x.set_ylabel(r'$\Delta J_r/J_r\,[\,\%\,]$')
 for x in ax[:,1]:
-    x.set_ylabel(r'$\Delta L_z/L_z\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+    x.set_ylabel(r'$\Delta L_z/L_z\,[\,\%\,]$')
 for x in ax[:,2]:
-    x.set_ylabel(r'$\Delta J_z/J_z\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+    x.set_ylabel(r'$\Delta J_z/J_z\,[\,\%\,]$')
 
 ax[0][1].legend(frameon=False, title='orbit')
 

### 2019-01-23 paper/ms.tex
Update paper to include correct figures
Add table with orbits
Update ref to table

@@
 \eeq
-where $i=R,z,\phi$. Note that $J_{\phi} \equiv L_z$. In a slowly-evolving axisymmetric potential, the actions
-of stars' orbits are conserved quantities
+where $i=R,z,\phi$. Note that $J_{\phi} \equiv L_z$. In a slowly-evolving
+axisymmetric potential, the actions of stars' orbits are conserved quantities
 \citep{2008gady.book.....B,2014RvMP...86....1S}. The non-axisymmetries of a
@@
 First, we illustrate the time-dependence of one possible inaccurate reference
-frame. We begin by taking an orbit with initial position $(8, 0, 0)\,\kpc$ and
-initial velocity $(0, -190, 30)\,\kms$ --- this corresponds to an orbit with
-actions $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$ and
-$z_{\text{max}}=425\,\pc$. The orbit is plotted in \Gus{Appendix x}. We then
-suppose that an observer makes observations of this orbit in a coordinate
-system that is incorrect in height by $100\,\pc$ --- i.e. we subtract the
-vector $(0, 0, 100)\,\pc$ from each position in the orbit. This corresponds to
-an observer physically located at the position $(8, 0, 0)\,\kpc$ but
-erroneously thinking they are located at $(8, 0, 0.1)\,\kpc$.
+frame. Throughout this work we will consider three orbits typical of the
+stellar thin disk, stellar thick disk, and the stellar halo. We summarize
+their initial position in phase space and the actions of their integrated
+orbits in Table~\ref{tab:orbits}. Each orbit is plotted in \Gus{Appendix X}.
+We will refer to these orbits by their names henceforth.
+
+We begin by considering the thick orbit. Suppose that an observer makes
+observations of this orbit in a coordinate system that is incorrect in height
+by $100\,\pc$ --- i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each
+position in the orbit. This corresponds to an observer physically located at
+the position $(8, 0, 0)\,\kpc$ but erroneously thinking they are located at
+$(8, 0, 0.1)\,\kpc$.
+
+\begin{deluxetable*}{ccccccc}
+\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
+velocity} & \colhead{$J_r$} &
+\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$}\\ \colhead{ } &
+\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
+\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
+\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
+\colhead{$\mathrm{pc}$}}
+\startdata 
+thin & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 122 \\
+thick & $(8, 0, 0)$ & $(0, -190, 30)$ & 32.5 & -1520 & 23.0 & 425 \\ 
+halo & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.8 & -1520 & 529.1 & 854
+\enddata
+\label{tab:orbits}
+\end{deluxetable*}
 
@@
 \begin{figure*}
-\plotone{fig/one_orbit.pdf}
+\plotone{fig/schmactions_one_orbit.pdf}
 \caption{The artificial phase-dependence in the observed actions induced by an
@@
 \begin{figure*}
-\plotone{fig/many_orbits_schmactions.pdf}
+\plotone{fig/schmactions_many_orbits.pdf}
 \caption{\Gus{update figure} We report the 95\uth minus 5\uth percentile of
@@
 \end{figure*}
+
+\begin{figure}
+\plotone{fig/schmactions_Jz_hist.pdf}
+\caption{Caption.}
+\label{fig:Jz_hist}
+\end{figure}
 

### 2019-01-23 paper/ms.tex
Add reminder to fill in numbers

@@
 
-For an error in the $x$ component (or Solar radius), the
+\Gus{For an error in the $x$ component (or Solar radius)}, the
 95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $x$, $y$,

### 2019-01-23 real_cluster/compute_real_cluster_actions.py
Add gc coordinates for real clusters

@@ -57,3 +57,10 @@ r['zmax'].unit = u.pc
 
 r.write('real_cluster_actions.fits', format='fits')
 r.write('real_cluster_actions.tex', format='ascii.aastex')
+
+pos = np.transpose(scdyn.pos.xyz.to_value(u.pc))
+vel = np.transpose(scdyn.vel.d_xyz.to_value(u.km/u.s))
+
+cart = np.concatenate((m,pos,vel), axis=1)
+r2 = Table(cart, names=('cluster', 'distance', 'pos.x', 'pos.y', 'pos.z', 'vel.x', 'vel.y', 'vel.z'))
+r2.write('real_cluster_gc.fits', format='fits')

### 2019-01-28 paper/ms.tex
Add some masses of clusters to table

@@
 
-\begin{deluxetable*}{cccccc}
+\begin{deluxetable*}{ccccccc}
+\tablecaption{Relevant dynamical parameters for the $9$ open clusters within $250\,\pc$ of the Sun.}
 \tablehead{\colhead{cluster} & \colhead{distance} & \colhead{$J_r$} &
-\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$}\\ \colhead{ } &
+\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} \\ \colhead{ } &
 \colhead{$\mathrm{pc}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
 \colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{pc}$}}
+\colhead{$\mathrm{pc}$} & \colhead{$M_{\odot}$}}f
 \startdata
-alphaPer & 174.9 & 14.0 & -1739 & 0.0045 & 10.5 \\
-Blanco1 & 237.2 & 1.6 & -1841 & 1.53 & 207.0 \\
-ComaBer & 85.9 & 1.6 & -1860 & 0.71 & 139.7 \\
-Hyades & 47.5 & 20.3 & -1757 & 0.26 & 81.5 \\
-IC2391 & 151.6 & 5.5 & -1790 & 0.018 & 21.2 \\
-IC2602 & 152.2 & 11.5 & -1718 & 0.32 & 88.6 \\
-NGC2451 & 193.7 & 5.8 & -1805 & 0.21 & 73.4 \\
-Pleiades & 135.8 & 19.3 & -1696 & 0.38 & 96.3 \\
-Praesepe & 186.2 & 21.2 & -1766 & 0.58 & 122.9
+alphaPer & 174.9 & 14.0 & -1739 & 0.0045 & 10.5 & 352\tablenotemark{a} \\
+Blanco1 & 237.2 & 1.6 & -1841 & 1.53 & 207.0 & 410\tablenotemark{b} \\
+ComaBer & 85.9 & 1.6 & -1860 & 0.71 & 139.7 &  \\
+Hyades & 47.5 & 20.3 & -1757 & 0.26 & 81.5 & 400\tablenotemark{d} \\
+IC2391 & 151.6 & 5.5 & -1790 & 0.018 & 21.2 & \\
+IC2602 & 152.2 & 11.5 & -1718 & 0.32 & 88.6 & \\
+NGC2451 & 193.7 & 5.8 & -1805 & 0.21 & 73.4 & \\
+Pleiades & 135.8 & 19.3 & -1696 & 0.38 & 96.3 & \\
+Praesepe & 186.2 & 21.2 & -1766 & 0.58 & 122.9 & 
 \enddata
+
+\tablenotetext{a}{\citet{2016MNRAS.457.1028S}}
+\tablenotetext{b}{\citet{2007AA...471..499M}}
+\tablenotetext{c}{}
+\tablenotetext{d}{\citet{1998AA...331...81P}}
+
 \tablerefs{\citet{2018AA...616A..10G}}

### 2019-01-28 paper/ms.tex
Update fig and caption

@@
 \begin{deluxetable*}{ccccccc}
+\tablecaption{Description and names of the three orbits considered in this work.}
 \tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
@@
 \caption{The artificial phase-dependence in the observed actions induced by an
-error in the Galactocentric coordinate system. We consider here an orbit with
-initial position $(8,0,0)\,\kpc$ and initial velocity $(0,-190,30)\,\kms$.
-This corresponds to an orbit with actions $(J_r, L_z, J_z) = (37.9, -1520,
-7.0)\,\actunit$ and $z_{\text{max}}=425\,\pc$. We integrate the orbit
-according to the procedure laid out in Section~\ref{ssec:action_comp}. Then,
-we subtract $100\,\pc$ from the $z$ value of each position in the orbit,
-corresponding to an erroneous observer assuming a midplane that is off by
-$100\,\pc$. We then allow our (immortal) observer to observe the orbit over
-$1\,\Gyr$ and perform the same orbit integration procedure at each timestep,
-and report the values of the actions. The computation of $L_z$ is pristine to
-errors in $z$, with only numerical artifacts remaining. Only small errors are
-induced in $J_r$, with the middle $90\%$ of values over the $\Gyr$ being
-$\sim6\%$ of the true $J_r$. As expected, large errors are induced in $J_z$,
-with the middle $90\%$ of values being $\sim85\%$ of the true $J_z$.}
+error in the Galactocentric coordinate system. We consider here the thick disk
+orbit, which has actions of $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$
+and $z_{\text{max}}=425\,\pc$ (see Table~\ref{tab:orbits}). We integrate the
+orbit according to the procedure laid out in Section~\ref{ssec:action_comp}.
+Then, we subtract $100\,\pc$ from the $z$ value ({\em upper panels}) or the
+$x$ value ({\em lower panels}) of each position in the orbit, corresponding to
+an erroneous observer assuming a midplane ({\em upper}) or solar radius ({\em
+lower}) that is off by $100\,\pc$. We then allow our (immortal) observer to
+observe the orbit over $1\,\Gyr$ and perform the same orbit integration
+procedure at each timestep, and report the values of the actions. The
+computation of $L_z$ is pristine to errors in $z$, with only numerical
+artifacts remaining. Only small errors are induced in $J_r$, with the middle
+$90\%$ of values over the $\Gyr$ being $\sim6\%$ of the true $J_r$. As
+expected, large errors are induced in $J_z$, with the middle $90\%$ of values
+being $\sim85\%$ of the true $J_z$.}
 \label{fig:one_orbit_wrong_ref}

### 2019-01-28 paper/ms.tex
More updates to cluster table

@@
 \startdata
-alphaPer & 174.9 & 14.0 & -1739 & 0.0045 & 10.5 & 352\tablenotemark{a} \\
-Blanco1 & 237.2 & 1.6 & -1841 & 1.53 & 207.0 & 410\tablenotemark{b} \\
-ComaBer & 85.9 & 1.6 & -1860 & 0.71 & 139.7 &  \\
-Hyades & 47.5 & 20.3 & -1757 & 0.26 & 81.5 & 400\tablenotemark{d} \\
+alphaPer & 174.9 & 14.0 & -1739 & 0.0045 & 10.5 & 352\tablenotemark{b} \\
+Blanco1 & 237.2 & 1.6 & -1841 & 1.53 & 207.0 & 410\tablenotemark{c} \\
+ComaBer & 85.9 & 1.6 & -1860 & 0.71 & 139.7 & 112\tablenotemark{d} \\
+Hyades & 47.5 & 20.3 & -1757 & 0.26 & 81.5 & 400\tablenotemark{e} \\
 IC2391 & 151.6 & 5.5 & -1790 & 0.018 & 21.2 &  \\
 IC2602 & 152.2 & 11.5 & -1718 & 0.32 & 88.6 & \\
-NGC2451 & 193.7 & 5.8 & -1805 & 0.21 & 73.4 & \\
+NGC2451A\tablenotetext{a} & 193.7 & 5.8 & -1805 & 0.21 & 73.4 & \\
 Pleiades & 135.8 & 19.3 & -1696 & 0.38 & 96.3 & \\
-Praesepe & 186.2 & 21.2 & -1766 & 0.58 & 122.9 & 
+Praesepe & 186.2 & 21.2 & -1766 & 0.58 & 122.9 & 550\tablenotemark{d}
 \enddata
 
-\tablenotetext{a}{\citet{2016MNRAS.457.1028S}}
-\tablenotetext{b}{\citet{2007AA...471..499M}}
-\tablenotetext{c}{}
-\tablenotetext{d}{\citet{1998AA...331...81P}}
+\tablenotetext{a}{This cluster is labeled NGC2451, which we interpret as NGC2451A since NGC2451B lies further than $250\,\pc$ away.}
+\tablenotetext{b}{\citet{2016MNRAS.457.1028S}}
+\tablenotetext{c}{\citet{2007AA...471..499M}}
+\tablenotetext{d}{\citet{2007AJ....134.2340K}}
+\tablenotetext{e}{\citet{1998AA...331...81P}}
 

### 2019-01-28 paper/ms.tex
Update orbits table
Fixing references

@@
 be $\sim 8.1\text{--}8.3\,\kpc$
-\citep{2009ApJ...692.1075G,2018A&A...615L..15G}. The vertical height is found
+\citep{2009ApJ...692.1075G,2018AA...615L..15G}. The vertical height is found
 by identifying where the stellar positions and velocities reach a maximum
@@
 nine open clusters within $250\,\pc$ as reported by the {\em Gaia}
-collaboration \citep{2018A&A...616A..10G}. In Table~\ref{tab:real_clusters},
+collaboration \citep{2018AA...616A..10G}. In Table~\ref{tab:real_clusters},
 we report the distance to each cluster, the three actions, and the maximum
@@
 \colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{pc}$}}
+\colhead{$\mathrm{kpc}$}}
 \startdata 
-thin & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 122 \\
-thick & $(8, 0, 0)$ & $(0, -190, 30)$ & 32.5 & -1520 & 23.0 & 425 \\ 
-halo & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.8 & -1520 & 529.1 & 854
+thin & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 \\
+thick & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 \\ 
+halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16
 \enddata

### 2019-01-28 paper/ms.tex
Fixing table labels, captions

@@
 \begin{deluxetable*}{ccccccc}
-\tablecaption{Relevant dynamical parameters for the $9$ open clusters within $250\,\pc$ of the Sun.}
+\tablecaption{Relevant dynamical parameters for the $9$ open clusters within
+$250\,\pc$ of the Sun.\label{tab:real_clusters}}
 \tablehead{\colhead{cluster} & \colhead{distance} & \colhead{$J_r$} &
@@
 IC2602 & 152.2 & 11.5 & -1718 & 0.32 & 88.6 & \\
-NGC2451A\tablenotetext{a} & 193.7 & 5.8 & -1805 & 0.21 & 73.4 & \\
+NGC2451A\tablenotemark{a} & 193.7 & 5.8 & -1805 & 0.21 & 73.4 & \\
 Pleiades & 135.8 & 19.3 & -1696 & 0.38 & 96.3 & \\
@@
 
-\tablenotetext{a}{This cluster is labeled NGC2451, which we interpret as NGC2451A since NGC2451B lies further than $250\,\pc$ away.}
+\tablenotetext{a}{This cluster is labeled NGC2451, which we interpret as
+NGC2451A since NGC2451B lies further than $250\,\pc$ away.}
 \tablenotetext{b}{\citet{2016MNRAS.457.1028S}}
@@
 \tablerefs{\citet{2018AA...616A..10G}}
-\label{tab:real_clusters}
 \end{deluxetable*}
@@
 \begin{deluxetable*}{ccccccc}
-\tablecaption{Description and names of the three orbits considered in this work.}
+\tablecaption{Description and names of the three orbits considered in this
+work.\label{tab:orbits}}
 \tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
@@
 \enddata
-\label{tab:orbits}
 \end{deluxetable*}
@@
 \begin{deluxetable}{ccc}
-\tablecaption{Stellar Disk Scale Heights}
+\tablecaption{Stellar disk scale heights of the real Milky Way and the
+simulated galaxies considered in this work.\label{tab:scale_height}}
 \tablehead{\colhead{galaxy} & \colhead{thin disk} & \colhead{thick disk} \\ 
@@
 \tablenotetext{b}{\citet{2018arXiv180610564S}}
-
-\label{tab:scale_height}
 \end{deluxetable}

### 2019-01-28 schmactions/cluster_offset_real/Rn_eval.py
Add calc of neighborhood

@@ -1,0 +1,24 @@
+import pickle
+import numpy as np
+from astropy.table import Table
+from scipy.interpolate import interp1d
+
+mlist = np.array([ 352, 410, 112, 400, np.nan, np.nan, np.nan, np.nan, 550])
+t = Table.read('real_cluster_gc.fits', format='fits')
+name_list = t['cluster']
+
+gal_list = ['m12i', 'm12f', 'm12m']
+
+for name, mass in zip(name_list, mlist):
+    for gal in gal_list:
+        t = pickle.load(open('Rn_vs_mc_'+gal+'_'+name+'.p', 'rb'))
+        mlist = t[:,0]
+        Rn = t[:,1]
+        i = interp1d(mlist, Rn)
+        if np.isnan(mass):
+            this_Rn = np.nan
+        else:
+            this_Rn = i(mass)
+        print(name, gal, this_Rn*1000)
+
+

### 2019-01-28 paper/ms.tex
Line length fix
Add neighborhood column to table

@@
 
-\begin{deluxetable*}{ccccccc}
+\begin{deluxetable*}{cccccccc}
 \tablecaption{Relevant dynamical parameters for the $9$ open clusters within
@@
 \tablehead{\colhead{cluster} & \colhead{distance} & \colhead{$J_r$} &
-\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} \\ \colhead{ } &
+\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} & \colhead{neighborhood} \\ 
+\colhead{ } &
 \colhead{$\mathrm{pc}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
 \colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{pc}$} & \colhead{$M_{\odot}$}}f
+\colhead{$\mathrm{pc}$} & \colhead{$M_{\odot}$} & \colhead{$\mathrm{pc}$}}
 \startdata
-alphaPer & 174.9 & 14.0 & -1739 & 0.0045 & 10.5 & 352\tablenotemark{b} \\
-Blanco1 & 237.2 & 1.6 & -1841 & 1.53 & 207.0 & 410\tablenotemark{c} \\
-ComaBer & 85.9 & 1.6 & -1860 & 0.71 & 139.7 & 112\tablenotemark{d} \\
-Hyades & 47.5 & 20.3 & -1757 & 0.26 & 81.5 & 400\tablenotemark{e} \\
-IC2391 & 151.6 & 5.5 & -1790 & 0.018 & 21.2 &  \\
-IC2602 & 152.2 & 11.5 & -1718 & 0.32 & 88.6 & \\
-NGC2451A\tablenotemark{a} & 193.7 & 5.8 & -1805 & 0.21 & 73.4 & \\
-Pleiades & 135.8 & 19.3 & -1696 & 0.38 & 96.3 & \\
-Praesepe & 186.2 & 21.2 & -1766 & 0.58 & 122.9 & 550\tablenotemark{d}
+alphaPer & 174.9 & 14.0 & -1739 & 0.0045 & 10.5 & 352\tablenotemark{b} & 57 \\
+Blanco1 & 237.2 & 1.6 & -1841 & 1.53 & 207.0 & 410\tablenotemark{c} & 1460 \\
+ComaBer & 85.9 & 1.6 & -1860 & 0.71 & 139.7 & 112\tablenotemark{d} & 522\\
+Hyades & 47.5 & 20.3 & -1757 & 0.26 & 81.5 & 400\tablenotemark{e} & 462 \\
+IC2391 & 151.6 & 5.5 & -1790 & 0.018 & 21.2 & & \\
+IC2602 & 152.2 & 11.5 & -1718 & 0.32 & 88.6 & & \\
+NGC2451A\tablenotemark{a} & 193.7 & 5.8 & -1805 & 0.21 & 73.4 & & \\
+Pleiades & 135.8 & 19.3 & -1696 & 0.38 & 96.3 & & \\
+Praesepe & 186.2 & 21.2 & -1766 & 0.58 & 122.9 & 550\tablenotemark{d} & 781
 \enddata
@@
 Realistic Environments (FIRE)
-collaboration\footnote{\url{https://fire.northwestern.edu}} (the {\em Latte} suite
-of simulations).
+collaboration\footnote{\url{https://fire.northwestern.edu}} (the {\em Latte}
+suite of simulations).
 

### 2019-01-28 paper/ms.tex
Add Pleiades to neighborhood calc

@@
 NGC2451A\tablenotemark{a} & 193.7 & 5.8 & -1805 & 0.21 & 73.4 & & \\
-Pleiades & 135.8 & 19.3 & -1696 & 0.38 & 96.3 & & \\
+Pleiades & 135.8 & 19.3 & -1696 & 0.38 & 96.3 & 800\tablenotemark{f} & 685 \\
 Praesepe & 186.2 & 21.2 & -1766 & 0.58 & 122.9 & 550\tablenotemark{d} & 781
@@
 \tablenotetext{e}{\citet{1998AA...331...81P}}
+\tablenotetext{f}{\citet{2001AJ....121.2053A}}
 

### 2019-01-28 schmactions/cluster_offset_real/Rn_eval.py

@@ -3,7 +3,7 @@ import numpy as np
 from astropy.table import Table
 from scipy.interpolate import interp1d
 
-mlist = np.array([ 352, 410, 112, 400, np.nan, np.nan, np.nan, np.nan, 550])
+mlist = np.array([ 352, 410, 112, 400, np.nan, np.nan, np.nan, 800, 550])
 t = Table.read('real_cluster_gc.fits', format='fits')
 name_list = t['cluster']
 

### 2019-01-28 paper/ms.tex
Add extra cite to Gaia paper

@@
 \tablecaption{Relevant dynamical parameters for the $9$ open clusters within
-$250\,\pc$ of the Sun.\label{tab:real_clusters}}
+$250\,\pc$ of the Sun, using the relevant initial positions and velocities from \citet{2018AA...616A..10G}.\label{tab:real_clusters}}
 \tablehead{\colhead{cluster} & \colhead{distance} & \colhead{$J_r$} &

### 2019-01-28 paper/ms.tex
Add a little something about Stackel Fudge in the disk
Small update to numbers in text

@@
 \citep{1985MNRAS.216..273D,2012MNRAS.426.1324B} is commonly used
-\citep{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}.
+\citep{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}. For
+applications in the disk, the St\"ackel Fudge method is of acceptable
+accuracy, but since we also consider halo orbits in this work we choose to use
+orbit integration throughout \citep{2016MNRAS.457.2107S}.
 
@@
 $L_z$ does not depend on the value of $z$. It is worth pointing out that a
-$100\,\pc$ error in an orbit with $z_{\text{max}}=425\,\pc$ --- a $24\%$ error
---- induced an $85.7\%$ error in the computation of $J_z$.
+$100\,\pc$ error in an orbit with $z_{\text{max}}=850\,\pc$ --- a $12\%$ error
+--- induced a $43\%$ error in the computation of $J_z$.
 
-\Gus{For an error in the $x$ component (or Solar radius)}, the
-95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $x$, $y$,
-and $z \actunit$ for $J_r$, $L_z$ and $J_z$, respectively. These are
-fractional errors of $x\%$, $y\%$ and $z\%$, respectively.
+For an error in the $x$ component (or Solar radius), the
+95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$, $47$,
+and $0.71 \actunit$ for $J_r$, $L_z$ and $J_z$, respectively. These are
+fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively.
 

### 2019-01-29 paper/ms.tex
More small updates to text

@@
 orbit, which has actions of $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$
-and $z_{\text{max}}=425\,\pc$ (see Table~\ref{tab:orbits}). We integrate the
+and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We integrate the
 orbit according to the procedure laid out in Section~\ref{ssec:action_comp}.
@@
 artifacts remaining. Only small errors are induced in $J_r$, with the middle
-$90\%$ of values over the $\Gyr$ being $\sim6\%$ of the true $J_r$. As
-expected, large errors are induced in $J_z$, with the middle $90\%$ of values
-being $\sim85\%$ of the true $J_z$.}
+$90\%$ of values over the $\Gyr$ being $\sim8\%$ of the true $J_r$. As
+expected, large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$,
+with the middle $90\%$ of values being $\sim43\%$ of the true $J_z$. The $x$
+offset induces uncertainties in $J_r$ and $L_z$ of $\sim21\%$ and $\sim3\%$,
+respectively. A $\sim3\%$ error in $J_z$ is also induced.}
 \label{fig:one_orbit_wrong_ref}
@@
 minus 5\uth percentile divided by the true action value for three different
-orbits. The \Gus{orange} orbit is exactly the same as in
-Figure~\ref{fig:one_orbit_wrong_ref}, and we also consider two more planar
-orbits with initial vertical velocities of $50\,\kms$ (\Gus{red}) and
-$10\,\kms$ (\Gus{blue}). For these two new orbits, the true actions are $(J_r,
-L_z, J_z) = (32.6, -1520, 22.9)\,\actunit$ and $(40.4, -1520, 0.7)\,\actunit$
-for the \Gus{red} and \Gus{blue} orbits, respectively. The value of
-$z_{\text{max}}$ for these orbits are $854\,\pc$ and $122\,\pc$, respectively
-(with the orange orbit having $z_{\text{max}}$ of $425\,\pc$.) \Gus{Easy way
-to explain change in $J_r$?}
+orbits. The thick (brown) orbit is exactly the same as in
+Figure~\ref{fig:one_orbit_wrong_ref}, and we also consider the thin and halo
+orbits (pink and blue, respectively) from Table~\ref{tab:orbits}.
 
@@
 The {\em upper right} panel shows that the fractional error in $J_z$ is more
-exaggerated for more planar orbits. For the nearly planar (blue) orbit,
-deviations that are $\sim30\,\pc$ already give $100\%$ deviations in the
-actions. The required offset for $100\%$ deviation is $\sim120$ and
-$\sim225\,\pc$ for the orange and red orbits, respectively. \Gus{Make these
-numbers exact.}
+exaggerated for more planar orbits. For the thin orbit, deviations that are
+$\sim30\,\pc$ already give $100\%$ deviations in the actions. The required
+offset for $100\%$ deviation is $\sim225\,\pc$ for the thick orbit. The halo
+orbit is relatively resistant to errors in the midplane, with only $\sim25\%$
+error in $J_z$ out to a midplane offset of $500\,\pc$.
 

### 2019-01-29 paper/ms.tex
More small updates to text
Update to many orbits caption

@@
 for $J_r$, with some deviations resulting in $L_z$ and relatively small
-deviations in $J_z$. In the {\em bottom left} panel the orange and blue lines
-overlap, and in the {\em bottom middle} and {\em bottom right} panels all
-three lines overlap.
+deviations in $J_z$. In the {\em bottom middle} and {\em bottom right} panels
+all three lines overlap.
 
@@
 \plotone{fig/schmactions_many_orbits.pdf}
-\caption{\Gus{update figure} We report the 95\uth minus 5\uth percentile of
-the error in the measured action ($\Delta J_i$) from coordinate system errors
-for three orbits with low-, medium-, and high-$J_z$ values (blue, orange, and
-red, respectively). The {\em left}, {\em center}, and {\em right} panels show
-the result for $J_r$, $L_z$, and $J_z$, respectively. The {\em upper} panels
-consider an offset in $z$ and the {\em lower} panels consider an offset in $x$
-(equivalently, an offset in the Solar radius).}
+\caption{We report the 95\uth minus 5\uth percentile of the error in the
+measured action ($\Delta J_i$) from coordinate system errors for the thin,
+thick, and halo orbits (Table~\ref{tab:orbits}). The {\em left}, {\em center},
+and {\em right} panels show the result for $J_r$, $L_z$, and $J_z$,
+respectively. The {\em upper} panels consider an offset in $z$ and the {\em
+lower} panels consider an offset in $x$ (equivalently, an offset in the Solar
+radius).}
 \label{fig:many_orbit_wrong_ref}

### 2019-01-29 paper/ms.tex
Add caption to histogram plot

@@
 \plotone{fig/schmactions_Jz_hist.pdf}
-\caption{Caption.}
+\caption{A histogram of the observed values in $J_z$ for the thick orbit
+assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
+values have a bias (from e.g. an incorrectly computed midplane), then the
+induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
+sort of error propagation must take this into account.}
 \label{fig:Jz_hist}

### 2019-01-29 paper/ms.tex
Add thincolor, thickcolor, and halocolor commands

@@
 \newcommand{\eeq}{\end{equation}}
+
+\newcommand{\thincolor}{pink}
+\newcommand{\thickcolor}{brown}
+\newcommand{\halocolor}{blue}
 
@@
 minus 5\uth percentile divided by the true action value for three different
-orbits. The thick (brown) orbit is exactly the same as in
+orbits. The thick (\thickcolor) orbit is exactly the same as in
 Figure~\ref{fig:one_orbit_wrong_ref}, and we also consider the thin and halo
-orbits (pink and blue, respectively) from Table~\ref{tab:orbits}.
+orbits (\thincolor and \halocolor, respectively) from Table~\ref{tab:orbits}.
 
@@
 
-Figure~\ref{fig:Rn_mc} is most informative for the low $J_z$ orbit (blue). The
-neighborhood around $\sim100\Msun$ clusters can be as small as $\sim300\,\pc$
-but up to $\sim1.2\,\kpc$ for m12i and m12f, though the situation is
-marginally better for m12m with values ranging from $\sim0.8\,\kpc$ to
-$\sim3\,\kpc$. The neighborhood values steadily increase with $m_c$, topping
-out at $\sim5\,\kpc$ for $\sim10^4\,\Msun$ clusters in m12i and m12f. In m12m,
-the neighborhood value reaches the maximum allowed of $16.4\,\kpc$ for
-clusters of mass $\sim800\,\Msun$. The interpretation of this is that the
-expected action space dispersion is so high for such high mass clusters that
-the measured $z$~offset is not large enough throughout the entire galaxy to be
-important.
+Figure~\ref{fig:Rn_mc} is most informative for the thin orbit (\thincolor).
+The neighborhood around $\sim100\Msun$ clusters can be as small as
+$\sim300\,\pc$ but up to $\sim1.2\,\kpc$ for m12i and m12f, though the
+situation is marginally better for m12m with values ranging from
+$\sim0.8\,\kpc$ to $\sim3\,\kpc$. The neighborhood values steadily increase
+with $m_c$, topping out at $\sim5\,\kpc$ for $\sim10^4\,\Msun$ clusters in
+m12i and m12f. In m12m, the neighborhood value reaches the maximum allowed of
+$16.4\,\kpc$ for clusters of mass $\sim800\,\Msun$. The interpretation of this
+is that the expected action space dispersion is so high for such high mass
+clusters that the measured $z$~offset is not large enough throughout the
+entire galaxy to be important.
 
@@
 \plotone{fig/Rn_vs_mc.pdf}
-\caption{The neighborhood around an open cluster of mass $m_c$ for our three
-orbits (blue, orange, and red) and for each of the FIRE galaxies considered
-here --- m12i, m12f, and m12m ({\em left}, {\em center}, and {\em right}
-panels, respectively). Given the $z$ offset for a cluster of mass $m_c$
-computed in Figure~\ref{fig:cluster_offset}, we convert this to a chord length
-using the result from Figure~\ref{fig:range_deltaphi} and plot as a solid blue
-line. We interpret this chord length as the ``neighborhood'' around the
-cluster, i.e. the distance from the cluster one could go before the $J_z$
-error induced by our midplane effect is comparable to the intrinsic action
-dispersion of the cluster. When the $z$ offset is larger than the maximum
-range in Figure~\ref{fig:range_deltaphi}, we report the maximum chord length
+\caption{The neighborhood around an open cluster of mass $m_c$ for the thin,
+and thick orbits (\thincolor and \thickcolor, respectively; see
+Table~\ref{tab:orbits}) and for each of the FIRE galaxies considered here ---
+m12i, m12f, and m12m ({\em left}, {\em center}, and {\em right} panels,
+respectively). Given the $z$ offset for a cluster of mass $m_c$ computed in
+Figure~\ref{fig:cluster_offset}, we convert this to a chord length using the
+result from Figure~\ref{fig:range_deltaphi} and plot as a solid blue line. We
+interpret this chord length as the ``neighborhood'' around the cluster, i.e.
+the distance from the cluster one could go before the $J_z$ error induced by
+our midplane effect is comparable to the intrinsic action dispersion of the
+cluster. When the $z$ offset is larger than the maximum range in
+Figure~\ref{fig:range_deltaphi}, we report the maximum chord length
 ($2\times8.2\,\kpc$). We aso report the same procedure but using the

### 2019-01-29 paper/ms.tex
Fix caption

@@
 the result from Figure~\ref{fig:many_orbit_wrong_ref}. We consider the result
-for three different orbits with varying vertical actions of $J_z = 0.7, 7.0,
-22.9\,\actunit$ ($z_{\text{max}} = 122,425,854\,\pc$) --- colored by
-\Gus{blue, orange, red}, respectively. Nearby open clusters have actions
-closest to the $J_z=0.7\,\actunit$ (blue) orbit, see Table~\Gus{x}.}
+for the thin and thick orbit (Table~\ref{tab:orbits}). Nearby open clusters
+(Table~\ref{tab:real_clusters}) have orbits closest to the thin orbit .}
 \label{fig:cluster_offset}

### 2019-01-29 paper/ms.tex
Update Rn plot caption

@@
 Figure~\ref{fig:cluster_offset}, we convert this to a chord length using the
-result from Figure~\ref{fig:range_deltaphi} and plot as a solid blue line. We
-interpret this chord length as the ``neighborhood'' around the cluster, i.e.
-the distance from the cluster one could go before the $J_z$ error induced by
-our midplane effect is comparable to the intrinsic action dispersion of the
-cluster. When the $z$ offset is larger than the maximum range in
-Figure~\ref{fig:range_deltaphi}, we report the maximum chord length
-($2\times8.2\,\kpc$). We aso report the same procedure but using the
-$\pm1\sigma$ lines from Figure~\ref{fig:range_deltaphi}, which we plot here as
-dashed lines. For the medium- and high-$J_z$ orbits (orange and red), the
-neighborhood is quite large and thus the effect is negligible. However, for
-the low-$J_z$ orbit (blue), which is closest to the orbits of nearby open
-clusters (see Table~\Gus{x}), the neighborhood is only a few hundred $\pc$ for
-low-mass clusters.}
+result from Figure~\ref{fig:range_deltaphi}. We interpret this chord length as
+the ``neighborhood'' around the cluster, i.e. the distance from the cluster
+one could go before the $J_z$ error induced by our midplane effect is
+comparable to the intrinsic action dispersion of the cluster. When the $z$
+offset is larger than the maximum range in Figure~\ref{fig:range_deltaphi}, we
+report the maximum chord length ($2\times8.2\,\kpc$). We aso report the same
+procedure but using the $\pm1\sigma$ lines from
+Figure~\ref{fig:range_deltaphi}, which we plot here as dashed lines. For the
+thick orbit (\thickcolor), the neighborhood is quite large and thus the effect
+is negligible. However, for the thin orbit (\thincolor), which is closest to
+the orbits of nearby open clusters (see Table~\ref{tab:real_clusters}), the
+neighborhood is only a few hundred $\pc$ for low-mass clusters.}
 \label{fig:Rn_mc}

### 2019-01-29 midplane_test/plot_midplane.py
Print out midplane mid90th values upon run

@@ -61,6 +61,10 @@ for gal,ax_col in zip(glist, ax.transpose()):
     ax_col[1].plot(theta/np.pi, (err_high-fit)*1000, c=tb_c[0], ls='dashed', alpha=0.5)
     ax_col[1].fill_between(theta/np.pi, (err_high-fit)*1000, (err_low-fit)*1000, color=tb_c[0], alpha=0.25)
 
+    midplane_up = np.percentile(midplane_est - fit, 95)
+    midplane_low = np.percentile(midplane_est - fit, 5)
+    print(gal, midplane_up - midplane_low)
+
 ax[0][0].set_ylabel(r'$\text{midplane}\,[\,\text{pc}\,]$')
 ax[1][0].set_ylabel(r'$\text{midplane}\,[\,\text{pc}\,]$')
 

### 2019-01-29 paper/ms.tex
Update numbers

@@
 we also subtract the best fit $\text{A} \sin{\left(\phi + \text{B}\right)} +
-\text{C}$ curve from the midplane as a function of azimuth. For $\text{A}$,
-the values are $-165$, $45$, $9\,\pc$ and for $\text{C}$ the values are $-69$,
-$19$, and $-18\,\pc$ for m12i, m12f, and m12m, respectively \Gus{Double check
-numbers}. For the assumed solar radius of $8.2\,\kpc$, we can approximate the
-angle offset for the $z$-axis from the values of $\text{A}$ --- we compute
-\Gus{$1.15$, $0.31$, and $0.06\,\deg$} for m12i, m12f, and m12m. These angle
-offsets are consistent with the values given in \citet{2018arXiv180610564S}
-for the difference between the $z$-axis as defined by the gas and stars.
+\text{C}$ curve from the midplane as a function of azimuth. For $\text{A}$ the
+values are $-174$, $62$, and $3.8\,\pc$, for $\text{B}$ the values are $0.3$,
+$0.65$, and $0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$,
+and $-15\,\pc$ for m12i, m12f, and m12m, respectively. For the assumed solar
+radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
+from the values of $\text{A}$ --- we compute $1.22$, $0.43$, and $0.03\,\deg$
+for m12i, m12f, and m12m. These angle offsets are consistent with the values
+given in \citet{2018arXiv180610564S} for the difference between the $z$-axis
+as defined by the gas and stars.
 

### 2019-02-04 paper/ms.tex
Add figure to explain range-dphi plot

@@
 \begin{figure*}
+\plotone{fig/fig_to_explain.pdf}
+\caption{A cartoon explanation of Figure~\ref{fig:range_deltaphi}. An observer
+is placed at the blue x in the plane of a backgrround galaxy (lime). For a
+given $\Delta \phi$ (blue) centered on the observer, we record the range of
+midplane values for this section of the Solar circle (orange). We plot the
+value of the range against the value of $\Delta \phi$. We are also able to
+convert the value of $\Delta \phi$ into a chord length (pink), which we plot
+as a secondary, upper $x$-axis. We repeat the procedure for each initial
+$\phi$ (gray lines), and also compute the mean and $\pm1\,\sigma$ values (dark
+blue).}
+\label{fig:fig_to_explain}
+\end{figure*}
+
+\begin{figure*}
 \plotone{fig/range_dphi.pdf}
@@
 a solid blue line and the $\pm1\sigma$ lines as dashed blue lines. The upper
-$x$-axis shows the chord length from the position $\phi$ to
-$\pm\Delta\phi/2$.}
+$x$-axis shows the chord length from the position $\phi$ to $\pm\Delta\phi/2$.
+A cartoon explanation of this Figure is given in
+Figure~\ref{fig:fig_to_explain}.}
 \label{fig:range_deltaphi}

### 2019-02-04 paper/ms.tex
Move Megan to ack

@@
 \author{Daniel Angl\'es-Alc\'azar}
-\affil{\cca}
-
-\author{Megan Bedell}
 \affil{\cca}
@@
 
+\acknowledgments
+We would like to thank David W. Hogg, Kathryn Johnston, and Megan Bedell for
+many useful discussions. A.B. would like to thank Todd Phillips for helpful
+discussions. A.B. was supported in part by the Roy \& Diana Vagelos Program in
+the Molecular Life Sciences and the Roy \& Diana Vagelos Challenge Award. The
+work of \Gus{everyone} is supported by the Simons Foundation.
+
 \appendix \section{Appendix 1}

### 2019-02-04 paper/ms.tex
Add barebones conclusion
Add orbit plot to appendix

@@
 
+\section{Conclusions}\label{sec:conclusion}
+Let's conclude, shall we.
+
+\begin{itemize}
+\item
+
+\end{itemize}
+
 \acknowledgments
@@
 
-\appendix \section{Appendix 1}
-Gonna write an appendix probably
+\appendix
+\section{Orbits} \label{app:orbits}
+We plot the three orbits considered throughout the work
+(Table~\ref{tab:orbits}) in Figure~\ref{fig:plot_orbits}.
+
+\begin{figure*}
+\plotone{fig/orbits.pdf}
+\caption{The three orbits presented in Table~\ref{tab:orbits} and considered
+throughout the work. We plot the thin, thick, and halo orbits in the {\em
+left}, {\em center}, and {\em right} columns, respectively. The {\em upper}
+row shows a plot of $x$ vs. $y$ while the {\em lower} row shows $R$ vs. $z$.}
+\label{fig:plot_orbits}
+\end{figure*}
 

### 2019-02-04 paper/ms.tex
Make ack in alpha order
Fix reference to appendix

@@
 their initial position in phase space and the actions of their integrated
-orbits in Table~\ref{tab:orbits}. Each orbit is plotted in \Gus{Appendix X}.
-We will refer to these orbits by their names henceforth.
+orbits in Table~\ref{tab:orbits}. Each orbit is plotted in
+Appendix~\ref{app:orbits}. We will refer to these orbits by their names
+henceforth.
 
@@
 \acknowledgments
-We would like to thank David W. Hogg, Kathryn Johnston, and Megan Bedell for
+We would like to thank Megan Bedell, David W. Hogg, and Kathryn Johnston for
 many useful discussions. A.B. would like to thank Todd Phillips for helpful

### 2019-02-04 paper/ms.tex
Fix reference to old lines in fig

@@
 
-For the medium $J_z$ orbit (orange), the situation is better, although such an
-orbit is not a good fit to any of the nearby open clusters \Gus{Table x}. For
-m12i and m12f, the neighborhood is $\sim \text{few}\,\kpc$ for
-$m_c\sim10^2\,\Msun$. At $\sim10^3\,\Msun$ the curves reach their maximum
-value. For m12m, the neighborhood starts at $\sim10\,\kpc$ and quickly rises
-to the maximum value. For the high $J_z$ orbit (red), the neighborhood is at
-the maximum value for nearly all masses, except at the very low-mass end in
-m12i.
+For the thick orbit (brown), the neighborhood is at the maximum value for
+nearly all masses, except at the very low-mass end in m12i.
 

### 2019-02-10 paper/ms.tex
Add lsr to paper in Appendix

@@
 
-\subsection{Local Midplane} \label{sec:local_midplane}
+\subsection{Local Midplane} \label{ssec:local_midplane}
 We now turn to defining the local midplane that an observer might measure if
@@
 
+\section{LSR Variations} \label{app:lsr}
+\begin{figure*}
+\plotone{fig/lsr.pdf}
+\caption{The local standard of rest (LSR) as a function of azimuth at the
+fiducial Solar circle ($R_{\odot} = 8.2\,\kpc$). No best-fit subtraction is
+performed here as we did in the case of the midplane
+(Section~\ref{ssec:local_midplane}).}
+\end{figure*}
+
 \bibliography{references}

### 2019-02-10 paper/ms.tex
Add text to introduce lsr in app

@@
 \section{LSR Variations} \label{app:lsr}
+We consider the variations of the LSR as a function of azimuth at the fiducial
+solar circle ($R_{\odot} = 8.2\,\kpc$). At each azimuth, $\phi$, we take the
+median velocity in cylindrical coordinates of all stars within $200\,\pc$ of
+the position, following \citet{2018arXiv180610564S}. No best-fit subtraction
+was performed as in Figure~\ref{fig:midplane}.
+
 \begin{figure*}

### 2019-02-10 paper/ms.tex
Add stubsection referring to lsr plot in app

@@
 
+\subsection{LSR Variations} \label{ssec:lsr_var}
+For similar reasons to the midplane varying across the disk, we also expect
+that the local standard of rest (LSR) is a function of azimuth as well. We
+perform this calculation in Appendix~\ref{app:lsr}, but in order to perform a
+best-fit subtraction as in the previous section is more involved, which we
+delay to future work. We do note, however, that the variation in the LSR is
+less pronounced than for the midplane, at least for how we calculate the LSR.
+
 \section{Discussion} \label{sec:discussion}

### 2019-02-11 paper/ms.tex
Move discussions of actions up in intro

@@
 
-These measurements of the galactocentric coordinate system rely on the
-assumption of an axisymmetric Milky Way. Under this assumption, these
-estimates would provide the correct parameters for the Galactocentric
-coordinate system up to measurement error. However, the Milky Way is not
-axisymmetric in detail. Spiral arms, bar(s), and infalling satellite galaxies
-such as Sagittarius and the Large Magellanic Cloud all induce
-non-axisymmetries.
+Once a Galactocentric coordinate system has been established and a 6D phase
+space measurement of a star has been made, it is desirable to convert this
+measurement into action space. Actions are quantities which describe the orbit
+of a star and are conserved quantities under some assumptions which we
+describe next. They are the typical integral of the Hamiltonian in cylindrical
+coordinates and are given by,
+\beq\label{eq:actions}
+J_i \equiv
+\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
+\eeq
+where $i=R,z,\phi$.\footnote{We typically write $J_r$ instead of $J_R$ for
+aesthetic reasons.} Note that $J_{\phi} \equiv L_z$. In a slowly-evolving
+axisymmetric potential, the actions of stars' orbits are conserved quantities
+\citep{2008gady.book.....B,2014RvMP...86....1S}. The non-axisymmetries of a
+more realistic potential induce time evolution in the {\em measured} actions
+--- the true orbital invariants (if they exist) are invariant by definition.
 
@@
 modelling across large distances.
-
-Since an azimuthal-dependent midplane is effectively a departure from a
-perfectly axisymmetric potential, we consider the effect that an erroneous
-midplane has on the conservations of actions. The actions of an orbit are
-typically computed in a cylindrical coordinate system as,
-\beq\label{eq:actions}
-J_i \equiv
-\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
-\eeq
-where $i=R,z,\phi$. Note that $J_{\phi} \equiv L_z$. In a slowly-evolving
-axisymmetric potential, the actions of stars' orbits are conserved quantities
-\citep{2008gady.book.....B,2014RvMP...86....1S}. The non-axisymmetries of a
-more realistic potential induce time evolution in the {\em measured} actions
---- the true orbital invariants (if they exist) are invariant by definition.
-Through simple tests, we will see that even in a truly axisymmetric potential,
-if the wrong coordinate system is assumed then there will be additional time
-evolution in the measured actions.
 

### 2019-02-11 paper/ms.tex
Add summary of actions + Gaia dr2

@@
 --- the true orbital invariants (if they exist) are invariant by definition.
+
+In a slowly-evolving, axisymmetric potential, the actions of each star are
+time-invariant \citep{2008gady.book.....B}. In the Milky Way, stellar actions
+are expected to evolve on short time scales due to scattering with gas clouds
+and on long time scales due to resonances with spiral arms, bar(s), and other
+large scale perturbations \citep{2014RvMP...86....1S}. For this reason, we and
+other authors have used actions to study stellar scattering in the Milky Way
+disk using the improved astrometry of {\em Gaia} DR2 and various age
+catalogues \citep{2018ApJ...867...31B,2018arXiv180803278T}. Actions have also
+been used to study different models of spiral structure in the Milky Way
+\citep{2019MNRAS.tmp..155S}. A summary of the disk in action space is given by
+\citet{2018arXiv180503653T}.
 

### 2019-02-11 paper/ms.tex
Add list of conclusions

@@
 \begin{itemize}
-\item
+\item We demonstrated that an inaccuracy in the Galactocentric coordinate
+system induces time evolution in the {\em observed} actions
+(Figures~\ref{fig:cartoon}\&\ref{fig:one_orbit_wrong_ref}).
+
+\item We interpret this time evolution as an error in the computed action
+(Figure~\ref{fig:many_orbit_wrong_ref}). We focused on the effect that an
+inaccuracy in the midplane has on the vertical action $J_z$. We found that in
+order for there to be a $100\%$ error in $J_z$ (as defined by the middle
+$90\uth$ percentile) one needs a midplane error of $\sim30\,\pc$ and
+$\sim225\,\pc$ for a typical thin and thick disk orbit, respectively. The
+fractional error is significantly less for halo orbits.
+
+\item We pointed out that the error distribution of the actions induced by a
+coordinate system error is very non-Gaussian. The distribution is bimodal with
+{\em neither of the modes peaking at the true value}. As a result, error
+propagation is very involved when considering actions.
+
+\item We proposed that dynamical modelling across large regions of the disk is
+susceptible to this type of error, since the assumption that our local
+Galactic midplane is the global Galactic midplane is not true {\em a priori}.
+
+\item To demonstrate the previous point, we measured the local Galactic
+midplane along the Solar circle in three different high-resolution, zoom-in
+simulations of Milky Way mass galaxies from the FIRE collaboration. We found
+that the midplane varies by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
+three galaxies we considered (m12i, m12f, and m12m, respectively).
 

### 2019-02-11 paper/ms.tex
Add action-orbit integration sentence
Just blow the intro up

@@
 
-We define the ``global Galactic midplane'' to be the plane of $z$ symmetry in
-the axisymmetric model of the Milky Way. We define the ``local Galactic
-midplane'' to be the plane of $z$ symmetry inferred by an observer at a
-particular location. The true global Galactic midplane is the plane of $z$
-symmetry in the best-fit axisymmetric model. It is not guaranteed that the
-local Galactic midplane coincide with the global Galactic midplane. Even
-without a high-precision definition and measurement of the global Galactic
-midplane, we can still assess how different the two should be by measuring the
-azimuthal dependence of the local Galactic midplane. This procedure may be
-readily tractable with current datasets. However, we turn to simulation in
-this work.
-
-It's important for us to emphasize that when we state our local Galactic
-midplane may be erroneous, we do not mean that past measurements/estimates of
-it are wrong. Rather, what we mean is that it may vary from the local midplane
-at other points in the Galaxy. In the face of {\em Gaia} DR2 (and soon DR3/4),
-this may be an important effect to consider when performing dynamical
-modelling across large distances.
-
 The transformation to action space is not trivial, and many of the assumptions
@@
 
-To provide a realization of a realistic Galactic potential, we turn to
-cosmological zoom-ins of Milky Way-mass galaxies from the Feedback in
-Realistic Environments (FIRE)
+We propose a new type of error in the computation of actions which has
+implications for general dynamical modelling across the disk. Our main point
+is that the ``local'' Galactic midplane is not the same as the ``global''
+Galactic midplane. If we observe a group of stars that are
+non-local\footnote{i.e. a group of stars that is far enough away. The exact
+distance depends on the accuracy of the measurement being made and the goals
+of the particular problem.}, they will know of a different local midplane than
+our local midplane extrapolated onto their position. This means that when we
+convert those star's positions and velocities from a heliocentric to a
+Galactocentric coordinate system, we are inducing a systematic bias in those
+stars' $z$ values. The same applies to inaccuracies in measurements of
+$R_{\odot}$ and variations in the local standard of rest (LSR).
+
+We do not propose what a modeller should use for the ``true'' global midplane,
+since such a definition probably depends on the particular problem such a
+modeller is facing. Even without a high-precision definition and measurement
+of the global Galactic midplane, we can still assess how different we should
+expect the local midplane to be by measuring the azimuthal dependence of the
+local midplane. This procedure may be readily tractable over some of the disk
+with current datasets. However, we turn to simulation in this work because we
+can investigate the entire disk. Specifically, we use cosmological zoom-ins of
+Milky Way-mass galaxies from the Feedback in Realistic Environments (FIRE)
 collaboration\footnote{\url{https://fire.northwestern.edu}} (the {\em Latte}
 suite of simulations).
+
+It is important for us to emphasize that when we state our local Galactic
+midplane may be erroneous, we do not mean that past measurements/estimates of
+it are wrong. Rather, what we mean is that it may vary from the local midplane
+at other points in the Galaxy. In the face of {\em Gaia} DR2 (and soon DR3/4),
+this may be an important effect to consider when performing dynamical
+modelling across large distances.
 
@@
 
+While in this work we have focused on errors in action computation, all of our
+conclusions also extend to studies of stars that simply rely on orbit
+integration, since the computation of actions and orbit integrations are
+essentially equivalent.
+
 \acknowledgments

### 2019-02-11 paper/ms.tex
Rewrite abstract

@@
 
-In a completely axisymmetric potential, the notion of a global Galactic
-midplane is well defined. However, since the Milky Way is not completely
-axisymmetric, the global Galactic midplane is to some level of precision not
-well defined. The local Galactic midplane has been found by determining the
-height at which the number counts of stars reaches a maximum. This local
-midplane is then used as the global midplane, an assumption which would be
-justified in a completely axisymmetric potential. Assuming a global Galactic
-midplane exists, we show in three cosmological zoom-in simulations from the
-FIRE collaboration that local determinations of the midplane can vary by up to
-a thin disk scale height. The effect is more pronounced in galaxies with less
-quiescent histories (e.g. Sagittarius-like encounters). We explore one
-consequence of such an inaccuracy of the midplane in the context of orbital
-actions. Using standard orbit integration techniques, we show that midplane
-inaccuracy induces artifical time evolution in the vertical action that makes
-it virtually unusable for individual stars in the disk. The situation is less
-dire for angular momentum and radial action determinations, though we also
-show that inaccuracies in the solar radius can have a similar effect on the
-latter. As a result, we question the use of actions as dynamical invariants in
-the Milky Way disk.
+We propose a new type of error in the computation of stellar actions. This
+error is induced by the fact that the local Galactic midplane varies across
+the disk. As a result, when we make 6D phase space measurements of stars that
+are non-local (i.e. far away), we are observing those stars with a systematic
+offset in $z$. We show that this systematic offset induces time evolution in
+the observed actions which we interpret as an error. We note that this error
+is non-Gaussian and bimodal, with neither of the modes peaking on the true
+action value. A midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is
+necessary to induce a $100\,\%$ error in the vertical action $J_z$ for thin
+and thick disk orbits, respectively. Using cosmological zoom-in simulations
+from the FIRE collaboration, we show that the midplane can vary by
+$\sim100\,\pc$ around the Solar circle, indicating that such an effect is
+important for understanding the thin and thick disk. Our discussion is also
+directly applicable to variations in the local standard of rest. The variation
+of the midplane must be taken into account when performing dynamic modelling
+across large regions of the disk accessible with the {\em Gaia} mission.
 

### 2019-02-11 paper/ms.tex
Change title
Tweaks to abstract

@@
 
-\title{Our Galactic Midplane Is Probably Wrong, and Not by a Little}
+\title{Issues with Dynamical Measurements Across the Disk}
 
@@
 
-We propose a new type of error in the computation of stellar actions. This
-error is induced by the fact that the local Galactic midplane varies across
-the disk. As a result, when we make 6D phase space measurements of stars that
-are non-local (i.e. far away), we are observing those stars with a systematic
-offset in $z$. We show that this systematic offset induces time evolution in
-the observed actions which we interpret as an error. We note that this error
-is non-Gaussian and bimodal, with neither of the modes peaking on the true
-action value. A midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is
+We demonstrate the effect of midplane errors on the computation of stellar
+actions. This error is induced by the fact that the local Galactic midplane
+varies across the disk. As a result, when we make 6D phase space measurements
+of stars that are non-local (i.e. far away), we are observing those stars with
+a systematic offset in $z$. We show that this systematic offset induces time
+evolution in the observed actions which we interpret as an error. We note that
+this error is non-Gaussian and bimodal, with neither of the modes peaking on
+the true action value. A midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is
 necessary to induce a $100\,\%$ error in the vertical action $J_z$ for thin
@@
 directly applicable to variations in the local standard of rest. The variation
-of the midplane must be taken into account when performing dynamic modelling
+of the midplane must be taken into account when performing dynamical modelling
 across large regions of the disk accessible with the {\em Gaia} mission.

### 2019-02-12 paper/ms.tex
add selection function statement
revisions to sections 2.1 and 2.2

@@
 local midplane. This procedure may be readily tractable over some of the disk
-with current datasets. However, we turn to simulation in this work because we
-can investigate the entire disk. Specifically, we use cosmological zoom-ins of
-Milky Way-mass galaxies from the Feedback in Realistic Environments (FIRE)
+with current datasets, using appropriate selection functions. However, we turn
+to simulation in this work because we can investigate the entire disk.
+Specifically, we use cosmological zoom-ins of Milky Way-mass galaxies from the
+Feedback in Realistic Environments (FIRE)
 collaboration\footnote{\url{https://fire.northwestern.edu}} (the {\em Latte}
@@
 We present a cartoon in Figure~\ref{fig:cartoon} to show that an inaccurate
-determination of the midplane can lead to a phase-dependence of the {\em
-observed} actions. The solid gray line indicates a hypothetical ``true''
-orbit. The y-axis corresponds to the vertical height of the orbit and the
-x-axis orbital phase. The dashed gray line is an incorrect midplane which an
-observer uses to compute orbits.
+determination of the midplane can lead to a phase-dependence of the actions calculated from an orbit specified by a phase-space point and an assumed potential.
+The y-axis corresponds to the vertical height of the orbit and the x-axis orbital phase.
+The solid gray line indicates the ``true'' orbit of the star; that is, its orbit as it oscillates around the true midplane.
+The dashed gray line is an erroneous midplane assigned to the potential model which an observer uses to integrate the orbit of the star and hence estimate its actions.
+The model potential is otherwise identical to the one in which the star is actually moving.
 
 Now suppose this observer makes a measurement of the true orbit at the blue
-point or the pink point. Then the blue and pink lines correspond to the orbits
-that the observer would compute for each point. In action space, this would
-correspond to a different value of $J_z$ for the blue and pink points. In this
-way, assuming the wrong coordinate system induces time dependence in the
-actions {\em computed by an observer}.
+point or the pink point (i.e.
+at two different orbital phases).
+Then the blue and pink lines correspond to the orbits that the observer would compute for each point based on the potential model with the erroneous midplane.
+In action space, this would correspond to a different value of $J_z$ for the blue and pink points.
+In this way, assuming the wrong coordinate system induces a phase-dependence in the
+actions estimated for the star, which in the correct potential (in this example, the one with the correct midplane) should be phase-independent.
 
@@
 \plotone{fig/cartoon.pdf}
-\caption{Cartoon showing the effect an incorrect midplane can have on orbit
-integrations and as an extension on action computations. The x-axis shows the
-orbital phase and the y-axis the vertical height. The top gray curve depicts a
-hypothetical true orbit oscillating about the true midplane, the horizontal
-solid gray line. Consider an immortal observer who erroneously assumes the
-midplane is located at the horizontal dashed line. Suppose this immortal
-observer measures the orbit at the blue and pink points. If the immortal
-observer were to perform orbit integrations at each observation, they would
-assume the blue and pink curves for the respective orbits. In this way, an
-incorrect midplane will induce phase-dependence in the {\em observed}
-actions.}
+\caption{Cartoon showing the effect an error in the determination of the coordinate midplane can have on orbit integration and action estimation. The x-axis shows the
+orbital phase and the y-axis the vertical height. The top gray curve depicts an example ``true'' orbit oscillating about the true midplane (horizontal
+solid gray line). Consider an observer who erroneously assumes the
+midplane is located at the horizontal dashed line. Suppose this
+observer measures the phase-space position of the star at two different orbital phases (blue and pink points).
+If the observer were then to integrate the star's orbit using a model potential with the erroneous midplane, they would
+obtain the blue and pink curves for the star's orbit, respectively.
+The actions estimated from these two  erroneous orbits would subsequently differ, both from each other and from the actions estimated for the true orbit (in the potential with the correct midplane).
+Hence an
+incorrect midplane in the potential model assumed will induce phase-dependence in the actions estimated for a given star in that potential.}
 \label{fig:cartoon}
@@
 \subsection{Summary of Action Computation} \label{ssec:action_comp}
-We follow our previous work in computing actions \citep{2018ApJ...867...31B}.
-We use the code \texttt{gala} v0.3 to perform orbit integrations and
-conversion to action space \citep{2017JOSS....2..388P,Price-Whelan:2018},
-which uses an action estimator technique presented by
-\citet{2014MNRAS.441.3284S}. We use the default \texttt{MWPotential} as our
+We compute actions as in our previous work \citep{2018ApJ...867...31B}, using the code \texttt{gala} v0.3 to perform orbit integrations and
+conversion to action space \citep{2017JOSS....2..388P,Price-Whelan:2018}.
+To compute actions we use the torus-mapping technique first presented by \citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to calculate actions for an orbital time-series starting from a phase-space position $(x, v)$ and integrated in a potential $\Phi$.
+We use the default \texttt{MWPotential} as our
 potential, which is based on the Milky Way potential available in
@@
 is fit to empirically match some observations. We use the Dormand-Prince
-8(5,3) integration scheme \citep{Dormand80:integrator}. We use a timestep of
+8(5,3) integration scheme \citep{Dormand80:integrator} with a timestep of
 $1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
@@
 Other methods for computing actions are used in the literature. For example,
-the St\"ackel Fudge method \citep{2016MNRAS.457.2107S} which uses the
-St\"ackel potential to approximate the Galactic potential
-\citep{1985MNRAS.216..273D,2012MNRAS.426.1324B} is commonly used
-\citep{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}. For
-applications in the disk, the St\"ackel Fudge method is of acceptable
-accuracy, but since we also consider halo orbits in this work we choose to use
-orbit integration throughout \citep{2016MNRAS.457.2107S}.
+the St\"ackel Fudge method \citep{2016MNRAS.457.2107S}, which uses a single
+St\"ackel potential (with analytic actions) to approximate the Galactic potential
+\citep{1985MNRAS.216..273D,2012MNRAS.426.1324B}, was used in many recent works exploring actions in the Galactic disk
+\citep[e.g.][]{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}. For
+disk-like orbits, existing implementations of the St\"ackel Fudge method are of acceptable
+accuracy, but since we also consider halo-like orbits in this work we choose to use
+orbit integration and torus mapping throughout \citep{2016MNRAS.457.2107S}.
 

### 2019-02-12 paper/ms.tex
edits to the first half of 2.3, and some suggestions for different section headings

@@
 
-\section{Impact on Actions} \label{sec:ref_frame}
-We first demonstrate the importance of measuring the Galactic midplane and
-Solar radius to very high accuracy. To be clear, the consequences we explore
-here may also arise from various other observational errors. For instance, the
-symmetrized Galactic potential used by observers may not be a good description
-of the true potential --- or the parameters used may be incorrect. In this
-Section, we assume these approximations are valid and simply explore the
-consequences of an inaccurate Galactocentric coordinate system.
+\section{Motivation} \label{sec:ref_frame}
+We first demonstrate the significance to action computations, especially for disk-like orbits, of systematic errors in the determination of the Galactic midplane and distance to the Galactic center.
+The consequences we explore
+here may also arise from various other systematic errors. For instance, the
+axisymmetric Galactic potential model used in many works to compute actions may not be a good description of the true potential --- or the parameters used may yield a potential that is systematically incorrect outside an original fitted region.
+In this section, we assume that the Galaxy is perfectly described by our model axisymmetric potential, and simply explore the consequences of systematic errors in the determination of the Galactocentric coordinate system.
 
-\subsection{Cartoon} \label{ssec:cartoon}
-We present a cartoon in Figure~\ref{fig:cartoon} to show that an inaccurate
-determination of the midplane can lead to a phase-dependence of the actions calculated from an orbit specified by a phase-space point and an assumed potential.
+\subsection{The effect of midplane determination on orbit and action calculations} \label{ssec:cartoon}
+We present a cartoon in Figure~\ref{fig:cartoon} to show how an inaccurate
+determination of the midplane leads to a phase-dependence of the actions calculated from an orbit specified by a phase-space point and an assumed potential model.
 The y-axis corresponds to the vertical height of the orbit and the x-axis orbital phase.
@@
 
-\subsection{Summary of Action Computation} \label{ssec:action_comp}
+\subsection{Numerical methods} \label{ssec:action_comp}
 We compute actions as in our previous work \citep{2018ApJ...867...31B}, using the code \texttt{gala} v0.3 to perform orbit integrations and
@@
 
-\subsection{Quantification} \label{ssec:quant}
-First, we illustrate the time-dependence of one possible inaccurate reference
-frame. Throughout this work we will consider three orbits typical of the
+\subsection{Quantification of the effect of systematic coordinate-system errors} \label{ssec:quant}
+First, we illustrate the phase-dependence of an example systematic error in the Galactocentric coordinate system.
+We will consider three orbits in the model potential described in \S \ref{ssec:action_comp} that are typical of stars in the
 stellar thin disk, stellar thick disk, and the stellar halo. We summarize
-their initial position in phase space and the actions of their integrated
-orbits in Table~\ref{tab:orbits}. Each orbit is plotted in
-Appendix~\ref{app:orbits}. We will refer to these orbits by their names
+their initial positions in phase space and the actions computed by integrating their
+orbits in the correct potential in Table~\ref{tab:orbits}. Each orbit, integrated without systematic coordinate errors, is plotted in
+Appendix~\ref{app:orbits}. We will refer to these orbits by their names (thin-disk, thick-disk, halo)
 henceforth.
 
-We begin by considering the thick orbit. Suppose that an observer makes
-observations of this orbit in a coordinate system that is incorrect in height
-by $100\,\pc$ --- i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each
+We begin by considering the thick-disk orbit.
+Suppose that we could observe the thick-disk star's phase-space position at many different times (and hence different orbital phases), but used a coordinate system in which the midplane is systematically offset in height by $100\,\pc$ from its actual location --- i.e.
+we subtract the vector $(0, 0, 100)\,\pc$ from each
 position in the orbit. This corresponds to an observer physically located at
-the position $(8, 0, 0)\,\kpc$ but erroneously thinking they are located at
+the position $(8, 0, 0)\,\kpc$ in the coordinate system of the true potential, but erroneously thinking they are located at
 $(8, 0, 0.1)\,\kpc$.
@@
 
-Using the observations of the orbit that this erroneous (and immortal)
-observer makes every $\Myr$, we perform the same orbit integration procedure
-as before to compute the actions our observer would observe in the incorrect
-coordinate system over the duration of the orbit. The erroneous computed
-actions are shown for the first $\Gyr$ of the orbit in the {\em upper} panels
-Figure~\ref{fig:one_orbit_wrong_ref}. Occasionally the numerical scheme fails
-and very large actions are reported by \texttt{gala}
---- we perform a $5\sigma$ clip on each action to exclude such orbits, but
-    this only excludes a total of $5$ orbits out of the $1000$ considered for
+Starting from each observation of the star's phase-space position that such an (extremely long-lived)
+observer makes every $\Myr$, we then compute the actions as outlined above using an integration from that starting point in phase space carried out using the true potential (but specifying the star's present-day position using the systematically offset coordinate system).
+The erroneous computed
+actions for each phase-space starting point are shown for the first $\Gyr$ of the orbit in the {\em upper} panels Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical scheme fails
+and very large actions are reported by \texttt{gala}---we perform a $5\sigma$ clip on each action to exclude such orbits, but this only excludes a total of $5$ orbits out of the $1000$ considered for
     Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but
-    the vast majority of orbits are computed properly. We perform the same
+    the vast majority of orbits are computed properly.}
+
+We also perform the same
     procedure but assume a $100\,\pc$ offset in the $x$ component (i.e.
     subtracting the vector $(100, 0, 0)\,\pc$) in the {\em lower} panels. This
-    is equivalent to an observer making an incorrect measurement of the Solar
-    radius.
+    is equivalent to a systematic offset in the determination of the distance from the Sun to the Galactic center.
 
-Figure~\ref{fig:one_orbit_wrong_ref} shows that the erroneous observer would
-infer significant time evolution in their computed actions. For an error in
-$z$ ({\em upper} panels), the 95\textsuperscript{th} minus
+Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the offset coordinate systems depend on the time (i.e.
+orbital phase) at which the star's phase-space position is observed.
+(To our extremely long-lived observer this would appear as time-dependence in the actions, which appear to oscillate around their value using the correct coordinate system even though in this example the Galaxy is described perfectly by a static, axisymmetric potential.) The relative size of the phase variation in each action depends on the direction of the systematic offset as well as the true values of the actions (i.e.
+the type of orbit).
+For a systematic offset in $z$ ({\em upper} panels), the 95\textsuperscript{th} minus
 5\textsuperscript{th} percentiles are $2.2$ and $6.2 \actunit$ for $J_r$ and
@@
 respectively. The error induced in $L_z$ is negligible, as expected since
-$L_z$ does not depend on the value of $z$. It is worth pointing out that a
+$L_z$ is independent of $z$. It is worth pointing out that a
 $100\,\pc$ error in an orbit with $z_{\text{max}}=850\,\pc$ --- a $12\%$ error
@@
 
-For an error in the $x$ component (or Solar radius), the
+For a systematic offset in $x$ (or Galactic center distance), the
 95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$, $47$,

### 2019-02-12 paper/ms.tex
edits to the second half of 2.3, some suggestions for terminology changes, and a new proposal for the name of section 3

@@
 \startdata 
-thin & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 \\
-thick & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 \\ 
+thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 \\
+thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 \\ 
 halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16
@@
 the type of orbit).
+In reality we will have one measurement of the phase-space position to work with, in which case the determination of the orbital phase in $z$ or $R$ is degenerate with the degree of systematic offset in that coordinate (see Figure \ref{fig:cartoon}).
+Additionally, in searching for migrated members of known open clusters, we would compute actions for stars with a wide range of orbital phases and locations in the Galaxy.
+In the following we therefore quote percentile ranges for the possible values computed for each action as a proxy for the effect of these systematic errors in the coordinate system.
+
 For a systematic offset in $z$ ({\em upper} panels), the 95\textsuperscript{th} minus
@@
 artifacts remaining. Only small errors are induced in $J_r$, with the middle
-$90\%$ of values over the $\Gyr$ being $\sim8\%$ of the true $J_r$. As
+$90\%$ of values over the $\Gyr$ being within $\sim8\%$ of the true $J_r$. As
 expected, large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$,
-with the middle $90\%$ of values being $\sim43\%$ of the true $J_z$. The $x$
+with the middle $90\%$ of values being within $\sim43\%$ of the true $J_z$. The $x$
 offset induces uncertainties in $J_r$ and $L_z$ of $\sim21\%$ and $\sim3\%$,
@@
 
-We now repeat the procedure but continuously changing the error in the $z$ and
+We now repeat the same procedure for systematic offsets between 0 and 500 pc in the $z$ and
 $x$ components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth
-minus 5\uth percentile divided by the true action value for three different
-orbits. The thick (\thickcolor) orbit is exactly the same as in
-Figure~\ref{fig:one_orbit_wrong_ref}, and we also consider the thin and halo
-orbits (\thincolor and \halocolor, respectively) from Table~\ref{tab:orbits}.
+minus 5\uth percentile divided by the true action value for the three different fiducial
+orbits in Table \ref{tab:orbits}. The thick-disk (\thickcolor) orbit is the one from
+Figure~\ref{fig:one_orbit_wrong_ref}, but we also now consider the effect on the action determined for the thin-disk (\thincolor) and halo (\halocolor) orbits.
 
-The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the error in each
-action assuming an offset in the $z$ component (i.e. the Galactic midplane).
-In the {\em bottom row} we also consider offsets in the $x$ component (i.e.
+The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced in each
+action for an offset in the $z$ component (i.e. the Galactic midplane).
+In the {\em bottom row} we consider offsets in the $x$ component (i.e.
 the Solar radius). The {\em left}, {\em middle}, and {\em right} columns show
-the relative error in $J_z$, $L_z$, and $J_r$, respectively. We compute the
-relative error as the  95\uth minus 5\uth percentile of the measured action
-($\Delta J_i$) over the course of the first $\Gyr$ of the incorrectly observed
-orbit divided by the true action value as a percentage.
+the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$, respectively. We compute the
+fractional spread $\Delta J_i/J_i$ as the 95\uth minus 5\uth percentile of the action computed using the 
+offset coordinate system over the course of the first $\Gyr$, divided by the true action (computed in the correct coordinate system).
 
-In the {\em upper middle} panel, there is essentially no error in the
+In the {\em upper middle} panel, there is essentially no spread in the
 determination of $L_z$. This is expected since, in principle, $L_z$ only
 depends on the $x$- and $y$-components of the position and velocity of the
-stars\footnote{Though this is not how we compute $L_z$ in practice.}, and is
-thus unaffected by offsets in $z$. This also agrees with the result we found
-in Figure~\ref{fig:one_orbit_wrong_ref}.
+stars\footnote{In practice, however, $L_z$ is computed as part of the torus-fitting method.}, and is
+thus unaffected by offsets in $z$. Indeed, the result we found
+in Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all orbit types.
 
 The {\em upper right} panel shows that the fractional error in $J_z$ is more
-exaggerated for more planar orbits. For the thin orbit, deviations that are
-$\sim30\,\pc$ already give $100\%$ deviations in the actions. The required
-offset for $100\%$ deviation is $\sim225\,\pc$ for the thick orbit. The halo
+exaggerated for more planar (disk-like) orbits. For the thin-disk orbit, a systematic offset of
+$\sim30\,\pc$ in the $z$ coordinate already results in $100\%$ deviations in the actions. The required
+offset for $100\%$ deviation is $\sim225\,\pc$ for the thick-disk orbit. The halo
 orbit is relatively resistant to errors in the midplane, with only $\sim25\%$
@@
 deviations in $J_z$. In the {\em bottom middle} and {\em bottom right} panels
-all three lines overlap.
+all three lines nearly overlap.
 
@@
 
-\section{Local Midplane in FIRE} \label{sec:local_fire}
+%where is Figure 4 referred to? we should include a discussion of it before starting
+
+\section{Azimuthal midplane variations at the Solar Circle: estimates from FIRE} \label{sec:local_fire}
 \subsection{Description of Simulation} \label{ssec:cosmozoom}

### 2019-02-15 paper/ms.tex
one last edit, to a comment at the end of 2.3
Fix line formatting

@@
 \section{Motivation} \label{sec:ref_frame}
-We first demonstrate the significance to action computations, especially for disk-like orbits, of systematic errors in the determination of the Galactic midplane and distance to the Galactic center.
-The consequences we explore
-here may also arise from various other systematic errors. For instance, the
-axisymmetric Galactic potential model used in many works to compute actions may not be a good description of the true potential --- or the parameters used may yield a potential that is systematically incorrect outside an original fitted region.
-In this section, we assume that the Galaxy is perfectly described by our model axisymmetric potential, and simply explore the consequences of systematic errors in the determination of the Galactocentric coordinate system.
+We first demonstrate the significance to action computations, especially for
+disk-like orbits, of systematic errors in the determination of the Galactic
+midplane and distance to the Galactic center. The consequences we explore here
+may also arise from various other systematic errors. For instance, the
+axisymmetric Galactic potential model used in many works to compute actions
+may not be a good description of the true potential --- or the parameters used
+may yield a potential that is systematically incorrect outside an original
+fitted region. In this section, we assume that the Galaxy is perfectly
+described by our model axisymmetric potential, and simply explore the
+consequences of systematic errors in the determination of the Galactocentric
+coordinate system.
 
-\subsection{The effect of midplane determination on orbit and action calculations} \label{ssec:cartoon}
-We present a cartoon in Figure~\ref{fig:cartoon} to show how an inaccurate
-determination of the midplane leads to a phase-dependence of the actions calculated from an orbit specified by a phase-space point and an assumed potential model.
-The y-axis corresponds to the vertical height of the orbit and the x-axis orbital phase.
-The solid gray line indicates the ``true'' orbit of the star; that is, its orbit as it oscillates around the true midplane.
-The dashed gray line is an erroneous midplane assigned to the potential model which an observer uses to integrate the orbit of the star and hence estimate its actions.
-The model potential is otherwise identical to the one in which the star is actually moving.
+\subsection{The effect of midplane determination on orbit and action
+calculations} \label{ssec:cartoon} We present a cartoon in
+Figure~\ref{fig:cartoon} to show how an inaccurate determination of the
+midplane leads to a phase-dependence of the actions calculated from an orbit
+specified by a phase-space point and an assumed potential model. The y-axis
+corresponds to the vertical height of the orbit and the x-axis orbital phase.
+The solid gray line indicates the ``true'' orbit of the star; that is, its
+orbit as it oscillates around the true midplane. The dashed gray line is an
+erroneous midplane assigned to the potential model which an observer uses to
+integrate the orbit of the star and hence estimate its actions. The model
+potential is otherwise identical to the one in which the star is actually
+moving.
 
 Now suppose this observer makes a measurement of the true orbit at the blue
-point or the pink point (i.e.
-at two different orbital phases).
-Then the blue and pink lines correspond to the orbits that the observer would compute for each point based on the potential model with the erroneous midplane.
-In action space, this would correspond to a different value of $J_z$ for the blue and pink points.
-In this way, assuming the wrong coordinate system induces a phase-dependence in the
-actions estimated for the star, which in the correct potential (in this example, the one with the correct midplane) should be phase-independent.
+point or the pink point (i.e. at two different orbital phases). Then the blue
+and pink lines correspond to the orbits that the observer would compute for
+each point based on the potential model with the erroneous midplane. In action
+space, this would correspond to a different value of $J_z$ for the blue and
+pink points. In this way, assuming the wrong coordinate system induces a
+phase-dependence in the actions estimated for the star, which in the correct
+potential (in this example, the one with the correct midplane) should be
+phase-independent.
 
@@
 \plotone{fig/cartoon.pdf}
-\caption{Cartoon showing the effect an error in the determination of the coordinate midplane can have on orbit integration and action estimation. The x-axis shows the
-orbital phase and the y-axis the vertical height. The top gray curve depicts an example ``true'' orbit oscillating about the true midplane (horizontal
-solid gray line). Consider an observer who erroneously assumes the
-midplane is located at the horizontal dashed line. Suppose this
-observer measures the phase-space position of the star at two different orbital phases (blue and pink points).
-If the observer were then to integrate the star's orbit using a model potential with the erroneous midplane, they would
-obtain the blue and pink curves for the star's orbit, respectively.
-The actions estimated from these two  erroneous orbits would subsequently differ, both from each other and from the actions estimated for the true orbit (in the potential with the correct midplane).
-Hence an
-incorrect midplane in the potential model assumed will induce phase-dependence in the actions estimated for a given star in that potential.}
+\caption{Cartoon showing the effect an error in the determination of the
+coordinate midplane can have on orbit integration and action estimation. The
+x-axis shows the orbital phase and the y-axis the vertical height. The top
+gray curve depicts an example ``true'' orbit oscillating about the true
+midplane (horizontal solid gray line). Consider an observer who erroneously
+assumes the midplane is located at the horizontal dashed line. Suppose this
+observer measures the phase-space position of the star at two different
+orbital phases (blue and pink points). If the observer were then to integrate
+the star's orbit using a model potential with the erroneous midplane, they
+would obtain the blue and pink curves for the star's orbit, respectively. The
+actions estimated from these two  erroneous orbits would subsequently differ,
+both from each other and from the actions estimated for the true orbit (in the
+potential with the correct midplane). Hence an incorrect midplane in the
+potential model assumed will induce phase-dependence in the actions estimated
+for a given star in that potential.}
 \label{fig:cartoon}
@@
 \subsection{Numerical methods} \label{ssec:action_comp}
-We compute actions as in our previous work \citep{2018ApJ...867...31B}, using the code \texttt{gala} v0.3 to perform orbit integrations and
-conversion to action space \citep{2017JOSS....2..388P,Price-Whelan:2018}.
-To compute actions we use the torus-mapping technique first presented by \citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to calculate actions for an orbital time-series starting from a phase-space position $(x, v)$ and integrated in a potential $\Phi$.
-We use the default \texttt{MWPotential} as our
-potential, which is based on the Milky Way potential available in
-\texttt{galpy} \citep{2015ApJS..216...29B}. This potential includes a
-Hernquist bulge and nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
-\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
-is fit to empirically match some observations. We use the Dormand-Prince
-8(5,3) integration scheme \citep{Dormand80:integrator} with a timestep of
-$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
-Sun-like star.
+We compute actions as in our previous work \citep{2018ApJ...867...31B}, using
+the code \texttt{gala} v0.3 to perform orbit integrations and conversion to
+action space \citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute actions
+we use the torus-mapping technique first presented by
+\citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to
+calculate actions for an orbital time-series starting from a phase-space
+position $(x, v)$ and integrated in a potential $\Phi$. We use the default
+\texttt{MWPotential} as our potential, which is based on the Milky Way
+potential available in \texttt{galpy} \citep{2015ApJS..216...29B}. This
+potential includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
+a Miyamoto-Nagai disk \citep{1975PASJ...27..533M}, and an NFW halo
+\citep{1997ApJ...490..493N}, and is fit to empirically match some
+observations. We use the Dormand-Prince 8(5,3) integration scheme
+\citep{Dormand80:integrator} with a timestep of $1\,\Myr$ and integrate for
+$5\,\Gyr$, corresponding to $\sim 20$ orbits for a Sun-like star.
 
@@
 the St\"ackel Fudge method \citep{2016MNRAS.457.2107S}, which uses a single
-St\"ackel potential (with analytic actions) to approximate the Galactic potential
-\citep{1985MNRAS.216..273D,2012MNRAS.426.1324B}, was used in many recent works exploring actions in the Galactic disk
-\citep[e.g.][]{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}. For
-disk-like orbits, existing implementations of the St\"ackel Fudge method are of acceptable
-accuracy, but since we also consider halo-like orbits in this work we choose to use
-orbit integration and torus mapping throughout \citep{2016MNRAS.457.2107S}.
+St\"ackel potential (with analytic actions) to approximate the Galactic
+potential \citep{1985MNRAS.216..273D,2012MNRAS.426.1324B}, was used in many
+recent works exploring actions in the Galactic disk
+\citep[e.g.][]{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}.
+For disk-like orbits, existing implementations of the St\"ackel Fudge method
+are of acceptable accuracy, but since we also consider halo-like orbits in
+this work we choose to use orbit integration and torus mapping throughout
+\citep{2016MNRAS.457.2107S}.
 
 \subsection{Quantification of the effect of systematic coordinate-system errors} \label{ssec:quant}
-First, we illustrate the phase-dependence of an example systematic error in the Galactocentric coordinate system.
-We will consider three orbits in the model potential described in \S \ref{ssec:action_comp} that are typical of stars in the
-stellar thin disk, stellar thick disk, and the stellar halo. We summarize
-their initial positions in phase space and the actions computed by integrating their
-orbits in the correct potential in Table~\ref{tab:orbits}. Each orbit, integrated without systematic coordinate errors, is plotted in
-Appendix~\ref{app:orbits}. We will refer to these orbits by their names (thin-disk, thick-disk, halo)
-henceforth.
+First, we illustrate the phase-dependence of an example systematic error in
+the Galactocentric coordinate system. We will consider three orbits in the
+model potential described in \S \ref{ssec:action_comp} that are typical of
+stars in the stellar thin disk, stellar thick disk, and the stellar halo. We
+summarize their initial positions in phase space and the actions computed by
+integrating their orbits in the correct potential in Table~\ref{tab:orbits}.
+Each orbit, integrated without systematic coordinate errors, is plotted in
+Appendix~\ref{app:orbits}. We will refer to these orbits by their names
+(thin-disk, thick-disk, halo) henceforth.
 
-We begin by considering the thick-disk orbit.
-Suppose that we could observe the thick-disk star's phase-space position at many different times (and hence different orbital phases), but used a coordinate system in which the midplane is systematically offset in height by $100\,\pc$ from its actual location --- i.e.
-we subtract the vector $(0, 0, 100)\,\pc$ from each
-position in the orbit. This corresponds to an observer physically located at
-the position $(8, 0, 0)\,\kpc$ in the coordinate system of the true potential, but erroneously thinking they are located at
-$(8, 0, 0.1)\,\kpc$.
+We begin by considering the thick-disk orbit. Suppose that we could observe
+the thick-disk star's phase-space position at many different times (and hence
+different orbital phases), but used a coordinate system in which the midplane
+is systematically offset in height by $100\,\pc$ from its actual location ---
+i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position in the
+orbit. This corresponds to an observer physically located at the position $(8,
+0, 0)\,\kpc$ in the coordinate system of the true potential, but erroneously
+thinking they are located at $(8, 0, 0.1)\,\kpc$.
 
@@
 
-Starting from each observation of the star's phase-space position that such an (extremely long-lived)
-observer makes every $\Myr$, we then compute the actions as outlined above using an integration from that starting point in phase space carried out using the true potential (but specifying the star's present-day position using the systematically offset coordinate system).
-The erroneous computed
-actions for each phase-space starting point are shown for the first $\Gyr$ of the orbit in the {\em upper} panels Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical scheme fails
-and very large actions are reported by \texttt{gala}---we perform a $5\sigma$ clip on each action to exclude such orbits, but this only excludes a total of $5$ orbits out of the $1000$ considered for
-    Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but
-    the vast majority of orbits are computed properly.}
+Starting from each observation of the star's phase-space position that such an
+(extremely long-lived) observer makes every $\Myr$, we then compute the
+actions as outlined above using an integration from that starting point in
+phase space carried out using the true potential (but specifying the star's
+present-day position using the systematically offset coordinate system). The
+erroneous computed actions for each phase-space starting point are shown for
+the first $\Gyr$ of the orbit in the {\em upper} panels
+Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
+scheme fails and very large actions are reported by \texttt{gala}---we perform
+a $5\sigma$ clip on each action to exclude such orbits, but this only excludes
+a total of $5$ orbits out of the $1000$ considered for
+Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
+vast majority of orbits are computed properly.}
 
-We also perform the same
-    procedure but assume a $100\,\pc$ offset in the $x$ component (i.e.
-    subtracting the vector $(100, 0, 0)\,\pc$) in the {\em lower} panels. This
-    is equivalent to a systematic offset in the determination of the distance from the Sun to the Galactic center.
+We also perform the same procedure but assume a $100\,\pc$ offset in the $x$
+    component (i.e. subtracting the vector $(100, 0, 0)\,\pc$) in the {\em
+    lower} panels. This is equivalent to a systematic offset in the
+    determination of the distance from the Sun to the Galactic center.
 
-Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the offset coordinate systems depend on the time (i.e.
-orbital phase) at which the star's phase-space position is observed.
-(To our extremely long-lived observer this would appear as time-dependence in the actions, which appear to oscillate around their value using the correct coordinate system even though in this example the Galaxy is described perfectly by a static, axisymmetric potential.) The relative size of the phase variation in each action depends on the direction of the systematic offset as well as the true values of the actions (i.e.
-the type of orbit).
-In reality we will have one measurement of the phase-space position to work with, in which case the determination of the orbital phase in $z$ or $R$ is degenerate with the degree of systematic offset in that coordinate (see Figure \ref{fig:cartoon}).
-Additionally, in searching for migrated members of known open clusters, we would compute actions for stars with a wide range of orbital phases and locations in the Galaxy.
-In the following we therefore quote percentile ranges for the possible values computed for each action as a proxy for the effect of these systematic errors in the coordinate system.
+Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
+offset coordinate systems depend on the time (i.e. orbital phase) at which the
+star's phase-space position is observed. (To our extremely long-lived observer
+this would appear as time-dependence in the actions, which appear to oscillate
+around their value using the correct coordinate system even though in this
+example the Galaxy is described perfectly by a static, axisymmetric
+potential.) The relative size of the phase variation in each action depends on
+the direction of the systematic offset as well as the true values of the
+actions (i.e. the type of orbit). In reality we will have one measurement of
+the phase-space position to work with, in which case the determination of the
+orbital phase in $z$ or $R$ is degenerate with the degree of systematic offset
+in that coordinate (see Figure \ref{fig:cartoon}). Additionally, in searching
+for migrated members of known open clusters, we would compute actions for
+stars with a wide range of orbital phases and locations in the Galaxy. In the
+following we therefore quote percentile ranges for the possible values
+computed for each action as a proxy for the effect of these systematic errors
+in the coordinate system.
 
@@
 
-We now repeat the same procedure for systematic offsets between 0 and 500 pc in the $z$ and
-$x$ components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth
-minus 5\uth percentile divided by the true action value for the three different fiducial
-orbits in Table \ref{tab:orbits}. The thick-disk (\thickcolor) orbit is the one from
-Figure~\ref{fig:one_orbit_wrong_ref}, but we also now consider the effect on the action determined for the thin-disk (\thincolor) and halo (\halocolor) orbits.
+We now repeat the same procedure for systematic offsets between 0 and 500 pc
+in the $z$ and $x$ components. In Figure~\ref{fig:many_orbit_wrong_ref}, we
+report the 95\uth minus 5\uth percentile divided by the true action value for
+the three different fiducial orbits in Table \ref{tab:orbits}. The thick-disk
+(\thickcolor) orbit is the one from Figure~\ref{fig:one_orbit_wrong_ref}, but
+we also now consider the effect on the action determined for the thin-disk
+(\thincolor) and halo (\halocolor) orbits.
 
-The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced in each
-action for an offset in the $z$ component (i.e. the Galactic midplane).
-In the {\em bottom row} we consider offsets in the $x$ component (i.e.
-the Solar radius). The {\em left}, {\em middle}, and {\em right} columns show
-the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$, respectively. We compute the
-fractional spread $\Delta J_i/J_i$ as the 95\uth minus 5\uth percentile of the action computed using the 
-offset coordinate system over the course of the first $\Gyr$, divided by the true action (computed in the correct coordinate system).
+The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
+in each action for an offset in the $z$ component (i.e. the Galactic
+midplane). In the {\em bottom row} we consider offsets in the $x$ component
+(i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
+show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
+respectively. We compute the fractional spread $\Delta J_i/J_i$ as the 95\uth
+minus 5\uth percentile of the action computed using the offset coordinate
+system over the course of the first $\Gyr$, divided by the true action
+(computed in the correct coordinate system).
 
@@
 depends on the $x$- and $y$-components of the position and velocity of the
-stars\footnote{In practice, however, $L_z$ is computed as part of the torus-fitting method.}, and is
-thus unaffected by offsets in $z$. Indeed, the result we found
-in Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all orbit types.
+stars\footnote{In practice, however, $L_z$ is computed as part of the
+torus-fitting method.}, and is thus unaffected by offsets in $z$. Indeed, the
+result we found in Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk
+orbit holds for all orbit types.
 
 The {\em upper right} panel shows that the fractional error in $J_z$ is more
-exaggerated for more planar (disk-like) orbits. For the thin-disk orbit, a systematic offset of
-$\sim30\,\pc$ in the $z$ coordinate already results in $100\%$ deviations in the actions. The required
-offset for $100\%$ deviation is $\sim225\,\pc$ for the thick-disk orbit. The halo
-orbit is relatively resistant to errors in the midplane, with only $\sim25\%$
-error in $J_z$ out to a midplane offset of $500\,\pc$.
+exaggerated for more planar (disk-like) orbits. For the thin-disk orbit, a
+systematic offset of $\sim30\,\pc$ in the $z$ coordinate already results in
+$100\%$ deviations in the actions. The required offset for $100\%$ deviation
+is $\sim225\,\pc$ for the thick-disk orbit. The halo orbit is relatively
+resistant to errors in the midplane, with only $\sim25\%$ error in $J_z$ out
+to a midplane offset of $500\,\pc$.
 
@@
 
-%where is Figure 4 referred to? we should include a discussion of it before starting
+%where is Figure 4 referred to? we should include a discussion of it before starting the next section
 
-\section{Azimuthal midplane variations at the Solar Circle: estimates from FIRE} \label{sec:local_fire}
+\section{Azimuthal midplane variations at the Solar Circle: estimates from
+FIRE} \label{sec:local_fire}
 \subsection{Description of Simulation} \label{ssec:cosmozoom}

### 2019-02-15 paper/ms.tex
Fix minor formatting things

@@
 midplane leads to a phase-dependence of the actions calculated from an orbit
-specified by a phase-space point and an assumed potential model. The y-axis
-corresponds to the vertical height of the orbit and the x-axis orbital phase.
-The solid gray line indicates the ``true'' orbit of the star; that is, its
-orbit as it oscillates around the true midplane. The dashed gray line is an
-erroneous midplane assigned to the potential model which an observer uses to
-integrate the orbit of the star and hence estimate its actions. The model
+specified by a phase-space point and an assumed potential model. The $y$-axis
+corresponds to the vertical height of the orbit and the $x$-axis orbital
+phase. The solid gray line indicates the ``true'' orbit of the star; that is,
+its orbit as it oscillates around the true midplane. The dashed gray line is
+an erroneous midplane assigned to the potential model which an observer uses
+to integrate the orbit of the star and hence estimate its actions. The model
 potential is otherwise identical to the one in which the star is actually
@@
 coordinate midplane can have on orbit integration and action estimation. The
-x-axis shows the orbital phase and the y-axis the vertical height. The top
+$x$-axis shows the orbital phase and the $y$-axis the vertical height. The top
 gray curve depicts an example ``true'' orbit oscillating about the true
@@
 
-\subsection{Quantification of the effect of systematic coordinate-system errors} \label{ssec:quant}
-First, we illustrate the phase-dependence of an example systematic error in
-the Galactocentric coordinate system. We will consider three orbits in the
-model potential described in \S \ref{ssec:action_comp} that are typical of
-stars in the stellar thin disk, stellar thick disk, and the stellar halo. We
-summarize their initial positions in phase space and the actions computed by
-integrating their orbits in the correct potential in Table~\ref{tab:orbits}.
-Each orbit, integrated without systematic coordinate errors, is plotted in
-Appendix~\ref{app:orbits}. We will refer to these orbits by their names
-(thin-disk, thick-disk, halo) henceforth.
+\subsection{Quantification of the effect of systematic coordinate-system
+errors} \label{ssec:quant}
+First, we illustrate the phase-dependence of an
+example systematic error in the Galactocentric coordinate system. We will
+consider three orbits in the model potential described in \S
+\ref{ssec:action_comp} that are typical of stars in the stellar thin disk,
+stellar thick disk, and the stellar halo. We summarize their initial positions
+in phase space and the actions computed by integrating their orbits in the
+correct potential in Table~\ref{tab:orbits}. Each orbit, integrated without
+systematic coordinate errors, is plotted in Appendix~\ref{app:orbits}. We will
+refer to these orbits by their names (thin-disk, thick-disk, halo) henceforth.
 
@@
 
-We now repeat the same procedure for systematic offsets between 0 and 500 pc
-in the $z$ and $x$ components. In Figure~\ref{fig:many_orbit_wrong_ref}, we
-report the 95\uth minus 5\uth percentile divided by the true action value for
-the three different fiducial orbits in Table \ref{tab:orbits}. The thick-disk
-(\thickcolor) orbit is the one from Figure~\ref{fig:one_orbit_wrong_ref}, but
-we also now consider the effect on the action determined for the thin-disk
-(\thincolor) and halo (\halocolor) orbits.
+We now repeat the same procedure for systematic offsets between $0$ and
+$500\,\pc$ in the $z$ and $x$ components. In
+Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth minus 5\uth
+percentile divided by the true action value for the three different fiducial
+orbits in Table \ref{tab:orbits}. The thick-disk (\thickcolor) orbit is the
+one from Figure~\ref{fig:one_orbit_wrong_ref}, but we also now consider the
+effect on the action determined for the thin-disk (\thincolor) and halo
+(\halocolor) orbits.
 

### 2019-02-15 paper/ms.tex
Add text introducing Fig 4

@@
 \plotone{fig/schmactions_Jz_hist.pdf}
-\caption{A histogram of the observed values in $J_z$ for the thick orbit
+\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
 assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
@@
 \end{figure}
+
+In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
+$J_z$ for the thick-disk orbit assuming a $z$ offset of $100\,\pc$ (i.e. the
+{\em right} panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of
+$J_z$ is plotted as a vertical dashed line. Here we see that the spread in
+$J_z$ induced by a systematic offset in $z$ is non-Gaussian and bimodal.
+Furthermore, neither of the modes lie at the true value of $J_z$.
 

### 2019-02-15 paper/ms.tex
Small changes

@@
 \subsection{Quantification of the effect of systematic coordinate-system
-errors} \label{ssec:quant}
-First, we illustrate the phase-dependence of an
+errors} \label{ssec:quant} First, we illustrate the phase-dependence of an
 example systematic error in the Galactocentric coordinate system. We will
-consider three orbits in the model potential described in \S
-\ref{ssec:action_comp} that are typical of stars in the stellar thin disk,
-stellar thick disk, and the stellar halo. We summarize their initial positions
-in phase space and the actions computed by integrating their orbits in the
-correct potential in Table~\ref{tab:orbits}. Each orbit, integrated without
-systematic coordinate errors, is plotted in Appendix~\ref{app:orbits}. We will
-refer to these orbits by their names (thin-disk, thick-disk, halo) henceforth.
+consider three orbits in the model potential described in
+Section~\ref{ssec:action_comp} that are typical of stars in the stellar thin
+disk, stellar thick disk, and the stellar halo. We summarize their initial
+positions in phase space and the actions computed by integrating their orbits
+in the correct potential in Table~\ref{tab:orbits}. Each orbit, integrated
+without systematic coordinate errors, is plotted in Appendix~\ref{app:orbits}.
+We will refer to these orbits by their names (thin-disk, thick-disk, halo)
+henceforth.
 
@@
 dispersion of a disrupted cluster. This dispersion can be estimated as
-\citep[\S~8.3.3][]{2008gady.book.....B}
+\citep[Section~8.3.3][]{2008gady.book.....B}
 \beq \label{eq:action_disp}

### 2019-02-15 paper/ms.tex
Various small fixes

@@
 midplane is insufficient for extracting the most dynamical information from
-the {\em Gaia} mission. From now on, we set aside the discussion of
-inaccuracies in the Solar radius. While all that we state about the midplane
-applies to the Solar radius as well, it is not clear the best way to determine
-the dynamical center of the galaxy in the FIRE simulations.
+the {\em Gaia} mission. We focus our discussion now solely on the midplane.
 
@@
 Under the assumption that this interpretation is correct, we now consider the
-implications for dynamical studies of disrupting open clusters (the original
-motivation of this study).
+implications for attempts to reconstruct disrupting open clusters by selecting
+ejected members based on their actions.
 
 Suppose an open cluster of mass $m_c$ on an orbit with actions $\bm{J}_c$ is
-disrupted, and that the cluster is currently at an orbit phase of
-$\bm{\theta}_0$. We would like to determine all stars in the {\em Gaia}
+disrupted. We would like to determine all stars in the {\em Gaia}
 dataset\footnote{This is likely to be more successful with DR3/4 than DR2,
@@
 allow this program to be performed sooner.} that are likely to be ejected
-members from that cluster. One could proceed by integrating the cluster's
-orbit and selecting all stars within a certain distance of that orbit,
-accounting for the fact that the stream will not exactly follow the orbit of
-the cluster \citep[e.g.][]{2011MNRAS.413.1852E}.
+members from that cluster. One could generate an initial catalogue of
+candidate members by integrating the cluster's orbit and selecting stars close
+to that orbit, accounting for the fact that the stream will not exactly follow
+the orbit of the cluster \citep[e.g.][]{2011MNRAS.413.1852E}.
 
@@
 
-For the thick orbit (brown), the neighborhood is at the maximum value for
+For the thick-disk orbit (brown), the neighborhood is at the maximum value for
 nearly all masses, except at the very low-mass end in m12i.

### 2019-02-15 paper/ms.tex
revisions to section 3,3.1

--- 2019-02-15 paper/ms.tex
@@
 
-\newcommand{\Msun}{\text{M}_\odot}
+\newcommand{\Msun}{\ensuremath{\text{M}_\odot}}
 \newcommand{\pc}{\text{pc}}
@@
 \newcommand{\actunit}{\text{kpc}\,\kms}
+
+\newcommand{\unit}[2]{\ensuremath{\textrm{#1}^{\mathrm{#2}}}}
+\newcommand{\m12i}{\textsf{m12i}}
+\newcommand{\m12f}{\textsf{m12f}}
+\newcommand{\m12m}{\textsf{m12m}}
 
@@
 errors} \label{ssec:quant}
-First, we illustrate the phase-dependence of an
+First, we illustrate the phase-dependence induced by an
 example systematic error in the Galactocentric coordinate system. We will
@@
 \subsection{Description of Simulation} \label{ssec:cosmozoom}
-Here we briefly describe the cosmological zoom-ins used in this work.
-Cosmological zoom-ins allow one to simulate a selected region at high
-resolution embedded in a low-resolution cosmological background
-\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}. The FIRE
-collaboration has simulated a number of Milky Way-mass zoom-ins as part of the
-{\em Latte} suite of FIRE-2 simulations
-\citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. We use the three zoom-ins
-considered in \citet{2018arXiv180610564S} --- m12i, m12f, m12m, since they
-have associated mock {\em Gaia} DR2 catalogues and the $z=0$ snapshots are
-publicly available.
+Here we briefly describe the cosmological-hydrodynamical zoom-in simulations used in this work.
+Cosmological zoom-ins allow one to simulate a selected region at high resolution embedded in a low-resolution cosmological background \citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added a hydrodynamical simulation of the baryonic component with recipes for star formation and feedback.
+The FIRE collaboration has simulated a number of Milky Way-mass galaxies using this technique as part of the
+{\em Latte} suite of FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. 
+For this work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S}, which show broad agreement of many of their global properties with observations of the Milky Way.
+The $\z=0$ snapshots\footnote{In this work, to avoid confusion with the vertical height $z$, we refer to cosmological redshift as $\z$.} of these three simulations, called m12i, m12f, and m12m for shorthand, are publicly available alongside associated mock {\em Gaia} DR2 catalogues generated from them.\footnote{available at \url{http://ananke.hub.yt}}
 
-First, a dark matter only simulation was run with $\Lambda$CDM cosmology
-parameterized by $\Omega_m = 0.272$, $\Omega_{\Lambda} = 0.728$, $\Omega_b =
-0.0455$, $h = 0.702$, $\sigma_8 = 0.807$, and $n_s = 0.961$, consistent with
-current constraints \citep{2018arXiv180706209P}. Halos are selected at $\z=0$
-based solely on their mass ($\sim 1\text{--}2 \times 10^{12} \Msun$). In this
-work, to avoid confusion with the vertical height $z$, we refer to
-cosmological redshift as $\z$. Particles within $5 R_{200\text{m}}$ of each
-halo are traced back to $\z=99$ and the initial conditions are regenerated at
-higher resolution using MUSIC \citep{2011MNRAS.415.2101H}.
+%First, a dark matter only simulation was run with $\Lambda$CDM cosmology parameterized by $\Omega_m = 0.272$, $\Omega_{\Lambda} = 0.728$, $\Omega_b = 0.0455$, $h = 0.702$, $\sigma_8 = 0.807$, and $n_s = 0.961$, consistent with current constraints \citep{2018arXiv180706209P}.
+Halos are selected at $\z=0$ based solely on their mass ($\sim 1\text{--}2 \times 10^{12} \Msun$).
+Particles within $5 R_{200\text{m}}$ of each halo are traced back to $\z=99$ and the initial conditions are regenerated at higher resolution using MUSIC \citep{2011MNRAS.415.2101H}.
 
-Simulations were run using
-\texttt{GIZMO}\footnote{\url{http://www.tapir.caltech.edu/~phopkins/Site/GIZMO.html}}
-\citep{2015MNRAS.450...53H}. Hydrodynamics is solved using the meshless
-finite-mass method and gravity is solved using a modified version of the
-Tree-PM solver of \texttt{GADGET-3}, using fully adaptive and fully
-conservative gravitational force softening for gas.
+%Simulations were run using \texttt{GIZMO}\footnote{\url{http://www.tapir.caltech.edu/~phopkins/Site/GIZMO.html}} \citep{2015MNRAS.450...53H}.
+Hydrodynamics is solved using the meshless finite-mass method and gravity is solved using a modified version of the Tree-PM solver of \texttt{GADGET-3}, using fully adaptive and fully conservative gravitational force softening for gas.
 
-These halos are uncontaminated and contain gas particles of mass $\sim 7000
-\text{--} 20,000\,\Msun$ and star particles of mass $\sim 5000 \text{--} 7000\,
-\Msun$\footnote{In fact, the simulation m12i contained a gas
+These simulations contain dark matter particles of mass $\sim$35,000 \Msun, gas particles of mass $\sim 7000
+\text{--} 20,000\,\Msun$, and star particles of mass $\sim 5000 \text{--} 7000\,
+\Msun$,\footnote{In fact, the simulation m12i contained a gas
 splitting bug which causes a small number of gas and star particles to have
 larger than typical masses. These constitute only $\sim0.2\%$ of the star
-particles in the $z=0$ snapshots, and are only a factor of a few more massive
+particles in the $\z=0$ snapshots, and are only a factor of a few more massive
 than typical star particles. Since our results are broadly consistent between
-the three simulations, we ignore this minor complication.}, with the lower end
+the three simulations, we ignore this minor complication.} with the lower end
 coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
@@
 
-In this work, we take the coordinate systems used in
-\citet{2018arXiv180610564S} as our fiducial coordinate systems for each
-galaxy. The center of the galaxy is found through an iterative ``shrinking
-spheres'' method, where the center is calculated for a given radius which is
-then reduced by $50\%$ until converging on the galaxy's center. The center of
+The softening lengths used in the simulations can affect the ability to resolve the very thinnest planar structures, which in turn can affect how much the density-based midplane varies as a function of azimuth.
+The Milky Way's gas disk is thought to have a scale height of {\bf XXX} pc (XX times the median gas softening length) and the thin stellar disk a scale height of {\bf XXX} pc (XX times the stellar softening length).
+We therefore expect that resolution effects may still be affecting the scale heights of these components in the simulations, and indeed the gas and stellar scale heights of the simulated galaxies are all somewhat larger than the MW's.
+However, in this work we are considering azimuthal variations in the midplane \emph{as defined by the mass density}, which will subsequently be lower in the simulations than in the MW: within 200 pc of the Sun the measured gas and stellar volume densities in the MW are roughly comparable at $\sim .04\ \Msun\ \unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see Table 3 of \citealt{2018arXiv180610564S}).
+We thus expect that the degree of midplane variation we find in the simulated systems is likely to be \emph{less} than in the real MW, since on the smallest scales the simulations are \emph{smoother} than the real Galaxy.
+
+Cosmological simulations of Milky Way-mass galaxies are
+not perfect representations of the true Milky Way in other ways as well, as discussed in
+\citet{2018arXiv180610564S}. For instance, the velocity structure of m12i is
+closer to M31's than the Milky Way's (Loebman et al. in prep). However, in this work we are most interested
+in the global properties of the potential, and specifically in deviations
+from axisymmetry. From this perspective, the simulated galaxies are actually \emph{more}
+axisymmetric than we might expect of the Milky Way. While they have prominent
+spiral arms, none has as strong a bar as the MW does at present day, and none has a nearby companion like the Large Magellanic Cloud.
+One of the three we consider (\m12f) does have an ongoing interaction with a satellite galaxy similar to Sagittarius, which has punched through the galactic disk outside the solar circle and induced some warping.
+However, we expect that, as with the use of finite smoothing lengths, the degree of midplane variation in the simulations is likely a factor of a few lower than what could be expected in the MW.
+
+In this work, we take the galactocentric coordinate system described in Section 3 of
+\citet{2018arXiv180610564S} as our fiducial coordinate system for each
+galaxy. In short, the center of the galaxy is found iteratively, the center of
 mass velocity is then determined by all star particles within $15\,\kpc$ of
-this center. The galaxy is then rotated onto the principal axis frame as
+this center, and the galaxy is then rotated onto a principal-axis frame
 determined by stars younger than $1\,\Gyr$ inside of the fiducial solar radius
-$R_{\odot} = 8.2\,\kpc$.
-
-It is important to remember that simulations of Milky Way-mass galaxies are
-not perfect representations of the true Milky Way, as discussed in
-\citet{2018arXiv180610564S}. For instance, the velocity structure of m12i is
-closer to M31 than the Milky Way. However, in this work we are most interested
-in the global properties of the m12i potential, and specifically in deviations
-from axisymmetry. From this perspective, m12i is actually far more
-axisymmetric than we might expect of the Milky Way. While m12i has prominent
-spiral arms, it lacks a bar and a companion like the Large Magellanic Cloud.
-In future work, these additional complications also need to be addressed.
+$R_{\odot} = 8.2\,\kpc$, such that the disk plane is the $x$--$y$ plane.
 
 \subsection{Local Midplane} \label{ssec:local_midplane}
-We now turn to defining the local midplane that an observer might measure if
+Using the three simulations, we determine the \emph{local} midplane that an observer might measure if
 they were situated in each of these galaxies. We place our imaginary observer
-at the solar radius of $8.2\,\kpc$ and azimuth $\phi$\footnote{Our erroneous,
-immortal observer is also warp-drive capable.}, and consider stars within a
+at the solar radius of $8.2\,\kpc$ and azimuth $0<\phi<2\pi$\footnote{Our 
+nearly immortal observer also has warp-drive.}, and consider stars within a
 cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to the

### 2019-02-15 paper/ms.tex
Various small fixes to Discussion

--- 2019-02-15 paper/ms.tex
@@
 have to account for the different midplane of the field star and the cluster.
-We repeat this analysis for each of the three orbits considered in
-Section~\ref{ssec:quant}.
+We perform this analysis for the thin-disk and thick-disk orbits considered in
+Section~\ref{ssec:quant}, since the halo orbit is unrealistic for known open
+clusters (Table~\ref{tab:real_clusters}).
 
@@
 Figure~\ref{fig:range_deltaphi} to convert this $z$ offset into a $\Delta
-\phi$, which we convert to an arclength or characteristic distance (the
-difference between the two will be negligible for the small $\Delta \phi$'s we
-will compute). Therefore, if an observer is attempting to reconstruct a
-cluster with a given mass on a given orbit, we will give a characteristic
-distance between a field star and the cluster within which an observer might
-not need to consider differences in their respective local midplanes.
+\phi$, which we convert to an arclength or characteristic distance, which we
+refer to as the ``neighborhood radius'' or $R_{\n}$. Therefore, if an observer
+is attempting to reconstruct a cluster with a given mass on a given orbit, we
+will give a characteristic distance between a field star and the cluster
+within which an observer might not need to consider differences in their
+respective local midplanes.
 
@@
 We now convert the $z$~offset-$m_c$ relation to an $R_{\n}$-$m_c$ relation,
-where $R_{\n}$ indicates a sort of ``solar neighborhood'' radius. An observer
+where $R_{\n}$ indicates the neighborhood radius. An observer
 attempting to reconstruct an open cluster of mass $m_c$ with perfect knowledge
@@
 are at a distance greater than $R_{\n}$, a more sophisticated analysis would
-be required, the details of which are not apparent to the authors of this
-work.
+be required.
 
 We use the relation in Figure~\ref{fig:range_deltaphi}, which shows the
-typical $z$ offset at each $\Delta \phi$, to convert a given $z$ offset into a
-$\Delta\phi$, which we then convert into a chord length which we interpret as
-$R_{\n}$, the neighborhood around the cluster. If the $z$~offset is greater
-than the maximum range value of the galaxy, we simply report the maximum chord
-length ($2\times8.2\,\kpc$). The result of this procedure is shown in
+typical $z$ offset at each $\Delta \phi$, to convert a given $z$~offset into a
+typical $\Delta\phi$. We then convert this $\Delta \phi$ into a chord length
+(see Figure~\ref{fig:fig_to_explain}) which we interpret as $R_{\n}$, the
+neighborhood around the cluster. If the $z$~offset is greater than the maximum
+range value of the galaxy, we simply report the maximum chord length
+($2\times8.2\,\kpc$). The result of this procedure is shown in
 Figure~\ref{fig:Rn_mc}. The solid lines indicate the neighborhood values
@@
 \acknowledgments
-We would like to thank Megan Bedell, David W. Hogg, and Kathryn Johnston for
-many useful discussions. A.B. would like to thank Todd Phillips for helpful
-discussions. A.B. was supported in part by the Roy \& Diana Vagelos Program in
-the Molecular Life Sciences and the Roy \& Diana Vagelos Challenge Award. The
-work of \Gus{everyone} is supported by the Simons Foundation.
+We would like to thank Megan Bedell, Tobias Buck, David W. Hogg, and Kathryn
+Johnston for many useful discussions. A.B. would like to thank Todd Phillips
+for helpful discussions. A.B. was supported in part by the Roy \& Diana
+Vagelos Program in the Molecular Life Sciences and the Roy \& Diana Vagelos
+Challenge Award. The work of \Gus{everyone} is supported by the Simons
+Foundation. \Gus{remove last sentence if everyone CCA affiliated}
 

### 2019-02-15 paper/ms.tex
Rewrite abstract

@@
 
-We demonstrate the effect of midplane errors on the computation of stellar
-actions. This error is induced by the fact that the local Galactic midplane
-varies across the disk. As a result, when we make 6D phase space measurements
-of stars that are non-local (i.e. far away), we are observing those stars with
-a systematic offset in $z$. We show that this systematic offset induces time
-evolution in the observed actions which we interpret as an error. We note that
-this error is non-Gaussian and bimodal, with neither of the modes peaking on
-the true action value. A midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is
-necessary to induce a $100\,\%$ error in the vertical action $J_z$ for thin
-and thick disk orbits, respectively. Using cosmological zoom-in simulations
-from the FIRE collaboration, we show that the midplane can vary by
-$\sim100\,\pc$ around the Solar circle, indicating that such an effect is
-important for understanding the thin and thick disk. Our discussion is also
-directly applicable to variations in the local standard of rest. The variation
-of the midplane must be taken into account when performing dynamical modelling
-across large regions of the disk accessible with the {\em Gaia} mission.
+Stellar actions, computed for stars from their 6D phase space measurement and
+an assumed Galactic potential, are used to label orbits. In principle, actions
+have the attractive quality of being invariant and are thus good labels of a
+star's orbit. Of course, inaccurate measurements of the 6D phase space
+position and uncertainities in the assumed Galactic potential induce an error
+in the computed action. We show that, in addition to these complications, a
+systematic bias in the 6D phase space position induces a phase-dependence in
+the actions, which we interpret as an error. This error is non-Gaussian and
+bimodal, with neither of the modes peaking on the true action value. A
+midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is necessary to induce a
+$100\,\%$ error in the vertical action $J_z$ for thin and thick disk orbits,
+respectively. Furthermore, we show that the local midplane varies by
+$\sim100\,\pc$ at the Solar circle in cosmological zoom-in simulations from
+the FIRE collaboration. Thus, current state-of-the-art action calculations ---
+which assume the global and local midplanes are the same --- include a
+systematic offset in $z$. Variation in the local standard of rest will induce
+similar issues. The variation of the midplane must be taken into account when
+performing dynamical modelling across large regions of the disk accessible
+with the {\em Gaia} mission.
 

### 2019-02-15 paper/ms.tex
Some edits from melissa

@@
 
-Underlying all of these discoveries is the assumption of a Galactocentric
-coordinate system \citep{2008gady.book.....B}. In angular
-coordinates, the center of the galaxy is taken to be the location of
+These discoveries are all indicative that the Milky Way departs from
+axisymmetry and is undergoing phase mixing and dynamical interactions across a
+range of spatial and temporal scales. Underlying the quantitative description
+of the mechanisms at work that give rise to these signatures is the assumption
+of a Galactocentric coordinate system \citep{2008gady.book.....B} which can
+both precisely and accurately be defined and measured. In angular coordinates,
+the center of the galaxy is taken to be the location of
 Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From stellar
@@
 nine open clusters within $250\,\pc$ as reported by the {\em Gaia}
-collaboration \citep{2018AA...616A..10G}. In Table~\ref{tab:real_clusters},
-we report the distance to each cluster, the three actions, and the maximum
+collaboration \citep{2018AA...616A..10G}. In Table~\ref{tab:real_clusters}, we
+report the distance to each cluster, the three actions, and the maximum
 vertical extent of their orbits ($z_{\text{max}}$). The actions are computed
-using the technique described in Section~\ref{ssec:action_comp}.
+using the technique described in Section~\ref{ssec:action_comp}. We also show
+the mass of each cluster and the ``neighborhood'' we compute for each cluster,
+which gives a rough sense of how far one can move away from the cluster before
+the issue which we discuss next becomes important. This neighborhood is
+discussed in more detail in Section~\ref{sssec:neighborhood}.
 
@@
 
-\subsection{Solar Neighborhood} \label{ssec:neighborhood}
+\subsection{Impact on Milky Way observations} \label{ssec:impact_on_mw}
+\subsubsection{The Solar neighborhood} \label{sssec:neighborhood{}}
 To gain further handle on the midplane behavior as a function of azimuth, we
@@
 \end{figure*}
+
+\subsubsection{Consequences for disrupted cluster reconstruction}
+\label{sssec:reconstruction}
 

### 2019-02-15 paper/ms.tex
Add paragraph to conclusions

@@
 
-\begin{deluxetable*}{cccccccc}
+\begin{deluxetable*}{cCCCCCCC}
 \tablecaption{Relevant dynamical parameters for the $9$ open clusters within
@@
 \section{Conclusions}\label{sec:conclusion}
-Let's conclude, shall we.
+Actions have promise as excellent orbit labels. If the Galaxy can be
+approximated as axisymmetric and 6D phase space positions can be measured
+accurately and precisely enough, then the computed actions are invariant with
+orbital phase. However, we hvae shown that the fact that the Galactic midplane
+is not constant across the disk presents a significant complication to
+computed actions actually being invariant. Our main conclusions are as
+follows:
 

### 2019-02-15 paper/ms.tex
Add paragraph explaining neighborhood column

@@
 \subsection{Impact on Milky Way observations} \label{ssec:impact_on_mw}
-\subsubsection{The Solar neighborhood} \label{sssec:neighborhood{}}
+\subsubsection{The Solar neighborhood} \label{sssec:neighborhood}
 To gain further handle on the midplane behavior as a function of azimuth, we
@@
 
+We repeat this entire procedure for each of the orbits of the real open
+clusters in Table~\ref{tab:real_clusters} using the midplane variations from
+m12i. We use m12i as a conservative choice, since it provides the greatest
+midplane variation (Figure~\ref{fig:midplane}). We report the neighborhood
+values for the mass of each cluster in the neighborhood column of
+Table~\ref{tab:real_clusters}. Except for alphaPer, we find that the
+neigbhorhood around each cluster is typically $\gtrsim 500\,\pc$. This
+analysis assumes that the midplane from m12i can be accurately applied to
+Milky Way orbits, which is not necessarily true. However, our analysis is
+meant to be more illustrative than quantitative and these neighborhood numbers
+should be understood as such.
+
 \begin{figure*}

### 2019-02-15 paper/ms.tex
Add additional complications to midplane meas.

@@
 
-It is important for us to emphasize that when we state our local Galactic
-midplane may be erroneous, we do not mean that past measurements/estimates of
-it are wrong. Rather, what we mean is that it may vary from the local midplane
-at other points in the Galaxy. In the face of {\em Gaia} DR2 (and soon DR3/4),
-this may be an important effect to consider when performing dynamical
-modelling across large distances.
+Our main point is that the local midplane varies between different points in
+the Galaxy. This is a different statement than that our measurements of the
+local midplane are erroneous. Indeed, the local midplane has recently been
+measured to a claimed high level of precision, setting $z_{\odot} = 20.8 \pm
+0.3\,\pc$ \citep{2019MNRAS.482.1417B}. However, we also do not mean to say
+that determinations of the local midplane are perfect, since midplane
+measurements from H~\textsc{ii} regions place $z_{\odot} \sim 5-10\,\pc$. An
+additional complication is the fact that non-zero Solar heights induce a tilt
+in the Galactic midplane \citep{2014ApJ...797...53G,2016ARAA..54..529B}.
 

### 2019-02-15 paper/ms.tex
Small fixes

@@
 the issue which we discuss next becomes important. This neighborhood is
-discussed in more detail in Section~\ref{sssec:neighborhood}.
+discussed in more detail in Section~\ref{sssec:reconstruction}.
 
@@
 \tablecaption{Relevant dynamical parameters for the $9$ open clusters within
-$250\,\pc$ of the Sun, using the relevant initial positions and velocities from \citet{2018AA...616A..10G}.\label{tab:real_clusters}}
+$250\,\pc$ of the Sun, using the relevant initial positions and velocities from \citet{2018AA...616A..10G}.
+The neighborhood column is explained in more detail in Section~\ref{sssec:reconstruction}.\label{tab:real_clusters}}
 \tablehead{\colhead{cluster} & \colhead{distance} & \colhead{$J_r$} &

### 2019-02-17 paper/ms.tex
Update citations in first paragraph

@@
 era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
-major merger \citep{2018Natur.563...85H}, a phase-space spiral (\Gus{cites})
-indicating local substructure infall (\Gus{cites}), a gap indicative of a dark
-matter subtructure in GD1 \citep{2018arXiv181103631B}, and \Gus{others}.
+major merger
+\citep{2018ApJ...860L..11K,2018Natur.563...85H,2018arXiv180704290L,2019MNRAS.482.3426M},
+a phase-space spiral \citep{2018Natur.561..360A} indicating either local
+substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L} or bar
+buckling \citep{2018arXiv181109205K}, a gap indicative of a dark matter
+subtructure in GD1 \citep{2018arXiv181103631B}, among others.
 

### 2019-02-17 paper/ms.tex
Add citation to dynamical center

@@
 midplane is insufficient for extracting the most dynamical information from
-the {\em Gaia} mission. We focus our discussion now solely on the midplane.
+the {\em Gaia} mission.
+
+While we focus our discussion now solely on the midplane, there is an
+additional complication in the measurement of the distance from the Sun to the
+Galactic center that may also be important. When modelling stars at Galactic
+radii close to the Sun's radius, it is in principle possible to determine what
+point they are orbitting about, or in other words the ``dynamic Galactic
+center.'' It is not guaranteed that this dynamic Galactic center will coincide
+with the distance to Sgr~A\textsuperscript{*} \citep{2018AA...615L..15G}. In
+fact, \citet{2015ApJ...803...80K} were able to measure the distance of the Sun
+to the Galactic center using dynamical information on Palomar 5, though not to
+a precision capable of competing with the distance to
+Sgr~A\textsuperscript{*}.
 

### 2019-02-17 paper/ms.tex
Add citation to Bland-Hawthorn
Add more references to dynamical R0 measurements

@@
 $\sim 25\,\pc$ \citep{2001ApJ...553..184C}, with more recent measurements from
-{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}.
+{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}. A
+pre-{\em Gaia} review of these measurements is given by
+\citet{2016ARAA..54..529B}.
 
@@
 fact, \citet{2015ApJ...803...80K} were able to measure the distance of the Sun
-to the Galactic center using dynamical information on Palomar 5, though not to
-a precision capable of competing with the distance to
-Sgr~A\textsuperscript{*}.
+to the Galactic center using the dynamical information of Palomar 5. Other
+dynamical measurements of the Solar distance have been made, though none with
+a precision capable of competing with the distance to Sgr~A\textsuperscript{*}
+\citep{1981gask.book.....M,2011PASJ...63..867S,2012MNRAS.427..274S,2013AstL...39...95B,2013IAUS..289..444Z}.
 

### 2019-02-17 paper/ms.tex
Minor fix to thick-disk label

@@
 \caption{The artificial phase-dependence in the observed actions induced by an
-error in the Galactocentric coordinate system. We consider here the thick disk
+error in the Galactocentric coordinate system. We consider here the thick-disk
 orbit, which has actions of $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$

### 2019-02-20 paper/ms.tex

@@
 \subsection{Quantification of the effect of systematic coordinate-system
-errors} \label{ssec:quant} First, we illustrate the phase-dependence of an
+errors} \label{ssec:quant}
+First, we illustrate the phase-dependence induced by an
 example systematic error in the Galactocentric coordinate system. We will

### 2019-02-20 paper/ms.tex
Fix line breaks

@@
 \tablecaption{Relevant dynamical parameters for the $9$ open clusters within
-$250\,\pc$ of the Sun, using the relevant initial positions and velocities from \citet{2018AA...616A..10G}.
-The neighborhood column is explained in more detail in Section~\ref{sssec:reconstruction}.\label{tab:real_clusters}}
+$250\,\pc$ of the Sun, using the relevant initial positions and velocities
+from \citet{2018AA...616A..10G}. The neighborhood column is explained in more
+detail in Section~\ref{sssec:reconstruction}.\label{tab:real_clusters}}
 \tablehead{\colhead{cluster} & \colhead{distance} & \colhead{$J_r$} &
@@
 
-For a systematic offset in $z$ ({\em upper} panels), the 95\textsuperscript{th} minus
-5\textsuperscript{th} percentiles are $2.2$ and $6.2 \actunit$ for $J_r$ and
-$J_z$, respectively. These are fractional errors of $5.7\%$ and $85.7\%$,
-respectively. The error induced in $L_z$ is negligible, as expected since
-$L_z$ is independent of $z$. It is worth pointing out that a
-$100\,\pc$ error in an orbit with $z_{\text{max}}=850\,\pc$ --- a $12\%$ error
+For a systematic offset in $z$ ({\em upper} panels), the
+95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
+$6.2 \actunit$ for $J_r$ and $J_z$, respectively. These are fractional errors
+of $5.7\%$ and $85.7\%$, respectively. The error induced in $L_z$ is
+negligible, as expected since $L_z$ is independent of $z$. It is worth
+pointing out that a $100\,\pc$ error in an orbit with
+$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
 --- induced a $43\%$ error in the computation of $J_z$.
@@
 For a systematic offset in $x$ (or Galactic center distance), the
-95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$, $47$,
-and $0.71 \actunit$ for $J_r$, $L_z$ and $J_z$, respectively. These are
+95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
+$47$, and $0.71 \actunit$ for $J_r$, $L_z$ and $J_z$, respectively. These are
 fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively.
@@
 expected, large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$,
-with the middle $90\%$ of values being within $\sim43\%$ of the true $J_z$. The $x$
-offset induces uncertainties in $J_r$ and $L_z$ of $\sim21\%$ and $\sim3\%$,
-respectively. A $\sim3\%$ error in $J_z$ is also induced.}
+with the middle $90\%$ of values being within $\sim43\%$ of the true $J_z$.
+The $x$ offset induces uncertainties in $J_r$ and $L_z$ of $\sim21\%$ and
+$\sim3\%$, respectively. A $\sim3\%$ error in $J_z$ is also induced.}
 \label{fig:one_orbit_wrong_ref}
@@
 \subsection{Description of Simulation} \label{ssec:cosmozoom}
-Here we briefly describe the cosmological-hydrodynamical zoom-in simulations used in this work.
-Cosmological zoom-ins allow one to simulate a selected region at high resolution embedded in a low-resolution cosmological background \citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added a hydrodynamical simulation of the baryonic component with recipes for star formation and feedback.
-The FIRE collaboration has simulated a number of Milky Way-mass galaxies using this technique as part of the
-{\em Latte} suite of FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. 
-For this work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S}, which show broad agreement of many of their global properties with observations of the Milky Way.
-The $\z=0$ snapshots\footnote{In this work, to avoid confusion with the vertical height $z$, we refer to cosmological redshift as $\z$.} of these three simulations, called m12i, m12f, and m12m for shorthand, are publicly available alongside associated mock {\em Gaia} DR2 catalogues generated from them.\footnote{available at \url{http://ananke.hub.yt}}
+Here we briefly describe the cosmological-hydrodynamical zoom-in simulations
+used in this work. Cosmological zoom-ins allow one to simulate a selected
+region at high resolution embedded in a low-resolution cosmological background
+\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added
+a hydrodynamical simulation of the baryonic component with recipes for star
+formation and feedback. The FIRE collaboration has simulated a number of Milky
+Way-mass galaxies using this technique as part of the {\em Latte} suite of
+FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. For this
+work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S},
+which show broad agreement of many of their global properties with
+observations of the Milky Way. The $\z=0$ snapshots\footnote{In this work, to
+avoid confusion with the vertical height $z$, we refer to cosmological
+redshift as $\z$.} of these three simulations, called m12i, m12f, and m12m for
+shorthand, are publicly available alongside associated mock {\em Gaia} DR2
+catalogues generated from them.\footnote{available at
+\url{http://ananke.hub.yt}}
 
@@
 
-These simulations contain dark matter particles of mass $\sim$35,000 \Msun, gas particles of mass $\sim 7000
-\text{--} 20,000\,\Msun$, and star particles of mass $\sim 5000 \text{--} 7000\,
+These simulations contain dark matter particles of mass $\sim$35,000 \Msun,
+gas particles of mass $\sim 7000 \text{--} 20,000\,\Msun$, and star particles
+of mass $\sim 5000 \text{--} 7000\,
 \Msun$,\footnote{In fact, the simulation m12i contained a gas
@@
 
-The softening lengths used in the simulations can affect the ability to resolve the very thinnest planar structures, which in turn can affect how much the density-based midplane varies as a function of azimuth.
-The Milky Way's gas disk is thought to have a scale height of {\bf XXX} pc (XX times the median gas softening length) and the thin stellar disk a scale height of {\bf XXX} pc (XX times the stellar softening length).
-We therefore expect that resolution effects may still be affecting the scale heights of these components in the simulations, and indeed the gas and stellar scale heights of the simulated galaxies are all somewhat larger than the MW's.
-However, in this work we are considering azimuthal variations in the midplane \emph{as defined by the mass density}, which will subsequently be lower in the simulations than in the MW: within 200 pc of the Sun the measured gas and stellar volume densities in the MW are roughly comparable at $\sim .04\ \Msun\ \unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see Table 3 of \citealt{2018arXiv180610564S}).
-We thus expect that the degree of midplane variation we find in the simulated systems is likely to be \emph{less} than in the real MW, since on the smallest scales the simulations are \emph{smoother} than the real Galaxy.
+The softening lengths used in the simulations can affect the ability to
+resolve the very thinnest planar structures, which in turn can affect how much
+the density-based midplane varies as a function of azimuth. The Milky Way's
+gas disk is thought to have a scale height of {\bf XXX} pc (XX times the
+median gas softening length) and the thin stellar disk a scale height of {\bf
+XXX} pc (XX times the stellar softening length). We therefore expect that
+resolution effects may still be affecting the scale heights of these
+components in the simulations, and indeed the gas and stellar scale heights of
+the simulated galaxies are all somewhat larger than the MW's. However, in this
+work we are considering azimuthal variations in the midplane \emph{as defined
+by the mass density}, which will subsequently be lower in the simulations than
+in the MW: within 200 pc of the Sun the measured gas and stellar volume
+densities in the MW are roughly comparable at $\sim .04\ \Msun\
+\unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
+Table 3 of \citealt{2018arXiv180610564S}). We thus expect that the degree of
+midplane variation we find in the simulated systems is likely to be
+\emph{less} than in the real MW, since on the smallest scales the simulations
+are \emph{smoother} than the real Galaxy.
 
-Cosmological simulations of Milky Way-mass galaxies are
-not perfect representations of the true Milky Way in other ways as well, as discussed in
+Cosmological simulations of Milky Way-mass galaxies are not perfect
+representations of the true Milky Way in other ways as well, as discussed in
 \citet{2018arXiv180610564S}. For instance, the velocity structure of m12i is
-closer to M31's than the Milky Way's (Loebman et al. in prep). However, in this work we are most interested
-in the global properties of the potential, and specifically in deviations
-from axisymmetry. From this perspective, the simulated galaxies are actually \emph{more}
-axisymmetric than we might expect of the Milky Way. While they have prominent
-spiral arms, none has as strong a bar as the MW does at present day, and none has a nearby companion like the Large Magellanic Cloud.
-One of the three we consider (\m12f) does have an ongoing interaction with a satellite galaxy similar to Sagittarius, which has punched through the galactic disk outside the solar circle and induced some warping.
-However, we expect that, as with the use of finite smoothing lengths, the degree of midplane variation in the simulations is likely a factor of a few lower than what could be expected in the MW.
+closer to M31's than the Milky Way's (Loebman et al. in prep). However, in
+this work we are most interested in the global properties of the potential,
+and specifically in deviations from axisymmetry. From this perspective, the
+simulated galaxies are actually \emph{more} axisymmetric than we might expect
+of the Milky Way. While they have prominent spiral arms, none has as strong a
+bar as the MW does at present day, and none has a nearby companion like the
+Large Magellanic Cloud. One of the three we consider (\m12f) does have an
+ongoing interaction with a satellite galaxy similar to Sagittarius, which has
+punched through the galactic disk outside the solar circle and induced some
+warping. However, we expect that, as with the use of finite smoothing lengths,
+the degree of midplane variation in the simulations is likely a factor of a
+few lower than what could be expected in the MW.
 
-In this work, we take the galactocentric coordinate system described in Section 3 of
-\citet{2018arXiv180610564S} as our fiducial coordinate system for each
-galaxy. In short, the center of the galaxy is found iteratively, the center of
-mass velocity is then determined by all star particles within $15\,\kpc$ of
-this center, and the galaxy is then rotated onto a principal-axis frame
-determined by stars younger than $1\,\Gyr$ inside of the fiducial solar radius
-$R_{\odot} = 8.2\,\kpc$, such that the disk plane is the $x$--$y$ plane.
+In this work, we take the galactocentric coordinate system described in
+Section 3 of \citet{2018arXiv180610564S} as our fiducial coordinate system for
+each galaxy. In short, the center of the galaxy is found iteratively, the
+center of mass velocity is then determined by all star particles within
+$15\,\kpc$ of this center, and the galaxy is then rotated onto a
+principal-axis frame determined by stars younger than $1\,\Gyr$ inside of the
+fiducial solar radius $R_{\odot} = 8.2\,\kpc$, such that the disk plane is the
+$x$--$y$ plane.
 
 \subsection{Local Midplane} \label{ssec:local_midplane}
-Using the three simulations, we determine the \emph{local} midplane that an observer might measure if
-they were situated in each of these galaxies. We place our imaginary observer
-at the solar radius of $8.2\,\kpc$ and azimuth $0<\phi<2\pi$\footnote{Our 
-nearly immortal observer also has warp-drive.}, and consider stars within a
-cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to the
-fiducial disk.
+Using the three simulations, we determine the \emph{local} midplane that an
+observer might measure if they were situated in each of these galaxies. We
+place our imaginary observer at the solar radius of $8.2\,\kpc$ and azimuth
+$0<\phi<2\pi$\footnote{Our nearly immortal observer also has warp-drive.}, and
+consider stars within a cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$
+perpendicular to the fiducial disk.
 

### 2019-02-20 paper/ms.tex
Small edits, remove comments

@@
 
-%where is Figure 4 referred to? we should include a discussion of it before starting the next section
-
 \section{Azimuthal midplane variations at the Solar Circle: estimates from
@@
 
-%First, a dark matter only simulation was run with $\Lambda$CDM cosmology parameterized by $\Omega_m = 0.272$, $\Omega_{\Lambda} = 0.728$, $\Omega_b = 0.0455$, $h = 0.702$, $\sigma_8 = 0.807$, and $n_s = 0.961$, consistent with current constraints \citep{2018arXiv180706209P}.
-Halos are selected at $\z=0$ based solely on their mass ($\sim 1\text{--}2 \times 10^{12} \Msun$).
-Particles within $5 R_{200\text{m}}$ of each halo are traced back to $\z=99$ and the initial conditions are regenerated at higher resolution using MUSIC \citep{2011MNRAS.415.2101H}.
-
-%Simulations were run using \texttt{GIZMO}\footnote{\url{http://www.tapir.caltech.edu/~phopkins/Site/GIZMO.html}} \citep{2015MNRAS.450...53H}.
-Hydrodynamics is solved using the meshless finite-mass method and gravity is solved using a modified version of the Tree-PM solver of \texttt{GADGET-3}, using fully adaptive and fully conservative gravitational force softening for gas.
-
-These simulations contain dark matter particles of mass $\sim$35,000 \Msun,
+These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
 gas particles of mass $\sim 7000 \text{--} 20,000\,\Msun$, and star particles

### 2019-02-20 paper/ms.tex
Fix m12x macros

@@
 \newcommand{\unit}[2]{\ensuremath{\textrm{#1}^{\mathrm{#2}}}}
-\newcommand{\m12i}{\textsf{m12i}}
-\newcommand{\m12f}{\textsf{m12f}}
-\newcommand{\m12m}{\textsf{m12m}}
+
+\newcommand{\mi}{\textsf{m12i}}
+\newcommand{\mf}{\textsf{m12f}}
+\newcommand{\mm}{\textsf{m12m}}
 
@@
 bar as the MW does at present day, and none has a nearby companion like the
-Large Magellanic Cloud. One of the three we consider (\m12f) does have an
+Large Magellanic Cloud. One of the three we consider (\mf) does have an
 ongoing interaction with a satellite galaxy similar to Sagittarius, which has

### 2019-02-23 paper/ms.tex
Add new cold gas soft. length

@@
 Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
-median softening length for gas particles around roughly solar positions (with
-cylindrical radii within $500\,\pc$ of $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is
-$98\,\pc$.
+median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
+roughly solar positions (with cylindrical radii within $500\,\pc$ of
+$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for m12i, m12f,
+and m12m, respectively.
 

### 2019-02-25 paper/ms.tex
more changes to section 3, and added some aux files to gitignore

--- 2019-02-20 paper/ms.tex
@@
 cylindrical radii within $500\,\pc$ of $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is
-$98\,\pc$.
+$98\,\pc$. %***see Issue***
 
@@
 the density-based midplane varies as a function of azimuth. The Milky Way's
-gas disk is thought to have a scale height of {\bf XXX} pc (XX times the
-median gas softening length) and the thin stellar disk a scale height of {\bf
-XXX} pc (XX times the stellar softening length). We therefore expect that
+dense, star-forming gas disk is thought to have a scale height of about 40 pc (XXX times the
+median gas softening length; \citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of about 300 pc (30 times the stellar softening length; \citealt{2008ApJ...673..864J}).
+We therefore expect that
 resolution effects may still be affecting the scale heights of these
 components in the simulations, and indeed the gas and stellar scale heights of
-the simulated galaxies are all somewhat larger than the MW's. However, in this
+the simulated galaxies are equal to or larger than the MW's. However, in this
 work we are considering azimuthal variations in the midplane \emph{as defined
@@
 \subsection{Local Midplane} \label{ssec:local_midplane}
-Using the three simulations, we determine the \emph{local} midplane that an
-observer might measure if they were situated in each of these galaxies. We
-place our imaginary observer at the solar radius of $8.2\,\kpc$ and azimuth
-$0<\phi<2\pi$\footnote{Our nearly immortal observer also has warp-drive.}, and
-consider stars within a cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$
+Using the three simulations, we determine the local midplane that an
+observer might measure if they were situated in each of these galaxies, as a function of azimuth at the solar circle.
+Starting from the coordinate system described in the previous section, which is approximately aligned so that the $z$ coordinate is perpendicular to the disk plane, we
+place our imaginary observer at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth between 
+$0<\phi<2\pi$\footnote{Our very long-lived observer also has warp-drive.}. At each value of $\phi$ we then
+compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$
 perpendicular to the fiducial disk.
-
-This median height is taken to be what our observer would measure as the local
-Galactic midplane at each $\phi$, and is reported for each galaxy in
-Figure~\ref{fig:midplane}. We find that only $10$ iterations of this procedure
-are necessary for convergence. We bootstrap resample $1000$ times and report
-the $95^{\text{th}}$ minus $5^{\text{th}}$ percentile as the dashed-line error
-bars. This bootstrap resampling is performed by first selecting all stars
+We then re-define this median $z$ as the new midplane of the cylinder, re-select stars, and iterate until the median $z$ value converges.
+We find that only $10$ iterations of this procedure
+are necessary for convergence. The resulting median $z$ is taken to be what our observer would measure as the local
+Galactic midplane at each $\phi$.
+To account for the effect of particle noise, we bootstrap-resample stars within the cylinder $1000$ times and determine
+the $95^{\text{th}}$ minus $5^{\text{th}}$ percentile range. This bootstrap resampling is performed by first selecting all stars
 within a height of $2\,\kpc$, resampling that selection, and then repeating
@@
 
-To allow for potential small inaccuracies in the fiducial coordinate systems,
+To allow for potential small inaccuracies in the determination of the original fiducial coordinate system,
 we also subtract the best fit $\text{A} \sin{\left(\phi + \text{B}\right)} +
@@
 Figure~\ref{fig:midplane} shows the inferred midplane our imaginary observer
-would make as a function of azimuth for each galaxy. The $90\%$ interquartile
+would make as a function of azimuth for each galaxy. The 5--95 percentile range from the bootstrap procedure is shown as the dashed-line error bars. The $90\%$ interquartile
 range for each galaxy is $185$, $162$, $84\,\pc$ for m12i, m12f, and m12m.
@@
 that the local standard of rest (LSR) is a function of azimuth as well. We
-perform this calculation in Appendix~\ref{app:lsr}, but in order to perform a
-best-fit subtraction as in the previous section is more involved, which we
+perform this calculation in Appendix~\ref{app:lsr}, but performing a
+best-fit subtraction to correct for misalignment of the original coordinate system (as in the previous section) is more involved, and we
 delay to future work. We do note, however, that the variation in the LSR is

### 2019-02-25 paper/ms.tex

@@
 Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
-median softening length for gas particles around roughly solar positions (with
-cylindrical radii within $500\,\pc$ of $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is
-$98\,\pc$. %***see Issue***
+median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
+roughly solar positions (with cylindrical radii within $500\,\pc$ of
+$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for m12i, m12f,
+and m12m, respectively.
 

### 2019-02-25 paper/ms.tex
changed simulation names to sans serif

--- 2019-02-25 paper/ms.tex
@@
 avoid confusion with the vertical height $z$, we refer to cosmological
-redshift as $\z$.} of these three simulations, called m12i, m12f, and m12m for
+redshift as $\z$.} of these three simulations, called \mi , \mf , and \mm\ for
 shorthand, are publicly available alongside associated mock {\em Gaia} DR2
@@
 of mass $\sim 5000 \text{--} 7000\,
-\Msun$,\footnote{In fact, the simulation m12i contained a gas
+\Msun$,\footnote{In fact, the simulation \mi\ contained a gas
 splitting bug which causes a small number of gas and star particles to have
@@
 Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
-median softening length for gas particles around roughly solar positions (with
-cylindrical radii within $500\,\pc$ of $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is
-$98\,\pc$. %***see Issue***
+median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
+roughly solar positions (with cylindrical radii within $500\,\pc$ of
+$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi , \mf ,
+and \mm , respectively.
 
@@
 the density-based midplane varies as a function of azimuth. The Milky Way's
-dense, star-forming gas disk is thought to have a scale height of about 40 pc (XXX times the
-median gas softening length; \citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of about 300 pc (30 times the stellar softening length; \citealt{2008ApJ...673..864J}).
+dense, star-forming gas disk is thought to have a scale height of about 40 pc (3--4 times the
+cold gas softening length; \citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of about 300 pc (30 times the stellar softening length; \citealt{2008ApJ...673..864J}).
 We therefore expect that
@@
 representations of the true Milky Way in other ways as well, as discussed in
-\citet{2018arXiv180610564S}. For instance, the velocity structure of m12i is
+\citet{2018arXiv180610564S}. For instance, the velocity structure of \mi\ is
 closer to M31's than the Milky Way's (Loebman et al. in prep). However, in
@@
 $0.65$, and $0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$,
-and $-15\,\pc$ for m12i, m12f, and m12m, respectively. For the assumed solar
+and $-15\,\pc$ for \mi , \mf , and \mm , respectively. For the assumed solar
 radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
 from the values of $\text{A}$ --- we compute $1.22$, $0.43$, and $0.03\,\deg$
-for m12i, m12f, and m12m. These angle offsets are consistent with the values
+for \mi , \mf , and \mm . These angle offsets are consistent with the values
 given in \citet{2018arXiv180610564S} for the difference between the $z$-axis
@@
 would make as a function of azimuth for each galaxy. The 5--95 percentile range from the bootstrap procedure is shown as the dashed-line error bars. The $90\%$ interquartile
-range for each galaxy is $185$, $162$, $84\,\pc$ for m12i, m12f, and m12m.
+range for each galaxy is $185$, $162$, $84\,\pc$ for \mi , \mf , and \mm .
 
@@
 \caption{The local midplane determined at the fiducial Solar radius
-($8.2\,\kpc$) for the three FIRE galaxies m12i, m21f, and m12m ({\em left},
+($8.2\,\kpc$) for the three FIRE galaxies \mi , \mf , and \mm\ ({\em left},
 {\em center}, and {\em right} panels). The local midplane is determined at a
@@
 Milky Way, and do not reproduce every available observation of the Milky Way.
-For instance, the velocity dispersion-age relation in m12i, m12f, and m12m is
+For instance, the velocity dispersion-age relation in \mi , \mf , and \mm\ is
 closer to that of M31 \citep{2018arXiv180610564S}.
@@
 consider a subset of Table~3 from \citet{2018arXiv180610564S} --- our
-Table~\ref{tab:scale_height} shows the stellar scale heights of m12i, m12f,
-and m12m and the current best estimates for the Milky Way. One noticeable
+Table~\ref{tab:scale_height} shows the stellar scale heights of \mi , \mf ,
+and \mm\ and the current best estimates for the Milky Way. One noticeable
 feature is that the FIRE galaxies are more ``puffy'' than the Milky Way.
@@
 MW\tablenotemark{a} & 300 & 900 \\
-m12i\tablenotemark{b} & 480 & 2000 \\
-m12f\tablenotemark{b} & 440 & 1280 \\
-m12m\tablenotemark{b} & 290 & 1030 \\
+\mi\tablenotemark{b} & 480 & 2000 \\
+\mf\tablenotemark{b} & 440 & 1280 \\
+\mm\tablenotemark{b} & 290 & 1030 \\
 \enddata
@@
 Another important aspect to consider when comparing the simulated galaxies to
-the Milky Way is the environment/merger history. For instance, m12i and m12m
-both lack a massive neighbor analogous to the LMC. On the other hand, m12f has
+the Milky Way is the environment/merger history. For instance, \mi\ and \mm\
+both lack a massive neighbor analogous to the LMC. On the other hand, \mf\ has
 had a recenter merger with \Gus{something}. \Gus{Expand more here - how to
@@
 The neighborhood around $\sim100\Msun$ clusters can be as small as
-$\sim300\,\pc$ but up to $\sim1.2\,\kpc$ for m12i and m12f, though the
-situation is marginally better for m12m with values ranging from
+$\sim300\,\pc$ but up to $\sim1.2\,\kpc$ for \mi\ and \mf , though the
+situation is marginally better for \mm\ with values ranging from
 $\sim0.8\,\kpc$ to $\sim3\,\kpc$. The neighborhood values steadily increase
 with $m_c$, topping out at $\sim5\,\kpc$ for $\sim10^4\,\Msun$ clusters in
-m12i and m12f. In m12m, the neighborhood value reaches the maximum allowed of
+\mi\ and \mf . In \mm , the neighborhood value reaches the maximum allowed of
 $16.4\,\kpc$ for clusters of mass $\sim800\,\Msun$. The interpretation of this
@@
 For the thick-disk orbit (brown), the neighborhood is at the maximum value for
-nearly all masses, except at the very low-mass end in m12i.
+nearly all masses, except at the very low-mass end in \mi .
 
@@
 clusters in Table~\ref{tab:real_clusters} using the midplane variations from
-m12i. We use m12i as a conservative choice, since it provides the greatest
+\mi . We use \mi\ as a conservative choice, since it provides the greatest
 midplane variation (Figure~\ref{fig:midplane}). We report the neighborhood
@@
 neigbhorhood around each cluster is typically $\gtrsim 500\,\pc$. This
-analysis assumes that the midplane from m12i can be accurately applied to
+analysis assumes that the midplane from \mi\ can be accurately applied to
 Milky Way orbits, which is not necessarily true. However, our analysis is
@@
 Table~\ref{tab:orbits}) and for each of the FIRE galaxies considered here ---
-m12i, m12f, and m12m ({\em left}, {\em center}, and {\em right} panels,
+\mi , \mf , and \mm\ ({\em left}, {\em center}, and {\em right} panels,
 respectively). Given the $z$ offset for a cluster of mass $m_c$ computed in
@@
 that the midplane varies by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
-three galaxies we considered (m12i, m12f, and m12m, respectively).
+three galaxies we considered (\mi , \mf , and \mm , respectively).
 

### 2019-02-25 paper/ms.tex
last changes to section 3 (for now)

@@
 Galactic midplane at each $\phi$.
+This procedure assumes perfect density estimation, and therefore perfect corrections for extinction, within the cylinder defining the ``solar neighborhood.'' Imperfect extinction correction is likely to increase the amplitude of the estimated fluctuations in $z$.
 To account for the effect of particle noise, we bootstrap-resample stars within the cylinder $1000$ times and determine
@@
 we also subtract the best fit $\text{A} \sin{\left(\phi + \text{B}\right)} +
-\text{C}$ curve from the midplane as a function of azimuth. For $\text{A}$ the
+\text{C}$ curve from the midplane as a function of azimuth to account for an overall tilt of the midplane (a simplified version of the strategy described in \citealt{2019ApJ...871..145A}).
+For $\text{A}$ the
 values are $-174$, $62$, and $3.8\,\pc$, for $\text{B}$ the values are $0.3$,
@@
 
-Figure~\ref{fig:midplane} shows the inferred midplane our imaginary observer
-would make as a function of azimuth for each galaxy. The 5--95 percentile range from the bootstrap procedure is shown as the dashed-line error bars. The $90\%$ interquartile
+Figure~\ref{fig:midplane} shows the relative $z$ location of the inferred midplane our imaginary observer
+would determine as a function of azimuth for each galaxy, using their local ``solar neighborhood'' (the cylinder defined above).
+The 5--95 percentile range from the bootstrap procedure is shown as the dashed-line error bars.
+The $90\%$ interquartile
 range for each galaxy is $185$, $162$, $84\,\pc$ for \mi , \mf , and \mm .
+In two of the three cases this varies by $\pm 100$pc depending on the azimuth along the Solar circle; in the third (\mm , which has the thinnest ``thin disk'' of stars, but the largest stellar mass) the variation is more like $\pm 50$ pc.
 
@@
 determination (i.e. before subtracting the best fit sine curve) in
-Appendix~\Gus{x}. We then bootstrap resample $1000$ times on all stars within
+Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars within
 a $2\,\kpc$ height of the fiducial midplane to determine error bars (95\uth
@@
 \subsection{LSR Variations} \label{ssec:lsr_var}
-For similar reasons to the midplane varying across the disk, we also expect
-that the local standard of rest (LSR) is a function of azimuth as well. We
-perform this calculation in Appendix~\ref{app:lsr}, but performing a
-best-fit subtraction to correct for misalignment of the original coordinate system (as in the previous section) is more involved, and we
-delay to future work. We do note, however, that the variation in the LSR is
-less pronounced than for the midplane, at least for how we calculate the LSR.
+We also expect
+that the local standard of rest (LSR) should vary as a function of azimuth. We
+perform this calculation in Appendix~\ref{app:lsr} to estimate , but performing a
+best-fit subtraction to correct for misalignment of the original coordinate system (as in the previous section) is more involved. Since the variation in the LSR is
+less pronounced than for the midplane, and since the dominant effect is limited to $J_r$ (Figure \ref{fig:many_orbit_wrong_ref}), we defer this calculation to future work.
 

### 2019-02-25 paper/ms.tex
moved table 3 earlier and removed redundant discussion in Section 4

@@
 shorthand, are publicly available alongside associated mock {\em Gaia} DR2
-catalogues generated from them.\footnote{available at
+catalogues generated from them.\footnote{at
 \url{http://ananke.hub.yt}}
@@
 $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi , \mf ,
-and \mm , respectively.
+and \mm , respectively. These values are summarized in Table \ref{tab:scale_height}.
+
+\begin{deluxetable*}{cccccc}
+\tablecaption{Stellar disk scale heights of the real Milky Way and the
+simulated galaxies considered in this work.\label{tab:scale_height}}
+\tablehead{\colhead{galaxy} & \colhead{cold\tablenotemark{a} gas disk} & \colhead{stellar thin disk} & \colhead{stellar thick disk} & \colhead{cold gas softening length} & \colhead{stellar softening length} \\
+\colhead{} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} } 
+\startdata
+MW\tablenotemark{b} & 40 & 300 & 900 & \nodata & \nodata \\
+\mi\tablenotemark{c} & 800\tablenotemark{d} & 480 & 2000 & 12.4 & 11.2 \\
+\mf\tablenotemark{c} & 360 & 440 & 1280 & 12.5 & 11.2 \\
+\mm\tablenotemark{c} & 250 & 290 & 1030 &10.3 & 11.2 \\
+\enddata
+% \tablerefs{\citet{2016ARAA..54..529B,2018arXiv180610564S,2008ApJ...673..864J}}
+\tablenotetext{a}{$T<1000$ K}
+\tablenotetext{b}{\citet{2008ApJ...673..864J,2016ARAA..54..529B}}
+\tablenotetext{c}{\citet{2018arXiv180610564S}}
+\tablenotetext{d}{The azimuthally averaged gas vertical density profile in m12i is nearly constant to this height, though individual regions show smaller scale heights and dense clouds.}
+\end{deluxetable*}
 
@@
 resolution effects may still be affecting the scale heights of these
-components in the simulations, and indeed the gas and stellar scale heights of
-the simulated galaxies are equal to or larger than the MW's. However, in this
-work we are considering azimuthal variations in the midplane \emph{as defined
-by the mass density}, which will subsequently be lower in the simulations than
+components in the simulations, especially the cold gas. Indeed, the stellar scale heights of
+the simulated galaxies are equal to or larger than the MW's while the gas scale heights are significantly larger (although the proper basis comparison is less clear in the case of the gas; the quoted value for the Milky Way comes from studies of high-mass star-forming regions).
+As noted above, the midplanes defined by gas and stars can be tilted with respect to one another as well, precluding extending the precision of the gas midplane definition to the stellar component.
+In this
+work we specifically consider azimuthal variations in the midplane \emph{as defined
+by the stellar mass density}, which will subsequently be lower in the simulations than
 in the MW: within 200 pc of the Sun the measured gas and stellar volume
@@
 \section{Discussion} \label{sec:discussion}
-We have shown that, in simulation, the ``local midplane'' can have strong
-azimuthal dependence. The implications of this finding are not clear, but
-clearly demonstrate that assuming our local midplane is the same as the global
-midplane is insufficient for extracting the most dynamical information from
-the {\em Gaia} mission.
+We have used high-resolution cosmologial simulations to illustrate that the ``local midplane'' defined by stellar density can have a strong azimuthal dependence, as a natural consequence of the non-axisymmetry of the Galactic disk at small scales.
+While this is not in itself surprising or new, we also demonstrate that extending the coordinate system established by our local midplane to a global axisymmetric coordinate system spanning the entirety of the Galactic disk introduces systematic error in computations of the actions under this symmetry assumption, when starting from the present-day positions and velocities of stars as measured, for example, by \emph{Gaia}.
 
@@
 
-\subsection{Applicability to the Milky Way} \label{ssec:is_it_real}
-As we briefly discussed in Section~\ref{ssec:cosmozoom}, cosmological zoom-in
-simulations of Milky Way-mass galaxies are not perfect representations of the
-Milky Way, and do not reproduce every available observation of the Milky Way.
-For instance, the velocity dispersion-age relation in \mi , \mf , and \mm\ is
-closer to that of M31 \citep{2018arXiv180610564S}.
-
-In order to assess the applicability of our work to the real Milky Way, we
-consider a subset of Table~3 from \citet{2018arXiv180610564S} --- our
-Table~\ref{tab:scale_height} shows the stellar scale heights of \mi , \mf ,
-and \mm\ and the current best estimates for the Milky Way. One noticeable
-feature is that the FIRE galaxies are more ``puffy'' than the Milky Way.
-\Gus{Maybe FIRE collaboration has some insight into the mechanism for this?}.
-This would seem to suggest that the midplane estimates from
-Figure~\ref{fig:midplane} over-estimate the true variation in the Milky Way.
-
-\begin{deluxetable}{ccc}
-\tablecaption{Stellar disk scale heights of the real Milky Way and the
-simulated galaxies considered in this work.\label{tab:scale_height}}
-\tablehead{\colhead{galaxy} & \colhead{thin disk} & \colhead{thick disk} \\ 
-\colhead{} & \colhead{(pc)} & \colhead{(pc)} } 
-\startdata
-MW\tablenotemark{a} & 300 & 900 \\
-\mi\tablenotemark{b} & 480 & 2000 \\
-\mf\tablenotemark{b} & 440 & 1280 \\
-\mm\tablenotemark{b} & 290 & 1030 \\
-\enddata
-% \tablerefs{\citet{2016ARAA..54..529B,2018arXiv180610564S,2008ApJ...673..864J}}
-
-\tablenotetext{a}{\citet{2008ApJ...673..864J,2016ARAA..54..529B}}
-\tablenotetext{b}{\citet{2018arXiv180610564S}}
-\end{deluxetable}
-
-Another important aspect to consider when comparing the simulated galaxies to
-the Milky Way is the environment/merger history. For instance, \mi\ and \mm\
-both lack a massive neighbor analogous to the LMC. On the other hand, \mf\ has
-had a recenter merger with \Gus{something}. \Gus{Expand more here - how to
-incorporate merger trees?}
+%%% this stuff is now in section 3 mostly
+%\subsection{Applicability to the Milky Way} \label{ssec:is_it_real}
+%As we briefly discussed in Section~\ref{ssec:cosmozoom}, cosmological zoom-in
+%simulations of Milky Way-mass galaxies are not perfect representations of the
+%Milky Way, and do not reproduce every available observation of the Milky Way.
+%For instance, the velocity dispersion-age relation in \mi , \mf , and \mm\ is
+%closer to that of M31 \citep{2018arXiv180610564S}.
+%
+%In order to assess the applicability of our work to the real Milky Way, we
+%consider a subset of Table~3 from \citet{2018arXiv180610564S} --- our
+%Table~\ref{tab:scale_height} shows the stellar scale heights of \mi , \mf ,
+%and \mm\ and the current best estimates for the Milky Way. One noticeable
+%feature is that the FIRE galaxies are more ``puffy'' than the Milky Way.
+%\Gus{Maybe FIRE collaboration has some insight into the mechanism for this?}.
+%This would seem to suggest that the midplane estimates from
+%Figure~\ref{fig:midplane} over-estimate the true variation in the Milky Way.
+
+%Another important aspect to consider when comparing the simulated galaxies to
+%the Milky Way is the environment/merger history. For instance, \mi\ and \mm\
+%both lack a massive neighbor analogous to the LMC. On the other hand, \mf\ has
+%had a recenter merger with \Gus{something}. \Gus{Expand more here - how to
+%incorporate merger trees?}
 

### 2019-02-25 paper/ms.tex
added to intro of section 4

@@
 \section{Discussion} \label{sec:discussion}
-We have used high-resolution cosmologial simulations to illustrate that the ``local midplane'' defined by stellar density can have a strong azimuthal dependence, as a natural consequence of the non-axisymmetry of the Galactic disk at small scales.
+We have used high-resolution cosmological simulations to illustrate that we expect the ``local midplane'' defined by stellar density to vary with azimuth by up to $\pm 100$ pc, as a natural consequence of the non-axisymmetry of the Galactic disk at small scales.
 While this is not in itself surprising or new, we also demonstrate that extending the coordinate system established by our local midplane to a global axisymmetric coordinate system spanning the entirety of the Galactic disk introduces systematic error in computations of the actions under this symmetry assumption, when starting from the present-day positions and velocities of stars as measured, for example, by \emph{Gaia}.
-
-While we focus our discussion now solely on the midplane, there is an
-additional complication in the measurement of the distance from the Sun to the
-Galactic center that may also be important. When modelling stars at Galactic
-radii close to the Sun's radius, it is in principle possible to determine what
-point they are orbitting about, or in other words the ``dynamic Galactic
-center.'' It is not guaranteed that this dynamic Galactic center will coincide
-with the distance to Sgr~A\textsuperscript{*} \citep{2018AA...615L..15G}. In
-fact, \citet{2015ApJ...803...80K} were able to measure the distance of the Sun
-to the Galactic center using the dynamical information of Palomar 5. Other
-dynamical measurements of the Solar distance have been made, though none with
-a precision capable of competing with the distance to Sgr~A\textsuperscript{*}
-\citep{1981gask.book.....M,2011PASJ...63..867S,2012MNRAS.427..274S,2013AstL...39...95B,2013IAUS..289..444Z}.
+These systematic errors are most important for stars on thin-disk-like orbits, where they can be large enough to yield actions representative of orbits in the thick disk.
+This effect is entirely due to the extension of a local to a global coordinate system, and is separate from real diffusion in stellar orbits caused by interactions with these same deviations from axisymmetry such as resonant perturbation by spiral arms or scattering from molecular clouds.
+
+This finding has many implications for the study of the dynamics of stars in what is normally considered the regime of the epicyclic approximation: quasi-harmonic oscillations around a circular ``guiding center orbit.'' Here we consider a few of those implications, starting with an estimate of the region around the Sun where we expect an extension of the local axisymmetric coordinate system, and therefore the epicyclic approximation, to still be valid.
+This region is also where it should be feasible to perform ``dynamical tagging'' in the thin disk; that is, using the computed orbits or integrals of motion of field stars to match them with known open clusters.
+
+%%%move this later - decide where
+%While we focus our discussion now solely on the midplane, there is an
+%additional complication in the measurement of the distance from the Sun to the
+%Galactic center that may also be important. When modelling stars at Galactic
+%radii close to the Sun's radius, it is in principle possible to determine what
+%point they are orbiting about: the ``dynamical Galactic
+%center.'' It is not guaranteed that this dynamical center will coincide
+%with the distance to Sgr~A\textsuperscript{*} \citep{2018AA...615L..15G}. In
+%fact, \citet{2015ApJ...803...80K} were able to measure the distance of the Sun
+%to the Galactic center using the dynamical information of Palomar 5. Other
+%dynamical measurements of the Solar distance have been made, though none with
+%a precision capable of competing with the distance to Sgr~A\textsuperscript{*}
+%\citep{1981gask.book.....M,2011PASJ...63..867S,2012MNRAS.427..274S,2013AstL...39...95B,2013IAUS..289..444Z}.
 

### 2019-02-25 paper/ms.tex
revisions to new section 4.1, also moved Table 3 earlier (to section 3) and added a bit more detail there. See related issues to be raised on Figure 7 and accompanying discussion

@@
 range for each galaxy is $185$, $162$, $84\,\pc$ for \mi , \mf , and \mm .
-In two of the three cases this varies by $\pm 100$pc depending on the azimuth along the Solar circle; in the third (\mm , which has the thinnest ``thin disk'' of stars, but the largest stellar mass) the variation is more like $\pm 50$ pc.
+In two of the three cases the midplane therefore varies by more than $\pm 100$pc depending on the azimuth along the Solar circle; in the third (\mm , which has the thinnest ``thin disk'' of stars, but the largest stellar mass) the variation is more like $\pm 50$ pc.
 
@@
 These systematic errors are most important for stars on thin-disk-like orbits, where they can be large enough to yield actions representative of orbits in the thick disk.
-This effect is entirely due to the extension of a local to a global coordinate system, and is separate from real diffusion in stellar orbits caused by interactions with these same deviations from axisymmetry such as resonant perturbation by spiral arms or scattering from molecular clouds.
+This effect is entirely due to the extension of a local to a global coordinate system, and is separate from real diffusion in stellar integrals of motion caused by interactions with these same deviations from axisymmetry, such as resonant perturbation by spiral arms or scattering from molecular clouds.
 
@@
 
-%%%move this later - decide where
+%%%move this later - decide where, probably conclusions
 %While we focus our discussion now solely on the midplane, there is an
@@
 
-\subsection{Impact on Milky Way observations} \label{ssec:impact_on_mw}
-\subsubsection{The Solar neighborhood} \label{sssec:neighborhood}
-To gain further handle on the midplane behavior as a function of azimuth, we
-compute the range of midplane values for a given $\Delta \phi$ segment of the
-midplane in Figure~\ref{fig:range_deltaphi}. With the observer's azimuth set
+%\subsection{Impact on Milky Way observations} \label{ssec:impact_on_mw}
+\subsection{The size of the Solar neighborhood} \label{sssec:neighborhood}
+
+Given a local coordinate system, the amplitude and characteristic length scale of the density fluctuations in the Galactic disk will determine how far this system can be extended before introducing $\mathcal{O}(1)$ systematic errors in the computation of actions for thin-disk stars.
+We consider this a criterion for the region around the Sun where the analysis of the action-space distribution of stars on quasi-circular orbits, and therefore the epicyclic approximation, can be expected to be valid.
+This leads to a somewhat dynamically motivated definition of the ``solar neighborhood.''
+
+To illustrate this concept, we first
+compute the absolute range of midplane values for a given $\Delta \phi$ segment of the
+midplane. With the observer's azimuth set
 to zero, we consider a region extending from $-\Delta \phi/2$ to
-$\Delta\phi/2$. We also compute the chord length for the chord extending from
-the observer to $\pm \Delta\phi/2$. We repeat this for each of the $50$ bins
-in azimuth. Each bin is plotted as a gray line, with the mean value in solid
-blue and the $1\,\sigma$ dispersion as the dashed blue lines.
+$\Delta\phi/2$.
+To translate this angular scale into a physical scale, we also compute the chord length $\lambda$ for the chord extending from the observer to $\pm \Delta\phi/2$ (see Figure \ref{fig:fig_to_explain}).
+We repeat this for each of $50$ bins in azimuth.
+Each bin is plotted in Figure~\ref{fig:range_deltaphi} as a gray line, with the mean value in solid blue and the $1\,\sigma$ dispersion as the dashed blue lines.
 
@@
 
-\subsubsection{Consequences for disrupted cluster reconstruction}
+From Figure \ref{fig:range_deltaphi} it is clear that the range of midplane values spanned in a given $\Delta \phi$ quickly saturates at $\Delta \phi \gtrsim \pi/2$ radians (i.e.
+one quarter of the way around the Galactic disk from the Sun).
+At far smaller $\Delta \phi$ it reaches the level of deviation required for an $\mathcal{O}(1)$ systematic error in $J_z$ for our fiducial thin-disk-like orbit (*color1* horizontal line in Figure \ref{fig:range_deltaphi}).
+For the three simulations considered, the median range exceeds this value at $\Delta \phi >$ XX, YY, and ZZ radians, with corresponding physical scales of $\lambda \sim$ XX, YY, and ZZ parsecs.
+These can be considered estimates of the range of validity for an extension of the locally determined axisymmetric coordinate system, at least when it comes to the study of stars on quasi-circular orbits.
+
+The situation is more forgiving for the study of thick-disk-like orbits, where an $\mathcal{O}(1)$ systematic error in $J_z$ requires a larger midplane deviation (*color2* horizontal line in Figure \ref{fig:range_deltaphi}) corresponding to $\Delta \phi >$ XX, YY, and ZZ radians or $\lambda \sim$ XX, YY, and ZZ parsecs.
+For stars on halo-like orbits, extending the local coordinate system globally is not likely to cause serious problems, since the midplane deviations are relatively small compared to those needed to produce $\mathcal{O}(1)$ systematic errors in the actions (*color3* horizontal line in Figure \ref{fig:range_deltaphi}).
+However, we caution that the tendency of these midplane deviations is to \emph{inflate} the value of $J_z$ for thin disk stars in an extremely non-Gaussian way (Figure \ref{fig:Jz_hist}), scattering them into the region of action space normally associated with different stellar populations and possibly contributing to confusion there.
+
+\subsection{Consequences for disrupted cluster reconstruction}
 \label{sssec:reconstruction}
 
-A possible interpretation of our finding is that one should use a star's local
-midplane in order to compute dynamical quantities like the vertical action.
-Under the assumption that this interpretation is correct, we now consider the
-implications for attempts to reconstruct disrupting open clusters by selecting
+We now consider the implications for attempts to reconstruct disrupting open clusters by selecting
 ejected members based on their actions.

### 2019-02-26 paper/ms.tex
Fix line indenting

@@
 \tablenotetext{c}{\citet{2018arXiv180610564S}}
-\tablenotetext{d}{The azimuthally averaged gas vertical density profile in m12i is nearly constant to this height, though individual regions show smaller scale heights and dense clouds.}
+\tablenotetext{d}{The azimuthally averaged gas vertical density profile in
+m12i is nearly constant to this height, though individual regions show smaller
+scale heights and dense clouds.}
 \end{deluxetable*}
@@
 the density-based midplane varies as a function of azimuth. The Milky Way's
-dense, star-forming gas disk is thought to have a scale height of about 40 pc (3--4 times the
-cold gas softening length; \citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of about 300 pc (30 times the stellar softening length; \citealt{2008ApJ...673..864J}).
-We therefore expect that
+dense, star-forming gas disk is thought to have a scale height of about $40\,\pc$
+($3\text{-}4$ times the cold gas softening length; \citealt{2019ApJ...871..145A}) and
+the thin stellar disk a scale height of about 300 pc (30 times the stellar
+softening length; \citealt{2008ApJ...673..864J}). We therefore expect that
 resolution effects may still be affecting the scale heights of these
-components in the simulations, especially the cold gas. Indeed, the stellar scale heights of
-the simulated galaxies are equal to or larger than the MW's while the gas scale heights are significantly larger (although the proper basis comparison is less clear in the case of the gas; the quoted value for the Milky Way comes from studies of high-mass star-forming regions).
-As noted above, the midplanes defined by gas and stars can be tilted with respect to one another as well, precluding extending the precision of the gas midplane definition to the stellar component.
-In this
-work we specifically consider azimuthal variations in the midplane \emph{as defined
-by the stellar mass density}, which will subsequently be lower in the simulations than
-in the MW: within 200 pc of the Sun the measured gas and stellar volume
-densities in the MW are roughly comparable at $\sim .04\ \Msun\
+components in the simulations, especially the cold gas. Indeed, the stellar
+scale heights of the simulated galaxies are equal to or larger than the MW's
+while the gas scale heights are significantly larger (although the proper
+basis comparison is less clear in the case of the gas; the quoted value for
+the Milky Way comes from studies of high-mass star-forming regions). As noted
+above, the midplanes defined by gas and stars can be tilted with respect to
+one another as well, precluding extending the precision of the gas midplane
+definition to the stellar component. In this work we specifically consider
+azimuthal variations in the midplane \emph{as defined by the stellar mass
+density}, which will subsequently be lower in the simulations than in the MW:
+within 200 pc of the Sun the measured gas and stellar volume densities in the
+MW are roughly comparable at $\sim .04\ \Msun\
 \unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
@@
 \subsection{Local Midplane} \label{ssec:local_midplane}
-Using the three simulations, we determine the local midplane that an
-observer might measure if they were situated in each of these galaxies, as a function of azimuth at the solar circle.
-Starting from the coordinate system described in the previous section, which is approximately aligned so that the $z$ coordinate is perpendicular to the disk plane, we
-place our imaginary observer at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth between 
-$0<\phi<2\pi$\footnote{Our very long-lived observer also has warp-drive.}. At each value of $\phi$ we then
-compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$
-perpendicular to the fiducial disk.
-We then re-define this median $z$ as the new midplane of the cylinder, re-select stars, and iterate until the median $z$ value converges.
-We find that only $10$ iterations of this procedure
-are necessary for convergence. The resulting median $z$ is taken to be what our observer would measure as the local
-Galactic midplane at each $\phi$.
-This procedure assumes perfect density estimation, and therefore perfect corrections for extinction, within the cylinder defining the ``solar neighborhood.'' Imperfect extinction correction is likely to increase the amplitude of the estimated fluctuations in $z$.
-To account for the effect of particle noise, we bootstrap-resample stars within the cylinder $1000$ times and determine
-the $95^{\text{th}}$ minus $5^{\text{th}}$ percentile range. This bootstrap resampling is performed by first selecting all stars
-within a height of $2\,\kpc$, resampling that selection, and then repeating
-the $10$ iterations.
-
-To allow for potential small inaccuracies in the determination of the original fiducial coordinate system,
-we also subtract the best fit $\text{A} \sin{\left(\phi + \text{B}\right)} +
-\text{C}$ curve from the midplane as a function of azimuth to account for an overall tilt of the midplane (a simplified version of the strategy described in \citealt{2019ApJ...871..145A}).
-For $\text{A}$ the
-values are $-174$, $62$, and $3.8\,\pc$, for $\text{B}$ the values are $0.3$,
-$0.65$, and $0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$,
-and $-15\,\pc$ for \mi , \mf , and \mm , respectively. For the assumed solar
+Using the three simulations, we determine the local midplane that an observer
+might measure if they were situated in each of these galaxies, as a function
+of azimuth at the solar circle. Starting from the coordinate system described
+in the previous section, which is approximately aligned so that the $z$
+coordinate is perpendicular to the disk plane, we place our imaginary observer
+at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth between
+$0<\phi<2\pi$\footnote{Our very long-lived observer also has warp-drive.}. At
+each value of $\phi$ we then compute the median $z$ for stars within a
+cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to the
+fiducial disk. We then re-define this median $z$ as the new midplane of the
+cylinder, re-select stars, and iterate until the median $z$ value converges.
+We find that only $10$ iterations of this procedure are necessary for
+convergence. The resulting median $z$ is taken to be what our observer would
+measure as the local Galactic midplane at each $\phi$. This procedure assumes
+perfect density estimation, and therefore perfect corrections for extinction,
+within the cylinder defining the ``solar neighborhood.'' Imperfect extinction
+correction is likely to increase the amplitude of the estimated fluctuations
+in $z$. To account for the effect of particle noise, we bootstrap-resample
+stars within the cylinder $1000$ times and determine the $95^{\text{th}}$
+minus $5^{\text{th}}$ percentile range. This bootstrap resampling is performed
+by first selecting all stars within a height of $2\,\kpc$, resampling that
+selection, and then repeating the $10$ iterations.
+
+To allow for potential small inaccuracies in the determination of the original
+fiducial coordinate system, we also subtract the best fit $\text{A}
+\sin{\left(\phi + \text{B}\right)} + \text{C}$ curve from the midplane as a
+function of azimuth to account for an overall tilt of the midplane (a
+simplified version of the strategy described in
+\citealt{2019ApJ...871..145A}). For $\text{A}$ the values are $-174$, $62$,
+and $3.8\,\pc$, for $\text{B}$ the values are $0.3$, $0.65$, and
+$0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$, and
+$-15\,\pc$ for \mi , \mf , and \mm , respectively. For the assumed solar
 radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
@@
 
-Figure~\ref{fig:midplane} shows the relative $z$ location of the inferred midplane our imaginary observer
-would determine as a function of azimuth for each galaxy, using their local ``solar neighborhood'' (the cylinder defined above).
-The 5--95 percentile range from the bootstrap procedure is shown as the dashed-line error bars.
-The $90\%$ interquartile
-range for each galaxy is $185$, $162$, $84\,\pc$ for \mi , \mf , and \mm .
-In two of the three cases the midplane therefore varies by more than $\pm 100$pc depending on the azimuth along the Solar circle; in the third (\mm , which has the thinnest ``thin disk'' of stars, but the largest stellar mass) the variation is more like $\pm 50$ pc.
+Figure~\ref{fig:midplane} shows the relative $z$ location of the inferred
+midplane our imaginary observer would determine as a function of azimuth for
+each galaxy, using their local ``solar neighborhood'' (the cylinder defined
+above). The 5--95 percentile range from the bootstrap procedure is shown as
+the dashed-line error bars. The $90\%$ interquartile range for each galaxy is
+$185$, $162$, $84\,\pc$ for \mi , \mf , and \mm . In two of the three cases
+the midplane therefore varies by more than $\pm 100$pc depending on the
+azimuth along the Solar circle; in the third (\mm , which has the thinnest
+``thin disk'' of stars, but the largest stellar mass) the variation is more
+like $\pm 50$ pc.
 
@@
 \subsection{LSR Variations} \label{ssec:lsr_var}
-We also expect
-that the local standard of rest (LSR) should vary as a function of azimuth. We
-perform this calculation in Appendix~\ref{app:lsr} to estimate , but performing a
-best-fit subtraction to correct for misalignment of the original coordinate system (as in the previous section) is more involved. Since the variation in the LSR is
-less pronounced than for the midplane, and since the dominant effect is limited to $J_r$ (Figure \ref{fig:many_orbit_wrong_ref}), we defer this calculation to future work.
+We also expect that the local standard of rest (LSR) should vary as a function
+of azimuth. We perform this calculation in Appendix~\ref{app:lsr} to estimate
+, but performing a best-fit subtraction to correct for misalignment of the
+original coordinate system (as in the previous section) is more involved.
+Since the variation in the LSR is less pronounced than for the midplane, and
+since the dominant effect is limited to $J_r$ (Figure
+\ref{fig:many_orbit_wrong_ref}), we defer this calculation to future work.
 
 \section{Discussion} \label{sec:discussion}
-We have used high-resolution cosmological simulations to illustrate that we expect the ``local midplane'' defined by stellar density to vary with azimuth by up to $\pm 100$ pc, as a natural consequence of the non-axisymmetry of the Galactic disk at small scales.
-While this is not in itself surprising or new, we also demonstrate that extending the coordinate system established by our local midplane to a global axisymmetric coordinate system spanning the entirety of the Galactic disk introduces systematic error in computations of the actions under this symmetry assumption, when starting from the present-day positions and velocities of stars as measured, for example, by \emph{Gaia}.
-These systematic errors are most important for stars on thin-disk-like orbits, where they can be large enough to yield actions representative of orbits in the thick disk.
-This effect is entirely due to the extension of a local to a global coordinate system, and is separate from real diffusion in stellar integrals of motion caused by interactions with these same deviations from axisymmetry, such as resonant perturbation by spiral arms or scattering from molecular clouds.
-
-This finding has many implications for the study of the dynamics of stars in what is normally considered the regime of the epicyclic approximation: quasi-harmonic oscillations around a circular ``guiding center orbit.'' Here we consider a few of those implications, starting with an estimate of the region around the Sun where we expect an extension of the local axisymmetric coordinate system, and therefore the epicyclic approximation, to still be valid.
-This region is also where it should be feasible to perform ``dynamical tagging'' in the thin disk; that is, using the computed orbits or integrals of motion of field stars to match them with known open clusters.
+We have used high-resolution cosmological simulations to illustrate that we
+expect the ``local midplane'' defined by stellar density to vary with azimuth
+by up to $\pm 100$ pc, as a natural consequence of the non-axisymmetry of the
+Galactic disk at small scales. While this is not in itself surprising or new,
+we also demonstrate that extending the coordinate system established by our
+local midplane to a global axisymmetric coordinate system spanning the
+entirety of the Galactic disk introduces systematic error in computations of
+the actions under this symmetry assumption, when starting from the present-day
+positions and velocities of stars as measured, for example, by \emph{Gaia}.
+These systematic errors are most important for stars on thin-disk-like orbits,
+where they can be large enough to yield actions representative of orbits in
+the thick disk. This effect is entirely due to the extension of a local to a
+global coordinate system, and is separate from real diffusion in stellar
+integrals of motion caused by interactions with these same deviations from
+axisymmetry, such as resonant perturbation by spiral arms or scattering from
+molecular clouds.
+
+This finding has many implications for the study of the dynamics of stars in
+what is normally considered the regime of the epicyclic approximation:
+quasi-harmonic oscillations around a circular ``guiding center orbit.'' Here
+we consider a few of those implications, starting with an estimate of the
+region around the Sun where we expect an extension of the local axisymmetric
+coordinate system, and therefore the epicyclic approximation, to still be
+valid. This region is also where it should be feasible to perform ``dynamical
+tagging'' in the thin disk; that is, using the computed orbits or integrals of
+motion of field stars to match them with known open clusters.
 
@@
 
-Given a local coordinate system, the amplitude and characteristic length scale of the density fluctuations in the Galactic disk will determine how far this system can be extended before introducing $\mathcal{O}(1)$ systematic errors in the computation of actions for thin-disk stars.
-We consider this a criterion for the region around the Sun where the analysis of the action-space distribution of stars on quasi-circular orbits, and therefore the epicyclic approximation, can be expected to be valid.
-This leads to a somewhat dynamically motivated definition of the ``solar neighborhood.''
-
-To illustrate this concept, we first
-compute the absolute range of midplane values for a given $\Delta \phi$ segment of the
-midplane. With the observer's azimuth set
-to zero, we consider a region extending from $-\Delta \phi/2$ to
-$\Delta\phi/2$.
-To translate this angular scale into a physical scale, we also compute the chord length $\lambda$ for the chord extending from the observer to $\pm \Delta\phi/2$ (see Figure \ref{fig:fig_to_explain}).
-We repeat this for each of $50$ bins in azimuth.
-Each bin is plotted in Figure~\ref{fig:range_deltaphi} as a gray line, with the mean value in solid blue and the $1\,\sigma$ dispersion as the dashed blue lines.
+Given a local coordinate system, the amplitude and characteristic length scale
+of the density fluctuations in the Galactic disk will determine how far this
+system can be extended before introducing $\mathcal{O}(1)$ systematic errors
+in the computation of actions for thin-disk stars. We consider this a
+criterion for the region around the Sun where the analysis of the action-space
+distribution of stars on quasi-circular orbits, and therefore the epicyclic
+approximation, can be expected to be valid. This leads to a somewhat
+dynamically motivated definition of the ``solar neighborhood.''
+
+To illustrate this concept, we first compute the absolute range of midplane
+values for a given $\Delta \phi$ segment of the midplane. With the observer's
+azimuth set to zero, we consider a region extending from $-\Delta \phi/2$ to
+$\Delta\phi/2$. To translate this angular scale into a physical scale, we also
+compute the chord length $\lambda$ for the chord extending from the observer
+to $\pm \Delta\phi/2$ (see Figure \ref{fig:fig_to_explain}). We repeat this
+for each of $50$ bins in azimuth. Each bin is plotted in
+Figure~\ref{fig:range_deltaphi} as a gray line, with the mean value in solid
+blue and the $1\,\sigma$ dispersion as the dashed blue lines.
 
@@
 
-From Figure \ref{fig:range_deltaphi} it is clear that the range of midplane values spanned in a given $\Delta \phi$ quickly saturates at $\Delta \phi \gtrsim \pi/2$ radians (i.e.
-one quarter of the way around the Galactic disk from the Sun).
-At far smaller $\Delta \phi$ it reaches the level of deviation required for an $\mathcal{O}(1)$ systematic error in $J_z$ for our fiducial thin-disk-like orbit (*color1* horizontal line in Figure \ref{fig:range_deltaphi}).
-For the three simulations considered, the median range exceeds this value at $\Delta \phi >$ XX, YY, and ZZ radians, with corresponding physical scales of $\lambda \sim$ XX, YY, and ZZ parsecs.
-These can be considered estimates of the range of validity for an extension of the locally determined axisymmetric coordinate system, at least when it comes to the study of stars on quasi-circular orbits.
-
-The situation is more forgiving for the study of thick-disk-like orbits, where an $\mathcal{O}(1)$ systematic error in $J_z$ requires a larger midplane deviation (*color2* horizontal line in Figure \ref{fig:range_deltaphi}) corresponding to $\Delta \phi >$ XX, YY, and ZZ radians or $\lambda \sim$ XX, YY, and ZZ parsecs.
-For stars on halo-like orbits, extending the local coordinate system globally is not likely to cause serious problems, since the midplane deviations are relatively small compared to those needed to produce $\mathcal{O}(1)$ systematic errors in the actions (*color3* horizontal line in Figure \ref{fig:range_deltaphi}).
-However, we caution that the tendency of these midplane deviations is to \emph{inflate} the value of $J_z$ for thin disk stars in an extremely non-Gaussian way (Figure \ref{fig:Jz_hist}), scattering them into the region of action space normally associated with different stellar populations and possibly contributing to confusion there.
+From Figure \ref{fig:range_deltaphi} it is clear that the range of midplane
+values spanned in a given $\Delta \phi$ quickly saturates at $\Delta \phi
+\gtrsim \pi/2$ radians (i.e. one quarter of the way around the Galactic disk
+from the Sun). At far smaller $\Delta \phi$ it reaches the level of deviation
+required for an $\mathcal{O}(1)$ systematic error in $J_z$ for our fiducial
+thin-disk-like orbit (*color1* horizontal line in Figure
+\ref{fig:range_deltaphi}). For the three simulations considered, the median
+range exceeds this value at $\Delta \phi >$ XX, YY, and ZZ radians, with
+corresponding physical scales of $\lambda \sim$ XX, YY, and ZZ parsecs. These
+can be considered estimates of the range of validity for an extension of the
+locally determined axisymmetric coordinate system, at least when it comes to
+the study of stars on quasi-circular orbits.
+
+The situation is more forgiving for the study of thick-disk-like orbits, where
+an $\mathcal{O}(1)$ systematic error in $J_z$ requires a larger midplane
+deviation (*color2* horizontal line in Figure \ref{fig:range_deltaphi})
+corresponding to $\Delta \phi >$ XX, YY, and ZZ radians or $\lambda \sim$ XX,
+YY, and ZZ parsecs. For stars on halo-like orbits, extending the local
+coordinate system globally is not likely to cause serious problems, since the
+midplane deviations are relatively small compared to those needed to produce
+$\mathcal{O}(1)$ systematic errors in the actions (*color3* horizontal line in
+Figure \ref{fig:range_deltaphi}). However, we caution that the tendency of
+these midplane deviations is to \emph{inflate} the value of $J_z$ for thin
+disk stars in an extremely non-Gaussian way (Figure \ref{fig:Jz_hist}),
+scattering them into the region of action space normally associated with
+different stellar populations and possibly contributing to confusion there.
 
@@
 
-We now consider the implications for attempts to reconstruct disrupting open clusters by selecting
-ejected members based on their actions.
+We now consider the implications for attempts to reconstruct disrupting open
+clusters by selecting ejected members based on their actions.
 

### 2019-02-26 paper/ms.tex
Fix technical piece of bootstrap resampling

@@
 in $z$. To account for the effect of particle noise, we bootstrap-resample
-stars within the cylinder $1000$ times and determine the $95^{\text{th}}$
-minus $5^{\text{th}}$ percentile range. This bootstrap resampling is performed
-by first selecting all stars within a height of $2\,\kpc$, resampling that
-selection, and then repeating the $10$ iterations.
+stars within a cylinder of stars with height $2\,\kpc$ and the same radius
+$1000$ times and determine the $95^{\text{th}}$ minus $5^{\text{th}}$
+percentile range. This bootstrap resampling is performed by first selecting
+all stars within a height of $2\,\kpc$, resampling that selection, and then
+repeating the $10$ iterations.
 

### 2019-02-26 paper/ms.tex
Change to typewriter
Small edits

@@
 
-\newcommand{\mi}{\textsf{m12i}}
-\newcommand{\mf}{\textsf{m12f}}
-\newcommand{\mm}{\textsf{m12m}}
+\newcommand{\mi}{\texttt{m12i}}
+\newcommand{\mf}{\texttt{m12f}}
+\newcommand{\mm}{\texttt{m12m}}
 
@@
 avoid confusion with the vertical height $z$, we refer to cosmological
-redshift as $\z$.} of these three simulations, called \mi , \mf , and \mm\ for
-shorthand, are publicly available alongside associated mock {\em Gaia} DR2
-catalogues generated from them.\footnote{at
-\url{http://ananke.hub.yt}}
+redshift as $\z$.} of these three simulations, called \mi , \mf , and \mm{}
+for shorthand, are publicly available alongside associated mock {\em Gaia} DR2
+catalogues generated from them.\footnote{at \url{http://ananke.hub.yt}}
 
 These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
-gas particles of mass $\sim 7000 \text{--} 20,000\,\Msun$, and star particles
+gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles
 of mass $\sim 5000 \text{--} 7000\,
@@
 $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi , \mf ,
-and \mm , respectively. These values are summarized in Table \ref{tab:scale_height}.
+and \mm , respectively. These values are summarized in Table
+\ref{tab:scale_height}.
 
@@
 each galaxy, using their local ``solar neighborhood'' (the cylinder defined
-above). The 5--95 percentile range from the bootstrap procedure is shown as
-the dashed-line error bars. The $90\%$ interquartile range for each galaxy is
-$185$, $162$, $84\,\pc$ for \mi , \mf , and \mm . In two of the three cases
-the midplane therefore varies by more than $\pm 100$pc depending on the
-azimuth along the Solar circle; in the third (\mm , which has the thinnest
-``thin disk'' of stars, but the largest stellar mass) the variation is more
-like $\pm 50$ pc.
+above). The $5\textup{--}95$ percentile range from the bootstrap procedure is
+shown as the dashed-line error bars. The $90\%$ interquartile range for each
+galaxy is $185$, $162$, $84\,\pc$ for \mi , \mf , and \mm . In two of the
+three cases the midplane therefore varies by more than $\pm 100$pc depending
+on the azimuth along the Solar circle; in the third (\mm , which has the
+thinnest ``thin disk'' of stars, but the largest stellar mass) the variation
+is more like $\pm 50$ pc.
 

### 2019-02-26 paper/ms.tex
Change to prevent overrunning margin

@@
 \subsection{Description of Simulation} \label{ssec:cosmozoom}
-Here we briefly describe the cosmological-hydrodynamical zoom-in simulations
-used in this work. Cosmological zoom-ins allow one to simulate a selected
-region at high resolution embedded in a low-resolution cosmological background
+We briefly describe the cosmological-hydrodynamical zoom-in simulations used
+in this work. Cosmological zoom-ins allow one to simulate a selected region at
+high resolution embedded in a low-resolution cosmological background
 \citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added

### 2019-02-26 paper/ms.tex
It's technically true to many OOM of precision
Small changes

@@
 of mass $\sim 5000 \text{--} 7000\,
-\Msun$,\footnote{In fact, the simulation \mi\ contained a gas
+\Msun$,\footnote{In fact, the simulation \mi{} contained a gas
 splitting bug which causes a small number of gas and star particles to have
@@
 $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi , \mf ,
-and \mm , respectively. These values are summarized in Table
-\ref{tab:scale_height}.
+and \mm , respectively. These values are summarized in
+Table~\ref{tab:scale_height}.
 
@@
 \startdata
-MW\tablenotemark{b} & 40 & 300 & 900 & \nodata & \nodata \\
+MW\tablenotemark{b} & 40 & 300 & 900 & 0 & 0 \\
 \mi\tablenotemark{c} & 800\tablenotemark{d} & 480 & 2000 & 12.4 & 11.2 \\
@@
 The neighborhood around $\sim100\Msun$ clusters can be as small as
-$\sim300\,\pc$ but up to $\sim1.2\,\kpc$ for \mi\ and \mf , though the
-situation is marginally better for \mm\ with values ranging from
+$\sim300\,\pc$ but up to $\sim1.2\,\kpc$ for \mi{} and \mf{}, though the
+situation is marginally better for \mm{} with values ranging from
 $\sim0.8\,\kpc$ to $\sim3\,\kpc$. The neighborhood values steadily increase
 with $m_c$, topping out at $\sim5\,\kpc$ for $\sim10^4\,\Msun$ clusters in
-\mi\ and \mf . In \mm , the neighborhood value reaches the maximum allowed of
+\mi{} and \mf{}. In \mm{}, the neighborhood value reaches the maximum allowed of
 $16.4\,\kpc$ for clusters of mass $\sim800\,\Msun$. The interpretation of this

### 2019-02-26 paper/ms.tex
Minor macro changes

@@
 avoid confusion with the vertical height $z$, we refer to cosmological
-redshift as $\z$.} of these three simulations, called \mi , \mf , and \mm{}
+redshift as $\z$.} of these three simulations, called \mi{}, \mf{}, and \mm{}
 for shorthand, are publicly available alongside associated mock {\em Gaia} DR2
@@
 roughly solar positions (with cylindrical radii within $500\,\pc$ of
-$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi , \mf ,
-and \mm , respectively. These values are summarized in
+$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi{},
+\mf{}, and \mm{}, respectively. These values are summarized in
 Table~\ref{tab:scale_height}.
@@
 representations of the true Milky Way in other ways as well, as discussed in
-\citet{2018arXiv180610564S}. For instance, the velocity structure of \mi\ is
+\citet{2018arXiv180610564S}. For instance, the velocity structure of \mi{} is
 closer to M31's than the Milky Way's (Loebman et al. in prep). However, in
@@
 $0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$, and
-$-15\,\pc$ for \mi , \mf , and \mm , respectively. For the assumed solar
+$-15\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. For the assumed solar
 radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
 from the values of $\text{A}$ --- we compute $1.22$, $0.43$, and $0.03\,\deg$
-for \mi , \mf , and \mm . These angle offsets are consistent with the values
-given in \citet{2018arXiv180610564S} for the difference between the $z$-axis
-as defined by the gas and stars.
+for \mi{}, \mf{}, and \mm{}. These angle offsets are consistent with the
+values given in \citet{2018arXiv180610564S} for the difference between the
+$z$-axis as defined by the gas and stars.
 
@@
 shown as the dashed-line error bars. The $90\%$ interquartile range for each
-galaxy is $185$, $162$, $84\,\pc$ for \mi , \mf , and \mm . In two of the
+galaxy is $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the
 three cases the midplane therefore varies by more than $\pm 100$pc depending
-on the azimuth along the Solar circle; in the third (\mm , which has the
+on the azimuth along the Solar circle; in the third (\mm{}, which has the
 thinnest ``thin disk'' of stars, but the largest stellar mass) the variation
@@
 \caption{The local midplane determined at the fiducial Solar radius
-($8.2\,\kpc$) for the three FIRE galaxies \mi , \mf , and \mm\ ({\em left},
+($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} ({\em left},
 {\em center}, and {\em right} panels). The local midplane is determined at a
@@
 determination (i.e. before subtracting the best fit sine curve) in
-Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars within
-a $2\,\kpc$ height of the fiducial midplane to determine error bars (95\uth
-and 5\uth percentiles), which we report as dashed lines.}
+Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
+within a $2\,\kpc$ height of the fiducial midplane to determine error bars
+(95\uth and 5\uth percentiles), which we report as dashed lines.}
 \label{fig:midplane}
@@
 For the thick-disk orbit (brown), the neighborhood is at the maximum value for
-nearly all masses, except at the very low-mass end in \mi .
+nearly all masses, except at the very low-mass end in \mi{}.
 
@@
 clusters in Table~\ref{tab:real_clusters} using the midplane variations from
-\mi . We use \mi\ as a conservative choice, since it provides the greatest
+\mi{}. We use \mi{} as a conservative choice, since it provides the greatest
 midplane variation (Figure~\ref{fig:midplane}). We report the neighborhood
@@
 neigbhorhood around each cluster is typically $\gtrsim 500\,\pc$. This
-analysis assumes that the midplane from \mi\ can be accurately applied to
+analysis assumes that the midplane from \mi{} can be accurately applied to
 Milky Way orbits, which is not necessarily true. However, our analysis is
@@
 Table~\ref{tab:orbits}) and for each of the FIRE galaxies considered here ---
-\mi , \mf , and \mm\ ({\em left}, {\em center}, and {\em right} panels,
+\mi{}, \mf{}, and \mm{} ({\em left}, {\em center}, and {\em right} panels,
 respectively). Given the $z$ offset for a cluster of mass $m_c$ computed in
@@
 that the midplane varies by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
-three galaxies we considered (\mi , \mf , and \mm , respectively).
+three galaxies we considered (\mi{}, \mf{}, and \mm{}, respectively).
 

### 2019-02-26 paper/ms.tex
Fix colors

@@
 required for an $\mathcal{O}(1)$ systematic error in $J_z$ for our fiducial
-thin-disk-like orbit (*color1* horizontal line in Figure
+thin-disk-like orbit (\thincolor horizontal line in Figure
 \ref{fig:range_deltaphi}). For the three simulations considered, the median
@@
 an $\mathcal{O}(1)$ systematic error in $J_z$ requires a larger midplane
-deviation (*color2* horizontal line in Figure \ref{fig:range_deltaphi})
+deviation (\thickcolor horizontal line in Figure \ref{fig:range_deltaphi})
 corresponding to $\Delta \phi >$ XX, YY, and ZZ radians or $\lambda \sim$ XX,
@@
 midplane deviations are relatively small compared to those needed to produce
-$\mathcal{O}(1)$ systematic errors in the actions (*color3* horizontal line in
-Figure \ref{fig:range_deltaphi}). However, we caution that the tendency of
+$\mathcal{O}(1)$ systematic errors in the actions (\halocolor horizontal line
+in Figure \ref{fig:range_deltaphi}). However, we caution that the tendency of
 these midplane deviations is to \emph{inflate} the value of $J_z$ for thin

### 2019-02-26 paper/ms.tex
Remove old section comments

@@
 
-%%% this stuff is now in section 3 mostly
-%\subsection{Applicability to the Milky Way} \label{ssec:is_it_real}
-%As we briefly discussed in Section~\ref{ssec:cosmozoom}, cosmological zoom-in
-%simulations of Milky Way-mass galaxies are not perfect representations of the
-%Milky Way, and do not reproduce every available observation of the Milky Way.
-%For instance, the velocity dispersion-age relation in \mi , \mf , and \mm\ is
-%closer to that of M31 \citep{2018arXiv180610564S}.
-%
-%In order to assess the applicability of our work to the real Milky Way, we
-%consider a subset of Table~3 from \citet{2018arXiv180610564S} --- our
-%Table~\ref{tab:scale_height} shows the stellar scale heights of \mi , \mf ,
-%and \mm\ and the current best estimates for the Milky Way. One noticeable
-%feature is that the FIRE galaxies are more ``puffy'' than the Milky Way.
-%\Gus{Maybe FIRE collaboration has some insight into the mechanism for this?}.
-%This would seem to suggest that the midplane estimates from
-%Figure~\ref{fig:midplane} over-estimate the true variation in the Milky Way.
-
-%Another important aspect to consider when comparing the simulated galaxies to
-%the Milky Way is the environment/merger history. For instance, \mi\ and \mm\
-%both lack a massive neighbor analogous to the LMC. On the other hand, \mf\ has
-%had a recenter merger with \Gus{something}. \Gus{Expand more here - how to
-%incorporate merger trees?}
-
-%\subsection{Impact on Milky Way observations} \label{ssec:impact_on_mw}
 \subsection{The size of the Solar neighborhood} \label{sssec:neighborhood}
-
 Given a local coordinate system, the amplitude and characteristic length scale

### 2019-02-26 paper/ms.tex
Minor fixes

@@
 galaxy is $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the
-three cases the midplane therefore varies by more than $\pm 100$pc depending
-on the azimuth along the Solar circle; in the third (\mm{}, which has the
-thinnest ``thin disk'' of stars, but the largest stellar mass) the variation
-is more like $\pm 50$ pc.
+three cases the midplane therefore varies by more than $\pm 100\,\pc$
+depending on the azimuth along the Solar circle; in the third (\mm{}, which
+has the thinnest ``thin disk'' of stars, but the largest stellar mass) the
+variation is more like $\pm 50\,\pc$.
 
@@
 Since the variation in the LSR is less pronounced than for the midplane, and
-since the dominant effect is limited to $J_r$ (Figure
-\ref{fig:many_orbit_wrong_ref}), we defer this calculation to future work.
+since the dominant effect is limited to $J_r$
+(Figure~\ref{fig:many_orbit_wrong_ref}), we defer this calculation to future
+work.
 

### 2019-02-26 paper/ms.tex
Add contrast to chemical solar neighborhood

@@
 approximation, can be expected to be valid. This leads to a somewhat
-dynamically motivated definition of the ``solar neighborhood.''
+dynamically motivated definition of the ``solar neighborhood.'' Since the
+motivation may be different for defining the solar neighborhood in e.g.
+studies of the chemical distribution of the Galaxy, we propose the phrase
+``dynamical solar neighborhood.''
 

### 2019-02-26 paper/ms.tex
Update abstract

@@
 Stellar actions, computed for stars from their 6D phase space measurement and
-an assumed Galactic potential, are used to label orbits. In principle, actions
-have the attractive quality of being invariant and are thus good labels of a
-star's orbit. Of course, inaccurate measurements of the 6D phase space
-position and uncertainities in the assumed Galactic potential induce an error
-in the computed action. We show that, in addition to these complications, a
-systematic bias in the 6D phase space position induces a phase-dependence in
-the actions, which we interpret as an error. This error is non-Gaussian and
-bimodal, with neither of the modes peaking on the true action value. A
-midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is necessary to induce a
-$100\,\%$ error in the vertical action $J_z$ for thin and thick disk orbits,
-respectively. Furthermore, we show that the local midplane varies by
-$\sim100\,\pc$ at the Solar circle in cosmological zoom-in simulations from
-the FIRE collaboration. Thus, current state-of-the-art action calculations ---
-which assume the global and local midplanes are the same --- include a
-systematic offset in $z$. Variation in the local standard of rest will induce
-similar issues. The variation of the midplane must be taken into account when
-performing dynamical modelling across large regions of the disk accessible
-with the {\em Gaia} mission.
+an assumed Galactic potential, are used to label and distinguish orbits. In
+principle, actions have the attractive quality of being invariant and are thus
+good labels of a star's orbit. Of course, inaccurate measurements of the 6D
+phase space position and uncertainities in the assumed Galactic potential
+induce an error in the computed action. We show that, in addition to these
+complications, a systematic bias in the 6D phase space position induces a
+phase-dependence in the actions, which we interpret as an error. This error is
+non-Gaussian and bimodal, with neither of the modes peaking on the true action
+value. A midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is necessary to
+induce a $100\,\%$ error in the vertical action $J_z$ for thin and thick disk
+orbits, respectively. Furthermore, we show that the local midplane varies by
+$\sim100\,\pc$ at the Solar circle in Milky Way-mass cosmological zoom-in
+simulations from the FIRE collaboration. Thus, current state-of-the-art action
+calculations --- which assume the global and local midplanes are the same ---
+include a systematic offset in $z$. Variation in the local standard of rest
+will induce similar issues. The variation of the midplane must be taken into
+account when performing dynamical modelling across large regions of the disk
+accessible with the {\em Gaia} mission.
 
@@
 orbital phase in $z$ or $R$ is degenerate with the degree of systematic offset
-in that coordinate (see Figure \ref{fig:cartoon}). Additionally, in searching
+in that coordinate (see Figure~\ref{fig:cartoon}). Additionally, in searching
 for migrated members of known open clusters, we would compute actions for
@@
 compute the chord length $\lambda$ for the chord extending from the observer
-to $\pm \Delta\phi/2$ (see Figure \ref{fig:fig_to_explain}). We repeat this
+to $\pm \Delta\phi/2$ (see Figure~\ref{fig:fig_to_explain}). We repeat this
 for each of $50$ bins in azimuth. Each bin is plotted in

### 2019-02-26 paper/ms.tex
Expand a little on effect on Jz

@@
 in Figure \ref{fig:range_deltaphi}). However, we caution that the tendency of
-these midplane deviations is to \emph{inflate} the value of $J_z$ for thin
-disk stars in an extremely non-Gaussian way (Figure \ref{fig:Jz_hist}),
-scattering them into the region of action space normally associated with
-different stellar populations and possibly contributing to confusion there.
+these midplane deviations is to \emph{inflate} the value of $J_z$ whenever the
+midplane deviation is $\gtrsim z_{\text{max}}$, scattering them into the
+region of action space normally associated with different stellar populations
+and possibly contributing to confusion there. This inflation is extremely
+non-Gaussian (as the induced error is for all orbits, e.g.
+Figure~\ref{fig:Jz_hist}), and is particularly troubling for the case of thin
+disk stars, where the necessary midplane deviation is
+$\sim50\textup{--}100\,\pc$.
 

### 2019-02-26 paper/ms.tex
Refine statement about the Galactic potential

@@
 offset coordinate systems depend on the time (i.e. orbital phase) at which the
-star's phase-space position is observed. (To our extremely long-lived observer
+star's phase-space position is observed. To our extremely long-lived observer
 this would appear as time-dependence in the actions, which appear to oscillate
 around their value using the correct coordinate system even though in this
-example the Galaxy is described perfectly by a static, axisymmetric
-potential.) The relative size of the phase variation in each action depends on
-the direction of the systematic offset as well as the true values of the
-actions (i.e. the type of orbit). In reality we will have one measurement of
-the phase-space position to work with, in which case the determination of the
-orbital phase in $z$ or $R$ is degenerate with the degree of systematic offset
-in that coordinate (see Figure~\ref{fig:cartoon}). Additionally, in searching
-for migrated members of known open clusters, we would compute actions for
-stars with a wide range of orbital phases and locations in the Galaxy. In the
-following we therefore quote percentile ranges for the possible values
-computed for each action as a proxy for the effect of these systematic errors
-in the coordinate system.
+example the observer is using the correctly constructed best-fit static,
+axisymmetric potential. The relative size of the phase variation in each
+action depends on the direction of the systematic offset as well as the true
+values of the actions (i.e. the type of orbit). In reality we will have one
+measurement of the phase-space position to work with, in which case the
+determination of the orbital phase in $z$ or $R$ is degenerate with the degree
+of systematic offset in that coordinate (see Figure~\ref{fig:cartoon}).
+Additionally, in searching for migrated members of known open clusters, we
+would compute actions for stars with a wide range of orbital phases and
+locations in the Galaxy. In the following we therefore quote percentile ranges
+for the possible values computed for each action as a proxy for the effect of
+these systematic errors in the coordinate system.
 

### 2019-02-26 paper/ms.tex
revisions to 4.2

@@
 Suppose an open cluster of mass $m_c$ on an orbit with actions $\bm{J}_c$ is
-disrupted. We would like to determine all stars in the {\em Gaia}
-dataset\footnote{This is likely to be more successful with DR3/4 than DR2,
+disrupted. We would like to determine which stars with complete 6D phase-space positions, for example in the {\em Gaia}
+dataset,\footnote{This is likely to be more successful with DR3/4 than DR2,
 since upcoming data releases will contain a far greater number of stars with
 radial velocities --- though targeted, follow-up radial velocity surveys may
-allow this program to be performed sooner.} that are likely to be ejected
+allow this program to be performed sooner.} are likely to be ejected
 members from that cluster. One could generate an initial catalogue of
 candidate members by integrating the cluster's orbit and selecting stars close
-to that orbit, accounting for the fact that the stream will not exactly follow
+to that orbit, accounting for the fact that the stream of ejected stars will not exactly follow
 the orbit of the cluster \citep[e.g.][]{2011MNRAS.413.1852E}.
@@
 field star $i$ be close to the cluster's actions to within some bound, i.e.
-$\bm{J}_i - \bm{J}_c \in V_e$ where $V_e$ is some volume in action space
-enclosing a region of acceptable error. A natural choice for $V_e$ is to
-center it at the origin and allow its extent to be the intrinsic action-space
-dispersion of a disrupted cluster. This dispersion can be estimated as
-\citep[Section~8.3.3][]{2008gady.book.....B}
+$\bm{J}_i - \bm{J}_c \in V_e$, where $V_e$ is centered on $J_c$ and has size proportional to the intrinsic action-space dispersion of the disrupting cluster. This dispersion can be estimated as
+\citep[Section~8.3.3 of][]{2008gady.book.....B}
 \beq \label{eq:action_disp}
@@
 where $m_c$ is the mass of the cluster and $M$ is the mass enclosed by the
-cluster's orbit. A more sophisticated treatment of the action-space dispersion
-of disrupted clusters can be obtained by following
-\citet{2011MNRAS.413.1852E}, but such a treatment is premature for this work.
-
-The act of comparing a field star's action to a cluster's action (i.e. in
-computing $\bm{J}_i - \bm{J}_c$) is subject to significant error if the
-cluster and field star have significantly different local midplanes and an
-observer assumed a single global midplane.
-
-To explore the magnitude of this effect, we consider the following program.
-First, we assume a cluster mass $m_c$ and estimate the expected action-space
+cluster's orbit around the Galaxy, in this case assumed to be the mass inside the Solar circle ($\sim 10^{11}\Msun$). Stars with actions within $V_e$ would then be considered cluster members.
+
+%A more sophisticated treatment of the action-space dispersion
+%of disrupted clusters can be obtained by following
+%\citet{2011MNRAS.413.1852E}, but such a treatment is premature for this work.
+
+Two issues are immediately apparent with this approach.
+The first is that it neglects the overall smooth distribution \emph{in action space} which will contribute some stars to the selection by chance.
+For dissolving open clusters in the thin disk, this background is not likely to be negligible, but we will assume for the sake of argument that it can be modeled within the ``dynamical solar neighborhood'' where it is safe to extend the local coordinate system.
+Since the effect of extending the coordinate frame beyond this region is generally to inflate the actions computed for thin-disk stars, using too large a volume is likely to under-estimate the density of the smooth thin-disk-like component in action space, while scattering more thin-disk stars onto thick-disk-like orbits and over-estimating the density of the smooth thick-disk-like component.
+
+The second issue, which is related to the point of this paper, is that based on the results in our previous sections computing $\bm{J}_i$ for a field star and directly comparing it to  $\bm{J}_c$ will have significant systematic error if the cluster and field star are far enough apart to have significantly different local midplanes (see Figure \ref{fig:range_deltaphi}).
+Since our criterion for cluster membership (distance in action space) scales with the cluster mass, the region where the search will succeed varies with cluster mass as well.
+To illustrate this effect, we assume a cluster mass $m_c$ and estimate the expected action-space
 dispersion from Equation~\ref{eq:action_disp}. We then consult
-Figure~\ref{fig:many_orbit_wrong_ref} to determine the $z$ offset necessary in
+Figure~\ref{fig:many_orbit_wrong_ref} to determine the midplane $z$ offset necessary in
 order for the induced error in the vertical action to be the same as the
-expected action-space dispersion. If this were the case, an observer would
-have to account for the different midplane of the field star and the cluster.
-We perform this analysis for the thin-disk and thick-disk orbits considered in
-Section~\ref{ssec:quant}, since the halo orbit is unrealistic for known open
-clusters (Table~\ref{tab:real_clusters}).
-
-With a characteristic $z$ offset in hand, we consult
-Figure~\ref{fig:range_deltaphi} to convert this $z$ offset into a $\Delta
-\phi$, which we convert to an arclength or characteristic distance, which we
-refer to as the ``neighborhood radius'' or $R_{\n}$. Therefore, if an observer
-is attempting to reconstruct a cluster with a given mass on a given orbit, we
-will give a characteristic distance between a field star and the cluster
-within which an observer might not need to consider differences in their
-respective local midplanes.
+expected action-space dispersion. We perform this analysis for the thin-disk-like and thick-disk-like orbits considered in
+Section~\ref{ssec:quant}, assuming that the halo-like orbit is unlikely for members of known open
+clusters (see Table~\ref{tab:real_clusters}). The result for the allowable midplane offset for these two orbits in shown in Figure \ref{fig:cluster_offset} as a function of cluster mass.
+
+With a characteristic midplane offset in hand, we use the relation in Figure~\ref{fig:range_deltaphi}, which shows the typical midplane offset at each $\Delta \phi$, to convert a given midplane offset into a
+typical $\Delta\phi$.
+As in this Figure we assume the cluster is at the Solar circle to obtain a physical scale, which we refer to as the ``neighborhood radius'' or $R_{\n}$.
+If the allowable midplane discrepancy is greater than the maximum excursions recorded in the simulated galaxy, we simply report the maximum chord length assuming an orbit on the Solar circle ($2\times8.2\,\kpc$).
+ If an observer is attempting to reconstruct a cluster with a given mass on a given orbit, $R_{\n}$ gives the characteristic search distance between a field star and the cluster within which an observer may not need to consider differences in their respective local midplanes when computing actions.
+The result is shown in
+Figure~\ref{fig:Rn_mc}. The solid lines indicate the neighborhood values
+assuming the mean relation between midplane offset and $m_c$, and the dashed lines
+show the $+1\,\sigma$ and $-1\,\sigma$ relations based on the 50 fiducial Solar positions in each simulation (as discussed in Figure \ref{fig:range_deltaphi}).
+That this range is relatively wide illustrates the strong
+$\phi$-dependence of the midplane offsets.
 
@@
 
-In Figure~\ref{fig:cluster_offset}, we show the $z$ offset necessary to induce
-the action space dispersion computed from Equation~\ref{eq:action_disp}, using
-the \texttt{gala} computed mass enclosed within a radius of $8.2\,\kpc$ ($\sim
-10^{11}\,M_{\odot}$).
-
-We now convert the $z$~offset-$m_c$ relation to an $R_{\n}$-$m_c$ relation,
-where $R_{\n}$ indicates the neighborhood radius. An observer
-attempting to reconstruct an open cluster of mass $m_c$ with perfect knowledge
-of the midplane at the location of the cluster would expect to find the
-ejected members using a dynamical cut for field stars within a distance
-$R_{\n}$ of the cluster. If such an observer wants to find ejected stars that
-are at a distance greater than $R_{\n}$, a more sophisticated analysis would
-be required.
-
-We use the relation in Figure~\ref{fig:range_deltaphi}, which shows the
-typical $z$ offset at each $\Delta \phi$, to convert a given $z$~offset into a
-typical $\Delta\phi$. We then convert this $\Delta \phi$ into a chord length
-(see Figure~\ref{fig:fig_to_explain}) which we interpret as $R_{\n}$, the
-neighborhood around the cluster. If the $z$~offset is greater than the maximum
-range value of the galaxy, we simply report the maximum chord length
-($2\times8.2\,\kpc$). The result of this procedure is shown in
-Figure~\ref{fig:Rn_mc}. The solid lines indicate the neighborhood values
-assuming the mean relation between $z$~offset and $m_c$, and the dashed lines
-show the $+1\,\sigma$ and $-1\,\sigma$ relations, indicative of the strong
-$\phi$-dependence of the midplane effect.
-
-Figure~\ref{fig:Rn_mc} is most informative for the thin orbit (\thincolor).
+%In Figure~\ref{fig:cluster_offset}, we show the $z$ offset necessary to induce
+%the action space dispersion computed from Equation~\ref{eq:action_disp}, using
+%the \texttt{gala} computed mass enclosed within a radius of $8.2\,\kpc$ ($\sim
+%10^{11}\,M_{\odot}$).
+
+%We now convert the $z$~offset-$m_c$ relation to an $R_{\n}$-$m_c$ relation,
+%where $R_{\n}$ indicates the neighborhood radius. An observer
+%attempting to reconstruct an open cluster of mass $m_c$ with perfect knowledge
+%of the midplane at the location of the cluster would expect to find the
+%ejected members using a dynamical cut for field stars within a distance
+%$R_{\n}$ of the cluster. If such an observer wants to find ejected stars that
+%are at a distance greater than $R_{\n}$, a more sophisticated analysis would
+%be required.
+
+Figure~\ref{fig:Rn_mc} shows that azimuthal midplane fluctuations will limit the volume over which actions can be used to match stars to open clusters on thin-disk-like orbits (\thincolor), depending on their mass.
+For thick-disk-like orbits (\thickcolor), the tolerances are much more forgiving: stars' actions can be compared over the entire Solar circle for nearly all cluster masses, except at the very low-mass end in \mi .
 The neighborhood around $\sim100\Msun$ clusters can be as small as
@@
 with $m_c$, topping out at $\sim5\,\kpc$ for $\sim10^4\,\Msun$ clusters in
-\mi{} and \mf{}. In \mm{}, the neighborhood value reaches the maximum allowed of
-$16.4\,\kpc$ for clusters of mass $\sim800\,\Msun$. The interpretation of this
-is that the expected action space dispersion is so high for such high mass
-clusters that the measured $z$~offset is not large enough throughout the
-entire galaxy to be important.
-
-For the thick-disk orbit (brown), the neighborhood is at the maximum value for
-nearly all masses, except at the very low-mass end in \mi{}.
-
-We repeat this entire procedure for each of the orbits of the real open
-clusters in Table~\ref{tab:real_clusters} using the midplane variations from
-\mi{}. We use \mi{} as a conservative choice, since it provides the greatest
+\mi{} and \mf{}. In \mm{}, the neighborhood size reaches the maximum allowed of
+$16.4\,\kpc$ for clusters of mass $\sim800\,\Msun$.
+This figure also underlines that the degree to which such an analysis will succeed depends crucially on the expected degree of midplane fluctuations in the real Galaxy.
+If the Milky Way is more like \mi , the range of action-space analysis techniques like this is far more limited than if it more closely resembles \mm .
+
+We also compute $R_{\n}$ for each of the orbits of the real open
+clusters in Table~\ref{tab:real_clusters}, using the midplane variations from
+\mi{} as a conservative choice, since it has the largest degree of
 midplane variation (Figure~\ref{fig:midplane}). We report the neighborhood
@@
 neigbhorhood around each cluster is typically $\gtrsim 500\,\pc$. This
-analysis assumes that the midplane from \mi{} can be accurately applied to
-Milky Way orbits, which is not necessarily true. However, our analysis is
-meant to be more illustrative than quantitative and these neighborhood numbers
-should be understood as such.
+analysis assumes that the MW resembles \mi{}, which is not necessarily true.
+However, this step of our analysis is meant to be more illustrative than quantitative: this calculation should actually be performed using real measurements of variations in the MW's stellar midplane.
+Such measurements should be possible in the near future with Gaia, with quality distance estimates the crucial limiting factor.
 

### 2019-03-07 schmactions/many_orbits/plot_many_orbits.py
Print zoff for 100perc Jz error

@@ -10,6 +10,9 @@ import matplotlib.pyplot as plt
 from scipy.stats import sigmaclip
 import numpy as np
 
+from scipy.interpolate import interp1d
+from scipy.optimize import minimize
+
 import matplotlib as mpl
 from matplotlib import rc
 rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
@@ -55,6 +58,13 @@ init_vel_list = [[0, -190, 10] * u.km/u.s,
 
 clist = [tb_c[7], tb_c[8], tb_c[0]]
 
+def print_100(zerr, dJz):
+    interp = interp1d(zerr, dJz, bounds_error=False, fill_value='extrapolate')
+    def to_min(z):
+        return np.abs(interp(z)-1)
+    res = minimize(to_min, 250)
+    print(res.x)
+
 for fname, name, init_vel, c in zip(fname_list, name_list, init_vel_list, clist):
     zout = pickle.load(open('zout_'+fname+'.p', 'rb'))
     xout = pickle.load(open('xout_'+fname+'.p', 'rb'))
@@ -85,6 +95,10 @@ for fname, name, init_vel, c in zip(fname_list, name_list, init_vel_list, clist)
     dx[:,0] /= J0
     dx[:,1] /= J1
     dx[:,2] /= J2
+
+    print_100(zoffset, dz[:,2])
+    print(zoffset, dz[:,2])
+
     
     for x,o,d in zip(ax, (zoffset, xoffset), (dz, dx)):
         # get keys corresponding to 4 sigmaclip

### 2019-03-07 midplane_test/plot_deltaphi.py
Print dphi needed for zoff

@@ -3,6 +3,8 @@ from tqdm import tqdm
 import matplotlib.pyplot as plt
 from scipy.optimize import minimize
 
+from scipy.interpolate import interp1d
+
 from matplotlib import rc
 import matplotlib as mpl
 rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
@@ -22,7 +24,17 @@ Rsolar = 8.2
 
 glist = ['m12i', 'm12f', 'm12m']
 
+zoff_100_list = [30.7101346, 237.79862443]
+clist = [tb_c[7], tb_c[8]]
+
 theta = np.load('output/theta.npy')
+
+def print_dphi(dphi, rng):
+    interp = interp1d(dphi, rng)
+    def to_min(p):
+        return np.abs(interp(p)-zoff_100_list[0]/1000)
+    res = minimize(to_min, 0, bounds=[[0, 2.*np.pi]])
+    print(res.x)
 
 def rotate(l, n):
     return np.append(l[n:], l[:n])
@@ -76,6 +88,10 @@ for gal,ax_col in zip(glist, ax.transpose()):
 
     ax_col.set_ylim(0, 400)
 
+    for z, c in zip(zoff_100_list, clist):
+        ax_col.axhline(z, color=c)
+
+
     m = midplane_est - fit
 
     dphi_list_list, r_list_list = get_range_vs_dphi(theta, m)
@@ -90,6 +106,8 @@ for gal,ax_col in zip(glist, ax.transpose()):
     ax_col.plot(dphi_mean/np.pi, r_mean*1000, c=tb_c[0])
     ax_col.plot(dphi_mean/np.pi, r_up*1000, alpha=0.75, ls='dashed', c=tb_c[0])
     ax_col.plot(dphi_mean/np.pi, r_low*1000, alpha=0.75, ls='dashed', c=tb_c[0])
+
+    print_dphi(dphi_mean, r_mean)
 
     out = np.transpose([dphi_mean, r_mean, r_sigma])
     np.save('output/r_vs_dphi_'+gal+'.npy', out)

### 2019-03-07 paper/ms.tex
Update dphi fig caption
Add thin-disk solar neighb

@@
 $x$-axis shows the chord length from the position $\phi$ to $\pm\Delta\phi/2$.
+The horizontal \thincolor{} and \thickcolor{} lines shows the midplane variation necessary to induce a $100\%$ error in $J_z$ for the thin-disk and thick-disk orbits, respectively.
 A cartoon explanation of this Figure is given in
@@
 \ref{fig:range_deltaphi}). For the three simulations considered, the median
-range exceeds this value at $\Delta \phi >$ XX, YY, and ZZ radians, with
-corresponding physical scales of $\lambda \sim$ XX, YY, and ZZ parsecs. These
+range exceeds this value at $\Delta \phi/\pi > 0.036, 0.047, 0.10\,\text{rad}$, with
+corresponding physical scales of $\lambda \sim 465, 610, 1290\,\pc$ for m12i, m12f, and m12m, respectively. These
 can be considered estimates of the range of validity for an extension of the

### 2019-03-07 paper/ms.tex
Add thick-disk discussion to neighborhood

@@
 an $\mathcal{O}(1)$ systematic error in $J_z$ requires a larger midplane
-deviation (\thickcolor horizontal line in Figure \ref{fig:range_deltaphi})
-corresponding to $\Delta \phi >$ XX, YY, and ZZ radians or $\lambda \sim$ XX,
-YY, and ZZ parsecs. For stars on halo-like orbits, extending the local
+deviation (\thickcolor{} horizontal line in Figure~\ref{fig:range_deltaphi}).
+In fact, the range of midplane variation in m12i just barely saturates to cause such an error in $J_z$, with the midplane variation in m12f and m12m never reaching such a value.
+For stars on halo-like orbits, extending the local
 coordinate system globally is not likely to cause serious problems, since the
 midplane deviations are relatively small compared to those needed to produce
-$\mathcal{O}(1)$ systematic errors in the actions (\halocolor horizontal line
+$\mathcal{O}(1)$ systematic errors in the actions (\halocolor{} horizontal line
 in Figure \ref{fig:range_deltaphi}). However, we caution that the tendency of

### 2019-03-07 paper/ms.tex
Fix line indentation

@@
 $x$-axis shows the chord length from the position $\phi$ to $\pm\Delta\phi/2$.
-The horizontal \thincolor{} and \thickcolor{} lines shows the midplane variation necessary to induce a $100\%$ error in $J_z$ for the thin-disk and thick-disk orbits, respectively.
-A cartoon explanation of this Figure is given in
-Figure~\ref{fig:fig_to_explain}.}
+The horizontal \thincolor{} and \thickcolor{} lines shows the midplane
+variation necessary to induce a $100\%$ error in $J_z$ for the thin-disk and
+thick-disk orbits, respectively. A cartoon explanation of this Figure is given
+in Figure~\ref{fig:fig_to_explain}.}
 \label{fig:range_deltaphi}
@@
 \ref{fig:range_deltaphi}). For the three simulations considered, the median
-range exceeds this value at $\Delta \phi/\pi > 0.036, 0.047, 0.10\,\text{rad}$, with
-corresponding physical scales of $\lambda \sim 465, 610, 1290\,\pc$ for m12i, m12f, and m12m, respectively. These
-can be considered estimates of the range of validity for an extension of the
-locally determined axisymmetric coordinate system, at least when it comes to
-the study of stars on quasi-circular orbits.
+range exceeds this value at $\Delta \phi/\pi > 0.036, 0.047,
+0.10\,\text{rad}$, with corresponding physical scales of $\lambda \sim 465,
+610, 1290\,\pc$ for m12i, m12f, and m12m, respectively. These can be
+considered estimates of the range of validity for an extension of the locally
+determined axisymmetric coordinate system, at least when it comes to the study
+of stars on quasi-circular orbits.
 
@@
 deviation (\thickcolor{} horizontal line in Figure~\ref{fig:range_deltaphi}).
-In fact, the range of midplane variation in m12i just barely saturates to cause such an error in $J_z$, with the midplane variation in m12f and m12m never reaching such a value.
-For stars on halo-like orbits, extending the local
-coordinate system globally is not likely to cause serious problems, since the
-midplane deviations are relatively small compared to those needed to produce
-$\mathcal{O}(1)$ systematic errors in the actions (\halocolor{} horizontal line
-in Figure \ref{fig:range_deltaphi}). However, we caution that the tendency of
-these midplane deviations is to \emph{inflate} the value of $J_z$ whenever the
-midplane deviation is $\gtrsim z_{\text{max}}$, scattering them into the
-region of action space normally associated with different stellar populations
-and possibly contributing to confusion there. This inflation is extremely
-non-Gaussian (as the induced error is for all orbits, e.g.
-Figure~\ref{fig:Jz_hist}), and is particularly troubling for the case of thin
-disk stars, where the necessary midplane deviation is
+In fact, the range of midplane variation in m12i just barely saturates to
+cause such an error in $J_z$, with the midplane variation in m12f and m12m
+never reaching such a value. For stars on halo-like orbits, extending the
+local coordinate system globally is not likely to cause serious problems,
+since the midplane deviations are relatively small compared to those needed to
+produce $\mathcal{O}(1)$ systematic errors in the actions (\halocolor{}
+horizontal line in Figure \ref{fig:range_deltaphi}). However, we caution that
+the tendency of these midplane deviations is to \emph{inflate} the value of
+$J_z$ whenever the midplane deviation is $\gtrsim z_{\text{max}}$, scattering
+them into the region of action space normally associated with different
+stellar populations and possibly contributing to confusion there. This
+inflation is extremely non-Gaussian (as the induced error is for all orbits,
+e.g. Figure~\ref{fig:Jz_hist}), and is particularly troubling for the case of
+thin disk stars, where the necessary midplane deviation is
 $\sim50\textup{--}100\,\pc$.
@@
 Suppose an open cluster of mass $m_c$ on an orbit with actions $\bm{J}_c$ is
-disrupted. We would like to determine which stars with complete 6D phase-space positions, for example in the {\em Gaia}
-dataset,\footnote{This is likely to be more successful with DR3/4 than DR2,
-since upcoming data releases will contain a far greater number of stars with
-radial velocities --- though targeted, follow-up radial velocity surveys may
-allow this program to be performed sooner.} are likely to be ejected
-members from that cluster. One could generate an initial catalogue of
-candidate members by integrating the cluster's orbit and selecting stars close
-to that orbit, accounting for the fact that the stream of ejected stars will not exactly follow
-the orbit of the cluster \citep[e.g.][]{2011MNRAS.413.1852E}.
+disrupted. We would like to determine which stars with complete 6D phase-space
+positions, for example in the {\em Gaia} dataset,\footnote{This is likely to
+be more successful with DR3/4 than DR2, since upcoming data releases will
+contain a far greater number of stars with radial velocities --- though
+targeted, follow-up radial velocity surveys may allow this program to be
+performed sooner.} are likely to be ejected members from that cluster. One
+could generate an initial catalogue of candidate members by integrating the
+cluster's orbit and selecting stars close to that orbit, accounting for the
+fact that the stream of ejected stars will not exactly follow the orbit of the
+cluster \citep[e.g.][]{2011MNRAS.413.1852E}.
 
@@
 field star $i$ be close to the cluster's actions to within some bound, i.e.
-$\bm{J}_i - \bm{J}_c \in V_e$, where $V_e$ is centered on $J_c$ and has size proportional to the intrinsic action-space dispersion of the disrupting cluster. This dispersion can be estimated as
-\citep[Section~8.3.3 of][]{2008gady.book.....B}
+$\bm{J}_i - \bm{J}_c \in V_e$, where $V_e$ is centered on $J_c$ and has size
+proportional to the intrinsic action-space dispersion of the disrupting
+cluster. This dispersion can be estimated as \citep[Section~8.3.3
+of][]{2008gady.book.....B}
 \beq \label{eq:action_disp}
@@
 where $m_c$ is the mass of the cluster and $M$ is the mass enclosed by the
-cluster's orbit around the Galaxy, in this case assumed to be the mass inside the Solar circle ($\sim 10^{11}\Msun$). Stars with actions within $V_e$ would then be considered cluster members.
-
-%A more sophisticated treatment of the action-space dispersion
-%of disrupted clusters can be obtained by following
-%\citet{2011MNRAS.413.1852E}, but such a treatment is premature for this work.
-
-Two issues are immediately apparent with this approach.
-The first is that it neglects the overall smooth distribution \emph{in action space} which will contribute some stars to the selection by chance.
-For dissolving open clusters in the thin disk, this background is not likely to be negligible, but we will assume for the sake of argument that it can be modeled within the ``dynamical solar neighborhood'' where it is safe to extend the local coordinate system.
-Since the effect of extending the coordinate frame beyond this region is generally to inflate the actions computed for thin-disk stars, using too large a volume is likely to under-estimate the density of the smooth thin-disk-like component in action space, while scattering more thin-disk stars onto thick-disk-like orbits and over-estimating the density of the smooth thick-disk-like component.
-
-The second issue, which is related to the point of this paper, is that based on the results in our previous sections computing $\bm{J}_i$ for a field star and directly comparing it to  $\bm{J}_c$ will have significant systematic error if the cluster and field star are far enough apart to have significantly different local midplanes (see Figure \ref{fig:range_deltaphi}).
-Since our criterion for cluster membership (distance in action space) scales with the cluster mass, the region where the search will succeed varies with cluster mass as well.
-To illustrate this effect, we assume a cluster mass $m_c$ and estimate the expected action-space
-dispersion from Equation~\ref{eq:action_disp}. We then consult
-Figure~\ref{fig:many_orbit_wrong_ref} to determine the midplane $z$ offset necessary in
-order for the induced error in the vertical action to be the same as the
-expected action-space dispersion. We perform this analysis for the thin-disk-like and thick-disk-like orbits considered in
-Section~\ref{ssec:quant}, assuming that the halo-like orbit is unlikely for members of known open
-clusters (see Table~\ref{tab:real_clusters}). The result for the allowable midplane offset for these two orbits in shown in Figure \ref{fig:cluster_offset} as a function of cluster mass.
-
-With a characteristic midplane offset in hand, we use the relation in Figure~\ref{fig:range_deltaphi}, which shows the typical midplane offset at each $\Delta \phi$, to convert a given midplane offset into a
-typical $\Delta\phi$.
-As in this Figure we assume the cluster is at the Solar circle to obtain a physical scale, which we refer to as the ``neighborhood radius'' or $R_{\n}$.
-If the allowable midplane discrepancy is greater than the maximum excursions recorded in the simulated galaxy, we simply report the maximum chord length assuming an orbit on the Solar circle ($2\times8.2\,\kpc$).
- If an observer is attempting to reconstruct a cluster with a given mass on a given orbit, $R_{\n}$ gives the characteristic search distance between a field star and the cluster within which an observer may not need to consider differences in their respective local midplanes when computing actions.
-The result is shown in
-Figure~\ref{fig:Rn_mc}. The solid lines indicate the neighborhood values
-assuming the mean relation between midplane offset and $m_c$, and the dashed lines
-show the $+1\,\sigma$ and $-1\,\sigma$ relations based on the 50 fiducial Solar positions in each simulation (as discussed in Figure \ref{fig:range_deltaphi}).
-That this range is relatively wide illustrates the strong
-$\phi$-dependence of the midplane offsets.
+cluster's orbit around the Galaxy, in this case assumed to be the mass inside
+the Solar circle ($\sim 10^{11}\Msun$). Stars with actions within $V_e$ would
+then be considered cluster members.
+
+Two issues are immediately apparent with this approach. The first is that it
+neglects the overall smooth distribution \emph{in action space} which will
+contribute some stars to the selection by chance. For dissolving open clusters
+in the thin disk, this background is not likely to be negligible, but we will
+assume for the sake of argument that it can be modeled within the ``dynamical
+solar neighborhood'' where it is safe to extend the local coordinate system.
+Since the effect of extending the coordinate frame beyond this region is
+generally to inflate the actions computed for thin-disk stars, using too large
+a volume is likely to under-estimate the density of the smooth thin-disk-like
+component in action space, while scattering more thin-disk stars onto
+thick-disk-like orbits and over-estimating the density of the smooth
+thick-disk-like component.
+
+The second issue, which is related to the point of this paper, is that based
+on the results in our previous sections computing $\bm{J}_i$ for a field star
+and directly comparing it to  $\bm{J}_c$ will have significant systematic
+error if the cluster and field star are far enough apart to have significantly
+different local midplanes (see Figure \ref{fig:range_deltaphi}). Since our
+criterion for cluster membership (distance in action space) scales with the
+cluster mass, the region where the search will succeed varies with cluster
+mass as well. To illustrate this effect, we assume a cluster mass $m_c$ and
+estimate the expected action-space dispersion from
+Equation~\ref{eq:action_disp}. We then consult
+Figure~\ref{fig:many_orbit_wrong_ref} to determine the midplane $z$ offset
+necessary in order for the induced error in the vertical action to be the same
+as the expected action-space dispersion. We perform this analysis for the
+thin-disk-like and thick-disk-like orbits considered in
+Section~\ref{ssec:quant}, assuming that the halo-like orbit is unlikely for
+members of known open clusters (see Table~\ref{tab:real_clusters}). The result
+for the allowable midplane offset for these two orbits in shown in Figure
+\ref{fig:cluster_offset} as a function of cluster mass.
+
+With a characteristic midplane offset in hand, we use the relation in
+Figure~\ref{fig:range_deltaphi}, which shows the typical midplane offset at
+each $\Delta \phi$, to convert a given midplane offset into a typical
+$\Delta\phi$. As in this Figure we assume the cluster is at the Solar circle
+to obtain a physical scale, which we refer to as the ``neighborhood radius''
+or $R_{\n}$. If the allowable midplane discrepancy is greater than the maximum
+excursions recorded in the simulated galaxy, we simply report the maximum
+chord length assuming an orbit on the Solar circle ($2\times8.2\,\kpc$). If an
+observer is attempting to reconstruct a cluster with a given mass on a given
+orbit, $R_{\n}$ gives the characteristic search distance between a field star
+and the cluster within which an observer may not need to consider differences
+in their respective local midplanes when computing actions. The result is
+shown in Figure~\ref{fig:Rn_mc}. The solid lines indicate the neighborhood
+values assuming the mean relation between midplane offset and $m_c$, and the
+dashed lines show the $+1\,\sigma$ and $-1\,\sigma$ relations based on the 50
+fiducial Solar positions in each simulation (as discussed in Figure
+\ref{fig:range_deltaphi}). That this range is relatively wide illustrates the
+strong $\phi$-dependence of the midplane offsets.
 
@@
 
-Figure~\ref{fig:Rn_mc} shows that azimuthal midplane fluctuations will limit the volume over which actions can be used to match stars to open clusters on thin-disk-like orbits (\thincolor), depending on their mass.
-For thick-disk-like orbits (\thickcolor), the tolerances are much more forgiving: stars' actions can be compared over the entire Solar circle for nearly all cluster masses, except at the very low-mass end in \mi .
-The neighborhood around $\sim100\Msun$ clusters can be as small as
-$\sim300\,\pc$ but up to $\sim1.2\,\kpc$ for \mi{} and \mf{}, though the
-situation is marginally better for \mm{} with values ranging from
-$\sim0.8\,\kpc$ to $\sim3\,\kpc$. The neighborhood values steadily increase
-with $m_c$, topping out at $\sim5\,\kpc$ for $\sim10^4\,\Msun$ clusters in
+Figure~\ref{fig:Rn_mc} shows that azimuthal midplane fluctuations will limit
+the volume over which actions can be used to match stars to open clusters on
+thin-disk-like orbits (\thincolor), depending on their mass. For
+thick-disk-like orbits (\thickcolor), the tolerances are much more forgiving:
+stars' actions can be compared over the entire Solar circle for nearly all
+cluster masses, except at the very low-mass end in \mi . The neighborhood
+around $\sim100\Msun$ clusters can be as small as $\sim300\,\pc$ but up to
+$\sim1.2\,\kpc$ for \mi{} and \mf{}, though the situation is marginally better
+for \mm{} with values ranging from $\sim0.8\,\kpc$ to $\sim3\,\kpc$. The
+neighborhood values steadily increase with $m_c$, topping out at $\sim5\,\kpc$
+for $\sim10^4\,\Msun$ clusters in
 \mi{} and \mf{}. In \mm{}, the neighborhood size reaches the maximum allowed of
-$16.4\,\kpc$ for clusters of mass $\sim800\,\Msun$.
-This figure also underlines that the degree to which such an analysis will succeed depends crucially on the expected degree of midplane fluctuations in the real Galaxy.
-If the Milky Way is more like \mi , the range of action-space analysis techniques like this is far more limited than if it more closely resembles \mm .
-
-We also compute $R_{\n}$ for each of the orbits of the real open
-clusters in Table~\ref{tab:real_clusters}, using the midplane variations from
+$16.4\,\kpc$ for clusters of mass $\sim800\,\Msun$. This figure also
+underlines that the degree to which such an analysis will succeed depends
+crucially on the expected degree of midplane fluctuations in the real Galaxy.
+If the Milky Way is more like \mi , the range of action-space analysis
+techniques like this is far more limited than if it more closely resembles \mm
+.
+
+We also compute $R_{\n}$ for each of the orbits of the real open clusters in
+Table~\ref{tab:real_clusters}, using the midplane variations from
 \mi{} as a conservative choice, since it has the largest degree of
@@
 analysis assumes that the MW resembles \mi{}, which is not necessarily true.
-However, this step of our analysis is meant to be more illustrative than quantitative: this calculation should actually be performed using real measurements of variations in the MW's stellar midplane.
-Such measurements should be possible in the near future with Gaia, with quality distance estimates the crucial limiting factor.
+However, this step of our analysis is meant to be more illustrative than
+quantitative: this calculation should actually be performed using real
+measurements of variations in the MW's stellar midplane. Such measurements
+should be possible in the near future with Gaia, with quality distance
+estimates the crucial limiting factor.
 

### 2019-03-07 paper/ms.tex
Small edits

@@
 required for an $\mathcal{O}(1)$ systematic error in $J_z$ for our fiducial
-thin-disk-like orbit (\thincolor horizontal line in Figure
+thin-disk-like orbit (\thincolor{} horizontal line in Figure
 \ref{fig:range_deltaphi}). For the three simulations considered, the median
@@
 0.10\,\text{rad}$, with corresponding physical scales of $\lambda \sim 465,
-610, 1290\,\pc$ for m12i, m12f, and m12m, respectively. These can be
+610, 1290\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. These can be
 considered estimates of the range of validity for an extension of the locally

### 2019-03-18 paper/ms.tex
Small text fixes

@@
 95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
-$6.2 \actunit$ for $J_r$ and $J_z$, respectively. These are fractional errors
-of $5.7\%$ and $85.7\%$, respectively. The error induced in $L_z$ is
+$6.2\,\actunit$ for $J_r$ and $J_z$, respectively. These are fractional errors
+of $5.7\,\%$ and $85.7\,\%$, respectively. The error induced in $L_z$ is
 negligible, as expected since $L_z$ is independent of $z$. It is worth

### 2019-03-18 schmactions/many_orbits/plot_many_orbits_Jzfun.py
Update figure a little bit

@@ -22,8 +22,8 @@ mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
         '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
 
-xmin = 0
-xmax = 100
+xmin = 0.08
+xmax = 200
 
 ymin = 0
 ymax = 100
@@ -100,13 +100,13 @@ for n,c in zip(names, colors):
 
 ax.set_xscale('log')
 
-# ax.set_xlim([xmin, xmax])
+ax.set_xlim([xmin, xmax])
 ax.set_ylim([ymin, ymax])
 
 ax.set_xlabel(r'$J_z\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
 ax.set_ylabel(r'$\Delta J_z/J_z\,[\,\%\,]$')
 
-ax.legend()
+ax.legend(frameon=False, title=r'$z\,\text{offset}\,[\,\text{pc}\,]$')
 
 fig.tight_layout()
 plt.savefig('schmactions_many_orbits_Jz_fun.pdf')

### 2019-03-18 paper/ms.tex
Add dJz fun Jz plot to paper

@@
 \begin{figure}
+\plotone{fig/schmactions_many_orbits_Jz_fun.pdf}
+\caption{The fractional error in $J_z$ as a function of $J_z$ for a few
+different offsets in $z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
+(red, cyan, green respectively). As before, the error ($\Delta J_z$) is the
+95\uth{} minus 5\uth{} percentile of the distribution of observed $J_z$ values
+over the course of the orbit. There are large errors for thin-disk like orbits
+($J_z \sim 0.5\,\actunit$), even for a small midplane offset of $10\,\pc$. }
+\label{fig:dJz_fun_Jz}
+\end{figure}
+
+\begin{figure}
 \plotone{fig/schmactions_Jz_hist.pdf}
@@
 \end{figure}
+
+To further understand the effect of the midplane error, we also plot the
+fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of $10, 50,
+100\,\pc$ (red, cyan, green, respectively) in Figure~\ref{fig:dJz_fun_Jz}. For
+a thin disk-like orbit ($J_z\sim0.5\,\actunit$), even a $10\,\pc$ offset in
+$z$ is enough to induce a $\sim50\%$ error in $J_z$. For larger values of
+$J_z$, the fractional errors are suppressed, with the situation becoming worse
+the greater the $z$ offset is.
 

### 2019-03-18 paper/ms.tex
Change ref to m12i/f/m

@@
 \tablenotetext{d}{The azimuthally averaged gas vertical density profile in
-m12i is nearly constant to this height, though individual regions show smaller
+\mi{} is nearly constant to this height, though individual regions show smaller
 scale heights and dense clouds.}
@@
 deviation (\thickcolor{} horizontal line in Figure~\ref{fig:range_deltaphi}).
-In fact, the range of midplane variation in m12i just barely saturates to
-cause such an error in $J_z$, with the midplane variation in m12f and m12m
+In fact, the range of midplane variation in \mi{} just barely saturates to
+cause such an error in $J_z$, with the midplane variation in \mf{} and \mm{}
 never reaching such a value. For stars on halo-like orbits, extending the

### 2019-03-18 paper/ms.tex
Small fixes
Small changes to table comparing MW and simulation values, and minor updates to references

--- 2019-03-18 paper/ms.tex
@@
 95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
-$6.2\,\actunit$ for $J_r$ and $J_z$, respectively. These are fractional errors
-of $5.7\,\%$ and $85.7\,\%$, respectively. The error induced in $L_z$ is
+$6.2 \actunit$ for $J_r$ and $J_z$, respectively. These are fractional errors
+of $5.7\%$ and $85.7\%$, respectively. The error induced in $L_z$ is
 negligible, as expected since $L_z$ is independent of $z$. It is worth
@@
 \startdata
-MW\tablenotemark{b} & 40 & 300 & 900 & 0 & 0 \\
+MW\tablenotemark{b} & 40 & 300 & 900 & \nodata & \nodata \\
 \mi\tablenotemark{c} & 800\tablenotemark{d} & 480 & 2000 & 12.4 & 11.2 \\

### 2019-03-18 paper/ms.tex
first revisions to start of introduction

--- 2019-03-18 paper/ms.tex
@@
 \citep{2018ApJ...860L..11K,2018Natur.563...85H,2018arXiv180704290L,2019MNRAS.482.3426M},
-a phase-space spiral \citep{2018Natur.561..360A} indicating either local
+a phase-space ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly indicating either local
 substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L} or bar
-buckling \citep{2018arXiv181109205K}, a gap indicative of a dark matter
-subtructure in GD1 \citep{2018arXiv181103631B}, among others.
-
-These discoveries are all indicative that the Milky Way departs from
-axisymmetry and is undergoing phase mixing and dynamical interactions across a
-range of spatial and temporal scales. Underlying the quantitative description
-of the mechanisms at work that give rise to these signatures is the assumption
-of a Galactocentric coordinate system \citep{2008gady.book.....B} which can
-both precisely and accurately be defined and measured. In angular coordinates,
-the center of the galaxy is taken to be the location of
+buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation by a dark matter
+subtructure in the tidal stream GD1 \citep{2018arXiv181103631B}.
+These discoveries all indicate that the Milky Way's stellar distribution, which demonstrably departs significantly from axisymmetry, is undergoing phase mixing and dynamical interactions across a range of spatial and temporal scales.
+
+Underlying the quantitative analysis
+of many of the mechanisms at work that give rise to these signatures is the assumption of a global, axisymmetric Galactocentric coordinate system \citep{2008gady.book.....B} in which the Sun's position (and therefore the origin of the coordinate system) can be defined and measured both precisely and accurately.
+In this system,
+the center of the galaxy is usually taken to be the location of the central supermassive black hole,
 Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From stellar
-motions near Sgr~A\textsuperscript{*}, the distance to the center is taken to
+motions near Sgr~A\textsuperscript{*}, the distance from the Sun to the Galactic center is taken to
 be $\sim 8.1\text{--}8.3\,\kpc$
-\citep{2009ApJ...692.1075G,2018AA...615L..15G}. The vertical height is found
-by identifying where the stellar positions and velocities reach a maximum
+\citep{2009ApJ...692.1075G,2018AA...615L..15G}.
+The Sun's vertical distance from the Galactic midplane, the other independent measurement that sets this coordinate system, can be determined by identifying where the stellar density and velocities reach a maximum
 (effectively the median height of all disk stars), and is usually taken to be
 $\sim 25\,\pc$ \citep{2001ApJ...553..184C}, with more recent measurements from
-{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}. A
+{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}.
+Another strategy is to use the cold gas or HII regions in the disk to define the Galactic midplane, leading to slightly values ($\sim 5\,\pc$) for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}.
+A
 pre-{\em Gaia} review of these measurements is given by
@@
 Once a Galactocentric coordinate system has been established and a 6D phase
-space measurement of a star has been made, it is desirable to convert this
-measurement into action space. Actions are quantities which describe the orbit
-of a star and are conserved quantities under some assumptions which we
-describe next. They are the typical integral of the Hamiltonian in cylindrical
+space measurement of a star has been made, it is often desirable to convert this
+measurement into action space to concisely summarize its projected orbit, model the stellar distribution function, or find stars with similar dynamical properties.
+Actions are conserved quantities which describe the orbit
+of a star under the assumption of regular, bound orbits in a system where the equations of motion are separable in a particular coordinate system.
+They are the typical integral of the Hamiltonian in cylindrical
 coordinates and are given by,
@@
 If the Milky Way is more like \mi , the range of action-space analysis
-techniques like this is far more limited than if it more closely resembles \mm
-.
+techniques like this is far more limited than if it more closely resembles
+\mm{}.
 

### 2019-03-21 paper/ms.tex
Add ref to new paper

@@
 a phase-space ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly indicating either local
-substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L} or bar
-buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation by a dark matter
+substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L}, bar
+buckling \citep{2018arXiv181109205K}, although may just be contributions from moving groups \citep{2019arXiv190308325M} and a gap suggestive of perturbation by a dark matter
 subtructure in the tidal stream GD1 \citep{2018arXiv181103631B}.

### 2019-03-25 paper/ms.tex
Change title
Add info about Sag sim

@@
 \shortauthors{Beane et al.}
-\shorttitle{Our Galactic Midplane}
+\shorttitle{The Midplane is Not a Plane}
 
@@
 
-\title{Issues with Dynamical Measurements Across the Disk}
+\title{The Midplane is Not a Plane: Implications for {\em Gaia}}
 
@@
 
+\subsection{Description of Sagittarius Simulaton} \label{ssec:sag_sim}
+In addition to the cosmological zoom-in, we will also briefly consider results
+from a toy Sagittarius encounter simulation. This simulation offers us the
+ability to see how the midplane varies in a more controlled environment. The
+details of the simulation are given in \citet{2018MNRAS.481..286L}, but we
+briefly summarize the most relevant details here. For the Milky Way, the dark
+halo is modelled as a Hernquist sphere of mass $10^{12}\,\Msun$ and scale
+length of $52\,\kpc$, the disk modelled as an exponential disk with a scale
+radius of $3.5\,\kpc$, scale height $0.53\,\kpc$, and mass
+$6\times10^{10}\Msun$, a Hernquist bulge of mass $10^{10}\,\Msun$ and scale
+length $0.7\,\kpc$ \citep{1990ApJ...356..359H}. Sagittarius is modelled as a
+dark matter Hernquist sphere of mass \Gus{something} and scale length
+\Gus{something}, along with a stellar component modelled as a Hernquist sphere
+of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$.
+
+For the disk and bulge components, a softening length of $30\,\pc$ is used
+whereas for the halo a softening length of $60\,\pc$ is used. For Sagittarius,
+the softening length for the dark matter and the stars is $60$ and $40\,\pc$,
+respectively.
+
 \subsection{Local Midplane} \label{ssec:local_midplane}

### 2019-03-25 paper/ms.tex
Add chervin sim fig to paper

@@
 
+\begin{figure*}
+\plotone{fig/midplane_fit_chervinsim.pdf}
+\caption{The local midplane determined at the fiducial Solar radius
+($8.2\,\kpc$) for 4 different time steps from previously run toy simulations
+of a Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
+before, we have subtracted the best fit $A\sin{(\phi+B)}+C$ curve to account
+for inaccuracies in the galactocentric coordinate system. Error bars are
+calculated as in Figure~\ref{fig:midplane}. (\Gus{Guessing here, need to
+figure out what the panels actually show:}) The left two panels show the
+midplane as a function of azimuth before the first encounter. As later
+encounters occur, we can see that stronger midplane variations result. In
+fact, one can see that an $m=2$ mode is excited in the third panel --- this is
+consistent with Figure~17 of \citet{2018MNRAS.481..286L}, which shows that an
+$m=2$ mode is excited (note that $m=0$ and $m=1$ modes are stronger, but these
+are removed in our sine-curve subtraction).}
+\label{fig:midplane_chervin}
+\end{figure*}
+
 \subsection{LSR Variations} \label{ssec:lsr_var}

### 2019-03-25 paper/ms.tex
Add text for chervin midplane fig

@@
 
+We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
+but for the toy simulation of a Sagittarius encounter
+\citep{2018MNRAS.481..286L}. Again we have subtracted a best fit curve of the
+form $\text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}$, with the values
+for $\text{A}$, $\text{B}$, and $\text{C}$ being \Gus{something},
+\Gus{something}, and \Gus{something else}, respectively. In the first two
+panels, we see that the midplane is relatively flat in the inner galaxy. But
+additional encounters drive strong midplane variation. In the third panel, we
+see a strong $m=2$ mode develop, consistent with the $R=8\,\kpc$ panel of
+Figure~17 in \citet{2018MNRAS.481..286L}.
+
 \begin{figure*}

### 2019-03-25 paper/ms.tex
Change subsection title

@@
 FIRE} \label{sec:local_fire}
-\subsection{Description of Simulation} \label{ssec:cosmozoom}
+\subsection{Description of FIRE Simulation} \label{ssec:cosmozoom}
 We briefly describe the cosmological-hydrodynamical zoom-in simulations used

### 2019-03-25 paper/ms.tex
Fix indentation throughout

@@
 era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
-major merger
-\citep{2018ApJ...860L..11K,2018Natur.563...85H,2018arXiv180704290L,2019MNRAS.482.3426M},
-a phase-space ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly indicating either local
-substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L}, bar
-buckling \citep{2018arXiv181109205K}, although may just be contributions from moving groups \citep{2019arXiv190308325M} and a gap suggestive of perturbation by a dark matter
-subtructure in the tidal stream GD1 \citep{2018arXiv181103631B}.
-These discoveries all indicate that the Milky Way's stellar distribution, which demonstrably departs significantly from axisymmetry, is undergoing phase mixing and dynamical interactions across a range of spatial and temporal scales.
-
-Underlying the quantitative analysis
-of many of the mechanisms at work that give rise to these signatures is the assumption of a global, axisymmetric Galactocentric coordinate system \citep{2008gady.book.....B} in which the Sun's position (and therefore the origin of the coordinate system) can be defined and measured both precisely and accurately.
-In this system,
-the center of the galaxy is usually taken to be the location of the central supermassive black hole,
-Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From stellar
-motions near Sgr~A\textsuperscript{*}, the distance from the Sun to the Galactic center is taken to
-be $\sim 8.1\text{--}8.3\,\kpc$
-\citep{2009ApJ...692.1075G,2018AA...615L..15G}.
-The Sun's vertical distance from the Galactic midplane, the other independent measurement that sets this coordinate system, can be determined by identifying where the stellar density and velocities reach a maximum
-(effectively the median height of all disk stars), and is usually taken to be
-$\sim 25\,\pc$ \citep{2001ApJ...553..184C}, with more recent measurements from
-{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}.
-Another strategy is to use the cold gas or HII regions in the disk to define the Galactic midplane, leading to slightly values ($\sim 5\,\pc$) for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}.
-A
-pre-{\em Gaia} review of these measurements is given by
-\citet{2016ARAA..54..529B}.
+major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
+2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the
+solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
+local substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L}, bar
+buckling \citep{2018arXiv181109205K}, although may just be contributions from
+moving groups \citep{2019arXiv190308325M} and a gap suggestive of perturbation
+by a dark matter subtructure in the tidal stream GD1
+\citep{2018arXiv181103631B}. These discoveries all indicate that the Milky
+Way's stellar distribution, which demonstrably departs significantly from
+axisymmetry, is undergoing phase mixing and dynamical interactions across a
+range of spatial and temporal scales.
+
+Underlying the quantitative analysis of many of the mechanisms at work that
+give rise to these signatures is the assumption of a global, axisymmetric
+Galactocentric coordinate system \citep{2008gady.book.....B} in which the
+Sun's position (and therefore the origin of the coordinate system) can be
+defined and measured both precisely and accurately. In this system, the center
+of the galaxy is usually taken to be the location of the central supermassive
+black hole, Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From
+stellar motions near Sgr~A\textsuperscript{*}, the distance from the Sun to
+the Galactic center is taken to be $\sim 8.1\text{--}8.3\,\kpc$
+\citep{2009ApJ...692.1075G,2018AA...615L..15G}. The Sun's vertical distance
+from the Galactic midplane, the other independent measurement that sets this
+coordinate system, can be determined by identifying where the stellar density
+and velocities reach a maximum (effectively the median height of all disk
+stars), and is usually taken to be $\sim 25\,\pc$ \citep{2001ApJ...553..184C},
+with more recent measurements from {\em Gaia} DR2 placing it at $\sim 20\,\pc$
+\citep{2019MNRAS.482.1417B}. Another strategy is to use the cold gas or HII
+regions in the disk to define the Galactic midplane, leading to slightly
+values ($\sim 5\,\pc$) for the Sun's relative height
+\citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review of these
+measurements is given by \citet{2016ARAA..54..529B}.
 
 Once a Galactocentric coordinate system has been established and a 6D phase
-space measurement of a star has been made, it is often desirable to convert this
-measurement into action space to concisely summarize its projected orbit, model the stellar distribution function, or find stars with similar dynamical properties.
-Actions are conserved quantities which describe the orbit
-of a star under the assumption of regular, bound orbits in a system where the equations of motion are separable in a particular coordinate system.
-They are the typical integral of the Hamiltonian in cylindrical
-coordinates and are given by,
+space measurement of a star has been made, it is often desirable to convert
+this measurement into action space to concisely summarize its projected orbit,
+model the stellar distribution function, or find stars with similar dynamical
+properties. Actions are conserved quantities which describe the orbit of a
+star under the assumption of regular, bound orbits in a system where the
+equations of motion are separable in a particular coordinate system. They are
+the typical integral of the Hamiltonian in cylindrical coordinates and are
+given by,
 \beq\label{eq:actions}
@@
 the density-based midplane varies as a function of azimuth. The Milky Way's
-dense, star-forming gas disk is thought to have a scale height of about $40\,\pc$
-($3\text{-}4$ times the cold gas softening length; \citealt{2019ApJ...871..145A}) and
-the thin stellar disk a scale height of about 300 pc (30 times the stellar
-softening length; \citealt{2008ApJ...673..864J}). We therefore expect that
-resolution effects may still be affecting the scale heights of these
-components in the simulations, especially the cold gas. Indeed, the stellar
-scale heights of the simulated galaxies are equal to or larger than the MW's
-while the gas scale heights are significantly larger (although the proper
-basis comparison is less clear in the case of the gas; the quoted value for
-the Milky Way comes from studies of high-mass star-forming regions). As noted
-above, the midplanes defined by gas and stars can be tilted with respect to
-one another as well, precluding extending the precision of the gas midplane
-definition to the stellar component. In this work we specifically consider
-azimuthal variations in the midplane \emph{as defined by the stellar mass
-density}, which will subsequently be lower in the simulations than in the MW:
-within 200 pc of the Sun the measured gas and stellar volume densities in the
-MW are roughly comparable at $\sim .04\ \Msun\
+dense, star-forming gas disk is thought to have a scale height of about
+$40\,\pc$ ($3\text{-}4$ times the cold gas softening length;
+\citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of
+about 300 pc (30 times the stellar softening length;
+\citealt{2008ApJ...673..864J}). We therefore expect that resolution effects
+may still be affecting the scale heights of these components in the
+simulations, especially the cold gas. Indeed, the stellar scale heights of the
+simulated galaxies are equal to or larger than the MW's while the gas scale
+heights are significantly larger (although the proper basis comparison is less
+clear in the case of the gas; the quoted value for the Milky Way comes from
+studies of high-mass star-forming regions). As noted above, the midplanes
+defined by gas and stars can be tilted with respect to one another as well,
+precluding extending the precision of the gas midplane definition to the
+stellar component. In this work we specifically consider azimuthal variations
+in the midplane \emph{as defined by the stellar mass density}, which will
+subsequently be lower in the simulations than in the MW: within 200 pc of the
+Sun the measured gas and stellar volume densities in the MW are roughly
+comparable at $\sim .04\ \Msun\
 \unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
@@
 using the new height $10$ times to converge on the local midplane height. In
-order to allow for the possibility that the fiducial Galactocentric coordinate
+order to allow for the possibility that the fiducial galactocentric coordinate
 system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
@@
 
-%%%move this later - decide where, probably conclusions
-%While we focus our discussion now solely on the midplane, there is an
-%additional complication in the measurement of the distance from the Sun to the
-%Galactic center that may also be important. When modelling stars at Galactic
-%radii close to the Sun's radius, it is in principle possible to determine what
-%point they are orbiting about: the ``dynamical Galactic
-%center.'' It is not guaranteed that this dynamical center will coincide
-%with the distance to Sgr~A\textsuperscript{*} \citep{2018AA...615L..15G}. In
-%fact, \citet{2015ApJ...803...80K} were able to measure the distance of the Sun
-%to the Galactic center using the dynamical information of Palomar 5. Other
-%dynamical measurements of the Solar distance have been made, though none with
-%a precision capable of competing with the distance to Sgr~A\textsuperscript{*}
+%%%move this later - decide where, probably conclusions While we focus our
+%discussion now solely on the midplane, there is an additional complication in
+%the measurement of the distance from the Sun to the Galactic center that may
+%also be important. When modelling stars at Galactic radii close to the Sun's
+%radius, it is in principle possible to determine what point they are orbiting
+%about: the ``dynamical Galactic center.'' It is not guaranteed that this
+%dynamical center will coincide with the distance to Sgr~A\textsuperscript{*}
+%\citep{2018AA...615L..15G}. In fact, \citet{2015ApJ...803...80K} were able to
+%measure the distance of the Sun to the Galactic center using the dynamical
+%information of Palomar 5. Other dynamical measurements of the Solar distance
+%have been made, though none with a precision capable of competing with the
+%distance to Sgr~A\textsuperscript{*}
 %\citep{1981gask.book.....M,2011PASJ...63..867S,2012MNRAS.427..274S,2013AstL...39...95B,2013IAUS..289..444Z}.

### 2019-04-06 paper/ms.tex
more revisions to the introduction

--- 2019-03-25 paper/ms.tex
@@
 \shortauthors{Beane et al.}
-\shorttitle{The Midplane is Not a Plane}
+\shorttitle{Our Galactic Midplane}
 
@@
 
-\title{The Midplane is Not a Plane: Implications for {\em Gaia}}
+\title{Issues with Dynamical Measurements Across the Disk}
 
@@
 a phase-space ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly indicating either local
-substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L}, bar
-buckling \citep{2018arXiv181109205K}, although may just be contributions from moving groups \citep{2019arXiv190308325M} and a gap suggestive of perturbation by a dark matter
+substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L} or bar
+buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation by a dark matter
 subtructure in the tidal stream GD1 \citep{2018arXiv181103631B}.
@@
 Once a Galactocentric coordinate system has been established and a 6D phase
-space measurement of a star has been made, it is often desirable to convert this
-measurement into action space to concisely summarize its projected orbit, model the stellar distribution function, or find stars with similar dynamical properties.
-Actions are conserved quantities which describe the orbit
-of a star under the assumption of regular, bound orbits in a system where the equations of motion are separable in a particular coordinate system.
-They are the typical integral of the Hamiltonian in cylindrical
-coordinates and are given by,
+space measurement of a star has been made, it is often desirable to calculate canonical invariants, such as actions, from that measurement.
+Actions are conserved quantities which describe the orbit of a star under the assumption of regular, bound orbits in a system where the equations of motion are separable in a coordinate system that reflects the system's underlying symmetries.
+In cylindrical coordinates, which correspond to the axisymmetry commonly assumed for the Galactic disk, the are given by
 \beq\label{eq:actions}
@@
 \eeq
-where $i=R,z,\phi$.\footnote{We typically write $J_r$ instead of $J_R$ for
-aesthetic reasons.} Note that $J_{\phi} \equiv L_z$. In a slowly-evolving
-axisymmetric potential, the actions of stars' orbits are conserved quantities
-\citep{2008gady.book.....B,2014RvMP...86....1S}. The non-axisymmetries of a
-more realistic potential induce time evolution in the {\em measured} actions
---- the true orbital invariants (if they exist) are invariant by definition.
-
-In a slowly-evolving, axisymmetric potential, the actions of each star are
-time-invariant \citep{2008gady.book.....B}. In the Milky Way, stellar actions
-are expected to evolve on short time scales due to scattering with gas clouds
-and on long time scales due to resonances with spiral arms, bar(s), and other
+where $i=R,z,\phi$ are the cylindrical radial, vertical, and azimuthal coordinates respectively and the $p_i$ are the conjugate momenta.\footnote{We typically write $J_r$ instead of $J_R$ for
+aesthetic reasons.} In a slowly-evolving
+axisymmetric potential, these actions are invariant and $J_{\phi} \equiv L_z$
+\citep{2008gady.book.....B,2014RvMP...86....1S}.
+
+With the advent of 6D phase-space measurements over a relatively large ($\gtrsim 2$ kpc) volume from the {\em Gaia} satellite, the study of stars' actions has gained new popularity.
+One reason is dimensionality reduction---an individual stellar orbit is concisely described by three actions, as opposed to six dimensions of phase space.
+Under the assumption of a phase-mixed system, the dynamical properties of a population of stars should be uniquely a function of their actions and independent of the conjugate angles, allowing investigation of the
+relationship between {\em orbital} properties of stars and other intrinsic, and at least partially invariant, properties such as age or metallicity.
+Actions also provide a convenient basis for constructing models of the stellar distribution function, or for associating stars with similar dynamical properties (e.g.
+to determine membership in moving groups).
+
+If the system being considered departs from axisymmetry in a significant and non-adiabatic way, the actions computed using an axisymmetric approximation to the true potential can exhibit cyclic dependence on the orbital phase (or time at which the star's position and velocity are observed), large-scale migration, or diffusion from their initial values.
+In the Milky Way, stellar actions are expected to diffuse on short time scales due to scattering with gas clouds
+and to evolve on longer time scales in the case of orbits near resonances with spiral arms, bar(s), and other
 large scale perturbations \citep{2014RvMP...86....1S}. For this reason, we and
@@
 been used to study different models of spiral structure in the Milky Way
-\citep{2019MNRAS.tmp..155S}. A summary of the disk in action space is given by
+\citep{2019MNRAS.tmp..155S}. Characteristics of the distribution of stars in the extended Solar neighborhood in action space are discussed in
 \citet{2018arXiv180503653T}.
 
-The transformation to action space is not trivial, and many of the assumptions
-of the transformation are potentially incorrect. For instance, the assumption
-that the Galactic potential is axisymmetric may be inadequate with current
-measurement errors. Even if this assumption is valid, the parameters used in
-our axisymmetric models may be inadequately tuned to observations. In
-addition, the necessity of 6D phase space measurements of a star limits the
-Galactic volume observable in action space.
-
-With these limitations in mind, it is therefore natural to consider the reason
-to transform to action space. One reason is dimensionality reduction --- in
-the action space paradigm, a stellar orbit is completely described by three
-actions, as opposed to six dimensions of phase space. This means that the
-relationship between {\em orbital} properties of individual stars and stellar
-properties such as age or metallicity can be investigated.
-
-We explore one science goal that relies on the conservation of actions --- the
-reconstruction of open clusters across large distances using dynamical
-tagging. The basic program is to match stars with roughly the same actions as
-their supposed natal cluster. To set the scene, we consider the orbits of the
+The transformation to action space is nonlinear in both the phase-space position and the potential, and requires relatively strong assumptions about the system's symmetry. For instance, the 
+Galactic potential is not strictly axisymmetric; with the vast improvement in the quality of phase-space measurements due to {\em Gaia}, this assumption appears increasingly inadequate \citep[e.g.][]{2018Natur.561..360A}.
+Even if this assumption is close enough for many purposes, the parameters used in axisymmetric models may be inadequately constrained by current observations.
+
+%In addition, the necessity of 6D phase space measurements of a star limits the Galactic volume observable in action space.
+
+%With these limitations in mind, it is therefore natural to consider the reason to transform to action space.
+
+In this work we discuss how the imposition of a global axisymmetric coordinate system based on a local measurement of the midplane location and Galactocentric distance introduces systematic error in the computation of actions, especially for stars in the thin disk.
+Our main point
+is that the ``local'' Galactic midplane is not the same as the ``global''
+Galactic midplane. For stars relatively far from the Sun,
+the local midplane location will not agree with
+our local midplane extrapolated onto their position, thanks to relatively small-scale variations in the density of the gas and stellar distribution, not necessarily even as large as spiral arms.
+Converting positions and velocities of more distant stars from a heliocentric to a
+Galactocentric coordinate system introduces a systematic bias in the $z$ coordinate relative to $z$ velocity, and hence the actions computed for these stars.
+The further from the solar neighborhood the target star, the more likely the mismatch will result in large systematic uncertainty, especially in the vertical action $J_z$.
+Depending on the type of orbit being studied, this sets a limit on the ``dynamical solar neighborhood'' over which the local coordinate frame can be usefully extended for action-space analysis.
+A similar argument applies to uncertainty in measurements of
+$R_{\odot}$ and variations in the local standard of rest (LSR). In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
+system errors have on the measured actions.
+
+Lacking an empirical, nonlocal definition of the global Galactic midplane or a measurement of the expected midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire} we attempt to quantify the azimuthal dependence of the
+local midplane, and hence the expected size of the dynamical solar neighborhood, using cosmological-hydrodynamical zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic Environments (FIRE)
+collaboration\footnote{\url{https://fire.northwestern.edu}} \citep{2016ApJ...827L..23W,2017arXiv170206148H}.
+In Section \ref{sssec:neighborhood} we use these estimates to consider, as an example, the feasibility of one science goal that relies on the invariance of actions --- the reconstruction of open clusters across large distances using ``dynamical tagging;'' that is, the association of stars to a birth cluster based on their similar actions.
+In light of our findings, we considered the orbits of the
 nine open clusters within $250\,\pc$ as reported by the {\em Gaia}
@@
 report the distance to each cluster, the three actions, and the maximum
-vertical extent of their orbits ($z_{\text{max}}$). The actions are computed
-using the technique described in Section~\ref{ssec:action_comp}. We also show
-the mass of each cluster and the ``neighborhood'' we compute for each cluster,
-which gives a rough sense of how far one can move away from the cluster before
-the issue which we discuss next becomes important. This neighborhood is
-discussed in more detail in Section~\ref{sssec:reconstruction}.
+vertical extent of their orbits ($z_{\text{max}}$). 
+%The actions are computed using the technique described in Section~\ref{ssec:action_comp}. 
+ In Section~\ref{sssec:reconstruction} we estimate and discuss the size of the dynamical solar neighborhood for each cluster based on its mass, to show that action-space tagging is likely still a viable method if background is unimportant.
+In Section~\ref{sec:discussion} we discuss further implications for action-space analyses beyond the dynamical solar neighborhood.
+We summarize and conclude in Section~\ref{sec:conclusion}.
 
@@
 \end{deluxetable*}
-
-We propose a new type of error in the computation of actions which has
-implications for general dynamical modelling across the disk. Our main point
-is that the ``local'' Galactic midplane is not the same as the ``global''
-Galactic midplane. If we observe a group of stars that are
-non-local\footnote{i.e. a group of stars that is far enough away. The exact
-distance depends on the accuracy of the measurement being made and the goals
-of the particular problem.}, they will know of a different local midplane than
-our local midplane extrapolated onto their position. This means that when we
-convert those star's positions and velocities from a heliocentric to a
-Galactocentric coordinate system, we are inducing a systematic bias in those
-stars' $z$ values. The same applies to inaccuracies in measurements of
-$R_{\odot}$ and variations in the local standard of rest (LSR).
-
-We do not propose what a modeller should use for the ``true'' global midplane,
-since such a definition probably depends on the particular problem such a
-modeller is facing. Even without a high-precision definition and measurement
-of the global Galactic midplane, we can still assess how different we should
-expect the local midplane to be by measuring the azimuthal dependence of the
-local midplane. This procedure may be readily tractable over some of the disk
-with current datasets, using appropriate selection functions. However, we turn
-to simulation in this work because we can investigate the entire disk.
-Specifically, we use cosmological zoom-ins of Milky Way-mass galaxies from the
-Feedback in Realistic Environments (FIRE)
-collaboration\footnote{\url{https://fire.northwestern.edu}} (the {\em Latte}
-suite of simulations).
-
-Our main point is that the local midplane varies between different points in
-the Galaxy. This is a different statement than that our measurements of the
-local midplane are erroneous. Indeed, the local midplane has recently been
-measured to a claimed high level of precision, setting $z_{\odot} = 20.8 \pm
-0.3\,\pc$ \citep{2019MNRAS.482.1417B}. However, we also do not mean to say
-that determinations of the local midplane are perfect, since midplane
-measurements from H~\textsc{ii} regions place $z_{\odot} \sim 5-10\,\pc$. An
-additional complication is the fact that non-zero Solar heights induce a tilt
-in the Galactic midplane \citep{2014ApJ...797...53G,2016ARAA..54..529B}.
-
-In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
-system errors have on the measured actions. In Section~\ref{sec:local_fire},
-we measure the local midplane in cosmological zoom-in simulations from the
-FIRE collaboration. In Section~\ref{sec:discussion} we discuss and give some
-context for our results, and implications for the definition of the Solar
-neighborhood. We conclude in Section~\ref{sec:conclusion}.
 
@@
 
-\subsection{Description of Sagittarius Simulaton} \label{ssec:sag_sim}
-In addition to the cosmological zoom-in, we will also briefly consider results
-from a toy Sagittarius encounter simulation. This simulation offers us the
-ability to see how the midplane varies in a more controlled environment. The
-details of the simulation are given in \citet{2018MNRAS.481..286L}, but we
-briefly summarize the most relevant details here. For the Milky Way, the dark
-halo is modelled as a Hernquist sphere of mass $10^{12}\,\Msun$ and scale
-length of $52\,\kpc$, the disk modelled as an exponential disk with a scale
-radius of $3.5\,\kpc$, scale height $0.53\,\kpc$, and mass
-$6\times10^{10}\Msun$, a Hernquist bulge of mass $10^{10}\,\Msun$ and scale
-length $0.7\,\kpc$ \citep{1990ApJ...356..359H}. Sagittarius is modelled as a
-dark matter Hernquist sphere of mass \Gus{something} and scale length
-\Gus{something}, along with a stellar component modelled as a Hernquist sphere
-of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$.
-
-For the disk and bulge components, a softening length of $30\,\pc$ is used
-whereas for the halo a softening length of $60\,\pc$ is used. For Sagittarius,
-the softening length for the dark matter and the stars is $60$ and $40\,\pc$,
-respectively.
-
 \subsection{Local Midplane} \label{ssec:local_midplane}
@@
 
-We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
-but for the toy simulation of a Sagittarius encounter
-\citep{2018MNRAS.481..286L}. Again we have subtracted a best fit curve of the
-form $\text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}$, with the values
-for $\text{A}$, $\text{B}$, and $\text{C}$ being \Gus{something},
-\Gus{something}, and \Gus{something else}, respectively. In the first two
-panels, we see that the midplane is relatively flat in the inner galaxy. But
-additional encounters drive strong midplane variation. In the third panel, we
-see a strong $m=2$ mode develop, consistent with the $R=8\,\kpc$ panel of
-Figure~17 in \citet{2018MNRAS.481..286L}.
-
 \begin{figure*}
@@
 \label{fig:midplane}
-\end{figure*}
-
-\begin{figure*}
-\plotone{fig/midplane_fit_chervinsim.pdf}
-\caption{The local midplane determined at the fiducial Solar radius
-($8.2\,\kpc$) for 4 different time steps from previously run toy simulations
-of a Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
-before, we have subtracted the best fit $A\sin{(\phi+B)}+C$ curve to account
-for inaccuracies in the galactocentric coordinate system. Error bars are
-calculated as in Figure~\ref{fig:midplane}. (\Gus{Guessing here, need to
-figure out what the panels actually show:}) The left two panels show the
-midplane as a function of azimuth before the first encounter. As later
-encounters occur, we can see that stronger midplane variations result. In
-fact, one can see that an $m=2$ mode is excited in the third panel --- this is
-consistent with Figure~17 of \citet{2018MNRAS.481..286L}, which shows that an
-$m=2$ mode is excited (note that $m=0$ and $m=1$ modes are stronger, but these
-are removed in our sine-curve subtraction).}
-\label{fig:midplane_chervin}
 \end{figure*}
@@
 accurately and precisely enough, then the computed actions are invariant with
-orbital phase. However, we hvae shown that the fact that the Galactic midplane
+orbital phase. However, we have shown that the fact that the Galactic midplane
 is not constant across the disk presents a significant complication to
@@
 essentially equivalent.
+
+Add sentences about distribution function modeling
+
+Our main point is that the local midplane varies between different points in
+the Galaxy. This is a different statement than that our measurements of the
+local midplane are erroneous. Indeed, the local midplane has recently been
+measured to a claimed high level of precision, setting $z_{\odot} = 20.8 \pm
+0.3\,\pc$ \citep{2019MNRAS.482.1417B}. However, we also do not mean to say
+that determinations of the local midplane are perfect, since midplane
+measurements from H~\textsc{ii} regions place $z_{\odot} \sim 5-10\,\pc$. An
+additional complication is the fact that non-zero Solar heights induce a tilt
+in the Galactic midplane \citep{2014ApJ...797...53G,2016ARAA..54..529B}.
+We do not propose here what a modeler should use for the ``true'' global midplane, since it depends on the particular problem being studied.
+Even without a high-precision definition of the global Galactic midplane, current observations from Gaia should soon permit a measurement of the real azimuthal dependence of the midplane location.
 

### 2019-04-06 paper/ms.tex
final revisions to the introduction

@@
 collaboration\footnote{\url{https://fire.northwestern.edu}} \citep{2016ApJ...827L..23W,2017arXiv170206148H}.
-In Section \ref{sssec:neighborhood} we use these estimates to consider, as an example, the feasibility of one science goal that relies on the invariance of actions --- the reconstruction of open clusters across large distances using ``dynamical tagging;'' that is, the association of stars to a birth cluster based on their similar actions.
+
+In Section~\ref{sec:discussion} we discuss the implications of midplane variations, and the resulting systematic uncertainty in the vertical action, for action-space analyses beyond the dynamical solar neighborhood.
+As an illustration, in Section \ref{sssec:neighborhood} we use these estimates to consider the feasibility of one science goal that relies on the invariance of actions --- the reconstruction of open clusters across large distances using ``dynamical tagging;'' that is, the association of stars to a birth cluster based on their similar actions.
 In light of our findings, we considered the orbits of the
@@
  In Section~\ref{sssec:reconstruction} we estimate and discuss the size of the dynamical solar neighborhood for each cluster based on its mass, to show that action-space tagging is likely still a viable method if background is unimportant.
-In Section~\ref{sec:discussion} we discuss further implications for action-space analyses beyond the dynamical solar neighborhood.
 We summarize and conclude in Section~\ref{sec:conclusion}.

### 2019-04-08 paper/ms.tex

@@
 
-Underlying the quantitative analysis
-of many of the mechanisms at work that give rise to these signatures is the assumption of a global, axisymmetric Galactocentric coordinate system \citep{2008gady.book.....B} in which the Sun's position (and therefore the origin of the coordinate system) can be defined and measured both precisely and accurately.
-In this system,
-the center of the galaxy is usually taken to be the location of the central supermassive black hole,
-Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From stellar
-motions near Sgr~A\textsuperscript{*}, the distance from the Sun to the Galactic center is taken to
-be $\sim 8.1\text{--}8.3\,\kpc$
-\citep{2009ApJ...692.1075G,2018AA...615L..15G}.
-The Sun's vertical distance from the Galactic midplane, the other independent measurement that sets this coordinate system, can be determined by identifying where the stellar density and velocities reach a maximum
-(effectively the median height of all disk stars), and is usually taken to be
-$\sim 25\,\pc$ \citep{2001ApJ...553..184C}, with more recent measurements from
-{\em Gaia} DR2 placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}.
-Another strategy is to use the cold gas or HII regions in the disk to define the Galactic midplane, leading to slightly values ($\sim 5\,\pc$) for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}.
-A
-pre-{\em Gaia} review of these measurements is given by
-\citet{2016ARAA..54..529B}.
+Underlying the quantitative analysis of many of the mechanisms at work that
+give rise to these signatures is the assumption of a global, axisymmetric
+Galactocentric coordinate system \citep{2008gady.book.....B} in which the
+Sun's position (and therefore the origin of the coordinate system) can be
+defined and measured both precisely and accurately. In this system, the center
+of the galaxy is usually taken to be the location of the central supermassive
+black hole, Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From
+stellar motions near Sgr~A\textsuperscript{*}, the distance from the Sun to
+the Galactic center is taken to be $\sim 8.1\text{--}8.3\,\kpc$
+\citep{2009ApJ...692.1075G,2018AA...615L..15G}. The Sun's vertical distance
+from the Galactic midplane, the other independent measurement that sets this
+coordinate system, can be determined by identifying where the stellar density
+and velocities reach a maximum (effectively the median height of all disk
+stars), and is usually taken to be $\sim 25\,\pc$ \citep{2001ApJ...553..184C},
+with more recent measurements from {\em Gaia} DR2 placing it at $\sim 20\,\pc$
+\citep{2019MNRAS.482.1417B}. Another strategy is to use the cold gas or HII
+regions in the disk to define the Galactic midplane, leading to slightly
+values ($\sim 5\,\pc$) for the Sun's relative height
+\citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review of these
+measurements is given by \citet{2016ARAA..54..529B}.
 
 Once a Galactocentric coordinate system has been established and a 6D phase
-space measurement of a star has been made, it is often desirable to calculate canonical invariants, such as actions, from that measurement.
-Actions are conserved quantities which describe the orbit of a star under the assumption of regular, bound orbits in a system where the equations of motion are separable in a coordinate system that reflects the system's underlying symmetries.
-In cylindrical coordinates, which correspond to the axisymmetry commonly assumed for the Galactic disk, the are given by
+space measurement of a star has been made, it is often desirable to convert
+this measurement into action space to concisely summarize its projected orbit,
+model the stellar distribution function, or find stars with similar dynamical
+properties. Actions are conserved quantities which describe the orbit of a
+star under the assumption of regular, bound orbits in a system where the
+equations of motion are separable in a particular coordinate system. They are
+the typical integral of the Hamiltonian in cylindrical coordinates and are
+given by,
 \beq\label{eq:actions}
@@
 $x$--$y$ plane.
+
+\subsection{Description of Sagittarius Simulaton} \label{ssec:sag_sim}
+In addition to the cosmological zoom-in, we will also briefly consider results
+from a toy Sagittarius encounter simulation. This simulation offers us the
+ability to see how the midplane varies in a more controlled environment. The
+details of the simulation are given in \citet{2018MNRAS.481..286L}, but we
+briefly summarize the most relevant details here. For the Milky Way, the dark
+halo is modelled as a Hernquist sphere of mass $10^{12}\,\Msun$ and scale
+length of $52\,\kpc$, the disk modelled as an exponential disk with a scale
+radius of $3.5\,\kpc$, scale height $0.53\,\kpc$, and mass
+$6\times10^{10}\Msun$, a Hernquist bulge of mass $10^{10}\,\Msun$ and scale
+length $0.7\,\kpc$ \citep{1990ApJ...356..359H}. Sagittarius is modelled as a
+dark matter Hernquist sphere of mass \Gus{something} and scale length
+\Gus{something}, along with a stellar component modelled as a Hernquist sphere
+of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$.
+
+For the disk and bulge components, a softening length of $30\,\pc$ is used
+whereas for the halo a softening length of $60\,\pc$ is used. For Sagittarius,
+the softening length for the dark matter and the stars is $60$ and $40\,\pc$,
+respectively.
 
@@
 \label{fig:midplane}
+\end{figure*}
+
+\begin{figure*}
+\plotone{fig/midplane_fit_chervinsim.pdf}
+\caption{The local midplane determined at the fiducial Solar radius
+($8.2\,\kpc$) for 4 different time steps from previously run toy simulations
+of a Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
+before, we have subtracted the best fit $A\sin{(\phi+B)}+C$ curve to account
+for inaccuracies in the galactocentric coordinate system. Error bars are
+calculated as in Figure~\ref{fig:midplane}. (\Gus{Guessing here, need to
+figure out what the panels actually show:}) The left two panels show the
+midplane as a function of azimuth before the first encounter. As later
+encounters occur, we can see that stronger midplane variations result. In
+fact, one can see that an $m=2$ mode is excited in the third panel --- this is
+consistent with Figure~17 of \citet{2018MNRAS.481..286L}, which shows that an
+$m=2$ mode is excited (note that $m=0$ and $m=1$ modes are stronger, but these
+are removed in our sine-curve subtraction).}
+\label{fig:midplane_chervin}
 \end{figure*}

### 2019-04-08 paper/ms.tex
new title
title & abstract edits

@@
 
-\title{The Midplane is Not a Plane: Implications for {\em Gaia}}
+\title{The Galactic Midplane is Not a Plane: Implications for {\em Gaia}}
 
@@
 
-Stellar actions, computed for stars from their 6D phase space measurement and
+Stellar actions, computed for stars from their six-dimensional phase space measurement and
 an assumed Galactic potential, are used to label and distinguish orbits. In
-principle, actions have the attractive quality of being invariant and are thus
-good labels of a star's orbit. Of course, inaccurate measurements of the 6D
+principle, actions have the attractive quality of being invariant and thus
+good labels of a star's orbit. Of course, inaccurate measurements of the 
 phase space position and uncertainities in the assumed Galactic potential
 induce an error in the computed action. We show that, in addition to these
-complications, a systematic bias in the 6D phase space position induces a
+complications, a systematic bias in the phase space position induces a
 phase-dependence in the actions, which we interpret as an error. This error is
-non-Gaussian and bimodal, with neither of the modes peaking on the true action
-value. A midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is necessary to
-induce a $100\,\%$ error in the vertical action $J_z$ for thin and thick disk
+non-Gaussian and bimodal, 
+%mm [this sounds imprecise: an error wouldn't peak on a value...]
+with neither of the modes peaking on the true action value. 
+A midplane error of $\sim30\,\pc$ for a thin disk orbit or
+$\sim225\,\pc$ for a thick disk orbit is all that is necessary to
+induce a $100\,\%$ error in the vertical action $J_z$.
 orbits, respectively. Furthermore, we show that the local midplane varies by
@@
 era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
-major merger
-\citep{2018ApJ...860L..11K,2018Natur.563...85H,2018arXiv180704290L,2019MNRAS.482.3426M},
-a phase-space ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly indicating either local
-substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L} or bar
-buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation by a dark matter
-subtructure in the tidal stream GD1 \citep{2018arXiv181103631B}.
-These discoveries all indicate that the Milky Way's stellar distribution, which demonstrably departs significantly from axisymmetry, is undergoing phase mixing and dynamical interactions across a range of spatial and temporal scales.
+major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
+2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the
+solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
+local substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L}, bar
+buckling \citep{2018arXiv181109205K}, although may just be contributions from
+moving groups \citep{2019arXiv190308325M} and a gap suggestive of perturbation
+by a dark matter subtructure in the tidal stream GD1
+\citep{2018arXiv181103631B}. These discoveries all indicate that the Milky
+Way's stellar distribution, which demonstrably departs significantly from
+axisymmetry, is undergoing phase mixing and dynamical interactions across a
+range of spatial and temporal scales.
 
@@
 \eeq
-where $i=R,z,\phi$ are the cylindrical radial, vertical, and azimuthal coordinates respectively and the $p_i$ are the conjugate momenta.\footnote{We typically write $J_r$ instead of $J_R$ for
-aesthetic reasons.} In a slowly-evolving
-axisymmetric potential, these actions are invariant and $J_{\phi} \equiv L_z$
-\citep{2008gady.book.....B,2014RvMP...86....1S}.
-
-With the advent of 6D phase-space measurements over a relatively large ($\gtrsim 2$ kpc) volume from the {\em Gaia} satellite, the study of stars' actions has gained new popularity.
-One reason is dimensionality reduction---an individual stellar orbit is concisely described by three actions, as opposed to six dimensions of phase space.
-Under the assumption of a phase-mixed system, the dynamical properties of a population of stars should be uniquely a function of their actions and independent of the conjugate angles, allowing investigation of the
-relationship between {\em orbital} properties of stars and other intrinsic, and at least partially invariant, properties such as age or metallicity.
-Actions also provide a convenient basis for constructing models of the stellar distribution function, or for associating stars with similar dynamical properties (e.g.
-to determine membership in moving groups).
-
-If the system being considered departs from axisymmetry in a significant and non-adiabatic way, the actions computed using an axisymmetric approximation to the true potential can exhibit cyclic dependence on the orbital phase (or time at which the star's position and velocity are observed), large-scale migration, or diffusion from their initial values.
-In the Milky Way, stellar actions are expected to diffuse on short time scales due to scattering with gas clouds
-and to evolve on longer time scales in the case of orbits near resonances with spiral arms, bar(s), and other
+where $i=R,z,\phi$.\footnote{We typically write $J_r$ instead of $J_R$ for
+aesthetic reasons.} Note that $J_{\phi} \equiv L_z$. In a slowly-evolving
+axisymmetric potential, the actions of stars' orbits are conserved quantities
+\citep{2008gady.book.....B,2014RvMP...86....1S}. The non-axisymmetries of a
+more realistic potential induce time evolution in the {\em measured} actions
+--- the true orbital invariants (if they exist) are invariant by definition.
+
+In a slowly-evolving, axisymmetric potential, the actions of each star are
+time-invariant \citep{2008gady.book.....B}. In the Milky Way, stellar actions
+are expected to evolve on short time scales due to scattering with gas clouds
+and on long time scales due to resonances with spiral arms, bar(s), and other
 large scale perturbations \citep{2014RvMP...86....1S}. For this reason, we and
@@
 been used to study different models of spiral structure in the Milky Way
-\citep{2019MNRAS.tmp..155S}. Characteristics of the distribution of stars in the extended Solar neighborhood in action space are discussed in
+\citep{2019MNRAS.tmp..155S}. A summary of the disk in action space is given by
 \citet{2018arXiv180503653T}.
 
-The transformation to action space is nonlinear in both the phase-space position and the potential, and requires relatively strong assumptions about the system's symmetry. For instance, the 
-Galactic potential is not strictly axisymmetric; with the vast improvement in the quality of phase-space measurements due to {\em Gaia}, this assumption appears increasingly inadequate \citep[e.g.][]{2018Natur.561..360A}.
-Even if this assumption is close enough for many purposes, the parameters used in axisymmetric models may be inadequately constrained by current observations.
-
-%In addition, the necessity of 6D phase space measurements of a star limits the Galactic volume observable in action space.
-
-%With these limitations in mind, it is therefore natural to consider the reason to transform to action space.
-
-In this work we discuss how the imposition of a global axisymmetric coordinate system based on a local measurement of the midplane location and Galactocentric distance introduces systematic error in the computation of actions, especially for stars in the thin disk.
-Our main point
-is that the ``local'' Galactic midplane is not the same as the ``global''
-Galactic midplane. For stars relatively far from the Sun,
-the local midplane location will not agree with
-our local midplane extrapolated onto their position, thanks to relatively small-scale variations in the density of the gas and stellar distribution, not necessarily even as large as spiral arms.
-Converting positions and velocities of more distant stars from a heliocentric to a
-Galactocentric coordinate system introduces a systematic bias in the $z$ coordinate relative to $z$ velocity, and hence the actions computed for these stars.
-The further from the solar neighborhood the target star, the more likely the mismatch will result in large systematic uncertainty, especially in the vertical action $J_z$.
-Depending on the type of orbit being studied, this sets a limit on the ``dynamical solar neighborhood'' over which the local coordinate frame can be usefully extended for action-space analysis.
-A similar argument applies to uncertainty in measurements of
-$R_{\odot}$ and variations in the local standard of rest (LSR). In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
-system errors have on the measured actions.
-
-Lacking an empirical, nonlocal definition of the global Galactic midplane or a measurement of the expected midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire} we attempt to quantify the azimuthal dependence of the
-local midplane, and hence the expected size of the dynamical solar neighborhood, using cosmological-hydrodynamical zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic Environments (FIRE)
-collaboration\footnote{\url{https://fire.northwestern.edu}} \citep{2016ApJ...827L..23W,2017arXiv170206148H}.
-
-In Section~\ref{sec:discussion} we discuss the implications of midplane variations, and the resulting systematic uncertainty in the vertical action, for action-space analyses beyond the dynamical solar neighborhood.
-As an illustration, in Section \ref{sssec:neighborhood} we use these estimates to consider the feasibility of one science goal that relies on the invariance of actions --- the reconstruction of open clusters across large distances using ``dynamical tagging;'' that is, the association of stars to a birth cluster based on their similar actions.
-In light of our findings, we considered the orbits of the
+The transformation to action space is not trivial, and many of the assumptions
+of the transformation are potentially incorrect. For instance, the assumption
+that the Galactic potential is axisymmetric may be inadequate with current
+measurement errors. Even if this assumption is valid, the parameters used in
+our axisymmetric models may be inadequately tuned to observations. In
+addition, the necessity of 6D phase space measurements of a star limits the
+Galactic volume observable in action space.
+
+With these limitations in mind, it is therefore natural to consider the reason
+to transform to action space. One reason is dimensionality reduction --- in
+the action space paradigm, a stellar orbit is completely described by three
+actions, as opposed to six dimensions of phase space. This means that the
+relationship between {\em orbital} properties of individual stars and stellar
+properties such as age or metallicity can be investigated.
+
+We explore one science goal that relies on the conservation of actions --- the
+reconstruction of open clusters across large distances using dynamical
+tagging. The basic program is to match stars with roughly the same actions as
+their supposed natal cluster. To set the scene, we consider the orbits of the
 nine open clusters within $250\,\pc$ as reported by the {\em Gaia}
@@
 report the distance to each cluster, the three actions, and the maximum
-vertical extent of their orbits ($z_{\text{max}}$). 
-%The actions are computed using the technique described in Section~\ref{ssec:action_comp}. 
- In Section~\ref{sssec:reconstruction} we estimate and discuss the size of the dynamical solar neighborhood for each cluster based on its mass, to show that action-space tagging is likely still a viable method if background is unimportant.
-We summarize and conclude in Section~\ref{sec:conclusion}.
+vertical extent of their orbits ($z_{\text{max}}$). The actions are computed
+using the technique described in Section~\ref{ssec:action_comp}. We also show
+the mass of each cluster and the ``neighborhood'' we compute for each cluster,
+which gives a rough sense of how far one can move away from the cluster before
+the issue which we discuss next becomes important. This neighborhood is
+discussed in more detail in Section~\ref{sssec:reconstruction}.
 
@@
 \end{deluxetable*}
+
+We propose a new type of error in the computation of actions which has
+implications for general dynamical modelling across the disk. Our main point
+is that the ``local'' Galactic midplane is not the same as the ``global''
+Galactic midplane. If we observe a group of stars that are
+non-local\footnote{i.e. a group of stars that is far enough away. The exact
+distance depends on the accuracy of the measurement being made and the goals
+of the particular problem.}, they will know of a different local midplane than
+our local midplane extrapolated onto their position. This means that when we
+convert those star's positions and velocities from a heliocentric to a
+Galactocentric coordinate system, we are inducing a systematic bias in those
+stars' $z$ values. The same applies to inaccuracies in measurements of
+$R_{\odot}$ and variations in the local standard of rest (LSR).
+
+We do not propose what a modeller should use for the ``true'' global midplane,
+since such a definition probably depends on the particular problem such a
+modeller is facing. Even without a high-precision definition and measurement
+of the global Galactic midplane, we can still assess how different we should
+expect the local midplane to be by measuring the azimuthal dependence of the
+local midplane. This procedure may be readily tractable over some of the disk
+with current datasets, using appropriate selection functions. However, we turn
+to simulation in this work because we can investigate the entire disk.
+Specifically, we use cosmological zoom-ins of Milky Way-mass galaxies from the
+Feedback in Realistic Environments (FIRE)
+collaboration\footnote{\url{https://fire.northwestern.edu}} (the {\em Latte}
+suite of simulations).
+
+Our main point is that the local midplane varies between different points in
+the Galaxy. This is a different statement than that our measurements of the
+local midplane are erroneous. Indeed, the local midplane has recently been
+measured to a claimed high level of precision, setting $z_{\odot} = 20.8 \pm
+0.3\,\pc$ \citep{2019MNRAS.482.1417B}. However, we also do not mean to say
+that determinations of the local midplane are perfect, since midplane
+measurements from H~\textsc{ii} regions place $z_{\odot} \sim 5-10\,\pc$. An
+additional complication is the fact that non-zero Solar heights induce a tilt
+in the Galactic midplane \citep{2014ApJ...797...53G,2016ARAA..54..529B}.
+
+In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
+system errors have on the measured actions. In Section~\ref{sec:local_fire},
+we measure the local midplane in cosmological zoom-in simulations from the
+FIRE collaboration. In Section~\ref{sec:discussion} we discuss and give some
+context for our results, and implications for the definition of the Solar
+neighborhood. We conclude in Section~\ref{sec:conclusion}.
 
@@
 accurately and precisely enough, then the computed actions are invariant with
-orbital phase. However, we have shown that the fact that the Galactic midplane
+orbital phase. However, we hvae shown that the fact that the Galactic midplane
 is not constant across the disk presents a significant complication to
@@
 essentially equivalent.
-
-Add sentences about distribution function modeling
-
-Our main point is that the local midplane varies between different points in
-the Galaxy. This is a different statement than that our measurements of the
-local midplane are erroneous. Indeed, the local midplane has recently been
-measured to a claimed high level of precision, setting $z_{\odot} = 20.8 \pm
-0.3\,\pc$ \citep{2019MNRAS.482.1417B}. However, we also do not mean to say
-that determinations of the local midplane are perfect, since midplane
-measurements from H~\textsc{ii} regions place $z_{\odot} \sim 5-10\,\pc$. An
-additional complication is the fact that non-zero Solar heights induce a tilt
-in the Galactic midplane \citep{2014ApJ...797...53G,2016ARAA..54..529B}.
-We do not propose here what a modeler should use for the ``true'' global midplane, since it depends on the particular problem being studied.
-Even without a high-precision definition of the global Galactic midplane, current observations from Gaia should soon permit a measurement of the real azimuthal dependence of the midplane location.
 

### 2019-04-09 paper/ms.tex
Minor changes throughout

--- 2019-04-08 paper/ms.tex
@@
 \shortauthors{Beane et al.}
-\shorttitle{The Midplane is Not a Plane}
+\shorttitle{The dynamical solar neighborhood}
 
@@
 
-\title{The Midplane is Not a Plane: Implications for {\em Gaia}}
+\title{Symmetry-based constraints on the size of the dynamical Solar neighborhood}
 
@@
 
-If the system being considered departs from axisymmetry in a significant and non-adiabatic way, the actions computed using an axisymmetric approximation to the true potential can exhibit cyclic dependence on the orbital phase (or time at which the star's position and velocity are observed), large-scale migration, or diffusion from their initial values.
+If the system being considered departs from axisymmetry in a significant and/or non-adiabatic way, the actions computed using an axisymmetric approximation to the true potential can exhibit cyclic dependence on the orbital phase (or time at which the star's position and velocity are observed), large-scale migration, or diffusion from their initial values.
 In the Milky Way, stellar actions are expected to diffuse on short time scales due to scattering with gas clouds
@@
 the local midplane location will not agree with
-our local midplane extrapolated onto their position, thanks to relatively small-scale variations in the density of the gas and stellar distribution, not necessarily even as large as spiral arms.
+our local midplane extrapolated onto their position, thanks to relatively small-scale variations in the density of the gas and stellar distribution.
 Converting positions and velocities of more distant stars from a heliocentric to a
-Galactocentric coordinate system introduces a systematic bias in the $z$ coordinate relative to $z$ velocity, and hence the actions computed for these stars.
+Galactocentric coordinate system introduces a systematic bias in the $z$ coordinate, and hence the actions computed for these stars.
 The further from the solar neighborhood the target star, the more likely the mismatch will result in large systematic uncertainty, especially in the vertical action $J_z$.
@@
 %The actions are computed using the technique described in Section~\ref{ssec:action_comp}. 
- In Section~\ref{sssec:reconstruction} we estimate and discuss the size of the dynamical solar neighborhood for each cluster based on its mass, to show that action-space tagging is likely still a viable method if background is unimportant.
+ In Section~\ref{sssec:reconstruction} we estimate and discuss the size of the dynamical solar neighborhood for each cluster based on its mass, to show that action-space tagging is likely still a viable method if contamination from foreground or background stars is unimportant.
 We summarize and conclude in Section~\ref{sec:conclusion}.
@@
 We do not propose here what a modeler should use for the ``true'' global midplane, since it depends on the particular problem being studied.
-Even without a high-precision definition of the global Galactic midplane, current observations from Gaia should soon permit a measurement of the real azimuthal dependence of the midplane location.
+Even without a high-precision definition of the global Galactic midplane, current observations from {\em Gaia} should soon permit a measurement of the real azimuthal dependence of the midplane location.
 

### 2019-04-09 paper/ms.tex
Add citation to first GD1 paper

@@
 buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation by a dark matter
-subtructure in the tidal stream GD1 \citep{2018arXiv181103631B}.
+subtructure in the tidal stream GD1 \citep{2018ApJ...863L..20P, 2018arXiv181103631B}.
 These discoveries all indicate that the Milky Way's stellar distribution, which demonstrably departs significantly from axisymmetry, is undergoing phase mixing and dynamical interactions across a range of spatial and temporal scales.

### 2019-04-09 paper/ms.tex
Some small edits throughout

@@
 defined and measured both precisely and accurately. In this system, the center
-of the galaxy is usually taken to be the location of the central supermassive
+of the Galaxy is usually taken to be the location of the central supermassive
 black hole, Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From
@@
 regions in the disk to define the Galactic midplane, leading to slightly
-values ($\sim 5\,\pc$) for the Sun's relative height
+different values ($\sim 5\,\pc$) for the Sun's relative height
 \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review of these
@@
 \eeq
-where $i=R,z,\phi$ are the cylindrical radial, vertical, and azimuthal coordinates respectively and the $p_i$ are the conjugate momenta.\footnote{We typically write $J_r$ instead of $J_R$ for
+where $i=R,z,\phi$ are the radial, vertical, and azimuthal coordinates respectively in a cylindrical coordinate system and the $p_i$ are the conjugate momenta.\footnote{We typically write $J_r$ instead of $J_R$ for
 aesthetic reasons.} In a slowly-evolving
@@
 One reason is dimensionality reduction---an individual stellar orbit is concisely described by three actions, as opposed to six dimensions of phase space.
-Under the assumption of a phase-mixed system, the dynamical properties of a population of stars should be uniquely a function of their actions and independent of the conjugate angles, allowing investigation of the
+Second, under the assumption of a phase-mixed system, the dynamical properties of a population of stars should be uniquely a function of their actions and independent of the conjugate angles, allowing investigation of the
 relationship between {\em orbital} properties of stars and other intrinsic, and at least partially invariant, properties such as age or metallicity.
@@
 A similar argument applies to uncertainty in measurements of
-$R_{\odot}$ and variations in the local standard of rest (LSR). In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
+$R_{\odot}$\footnote{Here, and throughout, we use $R_{\odot}$ to mean the distance from the Sun to the Galactic center.} and variations in the local standard of rest (LSR).
+
+In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
 system errors have on the measured actions.
-
 Lacking an empirical, nonlocal definition of the global Galactic midplane or a measurement of the expected midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire} we attempt to quantify the azimuthal dependence of the
@@
 \section{Motivation} \label{sec:ref_frame}
-We first demonstrate the significance to action computations, especially for
-disk-like orbits, of systematic errors in the determination of the Galactic
-midplane and distance to the Galactic center. The consequences we explore here
+We first demonstrate the significance of systematic errors in the determination of the Galactic midplane and distance to the Galactic center to action computations.
+We will find that such errors are especially important for disk-like orbits.
+The consequences we explore here
 may also arise from various other systematic errors. For instance, the
@@
 may yield a potential that is systematically incorrect outside an original
-fitted region. In this section, we assume that the Galaxy is perfectly
+fitted region. In this work, we assume that the Galaxy is perfectly
 described by our model axisymmetric potential, and simply explore the
@@
 
-\subsection{The effect of midplane determination on orbit and action
-calculations} \label{ssec:cartoon} We present a cartoon in
+\subsection{Cartoon effect of midplane error} \label{ssec:cartoon}
+We present a cartoon in
 Figure~\ref{fig:cartoon} to show how an inaccurate determination of the
@@
 \subsection{Numerical methods} \label{ssec:action_comp}
+We now move on to quantifying the argument made in Section~\ref{ssec:cartoon}.
 We compute actions as in our previous work \citep{2018ApJ...867...31B}, using
@@
 are of acceptable accuracy, but since we also consider halo-like orbits in
-this work we choose to use orbit integration and torus mapping throughout
+this work (where the St\"ackel Fudge method is inaccurate) we choose to use orbit integration and torus mapping throughout
 \citep{2016MNRAS.457.2107S}.
 
-\subsection{Quantification of the effect of systematic coordinate-system
-errors} \label{ssec:quant}
-First, we illustrate the phase-dependence induced by an
-example systematic error in the Galactocentric coordinate system. We will
+\subsection{Quantification of the midplane effect} \label{ssec:quant}
+We quantify how a systematic error in the Galactocentric coordinate system induces phase-dependence in the {\em observed} actions.
+We will
 consider three orbits in the model potential described in
@@
 i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position in the
-orbit. This corresponds to an observer physically located at the position $(8,
+orbit. This corresponds to an observer physically located at e.g. the position $(8,
 0, 0)\,\kpc$ in the coordinate system of the true potential, but erroneously
@@
 actions as outlined above using an integration from that starting point in
-phase space carried out using the true potential (but specifying the star's
-present-day position using the systematically offset coordinate system). The
+phase space carried out using the true potential but specifying the star's
+present-day position using the systematically offset coordinate system.
+Essentially we are shifting the integrated orbit and then reintegrating at each point along the shifted orbit, where the shifted orbit is the observed orbit.
+The
 erroneous computed actions for each phase-space starting point are shown for
@@
 $z_{\text{max}}=850\,\pc$ --- a $12\%$ error
---- induced a $43\%$ error in the computation of $J_z$.
+--- induced an $86\%$ error in the computation of $J_z$.
 
@@
 95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
-$47$, and $0.71 \actunit$ for $J_r$, $L_z$ and $J_z$, respectively. These are
+$47$, and $0.71\,\actunit$ for $J_r$, $L_z$ and $J_z$, respectively. These are
 fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively.
@@
 $500\,\pc$ in the $z$ and $x$ components. In
-Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth minus 5\uth
+Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth{} minus 5\uth{}
 percentile divided by the true action value for the three different fiducial
@@
 The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
-in each action for an offset in the $z$ component (i.e. the Galactic
+in each action for an offset in the $z$-component (i.e. the Galactic
 midplane). In the {\em bottom row} we consider offsets in the $x$ component
@@
 show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
-respectively. We compute the fractional spread $\Delta J_i/J_i$ as the 95\uth
-minus 5\uth percentile of the action computed using the offset coordinate
+respectively. We compute the fractional spread $\Delta J_i/J_i$ as the 95\uth{}
+minus 5\uth{} percentile of the action computed using the offset coordinate
 system over the course of the first $\Gyr$, divided by the true action
@@
 
-In the {\em upper middle} panel, there is essentially no spread in the
+In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref}, there is essentially no spread in the
 determination of $L_z$. This is expected since, in principle, $L_z$ only
@@
 
-The {\em upper right} panel shows that the fractional error in $J_z$ is more
+The {\em upper right} panel of Figure~\ref{fig:many_orbit_wrong_ref} shows that the fractional error in $J_z$ is more
 exaggerated for more planar (disk-like) orbits. For the thin-disk orbit, a
@@
 
-For the offset in the Solar radius {\em bottom panels}, the error is largest
+For the offset in the Solar radius ({\em bottom panels}), the error is largest
 for $J_r$, with some deviations resulting in $L_z$ and relatively small
@@
 \plotone{fig/schmactions_many_orbits.pdf}
-\caption{We report the 95\uth minus 5\uth percentile of the error in the
+\caption{We report the 95\uth{} minus 5\uth{} percentile of the error in the
 measured action ($\Delta J_i$) from coordinate system errors for the thin,
@@
 respectively.\footnote{This is $2.8$ times the often-quoted
-Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
+Plummer-equivalent.} The gas softening length is adaptive, but at $\z=0$ the
 median softening length for cold ($T < 10^4\,\text{K}$) gas particles around

### 2019-04-09 paper/ms.tex
Move Jz hist fig up earlier, add extra panel

@@
 
-We now repeat the same procedure for systematic offsets between $0$ and
+\begin{figure}
+\plotone{fig/schmactions_Jz_hist.pdf}
+\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
+assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
+values have a bias (from e.g. an incorrectly computed midplane), then the
+induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
+sort of error propagation must take this into account.}
+\label{fig:Jz_hist}
+\end{figure}
+
+In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
+$J_z$ for the thick-disk orbit ({\em upper panel}) and thin-disk orbit ({\em
+lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. the {\em right} panel
+of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted
+as a vertical dashed line. Here we see that the spread in $J_z$ induced by a
+systematic offset in $z$ is non-Gaussian and bimodal. Furthermore, neither of
+the modes lie at the true value of $J_z$. In the case of the thin-disk orbit
+({\em lower panel}), we see that, in addition to the prior complications, the
+error distribution is not even centered on the true value. This comes about
+when the midplane error is $\gtrsim z_{\text{max}}$, and can be understood by
+realizing that in such a case the midplane offset always acts to increase the
+value of $J_z$, resulting in a biased estimate of $J_z$.
+
+Empirically describing the error distribution shown in
+Figure~\ref{fig:Jz_hist} is difficult. Of course, Gaussian summary statistics
+are not adequate. We therefore elect to measure this error by computing the
+95\uth{} percentile minus the 5\uth{} percentile of the distribution of
+observed action values. We refer to this quantity as $\Delta J_i$ for each
+action value. Because of the bimodality of the error distribution, this
+quantity roughly measures the distance between the peaks of the two modes.
+Furthermore, this bimodality also implies that $\Delta J_i$ is not very
+sensitive to the exact percentiles used. This summary statistic does not
+reflect the bias induced when the midplane error is $\gtrsim z_{\text{max}}$.
+
+We now repeat the same procedure as in Figure~\ref{fig:one_orbit_wrong_ref} but for systematic offsets between $0$ and
 $500\,\pc$ in the $z$ and $x$ components. In
@@
 
-\begin{figure}
-\plotone{fig/schmactions_Jz_hist.pdf}
-\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
-assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
-values have a bias (from e.g. an incorrectly computed midplane), then the
-induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
-sort of error propagation must take this into account.}
-\label{fig:Jz_hist}
-\end{figure}
-
 To further understand the effect of the midplane error, we also plot the
@@
 $z$ is enough to induce a $\sim50\%$ error in $J_z$. For larger values of
-$J_z$, the fractional errors are suppressed, with the situation becoming worse
-the greater the $z$ offset is.
-
-In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
-$J_z$ for the thick-disk orbit assuming a $z$ offset of $100\,\pc$ (i.e. the
-{\em right} panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of
-$J_z$ is plotted as a vertical dashed line. Here we see that the spread in
-$J_z$ induced by a systematic offset in $z$ is non-Gaussian and bimodal.
-Furthermore, neither of the modes lie at the true value of $J_z$.
+$J_z$, the fractional errors are suppressed, but the induced error can still be large
+depending on how great the $z$ offset is.
 

### 2019-04-09 paper/ms.tex
Minor changes throughout

@@
 \section{Azimuthal midplane variations at the Solar Circle: estimates from
-FIRE} \label{sec:local_fire}
+simulation} \label{sec:local_fire}
 \subsection{Description of FIRE Simulation} \label{ssec:cosmozoom}
@@
 for shorthand, are publicly available alongside associated mock {\em Gaia} DR2
-catalogues generated from them.\footnote{at \url{http://ananke.hub.yt}}
+catalogues generated from them.\footnote{\url{http://ananke.hub.yt}}
 
@@
 Sun the measured gas and stellar volume densities in the MW are roughly
-comparable at $\sim .04\ \Msun\
+comparable at $\sim 0.04\ \Msun\
 \unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
@@
 Using the three simulations, we determine the local midplane that an observer
-might measure if they were situated in each of these galaxies, as a function
+might measure if they were situated in each of these galaxies as a function
 of azimuth at the solar circle. Starting from the coordinate system described
-in the previous section, which is approximately aligned so that the $z$
-coordinate is perpendicular to the disk plane, we place our imaginary observer
+in the previous section, which is approximately aligned so that the $z$-coordinate is perpendicular to the disk plane, we place our imaginary observer
 at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth between
-$0<\phi<2\pi$\footnote{Our very long-lived observer also has warp-drive.}. At
+$0<\phi<2\pi$\footnote{Our very long-lived observer also has warp drive.}. At
 each value of $\phi$ we then compute the median $z$ for stars within a
@@
 stars within a cylinder of stars with height $2\,\kpc$ and the same radius
-$1000$ times and determine the $95^{\text{th}}$ minus $5^{\text{th}}$
-percentile range. This bootstrap resampling is performed by first selecting
+$1000$ times and determine the $1\,\sigma$ error bars. This bootstrap resampling is performed by first selecting
 all stars within a height of $2\,\kpc$, resampling that selection, and then
@@
 above). The $5\textup{--}95$ percentile range from the bootstrap procedure is
-shown as the dashed-line error bars. The $90\%$ interquartile range for each
-galaxy is $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the
+shown as the dashed-line error bars. The middle $90\%$ of midplane values across the Solar circle spans $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the
 three cases the midplane therefore varies by more than $\pm 100\,\pc$
@@
 of azimuth. We perform this calculation in Appendix~\ref{app:lsr} to estimate
-, but performing a best-fit subtraction to correct for misalignment of the
-original coordinate system (as in the previous section) is more involved.
-Since the variation in the LSR is less pronounced than for the midplane, and
-since the dominant effect is limited to $J_r$
-(Figure~\ref{fig:many_orbit_wrong_ref}), we defer this calculation to future
-work.
+the components of the LSR as a function of azimuth, but performing a best-fit
+subtraction to correct for misalignment of the original coordinate system (as
+in the previous section) is more involved. Since we find that the variation in
+the LSR is less pronounced than for the midplane, and since the dominant
+effect is limited to $J_r$ (Figure~\ref{fig:many_orbit_wrong_ref}), we defer
+this calculation to future work.
 
@@
 the actions under this symmetry assumption, when starting from the present-day
-positions and velocities of stars as measured, for example, by \emph{Gaia}.
-These systematic errors are most important for stars on thin-disk-like orbits,
+positions and velocities of stars as measured by, e.g., \emph{Gaia}.
+These systematic errors are most important for stars on thin disk-like orbits,
 where they can be large enough to yield actions representative of orbits in
@@
 dynamically motivated definition of the ``solar neighborhood.'' Since the
-motivation may be different for defining the solar neighborhood in e.g.
+motivation may be different for defining the solar neighborhood in, e.g.,
 studies of the chemical distribution of the Galaxy, we propose the phrase
@@
 
-From Figure \ref{fig:range_deltaphi} it is clear that the range of midplane
+From Figure~\ref{fig:range_deltaphi} it is clear that the range of midplane
 values spanned in a given $\Delta \phi$ quickly saturates at $\Delta \phi
@@
 them into the region of action space normally associated with different
-stellar populations and possibly contributing to confusion there. This
+stellar populations and possibly contributing to confusion there (see Figure~\ref{fig:Jz_hist} and discussion therein). This
 inflation is extremely non-Gaussian (as the induced error is for all orbits,
@@
 field star $i$ be close to the cluster's actions to within some bound, i.e.
-$\bm{J}_i - \bm{J}_c \in V_e$, where $V_e$ is centered on $J_c$ and has size
-proportional to the intrinsic action-space dispersion of the disrupting
+$\bm{J}_i \in V_e$, where $V_e$ is a volume centered on $\bm{J}_c$ and has size
+comparable to the intrinsic action-space dispersion of the disrupting
 cluster. This dispersion can be estimated as \citep[Section~8.3.3
@@
 contribute some stars to the selection by chance. For dissolving open clusters
-in the thin disk, this background is not likely to be negligible, but we will
-assume for the sake of argument that it can be modeled within the ``dynamical
+in the thin disk, this background is likely to be significant, but we will
+assume for the sake of argument that these foreground stars can be modeled within the ``dynamical
 solar neighborhood'' where it is safe to extend the local coordinate system.
@@
 Equation~\ref{eq:action_disp}. We then consult
-Figure~\ref{fig:many_orbit_wrong_ref} to determine the midplane $z$ offset
+Figure~\ref{fig:many_orbit_wrong_ref} to determine the midplane offset
 necessary in order for the induced error in the vertical action to be the same
@@
 members of known open clusters (see Table~\ref{tab:real_clusters}). The result
-for the allowable midplane offset for these two orbits in shown in Figure
-\ref{fig:cluster_offset} as a function of cluster mass.
+for the allowable midplane offset for these two orbits in shown in Figure~\ref{fig:cluster_offset} as a function of cluster mass.
 
@@
 measurements of variations in the MW's stellar midplane. Such measurements
-should be possible in the near future with Gaia, with quality distance
+should be possible in the near future with {\em Gaia}, with quality distance
 estimates the crucial limiting factor.

### 2019-04-09 paper/ms.tex
Add bits about Solar circle

@@
 blue and the $1\,\sigma$ dispersion as the dashed blue lines.
+We choose to only consider the Solar circle to make the point that the midplane variation as a function of azimuth (and not just Galactic radius) is important.
+However, we note that because we use the range of midplane values, including regions of each galaxy outside the Solar circle can only act to increase this range, and therefore reduce the size of the dynamical solar neighborhood.
 
@@
 $\phi$ (gray lines), and also compute the mean and $\pm1\,\sigma$ values (dark
-blue).}
+blue). We only consider the Solar circle to explore the impact of azimuthal midplane variations, but similar effects are expected for radial midplane variations.}
 \label{fig:fig_to_explain}
@@
 thick-disk orbits, respectively. A cartoon explanation of this Figure is given
-in Figure~\ref{fig:fig_to_explain}.}
+in Figure~\ref{fig:fig_to_explain}. We only consider the Solar circle to explore the impact of azimuthal midplane variations, but similar effects are expected for radial midplane variations.}
 \label{fig:range_deltaphi}

### 2019-04-09 paper/ms.tex
Change neighborhood to dynamical neighborhood

@@
 \plotone{fig/Rn_vs_mc.pdf}
-\caption{The neighborhood around an open cluster of mass $m_c$ for the thin,
-and thick orbits (\thincolor and \thickcolor, respectively; see
+\caption{The dynamical neighborhood (discussed in Section~\ref{sssec:neighborhood}) around an open cluster of mass $m_c$ for the thin-disk,
+and thick-disk orbits (\thincolor{} and \thickcolor{}, respectively; see
 Table~\ref{tab:orbits}) and for each of the FIRE galaxies considered here ---

### 2019-04-09 paper/ms.tex
Small changes

@@
 been used to study different models of spiral structure in the Milky Way
-\citep{2019MNRAS.tmp..155S}. Characteristics of the distribution of stars in the extended Solar neighborhood in action space are discussed in
+\citep{2019MNRAS.tmp..155S}. Characteristics of the distribution of stars in the extended solar neighborhood in action space are discussed in
 \citet{2018arXiv180503653T}.
@@
 
-\subsection{The size of the Solar neighborhood} \label{sssec:neighborhood}
+\subsection{The size of the solar neighborhood} \label{sssec:neighborhood}
 Given a local coordinate system, the amplitude and characteristic length scale

### 2019-04-10 paper/ms.tex

@@
 
-\title{Symmetry-based constraints on the size of the dynamical Solar neighborhood}
+\title{The Galactic Midplane is Not a Plane: Implications for {\em Gaia}}
 

### 2019-04-10 paper/ms.tex
Update midplane values

@@
 simplified version of the strategy described in
-\citealt{2019ApJ...871..145A}). For $\text{A}$ the values are $-174$, $62$,
-and $3.8\,\pc$, for $\text{B}$ the values are $0.3$, $0.65$, and
-$0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$, and
-$-15\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. For the assumed solar
+\citealt{2019ApJ...871..145A}). For $\text{A}$ the values are $-165$, $45$,
+and $8.8\,\pc$, for $\text{B}$ the values are $0.67$, $-0.088$, and
+$0.031\,\text{rad}$ and for $\text{C}$ the values are $-69$, $19$, and
+$-18\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. For the assumed Solar
 radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
-from the values of $\text{A}$ --- we compute $1.22$, $0.43$, and $0.03\,\deg$
+from the values of $\text{A}$ --- we compute $1.15$, $0.31$, and $0.06\,\deg$
 for \mi{}, \mf{}, and \mm{}. These angle offsets are consistent with the

### 2019-04-10 paper/ms.tex
Rerun to check numbers, minor changes to ttext

@@
 each galaxy, using their local ``solar neighborhood'' (the cylinder defined
-above). The $5\textup{--}95$ percentile range from the bootstrap procedure is
+above). The $1\,\sigma$ error from the bootstrap procedure is
 shown as the dashed-line error bars. The middle $90\%$ of midplane values across the Solar circle spans $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the
@@
 Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
-within a $2\,\kpc$ height of the fiducial midplane to determine error bars
-(95\uth and 5\uth percentiles), which we report as dashed lines.}
+within a $2\,\kpc$ height of the fiducial midplane to determine $1\,\sigma $error bars, which we report as dashed lines.}
 \label{fig:midplane}

### 2019-04-10 midplane_test/plot_midplane.py

@@ -56,6 +56,8 @@ for gal,ax_col in zip(glist, ax.transpose()):
     ax_col[0].set_ylim(-200, 200)
     ax_col[1].set_ylim(-200, 200)
 
+    print(gal, '90 perc: ', np.percentile((midplane_est-fit)*1000, 95) - np.percentile((midplane_est-fit)*1000, 5))
+
     ax_col[1].plot(theta/np.pi, (midplane_est-fit)*1000, c=tb_c[0])
     ax_col[1].plot(theta/np.pi, (err_low-fit)*1000, c=tb_c[0], ls='dashed', alpha=0.5)
     ax_col[1].plot(theta/np.pi, (err_high-fit)*1000, c=tb_c[0], ls='dashed', alpha=0.5)

### 2019-04-10 paper/ms.tex
Add keywords
Fix line indentation throughout

@@
 \shortauthors{Beane et al.}
-\shorttitle{The dynamical solar neighborhood}
+\shorttitle{The Galactic Midplane is Not a Plane}
 
@@
 
-\keywords{actions}
+\keywords{Galaxy:~disk -- solar~neighborhood --
+open~clusters~and~associations:~general -- Galaxy:~structure --
+Galaxy:~evolution}
 
@@
 era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
-major merger
-\citep{2018ApJ...860L..11K,2018Natur.563...85H,2018arXiv180704290L,2019MNRAS.482.3426M},
-a phase-space ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly indicating either local
-substructure infall \citep{2018MNRAS.481.1501B,2018arXiv180800451L} or bar
-buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation by a dark matter
-subtructure in the tidal stream GD1 \citep{2018ApJ...863L..20P, 2018arXiv181103631B}.
-These discoveries all indicate that the Milky Way's stellar distribution, which demonstrably departs significantly from axisymmetry, is undergoing phase mixing and dynamical interactions across a range of spatial and temporal scales.
+major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
+2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the
+solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
+local substructure infall \citep{2018MNRAS.481.1501B, 2018arXiv180800451L} or
+bar buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation
+by a dark matter subtructure in the tidal stream GD1
+\citep{2018ApJ...863L..20P, 2018arXiv181103631B}. These discoveries all
+indicate that the Milky Way's stellar distribution, which demonstrably departs
+significantly from axisymmetry, is undergoing phase mixing and dynamical
+interactions across a range of spatial and temporal scales.
 
@@
 \eeq
-where $i=R,z,\phi$ are the radial, vertical, and azimuthal coordinates respectively in a cylindrical coordinate system and the $p_i$ are the conjugate momenta.\footnote{We typically write $J_r$ instead of $J_R$ for
-aesthetic reasons.} In a slowly-evolving
-axisymmetric potential, these actions are invariant and $J_{\phi} \equiv L_z$
+where $i=R,z,\phi$ are the radial, vertical, and azimuthal coordinates
+respectively in a cylindrical coordinate system and the $p_i$ are the
+conjugate momenta.\footnote{We typically write $J_r$ instead of $J_R$ for
+aesthetic reasons.} In a slowly-evolving axisymmetric potential, these actions
+are invariant and $J_{\phi} \equiv L_z$
 \citep{2008gady.book.....B,2014RvMP...86....1S}.
 
-With the advent of 6D phase-space measurements over a relatively large ($\gtrsim 2$ kpc) volume from the {\em Gaia} satellite, the study of stars' actions has gained new popularity.
-One reason is dimensionality reduction---an individual stellar orbit is concisely described by three actions, as opposed to six dimensions of phase space.
-Second, under the assumption of a phase-mixed system, the dynamical properties of a population of stars should be uniquely a function of their actions and independent of the conjugate angles, allowing investigation of the
-relationship between {\em orbital} properties of stars and other intrinsic, and at least partially invariant, properties such as age or metallicity.
-Actions also provide a convenient basis for constructing models of the stellar distribution function, or for associating stars with similar dynamical properties (e.g.
-to determine membership in moving groups).
-
-If the system being considered departs from axisymmetry in a significant and/or non-adiabatic way, the actions computed using an axisymmetric approximation to the true potential can exhibit cyclic dependence on the orbital phase (or time at which the star's position and velocity are observed), large-scale migration, or diffusion from their initial values.
-In the Milky Way, stellar actions are expected to diffuse on short time scales due to scattering with gas clouds
-and to evolve on longer time scales in the case of orbits near resonances with spiral arms, bar(s), and other
-large scale perturbations \citep{2014RvMP...86....1S}. For this reason, we and
-other authors have used actions to study stellar scattering in the Milky Way
-disk using the improved astrometry of {\em Gaia} DR2 and various age
-catalogues \citep{2018ApJ...867...31B,2018arXiv180803278T}. Actions have also
-been used to study different models of spiral structure in the Milky Way
-\citep{2019MNRAS.tmp..155S}. Characteristics of the distribution of stars in the extended solar neighborhood in action space are discussed in
+With the advent of 6D phase-space measurements over a relatively large
+($\gtrsim 2$ kpc) volume from the {\em Gaia} satellite, the study of stars'
+actions has gained new popularity. One reason is dimensionality reduction---an
+individual stellar orbit is concisely described by three actions, as opposed
+to six dimensions of phase space. Second, under the assumption of a
+phase-mixed system, the dynamical properties of a population of stars should
+be uniquely a function of their actions and independent of the conjugate
+angles, allowing investigation of the relationship between {\em orbital}
+properties of stars and other intrinsic, and at least partially invariant,
+properties such as age or metallicity. Actions also provide a convenient basis
+for constructing models of the stellar distribution function, or for
+associating stars with similar dynamical properties (e.g. to determine
+membership in moving groups).
+
+If the system being considered departs from axisymmetry in a significant
+and/or non-adiabatic way, the actions computed using an axisymmetric
+approximation to the true potential can exhibit cyclic dependence on the
+orbital phase (or time at which the star's position and velocity are
+observed), large-scale migration, or diffusion from their initial values. In
+the Milky Way, stellar actions are expected to diffuse on short time scales
+due to scattering with gas clouds and to evolve on longer time scales in the
+case of orbits near resonances with spiral arms, bar(s), and other large scale
+perturbations \citep{2014RvMP...86....1S}. For this reason, we and other
+authors have used actions to study stellar scattering in the Milky Way disk
+using the improved astrometry of {\em Gaia} DR2 and various age catalogues
+\citep{2018ApJ...867...31B,2018arXiv180803278T}. Actions have also been used
+to study different models of spiral structure in the Milky Way
+\citep{2019MNRAS.tmp..155S}. Characteristics of the distribution of stars in
+the extended solar neighborhood in action space are discussed in
 \citet{2018arXiv180503653T}.
 
-The transformation to action space is nonlinear in both the phase-space position and the potential, and requires relatively strong assumptions about the system's symmetry. For instance, the 
-Galactic potential is not strictly axisymmetric; with the vast improvement in the quality of phase-space measurements due to {\em Gaia}, this assumption appears increasingly inadequate \citep[e.g.][]{2018Natur.561..360A}.
-Even if this assumption is close enough for many purposes, the parameters used in axisymmetric models may be inadequately constrained by current observations.
-
-%In addition, the necessity of 6D phase space measurements of a star limits the Galactic volume observable in action space.
-
-%With these limitations in mind, it is therefore natural to consider the reason to transform to action space.
-
-In this work we discuss how the imposition of a global axisymmetric coordinate system based on a local measurement of the midplane location and Galactocentric distance introduces systematic error in the computation of actions, especially for stars in the thin disk.
-Our main point
-is that the ``local'' Galactic midplane is not the same as the ``global''
-Galactic midplane. For stars relatively far from the Sun,
-the local midplane location will not agree with
-our local midplane extrapolated onto their position, thanks to relatively small-scale variations in the density of the gas and stellar distribution.
-Converting positions and velocities of more distant stars from a heliocentric to a
-Galactocentric coordinate system introduces a systematic bias in the $z$ coordinate, and hence the actions computed for these stars.
-The further from the solar neighborhood the target star, the more likely the mismatch will result in large systematic uncertainty, especially in the vertical action $J_z$.
-Depending on the type of orbit being studied, this sets a limit on the ``dynamical solar neighborhood'' over which the local coordinate frame can be usefully extended for action-space analysis.
-A similar argument applies to uncertainty in measurements of
-$R_{\odot}$\footnote{Here, and throughout, we use $R_{\odot}$ to mean the distance from the Sun to the Galactic center.} and variations in the local standard of rest (LSR).
+The transformation to action space is nonlinear in both the phase-space
+position and the potential, and requires relatively strong assumptions about
+the system's symmetry. For instance, the Galactic potential is not strictly
+axisymmetric; with the vast improvement in the quality of phase-space
+measurements due to {\em Gaia}, this assumption appears increasingly
+inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this assumption is
+close enough for many purposes, the parameters used in axisymmetric models may
+be inadequately constrained by current observations.
+
+In this work we discuss how the imposition of a global axisymmetric coordinate
+system based on a local measurement of the midplane location and
+Galactocentric distance introduces systematic error in the computation of
+actions, especially for stars in the thin disk. Our main point is that the
+``local'' Galactic midplane is not the same as the ``global'' Galactic
+midplane. For stars relatively far from the Sun, the local midplane location
+will not agree with our local midplane extrapolated onto their position,
+thanks to relatively small-scale variations in the density of the gas and
+stellar distribution. Converting positions and velocities of more distant
+stars from a heliocentric to a Galactocentric coordinate system introduces a
+systematic bias in the $z$ coordinate, and hence the actions computed for
+these stars. The further from the solar neighborhood the target star, the more
+likely the mismatch will result in large systematic uncertainty, especially in
+the vertical action $J_z$. Depending on the type of orbit being studied, this
+sets a limit on the ``dynamical solar neighborhood'' over which the local
+coordinate frame can be usefully extended for action-space analysis. A similar
+argument applies to uncertainty in measurements of $R_{\odot}$\footnote{Here,
+and throughout, we use $R_{\odot}$ to mean the distance from the Sun to the
+Galactic center.} and variations in the local standard of rest (LSR).
 
 In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
-system errors have on the measured actions.
-Lacking an empirical, nonlocal definition of the global Galactic midplane or a measurement of the expected midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire} we attempt to quantify the azimuthal dependence of the
-local midplane, and hence the expected size of the dynamical solar neighborhood, using cosmological-hydrodynamical zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic Environments (FIRE)
-collaboration\footnote{\url{https://fire.northwestern.edu}} \citep{2016ApJ...827L..23W,2017arXiv170206148H}.
-
-In Section~\ref{sec:discussion} we discuss the implications of midplane variations, and the resulting systematic uncertainty in the vertical action, for action-space analyses beyond the dynamical solar neighborhood.
-As an illustration, in Section \ref{sssec:neighborhood} we use these estimates to consider the feasibility of one science goal that relies on the invariance of actions --- the reconstruction of open clusters across large distances using ``dynamical tagging;'' that is, the association of stars to a birth cluster based on their similar actions.
-In light of our findings, we considered the orbits of the
-nine open clusters within $250\,\pc$ as reported by the {\em Gaia}
-collaboration \citep{2018AA...616A..10G}. In Table~\ref{tab:real_clusters}, we
-report the distance to each cluster, the three actions, and the maximum
-vertical extent of their orbits ($z_{\text{max}}$). 
-%The actions are computed using the technique described in Section~\ref{ssec:action_comp}. 
- In Section~\ref{sssec:reconstruction} we estimate and discuss the size of the dynamical solar neighborhood for each cluster based on its mass, to show that action-space tagging is likely still a viable method if contamination from foreground or background stars is unimportant.
-We summarize and conclude in Section~\ref{sec:conclusion}.
+system errors have on the measured actions. Lacking an empirical, nonlocal
+definition of the global Galactic midplane or a measurement of the expected
+midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire}
+we attempt to quantify the azimuthal dependence of the local midplane, and
+hence the expected size of the dynamical solar neighborhood, using
+cosmological-hydrodynamical zoom-in simulations of Milky Way-mass galaxies
+from the Feedback in Realistic Environments (FIRE)
+collaboration\footnote{\url{https://fire.northwestern.edu}}
+\citep{2016ApJ...827L..23W,2017arXiv170206148H}.
+
+In Section~\ref{sec:discussion} we discuss the implications of midplane
+variations, and the resulting systematic uncertainty in the vertical action,
+for action-space analyses beyond the dynamical solar neighborhood. As an
+illustration, in Section \ref{sssec:neighborhood} we use these estimates to
+consider the feasibility of one science goal that relies on the invariance of
+actions --- the reconstruction of open clusters across large distances using
+``dynamical tagging;'' that is, the association of stars to a birth cluster
+based on their similar actions. In light of our findings, we considered the
+orbits of the nine open clusters within $250\,\pc$ as reported by the {\em
+Gaia} collaboration \citep{2018AA...616A..10G}. In
+Table~\ref{tab:real_clusters}, we report the distance to each cluster, the
+three actions, and the maximum vertical extent of their orbits
+($z_{\text{max}}$). In Section~\ref{sssec:reconstruction} we estimate and
+discuss the size of the dynamical solar neighborhood for each cluster based on
+its mass, to show that action-space tagging is likely still a viable method if
+contamination from foreground or background stars is unimportant. We summarize
+and conclude in Section~\ref{sec:conclusion}.
 
@@
 \section{Motivation} \label{sec:ref_frame}
-We first demonstrate the significance of systematic errors in the determination of the Galactic midplane and distance to the Galactic center to action computations.
-We will find that such errors are especially important for disk-like orbits.
-The consequences we explore here
-may also arise from various other systematic errors. For instance, the
-axisymmetric Galactic potential model used in many works to compute actions
-may not be a good description of the true potential --- or the parameters used
-may yield a potential that is systematically incorrect outside an original
-fitted region. In this work, we assume that the Galaxy is perfectly
-described by our model axisymmetric potential, and simply explore the
-consequences of systematic errors in the determination of the Galactocentric
-coordinate system.
+We first demonstrate the significance of systematic errors in the
+determination of the Galactic midplane and distance to the Galactic center to
+action computations. We will find that such errors are especially important
+for disk-like orbits. The consequences we explore here may also arise from
+various other systematic errors. For instance, the axisymmetric Galactic
+potential model used in many works to compute actions may not be a good
+description of the true potential --- or the parameters used may yield a
+potential that is systematically incorrect outside an original fitted region.
+In this work, we assume that the Galaxy is perfectly described by our model
+axisymmetric potential, and simply explore the consequences of systematic
+errors in the determination of the Galactocentric coordinate system.
 
 \subsection{Cartoon effect of midplane error} \label{ssec:cartoon}
-We present a cartoon in
-Figure~\ref{fig:cartoon} to show how an inaccurate determination of the
-midplane leads to a phase-dependence of the actions calculated from an orbit
-specified by a phase-space point and an assumed potential model. The $y$-axis
-corresponds to the vertical height of the orbit and the $x$-axis orbital
-phase. The solid gray line indicates the ``true'' orbit of the star; that is,
-its orbit as it oscillates around the true midplane. The dashed gray line is
-an erroneous midplane assigned to the potential model which an observer uses
-to integrate the orbit of the star and hence estimate its actions. The model
-potential is otherwise identical to the one in which the star is actually
-moving.
+We present a cartoon in Figure~\ref{fig:cartoon} to show how an inaccurate
+determination of the midplane leads to a phase-dependence of the actions
+calculated from an orbit specified by a phase-space point and an assumed
+potential model. The $y$-axis corresponds to the vertical height of the orbit
+and the $x$-axis orbital phase. The solid gray line indicates the ``true''
+orbit of the star; that is, its orbit as it oscillates around the true
+midplane. The dashed gray line is an erroneous midplane assigned to the
+potential model which an observer uses to integrate the orbit of the star and
+hence estimate its actions. The model potential is otherwise identical to the
+one in which the star is actually moving.
 
@@
 are of acceptable accuracy, but since we also consider halo-like orbits in
-this work (where the St\"ackel Fudge method is inaccurate) we choose to use orbit integration and torus mapping throughout
-\citep{2016MNRAS.457.2107S}.
+this work (where the St\"ackel Fudge method is inaccurate) we choose to use
+orbit integration and torus mapping throughout \citep{2016MNRAS.457.2107S}.
 
 \subsection{Quantification of the midplane effect} \label{ssec:quant}
-We quantify how a systematic error in the Galactocentric coordinate system induces phase-dependence in the {\em observed} actions.
-We will
-consider three orbits in the model potential described in
-Section~\ref{ssec:action_comp} that are typical of stars in the stellar thin
-disk, stellar thick disk, and the stellar halo. We summarize their initial
-positions in phase space and the actions computed by integrating their orbits
-in the correct potential in Table~\ref{tab:orbits}. Each orbit, integrated
-without systematic coordinate errors, is plotted in Appendix~\ref{app:orbits}.
-We will refer to these orbits by their names (thin-disk, thick-disk, halo)
-henceforth.
+We quantify how a systematic error in the Galactocentric coordinate system
+induces phase-dependence in the {\em observed} actions. We will consider three
+orbits in the model potential described in Section~\ref{ssec:action_comp} that
+are typical of stars in the stellar thin disk, stellar thick disk, and the
+stellar halo. We summarize their initial positions in phase space and the
+actions computed by integrating their orbits in the correct potential in
+Table~\ref{tab:orbits}. Each orbit, integrated without systematic coordinate
+errors, is plotted in Appendix~\ref{app:orbits}. We will refer to these orbits
+by their names (thin-disk, thick-disk, halo) henceforth.
 
@@
 i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position in the
-orbit. This corresponds to an observer physically located at e.g. the position $(8,
-0, 0)\,\kpc$ in the coordinate system of the true potential, but erroneously
-thinking they are located at $(8, 0, 0.1)\,\kpc$.
+orbit. This corresponds to an observer physically located at e.g. the position
+$(8, 0, 0)\,\kpc$ in the coordinate system of the true potential, but
+erroneously thinking they are located at $(8, 0, 0.1)\,\kpc$.
 
@@
 present-day position using the systematically offset coordinate system.
-Essentially we are shifting the integrated orbit and then reintegrating at each point along the shifted orbit, where the shifted orbit is the observed orbit.
-The
-erroneous computed actions for each phase-space starting point are shown for
-the first $\Gyr$ of the orbit in the {\em upper} panels
+Essentially we are shifting the integrated orbit and then reintegrating at
+each point along the shifted orbit, where the shifted orbit is the observed
+orbit. The erroneous computed actions for each phase-space starting point are
+shown for the first $\Gyr$ of the orbit in the {\em upper} panels
 Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
@@
 
-We now repeat the same procedure as in Figure~\ref{fig:one_orbit_wrong_ref} but for systematic offsets between $0$ and
-$500\,\pc$ in the $z$ and $x$ components. In
-Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth{} minus 5\uth{}
-percentile divided by the true action value for the three different fiducial
-orbits in Table \ref{tab:orbits}. The thick-disk (\thickcolor) orbit is the
-one from Figure~\ref{fig:one_orbit_wrong_ref}, but we also now consider the
-effect on the action determined for the thin-disk (\thincolor) and halo
-(\halocolor) orbits.
+We now repeat the same procedure as in Figure~\ref{fig:one_orbit_wrong_ref}
+but for systematic offsets between $0$ and $500\,\pc$ in the $z$ and $x$
+components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth{}
+minus 5\uth{} percentile divided by the true action value for the three
+different fiducial orbits in Table \ref{tab:orbits}. The thick-disk
+(\thickcolor) orbit is the one from Figure~\ref{fig:one_orbit_wrong_ref}, but
+we also now consider the effect on the action determined for the thin-disk
+(\thincolor) and halo (\halocolor) orbits.
 
@@
 
-In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref}, there is essentially no spread in the
-determination of $L_z$. This is expected since, in principle, $L_z$ only
-depends on the $x$- and $y$-components of the position and velocity of the
-stars\footnote{In practice, however, $L_z$ is computed as part of the
-torus-fitting method.}, and is thus unaffected by offsets in $z$. Indeed, the
-result we found in Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk
-orbit holds for all orbit types.
-
-The {\em upper right} panel of Figure~\ref{fig:many_orbit_wrong_ref} shows that the fractional error in $J_z$ is more
-exaggerated for more planar (disk-like) orbits. For the thin-disk orbit, a
-systematic offset of $\sim30\,\pc$ in the $z$ coordinate already results in
-$100\%$ deviations in the actions. The required offset for $100\%$ deviation
-is $\sim225\,\pc$ for the thick-disk orbit. The halo orbit is relatively
-resistant to errors in the midplane, with only $\sim25\%$ error in $J_z$ out
-to a midplane offset of $500\,\pc$.
+In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref},
+there is essentially no spread in the determination of $L_z$. This is expected
+since, in principle, $L_z$ only depends on the $x$- and $y$-components of the
+position and velocity of the stars\footnote{In practice, however, $L_z$ is
+computed as part of the torus-fitting method.}, and is thus unaffected by
+offsets in $z$. Indeed, the result we found in
+Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
+orbit types.
+
+The {\em upper right} panel of Figure~\ref{fig:many_orbit_wrong_ref} shows
+that the fractional error in $J_z$ is more exaggerated for more planar
+(disk-like) orbits. For the thin-disk orbit, a systematic offset of
+$\sim30\,\pc$ in the $z$ coordinate already results in $100\%$ deviations in
+the actions. The required offset for $100\%$ deviation is $\sim225\,\pc$ for
+the thick-disk orbit. The halo orbit is relatively resistant to errors in the
+midplane, with only $\sim25\%$ error in $J_z$ out to a midplane offset of
+$500\,\pc$.
 
@@
 $z$ is enough to induce a $\sim50\%$ error in $J_z$. For larger values of
-$J_z$, the fractional errors are suppressed, but the induced error can still be large
-depending on how great the $z$ offset is.
+$J_z$, the fractional errors are suppressed, but the induced error can still
+be large depending on how great the $z$ offset is.
 
@@
 simulated galaxies considered in this work.\label{tab:scale_height}}
-\tablehead{\colhead{galaxy} & \colhead{cold\tablenotemark{a} gas disk} & \colhead{stellar thin disk} & \colhead{stellar thick disk} & \colhead{cold gas softening length} & \colhead{stellar softening length} \\
+\tablehead{\colhead{galaxy} & \colhead{cold\tablenotemark{a} gas disk} &
+\colhead{stellar thin disk} & \colhead{stellar thick disk} & \colhead{cold gas
+softening length} & \colhead{stellar softening length} \\
 \colhead{} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} } 
@@
 Using the three simulations, we determine the local midplane that an observer
-might measure if they were situated in each of these galaxies as a function
-of azimuth at the solar circle. Starting from the coordinate system described
-in the previous section, which is approximately aligned so that the $z$-coordinate is perpendicular to the disk plane, we place our imaginary observer
-at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth between
-$0<\phi<2\pi$\footnote{Our very long-lived observer also has warp drive.}. At
-each value of $\phi$ we then compute the median $z$ for stars within a
-cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to the
-fiducial disk. We then re-define this median $z$ as the new midplane of the
-cylinder, re-select stars, and iterate until the median $z$ value converges.
-We find that only $10$ iterations of this procedure are necessary for
-convergence. The resulting median $z$ is taken to be what our observer would
-measure as the local Galactic midplane at each $\phi$. This procedure assumes
-perfect density estimation, and therefore perfect corrections for extinction,
-within the cylinder defining the ``solar neighborhood.'' Imperfect extinction
-correction is likely to increase the amplitude of the estimated fluctuations
-in $z$. To account for the effect of particle noise, we bootstrap-resample
-stars within a cylinder of stars with height $2\,\kpc$ and the same radius
-$1000$ times and determine the $1\,\sigma$ error bars. This bootstrap resampling is performed by first selecting
-all stars within a height of $2\,\kpc$, resampling that selection, and then
-repeating the $10$ iterations.
+might measure if they were situated in each of these galaxies as a function of
+azimuth at the solar circle. Starting from the coordinate system described in
+the previous section, which is approximately aligned so that the
+$z$-coordinate is perpendicular to the disk plane, we place our imaginary
+observer at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth
+between $0<\phi<2\pi$\footnote{Our very long-lived observer also has warp
+drive.}. At each value of $\phi$ we then compute the median $z$ for stars
+within a cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to
+the fiducial disk. We then re-define this median $z$ as the new midplane of
+the cylinder, re-select stars, and iterate until the median $z$ value
+converges. We find that only $10$ iterations of this procedure are necessary
+for convergence. The resulting median $z$ is taken to be what our observer
+would measure as the local Galactic midplane at each $\phi$. This procedure
+assumes perfect density estimation, and therefore perfect corrections for
+extinction, within the cylinder defining the ``solar neighborhood.'' Imperfect
+extinction correction is likely to increase the amplitude of the estimated
+fluctuations in $z$. To account for the effect of particle noise, we
+bootstrap-resample stars within a cylinder of stars with height $2\,\kpc$ and
+the same radius $1000$ times and determine the $1\,\sigma$ error bars. This
+bootstrap resampling is performed by first selecting all stars within a height
+of $2\,\kpc$, resampling that selection, and then repeating the $10$
+iterations.
 
@@
 each galaxy, using their local ``solar neighborhood'' (the cylinder defined
-above). The $1\,\sigma$ error from the bootstrap procedure is
-shown as the dashed-line error bars. The middle $90\%$ of midplane values across the Solar circle spans $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the
-three cases the midplane therefore varies by more than $\pm 100\,\pc$
+above). The $1\,\sigma$ error from the bootstrap procedure is shown as the
+dashed-line error bars. The middle $90\%$ of midplane values across the Solar
+circle spans $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of
+the three cases the midplane therefore varies by more than $\pm 100\,\pc$
 depending on the azimuth along the Solar circle; in the third (\mm{}, which
@@
 Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
-within a $2\,\kpc$ height of the fiducial midplane to determine $1\,\sigma $error bars, which we report as dashed lines.}
+within a $2\,\kpc$ height of the fiducial midplane to determine $1\,\sigma
+$error bars, which we report as dashed lines.}
 \label{fig:midplane}
@@
 Figure~\ref{fig:range_deltaphi} as a gray line, with the mean value in solid
-blue and the $1\,\sigma$ dispersion as the dashed blue lines.
-We choose to only consider the Solar circle to make the point that the midplane variation as a function of azimuth (and not just Galactic radius) is important.
-However, we note that because we use the range of midplane values, including regions of each galaxy outside the Solar circle can only act to increase this range, and therefore reduce the size of the dynamical solar neighborhood.
+blue and the $1\,\sigma$ dispersion as the dashed blue lines. We choose to
+only consider the Solar circle to make the point that the midplane variation
+as a function of azimuth (and not just Galactic radius) is important. However,
+we note that because we use the range of midplane values, including regions of
+each galaxy outside the Solar circle can only act to increase this range, and
+therefore reduce the size of the dynamical solar neighborhood.
 
@@
 $\phi$ (gray lines), and also compute the mean and $\pm1\,\sigma$ values (dark
-blue). We only consider the Solar circle to explore the impact of azimuthal midplane variations, but similar effects are expected for radial midplane variations.}
+blue). We only consider the Solar circle to explore the impact of azimuthal
+midplane variations, but similar effects are expected for radial midplane
+variations.}
 \label{fig:fig_to_explain}
@@
 thick-disk orbits, respectively. A cartoon explanation of this Figure is given
-in Figure~\ref{fig:fig_to_explain}. We only consider the Solar circle to explore the impact of azimuthal midplane variations, but similar effects are expected for radial midplane variations.}
+in Figure~\ref{fig:fig_to_explain}. We only consider the Solar circle to
+explore the impact of azimuthal midplane variations, but similar effects are
+expected for radial midplane variations.}
 \label{fig:range_deltaphi}
@@
 them into the region of action space normally associated with different
-stellar populations and possibly contributing to confusion there (see Figure~\ref{fig:Jz_hist} and discussion therein). This
-inflation is extremely non-Gaussian (as the induced error is for all orbits,
-e.g. Figure~\ref{fig:Jz_hist}), and is particularly troubling for the case of
-thin disk stars, where the necessary midplane deviation is
+stellar populations and possibly contributing to confusion there (see
+Figure~\ref{fig:Jz_hist} and discussion therein). This inflation is extremely
+non-Gaussian (as the induced error is for all orbits, e.g.
+Figure~\ref{fig:Jz_hist}), and is particularly troubling for the case of thin
+disk stars, where the necessary midplane deviation is
 $\sim50\textup{--}100\,\pc$.
@@
 in the thin disk, this background is likely to be significant, but we will
-assume for the sake of argument that these foreground stars can be modeled within the ``dynamical
-solar neighborhood'' where it is safe to extend the local coordinate system.
-Since the effect of extending the coordinate frame beyond this region is
-generally to inflate the actions computed for thin-disk stars, using too large
-a volume is likely to under-estimate the density of the smooth thin-disk-like
-component in action space, while scattering more thin-disk stars onto
-thick-disk-like orbits and over-estimating the density of the smooth
-thick-disk-like component.
+assume for the sake of argument that these foreground stars can be modeled
+within the ``dynamical solar neighborhood'' where it is safe to extend the
+local coordinate system. Since the effect of extending the coordinate frame
+beyond this region is generally to inflate the actions computed for thin-disk
+stars, using too large a volume is likely to under-estimate the density of the
+smooth thin-disk-like component in action space, while scattering more
+thin-disk stars onto thick-disk-like orbits and over-estimating the density of
+the smooth thick-disk-like component.
 
@@
 members of known open clusters (see Table~\ref{tab:real_clusters}). The result
-for the allowable midplane offset for these two orbits in shown in Figure~\ref{fig:cluster_offset} as a function of cluster mass.
+for the allowable midplane offset for these two orbits in shown in
+Figure~\ref{fig:cluster_offset} as a function of cluster mass.
 
@@
 
-%In Figure~\ref{fig:cluster_offset}, we show the $z$ offset necessary to induce
-%the action space dispersion computed from Equation~\ref{eq:action_disp}, using
-%the \texttt{gala} computed mass enclosed within a radius of $8.2\,\kpc$ ($\sim
-%10^{11}\,M_{\odot}$).
-
-%We now convert the $z$~offset-$m_c$ relation to an $R_{\n}$-$m_c$ relation,
-%where $R_{\n}$ indicates the neighborhood radius. An observer
-%attempting to reconstruct an open cluster of mass $m_c$ with perfect knowledge
-%of the midplane at the location of the cluster would expect to find the
-%ejected members using a dynamical cut for field stars within a distance
-%$R_{\n}$ of the cluster. If such an observer wants to find ejected stars that
-%are at a distance greater than $R_{\n}$, a more sophisticated analysis would
-%be required.
-
 Figure~\ref{fig:Rn_mc} shows that azimuthal midplane fluctuations will limit
@@
 \plotone{fig/Rn_vs_mc.pdf}
-\caption{The dynamical neighborhood (discussed in Section~\ref{sssec:neighborhood}) around an open cluster of mass $m_c$ for the thin-disk,
-and thick-disk orbits (\thincolor{} and \thickcolor{}, respectively; see
-Table~\ref{tab:orbits}) and for each of the FIRE galaxies considered here ---
+\caption{The dynamical neighborhood (discussed in
+Section~\ref{sssec:neighborhood}) around an open cluster of mass $m_c$ for the
+thin-disk, and thick-disk orbits (\thincolor{} and \thickcolor{},
+respectively; see Table~\ref{tab:orbits}) and for each of the FIRE galaxies
+considered here ---
 \mi{}, \mf{}, and \mm{} ({\em left}, {\em center}, and {\em right} panels,
@@
 additional complication is the fact that non-zero Solar heights induce a tilt
-in the Galactic midplane \citep{2014ApJ...797...53G,2016ARAA..54..529B}.
-We do not propose here what a modeler should use for the ``true'' global midplane, since it depends on the particular problem being studied.
-Even without a high-precision definition of the global Galactic midplane, current observations from {\em Gaia} should soon permit a measurement of the real azimuthal dependence of the midplane location.
+in the Galactic midplane \citep{2014ApJ...797...53G,2016ARAA..54..529B}. We do
+not propose here what a modeler should use for the ``true'' global midplane,
+since it depends on the particular problem being studied. Even without a
+high-precision definition of the global Galactic midplane, current
+observations from {\em Gaia} should soon permit a measurement of the real
+azimuthal dependence of the midplane location.
 

### 2019-04-10 .ci/build-paper.sh
Add hogg paper

@@ -7,6 +7,7 @@ then
   git checkout --orphan $TRAVIS_BRANCH-pdf
   git rm -rf .
   git add -f paper/ms.pdf
+  git add -f paper/ms_hogg.pdf
   git -c user.name='travis' -c user.email='travis' commit -m "building the paper"
   git push -q -f https://$GITHUB_USER:$GITHUB_API_KEY@github.com/$TRAVIS_REPO_SLUG $TRAVIS_BRANCH-pdf
 fi

### 2019-04-10 .ci/gen-hogg.sh

@@ -1,0 +1,3 @@
+cd $TRAVIS_BUILD_DIR/paper
+cp ms.tex ms_hogg.tex
+sed -i 's/twocolumn/fancy/g' ms_hogg.tex

### 2019-04-10 .ci/gen-hogg.sh
Fix travis

@@ -1,3 +1,3 @@
 cd $TRAVIS_BUILD_DIR/paper
 cp ms.tex ms_hogg.tex
-sed -i 's/twocolumn/fancy/g' ms_hogg.tex
+sed -i 's/twocolumn/modern/g' ms_hogg.tex

### 2019-04-16 paper/ms.tex
Fix travis
Converted this text to an issue
Add citation to new GRAVITY measurement
Update citation to Trick et al paper
Add section of analytic action errors in epic. approx.

@@
 \usepackage{bm}
+\usepackage{mathtools}
 
@@
 the Galactic center is taken to be $\sim 8.1\text{--}8.3\,\kpc$
-\citep{2009ApJ...692.1075G,2018AA...615L..15G}. The Sun's vertical distance
+\citep{2009ApJ...692.1075G,2018AA...615L..15G}, with a recent $\sim0.3\%$ measurement from the GRAVITY collaboration \citep{2019arXiv190405721A}. The Sun's vertical distance
 from the Galactic midplane, the other independent measurement that sets this
@@
 the extended solar neighborhood in action space are discussed in
-\citet{2018arXiv180503653T}.
+\citet{2019MNRAS.484.3291T}.
 
@@
 
+\subsection{Epicyclic Approximation} \label{ssec:epi_action}
+Before turning to numerical methods, we first write down analytically the
+induced action error from either a midplane offset or a Galactocentric radius
+offset. We make the epicyclic approximation, which assumes that the motion in
+the $z$ and $R$ components of the orbit are decoupled and follow a simple
+harmonic oscillator. This approximation should be an excellent description of
+thin disk-like orbits and a good description of thick disk-like orbits. We
+also make the assumption of a perfectly flat circular velocity curve with
+velocity $v_c$.
+
+Under this approximation, we can write down the vertical and radial components
+of the orbits as,
+\beq\label{eq:orbits_epi}
+\begin{split}
+z(t) &= A_z \sin{(\kappa t + \alpha)}\\
+R(t) &= A_R \sin{(\nu t + \beta)}\text{,}
+\end{split}
+\eeq
+where $\kappa$ and $\nu$ are the vertical and radial (or epicyclic)
+frequencies, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
+vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
+constants. We can write down the actions in this approximation as
+\citep[Section~3.5.3b][]{2008gady.book.....B},
+\beq\label{eq:actions_epi}
+\begin{split}
+L_z &= R_g v_c \\
+J_r &= \kappa A_R^2 \\
+J_z &= \nu A_z^2\text{,}
+\end{split}
+\eeq
+where $R_g$ is the guiding radius of the orbit.
+
+An induced error in $z$ of $\Delta z$ will cause $\Delta A_z \sim \Delta z$.
+Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$. If
+we assume that $\kappa$ and $\nu$ are constant (\Gus{justify?}), then we have
+that
+\beq\label{eq:actions_epi_error}
+\begin{split}
+\Delta L_z = \Delta R v_c &\implies \frac{\Delta L_z}{L_z} = \frac{\Delta R}{R} \\
+\Delta J_r = 2 \kappa \Delta A_R &\implies \frac{\Delta J_r}{J_r} = \frac{2\Delta R}{A_R} \\
+\Delta J_z = 2 \nu \Delta A_z &\implies \frac{\Delta J_z}{J_z} = \frac{2\Delta z}{A_z}\text{.}
+\end{split}
+\eeq
+Therefore, errors in each action are linearly proportional to the errors in
+either the midplane or cylindrical radius. We will check these formulae
+against numerical tests next.
+
 \subsection{Numerical methods} \label{ssec:action_comp}
@@
 recent works exploring actions in the Galactic disk
-\citep[e.g.][]{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}.
+\citep[e.g.][]{2019MNRAS.484.3291T,2018MNRAS.481.4093S,2018arXiv180803278T}.
 For disk-like orbits, existing implementations of the St\"ackel Fudge method
@@
 
-%%%move this later - decide where, probably conclusions While we focus our
-%discussion now solely on the midplane, there is an additional complication in
-%the measurement of the distance from the Sun to the Galactic center that may
-%also be important. When modelling stars at Galactic radii close to the Sun's
-%radius, it is in principle possible to determine what point they are orbiting
-%about: the ``dynamical Galactic center.'' It is not guaranteed that this
-%dynamical center will coincide with the distance to Sgr~A\textsuperscript{*}
-%\citep{2018AA...615L..15G}. In fact, \citet{2015ApJ...803...80K} were able to
-%measure the distance of the Sun to the Galactic center using the dynamical
-%information of Palomar 5. Other dynamical measurements of the Solar distance
-%have been made, though none with a precision capable of competing with the
-%distance to Sgr~A\textsuperscript{*}
-%\citep{1981gask.book.....M,2011PASJ...63..867S,2012MNRAS.427..274S,2013AstL...39...95B,2013IAUS..289..444Z}.
-
 \subsection{The size of the solar neighborhood} \label{sssec:neighborhood}

### 2019-04-16 paper/ms.tex
Fix Jr error calc

@@
 \newcommand{\abs}[1]{\left| #1 \right|}
+\newcommand{\avg}[1]{\left< #1 \right>}
 \newcommand{\z}{z_r}
@@
 Therefore, errors in each action are linearly proportional to the errors in
-either the midplane or cylindrical radius. We will check these formulae
-against numerical tests next.
+either the midplane or cylindrical radius.
+
+There is a slight complication in the formula for computing $\Delta J_r/J_r$.
+In a moment we will consider the effect of an inaccuracy in the distance from
+the Sun to the Galactic center on actions. However, this inaccuracy induces an
+error in $x$, or $\Delta x$. This is not the same as an inaccuracy in $\Delta
+R$, but they are related by the following argument. An error $\Delta x$ will
+induce an erroneous radius $R_{\text{err}}$ related by the formula,
+\beq
+(x+\Delta x)^2 + y^2 = R_{\text{err}}^2\text{.}
+\eeq
+Keeping only terms to first order in $\Delta x$, we have that,
+\beq
+\begin{split}
+R_{\text{err}}^2 &= R^2 - 2 R \cos{\theta} \Delta x \\
+\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\theta}} \Delta x\text{.}
+\end{split}
+\eeq
+Averaging over the circle, we therefore have that
+\beq
+\avg{\Delta R} = \frac{2}{\pi} \Delta x\text{.}
+\eeq
+Therefore, the error in $J_r$ is related to $\Delta x$ by,
+\beq\label{eq:Jr_epi_dx}
+\frac{\Delta J_r}{J_r} = \frac{4 \Delta x}{\pi A_R}\text{.}
+\eeq
+
+We will check these formulae against numerical tests next.
 

### 2019-04-16 paper/ms.tex
Add AR to table
Update Delta J_i references in paper

@@
 
-\begin{deluxetable*}{ccccccc}
+\begin{deluxetable*}{cccccccc}
 \tablecaption{Description and names of the three orbits considered in this
@@
 velocity} & \colhead{$J_r$} &
-\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$}\\ \colhead{ } &
+\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
+\colhead{$A_R$}\\ \colhead{ } &
 \colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
@@
 \colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{kpc}$}}
+\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$}}
 \startdata 
-thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 \\
-thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 \\ 
-halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16
+thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 & 1.29 \\
+thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 & 1.19 \\ 
+halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16 & 2.34
 \enddata
@@
 induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
-sort of error propagation must take this into account.}
+sort of error propagation must take this into account. We also plot one half
+the 95\uth{} percentile minus the 5\uth{} percentile of each distribution as a
+horizontal arrow anchored on the true $J_z$ value. We call this $\Delta J_z$
+and will use it (along with the similarly defined $\Delta J_r$ and $\Delta
+L_z$) to empirically describe the error distribution. We see that $\Delta J_z$
+roughly corresponds to the distance from the true $J_z$ value to one of the
+modes of the error distribution.}
 \label{fig:Jz_hist}
@@
 Figure~\ref{fig:Jz_hist} is difficult. Of course, Gaussian summary statistics
-are not adequate. We therefore elect to measure this error by computing the
+are not adequate. We therefore elect to measure this error by computing one half the
 95\uth{} percentile minus the 5\uth{} percentile of the distribution of
 observed action values. We refer to this quantity as $\Delta J_i$ for each
-action value. Because of the bimodality of the error distribution, this
-quantity roughly measures the distance between the peaks of the two modes.
+action value. We plot this quantity in Figure~\ref{fig:Jz_hist} as a horizontal arrow anchored on the true action value. Because of the bimodality of the error distribution, this
+quantity roughly measures the distance from the true action value to the peak of one of the modes.
 Furthermore, this bimodality also implies that $\Delta J_i$ is not very
@@
 but for systematic offsets between $0$ and $500\,\pc$ in the $z$ and $x$
-components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth{}
-minus 5\uth{} percentile divided by the true action value for the three
+components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report one half the 95\uth{}
+minus 5\uth{} percentile divided by the true action value ($\Delta J_i/J_i$) for the three
 different fiducial orbits in Table \ref{tab:orbits}. The thick-disk
@@
 show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
-respectively. We compute the fractional spread $\Delta J_i/J_i$ as the 95\uth{}
+respectively. We compute the fractional spread $\Delta J_i/J_i$ as one half the 95\uth{}
 minus 5\uth{} percentile of the action computed using the offset coordinate
@@
 \plotone{fig/schmactions_many_orbits.pdf}
-\caption{We report the 95\uth{} minus 5\uth{} percentile of the error in the
+\caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error in the
 measured action ($\Delta J_i$) from coordinate system errors for the thin,
-thick, and halo orbits (Table~\ref{tab:orbits}). The {\em left}, {\em center},
+thick, and halo orbits (Table~\ref{tab:orbits}). See discussion in the text for the justification in using this to measure the magnitude of the induced error. The {\em left}, {\em center},
 and {\em right} panels show the result for $J_r$, $L_z$, and $J_z$,
@@
 different offsets in $z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
-(red, cyan, green respectively). As before, the error ($\Delta J_z$) is the
+(red, cyan, green respectively). As before, the error ($\Delta J_z$) is one half the
 95\uth{} minus 5\uth{} percentile of the distribution of observed $J_z$ values

### 2019-04-16 paper/ms.tex
Add Lz to deltaR-deltax discussion

@@
 
-There is a slight complication in the formula for computing $\Delta J_r/J_r$.
+There is a slight complication in the formula for computing $\Delta J_r/J_r$ and $\Delta L_z/L_z$.
 In a moment we will consider the effect of an inaccuracy in the distance from
@@
 \eeq
-Therefore, the error in $J_r$ is related to $\Delta x$ by,
+Therefore, the error in $J_r$ and $L_z$ are related to $\Delta x$ by,
 \beq\label{eq:Jr_epi_dx}
-\frac{\Delta J_r}{J_r} = \frac{4 \Delta x}{\pi A_R}\text{.}
+\begin{split}
+\frac{\Delta J_r}{J_r} &= \frac{4 \Delta x}{\pi A_R} \\
+\frac{\Delta L_z}{L_z} &= \frac{2 \Delta x}{\pi R}\text{.}
+\end{split}
 \eeq

### 2019-04-16 schmactions/many_orbits/plot_many_orbits_Jzfun.py
Update many orbits Jz fun plot with new delta Jz

@@ -65,7 +65,7 @@ for n,c in zip(names, colors):
             Jz = act[2]
             up = np.percentile(r, 95, axis=0)[2]
             low = np.percentile(r, 5, axis=0)[2]
-            to_plot.append([Jz, 100*(up-low)/Jz])
+            to_plot.append([Jz, 100*(up-low)/(2*Jz)])
         except:
             to_plot.append([np.nan, np.nan])
     to_plot = np.array(to_plot)

### 2019-04-16 paper/ms.tex
Update abstract

--- 2019-04-16 paper/ms.tex
@@
 \usepackage{bm}
-\usepackage{mathtools}
 
@@
 \newcommand{\abs}[1]{\left| #1 \right|}
-\newcommand{\avg}[1]{\left< #1 \right>}
 \newcommand{\z}{z_r}
@@
 complications, a systematic bias in the phase space position induces a
-phase-dependence in the actions, which we interpret as an error. This error is
+phase-dependence in the actions, which we interpret as an error. The induced error distribution is
 non-Gaussian and bimodal, 
-%mm [this sounds imprecise: an error wouldn't peak on a value...]
 with neither of the modes peaking on the true action value. 
@@
 \end{figure*}
-
-\subsection{Epicyclic Approximation} \label{ssec:epi_action}
-Before turning to numerical methods, we first write down analytically the
-induced action error from either a midplane offset or a Galactocentric radius
-offset. We make the epicyclic approximation, which assumes that the motion in
-the $z$ and $R$ components of the orbit are decoupled and follow a simple
-harmonic oscillator. This approximation should be an excellent description of
-thin disk-like orbits and a good description of thick disk-like orbits. We
-also make the assumption of a perfectly flat circular velocity curve with
-velocity $v_c$.
-
-Under this approximation, we can write down the vertical and radial components
-of the orbits as,
-\beq\label{eq:orbits_epi}
-\begin{split}
-z(t) &= A_z \sin{(\kappa t + \alpha)}\\
-R(t) &= A_R \sin{(\nu t + \beta)}\text{,}
-\end{split}
-\eeq
-where $\kappa$ and $\nu$ are the vertical and radial (or epicyclic)
-frequencies, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
-vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
-constants. We can write down the actions in this approximation as
-\citep[Section~3.5.3b][]{2008gady.book.....B},
-\beq\label{eq:actions_epi}
-\begin{split}
-L_z &= R_g v_c \\
-J_r &= \kappa A_R^2 \\
-J_z &= \nu A_z^2\text{,}
-\end{split}
-\eeq
-where $R_g$ is the guiding radius of the orbit.
-
-An induced error in $z$ of $\Delta z$ will cause $\Delta A_z \sim \Delta z$.
-Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$. If
-we assume that $\kappa$ and $\nu$ are constant (\Gus{justify?}), then we have
-that
-\beq\label{eq:actions_epi_error}
-\begin{split}
-\Delta L_z = \Delta R v_c &\implies \frac{\Delta L_z}{L_z} = \frac{\Delta R}{R} \\
-\Delta J_r = 2 \kappa \Delta A_R &\implies \frac{\Delta J_r}{J_r} = \frac{2\Delta R}{A_R} \\
-\Delta J_z = 2 \nu \Delta A_z &\implies \frac{\Delta J_z}{J_z} = \frac{2\Delta z}{A_z}\text{.}
-\end{split}
-\eeq
-Therefore, errors in each action are linearly proportional to the errors in
-either the midplane or cylindrical radius.
-
-There is a slight complication in the formula for computing $\Delta J_r/J_r$.
-In a moment we will consider the effect of an inaccuracy in the distance from
-the Sun to the Galactic center on actions. However, this inaccuracy induces an
-error in $x$, or $\Delta x$. This is not the same as an inaccuracy in $\Delta
-R$, but they are related by the following argument. An error $\Delta x$ will
-induce an erroneous radius $R_{\text{err}}$ related by the formula,
-\beq
-(x+\Delta x)^2 + y^2 = R_{\text{err}}^2\text{.}
-\eeq
-Keeping only terms to first order in $\Delta x$, we have that,
-\beq
-\begin{split}
-R_{\text{err}}^2 &= R^2 - 2 R \cos{\theta} \Delta x \\
-\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\theta}} \Delta x\text{.}
-\end{split}
-\eeq
-Averaging over the circle, we therefore have that
-\beq
-\avg{\Delta R} = \frac{2}{\pi} \Delta x\text{.}
-\eeq
-Therefore, the error in $J_r$ is related to $\Delta x$ by,
-\beq\label{eq:Jr_epi_dx}
-\frac{\Delta J_r}{J_r} = \frac{4 \Delta x}{\pi A_R}\text{.}
-\eeq
-
-We will check these formulae against numerical tests next.
 

### 2019-04-16 paper/ms.tex
Update description of orbits table

--- 2019-04-16 paper/ms.tex
@@
 \tablecaption{Description and names of the three orbits considered in this
-work.\label{tab:orbits}}
+work. For the last two columns: $z_{\text{max}}$ the maximum height of the
+orbit while $A_R$ is the magnitude of the radial excursions of the orbit,
+computed as one half the maximum radius minus the minimum
+radius.\label{tab:orbits}}
 \tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial

### 2019-04-16 paper/ms.tex
Change abstract to 10 perc error

@@
 with neither of the modes peaking on the true action value. 
-A midplane error of $\sim30\,\pc$ for a thin disk orbit or
-$\sim225\,\pc$ for a thick disk orbit is all that is necessary to
-induce a $100\,\%$ error in the vertical action $J_z$.
-orbits, respectively. Furthermore, we show that the local midplane varies by
+A midplane error of $\sim6\,\pc$ for a thin disk orbit or
+$\sim50\,\pc$ for a thick disk orbit is all that is necessary to
+induce a $10\,\%$ error in the vertical action $J_z$.
+Furthermore, we show that the local midplane varies by
 $\sim100\,\pc$ at the Solar circle in Milky Way-mass cosmological zoom-in

### 2019-04-16 schmactions/many_orbits/plot_many_orbits.py

@@ -64,9 +64,9 @@ clist = [tb_c[7], tb_c[8], tb_c[0]]
 def print_100(zerr, dJz):
     interp = interp1d(zerr, dJz, bounds_error=False, fill_value='extrapolate')
     def to_min(z):
-        return np.abs(interp(z)-1)
+        return np.abs(interp(z)-0.1)
     res = minimize(to_min, 250)
-    print(res.x)
+    print('zerr needed for 10perc:', res.x)
 
 for fname, name, Az, AR, Rg, init_vel, c in zip(fname_list, name_list, Az_list, AR_list, Rg_list, init_vel_list, clist):
     zout = pickle.load(open('zout_'+fname+'.p', 'rb'))

### 2019-04-16 schmactions/many_orbits/plot_many_orbits.py
Many orbits print 10perc and 100perc errors

@@ -61,12 +61,12 @@ init_vel_list = [[0, -190, 10] * u.km/u.s,
 
 clist = [tb_c[7], tb_c[8], tb_c[0]]
 
-def print_100(zerr, dJz):
+def print_100(zerr, dJz, f):
     interp = interp1d(zerr, dJz, bounds_error=False, fill_value='extrapolate')
     def to_min(z):
-        return np.abs(interp(z)-0.1)
+        return np.abs(interp(z)-f)
     res = minimize(to_min, 250)
-    print('zerr needed for 10perc:', res.x)
+    print('zerr needed for ', str(100*f), 'perc:', res.x)
 
 for fname, name, Az, AR, Rg, init_vel, c in zip(fname_list, name_list, Az_list, AR_list, Rg_list, init_vel_list, clist):
     zout = pickle.load(open('zout_'+fname+'.p', 'rb'))
@@ -102,7 +102,8 @@ for fname, name, Az, AR, Rg, init_vel, c in zip(fname_list, name_list, Az_list, 
     dx[:,1] /= J1
     dx[:,2] /= J2
 
-    print_100(zoffset, dz[:,2])
+    print_100(zoffset, dz[:,2], 1)
+    print_100(zoffset, dz[:,2], 0.1)
     print(zoffset, dz[:,2])
     
     for x,o,d in zip(ax, (zoffset, xoffset), (dz, dx)):

### 2019-04-16 midplane_test/plot_deltaphi.py
Update horizontal lines on deltaphi range plot

@@ -24,7 +24,8 @@ Rsolar = 8.2
 
 glist = ['m12i', 'm12f', 'm12m']
 
-zoff_100_list = [30.7101346, 237.79862443]
+zoff_100_list = [61.47799653, 925.58008488]
+zoff_10_list = [6.19967419, 47.17989471]
 clist = [tb_c[7], tb_c[8]]
 
 theta = np.load('output/theta.npy')
@@ -34,7 +35,8 @@ def print_dphi(dphi, rng):
     def to_min(p):
         return np.abs(interp(p)-zoff_100_list[0]/1000)
     res = minimize(to_min, 0, bounds=[[0, 2.*np.pi]])
-    print(res.x)
+    print('dphi/pi:', res.x/np.pi)
+    return res.x
 
 def rotate(l, n):
     return np.append(l[n:], l[:n])
@@ -107,14 +109,15 @@ for gal,ax_col in zip(glist, ax.transpose()):
     ax_col.plot(dphi_mean/np.pi, r_up*1000, alpha=0.75, ls='dashed', c=tb_c[0])
     ax_col.plot(dphi_mean/np.pi, r_low*1000, alpha=0.75, ls='dashed', c=tb_c[0])
 
-    print_dphi(dphi_mean, r_mean)
-
     out = np.transpose([dphi_mean, r_mean, r_sigma])
     np.save('output/r_vs_dphi_'+gal+'.npy', out)
 
     # chord length
     def tick_function(l):
         return np.round(2*Rsolar*np.sin(0.5*(l/2)*np.pi), 1)
+
+    t = print_dphi(dphi_mean, r_mean)
+    print(tick_function(t/np.pi))
 
     ax2 = ax_col.twiny()
     #ax2.invert_xaxis()

### 2019-04-16 paper/ms.tex
Update numbers in plot for deltaphi plot

@@
 \ref{fig:range_deltaphi}). For the three simulations considered, the median
-range exceeds this value at $\Delta \phi/\pi > 0.036, 0.047,
-0.10\,\text{rad}$, with corresponding physical scales of $\lambda \sim 465,
-610, 1290\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. These can be
+range exceeds this value at $\Delta \phi/\pi > 0.072, 0.11,
+0.34\,\text{rad}$, with corresponding physical scales of $\lambda \sim 900,
+1400, 4300\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. These can be
 considered estimates of the range of validity for an extension of the locally
@@
 for the thin and thick orbit (Table~\ref{tab:orbits}). Nearby open clusters
-(Table~\ref{tab:real_clusters}) have orbits closest to the thin orbit .}
+(Table~\ref{tab:real_clusters}) have orbits closest to the thin orbit.}
 \label{fig:cluster_offset}

### 2019-04-16 paper/ms.tex
Add schmactions many orbit plot to paper, and update caption

@@
 lower} panels consider an offset in $x$ (equivalently, an offset in the Solar
-radius).}
+radius).
+In some panels we also plot as dashed lines the epicyclic prediction of the induced action error (Equations~\ref{eq:actions_epi_error}~and~\ref{eq:Jr_epi_dx}).
+In the epicyclic approximation, a $z$ offset only induces an error in $J_z$ --- for all three orbits the epicyclic approximation is a good description of the $J_z$ error.
+An $x$ offset induces an error in $J_r$ and $L_z$.
+The error in $J_r$ is somewhat well-described for the thin-disk and thick-disk orbits, and a poor description for the halo orbit.
+For $L_z$, the epicyclic approximation is not a good description for any orbit.}
 \label{fig:many_orbit_wrong_ref}

### 2019-04-18 paper/ms.tex
Add new chervin midplane plot to paper

--- 2019-04-16 paper/ms.tex
@@
 with neither of the modes peaking on the true action value. 
-A midplane error of $\sim6\,\pc$ for a thin disk orbit or
-$\sim50\,\pc$ for a thick disk orbit is all that is necessary to
-induce a $10\,\%$ error in the vertical action $J_z$.
-Furthermore, we show that the local midplane varies by
+A midplane error of $\sim30\,\pc$ for a thin disk orbit or
+$\sim225\,\pc$ for a thick disk orbit is all that is necessary to
+induce a $100\,\%$ error in the vertical action $J_z$.
+orbits, respectively. Furthermore, we show that the local midplane varies by
 $\sim100\,\pc$ at the Solar circle in Milky Way-mass cosmological zoom-in
@@
 
-\subsection{Epicyclic Approximation} \label{ssec:epi_action}
-Before turning to numerical methods, we first write down analytically the
-induced action error from either a midplane offset or a Galactocentric radius
-offset. We make the epicyclic approximation, which assumes that the motion in
-the $z$ and $R$ components of the orbit are decoupled and follow a simple
-harmonic oscillator. This approximation should be an excellent description of
-thin disk-like orbits and a good description of thick disk-like orbits. We
-also make the assumption of a perfectly flat circular velocity curve with
-velocity $v_c$.
-
-Under this approximation, we can write down the vertical and radial components
-of the orbits as,
-\beq\label{eq:orbits_epi}
-\begin{split}
-z(t) &= A_z \sin{(\kappa t + \alpha)}\\
-R(t) &= A_R \sin{(\nu t + \beta)}\text{,}
-\end{split}
-\eeq
-where $\kappa$ and $\nu$ are the vertical and radial (or epicyclic)
-frequencies, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
-vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
-constants. We can write down the actions in this approximation as
-\citep[Section~3.5.3b][]{2008gady.book.....B},
-\beq\label{eq:actions_epi}
-\begin{split}
-L_z &= R_g v_c \\
-J_r &= \kappa A_R^2 \\
-J_z &= \nu A_z^2\text{,}
-\end{split}
-\eeq
-where $R_g$ is the guiding radius of the orbit.
-
-An induced error in $z$ of $\Delta z$ will cause $\Delta A_z \sim \Delta z$.
-Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$. If
-we assume that $\kappa$ and $\nu$ are constant (\Gus{justify?}), then we have
-that
-\beq\label{eq:actions_epi_error}
-\begin{split}
-\Delta L_z = \Delta R v_c &\implies \frac{\Delta L_z}{L_z} = \frac{\Delta R}{R} \\
-\Delta J_r = 2 \kappa \Delta A_R &\implies \frac{\Delta J_r}{J_r} = \frac{2\Delta R}{A_R} \\
-\Delta J_z = 2 \nu \Delta A_z &\implies \frac{\Delta J_z}{J_z} = \frac{2\Delta z}{A_z}\text{.}
-\end{split}
-\eeq
-Therefore, errors in each action are linearly proportional to the errors in
-either the midplane or cylindrical radius.
-
-There is a slight complication in the formula for computing $\Delta J_r/J_r$ and $\Delta L_z/L_z$.
-In a moment we will consider the effect of an inaccuracy in the distance from
-the Sun to the Galactic center on actions. However, this inaccuracy induces an
-error in $x$, or $\Delta x$. This is not the same as an inaccuracy in $\Delta
-R$, but they are related by the following argument. An error $\Delta x$ will
-induce an erroneous radius $R_{\text{err}}$ related by the formula,
-\beq
-(x+\Delta x)^2 + y^2 = R_{\text{err}}^2\text{.}
-\eeq
-Keeping only terms to first order in $\Delta x$, we have that,
-\beq
-\begin{split}
-R_{\text{err}}^2 &= R^2 - 2 R \cos{\theta} \Delta x \\
-\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\theta}} \Delta x\text{.}
-\end{split}
-\eeq
-Averaging over the circle, we therefore have that
-\beq
-\avg{\Delta R} = \frac{2}{\pi} \Delta x\text{.}
-\eeq
-Therefore, the error in $J_r$ and $L_z$ are related to $\Delta x$ by,
-\beq\label{eq:Jr_epi_dx}
-\begin{split}
-\frac{\Delta J_r}{J_r} &= \frac{4 \Delta x}{\pi A_R} \\
-\frac{\Delta L_z}{L_z} &= \frac{2 \Delta x}{\pi R}\text{.}
-\end{split}
-\eeq
-
-We will check these formulae against numerical tests next.
-
 \subsection{Numerical methods} \label{ssec:action_comp}
@@
 
-\begin{deluxetable*}{cccccccc}
+\begin{deluxetable*}{ccccccc}
 \tablecaption{Description and names of the three orbits considered in this
-work. For the last two columns: $z_{\text{max}}$ the maximum height of the
-orbit while $A_R$ is the magnitude of the radial excursions of the orbit,
-computed as one half the maximum radius minus the minimum
-radius.\label{tab:orbits}}
+work.\label{tab:orbits}}
 \tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
 velocity} & \colhead{$J_r$} &
-\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
-\colhead{$A_R$}\\ \colhead{ } &
+\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$}\\ \colhead{ } &
 \colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
@@
 \colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$}}
+\colhead{$\mathrm{kpc}$}}
 \startdata 
-thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 & 1.29 \\
-thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 & 1.19 \\ 
-halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16 & 2.34
+thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 \\
+thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 \\ 
+halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16
 \enddata
@@
 induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
-sort of error propagation must take this into account. We also plot one half
-the 95\uth{} percentile minus the 5\uth{} percentile of each distribution as a
-horizontal arrow anchored on the true $J_z$ value. We call this $\Delta J_z$
-and will use it (along with the similarly defined $\Delta J_r$ and $\Delta
-L_z$) to empirically describe the error distribution. We see that $\Delta J_z$
-roughly corresponds to the distance from the true $J_z$ value to one of the
-modes of the error distribution.}
+sort of error propagation must take this into account.}
 \label{fig:Jz_hist}
@@
 Figure~\ref{fig:Jz_hist} is difficult. Of course, Gaussian summary statistics
-are not adequate. We therefore elect to measure this error by computing one half the
+are not adequate. We therefore elect to measure this error by computing the
 95\uth{} percentile minus the 5\uth{} percentile of the distribution of
 observed action values. We refer to this quantity as $\Delta J_i$ for each
-action value. We plot this quantity in Figure~\ref{fig:Jz_hist} as a horizontal arrow anchored on the true action value. Because of the bimodality of the error distribution, this
-quantity roughly measures the distance from the true action value to the peak of one of the modes.
+action value. Because of the bimodality of the error distribution, this
+quantity roughly measures the distance between the peaks of the two modes.
 Furthermore, this bimodality also implies that $\Delta J_i$ is not very
@@
 but for systematic offsets between $0$ and $500\,\pc$ in the $z$ and $x$
-components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report one half the 95\uth{}
-minus 5\uth{} percentile divided by the true action value ($\Delta J_i/J_i$) for the three
+components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth{}
+minus 5\uth{} percentile divided by the true action value for the three
 different fiducial orbits in Table \ref{tab:orbits}. The thick-disk
@@
 show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
-respectively. We compute the fractional spread $\Delta J_i/J_i$ as one half the 95\uth{}
+respectively. We compute the fractional spread $\Delta J_i/J_i$ as the 95\uth{}
 minus 5\uth{} percentile of the action computed using the offset coordinate
@@
 \plotone{fig/schmactions_many_orbits.pdf}
-\caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error in the
+\caption{We report the 95\uth{} minus 5\uth{} percentile of the error in the
 measured action ($\Delta J_i$) from coordinate system errors for the thin,
-thick, and halo orbits (Table~\ref{tab:orbits}). See discussion in the text for the justification in using this to measure the magnitude of the induced error. The {\em left}, {\em center},
+thick, and halo orbits (Table~\ref{tab:orbits}). The {\em left}, {\em center},
 and {\em right} panels show the result for $J_r$, $L_z$, and $J_z$,
@@
 different offsets in $z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
-(red, cyan, green respectively). As before, the error ($\Delta J_z$) is one half the
+(red, cyan, green respectively). As before, the error ($\Delta J_z$) is the
 95\uth{} minus 5\uth{} percentile of the distribution of observed $J_z$ values
@@
 \begin{figure*}
-\plotone{fig/midplane_fit_chervinsim.pdf}
+\begin{center}
+\includegraphics[width=5.5in]{fig/midplane_fit_chervinsim.pdf}
+\end{center}
 \caption{The local midplane determined at the fiducial Solar radius

### 2019-04-19 paper/ms.tex
Resize FIRE midplane plot
Add other affiliations

@@
 120th St., New York, NY 10027, USA}
+\newcommand{\victoria}{Departmeny of Physics \& Astronomy, University of
+Victoria, 3800 Finnerty Rd., Victoria, BC, V8P 4HN, Canada}
+\newcommand{\nyuphys}{Center for Cosmology and Particle Physics, Department of
+Physics, New York University, 726 Broadway, New York, NY 10003, USA}
+\newcommand{\nyucds}{Center for Data Science, New York University, 60 Fifth
+Ave., New York, NY 10011, USA}
+\newcommand{\mpia}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17,
+69117 Heidelberg, Germany}
 
@@
 \affil{\cca}
+
+\authors{others}
 
@@
 \begin{figure*}
-\plotone{fig/midplane_fit.pdf}
+\begin{center}
+\includegraphics[width=7in]{fig/midplane_fit.pdf}
+\end{center}
 \caption{The local midplane determined at the fiducial Solar radius

### 2019-04-20 paper/ms.tex
Fix previous commit
Update text to change to 10% error as neighborhood

@@
 
-\authors{others}
+\author{others}
 
@@
 of the density fluctuations in the Galactic disk will determine how far this
-system can be extended before introducing $\mathcal{O}(1)$ systematic errors
-in the computation of actions for thin-disk stars. We consider this a
+coordinate system can be extended before introducing relevant systematic errors
+in the computation of actions for thin-disk stars. We consider a $10\%$ error to be a
 criterion for the region around the Sun where the analysis of the action-space
@@
 studies of the chemical distribution of the Galaxy, we propose the phrase
-``dynamical solar neighborhood.''
+``dynamical solar neighborhood.'' We also point out that our use of a $10\%$ error is arbitrary.
 
@@
 The horizontal \thincolor{} and \thickcolor{} lines shows the midplane
-variation necessary to induce a $100\%$ error in $J_z$ for the thin-disk and
+variation necessary to induce a $10\%$ error in $J_z$ for the thin-disk and
 thick-disk orbits, respectively. A cartoon explanation of this Figure is given
@@
 from the Sun). At far smaller $\Delta \phi$ it reaches the level of deviation
-required for an $\mathcal{O}(1)$ systematic error in $J_z$ for our fiducial
+required for a $10\%$ systematic error in $J_z$ for our fiducial
 thin-disk-like orbit (\thincolor{} horizontal line in Figure
 \ref{fig:range_deltaphi}). For the three simulations considered, the median
-range exceeds this value at $\Delta \phi/\pi > 0.072, 0.11,
-0.34\,\text{rad}$, with corresponding physical scales of $\lambda \sim 900,
-1400, 4300\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. These can be
+range exceeds this value at $\Delta \phi/\pi > x, y,
+z\,\text{rad}$, with corresponding physical scales of $\lambda \sim a,
+b, c\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. These can be
 considered estimates of the range of validity for an extension of the locally
@@
 The situation is more forgiving for the study of thick-disk-like orbits, where
-an $\mathcal{O}(1)$ systematic error in $J_z$ requires a larger midplane
+a $10\%$ systematic error in $J_z$ requires a larger midplane
 deviation (\thickcolor{} horizontal line in Figure~\ref{fig:range_deltaphi}).
@@
 since the midplane deviations are relatively small compared to those needed to
-produce $\mathcal{O}(1)$ systematic errors in the actions (\halocolor{}
+produce $10\%$ systematic errors in the actions (\halocolor{}
 horizontal line in Figure \ref{fig:range_deltaphi}). However, we caution that
@@
 disk stars, where the necessary midplane deviation is
-$\sim50\textup{--}100\,\pc$.
+$\sim5\textup{--}10\,\pc$.
 
@@
 cluster masses, except at the very low-mass end in \mi . The neighborhood
-around $\sim100\Msun$ clusters can be as small as $\sim300\,\pc$ but up to
-$\sim1.2\,\kpc$ for \mi{} and \mf{}, though the situation is marginally better
-for \mm{} with values ranging from $\sim0.8\,\kpc$ to $\sim3\,\kpc$. The
-neighborhood values steadily increase with $m_c$, topping out at $\sim5\,\kpc$
+around $\sim100\Msun$ clusters can be as small as $\sim x\,\pc$ but up to
+$\sim y\,\kpc$ for \mi{} and \mf{}, though the situation is marginally better
+for \mm{} with values ranging from $\sim a\,\kpc$ to $\sim b\,\kpc$. The
+neighborhood values steadily increase with $m_c$, topping out at $\sim c\,\kpc$
 for $\sim10^4\,\Msun$ clusters in
 \mi{} and \mf{}. In \mm{}, the neighborhood size reaches the maximum allowed of
-$16.4\,\kpc$ for clusters of mass $\sim800\,\Msun$. This figure also
+$16.4\,\kpc$ for clusters of mass $\sim m\,\Msun$. This figure also
 underlines that the degree to which such an analysis will succeed depends
@@
 Table~\ref{tab:real_clusters}. Except for alphaPer, we find that the
-neigbhorhood around each cluster is typically $\gtrsim 500\,\pc$. This
+neigbhorhood around each cluster is typically $\gtrsim x\,\pc$. This
 analysis assumes that the MW resembles \mi{}, which is not necessarily true.
@@
 should be possible in the near future with {\em Gaia}, with quality distance
-estimates the crucial limiting factor.
+estimates the crucial limiting factor, along with careful selction function modelling.
 
@@
 inaccuracy in the midplane has on the vertical action $J_z$. We found that in
-order for there to be a $100\%$ error in $J_z$ (as defined by the middle
-$90\uth$ percentile) one needs a midplane error of $\sim30\,\pc$ and
-$\sim225\,\pc$ for a typical thin and thick disk orbit, respectively. The
+order for there to be a $10\%$ error in $J_z$ (as defined by one half the middle
+$90\uth$ percentile) one needs a midplane error of $\sim6\,\pc$ and
+$\sim50\,\pc$ for a typical thin and thick disk orbit, respectively. The
 fractional error is significantly less for halo orbits.

### 2019-04-22 paper/ms.tex
revised abstract

--- 2019-04-19 paper/ms.tex
@@
+%mm substantive Mac Low changes marked so.  Minor changes done silently.
+
 \documentclass[twocolumn]{aastex62}
@@
 % \correspondingauthor{Angus Beane}
+%mm [will this address go away during the review process?.  Do you
+%want to give a "current address" as a second footnote?]
 \email{abeane@sas.upenn.edu}
@@
 principle, actions have the attractive quality of being invariant and thus
-good labels of a star's orbit. Of course, inaccurate measurements of the 
-phase space position and uncertainities in the assumed Galactic potential
+good labels of a star's orbit. 
+%mm Of course,
+    Such analyses are often used to interpret data from the {\em Gaia}
+    mission. However, 
+inaccurate measurements of the 
+phase space position and uncertainties in the assumed Galactic potential
 induce an error in the computed action. We show that, in addition to these
 complications, a systematic bias in the phase space position induces a
-phase-dependence in the actions, which we interpret as an error. The induced error distribution is
+phase-dependence in the actions, which we interpret as 
+%mm an 
+    a systematic
+error. The induced error distribution is
 non-Gaussian and bimodal, 
-with neither of the modes peaking on the true action value. 
+with neither of the modes peaking on
+%mm  the true action value. [the error in a Gaussian peaks on zero] 
+    the null value.
 A midplane error of $\sim30\,\pc$ for a thin disk orbit or
-$\sim225\,\pc$ for a thick disk orbit is all that is necessary to
-induce a $100\,\%$ error in the vertical action $J_z$.
-orbits, respectively. Furthermore, we show that the local midplane varies by
+$\sim225\,\pc$ for a thick disk orbit is 
+%mm all that is necessary 
+      sufficient
+to induce a $100\,\%$ error in the vertical action $J_z$.
+%mm [fragment?] orbits, respectively. 
+Furthermore, we show that the local midplane varies by
 $\sim100\,\pc$ at the Solar circle in Milky Way-mass cosmological zoom-in
@@
 calculations --- which assume the global and local midplanes are the same ---
-include a systematic offset in $z$. Variation in the local standard of rest
-will induce similar issues. The variation of the midplane must be taken into
-account when performing dynamical modelling across large regions of the disk
-accessible with the {\em Gaia} mission.
+include a systematic 
+%mm offset in $z$.
+   vertical offset variation. 
+Variation in the local standard of rest
+%mm will induce 
+     induces
+similar issues. The variation of the midplane must be taken into
+account when performing dynamical modelling across
+the large regions of the disk
+accessible 
+%mm with the {\em Gaia} mission. [moved first mention up to top of
+%abstract, given that this is in the title.]
+    to {\em Gaia}.
 

### 2019-04-22 paper/ms.tex
revisions in introduction

@@
 \section{Introduction} \label{sec:intro}
-Our understanding of the Milky Way is currently undergoing a revolution in the
-era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
+Our understanding of the Milky Way is currently undergoing a
+revolution 
+%mm in the era of {\em Gaia} DR2. 
+     as a result of {\em Gaia} Data Release 2 (DR2).  Recent major discoveries include the remnants of a
 major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
-2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the
+2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the 2019arXiv190405721A
 solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
@@
 stellar motions near Sgr~A\textsuperscript{*}, the distance from the Sun to
-the Galactic center is taken to be $\sim 8.1\text{--}8.3\,\kpc$
-\citep{2009ApJ...692.1075G,2018AA...615L..15G}, with a recent $\sim0.3\%$ measurement from the GRAVITY collaboration \citep{2019arXiv190405721A}. The Sun's vertical distance
+the Galactic center
+%mm is taken to be $\sim 8.1\text{--}8.3\,\kpc$
+    can be measured
+\citep{2009ApJ...692.1075G,2018AA...615L..15G}, with a recent 
+%mm $\sim0.3\%$ measurement from the GRAVITY collaboration
+    measurement using near-infrared interferometry of $8.178 \pm
+    0.035$~kpc 
+\citep{2019arXiv190405721A}. The Sun's vertical distance
 from the Galactic midplane, the other independent measurement that sets this
@@
 model the stellar distribution function, or find stars with similar dynamical
-properties. Actions are conserved quantities which describe the orbit of a
+properties. Actions are conserved quantities that describe the orbit of a
 star under the assumption of regular, bound orbits in a system where the
 equations of motion are separable in a particular coordinate system. They are
-the typical integral of the Hamiltonian in cylindrical coordinates and are
+the %mm typical
+ integral of the Hamiltonian in cylindrical coordinates and are
 given by,

### 2019-04-22 paper/ms.tex
completed initial revisions to introduction

@@
 
-Once a Galactocentric coordinate system has been established and a 6D phase
+Once a Galactocentric coordinate system has been established and a
+six-dimensional (6D) phase
 space measurement of a star has been made, it is often desirable to convert
@@
 equations of motion are separable in a particular coordinate system. They are
-the %mm typical
- integral of the Hamiltonian in cylindrical coordinates and are
-given by,
+the 
+%mm typical
+integral of the Hamiltonian in cylindrical coordinates 
+%mm and are given by,
 \beq\label{eq:actions}
@@
 respectively in a cylindrical coordinate system and the $p_i$ are the
-conjugate momenta.\footnote{We typically write $J_r$ instead of $J_R$ for
+conjugate momenta.\footnote{We 
+%mm typically [there are no other J_R's in manuscript]
+write $J_r$ instead of $J_R$ for
 aesthetic reasons.} In a slowly-evolving axisymmetric potential, these actions
@@
 individual stellar orbit is concisely described by three actions, as opposed
-to six dimensions of phase space. Second, under the assumption of a
+to six phase space coordinates. Second, under the assumption of a
 phase-mixed system, the dynamical properties of a population of stars should
@@
 position and the potential, and requires relatively strong assumptions about
-the system's symmetry. For instance, the Galactic potential is not strictly
+the system's symmetry. 
+%mm For instance, 
+    However, 
+the Galactic potential is not strictly
 axisymmetric; with the vast improvement in the quality of phase-space
 measurements due to {\em Gaia}, this assumption appears increasingly
-inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this assumption is
+inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this
+assumption 
+%mm is [counterfactual subjunctive, I believe]
+     were
 close enough for many purposes, the parameters used in axisymmetric models may
@@
 coordinate frame can be usefully extended for action-space analysis. A similar
-argument applies to uncertainty in measurements of $R_{\odot}$\footnote{Here,
-and throughout, we use $R_{\odot}$ to mean the distance from the Sun to the
-Galactic center.} and variations in the local standard of rest (LSR).
+argument applies to 
+%mm [given that this is down to 0.3% error, seems a bit academic]
+   any remaining 
+uncertainty in measurements of 
+%mm $R_{\odot}$\footnote{Here,
+%and throughout, we use $R_{\odot}$ to mean the distance from the Sun to the
+%Galactic center.}
+     the Galactocentric radius of the Sun $R_{\odot}$
+and to variations in the local standard of rest (LSR).
 
@@
 midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire}
-we attempt to quantify the azimuthal dependence of the local midplane, and
+we 
+%mm attempt to quantify the 
+    examine the expected 
+azimuthal dependence of the local midplane, and
 hence the expected size of the dynamical solar neighborhood, using
-cosmological-hydrodynamical zoom-in simulations of Milky Way-mass galaxies
+cosmological, hydrodynamical, zoom-in simulations of Milky Way-mass galaxies
 from the Feedback in Realistic Environments (FIRE)
@@
 consider the feasibility of one science goal that relies on the invariance of
-actions --- the reconstruction of open clusters across large distances using
+actions---the reconstruction of open clusters across large distances using
 ``dynamical tagging;'' that is, the association of stars to a birth cluster

### 2019-04-22 paper/ms.tex
Gaussian error distributions peak at zero, not the true value.  I have revised accordingly throughout

@@
 of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted
-as a vertical dashed line. Here we see that the spread in $J_z$ induced by a
-systematic offset in $z$ is non-Gaussian and bimodal. Furthermore, neither of
-the modes lie at the true value of $J_z$. In the case of the thin-disk orbit
+as a vertical dashed line. Here we see that the 
+%mm spread 
+    errors 
+in $J_z$ induced by a
+systematic offset in $z$ are non-Gaussian and bimodal. Furthermore, neither of
+the modes lie %mm at the true value of $J_z$.
+   at null. 
+In the case of the thin-disk orbit
 ({\em lower panel}), we see that, in addition to the prior complications, the
-error distribution is not even centered on the true value. This comes about
+error distribution is not even centered on 
+%mm the true value. 
+    null. 
+This comes about
 when the midplane error is $\gtrsim z_{\text{max}}$, and can be understood by
@@
 coordinate system error is very non-Gaussian. The distribution is bimodal with
-{\em neither of the modes peaking at the true value}. As a result, error
-propagation is very involved when considering actions.
+{\em neither of the modes peaking at 
+%mm the true value}.
+     null}. 
+As a result, error
+propagation is 
+%mm very involved 
+     complex
+when considering actions.
 

### 2019-04-22 paper/ms.tex
added M-M grant

@@
 
-Add sentences about distribution function modeling
+{\bf Add sentences about distribution function modeling}
 
@@
 \acknowledgments
-We would like to thank Megan Bedell, Tobias Buck, David W. Hogg, and Kathryn
-Johnston for many useful discussions. A.B. would like to thank Todd Phillips
+We would like to thank Megan Bedell, Tobias Buck, and Todd Phillips
+%mm [moving to author list] David W. Hogg, and Kathryn
+%Johnston 
+%for many useful discussions. A.B. would like to thank Todd Phillips
 for helpful discussions. A.B. was supported in part by the Roy \& Diana
 Vagelos Program in the Molecular Life Sciences and the Roy \& Diana Vagelos
-Challenge Award. The work of \Gus{everyone} is supported by the Simons
-Foundation. \Gus{remove last sentence if everyone CCA affiliated}
+Challenge Award. 
+%mm The work of \Gus{everyone} is supported by the Simons
+%Foundation. \Gus{remove last sentence if everyone CCA affiliated}
+M.-M.M.L. was partly supported by NSF grant AST18-15461.
 

### 2019-04-22 paper/ms.tex
wording changes to abstract and intro from meeting discussion

@@
     the null value.
-A midplane error of $\sim30\,\pc$ for a thin disk orbit or
+A variation in the vertical position midplane of $\sim30\,\pc$ for a thin disk orbit or
 $\sim225\,\pc$ for a thick disk orbit is 
@@
       sufficient
-to induce a $100\,\%$ error in the vertical action $J_z$.
+to induce a $100\,\%$ systematic error in the vertical action $J_z$ computed under the assumption of an axisymmetric potential.
 %mm [fragment?] orbits, respectively. 
@@
 %mm typical
-integral of the Hamiltonian in cylindrical coordinates 
+cyclical integral of the canonical momentum  over its conjugate position,
 %mm and are given by,
@@
 J_i \equiv
-\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
+\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,.}
 \eeq
-where $i=R,z,\phi$ are the radial, vertical, and azimuthal coordinates
+Under the assumption of axisymmetry, $i=R,z,\phi$ are the radial, vertical, and azimuthal coordinates
 respectively in a cylindrical coordinate system and the $p_i$ are the

### 2019-04-22 paper/ms.tex
change to title from meeting discussion
Change range dphi plot to 10% error

--- 2019-04-22 paper/ms.tex
@@
-%mm substantive Mac Low changes marked so.  Minor changes done silently.
-
 \documentclass[twocolumn]{aastex62}
@@
 120th St., New York, NY 10027, USA}
-\newcommand{\victoria}{Departmeny of Physics \& Astronomy, University of
+\newcommand{\victoria}{Department of Physics \& Astronomy, University of
 Victoria, 3800 Finnerty Rd., Victoria, BC, V8P 4HN, Canada}
@@
 % \correspondingauthor{Angus Beane}
-%mm [will this address go away during the review process?.  Do you
-%want to give a "current address" as a second footnote?]
 \email{abeane@sas.upenn.edu}
@@
 principle, actions have the attractive quality of being invariant and thus
-good labels of a star's orbit. 
-%mm Of course,
-    Such analyses are often used to interpret data from the {\em Gaia}
-    mission. However, 
-inaccurate measurements of the 
-phase space position and uncertainties in the assumed Galactic potential
+good labels of a star's orbit. Of course, inaccurate measurements of the 
+phase space position and uncertainities in the assumed Galactic potential
 induce an error in the computed action. We show that, in addition to these
 complications, a systematic bias in the phase space position induces a
-phase-dependence in the actions, which we interpret as 
-%mm an 
-    a systematic
-error. The induced error distribution is
+phase-dependence in the actions, which we interpret as an error. The induced error distribution is
 non-Gaussian and bimodal, 
-with neither of the modes peaking on
-%mm  the true action value. [the error in a Gaussian peaks on zero] 
-    the null value.
-A midplane error of $\sim30\,\pc$ for a thin disk orbit or
-$\sim225\,\pc$ for a thick disk orbit is 
-%mm all that is necessary 
-      sufficient
-to induce a $100\,\%$ error in the vertical action $J_z$.
-%mm [fragment?] orbits, respectively. 
+with neither of the modes peaking on the true action value. 
+A midplane error of $\sim6\,\pc$ for a thin disk orbit or
+$\sim50\,\pc$ for a thick disk orbit is all that is necessary to
+induce a $10\,\%$ error in the vertical action $J_z$.
 Furthermore, we show that the local midplane varies by
@@
 calculations --- which assume the global and local midplanes are the same ---
-include a systematic 
-%mm offset in $z$.
-   vertical offset variation. 
-Variation in the local standard of rest
-%mm will induce 
-     induces
-similar issues. The variation of the midplane must be taken into
-account when performing dynamical modelling across
-the large regions of the disk
-accessible 
-%mm with the {\em Gaia} mission. [moved first mention up to top of
-%abstract, given that this is in the title.]
-    to {\em Gaia}.
+include a systematic offset in $z$. Variation in the local standard of rest
+will induce similar issues. The variation of the midplane must be taken into
+account when performing dynamical modelling across large regions of the disk
+accessible with the {\em Gaia} mission.
 
@@
 \section{Introduction} \label{sec:intro}
-Our understanding of the Milky Way is currently undergoing a
-revolution 
-%mm in the era of {\em Gaia} DR2. 
-     as a result of {\em Gaia} Data Release 2 (DR2).  Recent major discoveries include the remnants of a
+Our understanding of the Milky Way is currently undergoing a revolution in the
+era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
 major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
-2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the 2019arXiv190405721A
+2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the
 solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
@@
 stellar motions near Sgr~A\textsuperscript{*}, the distance from the Sun to
-the Galactic center
-%mm is taken to be $\sim 8.1\text{--}8.3\,\kpc$
-    can be measured
-\citep{2009ApJ...692.1075G,2018AA...615L..15G}, with a recent 
-%mm $\sim0.3\%$ measurement from the GRAVITY collaboration
-    measurement using near-infrared interferometry of $8.178 \pm
-    0.035$~kpc 
-\citep{2019arXiv190405721A}. The Sun's vertical distance
+the Galactic center is taken to be $\sim 8.1\text{--}8.3\,\kpc$
+\citep{2009ApJ...692.1075G,2018AA...615L..15G}, with a recent $\sim0.3\%$ measurement from the GRAVITY collaboration \citep{2019arXiv190405721A}. The Sun's vertical distance
 from the Galactic midplane, the other independent measurement that sets this
@@
 
-Once a Galactocentric coordinate system has been established and a
-six-dimensional (6D) phase
+Once a Galactocentric coordinate system has been established and a 6D phase
 space measurement of a star has been made, it is often desirable to convert
@@
 model the stellar distribution function, or find stars with similar dynamical
-properties. Actions are conserved quantities that describe the orbit of a
+properties. Actions are conserved quantities which describe the orbit of a
 star under the assumption of regular, bound orbits in a system where the
 equations of motion are separable in a particular coordinate system. They are
-the 
-%mm typical
-integral of the Hamiltonian in cylindrical coordinates 
-%mm and are given by,
+the typical integral of the Hamiltonian in cylindrical coordinates and are
+given by,
 \beq\label{eq:actions}
@@
 respectively in a cylindrical coordinate system and the $p_i$ are the
-conjugate momenta.\footnote{We 
-%mm typically [there are no other J_R's in manuscript]
-write $J_r$ instead of $J_R$ for
+conjugate momenta.\footnote{We typically write $J_r$ instead of $J_R$ for
 aesthetic reasons.} In a slowly-evolving axisymmetric potential, these actions
@@
 individual stellar orbit is concisely described by three actions, as opposed
-to six phase space coordinates. Second, under the assumption of a
+to six dimensions of phase space. Second, under the assumption of a
 phase-mixed system, the dynamical properties of a population of stars should
@@
 position and the potential, and requires relatively strong assumptions about
-the system's symmetry. 
-%mm For instance, 
-    However, 
-the Galactic potential is not strictly
+the system's symmetry. For instance, the Galactic potential is not strictly
 axisymmetric; with the vast improvement in the quality of phase-space
 measurements due to {\em Gaia}, this assumption appears increasingly
-inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this
-assumption 
-%mm is [counterfactual subjunctive, I believe]
-     were
+inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this assumption is
 close enough for many purposes, the parameters used in axisymmetric models may
@@
 coordinate frame can be usefully extended for action-space analysis. A similar
-argument applies to 
-%mm [given that this is down to 0.3% error, seems a bit academic]
-   any remaining 
-uncertainty in measurements of 
-%mm $R_{\odot}$\footnote{Here,
-%and throughout, we use $R_{\odot}$ to mean the distance from the Sun to the
-%Galactic center.}
-     the Galactocentric radius of the Sun $R_{\odot}$
-and to variations in the local standard of rest (LSR).
+argument applies to uncertainty in measurements of $R_{\odot}$\footnote{Here,
+and throughout, we use $R_{\odot}$ to mean the distance from the Sun to the
+Galactic center.} and variations in the local standard of rest (LSR).
 
@@
 midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire}
-we 
-%mm attempt to quantify the 
-    examine the expected 
-azimuthal dependence of the local midplane, and
+we attempt to quantify the azimuthal dependence of the local midplane, and
 hence the expected size of the dynamical solar neighborhood, using
-cosmological, hydrodynamical, zoom-in simulations of Milky Way-mass galaxies
+cosmological-hydrodynamical zoom-in simulations of Milky Way-mass galaxies
 from the Feedback in Realistic Environments (FIRE)
@@
 consider the feasibility of one science goal that relies on the invariance of
-actions---the reconstruction of open clusters across large distances using
+actions --- the reconstruction of open clusters across large distances using
 ``dynamical tagging;'' that is, the association of stars to a birth cluster
@@
 
+\subsection{Epicyclic Approximation} \label{ssec:epi_action}
+Before turning to numerical methods, we first write down analytically the
+induced action error from either a midplane offset or a Galactocentric radius
+offset. We make the epicyclic approximation, which assumes that the motion in
+the $z$ and $R$ components of the orbit are decoupled and follow a simple
+harmonic oscillator. This approximation should be an excellent description of
+thin disk-like orbits and a good description of thick disk-like orbits. We
+also make the assumption of a perfectly flat circular velocity curve with
+velocity $v_c$.
+
+Under this approximation, we can write down the vertical and radial components
+of the orbits as,
+\beq\label{eq:orbits_epi}
+\begin{split}
+z(t) &= A_z \sin{(\kappa t + \alpha)}\\
+R(t) &= A_R \sin{(\nu t + \beta)}\text{,}
+\end{split}
+\eeq
+where $\kappa$ and $\nu$ are the vertical and radial (or epicyclic)
+frequencies, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
+vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
+constants. We can write down the actions in this approximation as
+\citep[Section~3.5.3b][]{2008gady.book.....B},
+\beq\label{eq:actions_epi}
+\begin{split}
+L_z &= R_g v_c \\
+J_r &= \kappa A_R^2 \\
+J_z &= \nu A_z^2\text{,}
+\end{split}
+\eeq
+where $R_g$ is the guiding radius of the orbit.
+
+An induced error in $z$ of $\Delta z$ will cause $\Delta A_z \sim \Delta z$.
+Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$. If
+we assume that $\kappa$ and $\nu$ are constant (\Gus{justify?}), then we have
+that
+\beq\label{eq:actions_epi_error}
+\begin{split}
+\Delta L_z = \Delta R v_c &\implies \frac{\Delta L_z}{L_z} = \frac{\Delta R}{R} \\
+\Delta J_r = 2 \kappa \Delta A_R &\implies \frac{\Delta J_r}{J_r} = \frac{2\Delta R}{A_R} \\
+\Delta J_z = 2 \nu \Delta A_z &\implies \frac{\Delta J_z}{J_z} = \frac{2\Delta z}{A_z}\text{.}
+\end{split}
+\eeq
+Therefore, errors in each action are linearly proportional to the errors in
+either the midplane or cylindrical radius.
+
+There is a slight complication in the formula for computing $\Delta J_r/J_r$ and $\Delta L_z/L_z$.
+In a moment we will consider the effect of an inaccuracy in the distance from
+the Sun to the Galactic center on actions. However, this inaccuracy induces an
+error in $x$, or $\Delta x$. This is not the same as an inaccuracy in $\Delta
+R$, but they are related by the following argument. An error $\Delta x$ will
+induce an erroneous radius $R_{\text{err}}$ related by the formula,
+\beq
+(x+\Delta x)^2 + y^2 = R_{\text{err}}^2\text{.}
+\eeq
+Keeping only terms to first order in $\Delta x$, we have that,
+\beq
+\begin{split}
+R_{\text{err}}^2 &= R^2 - 2 R \cos{\theta} \Delta x \\
+\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\theta}} \Delta x\text{.}
+\end{split}
+\eeq
+Averaging over the circle, we therefore have that
+\beq
+\avg{\Delta R} = \frac{2}{\pi} \Delta x\text{.}
+\eeq
+Therefore, the error in $J_r$ and $L_z$ are related to $\Delta x$ by,
+\beq\label{eq:Jr_epi_dx}
+\begin{split}
+\frac{\Delta J_r}{J_r} &= \frac{4 \Delta x}{\pi A_R} \\
+\frac{\Delta L_z}{L_z} &= \frac{2 \Delta x}{\pi R}\text{.}
+\end{split}
+\eeq
+
+We will check these formulae against numerical tests next.
+
 \subsection{Numerical methods} \label{ssec:action_comp}
@@
 
-\begin{deluxetable*}{ccccccc}
+\begin{deluxetable*}{cccccccc}
 \tablecaption{Description and names of the three orbits considered in this
-work.\label{tab:orbits}}
+work. For the last two columns: $z_{\text{max}}$ the maximum height of the
+orbit while $A_R$ is the magnitude of the radial excursions of the orbit,
+computed as one half the maximum radius minus the minimum
+radius.\label{tab:orbits}}
 \tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
 velocity} & \colhead{$J_r$} &
-\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$}\\ \colhead{ } &
+\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
+\colhead{$A_R$}\\ \colhead{ } &
 \colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
@@
 of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted
-as a vertical dashed line. Here we see that the 
-%mm spread 
-    errors 
-in $J_z$ induced by a
-systematic offset in $z$ are non-Gaussian and bimodal. Furthermore, neither of
-the modes lie %mm at the true value of $J_z$.
-   at null. 
-In the case of the thin-disk orbit
+as a vertical dashed line. Here we see that the spread in $J_z$ induced by a
+systematic offset in $z$ is non-Gaussian and bimodal. Furthermore, neither of
+the modes lie at the true value of $J_z$. In the case of the thin-disk orbit
 ({\em lower panel}), we see that, in addition to the prior complications, the
-error distribution is not even centered on 
-%mm the true value. 
-    null. 
-This comes about
+error distribution is not even centered on the true value. This comes about
 when the midplane error is $\gtrsim z_{\text{max}}$, and can be understood by
@@
 from the Sun). At far smaller $\Delta \phi$ it reaches the level of deviation
-required for an $\mathcal{O}(1)$ systematic error in $J_z$ for our fiducial
+required for a $10\%$ systematic error in $J_z$ for our fiducial
 thin-disk-like orbit (\thincolor{} horizontal line in Figure
 \ref{fig:range_deltaphi}). For the three simulations considered, the median
-range exceeds this value at $\Delta \phi/\pi > 0.036, 0.047,
-0.10\,\text{rad}$, with corresponding physical scales of $\lambda \sim 465,
-610, 1290\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. These can be
+range exceeds this value at $\Delta \phi/\pi > 0.0073, 0.0096,
+0.018\,\text{rad}$, with corresponding physical scales of $\lambda \sim 93.9,
+123, 233\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. These can be
 considered estimates of the range of validity for an extension of the locally
@@
 coordinate system error is very non-Gaussian. The distribution is bimodal with
-{\em neither of the modes peaking at 
-%mm the true value}.
-     null}. 
-As a result, error
-propagation is 
-%mm very involved 
-     complex
-when considering actions.
+{\em neither of the modes peaking at the true value}. As a result, error
+propagation is very involved when considering actions.
 

### 2019-04-22 midplane_test/FIRE/plot_deltaphi.py

@@ -33,7 +33,7 @@ theta = np.load('output/theta.npy')
 def print_dphi(dphi, rng):
     interp = interp1d(dphi, rng)
     def to_min(p):
-        return np.abs(interp(p)-zoff_100_list[0]/1000)
+        return np.abs(interp(p)-zoff_10_list[0]/1000)
     res = minimize(to_min, 0, bounds=[[0, 2.*np.pi]])
     print('dphi/pi:', res.x/np.pi)
     return res.x
@@ -90,7 +90,7 @@ for gal,ax_col in zip(glist, ax.transpose()):
 
     ax_col.set_ylim(0, 300)
 
-    for z, c in zip(zoff_100_list, clist):
+    for z, c in zip(zoff_10_list, clist):
         ax_col.axhline(z, color=c)
 
 
@@ -113,11 +113,14 @@ for gal,ax_col in zip(glist, ax.transpose()):
     np.save('output/r_vs_dphi_'+gal+'.npy', out)
 
     # chord length
-    def tick_function(l):
-        return np.round(2*Rsolar*np.sin(0.5*(l/2)*np.pi), 1)
+    def tick_function(l, round=True):
+        a = 2*Rsolar*np.sin(0.5*(l/2)*np.pi)
+        if round:
+            a = np.round(a, 1)
+        return a
 
     t = print_dphi(dphi_mean, r_mean)
-    print(tick_function(t/np.pi))
+    print(tick_function(t/np.pi, round=False))
 
     ax2 = ax_col.twiny()
     #ax2.invert_xaxis()

### 2019-04-22 paper/ms.tex
Add sentence about error distribution

--- 2019-04-22 paper/ms.tex
@@
 
-\title{The Galactic Midplane is Not a Plane: Implications for {\em Gaia}}
+\title{The Galactic Midplane is Not a Plane: Implications for dynamical analysis with {\em Gaia} data and beyond}
 
@@
 major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
-2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the 2019arXiv190405721A
+2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the
 solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
@@
 the modes lie %mm at the true value of $J_z$.
-   at null. 
+   at zero. In other words, neither of the modes of the distribution of observed $J_z$ values are centered on the true $J_z$ (i.e. the $J_z$ computed with no offset in $z$).
 In the case of the thin-disk orbit

### 2019-04-22 paper/ms.tex
Cleaning up from mm's changes

@@
 revolution 
-%mm in the era of {\em Gaia} DR2. 
      as a result of {\em Gaia} Data Release 2 (DR2).  Recent major discoveries include the remnants of a
@@
 the Galactic center
-%mm is taken to be $\sim 8.1\text{--}8.3\,\kpc$
     can be measured
 \citep{2009ApJ...692.1075G,2018AA...615L..15G}, with a recent 
-%mm $\sim0.3\%$ measurement from the GRAVITY collaboration
     measurement using near-infrared interferometry of $8.178 \pm
@@
 the 
-%mm typical
 cyclical integral of the canonical momentum  over its conjugate position,
-%mm and are given by,
 \beq\label{eq:actions}
@@
 conjugate momenta.\footnote{We 
-%mm typically [there are no other J_R's in manuscript]
 write $J_r$ instead of $J_R$ for
@@
 the system's symmetry. 
-%mm For instance, 
     However, 
@@
 assumption 
-%mm is [counterfactual subjunctive, I believe]
      were
@@
 uncertainty in measurements of 
-%mm $R_{\odot}$\footnote{Here,
-%and throughout, we use $R_{\odot}$ to mean the distance from the Sun to the
-%Galactic center.}
      the Galactocentric radius of the Sun $R_{\odot}$
@@
 we 
-%mm attempt to quantify the 
     examine the expected 
@@
 as a vertical dashed line. Here we see that the 
-%mm spread 
     errors 
@@
 systematic offset in $z$ are non-Gaussian and bimodal. Furthermore, neither of
-the modes lie %mm at the true value of $J_z$.
+the modes lie
    at zero. In other words, neither of the modes of the distribution of observed $J_z$ values are centered on the true $J_z$ (i.e. the $J_z$ computed with no offset in $z$).
@@
 error distribution is not even centered on 
-%mm the true value. 
     null. 
@@
 {\em neither of the modes peaking at 
-%mm the true value}.
      null}. 
@@
 propagation is 
-%mm very involved 
      complex
@@
 We would like to thank Megan Bedell, Tobias Buck, and Todd Phillips
-%mm [moving to author list] David W. Hogg, and Kathryn
-%Johnston 
-%for many useful discussions. A.B. would like to thank Todd Phillips
+A.B. would like to thank Todd Phillips
 for helpful discussions. A.B. was supported in part by the Roy \& Diana

### 2019-04-22 paper/ms.tex

--- 2019-04-22 paper/ms.tex
@@
 
-\title{The Galactic Midplane is Not a Plane: Implications for {\em Gaia}}
+\title{The Galactic Midplane is Not a Plane: Implications for dynamical analysis with {\em Gaia} data and beyond}
 
 % \correspondingauthor{Angus Beane}
+%mm [will this address go away during the review process?.  Do you
+%want to give a "current address" as a second footnote?]
 \email{abeane@sas.upenn.edu}
@@
 principle, actions have the attractive quality of being invariant and thus
-good labels of a star's orbit. Of course, inaccurate measurements of the 
-phase space position and uncertainities in the assumed Galactic potential
+good labels of a star's orbit. 
+%mm Of course,
+    Such analyses are often used to interpret data from the {\em Gaia}
+    mission. However, 
+inaccurate measurements of the 
+phase space position and uncertainties in the assumed Galactic potential
 induce an error in the computed action. We show that, in addition to these
 complications, a systematic bias in the phase space position induces a
-phase-dependence in the actions, which we interpret as an error. The induced error distribution is
+phase-dependence in the actions, which we interpret as 
+%mm an 
+    a systematic
+error. The induced error distribution is
 non-Gaussian and bimodal, 
-with neither of the modes peaking on the true action value. 
-A midplane error of $\sim6\,\pc$ for a thin disk orbit or
-$\sim50\,\pc$ for a thick disk orbit is all that is necessary to
-induce a $10\,\%$ error in the vertical action $J_z$.
+with neither of the modes peaking on
+%mm  the true action value. [the error in a Gaussian peaks on zero] 
+    the null value.
+A variation in the vertical position midplane of $\sim30\,\pc$ for a thin disk orbit or
+$\sim225\,\pc$ for a thick disk orbit is 
+%mm all that is necessary 
+      sufficient
+to induce a $100\,\%$ systematic error in the vertical action $J_z$ computed under the assumption of an axisymmetric potential.
+%mm [fragment?] orbits, respectively. 
 Furthermore, we show that the local midplane varies by
@@
 \section{Introduction} \label{sec:intro}
-Our understanding of the Milky Way is currently undergoing a revolution in the
-era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
+Our understanding of the Milky Way is currently undergoing a
+revolution 
+     as a result of {\em Gaia} Data Release 2 (DR2).  Recent major discoveries include the remnants of a
 major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
@@
 stellar motions near Sgr~A\textsuperscript{*}, the distance from the Sun to
-the Galactic center is taken to be $\sim 8.1\text{--}8.3\,\kpc$
-\citep{2009ApJ...692.1075G,2018AA...615L..15G}, with a recent $\sim0.3\%$ measurement from the GRAVITY collaboration \citep{2019arXiv190405721A}. The Sun's vertical distance
+the Galactic center
+    can be measured
+\citep{2009ApJ...692.1075G,2018AA...615L..15G}, with a recent 
+    measurement using near-infrared interferometry of $8.178 \pm
+    0.035$~kpc 
+\citep{2019arXiv190405721A}. The Sun's vertical distance
 from the Galactic midplane, the other independent measurement that sets this
@@
 model the stellar distribution function, or find stars with similar dynamical
-properties. Actions are conserved quantities which describe the orbit of a
+properties. Actions are conserved quantities that describe the orbit of a
 star under the assumption of regular, bound orbits in a system where the
 equations of motion are separable in a particular coordinate system. They are
-the typical integral of the Hamiltonian in cylindrical coordinates and are
-given by,
+the 
+cyclical integral of the canonical momentum  over its conjugate position,
 \beq\label{eq:actions}
 J_i \equiv
-\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
+\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,.}
 \eeq
-where $i=R,z,\phi$ are the radial, vertical, and azimuthal coordinates
+Under the assumption of axisymmetry, $i=R,z,\phi$ are the radial, vertical, and azimuthal coordinates
 respectively in a cylindrical coordinate system and the $p_i$ are the
-conjugate momenta.\footnote{We typically write $J_r$ instead of $J_R$ for
+conjugate momenta.\footnote{We 
+write $J_r$ instead of $J_R$ for
 aesthetic reasons.} In a slowly-evolving axisymmetric potential, these actions
@@
 position and the potential, and requires relatively strong assumptions about
-the system's symmetry. For instance, the Galactic potential is not strictly
+the system's symmetry. 
+    However, 
+the Galactic potential is not strictly
 axisymmetric; with the vast improvement in the quality of phase-space
 measurements due to {\em Gaia}, this assumption appears increasingly
-inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this assumption is
+inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this
+assumption 
+     were
 close enough for many purposes, the parameters used in axisymmetric models may
@@
 coordinate frame can be usefully extended for action-space analysis. A similar
-argument applies to uncertainty in measurements of $R_{\odot}$\footnote{Here,
-and throughout, we use $R_{\odot}$ to mean the distance from the Sun to the
-Galactic center.} and variations in the local standard of rest (LSR).
+argument applies to 
+%mm [given that this is down to 0.3% error, seems a bit academic]
+   any remaining 
+uncertainty in measurements of 
+     the Galactocentric radius of the Sun $R_{\odot}$
+and to variations in the local standard of rest (LSR).
 
@@
 midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire}
-we attempt to quantify the azimuthal dependence of the local midplane, and
+we 
+    examine the expected 
+azimuthal dependence of the local midplane, and
 hence the expected size of the dynamical solar neighborhood, using
-cosmological-hydrodynamical zoom-in simulations of Milky Way-mass galaxies
+cosmological, hydrodynamical, zoom-in simulations of Milky Way-mass galaxies
 from the Feedback in Realistic Environments (FIRE)
@@
 of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted
-as a vertical dashed line. Here we see that the spread in $J_z$ induced by a
-systematic offset in $z$ is non-Gaussian and bimodal. Furthermore, neither of
-the modes lie at the true value of $J_z$. In the case of the thin-disk orbit
+as a vertical dashed line. Here we see that the 
+    errors 
+in $J_z$ induced by a
+systematic offset in $z$ are non-Gaussian and bimodal. Furthermore, neither of
+the modes lie
+   at zero. In other words, neither of the modes of the distribution of observed $J_z$ values are centered on the true $J_z$ (i.e. the $J_z$ computed with no offset in $z$).
+In the case of the thin-disk orbit
 ({\em lower panel}), we see that, in addition to the prior complications, the
-error distribution is not even centered on the true value. This comes about
+error distribution is not even centered on 
+    null. 
+This comes about
 when the midplane error is $\gtrsim z_{\text{max}}$, and can be understood by
@@
 coordinate system error is very non-Gaussian. The distribution is bimodal with
-{\em neither of the modes peaking at the true value}. As a result, error
-propagation is very involved when considering actions.
+{\em neither of the modes peaking at 
+     null}. 
+As a result, error
+propagation is 
+     complex
+when considering actions.
 
@@
 \acknowledgments
-We would like to thank Megan Bedell, Tobias Buck, David W. Hogg, and Kathryn
-Johnston for many useful discussions. A.B. would like to thank Todd Phillips
+We would like to thank Megan Bedell, Tobias Buck, and Todd Phillips
+A.B. would like to thank Todd Phillips
 for helpful discussions. A.B. was supported in part by the Roy \& Diana
 Vagelos Program in the Molecular Life Sciences and the Roy \& Diana Vagelos
-Challenge Award. The work of \Gus{everyone} is supported by the Simons
-Foundation. \Gus{remove last sentence if everyone CCA affiliated}
+Challenge Award. 
+%mm The work of \Gus{everyone} is supported by the Simons
+%Foundation. \Gus{remove last sentence if everyone CCA affiliated}
+M.-M.M.L. was partly supported by NSF grant AST18-15461.
 

### 2019-04-22 paper/ms.tex
Add velocities to orbit
Add a little bit about constancy of frequencies

@@
 vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
-constants. We can write down the actions in this approximation as
+constants. Similarly, the velocities of the orbit are given by,
+\beq\label{eq:orbits_vel_epi}
+\begin{split}
+v_z(t) &= A_z \kappa \cos{(\kappa t + \alpha)}\\
+v_R(t) &= A_R \nu \cos{(\nu t + \beta)}\text{.}
+\end{split}
+\eeq
+
+We can write down the actions in this approximation as
 \citep[Section~3.5.3b][]{2008gady.book.....B},
@@
 An induced error in $z$ of $\Delta z$ will cause $\Delta A_z \sim \Delta z$.
-Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$. If
-we assume that $\kappa$ and $\nu$ are constant (\Gus{justify?}), then we have
+Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$. Making the assumption that $\kappa$ and $\nu$ are constant, we have
 that
@@
 Therefore, errors in each action are linearly proportional to the errors in
-either the midplane or cylindrical radius.
+either the midplane or cylindrical radius. While in detail $\kappa$ and $\nu$ are not constant, we will find that this order of magnitude estimate matches the numerical estimates quite well.
 

### 2019-04-22 paper/ms.tex
Add velocity epicyclic calc

@@
 
-We will check these formulae against numerical tests next.
+A similarly rough argument can be made to arrive at the following formulae for
+the impact of velocity offsets on the actions,
+\beq\label{eq:actions_epi_error}
+\begin{split}
+\frac{\Delta L_z}{L_z} &= \frac{\Delta v_{\phi}}{v_c} \\
+\frac{\Delta J_r}{J_r} &= \frac{\Delta v_R}{v_{R,\text{max}}} \\
+\frac{\Delta J_z}{J_z} &= \frac{\Delta v_z}{v_{z,\text{max}}}\text{,}
+\end{split}
+\eeq
+where $v_{R,\text{max}}$ and $v_{z,\text{max}}$ are the maximum radial and
+vertical velocities of the orbit, respectively.
+
+We will check the formulae for the midplane offset against numerical tests
+next. The effect of velocity offsets on actions is deferred to future work, as
+we discuss in Section~\ref{ssec:lsr_var}.
 

### 2019-04-22 paper/ms.tex
Change from 100% error to 25% error in abstract

@@
     the null value.
-A variation in the vertical position midplane of $\sim30\,\pc$ for a thin disk orbit or
-$\sim225\,\pc$ for a thick disk orbit is 
+A variation in the vertical position midplane of $\sim15\,\pc$ for a thin disk orbit or
+$\sim120\,\pc$ for a thick disk orbit is 
 %mm all that is necessary 
       sufficient
-to induce a $100\,\%$ systematic error in the vertical action $J_z$ computed under the assumption of an axisymmetric potential.
+to induce a $25\,\%$ systematic error in the vertical action $J_z$ computed under the assumption of an axisymmetric potential.
 %mm [fragment?] orbits, respectively. 

### 2019-04-22 schmactions/many_orbits/plot_many_orbits.py

@@ -103,7 +103,7 @@ for fname, name, Az, AR, Rg, init_vel, c in zip(fname_list, name_list, Az_list, 
     dx[:,2] /= J2
 
     print_100(zoffset, dz[:,2], 1)
-    print_100(zoffset, dz[:,2], 0.1)
+    print_100(zoffset, dz[:,2], 0.25)
     print(zoffset, dz[:,2])
     
     for x,o,d in zip(ax, (zoffset, xoffset), (dz, dx)):

### 2019-04-22 paper/ms.tex
Add Nmemb column, degrade sig figs in Table 1

@@
 
-\begin{deluxetable*}{cCCCCCCC}
+\begin{deluxetable*}{cCCCCCCCC}
 \tablecaption{Relevant dynamical parameters for the $9$ open clusters within
@@
 detail in Section~\ref{sssec:reconstruction}.\label{tab:real_clusters}}
-\tablehead{\colhead{cluster} & \colhead{distance} & \colhead{$J_r$} &
+\tablehead{\colhead{cluster} & \colhead{$N_{\text{memb}}$} & \colhead{distance} & \colhead{$J_r$} &
 \colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} & \colhead{neighborhood} \\ 
-\colhead{ } &
+\colhead{ } & \colhead{ } &
 \colhead{$\mathrm{pc}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
@@
 \startdata
-alphaPer & 174.9 & 14.0 & -1739 & 0.0045 & 10.5 & 352\tablenotemark{b} & 57 \\
-Blanco1 & 237.2 & 1.6 & -1841 & 1.53 & 207.0 & 410\tablenotemark{c} & 1460 \\
-ComaBer & 85.9 & 1.6 & -1860 & 0.71 & 139.7 & 112\tablenotemark{d} & 522\\
-Hyades & 47.5 & 20.3 & -1757 & 0.26 & 81.5 & 400\tablenotemark{e} & 462 \\
-IC2391 & 151.6 & 5.5 & -1790 & 0.018 & 21.2 & & \\
-IC2602 & 152.2 & 11.5 & -1718 & 0.32 & 88.6 & & \\
-NGC2451A\tablenotemark{a} & 193.7 & 5.8 & -1805 & 0.21 & 73.4 & & \\
-Pleiades & 135.8 & 19.3 & -1696 & 0.38 & 96.3 & 800\tablenotemark{f} & 685 \\
-Praesepe & 186.2 & 21.2 & -1766 & 0.58 & 122.9 & 550\tablenotemark{d} & 781
+alphaPer & 740 & 174.9 & \sim20 & \sim -1700 & \sim0.005 & \sim11 & 352\tablenotemark{b} & 57 \\
+Blanco1 & 489 & 237.2 & \sim2 & \sim-1800 & \sim2 & \sim210 & 410\tablenotemark{c} & 1460 \\
+ComaBer & 153 & 85.9 & \sim2 & \sim-1900 & \sim0.7 & \sim140 & 112\tablenotemark{d} & 522\\
+Hyades & 515 & 47.5 & \sim20 & \sim-1800 & \sim0.3 & \sim80 & 400\tablenotemark{e} & 462 \\
+IC2391 & 325 & 151.6 & \sim6 & \sim-1800 & \sim0.02 & \sim20 & \nodata & \nodata \\
+IC2602 & 492 & 152.2 & \sim10 & \sim-1700 & \sim0.3 & \sim89 & \nodata & \nodata \\
+NGC2451A\tablenotemark{a} & 400 & 193.7 & \sim6 & \sim-1800 & \sim0.2 & \sim73 & \nodata & \nodata \\
+Pleiades & 1326 & 135.8 & \sim20 & \sim-1700 & \sim0.4 & \sim96 & 800\tablenotemark{f} & 685 \\
+Praesepe & 938 & 186.2 & \sim20 & \sim-1800 & \sim0.6 & \sim120 & 550\tablenotemark{d} & 781
 \enddata

### 2019-04-22 paper/ms.tex
Change from cartoon to cartoon approximation

@@
 \subsection{Cartoon effect of midplane error} \label{ssec:cartoon}
-We present a cartoon in Figure~\ref{fig:cartoon} to show how an inaccurate
+We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an inaccurate
 determination of the midplane leads to a phase-dependence of the actions
@@
 \plotone{fig/cartoon.pdf}
-\caption{Cartoon showing the effect an error in the determination of the
+\caption{Cartoon approximation showing the effect an error in the determination of the
 coordinate midplane can have on orbit integration and action estimation. The

### 2019-04-22 paper/ms.tex
Change colors in cartoon image

@@
 
-Now suppose this observer makes a measurement of the true orbit at the blue
-point or the pink point (i.e. at two different orbital phases). Then the blue
-and pink lines correspond to the orbits that the observer would compute for
+Now suppose this observer makes a measurement of the true orbit at the purple
+point or the green point (i.e. at two different orbital phases). Then the purple
+and green lines correspond to the orbits that the observer would compute for
 each point based on the potential model with the erroneous midplane. In action
@@
 observer measures the phase-space position of the star at two different
-orbital phases (blue and pink points). If the observer were then to integrate
+orbital phases (purple and green points). If the observer were then to integrate
 the star's orbit using a model potential with the erroneous midplane, they
-would obtain the blue and pink curves for the star's orbit, respectively. The
+would obtain the purple and green curves for the star's orbit, respectively. The
 actions estimated from these two  erroneous orbits would subsequently differ,

### 2019-04-22 cartoon/cartoon.py

@@ -19,11 +19,11 @@ y3 = 2.25*np.cos(x) - 1.25
 
 ax.plot(x, y, c=tb_c[-1])
 
-ax.plot(x, y2, c=tb_c[0])
-ax.scatter(3.*np.pi, -1, c=tb_c[0])
+ax.plot(x, y2, c=tb_c[4])
+ax.scatter(3.*np.pi, -1, c=tb_c[4])
 
-ax.plot(x, y3, c=tb_c[7])
-ax.scatter(4.*np.pi, 1, c=tb_c[7])
+ax.plot(x, y3, c=tb_c[6])
+ax.scatter(4.*np.pi, 1, c=tb_c[6])
 
 ax.plot(x, np.full(len(x), 0), c=tb_c[-1])
 ax.plot(x, np.full(len(x), -1.25), c=tb_c[-1], ls='dashed')

### 2019-04-22 paper/ms.tex
Add intuition for Jz hist

@@
 induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
-sort of error propagation must take this into account. We also plot one half
+sort of error propagation must take this into account. Intuition for the shape of each panel is given in the text. We also plot one half
 the 95\uth{} percentile minus the 5\uth{} percentile of each distribution as a
@@
 value of $J_z$, resulting in a biased estimate of $J_z$.
+
+The intuition behind the shape of Figure~\ref{fig:Jz_hist} is relatively
+simple. Consider first the thick-disk orbit ({\em top} panel), where the
+offset in $z$ is much less than $z_{\text{max}}$ of the orbit. The peaks in
+the distribution correspond to the turning points of the orbit (or points of
+maximum vertical amplitude), where $\v_z \sim 0$ and where the star spends
+most of it orbit. This is why the distribution is peaked on these values. For
+the thin-disk orbit ({\em bottom} panel), the offset in $z$ is comparable to
+$z_{\text{max}}$. Now, there will be some points in the orbit where $v_z = 0$
+and $z=0$ (in the observed, erroneous coordinate system). At these points, the
+computed $J_z$ will vanish. The asymmetry and systematic offset then comes
+about because of the constraint that $J_z \geg 0$.\footnote{This is similar to
+arguments in cosmology for why gravity produces non-Gaussianity in the density
+field, since the density cannot become negative but it can grow arbitrarily
+large.}
 

### 2019-04-22 paper/ms.tex
Fix line indenting throughout

@@
 \section{Introduction} \label{sec:intro}
-Our understanding of the Milky Way is currently undergoing a
-revolution 
-     as a result of {\em Gaia} Data Release 2 (DR2).  Recent major discoveries include the remnants of a
-major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
-2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the
-solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
-local substructure infall \citep{2018MNRAS.481.1501B, 2018arXiv180800451L} or
-bar buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation
-by a dark matter subtructure in the tidal stream GD1
-\citep{2018ApJ...863L..20P, 2018arXiv181103631B}. These discoveries all
+Our understanding of the Milky Way is currently undergoing a revolution as a
+result of {\em Gaia} Data Release 2 (DR2).  Recent major discoveries include
+the remnants of a major merger \citep{2018ApJ...860L..11K,
+2018Natur.563...85H, 2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space
+``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly
+indicating either local substructure infall \citep{2018MNRAS.481.1501B,
+2018arXiv180800451L} or bar buckling \citep{2018arXiv181109205K}, and a gap
+suggestive of perturbation by a dark matter subtructure in the tidal stream
+GD1 \citep{2018ApJ...863L..20P, 2018arXiv181103631B}. These discoveries all
 indicate that the Milky Way's stellar distribution, which demonstrably departs
@@
 stellar motions near Sgr~A\textsuperscript{*}, the distance from the Sun to
-the Galactic center
-    can be measured
-\citep{2009ApJ...692.1075G,2018AA...615L..15G}, with a recent 
-    measurement using near-infrared interferometry of $8.178 \pm
-    0.035$~kpc 
-\citep{2019arXiv190405721A}. The Sun's vertical distance
-from the Galactic midplane, the other independent measurement that sets this
-coordinate system, can be determined by identifying where the stellar density
-and velocities reach a maximum (effectively the median height of all disk
-stars), and is usually taken to be $\sim 25\,\pc$ \citep{2001ApJ...553..184C},
-with more recent measurements from {\em Gaia} DR2 placing it at $\sim 20\,\pc$
-\citep{2019MNRAS.482.1417B}. Another strategy is to use the cold gas or HII
-regions in the disk to define the Galactic midplane, leading to slightly
-different values ($\sim 5\,\pc$) for the Sun's relative height
-\citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review of these
-measurements is given by \citet{2016ARAA..54..529B}.
+the Galactic center can be measured \citep{2009ApJ...692.1075G,
+2018AA...615L..15G}, with a recent measurement using near-infrared
+interferometry of $8.178 \pm 0.035$~kpc \citep{2019arXiv190405721A}. The Sun's
+vertical distance from the Galactic midplane, the other independent
+measurement that sets this coordinate system, can be determined by identifying
+where the stellar density and velocities reach a maximum (effectively the
+median height of all disk stars), and is usually taken to be $\sim 25\,\pc$
+\citep{2001ApJ...553..184C}, with more recent measurements from {\em Gaia} DR2
+placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}. Another strategy is
+to use the cold gas or HII regions in the disk to define the Galactic
+midplane, leading to slightly different values ($\sim 5\,\pc$) for the Sun's
+relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review
+of these measurements is given by \citet{2016ARAA..54..529B}.
 
 Once a Galactocentric coordinate system has been established and a
-six-dimensional (6D) phase
-space measurement of a star has been made, it is often desirable to convert
-this measurement into action space to concisely summarize its projected orbit,
-model the stellar distribution function, or find stars with similar dynamical
-properties. Actions are conserved quantities that describe the orbit of a
-star under the assumption of regular, bound orbits in a system where the
-equations of motion are separable in a particular coordinate system. They are
-the 
-cyclical integral of the canonical momentum  over its conjugate position,
+six-dimensional (6D) phase space measurement of a star has been made, it is
+often desirable to convert this measurement into action space to concisely
+summarize its projected orbit, model the stellar distribution function, or
+find stars with similar dynamical properties. Actions are conserved quantities
+that describe the orbit of a star under the assumption of regular, bound
+orbits in a system where the equations of motion are separable in a particular
+coordinate system. They are the cyclical integral of the canonical momentum
+over its conjugate position,
 \beq\label{eq:actions}
@@
 \eeq
-Under the assumption of axisymmetry, $i=R,z,\phi$ are the radial, vertical, and azimuthal coordinates
-respectively in a cylindrical coordinate system and the $p_i$ are the
-conjugate momenta.\footnote{We 
-write $J_r$ instead of $J_R$ for
-aesthetic reasons.} In a slowly-evolving axisymmetric potential, these actions
-are invariant and $J_{\phi} \equiv L_z$
+Under the assumption of axisymmetry, $i=R,z,\phi$ are the radial, vertical,
+and azimuthal coordinates respectively in a cylindrical coordinate system and
+the $p_i$ are the conjugate momenta.\footnote{We write $J_r$ instead of $J_R$
+for aesthetic reasons.} In a slowly-evolving axisymmetric potential, these
+actions are invariant and $J_{\phi} \equiv L_z$
 \citep{2008gady.book.....B,2014RvMP...86....1S}.
@@
 individual stellar orbit is concisely described by three actions, as opposed
-to six phase space coordinates. Second, under the assumption of a
-phase-mixed system, the dynamical properties of a population of stars should
-be uniquely a function of their actions and independent of the conjugate
-angles, allowing investigation of the relationship between {\em orbital}
-properties of stars and other intrinsic, and at least partially invariant,
-properties such as age or metallicity. Actions also provide a convenient basis
-for constructing models of the stellar distribution function, or for
-associating stars with similar dynamical properties (e.g. to determine
-membership in moving groups).
+to six phase space coordinates. Second, under the assumption of a phase-mixed
+system, the dynamical properties of a population of stars should be uniquely a
+function of their actions and independent of the conjugate angles, allowing
+investigation of the relationship between {\em orbital} properties of stars
+and other intrinsic, and at least partially invariant, properties such as age
+or metallicity. Actions also provide a convenient basis for constructing
+models of the stellar distribution function, or for associating stars with
+similar dynamical properties (e.g. to determine membership in moving groups).
 
@@
 using the improved astrometry of {\em Gaia} DR2 and various age catalogues
-\citep{2018ApJ...867...31B,2018arXiv180803278T}. Actions have also been used
+\citep{2018ApJ...867...31B, 2018arXiv180803278T}. Actions have also been used
 to study different models of spiral structure in the Milky Way
@@
 position and the potential, and requires relatively strong assumptions about
-the system's symmetry. 
-    However, 
-the Galactic potential is not strictly
+the system's symmetry. However, the Galactic potential is not strictly
 axisymmetric; with the vast improvement in the quality of phase-space
 measurements due to {\em Gaia}, this assumption appears increasingly
-inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this
-assumption 
-     were
+inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this assumption were
 close enough for many purposes, the parameters used in axisymmetric models may
@@
 coordinate frame can be usefully extended for action-space analysis. A similar
-argument applies to 
-%mm [given that this is down to 0.3% error, seems a bit academic]
-   any remaining 
-uncertainty in measurements of 
-     the Galactocentric radius of the Sun $R_{\odot}$
-and to variations in the local standard of rest (LSR).
+argument applies to any remaining uncertainty in measurements of the
+Galactocentric radius of the Sun $R_{\odot}$ and to variations in the local
+standard of rest (LSR).
 
@@
 midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire}
-we 
-    examine the expected 
-azimuthal dependence of the local midplane, and
-hence the expected size of the dynamical solar neighborhood, using
-cosmological, hydrodynamical, zoom-in simulations of Milky Way-mass galaxies
-from the Feedback in Realistic Environments (FIRE)
+we examine the expected azimuthal dependence of the local midplane, and hence
+the expected size of the dynamical solar neighborhood, using cosmological,
+hydrodynamical, zoom-in simulations of Milky Way-mass galaxies from the
+Feedback in Realistic Environments (FIRE)
 collaboration\footnote{\url{https://fire.northwestern.edu}}
-\citep{2016ApJ...827L..23W,2017arXiv170206148H}.
+\citep{2016ApJ...827L..23W, 2017arXiv170206148H}.
 
@@
 \subsection{Cartoon effect of midplane error} \label{ssec:cartoon}
-We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an inaccurate
-determination of the midplane leads to a phase-dependence of the actions
-calculated from an orbit specified by a phase-space point and an assumed
-potential model. The $y$-axis corresponds to the vertical height of the orbit
-and the $x$-axis orbital phase. The solid gray line indicates the ``true''
-orbit of the star; that is, its orbit as it oscillates around the true
-midplane. The dashed gray line is an erroneous midplane assigned to the
+We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
+inaccurate determination of the midplane leads to a phase-dependence of the
+actions calculated from an orbit specified by a phase-space point and an
+assumed potential model. The $y$-axis corresponds to the vertical height of
+the orbit and the $x$-axis orbital phase. The solid gray line indicates the
+``true'' orbit of the star; that is, its orbit as it oscillates around the
+true midplane. The dashed gray line is an erroneous midplane assigned to the
 potential model which an observer uses to integrate the orbit of the star and
@@
 Now suppose this observer makes a measurement of the true orbit at the purple
-point or the green point (i.e. at two different orbital phases). Then the purple
-and green lines correspond to the orbits that the observer would compute for
-each point based on the potential model with the erroneous midplane. In action
-space, this would correspond to a different value of $J_z$ for the blue and
-pink points. In this way, assuming the wrong coordinate system induces a
-phase-dependence in the actions estimated for the star, which in the correct
-potential (in this example, the one with the correct midplane) should be
-phase-independent.
+point or the green point (i.e. at two different orbital phases). Then the
+purple and green lines correspond to the orbits that the observer would
+compute for each point based on the potential model with the erroneous
+midplane. In action space, this would correspond to a different value of $J_z$
+for the blue and pink points. In this way, assuming the wrong coordinate
+system induces a phase-dependence in the actions estimated for the star, which
+in the correct potential (in this example, the one with the correct midplane)
+should be phase-independent.
 
@@
 observer measures the phase-space position of the star at two different
-orbital phases (purple and green points). If the observer were then to integrate
-the star's orbit using a model potential with the erroneous midplane, they
-would obtain the purple and green curves for the star's orbit, respectively. The
-actions estimated from these two  erroneous orbits would subsequently differ,
-both from each other and from the actions estimated for the true orbit (in the
-potential with the correct midplane). Hence an incorrect midplane in the
-potential model assumed will induce phase-dependence in the actions estimated
-for a given star in that potential.}
+orbital phases (purple and green points). If the observer were then to
+integrate the star's orbit using a model potential with the erroneous
+midplane, they would obtain the purple and green curves for the star's orbit,
+respectively. The actions estimated from these two  erroneous orbits would
+subsequently differ, both from each other and from the actions estimated for
+the true orbit (in the potential with the correct midplane). Hence an
+incorrect midplane in the potential model assumed will induce phase-dependence
+in the actions estimated for a given star in that potential.}
 \label{fig:cartoon}
@@
 An induced error in $z$ of $\Delta z$ will cause $\Delta A_z \sim \Delta z$.
-Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$. Making the assumption that $\kappa$ and $\nu$ are constant, we have
-that
+Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$.
+Making the assumption that $\kappa$ and $\nu$ are constant, we have that
 \beq\label{eq:actions_epi_error}
@@
 Therefore, errors in each action are linearly proportional to the errors in
-either the midplane or cylindrical radius. While in detail $\kappa$ and $\nu$ are not constant, we will find that this order of magnitude estimate matches the numerical estimates quite well.
-
-There is a slight complication in the formula for computing $\Delta J_r/J_r$ and $\Delta L_z/L_z$.
-In a moment we will consider the effect of an inaccuracy in the distance from
-the Sun to the Galactic center on actions. However, this inaccuracy induces an
-error in $x$, or $\Delta x$. This is not the same as an inaccuracy in $\Delta
-R$, but they are related by the following argument. An error $\Delta x$ will
-induce an erroneous radius $R_{\text{err}}$ related by the formula,
+either the midplane or cylindrical radius. While in detail $\kappa$ and $\nu$
+are not constant, we will find that this order of magnitude estimate matches
+the numerical estimates quite well.
+
+There is a slight complication in the formula for computing $\Delta J_r/J_r$
+and $\Delta L_z/L_z$. In a moment we will consider the effect of an inaccuracy
+in the distance from the Sun to the Galactic center on actions. However, this
+inaccuracy induces an error in $x$, or $\Delta x$. This is not the same as an
+inaccuracy in $\Delta R$, but they are related by the following argument. An
+error $\Delta x$ will induce an erroneous radius $R_{\text{err}}$ related by
+the formula,
 \beq
@@
 We also perform the same procedure but assume a $100\,\pc$ offset in the $x$
-    component (i.e. subtracting the vector $(100, 0, 0)\,\pc$) in the {\em
-    lower} panels. This is equivalent to a systematic offset in the
-    determination of the distance from the Sun to the Galactic center.
+component (i.e. subtracting the vector $(100, 0, 0)\,\pc$) in the {\em lower}
+panels. This is equivalent to a systematic offset in the determination of the
+distance from the Sun to the Galactic center.
 
@@
 induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
-sort of error propagation must take this into account. Intuition for the shape of each panel is given in the text. We also plot one half
-the 95\uth{} percentile minus the 5\uth{} percentile of each distribution as a
-horizontal arrow anchored on the true $J_z$ value. We call this $\Delta J_z$
-and will use it (along with the similarly defined $\Delta J_r$ and $\Delta
-L_z$) to empirically describe the error distribution. We see that $\Delta J_z$
-roughly corresponds to the distance from the true $J_z$ value to one of the
-modes of the error distribution.}
+sort of error propagation must take this into account. Intuition for the shape
+of each panel is given in the text. We also plot one half the 95\uth{}
+percentile minus the 5\uth{} percentile of each distribution as a horizontal
+arrow anchored on the true $J_z$ value. We call this $\Delta J_z$ and will use
+it (along with the similarly defined $\Delta J_r$ and $\Delta L_z$) to
+empirically describe the error distribution. We see that $\Delta J_z$ roughly
+corresponds to the distance from the true $J_z$ value to one of the modes of
+the error distribution.}
 \label{fig:Jz_hist}
@@
 of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted
-as a vertical dashed line. Here we see that the 
-    errors 
-in $J_z$ induced by a
+as a vertical dashed line. Here we see that the errors in $J_z$ induced by a
 systematic offset in $z$ are non-Gaussian and bimodal. Furthermore, neither of
-the modes lie
-   at zero. In other words, neither of the modes of the distribution of observed $J_z$ values are centered on the true $J_z$ (i.e. the $J_z$ computed with no offset in $z$).
-In the case of the thin-disk orbit
+the modes lie at zero. In other words, neither of the modes of the
+distribution of observed $J_z$ values are centered on the true $J_z$ (i.e. the
+$J_z$ computed with no offset in $z$). In the case of the thin-disk orbit
 ({\em lower panel}), we see that, in addition to the prior complications, the
-error distribution is not even centered on 
-    null. 
-This comes about
-when the midplane error is $\gtrsim z_{\text{max}}$, and can be understood by
-realizing that in such a case the midplane offset always acts to increase the
-value of $J_z$, resulting in a biased estimate of $J_z$.
+error distribution is not even centered on null. This comes about when the
+midplane error is $\gtrsim z_{\text{max}}$, and can be understood by realizing
+that in such a case the midplane offset always acts to increase the value of
+$J_z$, resulting in a biased estimate of $J_z$.
 
@@
 Figure~\ref{fig:Jz_hist} is difficult. Of course, Gaussian summary statistics
-are not adequate. We therefore elect to measure this error by computing one half the
-95\uth{} percentile minus the 5\uth{} percentile of the distribution of
-observed action values. We refer to this quantity as $\Delta J_i$ for each
-action value. We plot this quantity in Figure~\ref{fig:Jz_hist} as a horizontal arrow anchored on the true action value. Because of the bimodality of the error distribution, this
-quantity roughly measures the distance from the true action value to the peak of one of the modes.
-Furthermore, this bimodality also implies that $\Delta J_i$ is not very
-sensitive to the exact percentiles used. This summary statistic does not
-reflect the bias induced when the midplane error is $\gtrsim z_{\text{max}}$.
+are not adequate. We therefore elect to measure this error by computing one
+half the 95\uth{} percentile minus the 5\uth{} percentile of the distribution
+of observed action values. We refer to this quantity as $\Delta J_i$ for each
+action value. We plot this quantity in Figure~\ref{fig:Jz_hist} as a
+horizontal arrow anchored on the true action value. Because of the bimodality
+of the error distribution, this quantity roughly measures the distance from
+the true action value to the peak of one of the modes. Furthermore, this
+bimodality also implies that $\Delta J_i$ is not very sensitive to the exact
+percentiles used. This summary statistic does not reflect the bias induced
+when the midplane error is $\gtrsim z_{\text{max}}$.
 
@@
 but for systematic offsets between $0$ and $500\,\pc$ in the $z$ and $x$
-components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report one half the 95\uth{}
-minus 5\uth{} percentile divided by the true action value ($\Delta J_i/J_i$) for the three
-different fiducial orbits in Table \ref{tab:orbits}. The thick-disk
-(\thickcolor) orbit is the one from Figure~\ref{fig:one_orbit_wrong_ref}, but
-we also now consider the effect on the action determined for the thin-disk
-(\thincolor) and halo (\halocolor) orbits.
+components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report one half the
+95\uth{} minus 5\uth{} percentile divided by the true action value ($\Delta
+J_i/J_i$) for the three different fiducial orbits in Table \ref{tab:orbits}.
+The thick-disk (\thickcolor) orbit is the one from
+Figure~\ref{fig:one_orbit_wrong_ref}, but we also now consider the effect on
+the action determined for the thin-disk (\thincolor) and halo (\halocolor)
+orbits.
 
@@
 show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
-respectively. We compute the fractional spread $\Delta J_i/J_i$ as one half the 95\uth{}
-minus 5\uth{} percentile of the action computed using the offset coordinate
-system over the course of the first $\Gyr$, divided by the true action
-(computed in the correct coordinate system).
+respectively. We compute the fractional spread $\Delta J_i/J_i$ as one half
+the 95\uth{} minus 5\uth{} percentile of the action computed using the offset
+coordinate system over the course of the first $\Gyr$, divided by the true
+action (computed in the correct coordinate system).
 
@@
 \plotone{fig/schmactions_many_orbits.pdf}
-\caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error in the
-measured action ($\Delta J_i$) from coordinate system errors for the thin,
-thick, and halo orbits (Table~\ref{tab:orbits}). See discussion in the text for the justification in using this to measure the magnitude of the induced error. The {\em left}, {\em center},
-and {\em right} panels show the result for $J_r$, $L_z$, and $J_z$,
-respectively. The {\em upper} panels consider an offset in $z$ and the {\em
-lower} panels consider an offset in $x$ (equivalently, an offset in the Solar
-radius).
-In some panels we also plot as dashed lines the epicyclic prediction of the induced action error (Equations~\ref{eq:actions_epi_error}~and~\ref{eq:Jr_epi_dx}).
-In the epicyclic approximation, a $z$ offset only induces an error in $J_z$ --- for all three orbits the epicyclic approximation is a good description of the $J_z$ error.
-An $x$ offset induces an error in $J_r$ and $L_z$.
-The error in $J_r$ is somewhat well-described for the thin-disk and thick-disk orbits, and a poor description for the halo orbit.
-For $L_z$, the epicyclic approximation is not a good description for any orbit.}
+\caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error
+in the measured action ($\Delta J_i$) from coordinate system errors for the
+thin, thick, and halo orbits (Table~\ref{tab:orbits}). See discussion in the
+text for the justification in using this to measure the magnitude of the
+induced error. The {\em left}, {\em center}, and {\em right} panels show the
+result for $J_r$, $L_z$, and $J_z$, respectively. The {\em upper} panels
+consider an offset in $z$ and the {\em lower} panels consider an offset in $x$
+(equivalently, an offset in the Solar radius). In some panels we also plot as
+dashed lines the epicyclic prediction of the induced action error
+(Equations~\ref{eq:actions_epi_error}~and~\ref{eq:Jr_epi_dx}). In the
+epicyclic approximation, a $z$ offset only induces an error in $J_z$ --- for
+all three orbits the epicyclic approximation is a good description of the
+$J_z$ error. An $x$ offset induces an error in $J_r$ and $L_z$. The error in
+$J_r$ is somewhat well-described for the thin-disk and thick-disk orbits, and
+a poor description for the halo orbit. For $L_z$, the epicyclic approximation
+is not a good description for any orbit.}
 \label{fig:many_orbit_wrong_ref}
@@
 different offsets in $z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
-(red, cyan, green respectively). As before, the error ($\Delta J_z$) is one half the
-95\uth{} minus 5\uth{} percentile of the distribution of observed $J_z$ values
-over the course of the orbit. There are large errors for thin-disk like orbits
-($J_z \sim 0.5\,\actunit$), even for a small midplane offset of $10\,\pc$. }
+(red, cyan, green respectively). As before, the error ($\Delta J_z$) is one
+half the 95\uth{} minus 5\uth{} percentile of the distribution of observed
+$J_z$ values over the course of the orbit. There are large errors for
+thin-disk like orbits ($J_z \sim 0.5\,\actunit$), even for a small midplane
+offset of $10\,\pc$. }
 \label{fig:dJz_fun_Jz}
@@
 the actions under this symmetry assumption, when starting from the present-day
-positions and velocities of stars as measured by, e.g., \emph{Gaia}.
-These systematic errors are most important for stars on thin disk-like orbits,
-where they can be large enough to yield actions representative of orbits in
-the thick disk. This effect is entirely due to the extension of a local to a
+positions and velocities of stars as measured by, e.g., \emph{Gaia}. These
+systematic errors are most important for stars on thin disk-like orbits, where
+they can be large enough to yield actions representative of orbits in the
+thick disk. This effect is entirely due to the extension of a local to a
 global coordinate system, and is separate from real diffusion in stellar
@@
 of the density fluctuations in the Galactic disk will determine how far this
-coordinate system can be extended before introducing relevant systematic errors
-in the computation of actions for thin-disk stars. We consider a $10\%$ error to be a
-criterion for the region around the Sun where the analysis of the action-space
-distribution of stars on quasi-circular orbits, and therefore the epicyclic
-approximation, can be expected to be valid. This leads to a somewhat
-dynamically motivated definition of the ``solar neighborhood.'' Since the
-motivation may be different for defining the solar neighborhood in, e.g.,
+coordinate system can be extended before introducing relevant systematic
+errors in the computation of actions for thin-disk stars. We consider a $10\%$
+error to be a criterion for the region around the Sun where the analysis of
+the action-space distribution of stars on quasi-circular orbits, and therefore
+the epicyclic approximation, can be expected to be valid. This leads to a
+somewhat dynamically motivated definition of the ``solar neighborhood.'' Since
+the motivation may be different for defining the solar neighborhood in, e.g.,
 studies of the chemical distribution of the Galaxy, we propose the phrase
-``dynamical solar neighborhood.'' We also point out that our use of a $10\%$ error is arbitrary.
+``dynamical solar neighborhood.'' We also point out that our use of a $10\%$
+error is arbitrary.
 
@@
 Table~\ref{tab:real_clusters}. Except for alphaPer, we find that the
-neigbhorhood around each cluster is typically $\gtrsim x\,\pc$. This
-analysis assumes that the MW resembles \mi{}, which is not necessarily true.
-However, this step of our analysis is meant to be more illustrative than
-quantitative: this calculation should actually be performed using real
-measurements of variations in the MW's stellar midplane. Such measurements
-should be possible in the near future with {\em Gaia}, with quality distance
-estimates the crucial limiting factor, along with careful selction function modelling.
+neigbhorhood around each cluster is typically $\gtrsim x\,\pc$. This analysis
+assumes that the MW resembles \mi{}, which is not necessarily true. However,
+this step of our analysis is meant to be more illustrative than quantitative:
+this calculation should actually be performed using real measurements of
+variations in the MW's stellar midplane. Such measurements should be possible
+in the near future with {\em Gaia}, with quality distance estimates the
+crucial limiting factor, along with careful selction function modelling.
 

### 2019-04-22 paper/ms.tex
Add Sun position, vcirc, fix errors

@@
 
+We assume the Sun is located at $(8.2, 0, 0)\,\kpc$. None of our orbit
+calculations depend on the value of the LSR in this toy potential (though this
+is important when using real data, since the conversion from heliocentric to
+Galactocentric coordinates depends on the LSR). In this potential, we have
+that the circular velocity $v_{\text{circ}}$ is $231\,\kms$ on the Solar
+circle.
+
 Other methods for computing actions are used in the literature. For example,
@@
 the distribution correspond to the turning points of the orbit (or points of
-maximum vertical amplitude), where $\v_z \sim 0$ and where the star spends
+maximum vertical amplitude), where $v_z \sim 0$ and where the star spends
 most of it orbit. This is why the distribution is peaked on these values. For
@@
 computed $J_z$ will vanish. The asymmetry and systematic offset then comes
-about because of the constraint that $J_z \geg 0$.\footnote{This is similar to
+about because of the constraint that $J_z \geq 0$.\footnote{This is similar to
 arguments in cosmology for why gravity produces non-Gaussianity in the density

### 2019-04-22 paper/ms.tex
Create new section that compiles all Galactocentric stuff

@@
 Sun's position (and therefore the origin of the coordinate system) can be
-defined and measured both precisely and accurately. In this system, the center
-of the Galaxy is usually taken to be the location of the central supermassive
-black hole, Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From
-stellar motions near Sgr~A\textsuperscript{*}, the distance from the Sun to
-the Galactic center can be measured \citep{2009ApJ...692.1075G,
-2018AA...615L..15G}, with a recent measurement using near-infrared
-interferometry of $8.178 \pm 0.035$~kpc \citep{2019arXiv190405721A}. The Sun's
-vertical distance from the Galactic midplane, the other independent
-measurement that sets this coordinate system, can be determined by identifying
-where the stellar density and velocities reach a maximum (effectively the
-median height of all disk stars), and is usually taken to be $\sim 25\,\pc$
-\citep{2001ApJ...553..184C}, with more recent measurements from {\em Gaia} DR2
-placing it at $\sim 20\,\pc$ \citep{2019MNRAS.482.1417B}. Another strategy is
-to use the cold gas or HII regions in the disk to define the Galactic
-midplane, leading to slightly different values ($\sim 5\,\pc$) for the Sun's
-relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review
-of these measurements is given by \citet{2016ARAA..54..529B}.
+defined and measured both precisely and accurately.
 
@@
 
-\section{Azimuthal midplane variations at the Solar Circle: estimates from
-simulation} \label{sec:local_fire}
+\section{Coordinate System Offsets I. Measurement Errors} \label{sec:mes_err}
+
+\subsection{Galactic Center}
+First, one must define the center of the Galaxy. This is usually taken to be
+the location of the central supermassive black hole, Sgr~A\textsuperscript{*}
+\citep[e.g.][]{2004ApJ...616..872R}. From stellar motions near
+Sgr~A\textsuperscript{*}, the distance from the Sun to the Galactic center can
+be measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a recent
+measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
+\citep{2019arXiv190405721A}.
+
+We take the ``dynamical Galactic center'' to be the point in three-dimensional
+space about which the stars in the solar neighborhood are orbiting. An
+important point to make about the Galactic center is that while many authors
+assume Sgr~A\textsuperscript{*} is colocated with the dynamical Galactic
+center, this assumption has not yet been justified. One potential test is to
+perform some sort of dynamical modelling where the distance from the Sun to
+the Galactic center is a free parameter and can be constrained with real data.
+For example, \citet{2015ApJ...803...80K} were able to make this measurement by
+modelling the dynamics of the stream Palomar 5. Other dynamical measurements
+of the Solar distance have been made, though none with a precision capable of
+competing with the distance to Sgr~A\textsuperscript{*}
+\citep{1981gask.book.....M, 2011PASJ...63..867S, 2012MNRAS.427..274S,
+2013AstL...39...95B, 2013IAUS..289..444Z}.
+
+\subsection{Galactic Orientation}
+Second, one must define the orientation of the Galaxy about the center. This
+was defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
+defining the location of the Galactic center and the north Galactic pole.
+However, there is growing evidence that the true midplane is tilted when
+considering stars \citep{2014ApJ...797...53G, 2016ARAA..54..529B}, though not
+with H~\textsc{ii} \citep{2019ApJ...871..145A}.
+
+\subsection{Solar Height}
+Thirdly, one must define The Sun's vertical distance from the Galactic
+midplane, which can be determined by identifying where the stellar density
+and velocities reach a maximum (effectively the median height of all disk
+stars) The Solar height is usually taken to be $\sim 25\,\pc$
+\citep{2001ApJ...553..184C}, with a more recent measurement from {\em Gaia} DR2
+placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another strategy is
+to use the cold gas or H~\textsc{ii} regions in the disk to define the Galactic
+midplane, leading to slightly different values ($\sim 5\,\pc$) for the Sun's
+relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review
+of these measurements is given by \citet{2016ARAA..54..529B}.
+
+\subsection{Local Standard of Rest}
+Finally, one must define the local standard of rest (LSR). The radial
+($U_{\odot}$) and vertical ($W_{\odot}$)components are easily computed by
+taking the mean motions of different stellar groups. The azimuthal component
+($V_{\odot}$) is more difficult to measure, but can be modelled using the
+asymmetric drift relation \citep{2008gady.book.....B}. The values of the
+components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}. The
+value of the circular velocity is taken to be $\sim 220\kms$
+\citep[e.g.][]{2012ApJ...759..131B}.
+
+In order to convert a star's position and velocity from heliocentric to
+Galactocentric coordinates, a number of parameters must be measured. Given the
+great interest in each parameter, considerable effort has been placed on these
+measurements. However, uncertainties remain, and detailed dynamical modelling
+across the disk --- particularly for dynamically cold stars --- may have to
+take them into account. For instance, the discrepancy of $\sim 5\,\pc$ between
+the midplane defined by stars and by gas will induce a $\sim10\%$ error in
+$J_z$ for an orbit with $z_{\text{max}}\sim100\,\pc$ (see
+Section~\ref{sec:ref_frame}).
+
+% \section{Azimuthal midplane variations at the Solar Circle: estimates from
+% simulation} 
+\section{Coordinate System Offsets II. Azimuthal Midplane Variations in
+Simulations}
+\label{sec:local_fire}
 \subsection{Description of FIRE Simulation} \label{ssec:cosmozoom}
@@
 Our main point is that the local midplane varies between different points in
-the Galaxy. This is a different statement than that our measurements of the
-local midplane are erroneous. Indeed, the local midplane has recently been
-measured to a claimed high level of precision, setting $z_{\odot} = 20.8 \pm
-0.3\,\pc$ \citep{2019MNRAS.482.1417B}. However, we also do not mean to say
-that determinations of the local midplane are perfect, since midplane
-measurements from H~\textsc{ii} regions place $z_{\odot} \sim 5-10\,\pc$. An
-additional complication is the fact that non-zero Solar heights induce a tilt
-in the Galactic midplane \citep{2014ApJ...797...53G,2016ARAA..54..529B}. We do
+the Galaxy. We do
 not propose here what a modeler should use for the ``true'' global midplane,

### 2019-04-22 paper/ms.tex
Add bit more about galactocent. in intro

@@
 Sun's position (and therefore the origin of the coordinate system) can be
-defined and measured both precisely and accurately.
+defined and measured both precisely and accurately. This involves measuring
+the Galactic center, the orientation of the Galactic plane, the distance from
+the Sun to the Galactic center, the height of the Sun above the Galactic
+midplane, and the local standard of rest (LSR). We discuss the observational
+efforts to measure all these parameters in Section~\ref{sec:mes_err}.
 

### 2019-04-22 paper/ms.tex
Add little bit of pre-stuff in FIRE section

@@
 \label{sec:local_fire}
+
+We now focus our attention on understanding the intrinsic variations of the
+midplane across the disk in two sets of complementary simulations. We will
+briefly discuss a rough estimate of the midplane variation derived from the
+observed velocity distribution of the Milky Way, but defer a detailed
+treatment to future work.
+
 \subsection{Description of FIRE Simulation} \label{ssec:cosmozoom}

### 2019-04-22 paper/ms.tex
Add Friske+ velocity variations, OOM midplane estimate

@@
 
-\subsection{LSR Variations} \label{ssec:lsr_var}
+\subsection{Velocity Variations} \label{ssec:lsr_var}
 We also expect that the local standard of rest (LSR) should vary as a function
@@
 this calculation to future work.
+
+An interesting, qualitative calculation can be made based on the velocity
+variations seen in e.g. \citet{2019arXiv190209569F}. The velocity variations
+in this work (admittedly found with respect to $L_z$ or radius) are
+$\sim5\text{--}10\,\kms$. The vertical frequency of the thin-disk and
+thick-disk orbits are $\sim 0.09\,1/\Myr$ and $\sim0.06\,1/\Myr$,
+respectively. By dimensional analysis, we therefore expect the midplane
+offsets to be of order $\sim 57\text{--}170\,\pc$. We stress that this is a
+rough calculation, but it shows that already we see evidence in the data from
+velocities for midplane offsets on the order of what we see in both sets of
+simulations.
 

### 2019-04-22 paper/ms.tex
Add bit about chemo-dynamical tagging
Add values of A, B, C for Sag

@@
 form $\text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}$, with the values
-for $\text{A}$, $\text{B}$, and $\text{C}$ being \Gus{something},
-\Gus{something}, and \Gus{something else}, respectively. In the first two
-panels, we see that the midplane is relatively flat in the inner galaxy. But
-additional encounters drive strong midplane variation. In the third panel, we
-see a strong $m=2$ mode develop, consistent with the $R=8\,\kpc$ panel of
-Figure~17 in \citet{2018MNRAS.481..286L}.
+for $\text{A}$ being $9.5, 2.5, -212, -394\,\pc$, for $\text{B}$ being
+$0.0013, 0.0007, -0.63, -1.00\,\text{rad}$ and $\text{C}$ being $2.6, -7.8,
+-65, -53\,\pc$. Each of these values are for each successive timestep. In the
+first two panels, we see that the midplane is relatively flat in the inner
+galaxy. But additional encounters drive strong midplane variation. In the
+third panel, we see a strong $m=2$ mode develop, consistent with the
+$R=8\,\kpc$ panel of Figure~17 in \citet{2018MNRAS.481..286L}.
 
@@
 
+\item This work emphasizes the importance of combining chemistry and dynamics.
+Chemical tagging \citep{2002ARA&A..40..487F} and dynamical tagging must
+complement and confirm each other. This was also recently pointed out by
+\Gus{someone}.
+
 \end{itemize}

### 2019-04-22 paper/ms.tex
Change to J_R, J_phi throughout

@@
 and azimuthal coordinates respectively in a cylindrical coordinate system and
-the $p_i$ are the conjugate momenta.\footnote{We write $J_r$ instead of $J_R$
-for aesthetic reasons.} In a slowly-evolving axisymmetric potential, these
+the $p_i$ are the conjugate momenta. In a slowly-evolving axisymmetric potential, these
 actions are invariant and $J_{\phi} \equiv L_z$
@@
 detail in Section~\ref{sssec:reconstruction}.\label{tab:real_clusters}}
-\tablehead{\colhead{cluster} & \colhead{$N_{\text{memb}}$} & \colhead{distance} & \colhead{$J_r$} &
-\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} & \colhead{neighborhood} \\ 
+\tablehead{\colhead{cluster} & \colhead{$N_{\text{memb}}$} & \colhead{distance} & \colhead{$J_R$} &
+\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} & \colhead{neighborhood} \\ 
 \colhead{ } & \colhead{ } &
@@
 \begin{split}
-L_z &= R_g v_c \\
-J_r &= \kappa A_R^2 \\
+J_{\phi} &= R_g v_c \\
+J_R &= \kappa A_R^2 \\
 J_z &= \nu A_z^2\text{,}
@@
 \begin{split}
-\Delta L_z = \Delta R v_c &\implies \frac{\Delta L_z}{L_z} = \frac{\Delta R}{R} \\
-\Delta J_r = 2 \kappa \Delta A_R &\implies \frac{\Delta J_r}{J_r} = \frac{2\Delta R}{A_R} \\
+\Delta J_{\phi} = \Delta R v_c &\implies \frac{\Delta J_{\phi}}{J_{\phi}} = \frac{\Delta R}{R} \\
+\Delta J_R = 2 \kappa \Delta A_R &\implies \frac{\Delta J_R}{J_R} = \frac{2\Delta R}{A_R} \\
 \Delta J_z = 2 \nu \Delta A_z &\implies \frac{\Delta J_z}{J_z} = \frac{2\Delta z}{A_z}\text{.}
@@
 
-There is a slight complication in the formula for computing $\Delta J_r/J_r$
-and $\Delta L_z/L_z$. In a moment we will consider the effect of an inaccuracy
+There is a slight complication in the formula for computing $\Delta J_R/J_R$
+and $\Delta J_{\phi}/J_{\phi}$. In a moment we will consider the effect of an inaccuracy
 in the distance from the Sun to the Galactic center on actions. However, this
@@
 \eeq
-Therefore, the error in $J_r$ and $L_z$ are related to $\Delta x$ by,
+Therefore, the error in $J_R$ and $J_{\phi}$ are related to $\Delta x$ by,
 \beq\label{eq:Jr_epi_dx}
 \begin{split}
-\frac{\Delta J_r}{J_r} &= \frac{4 \Delta x}{\pi A_R} \\
-\frac{\Delta L_z}{L_z} &= \frac{2 \Delta x}{\pi R}\text{.}
+\frac{\Delta J_R}{J_R} &= \frac{4 \Delta x}{\pi A_R} \\
+\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{2 \Delta x}{\pi R}\text{.}
 \end{split}
@@
 \begin{split}
-\frac{\Delta L_z}{L_z} &= \frac{\Delta v_{\phi}}{v_c} \\
-\frac{\Delta J_r}{J_r} &= \frac{\Delta v_R}{v_{R,\text{max}}} \\
+\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta v_{\phi}}{v_c} \\
+\frac{\Delta J_R}{J_R} &= \frac{\Delta v_R}{v_{R,\text{max}}} \\
 \frac{\Delta J_z}{J_z} &= \frac{\Delta v_z}{v_{z,\text{max}}}\text{,}
@@
 \tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
-velocity} & \colhead{$J_r$} &
-\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
+velocity} & \colhead{$J_R$} &
+\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
 \colhead{$A_R$}\\ \colhead{ } &
@@
 95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
-$6.2\,\actunit$ for $J_r$ and $J_z$, respectively. These are fractional errors
-of $5.7\,\%$ and $85.7\,\%$, respectively. The error induced in $L_z$ is
-negligible, as expected since $L_z$ is independent of $z$. It is worth
+$6.2\,\actunit$ for $J_R$ and $J_z$, respectively. These are fractional errors
+of $5.7\,\%$ and $85.7\,\%$, respectively. The error induced in $J_{\phi}$ is
+negligible, as expected since $J_{\phi}$ is independent of $z$. It is worth
 pointing out that a $100\,\pc$ error in an orbit with
@@
 95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
-$47$, and $0.71\,\actunit$ for $J_r$, $L_z$ and $J_z$, respectively. These are
+$47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively. These are
 fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively.
@@
 error in the Galactocentric coordinate system. We consider here the thick-disk
-orbit, which has actions of $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$
+orbit, which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520, 7.0)\,\actunit$
 and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We integrate the
@@
 procedure at each timestep, and report the values of the actions. The
-computation of $L_z$ is pristine to errors in $z$, with only numerical
-artifacts remaining. Only small errors are induced in $J_r$, with the middle
-$90\%$ of values over the $\Gyr$ being within $\sim8\%$ of the true $J_r$. As
+computation of $J_{\phi}$ is pristine to errors in $z$, with only numerical
+artifacts remaining. Only small errors are induced in $J_R$, with the middle
+$90\%$ of values over the $\Gyr$ being within $\sim8\%$ of the true $J_R$. As
 expected, large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$,
 with the middle $90\%$ of values being within $\sim43\%$ of the true $J_z$.
-The $x$ offset induces uncertainties in $J_r$ and $L_z$ of $\sim21\%$ and
+The $x$ offset induces uncertainties in $J_R$ and $J_{\phi}$ of $\sim21\%$ and
 $\sim3\%$, respectively. A $\sim3\%$ error in $J_z$ is also induced.}
@@
 arrow anchored on the true $J_z$ value. We call this $\Delta J_z$ and will use
-it (along with the similarly defined $\Delta J_r$ and $\Delta L_z$) to
+it (along with the similarly defined $\Delta J_R$ and $\Delta J_{\phi}$) to
 empirically describe the error distribution. We see that $\Delta J_z$ roughly
@@
 (i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
-show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
+show the fractional spread in the values computed for $J_z$, $J_{\phi}$, and $J_R$,
 respectively. We compute the fractional spread $\Delta J_i/J_i$ as one half
@@
 In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref},
-there is essentially no spread in the determination of $L_z$. This is expected
-since, in principle, $L_z$ only depends on the $x$- and $y$-components of the
-position and velocity of the stars\footnote{In practice, however, $L_z$ is
+there is essentially no spread in the determination of $J_{\phi}$. This is expected
+since, in principle, $J_{\phi}$ only depends on the $x$- and $y$-components of the
+position and velocity of the stars\footnote{In practice, however, $J_{\phi}$ is
 computed as part of the torus-fitting method.}, and is thus unaffected by
@@
 For the offset in the Solar radius ({\em bottom panels}), the error is largest
-for $J_r$, with some deviations resulting in $L_z$ and relatively small
+for $J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
 deviations in $J_z$. In the {\em bottom middle} and {\em bottom right} panels
@@
 induced error. The {\em left}, {\em center}, and {\em right} panels show the
-result for $J_r$, $L_z$, and $J_z$, respectively. The {\em upper} panels
+result for $J_R$, $J_{\phi}$, and $J_z$, respectively. The {\em upper} panels
 consider an offset in $z$ and the {\em lower} panels consider an offset in $x$
@@
 all three orbits the epicyclic approximation is a good description of the
-$J_z$ error. An $x$ offset induces an error in $J_r$ and $L_z$. The error in
-$J_r$ is somewhat well-described for the thin-disk and thick-disk orbits, and
-a poor description for the halo orbit. For $L_z$, the epicyclic approximation
+$J_z$ error. An $x$ offset induces an error in $J_R$ and $J_{\phi}$. The error in
+$J_R$ is somewhat well-described for the thin-disk and thick-disk orbits, and
+a poor description for the halo orbit. For $J_{\phi}$, the epicyclic approximation
 is not a good description for any orbit.}
@@
 the LSR is less pronounced than for the midplane, and since the dominant
-effect is limited to $J_r$ (Figure~\ref{fig:many_orbit_wrong_ref}), we defer
+effect is limited to $J_R$ (Figure~\ref{fig:many_orbit_wrong_ref}), we defer
 this calculation to future work.
@@
 variations seen in e.g. \citet{2019arXiv190209569F}. The velocity variations
-in this work (admittedly found with respect to $L_z$ or radius) are
+in this work (admittedly found with respect to $J_{\phi}$ or radius) are
 $\sim5\text{--}10\,\kms$. The vertical frequency of the thin-disk and

### 2019-04-22 schmactions/many_orbits/plot_many_orbits.py

@@ -144,10 +144,10 @@ for x in ax[1]:
     x.set_xticks(np.arange(xmin,xmax,100), minor=True)
 
 for x in ax[:,0]:
-    x.set_ylabel(r'$\Delta J_r/J_r\,[\,\%\,]$')
+    x.set_ylabel(r'$\Delta J_R/J_R\,[\,\%\,]$')
     x.locator_params(axis='y', nbins=6)
 for x in ax[:,1]:
-    x.set_ylabel(r'$\Delta L_z/L_z\,[\,\%\,]$')
+    x.set_ylabel(r'$\Delta J_{\phi}/J_{\phi}\,[\,\%\,]$')
     x.locator_params(axis='y', nbins=6)
 for x in ax[:,2]:
     x.set_ylabel(r'$\Delta J_z/J_z\,[\,\%\,]$')

### 2019-04-22 schmactions/one_orbit/plot_one_orbit.py

@@ -99,9 +99,9 @@ for x in ax[1]:
     x.set_xticks(np.arange(0,1000,100), minor=True)
 
 for x in ax[:,0]:
-    x.set_ylabel(r'$J_{r,\text{obs}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+    x.set_ylabel(r'$J_{R,\text{obs}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
 for x in ax[:,1]:
-    x.set_ylabel(r'$L_{z,\text{obs}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+    x.set_ylabel(r'$J_{\phi,\text{obs}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
 for x in ax[:,2]:
     x.set_ylabel(r'$J_{z,\text{obs}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
 

### 2019-04-22 schmactions/many_orbits/plot_many_orbits.py
Change yaxis range in Fig 4

@@ -30,7 +30,7 @@ xmax = 500
 y0min = 0
 y0max = 50
 y1min = 0
-y1max = 25
+y1max = 10
 y2min = 0
 y2max = 150
 

### 2019-04-22 paper/ms.tex
Small typo fixes from MKN

@@
 
-Stellar actions, computed for stars from their six-dimensional phase space measurement and
+Stellar actions, computed for stars from their six-dimensional phase space measurements and
 an assumed Galactic potential, are used to label and distinguish orbits. In
@@
 
-\subsection{Local Midplane} \label{ssec:local_midplane}
+\subsection{The Local Midplane} \label{ssec:local_midplane}
 Using the three simulations, we determine the local midplane that an observer
@@
 within a $2\,\kpc$ height of the fiducial midplane to determine $1\,\sigma
-$error bars, which we report as dashed lines.}
+$ error bars, which we report as dashed lines.}
 \label{fig:midplane}
@@
 \begin{figure*}
-\plotone{fig/range_dphi.pdf}
+\begin{center}
+\includegraphics[width=6.8in]{fig/range_dphi.pdf}
+\end{center}
 \caption{The range of midplane heights encountered as a function of angular
@@
 Table~\ref{tab:real_clusters}. Except for alphaPer, we find that the
-neigbhorhood around each cluster is typically $\gtrsim x\,\pc$. This analysis
+neighborhood around each cluster is typically $\gtrsim x\,\pc$. This analysis
 assumes that the MW resembles \mi{}, which is not necessarily true. However,

### 2019-04-23 paper/ms.tex
Add ref to Kamdar paper
Update title, subsection titles

@@
 
-\title{The Galactic Midplane is Not a Plane: Implications for dynamical analysis with {\em Gaia} data and beyond}
-
-% \correspondingauthor{Angus Beane}
-%mm [will this address go away during the review process?.  Do you
-%want to give a "current address" as a second footnote?]
+\title{The Galactic Midplane Is Not a Plane: Implications for Dynamical
+Analysis with {\em Gaia} Data and Beyond}
+
+\correspondingauthor{Angus Beane}
 \email{abeane@sas.upenn.edu}
@@
 
-\subsection{Cartoon effect of midplane error} \label{ssec:cartoon}
+\subsection{Cartoon Effect of Midplane Error} \label{ssec:cartoon}
 We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
@@
 
-\subsection{Numerical methods} \label{ssec:action_comp}
+\subsection{Numerical Methods} \label{ssec:action_comp}
 We now move on to quantifying the argument made in Section~\ref{ssec:cartoon}.
@@
 
-\subsection{Quantification of the midplane effect} \label{ssec:quant}
+\subsection{Quantification of the Midplane Effect} \label{ssec:quant}
 We quantify how a systematic error in the Galactocentric coordinate system
@@
 
-% \section{Azimuthal midplane variations at the Solar Circle: estimates from
-% simulation} 
 \section{Coordinate System Offsets II. Azimuthal Midplane Variations in
@@
 
-\subsection{Description of FIRE Simulation} \label{ssec:cosmozoom}
+\subsection{Description of FIRE Simulations} \label{ssec:cosmozoom}
 We briefly describe the cosmological-hydrodynamical zoom-in simulations used
@@
 
-\subsection{The size of the solar neighborhood} \label{sssec:neighborhood}
+\subsection{The Size of the Solar Neighborhood} \label{sssec:neighborhood}
 Given a local coordinate system, the amplitude and characteristic length scale
@@
 
-\subsection{Consequences for disrupted cluster reconstruction}
+\subsection{Disrupted Cluster Reconstruction}
 \label{sssec:reconstruction}
@@
 Chemical tagging \citep{2002ARA&A..40..487F} and dynamical tagging must
-complement and confirm each other. This was also recently pointed out by
-\Gus{someone}.
+complement and confirm each other. The importance of combining chemistry and
+dynamics for Galactic archaeology was recently pointed out by
+\citet{2019arXiv190210719K}.
 

### 2019-04-23 paper/ms.tex
Tidy up list of GC params

@@
 defined and measured both precisely and accurately. This involves measuring
-the Galactic center, the orientation of the Galactic plane, the distance from
-the Sun to the Galactic center, the height of the Sun above the Galactic
-midplane, and the local standard of rest (LSR). We discuss the observational
-efforts to measure all these parameters in Section~\ref{sec:mes_err}.
+the angular position of and distance to the Galactic center, the orientation
+of the Galactic plane, the height of the Sun above the Galactic midplane, and
+the local standard of rest (LSR). We discuss the observational efforts of all
+of these parameters in Section~\ref{sec:mes_err}.
 

### 2019-04-23 paper/ms.tex
Add citations to stuff about uses of actions

@@
 and other intrinsic, and at least partially invariant, properties such as age
-or metallicity. Actions also provide a convenient basis for constructing
-models of the stellar distribution function, or for associating stars with
-similar dynamical properties (e.g. to determine membership in moving groups).
+or metallicity \citep{2018ApJ...867...31B, 2018arXiv180803278T,
+2018MNRAS.481.4093S, 2019arXiv190304030G, 2019arXiv190309320D,
+2019MNRAS.486.1167B}. Actions also provide a convenient basis for constructing
+models of the stellar distribution function
+\citep[e.g.][]{1915MNRAS..76...70J, 1985ApJ...295..388V, 2017ApJ...839...61T},
+or for associating stars with similar dynamical properties (e.g. to determine
+membership in moving groups).
 

### 2019-04-23 paper/ms.tex
Rephrase intro paragraph to fit new structure

@@
 will not agree with our local midplane extrapolated onto their position,
-thanks to relatively small-scale variations in the density of the gas and
+due to relatively small-scale variations in the density of the gas and
 stellar distribution. Converting positions and velocities of more distant
@@
 argument applies to any remaining uncertainty in measurements of the
-Galactocentric radius of the Sun $R_{\odot}$ and to variations in the local
+Galactocentric radius of the Sun and to variations in the local
 standard of rest (LSR).
@@
 In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
-system errors have on the measured actions. Lacking an empirical, nonlocal
-definition of the global Galactic midplane or a measurement of the expected
-midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire}
-we examine the expected azimuthal dependence of the local midplane, and hence
-the expected size of the dynamical solar neighborhood, using cosmological,
-hydrodynamical, zoom-in simulations of Milky Way-mass galaxies from the
-Feedback in Realistic Environments (FIRE)
+system errors have on the measured actions. We then discuss two potential
+sources of these coordinate system offsets in
+Sections~\ref{sec:mes_err}~and~\ref{sec:local_file}. In
+Section~\ref{sec:mes_err} we discuss the potential errors in measuring the
+parameters of the Galactocentric coordinate system. In
+Section~\ref{sec:local_fire} we examine the azimuthal variations of the
+midplane itself in two sets of simulations: cosmological, hydrodynamical,
+zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic
+Environments (FIRE)
 collaboration\footnote{\url{https://fire.northwestern.edu}}
-\citep{2016ApJ...827L..23W, 2017arXiv170206148H}.
+\citep{2016ApJ...827L..23W, 2017arXiv170206148H}, and a controlled N-body
+simulation of a Sagittarius encounter with a Milky Way-mass galaxy
+\citep{2018MNRAS.481..286L}.
 

### 2019-04-24 paper/ms.tex
Add transition par. to Section 3

@@
 \section{Coordinate System Offsets I. Measurement Errors} \label{sec:mes_err}
+We consider the first potential source of coordinate system offsets which
+would generate the action errors described in Section~\ref{sec:ref_frame}:
+errors in the determination of the Galactocentric parameters. Converting from
+heliocentric to Galactocentric coordinates relies on these measurements.
+Therefore, errors in their values will induce a systematic offset in the
+Galactocentric phase-space position of any observed star.
 

### 2019-04-24 paper/ms.tex
Add paragraph about coordinate system errors

@@
 
+Galactocentric coordinate system offsets can also be generated by errors in
+the parameters of the Galactocentric coordinate system. However, such
+parameters have been measured to high enough precision that they are likely
+subdominant to the intrinsic variation of the midplane across large distances.
+On shorter scales, they are likely the dominant source of error.
+
 This finding has many implications for the study of the dynamics of stars in

### 2019-04-24 paper/ms.tex
Update ack
Small changes throughout

@@
 assumed potential model. The $y$-axis corresponds to the vertical height of
-the orbit and the $x$-axis orbital phase. The solid gray line indicates the
+the orbit and the $x$-axis orbital phase. The solid gray curve indicates the
 ``true'' orbit of the star; that is, its orbit as it oscillates around the
@@
 point or the green point (i.e. at two different orbital phases). Then the
-purple and green lines correspond to the orbits that the observer would
+purple and green curves correspond to the orbits that the observer would
 compute for each point based on the potential model with the erroneous
 midplane. In action space, this would correspond to a different value of $J_z$
-for the blue and pink points. In this way, assuming the wrong coordinate
+for the purple and green points. In this way, assuming the wrong coordinate
 system induces a phase-dependence in the actions estimated for the star, which
@@
 \eeq
-Averaging over the circle, we therefore have that
+Averaging over the circle, we therefore have that,
 \beq
@@
 
-We begin by considering the thick-disk orbit. Suppose that we could observe
+We begin by considering the thick-disk orbit. Suppose an immortal observer who can observe
 the thick-disk star's phase-space position at many different times (and hence
@@
 i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position in the
-orbit. This corresponds to an observer physically located at e.g. the position
+orbit. This corresponds to the immortal observer physically located at e.g. the position
 $(8, 0, 0)\,\kpc$ in the coordinate system of the true potential, but
@@
 Starting from each observation of the star's phase-space position that such an
-(extremely long-lived) observer makes every $\Myr$, we then compute the
+immortal observer makes every $\Myr$, we then compute the
 actions as outlined above using an integration from that starting point in
@@
 offset coordinate systems depend on the time (i.e. orbital phase) at which the
-star's phase-space position is observed. To our extremely long-lived observer
+star's phase-space position is observed. To the immortal observer
 this would appear as time-dependence in the actions, which appear to oscillate
@@
 95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
-$47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively. These are
-fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively.
+$47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively.
+These are fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively ---
+despite only a $1.2\%$ error in the distance to the Galactic center.
 
@@
 $J_z$ for the thick-disk orbit ({\em upper panel}) and thin-disk orbit ({\em
-lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. the {\em right} panel
+lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. the {\em upper right} panel
 of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted
@@
 error distribution is not even centered on null. This comes about when the
-midplane error is $\gtrsim z_{\text{max}}$, and can be understood by realizing
-that in such a case the midplane offset always acts to increase the value of
-$J_z$, resulting in a biased estimate of $J_z$.
-
-The intuition behind the shape of Figure~\ref{fig:Jz_hist} is relatively
-simple. Consider first the thick-disk orbit ({\em top} panel), where the
+midplane error is $\gtrsim z_{\text{max}}$.
+
+We now describe the intuition behind the shape of Figure~\ref{fig:Jz_hist}.
+Consider first the thick-disk orbit ({\em top} panel), where the
 offset in $z$ is much less than $z_{\text{max}}$ of the orbit. The peaks in
@@
 95\uth{} minus 5\uth{} percentile divided by the true action value ($\Delta
-J_i/J_i$) for the three different fiducial orbits in Table \ref{tab:orbits}.
+J_i/J_i$) for the three different fiducial orbits in Table~\ref{tab:orbits}.
 The thick-disk (\thickcolor) orbit is the one from
@@
 (disk-like) orbits. For the thin-disk orbit, a systematic offset of
-$\sim30\,\pc$ in the $z$ coordinate already results in $100\%$ deviations in
-the actions. The required offset for $100\%$ deviation is $\sim225\,\pc$ for
+$\sim15\,\pc$ in the $z$ coordinate already results in $25\%$ deviations in
+the actions. The required offset for $25\%$ deviation is $\sim120\,\pc$ for
 the thick-disk orbit. The halo orbit is relatively resistant to errors in the
-midplane, with only $\sim25\%$ error in $J_z$ out to a midplane offset of
+midplane, with only $\sim15\%$ error in $J_z$ out to a midplane offset of
 $500\,\pc$.
@@
 fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of $10, 50,
-100\,\pc$ (red, cyan, green, respectively) in Figure~\ref{fig:dJz_fun_Jz}. For
-a thin disk-like orbit ($J_z\sim0.5\,\actunit$), even a $10\,\pc$ offset in
-$z$ is enough to induce a $\sim50\%$ error in $J_z$. For larger values of
-$J_z$, the fractional errors are suppressed, but the induced error can still
-be large depending on how great the $z$ offset is.
+100\,\pc$ (orange, cyan, and green, respectively) in
+Figure~\ref{fig:dJz_fun_Jz}. For a thin disk-like orbit
+($J_z\sim0.5\,\actunit$), even a $10\,\pc$ offset in $z$ is enough to induce a
+$\sim25\%$ error in $J_z$. For larger values of $J_z$, the fractional errors
+are suppressed, but the induced error can still be large depending on how
+great the $z$ offset is. We also plot the epicylic prediction for $\Delta J_z
+/ J_z$ from Equation~\ref{eq:actions_epi_error} as dashed lines for each $z$
+offset. We find that the epicyclic approximation matches the numerical
+estimate quite well.
 
@@
 was defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
-defining the location of the Galactic center and the north Galactic pole.
+defining the coordinates of the Galactic center in B1950 coordinates as
+(17:42:26.6, -28:55:00) and the north Galactic pole as (12:49:00, +27:24:00).
 However, there is growing evidence that the true midplane is tilted when
@@
 \subsection{Solar Height}
-Thirdly, one must define The Sun's vertical distance from the Galactic
+Thirdly, one must define the Sun's vertical distance from the Galactic
 midplane, which can be determined by identifying where the stellar density
 and velocities reach a maximum (effectively the median height of all disk
-stars) The Solar height is usually taken to be $\sim 25\,\pc$
+stars). The Solar height is usually taken to be $\sim 25\,\pc$
 \citep{2001ApJ...553..184C}, with a more recent measurement from {\em Gaia} DR2
@@
 
-We now focus our attention on understanding the intrinsic variations of the
-midplane across the disk in two sets of complementary simulations. We will
+We now consider a second source of systematic offsets in the Galactocentric
+coordinate system --- intrinsic variations of the midplane across the disk. We
+study this in two sets of simulations of Milky Way-mass galaxies. We will also
 briefly discuss a rough estimate of the midplane variation derived from the
@@
 radius of $3.5\,\kpc$, scale height $0.53\,\kpc$, and mass
-$6\times10^{10}\Msun$, a Hernquist bulge of mass $10^{10}\,\Msun$ and scale
+$6\times10^{10}\,\Msun$, a Hernquist bulge of mass $10^{10}\,\Msun$ and scale
 length $0.7\,\kpc$ \citep{1990ApJ...356..359H}. Sagittarius is modelled as a
@@
 
+The mass resolution of the simulation is $2.6\times10^4$, $1.2\times10^4$, and
+$1.0\times10^4$ for the dark matter, disk, and bulge components, respectively.
 For the disk and bulge components, a softening length of $30\,\pc$ is used
@@
 \subsection{The Local Midplane} \label{ssec:local_midplane}
-Using the three simulations, we determine the local midplane that an observer
+Using the two sets of simulations, we determine the local midplane that an observer
 might measure if they were situated in each of these galaxies as a function of
@@
 observer at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth
-between $0<\phi<2\pi$\footnote{Our very long-lived observer also has warp
-drive.}. At each value of $\phi$ we then compute the median $z$ for stars
+between $0<\phi<2\pi$.\footnote{The immortal observer also has warp
+drive.} At each value of $\phi$ we then compute the median $z$ for stars
 within a cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to
-the fiducial disk. We then re-define this median $z$ as the new midplane of
-the cylinder, re-select stars, and iterate until the median $z$ value
+the fiducial disk. We then re-define the new midplane of the cylinder to be the median $z$, re-select stars, and iterate until the median $z$ value
 converges. We find that only $10$ iterations of this procedure are necessary
@@
 bootstrap-resample stars within a cylinder of stars with height $2\,\kpc$ and
-the same radius $1000$ times and determine the $1\,\sigma$ error bars. This
-bootstrap resampling is performed by first selecting all stars within a height
-of $2\,\kpc$, resampling that selection, and then repeating the $10$
-iterations.
+the same radius $1000$ times and determine the $1\,\sigma$ error bars by repeating the midplane determination with those reselection of stars.
 
@@
 form $\text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}$, with the values
-for $\text{A}$ being $9.5, 2.5, -212, -394\,\pc$, for $\text{B}$ being
-$0.0013, 0.0007, -0.63, -1.00\,\text{rad}$ and $\text{C}$ being $2.6, -7.8,
--65, -53\,\pc$. Each of these values are for each successive timestep. In the
+for $\text{A}$ being $9.5$, $2.5$, $-212$, and $-394\,\pc$, for $\text{B}$ being
+$0.0013$, $0.0007$, $-0.63$, and $-1.00\,\text{rad}$ and $\text{C}$ being $2.6$, $-7.8$,
+$-65$, and $-53\,\pc$. Each of these values are for each successive timestep. In the
 first two panels, we see that the midplane is relatively flat in the inner
@@
 ($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} ({\em left},
-{\em center}, and {\em right} panels). The local midplane is determined at a
+{\em center}, and {\em right} panels). All calculations are done at cosmological redshift $\z =0$. The local midplane is determined at a
 position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
@@
 $\sim5\text{--}10\,\kms$. The vertical frequency of the thin-disk and
-thick-disk orbits are $\sim 0.09\,1/\Myr$ and $\sim0.06\,1/\Myr$,
+thick-disk orbits are $\sim 0.09\,\Myr^{-1}$ and $\sim0.06\,\Myr^{-1}$,
 respectively. By dimensional analysis, we therefore expect the midplane
@@
 disrupted. We would like to determine which stars with complete 6D phase-space
-positions, for example in the {\em Gaia} dataset,\footnote{This is likely to
+positions, for example in the {\em Gaia} dataset, are likely to be ejected members from that cluster.\footnote{This is likely to
 be more successful with DR3/4 than DR2, since upcoming data releases will
@@
 targeted, follow-up radial velocity surveys may allow this program to be
-performed sooner.} are likely to be ejected members from that cluster. One
+performed sooner.} One
 could generate an initial catalogue of candidate members by integrating the
@@
 comparable to the intrinsic action-space dispersion of the disrupting
-cluster. This dispersion can be estimated as \citep[Section~8.3.3
-of][]{2008gady.book.....B}
+cluster. This dispersion can be estimated as \citep[Section~8.3.3][]{2008gady.book.....B},
 \beq \label{eq:action_disp}
@@
 system induces time evolution in the {\em observed} actions
-(Figures~\ref{fig:cartoon}\&\ref{fig:one_orbit_wrong_ref}).
-
-\item We interpret this time evolution as an error in the computed action
-(Figure~\ref{fig:many_orbit_wrong_ref}). We focused on the effect that an
+(Figures~\ref{fig:cartoon}~and~\ref{fig:one_orbit_wrong_ref}). We interpret
+this time evolution as an error in the computed action
+(Figure~\ref{fig:many_orbit_wrong_ref}).
+
+\item  We focused on the effect that an
 inaccuracy in the midplane has on the vertical action $J_z$. We found that in
@@
 As a result, error
-propagation is 
+propagation of systematic coordinate system errors is 
      complex
@@
 
-\item To demonstrate the previous point, we measured the local Galactic
+\item To demonstrate the previous point, we measured the local galactic
 midplane along the Solar circle in three different high-resolution, zoom-in
@@
 \acknowledgments
-We would like to thank Megan Bedell, Tobias Buck, and Todd Phillips
-A.B. would like to thank Todd Phillips
-for helpful discussions. A.B. was supported in part by the Roy \& Diana
-Vagelos Program in the Molecular Life Sciences and the Roy \& Diana Vagelos
-Challenge Award. 
+We would like to thank Megan Bedell, Tobias Buck, Elena D'Onghia, Benoit
+Famaey, and Adrian Price-Whelan for helpful discussions. A.B. would like to
+thank Todd Phillips for helpful discussions. A.B. was supported in part by the
+Roy \& Diana Vagelos Program in the Molecular Life Sciences and the Roy \&
+Diana Vagelos Challenge Award.
 %mm The work of \Gus{everyone} is supported by the Simons

### 2019-04-24 cartoon/cartoon.py
Update zorder in cartoon

@@ -19,11 +19,11 @@ y3 = 2.25*np.cos(x) - 1.25
 
 ax.plot(x, y, c=tb_c[-1])
 
-ax.plot(x, y2, c=tb_c[4])
-ax.scatter(3.*np.pi, -1, c=tb_c[4])
+ax.plot(x, y2, c=tb_c[4], zorder=4)
+ax.scatter(3.*np.pi, -1, c=tb_c[4], zorder=4)
 
-ax.plot(x, y3, c=tb_c[6])
-ax.scatter(4.*np.pi, 1, c=tb_c[6])
+ax.plot(x, y3, c=tb_c[6], zorder=5)
+ax.scatter(4.*np.pi, 1, c=tb_c[6], zorder=5)
 
 ax.plot(x, np.full(len(x), 0), c=tb_c[-1])
 ax.plot(x, np.full(len(x), -1.25), c=tb_c[-1], ls='dashed')

### 2019-04-24 cartoon/cartoon.py
Standardize all fig widths

@@ -10,7 +10,10 @@ mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
         '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
 
-fig, ax = plt.subplots(1, figsize=(7, 3))
+textwidth = 7.10000594991
+columnwidth = 3.35224200913
+
+fig, ax = plt.subplots(1, figsize=(textwidth, 3))
 
 x = np.linspace(0, 7*np.pi, 5000)
 y = np.cos(x)

### 2019-04-24 lsr_test/plot_lsr.py

@@ -7,12 +7,15 @@ rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
+textwidth = 7.10000594991
+columnwidth = 3.35224200913
+
 tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
         '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
 
 gal_list = ['m12i', 'm12f', 'm12m']
 
-fig, ax = plt.subplots(3, 3, sharex=True, sharey='row')
+fig, ax = plt.subplots(3, 3, figsize=(textwidth, 5), sharex=True, sharey='row')
 
 for gal, x in zip(gal_list, ax.transpose()):
     fname = 'output/lsr_'+gal+'.npy'
@@ -43,7 +46,7 @@ ax[0][0].set_ylabel(r'$v_{R,\text{LSR}}\,[\,\text{km}/\text{s}\,]$')
 ax[0][0].set_ylim(-100, 100)
 
 ax[1][0].set_ylabel(r'$v_{z,\text{LSR}}\,[\,\text{km}/\text{s}\,]$')
-ax[1][0].set_ylim(-50, 50)
+ax[1][0].set_ylim(-20, 20)
 
 ax[2][0].set_ylabel(r'$v_{\phi,\text{LSR}}\,[\,\text{km}/\text{s}\,]$')
 ax[2][0].set_ylim(-300, -150)

### 2019-04-24 midplane_test/FIRE/plot_deltaphi.py

@@ -11,6 +11,9 @@ rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
+textwidth = 7.10000594991
+columnwidth = 3.35224200913
+
 tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
         '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
 
@@ -63,7 +66,7 @@ def get_range_vs_dphi(theta, midplane):
     return np.array(dphi_list_list), np.array(r_list_list)
 
 
-fig, ax = plt.subplots(1, 3, sharex=True, sharey=True, figsize=(8,2.25))
+fig, ax = plt.subplots(1, 3, sharex=True, sharey=True, figsize=(textwidth,2))
 # now make paper plot, with just fit
 for gal,ax_col in zip(glist, ax.transpose()):
     out = np.load('output/out_'+gal+'.npy')
@@ -122,6 +125,8 @@ for gal,ax_col in zip(glist, ax.transpose()):
     t = print_dphi(dphi_mean, r_mean)
     print(tick_function(t/np.pi, round=False))
 
+    ax_col.locator_params('x', nbins=5)
+
     ax2 = ax_col.twiny()
     #ax2.invert_xaxis()
     ax2.set_xticks(ax_col.get_xticks())
@@ -129,6 +134,7 @@ for gal,ax_col in zip(glist, ax.transpose()):
     ax2.set_xticklabels(tick_function(ax_col.get_xticks()))
     #ax2.set_xlabel(r'$\nu\,[\,\text{MHz}\,]$')
     ax2.set_xlabel(r'$\text{chord length}\,[\,\text{kpc}\,]$')
+
 
 
 ax[0].set_ylabel(r'$\text{range}\,[\,\text{pc}\,]$')

### 2019-04-24 midplane_test/FIRE/plot_midplane.py

@@ -9,6 +9,9 @@ rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
+textwidth = 7.10000594991
+columnwidth = 3.35224200913
+
 tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
         '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
 
@@ -22,7 +25,7 @@ nspoke = 50
 
 glist = ['m12i', 'm12f', 'm12m']
 
-fig, ax = plt.subplots(2, 3, sharex=True, sharey=True, figsize=(8,3.5))
+fig, ax = plt.subplots(2, 3, sharex=True, sharey=True, figsize=(textwidth,3.2))
 # first make midplane comparison plot
 for gal,ax_col in zip(glist, ax.transpose()):
     out = np.load('output/out_'+gal+'.npy')
@@ -56,6 +59,9 @@ for gal,ax_col in zip(glist, ax.transpose()):
     ax_col[0].set_ylim(-200, 200)
     ax_col[1].set_ylim(-200, 200)
 
+    ax_col[0].locator_params('x', nbins=5)
+    ax_col[1].locator_params('y', nbins=5)
+
     print(gal, '90 perc: ', np.percentile((midplane_est-fit)*1000, 95) - np.percentile((midplane_est-fit)*1000, 5))
 
     ax_col[1].plot(theta/np.pi, (midplane_est-fit)*1000, c=tb_c[0])
@@ -77,7 +83,7 @@ plt.savefig('midplane.pdf')
 
 
 
-fig, ax = plt.subplots(1, 3, sharex=True, figsize=(8,2))
+fig, ax = plt.subplots(1, 3, sharex=True, figsize=(textwidth,1.8))
 # now make paper plot, with just fit
 for gal,x in zip(glist,ax):
     out = np.load('output/out_'+gal+'.npy')
@@ -108,6 +114,9 @@ for gal,x in zip(glist,ax):
                horizontalalignment='left', 
                verticalalignment='center', 
                transform = x.transAxes)
+
+    x.locator_params('x', nbins=5)
+    x.locator_params('y', nbins=5)
 
     x.set_xlim(0, 2)
     x.set_xlim(0, 2)

### 2019-04-24 midplane_test/chervin_sag_sim/plot_midplane.py

@@ -10,6 +10,9 @@ rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
+textwidth = 7.10000594991
+columnwidth = 3.35224200913
+
 tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
 '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
 
@@ -39,7 +42,7 @@ glist = ['Disk_200', 'Disk_400', 'Disk_600', 'Disk_690']
 column = [0,1,2,3]
 time = [r'$t=2.0\,\text{Gyr}$', r'$t=4.0\,\text{Gyr}$', 
         r'$t=6.0\,\text{Gyr}$', r'$t=6.9\,\text{Gyr}$']
-fig,axs = plt.subplots(2,4,figsize=(11,3.5))
+fig,axs = plt.subplots(2,4,figsize=((2/3)*textwidth,4))
 for gal,col,tm in zip(glist,column,time):
     out = np.load('out_'+gal+'.npy')
     theta = out[:,0]
@@ -84,7 +87,7 @@ axs[0,0].set_title("Chervin's Simulation")
 fig.tight_layout()
 plt.savefig('midplane.pdf')
 
-fig, ax = plt.subplots(2, 2, sharex=True, figsize=(5.5,3.5))
+fig, ax = plt.subplots(2, 2, sharex=True, figsize=((2/3)*textwidth,3.5))
 # now make paper plot, with just fit
 for gal,x,tm in zip(glist,ax.flatten(),time):
     out = np.load('out_'+gal+'.npy')
@@ -110,6 +113,8 @@ for gal,x,tm in zip(glist,ax.flatten(),time):
 
     x.set_ylim(-200, 200)
     x.set_ylim(-200, 200)
+
+    x.locator_params('x', nbins=5)
 
     x.plot(theta/np.pi, (midplane_est-fit)*1000, c=tb_c[0])
     x.plot(theta/np.pi, (err_low-fit)*1000, c=tb_c[0], ls='dashed', alpha=0.5)

### 2019-04-24 schmactions/cluster_offset/cluster_eval.py

@@ -21,6 +21,9 @@ rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
+textwidth = 7.10000594991
+columnwidth = 3.35224200913
+
 tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
         '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
 
@@ -77,7 +80,7 @@ init_vel_list = [[0, -190, 10] * u.km/u.s,
 clist = [tb_c[7], tb_c[8], tb_c[0]]
 
 # set up figure
-fig, ax = plt.subplots(1, 1, figsize=(4, 3))
+fig, ax = plt.subplots(1, 1, figsize=(columnwidth, 2.5))
 
 for fname, name, init_vel, c in zip(fname_list, name_list, init_vel_list, clist):
     offlist, dJzJz = load_orbit(fname, init_pos, init_vel)
@@ -104,7 +107,7 @@ plt.close()
 # now begin making Rn vs mc plot
 
 glist = ['m12i', 'm12f', 'm12m']
-fig, ax_list = plt.subplots(1, 3, figsize=(8, 3))
+fig, ax_list = plt.subplots(1, 3, figsize=(textwidth, 2.7))
 
 def chord(l):
         return np.round(2*Rsolar*np.sin(0.5*(l/2)*np.pi), 1)
@@ -162,10 +165,10 @@ for ax, gal in zip(ax_list, glist):
         if gal=='m12f':
             l = 0.05
         elif gal=='m12m':
-            l = 0.15
+            l = 0.4
         elif gal=='m12i':
             l = 0.12
-        ax.text(l, 0.88, gal, 
+        ax.text(l, 0.88, r'\texttt{'+gal+r'}', 
                horizontalalignment='left', 
                verticalalignment='center', 
                transform = ax.transAxes)

### 2019-04-24 schmactions/many_orbits/plot_many_orbits.py

@@ -21,6 +21,9 @@ rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
+textwidth = 7.10000594991
+columnwidth = 3.35224200913
+
 tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
         '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
 
@@ -55,7 +58,7 @@ Az_list = np.array([120, 850, 6160]) # pc
 AR_list = np.array([1290, 1190, 2340]) # pc
 Rg_list = np.array([5420, 5610, 6200]) # pc
 
-fig, ax = plt.subplots(2, 3, figsize=(7, 4))
+fig, ax = plt.subplots(2, 3, figsize=(textwidth, 4))
 init_pos = [8, 0, 0] * u.kpc
 init_vel_list = [[0, -190, 10] * u.km/u.s,
                  [0, -190, 50] * u.km/u.s,

### 2019-04-24 schmactions/many_orbits/plot_many_orbits_Jzfun.py

@@ -21,6 +21,9 @@ rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
+textwidth = 7.10000594991
+columnwidth = 3.35224200913
+
 tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
         '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
 
@@ -45,7 +48,7 @@ def sclip(a, s=4):
     k2 = np.where(a2bool)[0]
     return k0, k1, k2
 
-fig, ax = plt.subplots(1, 1, figsize=(3.5, 3.5))
+fig, ax = plt.subplots(1, 1, figsize=(columnwidth, columnwidth))
 init_pos = [8, 0, 0] * u.kpc
 init_vel = [0, -190, 10] * u.km/u.s
 

### 2019-04-24 schmactions/one_orbit/plot_one_orbit.py

@@ -16,6 +16,9 @@ rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath} \usepackage{bm}']
 
+textwidth = 7.10000594991
+columnwidth = 3.35224200913
+
 tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
         '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
 
@@ -73,7 +76,7 @@ xtime = s.extract_time(xout)
 zact_thin = sthin.extract_actions(zout_thin)
 ztime_thin = sthin.extract_time(zout_thin)
 
-fig, ax = plt.subplots(2, 3, figsize=(7, 4), sharex=True)
+fig, ax = plt.subplots(2, 3, figsize=(textwidth, 4), sharex=True)
 
 for x,t,a in zip(ax, (ztime, xtime), (zact, xact)):
     # get keys corresponding to 4 sigmaclip
@@ -112,7 +115,7 @@ fig.tight_layout()
 plt.savefig('schmactions_one_orbit.pdf')
 plt.close()
 
-fig, ax = plt.subplots(2,1, figsize=(3,6))
+fig, ax = plt.subplots(2,1, figsize=(columnwidth,6.5))
 
 k0, k1, k2 = sclip(zact)
 dJzJz = 0.5*(np.percentile(zact[:,2][k2], 95) - np.percentile(zact[:,2][k2], 5))

### 2019-04-24 schmactions/orbits_plot/plot_orbits.py

@@ -12,6 +12,9 @@ rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
 rc('text', usetex=True)
 mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
 
+textwidth = 7.10000594991
+columnwidth = 3.35224200913
+
 tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
         '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
 
@@ -21,7 +24,7 @@ init_vel_list = [[0, -190, 10] * u.km/u.s,
                  [0, -190, 50] * u.km/u.s,
                  [0, -190, 190] * u.km/u.s]
 
-fig, ax = plt.subplots(2, 3, sharey='row', figsize=(7,5.5))
+fig, ax = plt.subplots(2, 3, sharey='row', figsize=(textwidth,5.5))
 
 for init_vel, x in zip(init_vel_list, np.transpose(ax)):
     s = schmactions(init_pos, init_vel, save_orbit=True)

### 2019-04-24 paper/ms.tex
Make figures in paper correct widths

@@
 \begin{figure*}
-\plotone{fig/cartoon.pdf}
+\begin{center}
+\includegraphics[width=\textwidth]{fig/cartoon.pdf}
+\end{center}
 \caption{Cartoon approximation showing the effect an error in the determination of the
@@
 \begin{figure*}
-\plotone{fig/schmactions_one_orbit.pdf}
+\begin{center}
+\includegraphics[width=\textwidth]{fig/schmactions_one_orbit.pdf}
+\end{center}
 \caption{The artificial phase-dependence in the observed actions induced by an
@@
 \begin{figure}
-\plotone{fig/schmactions_Jz_hist.pdf}
+\begin{center}
+\includegraphics[width=\columnwidth]{fig/schmactions_Jz_hist.pdf}
+\end{center}
 \caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
@@
 \begin{figure*}
-\plotone{fig/schmactions_many_orbits.pdf}
+\begin{center}
+\includegraphics[width=\textwidth]{fig/schmactions_many_orbits.pdf}
+\end{center}
 \caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error
@@
 \begin{figure}
-\plotone{fig/schmactions_many_orbits_Jz_fun.pdf}
+\begin{center}
+\includegraphics[width=\columnwidth]{fig/schmactions_many_orbits_Jz_fun.pdf}
+\end{center}
 \caption{The fractional error in $J_z$ as a function of $J_z$ for a few
@@
 \begin{center}
-\includegraphics[width=7in]{fig/midplane_fit.pdf}
+\includegraphics[width=\textwidth]{fig/midplane_fit.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=5.5in]{fig/midplane_fit_chervinsim.pdf}
+\includegraphics[width=342.078286667pt]{fig/midplane_fit_chervinsim.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=6.8in]{fig/range_dphi.pdf}
+\includegraphics[width=\textwidth]{fig/range_dphi.pdf}
 \end{center}
@@
 \begin{figure}
-\plotone{fig/cluster_offset.pdf}
+\begin{center}
+\includegraphics[width=\columnwidth]{fig/cluster_offset.pdf}
+\end{center}
 \caption{The $z$ offset necessary in order for the action error induced by our
@@
 \begin{figure*}
-\plotone{fig/Rn_vs_mc.pdf}
+\begin{center}
+\includegraphics[width=\textwidth]{fig/Rn_vs_mc.pdf}
+\end{center}
 \caption{The dynamical neighborhood (discussed in
@@
 \begin{figure*}
-\plotone{fig/orbits.pdf}
+\begin{center}
+\includegraphics[width=\textwidth]{fig/orbits.pdf}
+\end{center}
 \caption{The three orbits presented in Table~\ref{tab:orbits} and considered
@@
 \begin{figure*}
-\plotone{fig/lsr.pdf}
+\begin{center}
+\includegraphics[width=\textwidth]{fig/lsr.pdf}
+\end{center}
 \caption{The local standard of rest (LSR) as a function of azimuth at the

### 2019-04-24 paper/ms.tex
Remove neighborhood stuff in paper

@@
 variations, and the resulting systematic uncertainty in the vertical action,
-for action-space analyses beyond the dynamical solar neighborhood. As an
-illustration, in Section \ref{sssec:neighborhood} we use these estimates to
-consider the feasibility of one science goal that relies on the invariance of
-actions---the reconstruction of open clusters across large distances using
-``dynamical tagging;'' that is, the association of stars to a birth cluster
-based on their similar actions. In light of our findings, we considered the
+for action-space analyses beyond the dynamical solar neighborhood. In light of our findings, we considered the
 orbits of the nine open clusters within $250\,\pc$ as reported by the {\em
@@
 
-\subsection{The Size of the Solar Neighborhood} \label{sssec:neighborhood}
-Given a local coordinate system, the amplitude and characteristic length scale
-of the density fluctuations in the Galactic disk will determine how far this
-coordinate system can be extended before introducing relevant systematic
-errors in the computation of actions for thin-disk stars. We consider a $10\%$
-error to be a criterion for the region around the Sun where the analysis of
-the action-space distribution of stars on quasi-circular orbits, and therefore
-the epicyclic approximation, can be expected to be valid. This leads to a
-somewhat dynamically motivated definition of the ``solar neighborhood.'' Since
-the motivation may be different for defining the solar neighborhood in, e.g.,
-studies of the chemical distribution of the Galaxy, we propose the phrase
-``dynamical solar neighborhood.'' We also point out that our use of a $10\%$
-error is arbitrary.
-
-To illustrate this concept, we first compute the absolute range of midplane
-values for a given $\Delta \phi$ segment of the midplane. With the observer's
-azimuth set to zero, we consider a region extending from $-\Delta \phi/2$ to
-$\Delta\phi/2$. To translate this angular scale into a physical scale, we also
-compute the chord length $\lambda$ for the chord extending from the observer
-to $\pm \Delta\phi/2$ (see Figure~\ref{fig:fig_to_explain}). We repeat this
-for each of $50$ bins in azimuth. Each bin is plotted in
-Figure~\ref{fig:range_deltaphi} as a gray line, with the mean value in solid
-blue and the $1\,\sigma$ dispersion as the dashed blue lines. We choose to
-only consider the Solar circle to make the point that the midplane variation
-as a function of azimuth (and not just Galactic radius) is important. However,
-we note that because we use the range of midplane values, including regions of
-each galaxy outside the Solar circle can only act to increase this range, and
-therefore reduce the size of the dynamical solar neighborhood.
-
-\begin{figure*}
-\plotone{fig/fig_to_explain.pdf}
-\caption{A cartoon explanation of Figure~\ref{fig:range_deltaphi}. An observer
-is placed at the blue x in the plane of a backgrround galaxy (lime). For a
-given $\Delta \phi$ (blue) centered on the observer, we record the range of
-midplane values for this section of the Solar circle (orange). We plot the
-value of the range against the value of $\Delta \phi$. We are also able to
-convert the value of $\Delta \phi$ into a chord length (pink), which we plot
-as a secondary, upper $x$-axis. We repeat the procedure for each initial
-$\phi$ (gray lines), and also compute the mean and $\pm1\,\sigma$ values (dark
-blue). We only consider the Solar circle to explore the impact of azimuthal
-midplane variations, but similar effects are expected for radial midplane
-variations.}
-\label{fig:fig_to_explain}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{fig/range_dphi.pdf}
-\end{center}
-\caption{The range of midplane heights encountered as a function of angular
-width. At each angle $\phi$ from Figure~\ref{fig:midplane} we consider an
-angular width of $\Delta \phi$ centered on $\phi$ and report the range of
-midplane heights within that width. We repeat the procedure for each $\phi$
-and plot the result as translucent gray lines. We also plot the mean range as
-a solid blue line and the $\pm1\sigma$ lines as dashed blue lines. The upper
-$x$-axis shows the chord length from the position $\phi$ to $\pm\Delta\phi/2$.
-The horizontal \thincolor{} and \thickcolor{} lines shows the midplane
-variation necessary to induce a $10\%$ error in $J_z$ for the thin-disk and
-thick-disk orbits, respectively. A cartoon explanation of this Figure is given
-in Figure~\ref{fig:fig_to_explain}. We only consider the Solar circle to
-explore the impact of azimuthal midplane variations, but similar effects are
-expected for radial midplane variations.}
-\label{fig:range_deltaphi}
-\end{figure*}
-
-From Figure~\ref{fig:range_deltaphi} it is clear that the range of midplane
-values spanned in a given $\Delta \phi$ quickly saturates at $\Delta \phi
-\gtrsim \pi/2$ radians (i.e. one quarter of the way around the Galactic disk
-from the Sun). At far smaller $\Delta \phi$ it reaches the level of deviation
-required for a $10\%$ systematic error in $J_z$ for our fiducial
-thin-disk-like orbit (\thincolor{} horizontal line in Figure
-\ref{fig:range_deltaphi}). For the three simulations considered, the median
-range exceeds this value at $\Delta \phi/\pi > 0.0073, 0.0096,
-0.018\,\text{rad}$, with corresponding physical scales of $\lambda \sim 93.9,
-123, 233\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. These can be
-considered estimates of the range of validity for an extension of the locally
-determined axisymmetric coordinate system, at least when it comes to the study
-of stars on quasi-circular orbits.
-
-The situation is more forgiving for the study of thick-disk-like orbits, where
-a $10\%$ systematic error in $J_z$ requires a larger midplane
-deviation (\thickcolor{} horizontal line in Figure~\ref{fig:range_deltaphi}).
-In fact, the range of midplane variation in \mi{} just barely saturates to
-cause such an error in $J_z$, with the midplane variation in \mf{} and \mm{}
-never reaching such a value. For stars on halo-like orbits, extending the
-local coordinate system globally is not likely to cause serious problems,
-since the midplane deviations are relatively small compared to those needed to
-produce $10\%$ systematic errors in the actions (\halocolor{}
-horizontal line in Figure \ref{fig:range_deltaphi}). However, we caution that
-the tendency of these midplane deviations is to \emph{inflate} the value of
-$J_z$ whenever the midplane deviation is $\gtrsim z_{\text{max}}$, scattering
-them into the region of action space normally associated with different
-stellar populations and possibly contributing to confusion there (see
-Figure~\ref{fig:Jz_hist} and discussion therein). This inflation is extremely
-non-Gaussian (as the induced error is for all orbits, e.g.
-Figure~\ref{fig:Jz_hist}), and is particularly troubling for the case of thin
-disk stars, where the necessary midplane deviation is
-$\sim5\textup{--}10\,\pc$.
-
 \subsection{Disrupted Cluster Reconstruction}
@@
 
-With a characteristic midplane offset in hand, we use the relation in
-Figure~\ref{fig:range_deltaphi}, which shows the typical midplane offset at
-each $\Delta \phi$, to convert a given midplane offset into a typical
-$\Delta\phi$. As in this Figure we assume the cluster is at the Solar circle
-to obtain a physical scale, which we refer to as the ``neighborhood radius''
-or $R_{\n}$. If the allowable midplane discrepancy is greater than the maximum
-excursions recorded in the simulated galaxy, we simply report the maximum
-chord length assuming an orbit on the Solar circle ($2\times8.2\,\kpc$). If an
-observer is attempting to reconstruct a cluster with a given mass on a given
-orbit, $R_{\n}$ gives the characteristic search distance between a field star
-and the cluster within which an observer may not need to consider differences
-in their respective local midplanes when computing actions. The result is
-shown in Figure~\ref{fig:Rn_mc}. The solid lines indicate the neighborhood
-values assuming the mean relation between midplane offset and $m_c$, and the
-dashed lines show the $+1\,\sigma$ and $-1\,\sigma$ relations based on the 50
-fiducial Solar positions in each simulation (as discussed in Figure
-\ref{fig:range_deltaphi}). That this range is relatively wide illustrates the
-strong $\phi$-dependence of the midplane offsets.
-
 \begin{figure}
@@
 
-Figure~\ref{fig:Rn_mc} shows that azimuthal midplane fluctuations will limit
-the volume over which actions can be used to match stars to open clusters on
-thin-disk-like orbits (\thincolor), depending on their mass. For
-thick-disk-like orbits (\thickcolor), the tolerances are much more forgiving:
-stars' actions can be compared over the entire Solar circle for nearly all
-cluster masses, except at the very low-mass end in \mi . The neighborhood
-around $\sim100\Msun$ clusters can be as small as $\sim x\,\pc$ but up to
-$\sim y\,\kpc$ for \mi{} and \mf{}, though the situation is marginally better
-for \mm{} with values ranging from $\sim a\,\kpc$ to $\sim b\,\kpc$. The
-neighborhood values steadily increase with $m_c$, topping out at $\sim c\,\kpc$
-for $\sim10^4\,\Msun$ clusters in
-\mi{} and \mf{}. In \mm{}, the neighborhood size reaches the maximum allowed of
-$16.4\,\kpc$ for clusters of mass $\sim m\,\Msun$. This figure also
-underlines that the degree to which such an analysis will succeed depends
-crucially on the expected degree of midplane fluctuations in the real Galaxy.
-If the Milky Way is more like \mi , the range of action-space analysis
-techniques like this is far more limited than if it more closely resembles
-\mm{}.
-
-We also compute $R_{\n}$ for each of the orbits of the real open clusters in
-Table~\ref{tab:real_clusters}, using the midplane variations from
-\mi{} as a conservative choice, since it has the largest degree of
-midplane variation (Figure~\ref{fig:midplane}). We report the neighborhood
-values for the mass of each cluster in the neighborhood column of
-Table~\ref{tab:real_clusters}. Except for alphaPer, we find that the
-neighborhood around each cluster is typically $\gtrsim x\,\pc$. This analysis
-assumes that the MW resembles \mi{}, which is not necessarily true. However,
-this step of our analysis is meant to be more illustrative than quantitative:
-this calculation should actually be performed using real measurements of
-variations in the MW's stellar midplane. Such measurements should be possible
-in the near future with {\em Gaia}, with quality distance estimates the
-crucial limiting factor, along with careful selction function modelling.
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{fig/Rn_vs_mc.pdf}
-\end{center}
-\caption{The dynamical neighborhood (discussed in
-Section~\ref{sssec:neighborhood}) around an open cluster of mass $m_c$ for the
-thin-disk, and thick-disk orbits (\thincolor{} and \thickcolor{},
-respectively; see Table~\ref{tab:orbits}) and for each of the FIRE galaxies
-considered here ---
-\mi{}, \mf{}, and \mm{} ({\em left}, {\em center}, and {\em right} panels,
-respectively). Given the $z$ offset for a cluster of mass $m_c$ computed in
-Figure~\ref{fig:cluster_offset}, we convert this to a chord length using the
-result from Figure~\ref{fig:range_deltaphi}. We interpret this chord length as
-the ``neighborhood'' around the cluster, i.e. the distance from the cluster
-one could go before the $J_z$ error induced by our midplane effect is
-comparable to the intrinsic action dispersion of the cluster. When the $z$
-offset is larger than the maximum range in Figure~\ref{fig:range_deltaphi}, we
-report the maximum chord length ($2\times8.2\,\kpc$). We aso report the same
-procedure but using the $\pm1\sigma$ lines from
-Figure~\ref{fig:range_deltaphi}, which we plot here as dashed lines. For the
-thick orbit (\thickcolor), the neighborhood is quite large and thus the effect
-is negligible. However, for the thin orbit (\thincolor), which is closest to
-the orbits of nearby open clusters (see Table~\ref{tab:real_clusters}), the
-neighborhood is only a few hundred $\pc$ for low-mass clusters.}
-\label{fig:Rn_mc}
-\end{figure*}
-
 \section{Conclusions}\label{sec:conclusion}

### 2019-04-24 paper/ms.tex
found a few more leftover instances of error language in section 2, now fixed

--- 2019-04-24 paper/ms.tex
@@
 sources of these coordinate system offsets in
-Sections~\ref{sec:mes_err}~and~\ref{sec:local_file}. In
+Sections~\ref{sec:mes_err}~and~\ref{sec:local_fire}. In
 Section~\ref{sec:mes_err} we discuss the potential errors in measuring the
@@
 \section{Motivation} \label{sec:ref_frame}
-We first demonstrate the significance of systematic errors in the
-determination of the Galactic midplane and distance to the Galactic center to
-action computations. We will find that such errors are especially important
+We first demonstrate the significance  to
+action computations of a systematic offset in the
+determination of the Galactic midplane or distance to the Galactic center. We will find that such offsets are especially important
 for disk-like orbits. The consequences we explore here may also arise from
@@
 In this work, we assume that the Galaxy is perfectly described by our model
-axisymmetric potential, and simply explore the consequences of systematic
-errors in the determination of the Galactocentric coordinate system.
-
-\subsection{Cartoon Effect of Midplane Error} \label{ssec:cartoon}
+axisymmetric potential, and simply explore the consequences of offsets in the determination of the Galactocentric coordinate system.
+
+\subsection{Effect of Midplane Offset} \label{ssec:cartoon}
 We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
-inaccurate determination of the midplane leads to a phase-dependence of the
+inaccurate or erroneous determination of the midplane (and hence the zero-point of the $z$ coordinate) leads to a phase-dependence of the
 actions calculated from an orbit specified by a phase-space point and an
 assumed potential model. The $y$-axis corresponds to the vertical height of
-the orbit and the $x$-axis orbital phase. The solid gray curve indicates the
+the orbit and the $x$-axis shows orbital phase. The solid gray curve indicates the
 ``true'' orbit of the star; that is, its orbit as it oscillates around the
-true midplane. The dashed gray line is an erroneous midplane assigned to the
-potential model which an observer uses to integrate the orbit of the star and
-hence estimate its actions. The model potential is otherwise identical to the
-one in which the star is actually moving.
-
-Now suppose this observer makes a measurement of the true orbit at the purple
+true midplane. The dashed gray line is the midplane location in the
+potential model used to integrate the orbit of the star and
+hence estimate its actions, which is offset from the true midplane of the star's orbit. The model potential is otherwise identical to the one in which the star is actually moving.
+
+Now suppose this observer makes a measurement of the star's position and velocity at the purple
 point or the green point (i.e. at two different orbital phases). Then the
 purple and green curves correspond to the orbits that the observer would
-compute for each point based on the potential model with the erroneous
+compute for each point based on the potential model with the offset
 midplane. In action space, this would correspond to a different value of $J_z$

### 2019-04-24 paper/ms.tex
mostly superficial revisions to 2.2. Added a reference to a paper supporting the assumption of a flat rotation curve near the solar circle.

@@
 \subsection{Epicyclic Approximation} \label{ssec:epi_action}
-Before turning to numerical methods, we first write down analytically the
-induced action error from either a midplane offset or a Galactocentric radius
-offset. We make the epicyclic approximation, which assumes that the motion in
-the $z$ and $R$ components of the orbit are decoupled and follow a simple
-harmonic oscillator. This approximation should be an excellent description of
-thin disk-like orbits and a good description of thick disk-like orbits. We
+Before turning to numerical methods, we first write down analytically the systematic error in the actions induced from either a midplane offset or a Galactocentric radius
+offset. We use the epicyclic approximation, which assumes that the motion in
+the $z$ and $R$ components of the orbit are decoupled and follow simple
+harmonic motion \citep[][section 3.2 and references therein]{2008gady.book.....B}. This approximation is an excellent description of
+thin disk-like orbits, and a good description of thick disk-like orbits, in an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and spiral arms). We
 also make the assumption of a perfectly flat circular velocity curve with
-velocity $v_c$.
+velocity $v_c$ \citep[decent near the Solar circle, e.g.][]{2017MNRAS.465...76M}.
 
@@
 vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
-constants. Similarly, the velocities of the orbit are given by,
+constants. Similarly, the velocities of the orbit are given by
 \beq\label{eq:orbits_vel_epi}
@@
 We can write down the actions in this approximation as
-\citep[Section~3.5.3b][]{2008gady.book.....B},
+\citep[][Section~3.5.3b]{2008gady.book.....B},
 \beq\label{eq:actions_epi}
@@
 
-An induced error in $z$ of $\Delta z$ will cause $\Delta A_z \sim \Delta z$.
-Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$.
-Making the assumption that $\kappa$ and $\nu$ are constant, we have that
+An offset $\Delta z$ will induce $\Delta A_z \sim \Delta z$.
+Similarly, an offset $\Delta R$ will induce $\Delta A_R \sim \Delta R$.
+Assuming $\kappa$ and $\nu$ are constant, 
 \beq\label{eq:actions_epi_error}
@@
 \eeq
-Therefore, errors in each action are linearly proportional to the errors in
-either the midplane or cylindrical radius. While in detail $\kappa$ and $\nu$
-are not constant, we will find that this order of magnitude estimate matches
-the numerical estimates quite well.
-
-There is a slight complication in the formula for computing $\Delta J_R/J_R$
-and $\Delta J_{\phi}/J_{\phi}$. In a moment we will consider the effect of an inaccuracy
-in the distance from the Sun to the Galactic center on actions. However, this
-inaccuracy induces an error in $x$, or $\Delta x$. This is not the same as an
-inaccuracy in $\Delta R$, but they are related by the following argument. An
-error $\Delta x$ will induce an erroneous radius $R_{\text{err}}$ related by
+Therefore, systematic errors in each action are linearly proportional to offsets in
+either the midplane ($z$ coordinate) or cylindrical radius ($R$ coordinate).
+While in general $\kappa$ and $\nu$ are not constant, we find that this order of magnitude estimate matches our numerical calculations quite well.
+
+%There is a slight complication in the formula for computing $\Delta J_R/J_R$
+%and $\Delta J_{\phi}/J_{\phi}$. 
+We will later consider the effect on actions of an inaccuracy in the distance from the Sun to the Galactic center, which introduces an offset in the $x$ coordinate, $\Delta x$, of each star when converting to a Galactocentric coordinate system.
+An offset $\Delta x$ results in an erroneous radius $R_{\text{err}}$ related by
 the formula,
@@
 
-A similarly rough argument can be made to arrive at the following formulae for
-the impact of velocity offsets on the actions,
+A similar derivation results in formulae for
+the impact of velocity offsets on the actions:
 \beq\label{eq:actions_epi_error}

### 2019-04-24 paper/ms.tex
a few more small revisions to section 2 in the new analytic derivation. more to come

@@
 
-We will check the formulae for the midplane offset against numerical tests
-next. The effect of velocity offsets on actions is deferred to future work, as
-we discuss in Section~\ref{ssec:lsr_var}.
+In the remainder of this section, we compare our analytic estimates of the effect on actions of a midplane offset against numerical calculations.
+The effect of velocity offsets on actions is deferred to future work, as we discuss in Section~\ref{ssec:lsr_var}.
 
 \subsection{Numerical Methods} \label{ssec:action_comp}
-We now move on to quantifying the argument made in Section~\ref{ssec:cartoon}.
+We now quantify the argument made in Section~\ref{ssec:cartoon} using numerical computations of the actions for a range of orbits in a model Galactic potential.
 We compute actions as in our previous work \citep{2018ApJ...867...31B}, using

### 2019-04-24 paper/ms.tex
small edits and rephrasings to section 3, and a few reference additions/changes

@@
 would generate the action errors described in Section~\ref{sec:ref_frame}:
-errors in the determination of the Galactocentric parameters. Converting from
+uncertainties in measurements of the position and velocity of the Sun relative to the Galactic center. Converting from
 heliocentric to Galactocentric coordinates relies on these measurements.
@@
 \citep[e.g.][]{2004ApJ...616..872R}. From stellar motions near
-Sgr~A\textsuperscript{*}, the distance from the Sun to the Galactic center can
-be measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a recent
+Sgr~A\textsuperscript{*}, the distance from the Sun to Sgr~A\textsuperscript{*} can
+be precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a recent
 measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
@@
 
-We take the ``dynamical Galactic center'' to be the point in three-dimensional
-space about which the stars in the solar neighborhood are orbiting. An
-important point to make about the Galactic center is that while many authors
-assume Sgr~A\textsuperscript{*} is colocated with the dynamical Galactic
-center, this assumption has not yet been justified. One potential test is to
-perform some sort of dynamical modelling where the distance from the Sun to
-the Galactic center is a free parameter and can be constrained with real data.
-For example, \citet{2015ApJ...803...80K} were able to make this measurement by
-modelling the dynamics of the stream Palomar 5. Other dynamical measurements
-of the Solar distance have been made, though none with a precision capable of
-competing with the distance to Sgr~A\textsuperscript{*}
-\citep{1981gask.book.....M, 2011PASJ...63..867S, 2012MNRAS.427..274S,
-2013AstL...39...95B, 2013IAUS..289..444Z}.
+Importantly, the location of Sgr~A\textsuperscript{*} may not be equivalent to the location of the ``dynamical Galactic center:'' the point in three-dimensional
+space about which the stars in the solar neighborhood are orbiting.
+This assumption, although frequently made, has not yet been justified.
+It is tested in any construction of a  dynamical model where the distance from the Sun to
+the Galactic center, $R_0$, is a free parameter.
+For example, \citet{2015ApJ...803...80K} measured $R_0$ while
+modelling the dynamics of the stream Palomar 5. Many other dynamical measurements
+of $R_0$ have been made (\citealt{2016ARAA..54..529B} summarize many pre-{\em Gaia} results) but none have yet achieved a precision comparable to that of the distance to Sgr~A\textsuperscript{*}.
+% \citep{1981gask.book.....M, 2011PASJ...63..867S, 2012MNRAS.427..274S,2013AstL...39...95B, 2013IAUS..289..444Z}.
 
 \subsection{Galactic Orientation}
-Second, one must define the orientation of the Galaxy about the center. This
+Second, one must define the angular orientation of the Galaxy. This
 was defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
 defining the coordinates of the Galactic center in B1950 coordinates as
-(17:42:26.6, -28:55:00) and the north Galactic pole as (12:49:00, +27:24:00).
-However, there is growing evidence that the true midplane is tilted when
-considering stars \citep{2014ApJ...797...53G, 2016ARAA..54..529B}, though not
-with H~\textsc{ii} \citep{2019ApJ...871..145A}.
+(17:42:26.6, -28:55:00) and the north Galactic pole as (12:49:00, +27:24:00). These two quantities, together with $R_0$, define the orientation of the Galactic plane.
+However, there is growing evidence that the stellar midplane is tilted relative to this coordinate system \citep{2014ApJ...797...53G, 2016ARAA..54..529B}, though not
+the H~\textsc{ii} midplane \citep{2019ApJ...871..145A}.
 
 \subsection{Solar Height}
-Thirdly, one must define the Sun's vertical distance from the Galactic
+Third, one must define the Sun's vertical distance from the Galactic
 midplane, which can be determined by identifying where the stellar density
@@
 \subsection{Local Standard of Rest}
-Finally, one must define the local standard of rest (LSR). The radial
-($U_{\odot}$) and vertical ($W_{\odot}$)components are easily computed by
-taking the mean motions of different stellar groups. The azimuthal component
+Finally, one must define the local standard of rest (LSR), or mean velocity of stars near the Sun relative to the Galactic center (which is defined to have zero velocity), and the velocity of the Sun relative to the LSR.
+The radial
+($U_{\odot}$) and vertical ($W_{\odot}$) components are computed by
+taking the mean motions of different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal component
 ($V_{\odot}$) is more difficult to measure, but can be modelled using the
@@
 
-In order to convert a star's position and velocity from heliocentric to
-Galactocentric coordinates, a number of parameters must be measured. Given the
-great interest in each parameter, considerable effort has been placed on these
-measurements. However, uncertainties remain, and detailed dynamical modelling
+\vspace{6pt}
+
+Considerable effort has been placed on each of these
+measurements, but uncertainties remain, and detailed dynamical modelling
 across the disk --- particularly for dynamically cold stars --- may have to
@@
 We now consider a second source of systematic offsets in the Galactocentric
-coordinate system --- intrinsic variations of the midplane across the disk. We
-study this in two sets of simulations of Milky Way-mass galaxies. We will also
+coordinate system: intrinsic variations of the midplane across the disk. We
+use two sets of simulations of Milky Way-mass galaxies to estimate the size of this effect. We will also
 briefly discuss a rough estimate of the midplane variation derived from the
@@
 
-\subsection{Description of Sagittarius Simulaton} \label{ssec:sag_sim}
+\subsection{Description of MW-Sagittarius Interaction Simulation} \label{ssec:sag_sim}
 In addition to the cosmological zoom-in, we will also briefly consider results
@@
 briefly summarize the most relevant details here. For the Milky Way, the dark
-halo is modelled as a Hernquist sphere of mass $10^{12}\,\Msun$ and scale
-length of $52\,\kpc$, the disk modelled as an exponential disk with a scale
+halo is modeled as a Hernquist sphere of mass $10^{12}\,\Msun$ and scale
+length of $52\,\kpc$, the disk modeled as an exponential disk with a scale
 radius of $3.5\,\kpc$, scale height $0.53\,\kpc$, and mass
@@
 dark matter Hernquist sphere of mass \Gus{something} and scale length
-\Gus{something}, along with a stellar component modelled as a Hernquist sphere
+\Gus{something}, along with a stellar component modeled as a Hernquist sphere
 of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$.
+All components are realized with live distributions of N-body particles; the MW and Sgr are each initialized to be in equilibrium in isolation \Gus{check!}.
 

### 2019-04-25 paper/ms.tex
Update author list, acknowledgements
Update Table 1 w more info about actions

@@
 
+\author{Melissa K. Ness}
+\affil{\cca}
+\affil{\columbia}
+
+\author{Kathryn V. Johnston}
+\affil{\columbia}
+
 \author{Mordecai-Mark Mac Low}
@@
 
-\author{Melissa K. Ness}
-\affil{\cca}
-\affil{\columbia}
-
 \author{Daniel Angl\'es-Alc\'azar}
@@
 
-\author{others}
+\author{Douglas Filho}
+\affil{\columbia}
+
+\author{David W. Hogg}
+\affil{\nyuphys}
+\affil{\nyucds}
+\affil{\cca}
+\affil{\mpia}
+
+\author{Chervin F. P. Laporte}
+\altaffiliation{Simons Fellow}
+\affil{\victoria}
 
@@
 $250\,\pc$ of the Sun, using the relevant initial positions and velocities
-from \citet{2018AA...616A..10G}. The neighborhood column is explained in more
+from \citet{2018AA...616A..10G}.
+Actions are computed using the method outlined in Section~\ref{ssec:action_comp}, with typical values for the Galactocentric coordinate system parameters.
+These are only shown to illustrate the general types of orbits open clusters are on.
+The neighborhood column is explained in more
 detail in Section~\ref{sssec:reconstruction}.\label{tab:real_clusters}}
@@
 \startdata
-alphaPer & 740 & 174.9 & \sim20 & \sim -1700 & \sim0.005 & \sim11 & 352\tablenotemark{b} & 57 \\
-Blanco1 & 489 & 237.2 & \sim2 & \sim-1800 & \sim2 & \sim210 & 410\tablenotemark{c} & 1460 \\
-ComaBer & 153 & 85.9 & \sim2 & \sim-1900 & \sim0.7 & \sim140 & 112\tablenotemark{d} & 522\\
-Hyades & 515 & 47.5 & \sim20 & \sim-1800 & \sim0.3 & \sim80 & 400\tablenotemark{e} & 462 \\
-IC2391 & 325 & 151.6 & \sim6 & \sim-1800 & \sim0.02 & \sim20 & \nodata & \nodata \\
-IC2602 & 492 & 152.2 & \sim10 & \sim-1700 & \sim0.3 & \sim89 & \nodata & \nodata \\
-NGC2451A\tablenotemark{a} & 400 & 193.7 & \sim6 & \sim-1800 & \sim0.2 & \sim73 & \nodata & \nodata \\
-Pleiades & 1326 & 135.8 & \sim20 & \sim-1700 & \sim0.4 & \sim96 & 800\tablenotemark{f} & 685 \\
-Praesepe & 938 & 186.2 & \sim20 & \sim-1800 & \sim0.6 & \sim120 & 550\tablenotemark{d} & 781
+alphaPer & 740 & 174.9 & 20 & -1700 & 0.005 & 11 & 352\tablenotemark{b} & 57 \\
+Blanco1 & 489 & 237.2 & 2 & -1800 & 2 & 210 & 410\tablenotemark{c} & 1460 \\
+ComaBer & 153 & 85.9 & 2 & -1900 & 0.7 & 140 & 112\tablenotemark{d} & 522\\
+Hyades & 515 & 47.5 & 20 & -1800 & 0.3 & 80 & 400\tablenotemark{e} & 462 \\
+IC2391 & 325 & 151.6 & 6 & -1800 & 0.02 & 20 & \nodata & \nodata \\
+IC2602 & 492 & 152.2 & 10 & -1700 & 0.3 & 89 & \nodata & \nodata \\
+NGC2451A\tablenotemark{a} & 400 & 193.7 & 6 & -1800 & 0.2 & 73 & \nodata & \nodata \\
+Pleiades & 1326 & 135.8 & 20 & -1700 & 0.4 & 96 & 800\tablenotemark{f} & 685 \\
+Praesepe & 938 & 186.2 & 20 & -1800 & 0.6 & 120 & 550\tablenotemark{d} & 781
 \enddata
@@
 Roy \& Diana Vagelos Program in the Molecular Life Sciences and the Roy \&
-Diana Vagelos Challenge Award.
-%mm The work of \Gus{everyone} is supported by the Simons
-%Foundation. \Gus{remove last sentence if everyone CCA affiliated}
-M.-M.M.L. was partly supported by NSF grant AST18-15461.
+Diana Vagelos Challenge Award. M.-M.M.L. was partly supported by NSF grant
+AST18-15461. The work of D.A.-A., A.B., D.W.H., C.F.P.L., R.S., K.V.J.,
+M.-M.M.L., and M.K.N. is supported by the Simons Foundation.
 

### 2019-04-25 paper/ms.tex
Add circular guiding orbit earlier

@@
 the $z$ and $R$ components of the orbit are decoupled and follow simple
-harmonic motion \citep[][section 3.2 and references therein]{2008gady.book.....B}. This approximation is an excellent description of
-thin disk-like orbits, and a good description of thick disk-like orbits, in an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and spiral arms). We
+harmonic motion about a circular and planar ``guiding orbit'' \citep[][Section 3.2 and references therein]{2008gady.book.....B}.
+We refer to the radius of this orbit (the ``guiding radius'') as $R_g$.
+This approximation is an excellent description of
+thin disk-like orbits and a good description of thick disk-like orbits in an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and spiral arms). We
 also make the assumption of a perfectly flat circular velocity curve with
@@
 J_R &= \kappa A_R^2 \\
-J_z &= \nu A_z^2\text{,}
+J_z &= \nu A_z^2\text{.}
 \end{split}
 \eeq
-where $R_g$ is the guiding radius of the orbit.
 

### 2019-04-25 paper/ms.tex
Move deltax to appendix, add discussion of epi

@@
 %and $\Delta J_{\phi}/J_{\phi}$.
-We will later consider the effect on actions of an inaccuracy in the distance from the Sun to the Galactic center, which introduces an offset in the $x$ coordinate, $\Delta x$, of each star when converting to a Galactocentric coordinate system.
-An offset $\Delta x$ results in an erroneous radius $R_{\text{err}}$ related by
-the formula,
-\beq
-(x+\Delta x)^2 + y^2 = R_{\text{err}}^2\text{.}
-\eeq
-Keeping only terms to first order in $\Delta x$, we have that,
-\beq
-\begin{split}
-R_{\text{err}}^2 &= R^2 - 2 R \cos{\theta} \Delta x \\
-\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\theta}} \Delta x\text{.}
-\end{split}
-\eeq
-Averaging over the circle, we therefore have that,
-\beq
-\avg{\Delta R} = \frac{2}{\pi} \Delta x\text{.}
-\eeq
-Therefore, the error in $J_R$ and $J_{\phi}$ are related to $\Delta x$ by,
-\beq\label{eq:Jr_epi_dx}
-\begin{split}
-\frac{\Delta J_R}{J_R} &= \frac{4 \Delta x}{\pi A_R} \\
-\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{2 \Delta x}{\pi R}\text{.}
-\end{split}
-\eeq
 
@@
 all three lines nearly overlap.
+
+In each panel of Figure~\ref{fig:many_orbit_wrong_ref}, where relevant, we
+include the estimation of the action errors derived under the epicyclic
+approximation in Section~\ref{ssec:epi_action}. Note that we consider an error
+in the $x$-coordinate $\Delta x$, which is not quite the same as $\Delta R$.
+For observations of stars close to us, we have that $\Delta x \sim \Delta R$,
+but for the experiment performed in this Section we consider observations of
+the star throughout its entire orbit. This introduces a factor of $2/\pi$ when
+converting from $\Delta R$ to $\Delta x$, which we derive in
+Appendix~\ref{app:deltax}.
+
+The epicyclic approximation is a good predictor of $J_z$, even for halo
+orbits. It performs similarly for $J_R$, now underpredicting for halo orbits
+and slightly overpredicting for thin-disk orbits. Note that for the particular
+orbits we chose, the thin-disk orbit has slightly larger $A_R$ than the
+thick-disk orbit, and so we expect the epicyclic approximation to perform
+slightly worse. The epicyclic approximation underpredicts $J_{\phi}$ for all
+orbits.
 
@@
 \appendix
+\section{$\Delta R$-$\Delta x$ Relation} \label{app:deltax}
+We considered the effect on actions of an inaccuracy in the distance from the
+Sun to the Galactic center, which introduces an offset in the $x$ coordinate,
+$\Delta x$, of each star when converting to a Galactocentric coordinate
+system. In observations of nearby stars, we have that $\Delta x \sim \Delta
+R$. However, for the experiment we performed in Section~\ref{ssec:quant} we
+considered observations of a star throughout its entire orbit. Therefore, we
+must average $\Delta R$ over the course of the orbit. We derive this relation
+now.
+
+An offset $\Delta x$ results in an erroneous radius $R_{\text{err}}$ related by
+the formula,
+\beq
+(x+\Delta x)^2 + y^2 = R_{\text{err}}^2\text{.}
+\eeq
+Keeping only terms to first order in $\Delta x$, we have that,
+\beq
+\begin{split}
+R_{\text{err}}^2 &= R^2 - 2 R \cos{\theta} \Delta x \\
+\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\theta}} \Delta x\text{.}
+\end{split}
+\eeq
+Averaging over the circle, we therefore have that,
+\beq
+\avg{\Delta R} = \frac{2}{\pi} \Delta x\text{.}
+\eeq
+
 \section{Orbits} \label{app:orbits}

### 2019-04-25 paper/ms.tex
Rehaul analytic section

@@
 
-Under this approximation, we can write down the vertical and radial components
+Under this approximation, we can write down the cylindrical components
 of the orbits as,
@@
 \begin{split}
-z(t) &= A_z \sin{(\kappa t + \alpha)}\\
-R(t) &= A_R \sin{(\nu t + \beta)}\text{,}
+\phi(t) &= \Omega t\\
+R(t) &= R_g + A_R \sin{(\nu t + \beta)} \\
+z(t) &= A_z \sin{(\kappa t + \alpha)}
+\text{,}
 \end{split}
@@
 \begin{split}
-v_z(t) &= A_z \kappa \cos{(\kappa t + \alpha)}\\
-v_R(t) &= A_R \nu \cos{(\nu t + \beta)}\text{.}
+v_{\phi}(t) &= \Omega \\
+v_R(t) &= A_R \nu \cos{(\nu t + \beta)} \\
+v_z(t) &= A_z \kappa \cos{(\kappa t + \alpha)}
+\text{.}
 \end{split}
@@
 
-We can write down the actions in this approximation as
-\citep[][Section~3.5.3b]{2008gady.book.....B},
-\beq\label{eq:actions_epi}
+In this case, we have that for $J_{\phi}$ \citep[][Section~3.5.3b]{2008gady.book.....B}, 
+\beq\label{eq:Jphi_epi}
+J_{\phi} = R_g v_c\text{,}
+\eeq
+and for $J_R$ and $J_z$,
+\beq\label{eq:JR_Jz_epi}
 \begin{split}
-J_{\phi} &= R_g v_c \\
-J_R &= \kappa A_R^2 \\
-J_z &= \nu A_z^2\text{.}
+J_R &= \frac{E_R}{\kappa} \\
+J_z &= \frac{E_z}{\nu} \text{,}
 \end{split}
 \eeq
-
-An offset $\Delta z$ will induce $\Delta A_z \sim \Delta z$.
-Similarly, an offset $\Delta R$ will induce $\Delta A_R \sim \Delta R$.
-Assuming $\kappa$ and $\nu$ are constant, 
-\beq\label{eq:actions_epi_error}
+where $E_R$ and $E_z$ are the energy per unit mass in the radial and vertical
+coordinates, respectively. Since $a_R = -\kappa^2(R-R_g)$, and $a_z = - \nu^2
+z$, we have that,
+\beq\label{eq:JR_Jz_epi_energy}
 \begin{split}
-\Delta J_{\phi} = \Delta R v_c &\implies \frac{\Delta J_{\phi}}{J_{\phi}} = \frac{\Delta R}{R} \\
-\Delta J_R = 2 \kappa \Delta A_R &\implies \frac{\Delta J_R}{J_R} = \frac{2\Delta R}{A_R} \\
-\Delta J_z = 2 \nu \Delta A_z &\implies \frac{\Delta J_z}{J_z} = \frac{2\Delta z}{A_z}\text{.}
+J_R &= \frac{\frac{1}{2}v_R^2 + \kappa^2 (R-R_g)^2}{\kappa} \\
+J_z &= \frac{\frac{1}{2}v_z^2 + \nu^2 z^2}{\nu}
 \end{split}
 \eeq
-Therefore, systematic errors in each action are linearly proportional to offsets in
-either the midplane ($z$ coordinate) or cylindrical radius ($R$ coordinate).
-While in general $\kappa$ and $\nu$ are not constant, we find that this order of magnitude estimate matches our numerical calculations quite well.
-
-%There is a slight complication in the formula for computing $\Delta J_R/J_R$
-%and $\Delta J_{\phi}/J_{\phi}$.
-
-A similar derivation results in formulae for
-the impact of velocity offsets on the actions:
-\beq\label{eq:actions_epi_error}
+
+While the formula for $J_{\phi}$ is phase independent, $J_R$ and $J_z$ are
+both phase dependent. We consider two cases. First, at the point of maximum
+amplitude or when $(v_R, v_z) = (0, 0)$ and second, at the point of maximum
+velocity $(R, z) = (R_g, 0)$. For $J_R$ we have that,
+\beq\label{eq:JR_epi_final}
+    \begin{aligned}
+        J_R &= \kappa A_R^2 && \text{if $v_R = 0$} \\
+    J_R &= \frac{v_{R,\text{max}}^2}{2\kappa^2} && \text{if $R=R_g$} \text{.}
+    \end{aligned}
+\eeq
+For $J_z$, we have that,
+\beq\label{eq:Jz_epi_final}
+    \begin{aligned}
+        J_z &= \kappa A_z^2 && \text{if $v_z = 0$} \\
+        J_z &= \frac{v_{z,\text{max}}^2}{2\kappa^2} && \text{if $z=0$} \text{.}
+    \end{aligned}
+\eeq
+
+Now assume that the coordinates $(R, z, v_{\phi}, v_R, v_z)$ are offset by
+$(\Delta R, \Delta z, \Delta v_{\phi}, \Delta v_R, \Delta v_z)$. Then, the
+induced error in $J_{\phi}$ is,
+\beq\label{eq:induced_Jphi}
+\frac{\Delta J_{\phi}}{J_{\phi}} = \frac{\Delta R}{R_g}\text{.}
+\eeq
+For $J_R$, the induced error is,
+\beq\label{eq:induced_JR}
+\begin{aligned}
+    \frac{\Delta J_R}{J_R} &= \frac{2\Delta R}{A_R} + 
+        \left(\frac{\Delta v_R}{v_{R,\text{max}}}\right)^2 && \text{if $v_R = 0$} \\
+    \frac{\Delta J_R}{J_R} &= \frac{2\Delta v_R}{v_{R,\text{max}}} + 
+        \left(\frac{\Delta R}{A_R}\right)^2  && \text{if $R=R_g$} \text{.}
+\end{aligned}
+\eeq
+For $J_{z}$, the induced error is,
+\beq\label{eq:induced_Jz}
+\begin{aligned}
+    \frac{\Delta J_z}{J_z} &= \frac{2\Delta z}{A_z} + 
+        \left(\frac{\Delta v_z}{v_{z,\text{max}}}\right)^2 && \text{if $v_z = 0$} \\
+    \frac{\Delta J_z}{J_z} &= \frac{2\Delta v_z}{v_{z,\text{max}}} + 
+        \left(\frac{\Delta z}{A_z}\right)^2  && \text{if $z=0$} \text{.}
+\end{aligned}
+\eeq
+
+Since most of the time stars will be at maximum amplitude in both the $R$ and
+$z$ components, we can assume for most stars that
+\beq\label{eq:Ji_err_mosttime}
 \begin{split}
-\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta v_{\phi}}{v_c} \\
-\frac{\Delta J_R}{J_R} &= \frac{\Delta v_R}{v_{R,\text{max}}} \\
-\frac{\Delta J_z}{J_z} &= \frac{\Delta v_z}{v_{z,\text{max}}}\text{,}
+\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta R}{R_g} \\
+\frac{\Delta J_{R}}{J_{R}} &= \frac{2\Delta R}{A_R} \\
+\frac{\Delta J_{z}}{J_{z}} &= \frac{2\Delta z}{A_z} \text{,}
 \end{split}
 \eeq
-where $v_{R,\text{max}}$ and $v_{z,\text{max}}$ are the maximum radial and
-vertical velocities of the orbit, respectively.
+where we have also ignored second order terms.
 

### 2019-04-25 paper/ms.tex
Add justification for spec. eq

@@
 In the remainder of this section, we compare our analytic estimates of the effect on actions of a midplane offset against numerical calculations.
-The effect of velocity offsets on actions is deferred to future work, as we discuss in Section~\ref{ssec:lsr_var}.
+Detailed worked on the effect of velocity offsets on actions is deferred to future work, as we discuss in Section~\ref{ssec:lsr_var}.
 
@@
 include the estimation of the action errors derived under the epicyclic
-approximation in Section~\ref{ssec:epi_action}. Note that we consider an error
+approximation from Equation~\ref{eq:Ji_err_mosttime}.
+This Equation is relevant since during most of the orbit the star will be at maximum radial and vertical amplitude and we do not consider velocity offsets.
+Note that we consider an error
 in the $x$-coordinate $\Delta x$, which is not quite the same as $\Delta R$.
@@
 \begin{split}
-R_{\text{err}}^2 &= R^2 - 2 R \cos{\theta} \Delta x \\
-\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\theta}} \Delta x\text{.}
+R_{\text{err}}^2 &= R^2 - 2 R \cos{\phi} \Delta x \\
+\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\phi}} \Delta x\text{.}
 \end{split}

### 2019-04-25 paper/ms.tex
Remove neighborhood column from table

@@
 \tablehead{\colhead{cluster} & \colhead{$N_{\text{memb}}$} & \colhead{distance} & \colhead{$J_R$} &
-\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} & \colhead{neighborhood} \\ 
+\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} \\ 
 \colhead{ } & \colhead{ } &
@@
 \colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{pc}$} & \colhead{$M_{\odot}$} & \colhead{$\mathrm{pc}$}}
+\colhead{$\mathrm{pc}$} & \colhead{$M_{\odot}$}}
 \startdata
-alphaPer & 740 & 174.9 & 20 & -1700 & 0.005 & 11 & 352\tablenotemark{b} & 57 \\
-Blanco1 & 489 & 237.2 & 2 & -1800 & 2 & 210 & 410\tablenotemark{c} & 1460 \\
-ComaBer & 153 & 85.9 & 2 & -1900 & 0.7 & 140 & 112\tablenotemark{d} & 522\\
-Hyades & 515 & 47.5 & 20 & -1800 & 0.3 & 80 & 400\tablenotemark{e} & 462 \\
-IC2391 & 325 & 151.6 & 6 & -1800 & 0.02 & 20 & \nodata & \nodata \\
-IC2602 & 492 & 152.2 & 10 & -1700 & 0.3 & 89 & \nodata & \nodata \\
-NGC2451A\tablenotemark{a} & 400 & 193.7 & 6 & -1800 & 0.2 & 73 & \nodata & \nodata \\
-Pleiades & 1326 & 135.8 & 20 & -1700 & 0.4 & 96 & 800\tablenotemark{f} & 685 \\
-Praesepe & 938 & 186.2 & 20 & -1800 & 0.6 & 120 & 550\tablenotemark{d} & 781
+alphaPer & 740 & 170 & 20 & -1700 & 0.005 & 11 & 352\tablenotemark{b} \\
+Blanco1 & 489 & 240 & 2 & -1800 & 2 & 210 & 410\tablenotemark{c} \\
+ComaBer & 153 & 90 & 2 & -1900 & 0.7 & 140 & 112\tablenotemark{d} \\
+Hyades & 515 & 50 & 20 & -1800 & 0.3 & 80 & 400\tablenotemark{e} \\
+IC2391 & 325 & 150 & 6 & -1800 & 0.02 & 20 & \nodata \\
+IC2602 & 492 & 150 & 10 & -1700 & 0.3 & 89 & \nodata \\
+NGC2451A\tablenotemark{a} & 400 & 190 & 6 & -1800 & 0.2 & 73 & \nodata \\
+Pleiades & 1326 & 140 & 20 & -1700 & 0.4 & 96 & 800\tablenotemark{f} \\
+Praesepe & 938 & 190 & 20 & -1800 & 0.6 & 120 & 550\tablenotemark{d}
 \enddata

### 2019-04-26 paper/ms.tex
Change Omega to vc
cosmetic changes to intro and fix to typo in Eq 2

@@
 of the Galactic plane, the height of the Sun above the Galactic midplane, and
-the local standard of rest (LSR). We discuss the observational efforts of all
-of these parameters in Section~\ref{sec:mes_err}.
+the local standard of rest (LSR). We review the observational efforts to measure these parameters in Section~\ref{sec:mes_err}.
 
@@
 \citep{2016ApJ...827L..23W, 2017arXiv170206148H}, and a controlled N-body
-simulation of a Sagittarius encounter with a Milky Way-mass galaxy
-\citep{2018MNRAS.481..286L}.
+simulation of a Sagittarius encounter with a galaxy otherwise tailored to the stellar mass, scale length, and scale height of the Milky Way \citep{2018MNRAS.481..286L}.
 
@@
 \begin{split}
-\phi(t) &= \Omega t\\
+\phi(t) &= \Omega_c t \\
 R(t) &= R_g + A_R \sin{(\nu t + \beta)} \\
@@
 where $\kappa$ and $\nu$ are the vertical and radial (or epicyclic)
-frequencies, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
+frequencies, $\Omega_c \equiv v_c/R_g$ the rotation frequency, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
 vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
@@
 \begin{split}
-v_{\phi}(t) &= \Omega \\
+v_{\phi}(t) &= v_c \\
 v_R(t) &= A_R \nu \cos{(\nu t + \beta)} \\

### 2019-04-26 paper/ms.tex
cosmetic changes to derivation in section 2 - see comments in slack for other needed checks

@@
 While the formula for $J_{\phi}$ is phase independent, $J_R$ and $J_z$ are
-both phase dependent. We consider two cases. First, at the point of maximum
+both phase dependent via Eq. \eqref{eq:orbits_epi}. We consider two cases. First, at the point of maximum
 amplitude or when $(v_R, v_z) = (0, 0)$ and second, at the point of maximum
@@
 $(\Delta R, \Delta z, \Delta v_{\phi}, \Delta v_R, \Delta v_z)$. Then, the
-induced error in $J_{\phi}$ is,
+induced error in $J_{\phi}$ is, regardless of phase,
 \beq\label{eq:induced_Jphi}
@@
 
-Since most of the time stars will be at maximum amplitude in both the $R$ and
-$z$ components, we can assume for most stars that
+Since most of the time stars will be at maximum amplitude (i.e. turnaround) in both $R$ and $z$, we can approximate the order of magnitude of the systematic error in the actions by
 \beq\label{eq:Ji_err_mosttime}
@@
 In the remainder of this section, we compare our analytic estimates of the effect on actions of a midplane offset against numerical calculations.
-Detailed worked on the effect of velocity offsets on actions is deferred to future work, as we discuss in Section~\ref{ssec:lsr_var}.
+A numerical evaluation of the effect of velocity offsets on actions is deferred to future work, as we discuss in Section~\ref{ssec:lsr_var}.
 

### 2019-04-26 paper/ms.tex
changes to conclusions wording

@@
 great the $z$ offset is. We also plot the epicylic prediction for $\Delta J_z
-/ J_z$ from Equation~\ref{eq:actions_epi_error} as dashed lines for each $z$
+/ J_z$ from Equation~\ref{eq:Ji_err_mosttime} as dashed lines for each $z$
 offset. We find that the epicyclic approximation matches the numerical
@@
 error if the cluster and field star are far enough apart to have significantly
-different local midplanes (see Figure \ref{fig:range_deltaphi}). Since our
-criterion for cluster membership (distance in action space) scales with the
+different local midplanes (see Figures \ref{fig:midplane} and \ref{fig:midplane_chervin}). Since our criterion for cluster membership (distance in action space) scales with the
 cluster mass, the region where the search will succeed varies with cluster
@@
 \begin{itemize}
-\item We demonstrated that an inaccuracy in the Galactocentric coordinate
-system induces time evolution in the {\em observed} actions
-(Figures~\ref{fig:cartoon}~and~\ref{fig:one_orbit_wrong_ref}). We interpret
-this time evolution as an error in the computed action
+\item Inaccuracy in the Galactocentric coordinate
+system induces phase-dependence in the actions calculated from the observed positions and velocities of stars
+(Figures~\ref{fig:cartoon}~and~\ref{fig:one_orbit_wrong_ref}).
+Since stars' instantaneous phase-space positions are measured without prior knowledge of their orbital phases, this results in systematic error in the computed actions
 (Figure~\ref{fig:many_orbit_wrong_ref}).
 
-\item  We focused on the effect that an
-inaccuracy in the midplane has on the vertical action $J_z$. We found that in
-order for there to be a $10\%$ error in $J_z$ (as defined by one half the middle
-$90\uth$ percentile) one needs a midplane error of $\sim6\,\pc$ and
-$\sim50\,\pc$ for a typical thin and thick disk orbit, respectively. The
-fractional error is significantly less for halo orbits.
-
-\item We pointed out that the error distribution of the actions induced by a
-coordinate system error is very non-Gaussian. The distribution is bimodal with
-{\em neither of the modes peaking at 
-     null}. 
-As a result, error
-propagation of systematic coordinate system errors is 
-     complex
-when considering actions.
-
-\item We proposed that dynamical modelling across large regions of the disk is
-susceptible to this type of error, since the assumption that our local
+\item Inaccuracy in the midplane location most severely affects computation of the vertical action $J_z$.
+A midplane offset of $\sim6\,\pc$ ($\sim50\,\pc$) for a typical thin (thick) disk orbit results in a  $10\%$ error in $J_z$ (as defined by one half the middle
+$90\uth$ percentile). The fractional error is significantly less for halo orbits.
+
+\item The distribution of systematic errors in the actions induced by a
+coordinate system offset is very non-Gaussian. The distribution is bimodal with
+{\em neither mode at null}. 
+As a result, error propagation of coordinate system offsets is complex when considering actions, and is likely to significantly deform the action-space distribution function.
+
+\item Dynamical modelling across large regions of the disk, over which the midplane location varies by more than the limits discussed above, is
+susceptible to this type of systematic error, since the assumption that our local
 Galactic midplane is the global Galactic midplane is not true {\em a priori}.
@@
 simulations of Milky Way mass galaxies from the FIRE collaboration. We found
-that the midplane varies by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
+that the midplane varies as a function of azimuth by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
 three galaxies we considered (\mi{}, \mf{}, and \mm{}, respectively).
@@
 
+\item While in this work we have focused on systematic errors in action computation, all of our conclusions also extend to studies of stars that simply rely on orbit
+integration, since the computation of actions and orbit integrations are
+mathematically equivalent.
+
 \end{itemize}
 
-While in this work we have focused on errors in action computation, all of our
-conclusions also extend to studies of stars that simply rely on orbit
-integration, since the computation of actions and orbit integrations are
-essentially equivalent.
-
-{\bf Add sentences about distribution function modeling}
-
 Our main point is that the local midplane varies between different points in
-the Galaxy. We do
+the Galaxy, and that this variation can lead to significant systematic errors in the computation of actions under the assumption of a global axisymmetric potential. We do
 not propose here what a modeler should use for the ``true'' global midplane,
@@
 azimuthal dependence of the midplane location.
+For some applications, such as those using actions as labels to group stars on similar orbits, using such a measurement to shift stars to a consistent midplane height as a function of azimuth before using a global axisymmetric approximation to the potential may be sufficient, although this ignores the \emph{dynamical} implications of shifts in the midplane height (which result from fluctuations in the local density).
+However, for other applications, such as the study of action diffusion which motivated this paper in the first place, a more extensive perturbative approach is likely to be needed.
+We plan to explore the mitigation of these effects in future work.
 

### 2019-04-26 paper/ms.tex
Fix equation refs throughout

@@
 While the formula for $J_{\phi}$ is phase independent, $J_R$ and $J_z$ are
-both phase dependent via Eq. \eqref{eq:orbits_epi}. We consider two cases. First, at the point of maximum
+both phase dependent via Equation~\eqref{eq:orbits_epi}. We consider two cases. First, at the point of maximum
 amplitude or when $(v_R, v_z) = (0, 0)$ and second, at the point of maximum
@@
 include the estimation of the action errors derived under the epicyclic
-approximation from Equation~\ref{eq:Ji_err_mosttime}.
+approximation from Equation~\eqref{eq:Ji_err_mosttime}.
 This Equation is relevant since during most of the orbit the star will be at maximum radial and vertical amplitude and we do not consider velocity offsets.
@@
 dashed lines the epicyclic prediction of the induced action error
-(Equations~\ref{eq:actions_epi_error}~and~\ref{eq:Jr_epi_dx}). In the
+(Equations~\eqref{eq:actions_epi_error}~and~\eqref{eq:Jr_epi_dx}). In the
 epicyclic approximation, a $z$ offset only induces an error in $J_z$ --- for
@@
 great the $z$ offset is. We also plot the epicylic prediction for $\Delta J_z
-/ J_z$ from Equation~\ref{eq:Ji_err_mosttime} as dashed lines for each $z$
+/ J_z$ from Equation~\eqref{eq:Ji_err_mosttime} as dashed lines for each $z$
 offset. We find that the epicyclic approximation matches the numerical
@@
 estimate the expected action-space dispersion from
-Equation~\ref{eq:action_disp}. We then consult
+Equation~\eqref{eq:action_disp}. We then consult
 Figure~\ref{fig:many_orbit_wrong_ref} to determine the midplane offset
@@
 cluster of mass $m_c$. We compute the intrinsic dispersion from
-Equation~\ref{eq:action_disp} and convert the dispersion to a $z$ offset using
+Equation~\eqref{eq:action_disp} and convert the dispersion to a $z$ offset using
 the result from Figure~\ref{fig:many_orbit_wrong_ref}. We consider the result

### 2019-04-26 paper/ms.tex
Small changes throughout

@@
 error if the cluster and field star are far enough apart to have significantly
-different local midplanes (see Figures \ref{fig:midplane} and \ref{fig:midplane_chervin}). Since our criterion for cluster membership (distance in action space) scales with the
+different local midplanes (see Figures~\ref{fig:midplane}~and~\ref{fig:midplane_chervin}). Since our criterion for cluster membership (distance in action space) scales with the
 cluster mass, the region where the search will succeed varies with cluster
@@
 simulations of Milky Way mass galaxies from the FIRE collaboration. We found
-that the midplane varies as a function of azimuth by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
+that the midplane varies as a function of azimuth at the Solar circle by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
 three galaxies we considered (\mi{}, \mf{}, and \mm{}, respectively).
@@
 \acknowledgments
-We would like to thank Megan Bedell, Tobias Buck, Elena D'Onghia, Benoit
+We would like to thank Megan Bedell, Robert A. Benjamin, Tobias Buck, Elena D'Onghia, Benoit
 Famaey, and Adrian Price-Whelan for helpful discussions. A.B. would like to

### 2019-04-26 paper/ms.tex
Remove ref to neighborhood column

@@
 Actions are computed using the method outlined in Section~\ref{ssec:action_comp}, with typical values for the Galactocentric coordinate system parameters.
-These are only shown to illustrate the general types of orbits open clusters are on.
-The neighborhood column is explained in more
-detail in Section~\ref{sssec:reconstruction}.\label{tab:real_clusters}}
+These are only shown to illustrate the general types of orbits open clusters are on.\label{tab:real_clusters}}
 \tablehead{\colhead{cluster} & \colhead{$N_{\text{memb}}$} & \colhead{distance} & \colhead{$J_R$} &

### 2019-04-26 paper/ms.tex
Fix epicyclic section

@@
 \phi(t) &= \Omega_c t \\
-R(t) &= R_g + A_R \sin{(\nu t + \beta)} \\
-z(t) &= A_z \sin{(\kappa t + \alpha)}
+R(t) &= R_g + A_R \sin{(\kappa t + \alpha)} \\
+z(t) &= A_z \sin{(\nu t + \beta)}
 \text{,}
@@
 \eeq
-where $\kappa$ and $\nu$ are the vertical and radial (or epicyclic)
+where $\kappa$ and $\nu$ are the radial and vertical
 frequencies, $\Omega_c \equiv v_c/R_g$ the rotation frequency, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
@@
 v_{\phi}(t) &= v_c \\
-v_R(t) &= A_R \nu \cos{(\nu t + \beta)} \\
-v_z(t) &= A_z \kappa \cos{(\kappa t + \alpha)}
+v_R(t) &= \kappa A_R \cos{(\kappa t + \alpha)} \\
+v_z(t) &= \nu A_z \cos{(\nu t + \beta)}
 \text{.}
@@
 where $E_R$ and $E_z$ are the energy per unit mass in the radial and vertical
-coordinates, respectively. Since $a_R = -\kappa^2(R-R_g)$, and $a_z = - \nu^2
-z$, we have that,
+coordinates, respectively. Therefore,
 \beq\label{eq:JR_Jz_epi_energy}
 \begin{split}
-J_R &= \frac{\frac{1}{2}v_R^2 + \kappa^2 (R-R_g)^2}{\kappa} \\
-J_z &= \frac{\frac{1}{2}v_z^2 + \nu^2 z^2}{\nu}
+J_R &= \frac{v_R^2 + \kappa^2 (R-R_g)^2}{2\kappa} \\
+J_z &= \frac{v_z^2 + \nu^2 z^2}{2\nu}\text{.}
 \end{split}
 \eeq
-
-While the formula for $J_{\phi}$ is phase independent, $J_R$ and $J_z$ are
-both phase dependent via Equation~\eqref{eq:orbits_epi}. We consider two cases. First, at the point of maximum
-amplitude or when $(v_R, v_z) = (0, 0)$ and second, at the point of maximum
-velocity $(R, z) = (R_g, 0)$. For $J_R$ we have that,
-\beq\label{eq:JR_epi_final}
-    \begin{aligned}
-        J_R &= \kappa A_R^2 && \text{if $v_R = 0$} \\
-    J_R &= \frac{v_{R,\text{max}}^2}{2\kappa^2} && \text{if $R=R_g$} \text{.}
-    \end{aligned}
-\eeq
-For $J_z$, we have that,
-\beq\label{eq:Jz_epi_final}
-    \begin{aligned}
-        J_z &= \kappa A_z^2 && \text{if $v_z = 0$} \\
-        J_z &= \frac{v_{z,\text{max}}^2}{2\kappa^2} && \text{if $z=0$} \text{.}
-    \end{aligned}
-\eeq
-
-Now assume that the coordinates $(R, z, v_{\phi}, v_R, v_z)$ are offset by
-$(\Delta R, \Delta z, \Delta v_{\phi}, \Delta v_R, \Delta v_z)$. Then, the
-induced error in $J_{\phi}$ is, regardless of phase,
+We can simplify this and write,
+\beq\label{eq:JR_Jz_epi_final}
+\begin{split}
+J_R &= \frac{\kappa A_R^2}{2} = \frac{v_{R,\text{max}}^2}{2\kappa} \\
+J_z &= \frac{\nu A_z^2}{2} = \frac{v_{z,\text{max}}^2}{2\nu}\text{,}
+\end{split}
+\eeq
+where the last equality comes from the fact that $v_{R,\text{max}} = \kappa
+A_R$ and similarly for $z$.
+
+Notice that while the value for each of $J_{\phi}$, $J_R$, and $J_z$ is phase
+independent, the contribution from the kinetic and potential terms in
+Equation~\eqref{eq:JR_Jz_epi_energy} is phase dependent. Now assume that the
+coordinates $(R, z, v_{\phi}, v_R, v_z)$ are offset by $(\Delta R, \Delta z,
+\Delta v_{\phi}, \Delta v_R, \Delta v_z)$. We can then apply the standard
+propagation of errors formula to Equation~\eqref{eq:JR_Jz_epi_energy}
+determine the error in each of the actions. For $J_{\phi}$ we have that,
 \beq\label{eq:induced_Jphi}
-\frac{\Delta J_{\phi}}{J_{\phi}} = \frac{\Delta R}{R_g}\text{.}
+\frac{\Delta J_{\phi}}{J_{\phi}} = \frac{\Delta R}{R_g} + \frac{\Delta v_{\phi}}{v_c}\text{.}
 \eeq
@@
 \beq\label{eq:induced_JR}
-\begin{aligned}
-    \frac{\Delta J_R}{J_R} &= \frac{2\Delta R}{A_R} + 
-        \left(\frac{\Delta v_R}{v_{R,\text{max}}}\right)^2 && \text{if $v_R = 0$} \\
-    \frac{\Delta J_R}{J_R} &= \frac{2\Delta v_R}{v_{R,\text{max}}} + 
-        \left(\frac{\Delta R}{A_R}\right)^2  && \text{if $R=R_g$} \text{.}
-\end{aligned}
-\eeq
-For $J_{z}$, the induced error is,
+\frac{\Delta J_R}{J_R} = \frac{2(R-R_g)}{A_R^2}\Delta R
+                         + \frac{2v_R}{v_{R,\text{max}}^2} \Delta v_R \text{.}
+\eeq
+For $J_z$, the induced error is,
 \beq\label{eq:induced_Jz}
-\begin{aligned}
-    \frac{\Delta J_z}{J_z} &= \frac{2\Delta z}{A_z} + 
-        \left(\frac{\Delta v_z}{v_{z,\text{max}}}\right)^2 && \text{if $v_z = 0$} \\
-    \frac{\Delta J_z}{J_z} &= \frac{2\Delta v_z}{v_{z,\text{max}}} + 
-        \left(\frac{\Delta z}{A_z}\right)^2  && \text{if $z=0$} \text{.}
-\end{aligned}
-\eeq
+\frac{\Delta J_z}{J_z} = \frac{2z}{A_z^2}\Delta z
+                         + \frac{2v_z}{v_{z,\text{max}}^2} \Delta v_z \text{.}
+\eeq
+We have ignored second order contributions.
 

### 2019-04-26 paper/ms.tex
Fix line indentation throughout

@@
 
-Stellar actions, computed for stars from their six-dimensional phase space measurements and
-an assumed Galactic potential, are used to label and distinguish orbits. In
-principle, actions have the attractive quality of being invariant and thus
-good labels of a star's orbit. 
-%mm Of course,
-    Such analyses are often used to interpret data from the {\em Gaia}
-    mission. However, 
-inaccurate measurements of the 
-phase space position and uncertainties in the assumed Galactic potential
-induce an error in the computed action. We show that, in addition to these
-complications, a systematic bias in the phase space position induces a
-phase-dependence in the actions, which we interpret as 
-%mm an 
-    a systematic
-error. The induced error distribution is
-non-Gaussian and bimodal, 
-with neither of the modes peaking on
-%mm  the true action value. [the error in a Gaussian peaks on zero] 
-    the null value.
-A variation in the vertical position midplane of $\sim15\,\pc$ for a thin disk orbit or
-$\sim120\,\pc$ for a thick disk orbit is 
-%mm all that is necessary 
-      sufficient
-to induce a $25\,\%$ systematic error in the vertical action $J_z$ computed under the assumption of an axisymmetric potential.
-%mm [fragment?] orbits, respectively. 
-Furthermore, we show that the local midplane varies by
+Stellar actions, computed for stars from their six-dimensional phase space
+measurements and an assumed Galactic potential, are used to label and
+distinguish orbits. In principle, actions have the attractive quality of being
+invariant and thus good labels of a star's orbit. Such analyses are often used
+to interpret data from the {\em Gaia} mission. However, inaccurate
+measurements of the phase space position and uncertainties in the assumed
+Galactic potential induce an error in the computed action. We show that, in
+addition to these complications, a systematic bias in the phase space position
+induces a phase-dependence in the actions, which we interpret as a systematic
+error. The induced error distribution is non-Gaussian and bimodal, with
+neither of the modes peaking on the null value. A variation in the vertical
+position midplane of $\sim15\,\pc$ for a thin disk orbit or $\sim120\,\pc$ for
+a thick disk orbit is sufficient to induce a $25\,\%$ systematic error in the
+vertical action $J_z$ computed under the assumption of an axisymmetric
+potential. Furthermore, we show that the local midplane varies by
 $\sim100\,\pc$ at the Solar circle in Milky Way-mass cosmological zoom-in
@@
 calculations --- which assume the global and local midplanes are the same ---
-include a systematic 
-%mm offset in $z$.
-   vertical offset variation. 
-Variation in the local standard of rest
-%mm will induce 
-     induces
-similar issues. The variation of the midplane must be taken into
-account when performing dynamical modelling across
-the large regions of the disk
-accessible 
-%mm with the {\em Gaia} mission. [moved first mention up to top of
-%abstract, given that this is in the title.]
-    to {\em Gaia}.
+include a systematic vertical offset variation. Variation in the local
+standard of rest induces similar issues. The variation of the midplane must be
+taken into account when performing dynamical modelling across the large
+regions of the disk accessible to {\em Gaia}.
 
@@
 of the Galactic plane, the height of the Sun above the Galactic midplane, and
-the local standard of rest (LSR). We review the observational efforts to measure these parameters in Section~\ref{sec:mes_err}.
+the local standard of rest (LSR). We review the observational efforts to
+measure these parameters in Section~\ref{sec:mes_err}.
 
@@
 and azimuthal coordinates respectively in a cylindrical coordinate system and
-the $p_i$ are the conjugate momenta. In a slowly-evolving axisymmetric potential, these
-actions are invariant and $J_{\phi} \equiv L_z$
+the $p_i$ are the conjugate momenta. In a slowly-evolving axisymmetric
+potential, these actions are invariant and $J_{\phi} \equiv L_z$
 \citep{2008gady.book.....B,2014RvMP...86....1S}.
@@
 midplane. For stars relatively far from the Sun, the local midplane location
-will not agree with our local midplane extrapolated onto their position,
-due to relatively small-scale variations in the density of the gas and
-stellar distribution. Converting positions and velocities of more distant
-stars from a heliocentric to a Galactocentric coordinate system introduces a
-systematic bias in the $z$ coordinate, and hence the actions computed for
-these stars. The further from the solar neighborhood the target star, the more
-likely the mismatch will result in large systematic uncertainty, especially in
-the vertical action $J_z$. Depending on the type of orbit being studied, this
-sets a limit on the ``dynamical solar neighborhood'' over which the local
+will not agree with our local midplane extrapolated onto their position, due
+to relatively small-scale variations in the density of the gas and stellar
+distribution. Converting positions and velocities of more distant stars from a
+heliocentric to a Galactocentric coordinate system introduces a systematic
+bias in the $z$ coordinate, and hence the actions computed for these stars.
+The further from the solar neighborhood the target star, the more likely the
+mismatch will result in large systematic uncertainty, especially in the
+vertical action $J_z$. Depending on the type of orbit being studied, this sets
+a limit on the ``dynamical solar neighborhood'' over which the local
 coordinate frame can be usefully extended for action-space analysis. A similar
 argument applies to any remaining uncertainty in measurements of the
-Galactocentric radius of the Sun and to variations in the local
-standard of rest (LSR).
+Galactocentric radius of the Sun and to variations in the local standard of
+rest (LSR).
 
@@
 \citep{2016ApJ...827L..23W, 2017arXiv170206148H}, and a controlled N-body
-simulation of a Sagittarius encounter with a galaxy otherwise tailored to the stellar mass, scale length, and scale height of the Milky Way \citep{2018MNRAS.481..286L}.
+simulation of a Sagittarius encounter with a galaxy otherwise tailored to the
+stellar mass, scale length, and scale height of the Milky Way
+\citep{2018MNRAS.481..286L}.
 
@@
 variations, and the resulting systematic uncertainty in the vertical action,
-for action-space analyses beyond the dynamical solar neighborhood. In light of our findings, we considered the
-orbits of the nine open clusters within $250\,\pc$ as reported by the {\em
-Gaia} collaboration \citep{2018AA...616A..10G}. In
-Table~\ref{tab:real_clusters}, we report the distance to each cluster, the
-three actions, and the maximum vertical extent of their orbits
-($z_{\text{max}}$). In Section~\ref{sssec:reconstruction} we estimate and
-discuss the size of the dynamical solar neighborhood for each cluster based on
-its mass, to show that action-space tagging is likely still a viable method if
-contamination from foreground or background stars is unimportant. We summarize
-and conclude in Section~\ref{sec:conclusion}.
+for action-space analyses beyond the dynamical solar neighborhood. In light of
+our findings, we considered the orbits of the nine open clusters within
+$250\,\pc$ as reported by the {\em Gaia} collaboration
+\citep{2018AA...616A..10G}. In Table~\ref{tab:real_clusters}, we report the
+distance to each cluster, the three actions, and the maximum vertical extent
+of their orbits ($z_{\text{max}}$). In Section~\ref{sssec:reconstruction} we
+estimate and discuss the size of the dynamical solar neighborhood for each
+cluster based on its mass, to show that action-space tagging is likely still a
+viable method if contamination from foreground or background stars is
+unimportant. We summarize and conclude in Section~\ref{sec:conclusion}.
 
@@
 $250\,\pc$ of the Sun, using the relevant initial positions and velocities
-from \citet{2018AA...616A..10G}.
-Actions are computed using the method outlined in Section~\ref{ssec:action_comp}, with typical values for the Galactocentric coordinate system parameters.
-These are only shown to illustrate the general types of orbits open clusters are on.\label{tab:real_clusters}}
-\tablehead{\colhead{cluster} & \colhead{$N_{\text{memb}}$} & \colhead{distance} & \colhead{$J_R$} &
-\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} \\ 
+from \citet{2018AA...616A..10G}. Actions are computed using the method
+outlined in Section~\ref{ssec:action_comp}, with typical values for the
+Galactocentric coordinate system parameters. These are only shown to
+illustrate the general types of orbits open clusters are
+on.\label{tab:real_clusters}}
+\tablehead{\colhead{cluster} & \colhead{$N_{\text{memb}}$} &
+\colhead{distance} & \colhead{$J_R$} &
+\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
+\colhead{mass} \\
 \colhead{ } & \colhead{ } &
@@
 \section{Motivation} \label{sec:ref_frame}
-We first demonstrate the significance  to
-action computations of a systematic offset in the
-determination of the Galactic midplane or distance to the Galactic center. We will find that such offsets are especially important
-for disk-like orbits. The consequences we explore here may also arise from
-various other systematic errors. For instance, the axisymmetric Galactic
-potential model used in many works to compute actions may not be a good
-description of the true potential --- or the parameters used may yield a
-potential that is systematically incorrect outside an original fitted region.
-In this work, we assume that the Galaxy is perfectly described by our model
-axisymmetric potential, and simply explore the consequences of offsets in the determination of the Galactocentric coordinate system.
+We first demonstrate the significance  to action computations of a systematic
+offset in the determination of the Galactic midplane or distance to the
+Galactic center. We will find that such offsets are especially important for
+disk-like orbits. The consequences we explore here may also arise from various
+other systematic errors. For instance, the axisymmetric Galactic potential
+model used in many works to compute actions may not be a good description of
+the true potential --- or the parameters used may yield a potential that is
+systematically incorrect outside an original fitted region. In this work, we
+assume that the Galaxy is perfectly described by our model axisymmetric
+potential, and simply explore the consequences of offsets in the determination
+of the Galactocentric coordinate system.
 
@@
 We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
-inaccurate or erroneous determination of the midplane (and hence the zero-point of the $z$ coordinate) leads to a phase-dependence of the
-actions calculated from an orbit specified by a phase-space point and an
-assumed potential model. The $y$-axis corresponds to the vertical height of
-the orbit and the $x$-axis shows orbital phase. The solid gray curve indicates the
+inaccurate or erroneous determination of the midplane (and hence the
+zero-point of the $z$ coordinate) leads to a phase-dependence of the actions
+calculated from an orbit specified by a phase-space point and an assumed
+potential model. The $y$-axis corresponds to the vertical height of the orbit
+and the $x$-axis shows orbital phase. The solid gray curve indicates the
 ``true'' orbit of the star; that is, its orbit as it oscillates around the
-true midplane. The dashed gray line is the midplane location in the
-potential model used to integrate the orbit of the star and
-hence estimate its actions, which is offset from the true midplane of the star's orbit. The model potential is otherwise identical to the one in which the star is actually moving.
-
-Now suppose this observer makes a measurement of the star's position and velocity at the purple
-point or the green point (i.e. at two different orbital phases). Then the
-purple and green curves correspond to the orbits that the observer would
-compute for each point based on the potential model with the offset
-midplane. In action space, this would correspond to a different value of $J_z$
-for the purple and green points. In this way, assuming the wrong coordinate
-system induces a phase-dependence in the actions estimated for the star, which
-in the correct potential (in this example, the one with the correct midplane)
-should be phase-independent.
+true midplane. The dashed gray line is the midplane location in the potential
+model used to integrate the orbit of the star and hence estimate its actions,
+which is offset from the true midplane of the star's orbit. The model
+potential is otherwise identical to the one in which the star is actually
+moving.
+
+Now suppose this observer makes a measurement of the star's position and
+velocity at the purple point or the green point (i.e. at two different orbital
+phases). Then the purple and green curves correspond to the orbits that the
+observer would compute for each point based on the potential model with the
+offset midplane. In action space, this would correspond to a different value
+of $J_z$ for the purple and green points. In this way, assuming the wrong
+coordinate system induces a phase-dependence in the actions estimated for the
+star, which in the correct potential (in this example, the one with the
+correct midplane) should be phase-independent.
 
@@
 \end{center}
-\caption{Cartoon approximation showing the effect an error in the determination of the
-coordinate midplane can have on orbit integration and action estimation. The
-$x$-axis shows the orbital phase and the $y$-axis the vertical height. The top
-gray curve depicts an example ``true'' orbit oscillating about the true
-midplane (horizontal solid gray line). Consider an observer who erroneously
-assumes the midplane is located at the horizontal dashed line. Suppose this
-observer measures the phase-space position of the star at two different
-orbital phases (purple and green points). If the observer were then to
-integrate the star's orbit using a model potential with the erroneous
-midplane, they would obtain the purple and green curves for the star's orbit,
-respectively. The actions estimated from these two  erroneous orbits would
-subsequently differ, both from each other and from the actions estimated for
-the true orbit (in the potential with the correct midplane). Hence an
-incorrect midplane in the potential model assumed will induce phase-dependence
-in the actions estimated for a given star in that potential.}
+\caption{Cartoon approximation showing the effect an error in the
+determination of the coordinate midplane can have on orbit integration and
+action estimation. The $x$-axis shows the orbital phase and the $y$-axis the
+vertical height. The top gray curve depicts an example ``true'' orbit
+oscillating about the true midplane (horizontal solid gray line). Consider an
+observer who erroneously assumes the midplane is located at the horizontal
+dashed line. Suppose this observer measures the phase-space position of the
+star at two different orbital phases (purple and green points). If the
+observer were then to integrate the star's orbit using a model potential with
+the erroneous midplane, they would obtain the purple and green curves for the
+star's orbit, respectively. The actions estimated from these two  erroneous
+orbits would subsequently differ, both from each other and from the actions
+estimated for the true orbit (in the potential with the correct midplane).
+Hence an incorrect midplane in the potential model assumed will induce
+phase-dependence in the actions estimated for a given star in that potential.}
 \label{fig:cartoon}
@@
 \subsection{Epicyclic Approximation} \label{ssec:epi_action}
-Before turning to numerical methods, we first write down analytically the systematic error in the actions induced from either a midplane offset or a Galactocentric radius
-offset. We use the epicyclic approximation, which assumes that the motion in
-the $z$ and $R$ components of the orbit are decoupled and follow simple
-harmonic motion about a circular and planar ``guiding orbit'' \citep[][Section 3.2 and references therein]{2008gady.book.....B}.
-We refer to the radius of this orbit (the ``guiding radius'') as $R_g$.
-This approximation is an excellent description of
-thin disk-like orbits and a good description of thick disk-like orbits in an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and spiral arms). We
-also make the assumption of a perfectly flat circular velocity curve with
-velocity $v_c$ \citep[decent near the Solar circle, e.g.][]{2017MNRAS.465...76M}.
+Before turning to numerical methods, we first write down analytically the
+systematic error in the actions induced from either a midplane offset or a
+Galactocentric radius offset. We use the epicyclic approximation, which
+assumes that the motion in the $z$ and $R$ components of the orbit are
+decoupled and follow simple harmonic motion about a circular and planar
+``guiding orbit'' \citep[][Section 3.2 and references
+therein]{2008gady.book.....B}. We refer to the radius of this orbit (the
+``guiding radius'') as $R_g$. This approximation is an excellent description
+of thin disk-like orbits and a good description of thick disk-like orbits in
+an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and
+spiral arms). We also make the assumption of a perfectly flat circular
+velocity curve with velocity $v_c$ \citep[decent near the Solar circle,
+e.g.][]{2017MNRAS.465...76M}.
 
@@
 \eeq
-where $\kappa$ and $\nu$ are the radial and vertical
-frequencies, $\Omega_c \equiv v_c/R_g$ the rotation frequency, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
-vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
-constants. Similarly, the velocities of the orbit are given by
+where $\kappa$ and $\nu$ are the radial and vertical frequencies, $\Omega_c
+\equiv v_c/R_g$ the rotation frequency, $A_z$ and $A_R$ are the amplitudes of
+the stellar motion in the vertical and radial coordinates, and $\alpha$ and
+$\beta$ are arbitrary constants. Similarly, the velocities of the orbit are
+given by
 \beq\label{eq:orbits_vel_epi}
@@
 
-In this case, we have that for $J_{\phi}$ \citep[][Section~3.5.3b]{2008gady.book.....B}, 
+In this case, we have that for $J_{\phi}$
+\citep[][Section~3.5.3b]{2008gady.book.....B},
 \beq\label{eq:Jphi_epi}
@@
 \beq\label{eq:induced_Jphi}
-\frac{\Delta J_{\phi}}{J_{\phi}} = \frac{\Delta R}{R_g} + \frac{\Delta v_{\phi}}{v_c}\text{.}
+\frac{\Delta J_{\phi}}{J_{\phi}} = \frac{\Delta R}{R_g} 
+                                    + \frac{\Delta v_{\phi}}{v_c}\text{.}
 \eeq
@@
 
-Since most of the time stars will be at maximum amplitude (i.e. turnaround) in both $R$ and $z$, we can approximate the order of magnitude of the systematic error in the actions by
+Since most of the time stars will be at maximum amplitude (i.e. turnaround) in
+both $R$ and $z$, we can approximate the order of magnitude of the systematic
+error in the actions by
 \beq\label{eq:Ji_err_mosttime}
@@
 
-In the remainder of this section, we compare our analytic estimates of the effect on actions of a midplane offset against numerical calculations.
-A numerical evaluation of the effect of velocity offsets on actions is deferred to future work, as we discuss in Section~\ref{ssec:lsr_var}.
+In the remainder of this section, we compare our analytic estimates of the
+effect on actions of a midplane offset against numerical calculations. A
+numerical evaluation of the effect of velocity offsets on actions is deferred
+to future work, as we discuss in Section~\ref{ssec:lsr_var}.
 
 \subsection{Numerical Methods} \label{ssec:action_comp}
-We now quantify the argument made in Section~\ref{ssec:cartoon} using numerical computations of the actions for a range of orbits in a model Galactic potential.
-We compute actions as in our previous work \citep{2018ApJ...867...31B}, using
-the code \texttt{gala} v0.3 to perform orbit integrations and conversion to
-action space \citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute actions
-we use the torus-mapping technique first presented by
-\citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to
-calculate actions for an orbital time-series starting from a phase-space
-position $(x, v)$ and integrated in a potential $\Phi$. We use the default
-\texttt{MWPotential} as our potential, which is based on the Milky Way
-potential available in \texttt{galpy} \citep{2015ApJS..216...29B}. This
-potential includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
-a Miyamoto-Nagai disk \citep{1975PASJ...27..533M}, and an NFW halo
-\citep{1997ApJ...490..493N}, and is fit to empirically match some
-observations. We use the Dormand-Prince 8(5,3) integration scheme
-\citep{Dormand80:integrator} with a timestep of $1\,\Myr$ and integrate for
-$5\,\Gyr$, corresponding to $\sim 20$ orbits for a Sun-like star.
+We now quantify the argument made in Section~\ref{ssec:cartoon} using
+numerical computations of the actions for a range of orbits in a model
+Galactic potential. We compute actions as in our previous work
+\citep{2018ApJ...867...31B}, using the code \texttt{gala} v0.3 to perform
+orbit integrations and conversion to action space
+\citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute actions we use the
+torus-mapping technique first presented by \citet{1990MNRAS.244..634M} and
+adapted by \citet{2014MNRAS.441.3284S} to calculate actions for an orbital
+time-series starting from a phase-space position $(x, v)$ and integrated in a
+potential $\Phi$. We use the default \texttt{MWPotential} as our potential,
+which is based on the Milky Way potential available in \texttt{galpy}
+\citep{2015ApJS..216...29B}. This potential includes a Hernquist bulge and
+nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
+\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
+is fit to empirically match some observations. We use the Dormand-Prince
+8(5,3) integration scheme \citep{Dormand80:integrator} with a timestep of
+$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
+Sun-like star.
 
@@
 
-We begin by considering the thick-disk orbit. Suppose an immortal observer who can observe
-the thick-disk star's phase-space position at many different times (and hence
-different orbital phases), but used a coordinate system in which the midplane
-is systematically offset in height by $100\,\pc$ from its actual location ---
-i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position in the
-orbit. This corresponds to the immortal observer physically located at e.g. the position
-$(8, 0, 0)\,\kpc$ in the coordinate system of the true potential, but
-erroneously thinking they are located at $(8, 0, 0.1)\,\kpc$.
+We begin by considering the thick-disk orbit. Suppose an immortal observer who
+can observe the thick-disk star's phase-space position at many different times
+(and hence different orbital phases), but used a coordinate system in which
+the midplane is systematically offset in height by $100\,\pc$ from its actual
+location --- i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position
+in the orbit. This corresponds to the immortal observer physically located at
+e.g. the position $(8, 0, 0)\,\kpc$ in the coordinate system of the true
+potential, but erroneously thinking they are located at $(8, 0, 0.1)\,\kpc$.
 
@@
 Starting from each observation of the star's phase-space position that such an
-immortal observer makes every $\Myr$, we then compute the
-actions as outlined above using an integration from that starting point in
-phase space carried out using the true potential but specifying the star's
-present-day position using the systematically offset coordinate system.
-Essentially we are shifting the integrated orbit and then reintegrating at
-each point along the shifted orbit, where the shifted orbit is the observed
-orbit. The erroneous computed actions for each phase-space starting point are
-shown for the first $\Gyr$ of the orbit in the {\em upper} panels
+immortal observer makes every $\Myr$, we then compute the actions as outlined
+above using an integration from that starting point in phase space carried out
+using the true potential but specifying the star's present-day position using
+the systematically offset coordinate system. Essentially we are shifting the
+integrated orbit and then reintegrating at each point along the shifted orbit,
+where the shifted orbit is the observed orbit. The erroneous computed actions
+for each phase-space starting point are shown for the first $\Gyr$ of the
+orbit in the {\em upper} panels
 Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
@@
 offset coordinate systems depend on the time (i.e. orbital phase) at which the
-star's phase-space position is observed. To the immortal observer
-this would appear as time-dependence in the actions, which appear to oscillate
-around their value using the correct coordinate system even though in this
-example the observer is using the correctly constructed best-fit static,
-axisymmetric potential. The relative size of the phase variation in each
-action depends on the direction of the systematic offset as well as the true
-values of the actions (i.e. the type of orbit). In reality we will have one
-measurement of the phase-space position to work with, in which case the
-determination of the orbital phase in $z$ or $R$ is degenerate with the degree
-of systematic offset in that coordinate (see Figure~\ref{fig:cartoon}).
-Additionally, in searching for migrated members of known open clusters, we
-would compute actions for stars with a wide range of orbital phases and
-locations in the Galaxy. In the following we therefore quote percentile ranges
-for the possible values computed for each action as a proxy for the effect of
-these systematic errors in the coordinate system.
+star's phase-space position is observed. To the immortal observer this would
+appear as time-dependence in the actions, which appear to oscillate around
+their value using the correct coordinate system even though in this example
+the observer is using the correctly constructed best-fit static, axisymmetric
+potential. The relative size of the phase variation in each action depends on
+the direction of the systematic offset as well as the true values of the
+actions (i.e. the type of orbit). In reality we will have one measurement of
+the phase-space position to work with, in which case the determination of the
+orbital phase in $z$ or $R$ is degenerate with the degree of systematic offset
+in that coordinate (see Figure~\ref{fig:cartoon}). Additionally, in searching
+for migrated members of known open clusters, we would compute actions for
+stars with a wide range of orbital phases and locations in the Galaxy. In the
+following we therefore quote percentile ranges for the possible values
+computed for each action as a proxy for the effect of these systematic errors
+in the coordinate system.
 
@@
 error in the Galactocentric coordinate system. We consider here the thick-disk
-orbit, which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520, 7.0)\,\actunit$
-and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We integrate the
-orbit according to the procedure laid out in Section~\ref{ssec:action_comp}.
-Then, we subtract $100\,\pc$ from the $z$ value ({\em upper panels}) or the
-$x$ value ({\em lower panels}) of each position in the orbit, corresponding to
-an erroneous observer assuming a midplane ({\em upper}) or solar radius ({\em
-lower}) that is off by $100\,\pc$. We then allow our (immortal) observer to
-observe the orbit over $1\,\Gyr$ and perform the same orbit integration
-procedure at each timestep, and report the values of the actions. The
-computation of $J_{\phi}$ is pristine to errors in $z$, with only numerical
-artifacts remaining. Only small errors are induced in $J_R$, with the middle
-$90\%$ of values over the $\Gyr$ being within $\sim8\%$ of the true $J_R$. As
-expected, large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$,
-with the middle $90\%$ of values being within $\sim43\%$ of the true $J_z$.
-The $x$ offset induces uncertainties in $J_R$ and $J_{\phi}$ of $\sim21\%$ and
-$\sim3\%$, respectively. A $\sim3\%$ error in $J_z$ is also induced.}
+orbit, which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520,
+7.0)\,\actunit$ and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We
+integrate the orbit according to the procedure laid out in
+Section~\ref{ssec:action_comp}. Then, we subtract $100\,\pc$ from the $z$
+value ({\em upper panels}) or the $x$ value ({\em lower panels}) of each
+position in the orbit, corresponding to an erroneous observer assuming a
+midplane ({\em upper}) or solar radius ({\em lower}) that is off by
+$100\,\pc$. We then allow our (immortal) observer to observe the orbit over
+$1\,\Gyr$ and perform the same orbit integration procedure at each timestep,
+and report the values of the actions. The computation of $J_{\phi}$ is
+pristine to errors in $z$, with only numerical artifacts remaining. Only small
+errors are induced in $J_R$, with the middle $90\%$ of values over the $\Gyr$
+being within $\sim8\%$ of the true $J_R$. As expected, large errors are
+induced in $J_z$ with a $100\,\pc$ offset in $z$, with the middle $90\%$ of
+values being within $\sim43\%$ of the true $J_z$. The $x$ offset induces
+uncertainties in $J_R$ and $J_{\phi}$ of $\sim21\%$ and $\sim3\%$,
+respectively. A $\sim3\%$ error in $J_z$ is also induced.}
 \label{fig:one_orbit_wrong_ref}
@@
 $J_z$ for the thick-disk orbit ({\em upper panel}) and thin-disk orbit ({\em
-lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. the {\em upper right} panel
-of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted
-as a vertical dashed line. Here we see that the errors in $J_z$ induced by a
-systematic offset in $z$ are non-Gaussian and bimodal. Furthermore, neither of
-the modes lie at zero. In other words, neither of the modes of the
-distribution of observed $J_z$ values are centered on the true $J_z$ (i.e. the
-$J_z$ computed with no offset in $z$). In the case of the thin-disk orbit
-({\em lower panel}), we see that, in addition to the prior complications, the
-error distribution is not even centered on null. This comes about when the
-midplane error is $\gtrsim z_{\text{max}}$.
+lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. the {\em upper right}
+panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is
+plotted as a vertical dashed line. Here we see that the errors in $J_z$
+induced by a systematic offset in $z$ are non-Gaussian and bimodal.
+Furthermore, neither of the modes lie at zero. In other words, neither of the
+modes of the distribution of observed $J_z$ values are centered on the true
+$J_z$ (i.e. the $J_z$ computed with no offset in $z$). In the case of the
+thin-disk orbit ({\em lower panel}), we see that, in addition to the prior
+complications, the error distribution is not even centered on null. This comes
+about when the midplane error is $\gtrsim z_{\text{max}}$.
 
 We now describe the intuition behind the shape of Figure~\ref{fig:Jz_hist}.
-Consider first the thick-disk orbit ({\em top} panel), where the
-offset in $z$ is much less than $z_{\text{max}}$ of the orbit. The peaks in
-the distribution correspond to the turning points of the orbit (or points of
-maximum vertical amplitude), where $v_z \sim 0$ and where the star spends
-most of it orbit. This is why the distribution is peaked on these values. For
-the thin-disk orbit ({\em bottom} panel), the offset in $z$ is comparable to
+Consider first the thick-disk orbit ({\em top} panel), where the offset in $z$
+is much less than $z_{\text{max}}$ of the orbit. The peaks in the distribution
+correspond to the turning points of the orbit (or points of maximum vertical
+amplitude), where $v_z \sim 0$ and where the star spends most of it orbit.
+This is why the distribution is peaked on these values. For the thin-disk
+orbit ({\em bottom} panel), the offset in $z$ is comparable to
 $z_{\text{max}}$. Now, there will be some points in the orbit where $v_z = 0$
@@
 (i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
-show the fractional spread in the values computed for $J_z$, $J_{\phi}$, and $J_R$,
-respectively. We compute the fractional spread $\Delta J_i/J_i$ as one half
-the 95\uth{} minus 5\uth{} percentile of the action computed using the offset
-coordinate system over the course of the first $\Gyr$, divided by the true
-action (computed in the correct coordinate system).
+show the fractional spread in the values computed for $J_z$, $J_{\phi}$, and
+$J_R$, respectively. We compute the fractional spread $\Delta J_i/J_i$ as one
+half the 95\uth{} minus 5\uth{} percentile of the action computed using the
+offset coordinate system over the course of the first $\Gyr$, divided by the
+true action (computed in the correct coordinate system).
 
 In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref},
-there is essentially no spread in the determination of $J_{\phi}$. This is expected
-since, in principle, $J_{\phi}$ only depends on the $x$- and $y$-components of the
-position and velocity of the stars\footnote{In practice, however, $J_{\phi}$ is
-computed as part of the torus-fitting method.}, and is thus unaffected by
-offsets in $z$. Indeed, the result we found in
+there is essentially no spread in the determination of $J_{\phi}$. This is
+expected since, in principle, $J_{\phi}$ only depends on the $x$- and
+$y$-components of the position and velocity of the stars\footnote{In practice,
+however, $J_{\phi}$ is computed as part of the torus-fitting method.}, and is
+thus unaffected by offsets in $z$. Indeed, the result we found in
 Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
@@
 include the estimation of the action errors derived under the epicyclic
-approximation from Equation~\eqref{eq:Ji_err_mosttime}.
-This Equation is relevant since during most of the orbit the star will be at maximum radial and vertical amplitude and we do not consider velocity offsets.
-Note that we consider an error
-in the $x$-coordinate $\Delta x$, which is not quite the same as $\Delta R$.
-For observations of stars close to us, we have that $\Delta x \sim \Delta R$,
-but for the experiment performed in this Section we consider observations of
-the star throughout its entire orbit. This introduces a factor of $2/\pi$ when
-converting from $\Delta R$ to $\Delta x$, which we derive in
-Appendix~\ref{app:deltax}.
+approximation from Equation~\eqref{eq:Ji_err_mosttime}. This Equation is
+relevant since during most of the orbit the star will be at maximum radial and
+vertical amplitude and we do not consider velocity offsets. Note that we
+consider an error in the $x$-coordinate $\Delta x$, which is not quite the
+same as $\Delta R$. For observations of stars close to us, we have that
+$\Delta x \sim \Delta R$, but for the experiment performed in this Section we
+consider observations of the star throughout its entire orbit. This introduces
+a factor of $2/\pi$ when converting from $\Delta R$ to $\Delta x$, which we
+derive in Appendix~\ref{app:deltax}.
 
@@
 all three orbits the epicyclic approximation is a good description of the
-$J_z$ error. An $x$ offset induces an error in $J_R$ and $J_{\phi}$. The error in
-$J_R$ is somewhat well-described for the thin-disk and thick-disk orbits, and
-a poor description for the halo orbit. For $J_{\phi}$, the epicyclic approximation
-is not a good description for any orbit.}
+$J_z$ error. An $x$ offset induces an error in $J_R$ and $J_{\phi}$. The error
+in $J_R$ is somewhat well-described for the thin-disk and thick-disk orbits,
+and a poor description for the halo orbit. For $J_{\phi}$, the epicyclic
+approximation is not a good description for any orbit.}
 \label{fig:many_orbit_wrong_ref}
@@
 would generate the action errors described in Section~\ref{sec:ref_frame}:
-uncertainties in measurements of the position and velocity of the Sun relative to the Galactic center. Converting from
-heliocentric to Galactocentric coordinates relies on these measurements.
-Therefore, errors in their values will induce a systematic offset in the
-Galactocentric phase-space position of any observed star.
+uncertainties in measurements of the position and velocity of the Sun relative
+to the Galactic center. Converting from heliocentric to Galactocentric
+coordinates relies on these measurements. Therefore, errors in their values
+will induce a systematic offset in the Galactocentric phase-space position of
+any observed star.
 
@@
 \citep[e.g.][]{2004ApJ...616..872R}. From stellar motions near
-Sgr~A\textsuperscript{*}, the distance from the Sun to Sgr~A\textsuperscript{*} can
-be precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a recent
-measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
-\citep{2019arXiv190405721A}.
-
-Importantly, the location of Sgr~A\textsuperscript{*} may not be equivalent to the location of the ``dynamical Galactic center:'' the point in three-dimensional
-space about which the stars in the solar neighborhood are orbiting.
-This assumption, although frequently made, has not yet been justified.
-It is tested in any construction of a  dynamical model where the distance from the Sun to
-the Galactic center, $R_0$, is a free parameter.
-For example, \citet{2015ApJ...803...80K} measured $R_0$ while
-modelling the dynamics of the stream Palomar 5. Many other dynamical measurements
-of $R_0$ have been made (\citealt{2016ARAA..54..529B} summarize many pre-{\em Gaia} results) but none have yet achieved a precision comparable to that of the distance to Sgr~A\textsuperscript{*}.
-% \citep{1981gask.book.....M, 2011PASJ...63..867S, 2012MNRAS.427..274S,2013AstL...39...95B, 2013IAUS..289..444Z}.
+Sgr~A\textsuperscript{*}, the distance from the Sun to
+Sgr~A\textsuperscript{*} can be precisely measured \citep{2009ApJ...692.1075G,
+2018AA...615L..15G}, with a recent measurement using near-infrared
+interferometry of $8.178 \pm 0.035$~kpc \citep{2019arXiv190405721A}.
+
+Importantly, the location of Sgr~A\textsuperscript{*} may not be equivalent to
+the location of the ``dynamical Galactic center:'' the point in
+three-dimensional space about which the stars in the solar neighborhood are
+orbiting. This assumption, although frequently made, has not yet been
+justified. It is tested in any construction of a  dynamical model where the
+distance from the Sun to the Galactic center, $R_0$, is a free parameter. For
+example, \citet{2015ApJ...803...80K} measured $R_0$ while modelling the
+dynamics of the stream Palomar 5. Many other dynamical measurements of $R_0$
+have been made (\citealt{2016ARAA..54..529B} summarize many pre-{\em Gaia}
+results) but none have yet achieved a precision comparable to that of the
+distance to Sgr~A\textsuperscript{*}.
 
 \subsection{Galactic Orientation}
-Second, one must define the angular orientation of the Galaxy. This
-was defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
+Second, one must define the angular orientation of the Galaxy. This was
+defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
 defining the coordinates of the Galactic center in B1950 coordinates as
-(17:42:26.6, -28:55:00) and the north Galactic pole as (12:49:00, +27:24:00). These two quantities, together with $R_0$, define the orientation of the Galactic plane.
-However, there is growing evidence that the stellar midplane is tilted relative to this coordinate system \citep{2014ApJ...797...53G, 2016ARAA..54..529B}, though not
-the H~\textsc{ii} midplane \citep{2019ApJ...871..145A}.
+(17:42:26.6, -28:55:00) and the north Galactic pole as (12:49:00, +27:24:00).
+These two quantities, together with $R_0$, define the orientation of the
+Galactic plane. However, there is growing evidence that the stellar midplane
+is tilted relative to this coordinate system \citep{2014ApJ...797...53G,
+2016ARAA..54..529B}, though not the H~\textsc{ii} midplane
+\citep{2019ApJ...871..145A}.
 
 \subsection{Solar Height}
-Third, one must define the Sun's vertical distance from the Galactic
-midplane, which can be determined by identifying where the stellar density
-and velocities reach a maximum (effectively the median height of all disk
-stars). The Solar height is usually taken to be $\sim 25\,\pc$
-\citep{2001ApJ...553..184C}, with a more recent measurement from {\em Gaia} DR2
-placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another strategy is
-to use the cold gas or H~\textsc{ii} regions in the disk to define the Galactic
-midplane, leading to slightly different values ($\sim 5\,\pc$) for the Sun's
-relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review
-of these measurements is given by \citet{2016ARAA..54..529B}.
+Third, one must define the Sun's vertical distance from the Galactic midplane,
+which can be determined by identifying where the stellar density and
+velocities reach a maximum (effectively the median height of all disk stars).
+The Solar height is usually taken to be $\sim 25\,\pc$
+\citep{2001ApJ...553..184C}, with a more recent measurement from {\em Gaia}
+DR2 placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another
+strategy is to use the cold gas or H~\textsc{ii} regions in the disk to define
+the Galactic midplane, leading to slightly different values ($\sim 5\,\pc$)
+for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em
+Gaia} review of these measurements is given by \citet{2016ARAA..54..529B}.
 
 \subsection{Local Standard of Rest}
-Finally, one must define the local standard of rest (LSR), or mean velocity of stars near the Sun relative to the Galactic center (which is defined to have zero velocity), and the velocity of the Sun relative to the LSR.
-The radial
-($U_{\odot}$) and vertical ($W_{\odot}$) components are computed by
-taking the mean motions of different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal component
-($V_{\odot}$) is more difficult to measure, but can be modelled using the
-asymmetric drift relation \citep{2008gady.book.....B}. The values of the
-components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}. The
-value of the circular velocity is taken to be $\sim 220\kms$
-\citep[e.g.][]{2012ApJ...759..131B}.
+Finally, one must define the local standard of rest (LSR), or mean velocity of
+stars near the Sun relative to the Galactic center (which is defined to have
+zero velocity), and the velocity of the Sun relative to the LSR. The radial
+($U_{\odot}$) and vertical ($W_{\odot}$) components are computed by taking the
+mean motions of different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}.
+The azimuthal component ($V_{\odot}$) is more difficult to measure, but can be
+modelled using the asymmetric drift relation \citep{2008gady.book.....B}. The
+values of the components of the LSR are usually taken from
+\citet{2010MNRAS.403.1829S}. The value of the circular velocity is taken to be
+$\sim 220\kms$ \citep[e.g.][]{2012ApJ...759..131B}.
 
@@
 
-Considerable effort has been placed on each of these
-measurements, but uncertainties remain, and detailed dynamical modelling
-across the disk --- particularly for dynamically cold stars --- may have to
-take them into account. For instance, the discrepancy of $\sim 5\,\pc$ between
-the midplane defined by stars and by gas will induce a $\sim10\%$ error in
-$J_z$ for an orbit with $z_{\text{max}}\sim100\,\pc$ (see
-Section~\ref{sec:ref_frame}).
+Considerable effort has been placed on each of these measurements, but
+uncertainties remain, and detailed dynamical modelling across the disk ---
+particularly for dynamically cold stars --- may have to take them into
+account. For instance, the discrepancy of $\sim 5\,\pc$ between the midplane
+defined by stars and by gas will induce a $\sim10\%$ error in $J_z$ for an
+orbit with $z_{\text{max}}\sim100\,\pc$ (see Section~\ref{sec:ref_frame}).
 
@@
 coordinate system: intrinsic variations of the midplane across the disk. We
-use two sets of simulations of Milky Way-mass galaxies to estimate the size of this effect. We will also
-briefly discuss a rough estimate of the midplane variation derived from the
-observed velocity distribution of the Milky Way, but defer a detailed
-treatment to future work.
+use two sets of simulations of Milky Way-mass galaxies to estimate the size of
+this effect. We will also briefly discuss a rough estimate of the midplane
+variation derived from the observed velocity distribution of the Milky Way,
+but defer a detailed treatment to future work.
 
@@
 softening length} & \colhead{stellar softening length} \\
-\colhead{} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} } 
+\colhead{} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)}
+& \colhead{(pc)} }
 \startdata
@@
 
-\subsection{Description of MW-Sagittarius Interaction Simulation} \label{ssec:sag_sim}
+\subsection{Description of MW-Sagittarius Interaction Simulation} 
+\label{ssec:sag_sim}
 In addition to the cosmological zoom-in, we will also briefly consider results
@@
 \Gus{something}, along with a stellar component modeled as a Hernquist sphere
-of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$.
-All components are realized with live distributions of N-body particles; the MW and Sgr are each initialized to be in equilibrium in isolation \Gus{check!}.
+of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$. All components
+are realized with live distributions of N-body particles; the MW and Sgr are
+each initialized to be in equilibrium in isolation \Gus{check!}.
 
@@
 \subsection{The Local Midplane} \label{ssec:local_midplane}
-Using the two sets of simulations, we determine the local midplane that an observer
-might measure if they were situated in each of these galaxies as a function of
-azimuth at the solar circle. Starting from the coordinate system described in
-the previous section, which is approximately aligned so that the
+Using the two sets of simulations, we determine the local midplane that an
+observer might measure if they were situated in each of these galaxies as a
+function of azimuth at the solar circle. Starting from the coordinate system
+described in the previous section, which is approximately aligned so that the
 $z$-coordinate is perpendicular to the disk plane, we place our imaginary
 observer at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth
-between $0<\phi<2\pi$.\footnote{The immortal observer also has warp
-drive.} At each value of $\phi$ we then compute the median $z$ for stars
-within a cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to
-the fiducial disk. We then re-define the new midplane of the cylinder to be the median $z$, re-select stars, and iterate until the median $z$ value
-converges. We find that only $10$ iterations of this procedure are necessary
-for convergence. The resulting median $z$ is taken to be what our observer
-would measure as the local Galactic midplane at each $\phi$. This procedure
-assumes perfect density estimation, and therefore perfect corrections for
-extinction, within the cylinder defining the ``solar neighborhood.'' Imperfect
-extinction correction is likely to increase the amplitude of the estimated
-fluctuations in $z$. To account for the effect of particle noise, we
-bootstrap-resample stars within a cylinder of stars with height $2\,\kpc$ and
-the same radius $1000$ times and determine the $1\,\sigma$ error bars by repeating the midplane determination with those reselection of stars.
+between $0<\phi<2\pi$.\footnote{The immortal observer also has warp drive.} At
+each value of $\phi$ we then compute the median $z$ for stars within a
+cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to the
+fiducial disk. We then re-define the new midplane of the cylinder to be the
+median $z$, re-select stars, and iterate until the median $z$ value converges.
+We find that only $10$ iterations of this procedure are necessary for
+convergence. The resulting median $z$ is taken to be what our observer would
+measure as the local Galactic midplane at each $\phi$. This procedure assumes
+perfect density estimation, and therefore perfect corrections for extinction,
+within the cylinder defining the ``solar neighborhood.'' Imperfect extinction
+correction is likely to increase the amplitude of the estimated fluctuations
+in $z$. To account for the effect of particle noise, we bootstrap-resample
+stars within a cylinder of stars with height $2\,\kpc$ and the same radius
+$1000$ times and determine the $1\,\sigma$ error bars by repeating the
+midplane determination with those reselection of stars.
 
@@
 form $\text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}$, with the values
-for $\text{A}$ being $9.5$, $2.5$, $-212$, and $-394\,\pc$, for $\text{B}$ being
-$0.0013$, $0.0007$, $-0.63$, and $-1.00\,\text{rad}$ and $\text{C}$ being $2.6$, $-7.8$,
-$-65$, and $-53\,\pc$. Each of these values are for each successive timestep. In the
-first two panels, we see that the midplane is relatively flat in the inner
-galaxy. But additional encounters drive strong midplane variation. In the
-third panel, we see a strong $m=2$ mode develop, consistent with the
-$R=8\,\kpc$ panel of Figure~17 in \citet{2018MNRAS.481..286L}.
+for $\text{A}$ being $9.5$, $2.5$, $-212$, and $-394\,\pc$, for $\text{B}$
+being $0.0013$, $0.0007$, $-0.63$, and $-1.00\,\text{rad}$ and $\text{C}$
+being $2.6$, $-7.8$, $-65$, and $-53\,\pc$. Each of these values are for each
+successive timestep. In the first two panels, we see that the midplane is
+relatively flat in the inner galaxy. But additional encounters drive strong
+midplane variation. In the third panel, we see a strong $m=2$ mode develop,
+consistent with the $R=8\,\kpc$ panel of Figure~17 in
+\citet{2018MNRAS.481..286L}.
 
@@
 ($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} ({\em left},
-{\em center}, and {\em right} panels). All calculations are done at cosmological redshift $\z =0$. The local midplane is determined at a
-position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
-and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
+{\em center}, and {\em right} panels). All calculations are done at
+cosmological redshift $\z =0$. The local midplane is determined at a position
+$\phi$ by taking the median height of all stars within $R=0.5\,\kpc$ and
+$z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
 using the new height $10$ times to converge on the local midplane height. In
@@
 Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
-within a $2\,\kpc$ height of the fiducial midplane to determine $1\,\sigma
-$ error bars, which we report as dashed lines.}
+within a $2\,\kpc$ height of the fiducial midplane to determine $1\,\sigma $
+error bars, which we report as dashed lines.}
 \label{fig:midplane}
@@
 disrupted. We would like to determine which stars with complete 6D phase-space
-positions, for example in the {\em Gaia} dataset, are likely to be ejected members from that cluster.\footnote{This is likely to
-be more successful with DR3/4 than DR2, since upcoming data releases will
-contain a far greater number of stars with radial velocities --- though
-targeted, follow-up radial velocity surveys may allow this program to be
-performed sooner.} One
-could generate an initial catalogue of candidate members by integrating the
-cluster's orbit and selecting stars close to that orbit, accounting for the
-fact that the stream of ejected stars will not exactly follow the orbit of the
-cluster \citep[e.g.][]{2011MNRAS.413.1852E}.
+positions, for example in the {\em Gaia} dataset, are likely to be ejected
+members from that cluster.\footnote{This is likely to be more successful with
+DR3/4 than DR2, since upcoming data releases will contain a far greater number
+of stars with radial velocities --- though targeted, follow-up radial velocity
+surveys may allow this program to be performed sooner.} One could generate an
+initial catalogue of candidate members by integrating the cluster's orbit and
+selecting stars close to that orbit, accounting for the fact that the stream
+of ejected stars will not exactly follow the orbit of the cluster
+\citep[e.g.][]{2011MNRAS.413.1852E}.
 
@@
 field star $i$ be close to the cluster's actions to within some bound, i.e.
-$\bm{J}_i \in V_e$, where $V_e$ is a volume centered on $\bm{J}_c$ and has size
-comparable to the intrinsic action-space dispersion of the disrupting
-cluster. This dispersion can be estimated as \citep[Section~8.3.3][]{2008gady.book.....B},
+$\bm{J}_i \in V_e$, where $V_e$ is a volume centered on $\bm{J}_c$ and has
+size comparable to the intrinsic action-space dispersion of the disrupting
+cluster. This dispersion can be estimated as
+\citep[Section~8.3.3][]{2008gady.book.....B},
 \beq \label{eq:action_disp}
@@
 error if the cluster and field star are far enough apart to have significantly
-different local midplanes (see Figures~\ref{fig:midplane}~and~\ref{fig:midplane_chervin}). Since our criterion for cluster membership (distance in action space) scales with the
+different local midplanes (see
+Figures~\ref{fig:midplane}~and~\ref{fig:midplane_chervin}). Since our
+criterion for cluster membership (distance in action space) scales with the
 cluster mass, the region where the search will succeed varies with cluster
@@
 \item Inaccuracy in the Galactocentric coordinate
-system induces phase-dependence in the actions calculated from the observed positions and velocities of stars
-(Figures~\ref{fig:cartoon}~and~\ref{fig:one_orbit_wrong_ref}).
-Since stars' instantaneous phase-space positions are measured without prior knowledge of their orbital phases, this results in systematic error in the computed actions
+system induces phase-dependence in the actions calculated from the observed
+positions and velocities of stars
+(Figures~\ref{fig:cartoon}~and~\ref{fig:one_orbit_wrong_ref}). Since stars'
+instantaneous phase-space positions are measured without prior knowledge of
+their orbital phases, this results in systematic error in the computed actions
 (Figure~\ref{fig:many_orbit_wrong_ref}).
 
-\item Inaccuracy in the midplane location most severely affects computation of the vertical action $J_z$.
-A midplane offset of $\sim6\,\pc$ ($\sim50\,\pc$) for a typical thin (thick) disk orbit results in a  $10\%$ error in $J_z$ (as defined by one half the middle
-$90\uth$ percentile). The fractional error is significantly less for halo orbits.
+\item Inaccuracy in the midplane location most severely affects computation of
+the vertical action $J_z$.  A midplane offset of $\sim6\,\pc$ ($\sim50\,\pc$)
+for a typical thin (thick) disk orbit results in a  $10\%$ error in $J_z$ (as
+defined by one half the middle $90\uth$ percentile). The fractional error is
+significantly less for halo orbits.
 
 \item The distribution of systematic errors in the actions induced by a
-coordinate system offset is very non-Gaussian. The distribution is bimodal with
-{\em neither mode at null}. 
-As a result, error propagation of coordinate system offsets is complex when considering actions, and is likely to significantly deform the action-space distribution function.
-
-\item Dynamical modelling across large regions of the disk, over which the midplane location varies by more than the limits discussed above, is
-susceptible to this type of systematic error, since the assumption that our local
-Galactic midplane is the global Galactic midplane is not true {\em a priori}.
+coordinate system offset is very non-Gaussian. The distribution is bimodal
+with {\em neither mode at null}. As a result, error propagation of coordinate
+system offsets is complex when considering actions, and is likely to
+significantly deform the action-space distribution function.
+
+\item Dynamical modelling across large regions of the disk, over which the
+midplane location varies by more than the limits discussed above, is
+susceptible to this type of systematic error, since the assumption that our
+local Galactic midplane is the global Galactic midplane is not true {\em a
+priori}.
 
@@
 simulations of Milky Way mass galaxies from the FIRE collaboration. We found
-that the midplane varies as a function of azimuth at the Solar circle by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
-three galaxies we considered (\mi{}, \mf{}, and \mm{}, respectively).
+that the midplane varies as a function of azimuth at the Solar circle by
+$\sim185, 162, 84\,\pc$ (middle $90\%$) for the three galaxies we considered
+(\mi{}, \mf{}, and \mm{}, respectively).
 
@@
 
-\item While in this work we have focused on systematic errors in action computation, all of our conclusions also extend to studies of stars that simply rely on orbit
-integration, since the computation of actions and orbit integrations are
-mathematically equivalent.
+\item While in this work we have focused on systematic errors in action
+computation, all of our conclusions also extend to studies of stars that
+simply rely on orbit integration, since the computation of actions and orbit
+integrations are mathematically equivalent.
 
@@
 Our main point is that the local midplane varies between different points in
-the Galaxy, and that this variation can lead to significant systematic errors in the computation of actions under the assumption of a global axisymmetric potential. We do
-not propose here what a modeler should use for the ``true'' global midplane,
-since it depends on the particular problem being studied. Even without a
-high-precision definition of the global Galactic midplane, current
-observations from {\em Gaia} should soon permit a measurement of the real
-azimuthal dependence of the midplane location.
-For some applications, such as those using actions as labels to group stars on similar orbits, using such a measurement to shift stars to a consistent midplane height as a function of azimuth before using a global axisymmetric approximation to the potential may be sufficient, although this ignores the \emph{dynamical} implications of shifts in the midplane height (which result from fluctuations in the local density).
-However, for other applications, such as the study of action diffusion which motivated this paper in the first place, a more extensive perturbative approach is likely to be needed.
-We plan to explore the mitigation of these effects in future work.
+the Galaxy, and that this variation can lead to significant systematic errors
+in the computation of actions under the assumption of a global axisymmetric
+potential. We do not propose here what a modeler should use for the ``true''
+global midplane, since it depends on the particular problem being studied.
+Even without a high-precision definition of the global Galactic midplane,
+current observations from {\em Gaia} should soon permit a measurement of the
+real azimuthal dependence of the midplane location. For some applications,
+such as those using actions as labels to group stars on similar orbits, using
+such a measurement to shift stars to a consistent midplane height as a
+function of azimuth before using a global axisymmetric approximation to the
+potential may be sufficient, although this ignores the \emph{dynamical}
+implications of shifts in the midplane height (which result from fluctuations
+in the local density). However, for other applications, such as the study of
+action diffusion which motivated this paper in the first place, a more
+extensive perturbative approach is likely to be needed. We plan to explore the
+mitigation of these effects in future work.
 
 \acknowledgments
-We would like to thank Megan Bedell, Robert A. Benjamin, Tobias Buck, Elena D'Onghia, Benoit
-Famaey, and Adrian Price-Whelan for helpful discussions. A.B. would like to
-thank Todd Phillips for helpful discussions. A.B. was supported in part by the
-Roy \& Diana Vagelos Program in the Molecular Life Sciences and the Roy \&
-Diana Vagelos Challenge Award. M.-M.M.L. was partly supported by NSF grant
-AST18-15461. The work of D.A.-A., A.B., D.W.H., C.F.P.L., R.S., K.V.J.,
-M.-M.M.L., and M.K.N. is supported by the Simons Foundation.
+We would like to thank Megan Bedell, Robert A. Benjamin, Tobias Buck, Elena
+D'Onghia, Benoit Famaey, and Adrian Price-Whelan for helpful discussions. A.B.
+would like to thank Todd Phillips for helpful discussions. A.B. was supported
+in part by the Roy \& Diana Vagelos Program in the Molecular Life Sciences and
+the Roy \& Diana Vagelos Challenge Award. M.-M.M.L. was partly supported by
+NSF grant AST18-15461. The work of D.A.-A., A.B., D.W.H., C.F.P.L., R.S.,
+K.V.J., M.-M.M.L., and M.K.N. is supported by the Simons Foundation.
 
@@
 R_{\text{err}}^2 &= R^2 - 2 R \cos{\phi} \Delta x \\
-\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\phi}} \Delta x\text{.}
+\implies \Delta R &\equiv \abs{R_{\text{err}} - R} 
+        = \abs{\cos{\phi}} \Delta x\text{.}
 \end{split}

### 2019-04-26 paper/ms.tex
Shuffle author order
Update caption to chervin midplane plot

@@
 
+\author{Douglas Filho}
+\affil{\columbia}
+
 \author{Mordecai-Mark Mac Low}
@@
 \affil{\cca}
-
-\author{Douglas Filho}
-\affil{\columbia}
 
@@
 for inaccuracies in the galactocentric coordinate system. Error bars are
-calculated as in Figure~\ref{fig:midplane}. (\Gus{Guessing here, need to
-figure out what the panels actually show:}) The left two panels show the
-midplane as a function of azimuth before the first encounter. As later
-encounters occur, we can see that stronger midplane variations result. In
-fact, one can see that an $m=2$ mode is excited in the third panel --- this is
-consistent with Figure~17 of \citet{2018MNRAS.481..286L}, which shows that an
-$m=2$ mode is excited (note that $m=0$ and $m=1$ modes are stronger, but these
-are removed in our sine-curve subtraction).}
+calculated as in Figure~\ref{fig:midplane}. (\Gus{Check:}) The {\em upper}
+panels show the midplane as a function of azimuth before the first encounter
+near the Solar circle. As later encounters occur which bring the
+Sagittarius-like object closer to the Solar circle, we can see that stronger
+midplane variations result. In fact, one can see that an $m=2$ mode is excited
+in the third panel --- this is consistent with Figure~17 of
+\citet{2018MNRAS.481..286L}, which shows that an $m=2$ mode is excited (note
+that $m=0$ and $m=1$ modes are stronger, but these are removed in our
+sine-curve subtraction). The fact that the $t=6.9\,\Gyr$ panel looks
+qualitatively similar to the panels from the FIRE simulations
+(Figure~\ref{fig:midplane}) is evidence that the midplane variations are
+produced in part by mergers.}
 \label{fig:midplane_chervin}

### 2019-04-30 paper/ms.tex
Fix eq ref in Fig 4 caption
Update author information
Remove all mentions of clusters

@@
 
-\author{Angus Beane}
+\author[0000-0002-8658-1453]{Angus Beane}
 \affil{\cca}
@@
 
-\author{Robyn Sanderson}
+\author[0000-0003-3939-3297]{Robyn Sanderson}
 \affil{\cca}
@@
 
-\author{Douglas Filho}
+\author[0000-0001-8275-9181]{Douglas Grion Filho}
 \affil{\columbia}
 
-\author{Mordecai-Mark Mac Low}
+\author[0000-0003-0064-4060]{Mordecai-Mark Mac Low}
 \affil{\cca}
@@
 
-\author{David W. Hogg}
+\author[0000-0003-2866-9403]{David W. Hogg}
 \affil{\nyuphys}
@@
 
-\keywords{Galaxy:~disk -- solar~neighborhood --
-open~clusters~and~associations:~general -- Galaxy:~structure --
+\keywords{Galaxy:~disk -- solar~neighborhood -- Galaxy:~structure --
 Galaxy:~evolution}
@@
 variations, and the resulting systematic uncertainty in the vertical action,
-for action-space analyses beyond the dynamical solar neighborhood. In light of
-our findings, we considered the orbits of the nine open clusters within
-$250\,\pc$ as reported by the {\em Gaia} collaboration
-\citep{2018AA...616A..10G}. In Table~\ref{tab:real_clusters}, we report the
-distance to each cluster, the three actions, and the maximum vertical extent
-of their orbits ($z_{\text{max}}$). In Section~\ref{sssec:reconstruction} we
-estimate and discuss the size of the dynamical solar neighborhood for each
-cluster based on its mass, to show that action-space tagging is likely still a
-viable method if contamination from foreground or background stars is
-unimportant. We summarize and conclude in Section~\ref{sec:conclusion}.
-
-\begin{deluxetable*}{cCCCCCCCC}
-\tablecaption{Relevant dynamical parameters for the $9$ open clusters within
-$250\,\pc$ of the Sun, using the relevant initial positions and velocities
-from \citet{2018AA...616A..10G}. Actions are computed using the method
-outlined in Section~\ref{ssec:action_comp}, with typical values for the
-Galactocentric coordinate system parameters. These are only shown to
-illustrate the general types of orbits open clusters are
-on.\label{tab:real_clusters}}
-\tablehead{\colhead{cluster} & \colhead{$N_{\text{memb}}$} &
-\colhead{distance} & \colhead{$J_R$} &
-\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
-\colhead{mass} \\
-\colhead{ } & \colhead{ } &
-\colhead{$\mathrm{pc}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{pc}$} & \colhead{$M_{\odot}$}}
-\startdata
-alphaPer & 740 & 170 & 20 & -1700 & 0.005 & 11 & 352\tablenotemark{b} \\
-Blanco1 & 489 & 240 & 2 & -1800 & 2 & 210 & 410\tablenotemark{c} \\
-ComaBer & 153 & 90 & 2 & -1900 & 0.7 & 140 & 112\tablenotemark{d} \\
-Hyades & 515 & 50 & 20 & -1800 & 0.3 & 80 & 400\tablenotemark{e} \\
-IC2391 & 325 & 150 & 6 & -1800 & 0.02 & 20 & \nodata \\
-IC2602 & 492 & 150 & 10 & -1700 & 0.3 & 89 & \nodata \\
-NGC2451A\tablenotemark{a} & 400 & 190 & 6 & -1800 & 0.2 & 73 & \nodata \\
-Pleiades & 1326 & 140 & 20 & -1700 & 0.4 & 96 & 800\tablenotemark{f} \\
-Praesepe & 938 & 190 & 20 & -1800 & 0.6 & 120 & 550\tablenotemark{d}
-\enddata
-
-\tablenotetext{a}{This cluster is labeled NGC2451, which we interpret as
-NGC2451A since NGC2451B lies further than $250\,\pc$ away.}
-\tablenotetext{b}{\citet{2016MNRAS.457.1028S}}
-\tablenotetext{c}{\citet{2007AA...471..499M}}
-\tablenotetext{d}{\citet{2007AJ....134.2340K}}
-\tablenotetext{e}{\citet{1998AA...331...81P}}
-\tablenotetext{f}{\citet{2001AJ....121.2053A}}
-
-\tablerefs{\citet{2018AA...616A..10G}}
-\end{deluxetable*}
+for action-space analyses. In particular, we consider the impact on modelling
+the distribution function of the Milky Way in Section~\ref{ssec:df}. We
+summarize and conclude in Section~\ref{sec:conclusion}.
 
@@
 orbital phase in $z$ or $R$ is degenerate with the degree of systematic offset
-in that coordinate (see Figure~\ref{fig:cartoon}). Additionally, in searching
-for migrated members of known open clusters, we would compute actions for
-stars with a wide range of orbital phases and locations in the Galaxy. In the
-following we therefore quote percentile ranges for the possible values
-computed for each action as a proxy for the effect of these systematic errors
-in the coordinate system.
+in that coordinate (see Figure~\ref{fig:cartoon}).  In the following we
+therefore quote percentile ranges for the possible values computed for each
+action as a proxy for the effect of these systematic errors in the coordinate
+system.
 
@@
 dashed lines the epicyclic prediction of the induced action error
-(Equations~\eqref{eq:actions_epi_error}~and~\eqref{eq:Jr_epi_dx}). In the
-epicyclic approximation, a $z$ offset only induces an error in $J_z$ --- for
-all three orbits the epicyclic approximation is a good description of the
-$J_z$ error. An $x$ offset induces an error in $J_R$ and $J_{\phi}$. The error
-in $J_R$ is somewhat well-described for the thin-disk and thick-disk orbits,
-and a poor description for the halo orbit. For $J_{\phi}$, the epicyclic
-approximation is not a good description for any orbit.}
+(Equation~\eqref{eq:Ji_err_mosttime}). In the epicyclic approximation, a $z$
+offset only induces an error in $J_z$ --- for all three orbits the epicyclic
+approximation is a good description of the $J_z$ error. An $x$ offset induces
+an error in $J_R$ and $J_{\phi}$. The error in $J_R$ is somewhat
+well-described for the thin-disk and thick-disk orbits, and a poor description
+for the halo orbit. For $J_{\phi}$, the epicyclic approximation is not a good
+description for any orbit.}
 \label{fig:many_orbit_wrong_ref}
@@
 quasi-harmonic oscillations around a circular ``guiding center orbit.'' Here
-we consider a few of those implications, starting with an estimate of the
-region around the Sun where we expect an extension of the local axisymmetric
-coordinate system, and therefore the epicyclic approximation, to still be
-valid. This region is also where it should be feasible to perform ``dynamical
-tagging'' in the thin disk; that is, using the computed orbits or integrals of
-motion of field stars to match them with known open clusters.
-
-\subsection{Disrupted Cluster Reconstruction}
-\label{sssec:reconstruction}
-
-We now consider the implications for attempts to reconstruct disrupting open
-clusters by selecting ejected members based on their actions.
-
-Suppose an open cluster of mass $m_c$ on an orbit with actions $\bm{J}_c$ is
-disrupted. We would like to determine which stars with complete 6D phase-space
-positions, for example in the {\em Gaia} dataset, are likely to be ejected
-members from that cluster.\footnote{This is likely to be more successful with
-DR3/4 than DR2, since upcoming data releases will contain a far greater number
-of stars with radial velocities --- though targeted, follow-up radial velocity
-surveys may allow this program to be performed sooner.} One could generate an
-initial catalogue of candidate members by integrating the cluster's orbit and
-selecting stars close to that orbit, accounting for the fact that the stream
-of ejected stars will not exactly follow the orbit of the cluster
-\citep[e.g.][]{2011MNRAS.413.1852E}.
-
-One could then make a further selection by requiring that the action of each
-field star $i$ be close to the cluster's actions to within some bound, i.e.
-$\bm{J}_i \in V_e$, where $V_e$ is a volume centered on $\bm{J}_c$ and has
-size comparable to the intrinsic action-space dispersion of the disrupting
-cluster. This dispersion can be estimated as
-\citep[Section~8.3.3][]{2008gady.book.....B},
-\beq \label{eq:action_disp}
-\frac{\Delta J_i}{J_i} \sim \left(\frac{m_c}{M}\right)^{1/3}\text{,}
-\eeq
-where $m_c$ is the mass of the cluster and $M$ is the mass enclosed by the
-cluster's orbit around the Galaxy, in this case assumed to be the mass inside
-the Solar circle ($\sim 10^{11}\Msun$). Stars with actions within $V_e$ would
-then be considered cluster members.
-
-Two issues are immediately apparent with this approach. The first is that it
-neglects the overall smooth distribution \emph{in action space} which will
-contribute some stars to the selection by chance. For dissolving open clusters
-in the thin disk, this background is likely to be significant, but we will
-assume for the sake of argument that these foreground stars can be modeled
-within the ``dynamical solar neighborhood'' where it is safe to extend the
-local coordinate system. Since the effect of extending the coordinate frame
-beyond this region is generally to inflate the actions computed for thin-disk
-stars, using too large a volume is likely to under-estimate the density of the
-smooth thin-disk-like component in action space, while scattering more
-thin-disk stars onto thick-disk-like orbits and over-estimating the density of
-the smooth thick-disk-like component.
-
-The second issue, which is related to the point of this paper, is that based
-on the results in our previous sections computing $\bm{J}_i$ for a field star
-and directly comparing it to  $\bm{J}_c$ will have significant systematic
-error if the cluster and field star are far enough apart to have significantly
-different local midplanes (see
-Figures~\ref{fig:midplane}~and~\ref{fig:midplane_chervin}). Since our
-criterion for cluster membership (distance in action space) scales with the
-cluster mass, the region where the search will succeed varies with cluster
-mass as well. To illustrate this effect, we assume a cluster mass $m_c$ and
-estimate the expected action-space dispersion from
-Equation~\eqref{eq:action_disp}. We then consult
-Figure~\ref{fig:many_orbit_wrong_ref} to determine the midplane offset
-necessary in order for the induced error in the vertical action to be the same
-as the expected action-space dispersion. We perform this analysis for the
-thin-disk-like and thick-disk-like orbits considered in
-Section~\ref{ssec:quant}, assuming that the halo-like orbit is unlikely for
-members of known open clusters (see Table~\ref{tab:real_clusters}). The result
-for the allowable midplane offset for these two orbits in shown in
-Figure~\ref{fig:cluster_offset} as a function of cluster mass.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{fig/cluster_offset.pdf}
-\end{center}
-\caption{The $z$ offset necessary in order for the action error induced by our
-midplane effect to be comparable to the intrinsic action dispersion from a
-cluster of mass $m_c$. We compute the intrinsic dispersion from
-Equation~\eqref{eq:action_disp} and convert the dispersion to a $z$ offset using
-the result from Figure~\ref{fig:many_orbit_wrong_ref}. We consider the result
-for the thin and thick orbit (Table~\ref{tab:orbits}). Nearby open clusters
-(Table~\ref{tab:real_clusters}) have orbits closest to the thin orbit.}
-\label{fig:cluster_offset}
-\end{figure}
+we consider those implications for the prospect of distribution function
+fitting.
+
+\subsection{Distribution Function Fitting}
+\label{ssec:df}
 

### 2019-05-03 paper/ms.tex
Make changes to abstract

@@
 induces a phase-dependence in the actions, which we interpret as a systematic
-error. The induced error distribution is non-Gaussian and bimodal, with
-neither of the modes peaking on the null value. A variation in the vertical
-position midplane of $\sim15\,\pc$ for a thin disk orbit or $\sim120\,\pc$ for
+error. An offset in the vertical
+position of $\sim15\,\pc$ for a thin disk orbit or $\sim120\,\pc$ for
 a thick disk orbit is sufficient to induce a $25\,\%$ systematic error in the
 vertical action $J_z$ computed under the assumption of an axisymmetric
-potential. Furthermore, we show that the local midplane varies by
+potential. The induced error distribution is non-Gaussian and bimodal, with
+neither of the modes peaking on the null value. Furthermore, we show that the local midplane varies by
 $\sim100\,\pc$ at the Solar circle in Milky Way-mass cosmological zoom-in
-simulations from the FIRE collaboration. Thus, current state-of-the-art action
+simulations from the FIRE project.
+From observations that the mean vertical velocity of the Milky Way disk varies by $\sim5\,\kms$ with radius, we estimate that the true midplane variations are of order $\sim100\,\pc$.
+Thus, current state-of-the-art action
 calculations --- which assume the global and local midplanes are the same ---
-include a systematic vertical offset variation. Variation in the local
+are likely to include a systematic vertical offset variation depending on the volume considered. Variation in the local
 standard of rest induces similar issues. The variation of the midplane must be

### 2019-05-03 paper/ms.tex
Update acknowledgements

@@
 in part by the Roy \& Diana Vagelos Program in the Molecular Life Sciences and
-the Roy \& Diana Vagelos Challenge Award. M.-M.M.L. was partly supported by
-NSF grant AST18-15461. The work of D.A.-A., A.B., D.W.H., C.F.P.L., R.S.,
-K.V.J., M.-M.M.L., and M.K.N. is supported by the Simons Foundation.
+the Roy \& Diana Vagelos Challenge Award. K.V.J.'s work was performed in part
+during the Gaia19 workshop and the 2019 Santa Barbara Gaia Sprint (also
+supported by the Heising-Simons Foundation), both hosted by the Kavli
+Institute for Theoretical Physics at the University of California, Santa
+Barbara. K.V.J.'s contributions were supported in part by the National Science
+Foundation under grants NSF PHY-1748958 and NSF AST-1614743. M.-M.M.L. was
+partly supported by NSF grant AST18-15461. The work of D.A.-A., A.B., D.W.H.,
+C.F.P.L., R.S., K.V.J., M.-M.M.L., and M.K.N. is supported by the Simons
+Foundation.
 

### 2019-05-03 paper/ms.tex
Small changes throughout

@@
 J_i \equiv
-\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,.}
+\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{.}
 \eeq
@@
 the $p_i$ are the conjugate momenta. In a slowly-evolving axisymmetric
-potential, these actions are invariant and $J_{\phi} \equiv L_z$
+potential, these actions are invariant and $J_{\phi} \equiv L_z$, where $L_z$ is the $z$-component of the angular momentum
 \citep{2008gady.book.....B,2014RvMP...86....1S}.
@@
 mismatch will result in large systematic uncertainty, especially in the
-vertical action $J_z$. Depending on the type of orbit being studied, this sets
-a limit on the ``dynamical solar neighborhood'' over which the local
-coordinate frame can be usefully extended for action-space analysis. A similar
+vertical action $J_z$. A similar
 argument applies to any remaining uncertainty in measurements of the
@@
 collaboration\footnote{\url{https://fire.northwestern.edu}}
-\citep{2016ApJ...827L..23W, 2017arXiv170206148H}, and a controlled N-body
+\citep{2014MNRAS.445..581H, 2016ApJ...827L..23W, 2018MNRAS.480..800H}, and a controlled N-body
 simulation of a Sagittarius encounter with a galaxy otherwise tailored to the
@@
 the distribution function of the Milky Way in Section~\ref{ssec:df}. We
-summarize and conclude in Section~\ref{sec:conclusion}.
+summarize our main results and conclude in Section~\ref{sec:conclusion}.
 
@@
 true midplane. The dashed gray line is the midplane location in the potential
-model used to integrate the orbit of the star and hence estimate its actions,
+model used by an observer to integrate the orbit of the star and hence estimate its actions,
 which is offset from the true midplane of the star's orbit. The model
@@
 where $\kappa$ and $\nu$ are the radial and vertical frequencies, $\Omega_c
-\equiv v_c/R_g$ the rotation frequency, $A_z$ and $A_R$ are the amplitudes of
+\equiv v_c/R_g$ is the rotation frequency, $A_z$ and $A_R$ are the amplitudes of
 the stellar motion in the vertical and radial coordinates, and $\alpha$ and
@@
 propagation of errors formula to Equation~\eqref{eq:JR_Jz_epi_energy}
-determine the error in each of the actions. For $J_{\phi}$ we have that,
+to determine the error in each of the actions. For $J_{\phi}$ we have that,
 \beq\label{eq:induced_Jphi}
@@
 numerical computations of the actions for a range of orbits in a model
-Galactic potential. We compute actions as in our previous work
-\citep{2018ApJ...867...31B}, using the code \texttt{gala} v0.3 to perform
+Galactic potential. We compute actions as in
+\citet{2018ApJ...867...31B}, using the code \texttt{gala} v0.3 to perform
 orbit integrations and conversion to action space
@@
 We begin by considering the thick-disk orbit. Suppose an immortal observer who
-can observe the thick-disk star's phase-space position at many different times
+can measure the thick-disk star's phase-space position at many different times
 (and hence different orbital phases), but used a coordinate system in which
@@
 \tablecaption{Description and names of the three orbits considered in this
-work. For the last two columns: $z_{\text{max}}$ the maximum height of the
+work. For the last two columns: $z_{\text{max}}$ is the maximum height of the
 orbit while $A_R$ is the magnitude of the radial excursions of the orbit,
@@
 
-Starting from each observation of the star's phase-space position that such an
-immortal observer makes every $\Myr$, we then compute the actions as outlined
-above using an integration from that starting point in phase space carried out
-using the true potential but specifying the star's present-day position using
-the systematically offset coordinate system. Essentially we are shifting the
+We consider that the immortal observer makes a measurement every $\Myr$.
+At each measurement, we compute the actions as outlined above using the true potential but by specifiying the star's starting position using the systematically offset coordinate system.
+Essentially we are shifting the
 integrated orbit and then reintegrating at each point along the shifted orbit,
@@
 $6.2\,\actunit$ for $J_R$ and $J_z$, respectively. These are fractional errors
-of $5.7\,\%$ and $85.7\,\%$, respectively. The error induced in $J_{\phi}$ is
+of $5.7\,\%$ and $86\,\%$, respectively. The error induced in $J_{\phi}$ is
 negligible, as expected since $J_{\phi}$ is independent of $z$. It is worth
@@
 $J_z$ for the thick-disk orbit ({\em upper panel}) and thin-disk orbit ({\em
-lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. the {\em upper right}
-panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is
-plotted as a vertical dashed line. Here we see that the errors in $J_z$
+lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. as in the {\em upper
+right} panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$
+is plotted as a vertical dashed line. Here we see that the errors in $J_z$
 induced by a systematic offset in $z$ are non-Gaussian and bimodal.
@@
 complications, the error distribution is not even centered on null. This comes
-about when the midplane error is $\gtrsim z_{\text{max}}$.
+about when the midplane error is $\gtrsim z_{\text{max}}$, where
+$z_{\text{max}}$ is the maximum height of the orbit (equivalent to $A_z$ in
+Section~\ref{ssec:epi_action}).
 
@@
 show the fractional spread in the values computed for $J_z$, $J_{\phi}$, and
-$J_R$, respectively. We compute the fractional spread $\Delta J_i/J_i$ as one
-half the 95\uth{} minus 5\uth{} percentile of the action computed using the
-offset coordinate system over the course of the first $\Gyr$, divided by the
-true action (computed in the correct coordinate system).
+$J_R$, respectively. We compute the fractional spread $\Delta J_i/J_i$, where
+$\Delta J_i$ is defined in the previous paragraph and $J_i$ is the action
+value computed in the true coordinate system, over the course of the first
+$\Gyr$.
 
@@
 
-The epicyclic approximation is a good predictor of $J_z$, even for halo
-orbits. It performs similarly for $J_R$, now underpredicting for halo orbits
+The epicyclic approximation is a good predictor of $\Delta J_z$, even for halo
+orbits. It performs similarly for $\Delta J_R$, now underpredicting for halo orbits
 and slightly overpredicting for thin-disk orbits. Note that for the particular
@@
 thick-disk orbit, and so we expect the epicyclic approximation to perform
-slightly worse. The epicyclic approximation underpredicts $J_{\phi}$ for all
+slightly worse. The epicyclic approximation underpredicts $\Delta J_{\phi}$ for all
 orbits.
@@
 strategy is to use the cold gas or H~\textsc{ii} regions in the disk to define
-the Galactic midplane, leading to slightly different values ($\sim 5\,\pc$)
+the Galactic midplane, leading to slightly different values (by $\sim 5\,\pc$)
 for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em
@@
 \mf{}, and \mm{}, respectively. These values are summarized in
-Table~\ref{tab:scale_height}.
+Table~\ref{tab:scale_height}, along with measurements of the stellar and gas
+disk scale heights.
 

### 2019-05-03 paper/ms.tex
Put in Daniel's FIRE paragraph

@@
 \subsection{Description of FIRE Simulations} \label{ssec:cosmozoom}
-We briefly describe the cosmological-hydrodynamical zoom-in simulations used
-in this work. Cosmological zoom-ins allow one to simulate a selected region at
-high resolution embedded in a low-resolution cosmological background
-\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added
-a hydrodynamical simulation of the baryonic component with recipes for star
-formation and feedback. The FIRE collaboration has simulated a number of Milky
-Way-mass galaxies using this technique as part of the {\em Latte} suite of
-FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. For this
-work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S},
-which show broad agreement of many of their global properties with
-observations of the Milky Way. The $\z=0$ snapshots\footnote{In this work, to
-avoid confusion with the vertical height $z$, we refer to cosmological
-redshift as $\z$.} of these three simulations, called \mi{}, \mf{}, and \mm{}
-for shorthand, are publicly available alongside associated mock {\em Gaia} DR2
+The FIRE cosmological hydrodynamic simulations
+\citep{2014MNRAS.445..581H,2018MNRAS.480..800H} use the zoom-in technique
+\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O} to model the formation
+of a small group of galaxies at high resolution in a full cosmological
+context.  Feedback from supernovae, stellar winds, and radiation from massive
+stars is implemented at the scale of star forming regions following stellar
+population synthesis models, generating galactic winds self-consistently
+\citep{2015MNRAS.454.2691M, 2017MNRAS.470.4698A} while reproducing many
+observed galaxy properties, including stellar masses, star formation
+histories, metallicities, and morphologies and kinematics of thin and thick
+disks \citep{2014MNRAS.445..581H, 2016MNRAS.456.2140M, 2017MNRAS.467.2430M,
+2016ApJ...827L..23W, 2018MNRAS.481.4133G, 2018MNRAS.480..800H}. For this work,
+we focus on the three Milky Way-mass zoom-ins considered in
+\citet{2018arXiv180610564S}, which were simulated as part of the
+\textit{Latte} suite and show broad agreement of many of their global
+properties with observations of the Milky Way \citep{2016ApJ...827L..23W,
+2018MNRAS.481.4133G} The $\z = 0$ snapshots\footnote{In this work, to avoid
+confusion with the vertical height $z$, we refer to cosmological redshift as
+$\z$.} of these three simulations, named \mi{}, \mf{}, and \mm{} for
+shorthand, are publicly available alongside associated mock \textit{Gaia} DR2
 catalogues generated from them.\footnote{\url{http://ananke.hub.yt}}

### 2019-05-03 paper/ms.tex
Remove gas splitting bug footnote

@@
 of mass $\sim 5000 \text{--} 7000\,
-\Msun$,\footnote{In fact, the simulation \mi{} contained a gas
-splitting bug which causes a small number of gas and star particles to have
-larger than typical masses. These constitute only $\sim0.2\%$ of the star
-particles in the $\z=0$ snapshots, and are only a factor of a few more massive
-than typical star particles. Since our results are broadly consistent between
-the three simulations, we ignore this minor complication.} with the lower end
+\Msun$, with the lower end
 coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths

### 2019-05-03 cartoon/cartoon.py
Update color handling in cartoon

@@ -38,23 +38,23 @@ ax.set_xlim((-1.0995574287564276, 23.09070600388498))
 ax.set_ylim((-3.726564988290398, 1.2578647540983607))
 
 # plot orbit, midplane, erroneous midplane
-ax.plot(x, y, c=tb_c[-1])
-ax.plot(x, np.full(len(x), 0), c=tb_c[-1])
-ax.plot(x, np.full(len(x), -1.25), c=tb_c[-1], ls='dashed')
+ax.plot(x, y, c=tb_c['gray'])
+ax.plot(x, np.full(len(x), 0), c=tb_c['gray'])
+ax.plot(x, np.full(len(x), -1.25), c=tb_c['gray'], ls='dashed')
 
 fig.tight_layout()
 fig.savefig('cartoon_0.pdf', bbox_inches='tight', pad_inches=0)
 
 #plot measurements
-ax.scatter(3.*np.pi, -1, c=tb_c[4], zorder=4)
-ax.scatter(4.*np.pi, 1, c=tb_c[6], zorder=5)
+ax.scatter(3.*np.pi, -1, c=tb_c['orange'], zorder=4)
+ax.scatter(4.*np.pi, 1, c=tb_c['cyan'], zorder=5)
 
 fig.tight_layout()
 fig.savefig('cartoon_1.pdf', bbox_inches='tight', pad_inches=0)
 
 # plot "integrated" orbits
-ax.plot(x, y2, c=tb_c[4], zorder=4)
-ax.plot(x, y3, c=tb_c[6], zorder=5)
+ax.plot(x, y2, c=tb_c['orange'], zorder=4)
+ax.plot(x, y3, c=tb_c['cyan'], zorder=5)
 
 fig.tight_layout()
 fig.savefig('cartoon.pdf', bbox_inches='tight', pad_inches=0)

### 2019-05-03 paper/ms.tex
Add footnote about diversity of galaxies at MW mass

@@
 Cosmological simulations of Milky Way-mass galaxies are not perfect
-representations of the true Milky Way in other ways as well, as discussed in
+representations of the true Milky Way in other ways as well\footnote{The failure of cosmological simulations to exactly reproduce the Milky Way is not entirely due to limitations of the numerical model.
+Candidate Milky Way-like galaxies are chosen solely on their mass and isolation, for which there are a wide variety of possible galaxies with qualitatively different properties.}, as discussed in
 \citet{2018arXiv180610564S}. For instance, the velocity structure of \mi{} is

### 2019-05-03 paper/ms.tex
Small changes throughout

@@
 correspond to the turning points of the orbit (or points of maximum vertical
-amplitude), where $v_z \sim 0$ and where the star spends most of it orbit.
+amplitude), where $v_z \sim 0$ and where the star is on most of it orbit.
 This is why the distribution is peaked on these values. For the thin-disk
@@
 Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
-orbit types.
+orbit types. This is also a demonstration of the robustness of the orbit
+integration and action calculation methods we use.
 
@@
 \begin{deluxetable*}{cccccc}
-\tablecaption{Stellar disk scale heights of the real Milky Way and the
+\tablecaption{Stellar and gas disk scale heights of the real Milky Way and the
 simulated galaxies considered in this work.\label{tab:scale_height}}
@@
 
-\subsection{Description of MW-Sagittarius Interaction Simulation} 
+\subsection{Description of Milky Way-Sagittarius Interaction Simulation} 
 \label{ssec:sag_sim}
-In addition to the cosmological zoom-in, we will also briefly consider results
+In addition to the cosmological zoom-ins, we will also briefly consider results
 from a toy Sagittarius encounter simulation. This simulation offers us the
@@
 The mass resolution of the simulation is $2.6\times10^4$, $1.2\times10^4$, and
-$1.0\times10^4$ for the dark matter, disk, and bulge components, respectively.
+$1.0\times10^4\,\Msun$ for the dark matter, disk, and bulge components, respectively.
 For the disk and bulge components, a softening length of $30\,\pc$ is used
@@
 respectively.
+
+The fiducial coordinate system for these toy simulations is the rest frame of
+the aligned toy galaxy at the beginning of the simulation.
 
@@
 being $0.0013$, $0.0007$, $-0.63$, and $-1.00\,\text{rad}$ and $\text{C}$
-being $2.6$, $-7.8$, $-65$, and $-53\,\pc$. Each of these values are for each
-successive timestep. In the first two panels, we see that the midplane is
-relatively flat in the inner galaxy. But additional encounters drive strong
-midplane variation. In the third panel, we see a strong $m=2$ mode develop,
+being $2.6$, $-7.8$, $-65$, and $-53\,\pc$ chronologically for each of the four timesteps shown. In the {\em upper} two panels, we see that the midplane is
+relatively flat in the inner galaxy, but additional encounters drive strong
+midplane variation. In the {\em lower left} panel, we see a strong $m=2$ mode develop,
 consistent with the $R=8\,\kpc$ panel of Figure~17 in
@@
 ($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} ({\em left},
-{\em center}, and {\em right} panels). All calculations are done at
+{\em center}, and {\em right} panels) as a function of azimuthal angle. All calculations are done at
 cosmological redshift $\z =0$. The local midplane is determined at a position
@@
 \caption{The local midplane determined at the fiducial Solar radius
-($8.2\,\kpc$) for 4 different time steps from previously run toy simulations
+($8.2\,\kpc$) for 4 different time steps from toy simulations
 of a Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
@@
 midplane variations result. In fact, one can see that an $m=2$ mode is excited
-in the third panel --- this is consistent with Figure~17 of
+in the {\em lower left} panel --- this is consistent with Figure~17 of
 \citet{2018MNRAS.481..286L}, which shows that an $m=2$ mode is excited (note
@@
 We consider the variations of the LSR as a function of azimuth at the fiducial
-solar circle ($R_{\odot} = 8.2\,\kpc$). At each azimuth, $\phi$, we take the
-median velocity in cylindrical coordinates of all stars within $200\,\pc$ of
-the position, following \citet{2018arXiv180610564S}. No best-fit subtraction
-was performed as in Figure~\ref{fig:midplane}.
+solar circle ($R_{\odot} = 8.2\,\kpc$) in Figure~\ref{fig:lsr_variations}. At
+each azimuth, $\phi$, we take the median velocity in cylindrical coordinates
+of all stars within $200\,\pc$ of the position, following
+\citet{2018arXiv180610564S}. No best-fit subtraction was performed as in
+Figure~\ref{fig:midplane}.
 
@@
 (Section~\ref{ssec:local_midplane}).}
+\label{fig:lsr_variations}
 \end{figure*}

### 2019-05-04 paper/ms.tex
Move and expand observational discussion

@@
 
-An interesting, qualitative calculation can be made based on the velocity
-variations seen in e.g. \citet{2019arXiv190209569F}. The velocity variations
-in this work (admittedly found with respect to $J_{\phi}$ or radius) are
-$\sim5\text{--}10\,\kms$. The vertical frequency of the thin-disk and
-thick-disk orbits are $\sim 0.09\,\Myr^{-1}$ and $\sim0.06\,\Myr^{-1}$,
-respectively. By dimensional analysis, we therefore expect the midplane
-offsets to be of order $\sim 57\text{--}170\,\pc$. We stress that this is a
-rough calculation, but it shows that already we see evidence in the data from
-velocities for midplane offsets on the order of what we see in both sets of
-simulations.
-
 \section{Discussion} \label{sec:discussion}
@@
 
-\subsection{Distribution Function Fitting}
-\label{ssec:df}
+\subsection{Estimate of Milky Way Midplane Offsets from Current Data}
+\label{ssec:mw_data_midplane}
+
+Systematic variations in $v_z$ were first noted as asymmetries in the local
+velocity distribution towards the North and South Galactic Caps from the
+radial velocity surveys of the Sloan Digital Sky Surveys
+\citep{2012ApJ...750L..41W} and RAdial Velocity Experiment
+\citep{2013MNRAS.436..101W}. Subsequently, \citet{2013ApJ...777L...5C} pointed
+out suggestions of an oscillation in average vertical velocities of order
+$5\,\kms$  on $\sim\,\kpc$ scales looking toward the Galactic anticenter. Work
+by the {\em Gaia} collaboration confirmed these preliminary results on the
+velocity and spatial scales of these oscillation with clear spatial maps made
+using DR2 data of median $v_z$ over a significant Galactic volume
+\citep{2018A&A...616A..11G, 2019arXiv190209569F}. We also see in the FIRE
+simulation that the vertical velocity variation as a function of azimuth is
+$\sim5\text{--}10\,\kms$ (Figure~\ref{fig:lsr_variations}), consistent with
+these observations.
+
+The vertical frequency of the thin-disk and thick-disk orbits are $\sim
+0.09\,\Myr^{-1}$ and $\sim0.06\,\Myr^{-1}$, respectively. By dimensional
+analysis, and assuming a vertical velocity variation of $5\,\kms$, we
+therefore expect the midplane offsets to be $\sim 57\textup{--}85\,\pc$. We
+stress that this is a rough calculation, but it shows that already we see
+evidence in the data from velocities for midplane offsets on the order of what
+we saw in both sets of simulations.
 

### 2019-05-04 paper/ms.tex
Add observations to conclusions, update abstract number

@@
 simulations from the FIRE project.
-From observations that the mean vertical velocity of the Milky Way disk varies by $\sim5\,\kms$ with radius, we estimate that the true midplane variations are of order $\sim100\,\pc$.
+From observations that the mean vertical velocity of the Milky Way disk varies by $\sim5\,\kms$ with radius, we estimate that the true midplane variations are of order $\sim60\textup{--}90\,\pc$.
 Thus, current state-of-the-art action
@@
 
+\item Assuming a vertical velocity variation of the Milky Way of
+$\sim5\,\kms$, consistent with recent results from {\em Gaia} DR2
+\citep{2018A&A...616A..11G, 2019arXiv190209569F} and our results from the FIRE
+simulations (Figure~\ref{fig:lsr_variations}), we estimated that the
+corresponding midplane offsets are $\sim60\textup{--}90\,\pc$ by dimensional
+analysis with the vertical frequencies of disk-like orbits.
+
 \item This work emphasizes the importance of combining chemistry and dynamics.

### 2019-05-04 paper/ms.tex
Update rationale for ignoring LSR variations

@@
 in the previous section) is more involved. Since we find that the variation in
-the LSR is less pronounced than for the midplane, and since the dominant
-effect is limited to $J_R$ (Figure~\ref{fig:many_orbit_wrong_ref}), we defer
-this calculation to future work.
+the LSR is less pronounced than for the midplane, and since offsets in
+velocity only contribute to second order to $J_R$ and $J_z$ when a star is at
+maximum amplitude in $R$ or $z$ (where the majority of the orbit is, see
+Section~\ref{ssec:epi_action}), we defer this calculation to future work.
 

### 2019-05-04 paper/ms.tex
Update keywords
Small changes to abstract, R0 discussion

@@
 calculations --- which assume the global and local midplanes are the same ---
-are likely to include a systematic vertical offset variation depending on the volume considered. Variation in the local
+are likely to include a systematic vertical offset depending on the volume considered. Variation in the local
 standard of rest induces similar issues. The variation of the midplane must be
 taken into account when performing dynamical modelling across the large
-regions of the disk accessible to {\em Gaia}.
+regions of the disk accessible to {\em Gaia} and future missions.
 
@@
 
-\keywords{Galaxy:~disk -- solar~neighborhood -- Galaxy:~structure --
-Galaxy:~evolution}
+\keywords{Galaxy: disk -- Galaxy: evolution -- Galaxy: kinematics and dynamics
+-- Galaxy: structure -- solar neighborhood -- stars: kinematics and dynamics}
 
@@
 three-dimensional space about which the stars in the solar neighborhood are
-orbiting. This assumption, although frequently made, has not yet been
+orbiting. This assumption, although sensible and frequently made, has not yet been
 justified. It is tested in any construction of a  dynamical model where the

### 2019-05-04 paper/ms.tex
Update LSR abbreviatons

@@
 argument applies to any remaining uncertainty in measurements of the
-Galactocentric radius of the Sun and to variations in the local standard of
-rest (LSR).
+Galactocentric radius of the Sun and to variations in the LSR.
 
@@
 \subsection{Local Standard of Rest}
-Finally, one must define the local standard of rest (LSR), or mean velocity of
+Finally, one must define the LSR, or mean velocity of
 stars near the Sun relative to the Galactic center (which is defined to have
@@
 \subsection{Velocity Variations} \label{ssec:lsr_var}
-We also expect that the local standard of rest (LSR) should vary as a function
+We also expect that the LSR should vary as a function
 of azimuth. We perform this calculation in Appendix~\ref{app:lsr} to estimate
@@
 \end{center}
-\caption{The local standard of rest (LSR) as a function of azimuth at the
+\caption{The LSR as a function of azimuth at the
 fiducial Solar circle ($R_{\odot} = 8.2\,\kpc$). No best-fit subtraction is

### 2019-05-04 paper/ms.tex
Swap two appendix sections to match order in text
Fix some abbreviations

@@
 \newcommand{\mm}{\texttt{m12m}}
+
+\newcommand{\sgra}{Sgr~A\textsuperscript{*}}
 
@@
 First, one must define the center of the Galaxy. This is usually taken to be
-the location of the central supermassive black hole, Sgr~A\textsuperscript{*}
-\citep[e.g.][]{2004ApJ...616..872R}. From stellar motions near
-Sgr~A\textsuperscript{*}, the distance from the Sun to
-Sgr~A\textsuperscript{*} can be precisely measured \citep{2009ApJ...692.1075G,
+the location of the central supermassive black hole, Sagittarius~A\textsuperscript{*}
+\citep[\sgra{}, e.g.][]{2004ApJ...616..872R}. From stellar motions near
+\sgra{}, the distance from the Sun to
+\sgra{} can be precisely measured \citep{2009ApJ...692.1075G,
 2018AA...615L..15G}, with a recent measurement using near-infrared
@@
 
-Importantly, the location of Sgr~A\textsuperscript{*} may not be equivalent to
+Importantly, the location of \sgra{} may not be equivalent to
 the location of the ``dynamical Galactic center:'' the point in
@@
 results) but none have yet achieved a precision comparable to that of the
-distance to Sgr~A\textsuperscript{*}.
+distance to \sgra{}.
 
@@
 \startdata
-MW\tablenotemark{b} & 40 & 300 & 900 & \nodata & \nodata \\
+Milky Way\tablenotemark{b} & 40 & 300 & 900 & \nodata & \nodata \\
 \mi\tablenotemark{c} & 800\tablenotemark{d} & 480 & 2000 & 12.4 & 11.2 \\
@@
 simulations, especially the cold gas. Indeed, the stellar scale heights of the
-simulated galaxies are equal to or larger than the MW's while the gas scale
+simulated galaxies are equal to or larger than the Milky Way's while the gas scale
 heights are significantly larger (although the proper basis comparison is less
@@
 in the midplane \emph{as defined by the stellar mass density}, which will
-subsequently be lower in the simulations than in the MW: within 200 pc of the
-Sun the measured gas and stellar volume densities in the MW are roughly
+subsequently be lower in the simulations than in the Milky Way: within 200 pc of the
+Sun the measured gas and stellar volume densities in the Milky Way are roughly
 comparable at $\sim 0.04\ \Msun\
@@
 midplane variation we find in the simulated systems is likely to be
-\emph{less} than in the real MW, since on the smallest scales the simulations
+\emph{less} than in the real Milky Way, since on the smallest scales the simulations
 are \emph{smoother} than the real Galaxy.
@@
 of the Milky Way. While they have prominent spiral arms, none has as strong a
-bar as the MW does at present day, and none has a nearby companion like the
+bar as the Milky Way does at present day, and none has a nearby companion like the
 Large Magellanic Cloud. One of the three we consider (\mf) does have an
@@
 the degree of midplane variation in the simulations is likely a factor of a
-few lower than what could be expected in the MW.
+few lower than what could be expected in the Milky Way.
 
@@
 of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$. All components
-are realized with live distributions of N-body particles; the MW and Sgr are
+are realized with live distributions of N-body particles; the Milky Way and Sagittarius are
 each initialized to be in equilibrium in isolation \Gus{check!}.
@@
 \appendix
+\section{Orbits} \label{app:orbits}
+We plot the three orbits considered throughout the work
+(Table~\ref{tab:orbits}) in Figure~\ref{fig:plot_orbits}.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{fig/orbits.pdf}
+\end{center}
+\caption{The three orbits presented in Table~\ref{tab:orbits} and considered
+throughout the work. We plot the thin, thick, and halo orbits in the {\em
+left}, {\em center}, and {\em right} columns, respectively. The {\em upper}
+row shows a plot of $x$ vs. $y$ while the {\em lower} row shows $R$ vs. $z$.}
+\label{fig:plot_orbits}
+\end{figure*}
+
 \section{$\Delta R$-$\Delta x$ Relation} \label{app:deltax}
@@
 
-\section{Orbits} \label{app:orbits}
-We plot the three orbits considered throughout the work
-(Table~\ref{tab:orbits}) in Figure~\ref{fig:plot_orbits}.
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{fig/orbits.pdf}
-\end{center}
-\caption{The three orbits presented in Table~\ref{tab:orbits} and considered
-throughout the work. We plot the thin, thick, and halo orbits in the {\em
-left}, {\em center}, and {\em right} columns, respectively. The {\em upper}
-row shows a plot of $x$ vs. $y$ while the {\em lower} row shows $R$ vs. $z$.}
-\label{fig:plot_orbits}
-\end{figure*}
-
 \section{LSR Variations} \label{app:lsr}

### 2019-05-04 paper/ms.tex
Fix lead-ins to equations

@@
 coordinate system. They are the cyclical integral of the canonical momentum
-over its conjugate position,
+over its conjugate position:
 \beq\label{eq:actions}
@@
 Under this approximation, we can write down the cylindrical components
-of the orbits as,
+of the orbits as
 \beq\label{eq:orbits_epi}
@@
 $\beta$ are arbitrary constants. Similarly, the velocities of the orbit are
-given by
+given by:
 \beq\label{eq:orbits_vel_epi}
@@
 In this case, we have that for $J_{\phi}$
-\citep[][Section~3.5.3b]{2008gady.book.....B},
+\citep[][Section~3.5.3b]{2008gady.book.....B}:
 \beq\label{eq:Jphi_epi}
@@
 \eeq
-and for $J_R$ and $J_z$,
+and for $J_R$ and $J_z$:
 \beq\label{eq:JR_Jz_epi}
@@
 \eeq
-We can simplify this and write,
+We can simplify this and write
 \beq\label{eq:JR_Jz_epi_final}
@@
 propagation of errors formula to Equation~\eqref{eq:JR_Jz_epi_energy}
-to determine the error in each of the actions. For $J_{\phi}$ we have that,
+to determine the error in each of the actions. For $J_{\phi}$, the induced error is:
 \beq\label{eq:induced_Jphi}
@@
 \eeq
-For $J_R$, the induced error is,
+For $J_R$, the induced error is:
 \beq\label{eq:induced_JR}
@@
 \eeq
-For $J_z$, the induced error is,
+For $J_z$, the induced error is:
 \beq\label{eq:induced_Jz}

### 2019-05-04 paper/ms.tex
Fix simplified epicylic approx of Jphi
Move stuff about no error in Jphi earlier

@@
 \begin{split}
-\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta R}{R_g} \\
+\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta R}{R_g} + \frac{\Delta v_{\phi}}{v_c} \\
 \frac{\Delta J_{R}}{J_{R}} &= \frac{2\Delta R}{A_R} \\
@@
 of $5.7\,\%$ and $86\,\%$, respectively. The error induced in $J_{\phi}$ is
-negligible, as expected since $J_{\phi}$ is independent of $z$. It is worth
+negligible, as expected since $J_{\phi}$ only depends on the $x$- and
+$y$-components of the position and velocity of the stars.\footnote{In practice,
+however, $J_{\phi}$ is computed as part of the torus-fitting method.} It is worth
 pointing out that a $100\,\pc$ error in an orbit with
@@
 there is essentially no spread in the determination of $J_{\phi}$. This is
-expected since, in principle, $J_{\phi}$ only depends on the $x$- and
-$y$-components of the position and velocity of the stars\footnote{In practice,
-however, $J_{\phi}$ is computed as part of the torus-fitting method.}, and is
-thus unaffected by offsets in $z$. Indeed, the result we found in
+expected since $J_{\phi}$ is independent of $z$ and is
+thus unaffected by offsets in $z$, as discussed earlier. Indeed, the result we found in
 Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all

### 2019-05-04 paper/ms.tex
Remove volume density argument, move def of midplane to section intro

@@
 We now consider a second source of systematic offsets in the Galactocentric
-coordinate system: intrinsic variations of the midplane across the disk. We
-use two sets of simulations of Milky Way-mass galaxies to estimate the size of
-this effect. We will also briefly discuss a rough estimate of the midplane
-variation derived from the observed velocity distribution of the Milky Way,
-but defer a detailed treatment to future work.
+coordinate system: intrinsic variations of the midplane across the disk. In
+this work we specifically consider azimuthal variations in the midplane as
+defined by the stellar mass density. We use two sets of simulations of Milky
+Way-mass galaxies to estimate the size of this effect.
 
@@
 precluding extending the precision of the gas midplane definition to the
-stellar component. In this work we specifically consider azimuthal variations
-in the midplane \emph{as defined by the stellar mass density}, which will
-subsequently be lower in the simulations than in the Milky Way: within 200 pc of the
-Sun the measured gas and stellar volume densities in the Milky Way are roughly
-comparable at $\sim 0.04\ \Msun\
-\unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
-Table 3 of \citealt{2018arXiv180610564S}). We thus expect that the degree of
-midplane variation we find in the simulated systems is likely to be
-\emph{less} than in the real Milky Way, since on the smallest scales the simulations
-are \emph{smoother} than the real Galaxy.
+stellar component.
 

### 2019-05-04 schmactions/one_orbit/plot_one_orbit.py
Update Jz hist plot

@@ -233,25 +233,25 @@ k0, k1, k2 = sclip(zact)
 dJzJz = 0.5*(np.percentile(zact[:,2][k2], 95) - np.percentile(zact[:,2][k2], 5))
 
 ax[0].hist(zact[:,2][k2], bins=np.linspace(histmin, histmax, 60),
-        edgecolor='k', fc='none', histtype='stepfilled')
+        edgecolor=tb_c[0], fc='none', histtype='stepfilled', linewidth=1.5)
 ax[0].arrow(J2, 60, dJzJz, 0, head_width=1, head_length=0.4, length_includes_head=True, color='k')
-ax[0].text(J2+dJzJz/3.5, 61, r'$\Delta J_z$', color='k')
+ax[0].text(J2+dJzJz/4, 61, r'$\Delta J_z$', color='k')
 
 ax[0].axvline(x=J2, color='k', ls='dashed', lw=1)
 ax[0].set_ylabel(r'$\text{count}$')
-ax[0].text(14.5, 75, r'\text{thick-disk orbit}', color='k')
+ax[0].text(15, 78, r'\text{thick-disk orbit}', color='k')
 
 k0t, k1t, k2t = sclip(zact_thin)
 dJzJz_thin = 0.5*(np.percentile(zact_thin[:,2][k2t], 95) - np.percentile(zact_thin[:,2][k2t], 5))
 
 ax[1].hist(zact_thin[:,2][k2t], bins=np.linspace(histmin_thin, histmax_thin, 60),
-        edgecolor='k', fc='none', histtype='stepfilled')
+        edgecolor=tb_c[0], fc='none', histtype='stepfilled', linewidth=1.5)
 ax[1].arrow(J2_thin, 60, dJzJz_thin, 0, head_width=1, head_length=0.15, length_includes_head=True, color='k')
 ax[1].text(J2_thin+dJzJz_thin/3.5, 61.5, r'$\Delta J_z$', color='k')
 
 ax[1].axvline(x=J2_thin, color='k', ls='dashed', lw=1)
 ax[1].set_ylabel(r'$\text{count}$')
-ax[1].text(1.3, 100, r'\text{thin-disk orbit}', color='k')
+ax[1].text(1, 104, r'\text{thin-disk orbit}', color='k')
 
 ax[1].set_xlabel(r'$J_{z,\text{obs}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
 

### 2019-05-04 paper/ms.tex
Update Jz hist plot
Update panel referencing

@@
 for each phase-space starting point are shown for the first $\Gyr$ of the
-orbit in the {\em upper} panels
+orbit in the upper panels
 Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
@@
 
-For a systematic offset in $z$ ({\em upper} panels), the
+For a systematic offset in $z$ (upper panels), the
 95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
@@
 integrate the orbit according to the procedure laid out in
-Section~\ref{ssec:action_comp}. Then, we subtract $100\,\pc$ from the $z$
-value ({\em upper panels}) or the $x$ value ({\em lower panels}) of each
+Section~\ref{ssec:action_comp}, and which we plot in Appendix~\ref{app:orbits}. Then, we subtract $100\,\pc$ from the $z$
+value (upper panels) or the $x$ value (lower panels) of each
 position in the orbit, corresponding to an erroneous observer assuming a
-midplane ({\em upper}) or solar radius ({\em lower}) that is off by
-$100\,\pc$. We then allow our (immortal) observer to observe the orbit over
+midplane (upper) or solar radius (lower) that is off by
+$100\,\pc$. We then allow an immortal observer to observe the orbit over
 $1\,\Gyr$ and perform the same orbit integration procedure at each timestep,
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{fig/schmactions_Jz_hist.pdf}
+\includegraphics[width=\columnwidth]{fig/schmactions_Jz_zerr_hist.pdf}
 \end{center}
@@
 In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
-$J_z$ for the thick-disk orbit ({\em upper panel}) and thin-disk orbit ({\em
-lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. as in the {\em upper
-right} panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$
+$J_z$ for the thick-disk orbit (top panel) and thin-disk orbit (bottom panel) assuming a $z$ offset of $100\,\pc$ (i.e. as in the upper
+right panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$
 is plotted as a vertical dashed line. Here we see that the errors in $J_z$
@@
 $J_z$ (i.e. the $J_z$ computed with no offset in $z$). In the case of the
-thin-disk orbit ({\em lower panel}), we see that, in addition to the prior
+thin-disk orbit (bottom panel), we see that, in addition to the prior
 complications, the error distribution is not even centered on null. This comes
@@
 We now describe the intuition behind the shape of Figure~\ref{fig:Jz_hist}.
-Consider first the thick-disk orbit ({\em top} panel), where the offset in $z$
+Consider first the thick-disk orbit (top panel), where the offset in $z$
 is much less than $z_{\text{max}}$ of the orbit. The peaks in the distribution
@@
 This is why the distribution is peaked on these values. For the thin-disk
-orbit ({\em bottom} panel), the offset in $z$ is comparable to
+orbit (bottom panel), the offset in $z$ is comparable to
 $z_{\text{max}}$. Now, there will be some points in the orbit where $v_z = 0$
@@
 
-The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
+The upper row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
 in each action for an offset in the $z$-component (i.e. the Galactic
-midplane). In the {\em bottom row} we consider offsets in the $x$ component
-(i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
+midplane). In the lower row we consider offsets in the $x$ component
+(i.e. the Solar radius). The left, center, and right columns
 show the fractional spread in the values computed for $J_z$, $J_{\phi}$, and
@@
 
-In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref},
+In the upper middle panel of Figure~\ref{fig:many_orbit_wrong_ref},
 there is essentially no spread in the determination of $J_{\phi}$. This is
@@
 
-The {\em upper right} panel of Figure~\ref{fig:many_orbit_wrong_ref} shows
+The upper right panel of Figure~\ref{fig:many_orbit_wrong_ref} shows
 that the fractional error in $J_z$ is more exaggerated for more planar
@@
 
-For the offset in the Solar radius ({\em bottom panels}), the error is largest
+For the offset in the Solar radius (lower panels), the error is largest
 for $J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
-deviations in $J_z$. In the {\em bottom middle} and {\em bottom right} panels
+deviations in $J_z$. In the lower middle and lower right panels
 all three lines nearly overlap.
@@
 text for the justification in using this to measure the magnitude of the
-induced error. The {\em left}, {\em center}, and {\em right} panels show the
-result for $J_R$, $J_{\phi}$, and $J_z$, respectively. The {\em upper} panels
-consider an offset in $z$ and the {\em lower} panels consider an offset in $x$
+induced error. The left, center, and right panels show the
+result for $J_R$, $J_{\phi}$, and $J_z$, respectively. The upper panels
+consider an offset in $z$ and the lower panels consider an offset in $x$
 (equivalently, an offset in the Solar radius). In some panels we also plot as
@@
 relatively flat in the inner galaxy, but additional encounters drive strong
-midplane variation. In the {\em lower left} panel, we see a strong $m=2$ mode develop,
+midplane variation. In the lower left panel, we see a strong $m=2$ mode develop,
 consistent with the $R=8\,\kpc$ panel of Figure~17 in
@@
 \caption{The local midplane determined at the fiducial Solar radius
-($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} ({\em left},
-{\em center}, and {\em right} panels) as a function of azimuthal angle. All calculations are done at
+($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} (left,
+center, and right panels) as a function of azimuthal angle. All calculations are done at
 cosmological redshift $\z =0$. The local midplane is determined at a position
@@
 for inaccuracies in the galactocentric coordinate system. Error bars are
-calculated as in Figure~\ref{fig:midplane}. (\Gus{Check:}) The {\em upper}
+calculated as in Figure~\ref{fig:midplane}. (\Gus{Check:}) The upper
 panels show the midplane as a function of azimuth before the first encounter
@@
 midplane variations result. In fact, one can see that an $m=2$ mode is excited
-in the {\em lower left} panel --- this is consistent with Figure~17 of
+in the lower left panel at $t=6.0\,\Gyr$ --- this is consistent with Figure~17 of
 \citet{2018MNRAS.481..286L}, which shows that an $m=2$ mode is excited (note
@@
 \caption{The three orbits presented in Table~\ref{tab:orbits} and considered
-throughout the work. We plot the thin, thick, and halo orbits in the {\em
-left}, {\em center}, and {\em right} columns, respectively. The {\em upper}
-row shows a plot of $x$ vs. $y$ while the {\em lower} row shows $R$ vs. $z$.}
+throughout the work. We plot the thin-disk, thick-disk, and halo orbits in the left, center, and right columns, respectively. The upper
+row shows a plot of $x$ vs. $y$ while the lower row shows $R$ vs. $z$.}
 \label{fig:plot_orbits}

### 2019-05-04 paper/ms.tex
Small changes throughout

@@
 ``local'' Galactic midplane is not the same as the ``global'' Galactic
-midplane. For stars relatively far from the Sun, the local midplane location
-will not agree with our local midplane extrapolated onto their position, due
+midplane. For stars relatively far from the Sun, their local midplane location
+will not agree with our local midplane extrapolated onto their position due
 to relatively small-scale variations in the density of the gas and stellar
@@
 
-In Section~\ref{sec:discussion} we discuss the implications of midplane
+In Section~\ref{sec:discussion}, we discuss the implications of midplane
 variations, and the resulting systematic uncertainty in the vertical action,
-for action-space analyses. In particular, we consider the impact on modelling
-the distribution function of the Milky Way in Section~\ref{ssec:df}. We
+for action-space analyses. We also estimate the expected midplane variations of the Milky Way based on the observed velocity variations. We
 summarize our main results and conclude in Section~\ref{sec:conclusion}.
@@
 Now suppose this observer makes a measurement of the star's position and
-velocity at the purple point or the green point (i.e. at two different orbital
-phases). Then the purple and green curves correspond to the orbits that the
+velocity at the teal point or the orange point (i.e. at two different orbital
+phases). Then the teal and orange curves correspond to the orbits that the
 observer would compute for each point based on the potential model with the
 offset midplane. In action space, this would correspond to a different value
-of $J_z$ for the purple and green points. In this way, assuming the wrong
+of $J_z$ for the teal and orange points. In this way, assuming the wrong
 coordinate system induces a phase-dependence in the actions estimated for the
@@
 dashed line. Suppose this observer measures the phase-space position of the
-star at two different orbital phases (purple and green points). If the
+star at two different orbital phases (teal and orange points). If the
 observer were then to integrate the star's orbit using a model potential with
-the erroneous midplane, they would obtain the purple and green curves for the
+the erroneous midplane, they would obtain the teal and orange curves for the
 star's orbit, respectively. The actions estimated from these two  erroneous
@@
 decoupled and follow simple harmonic motion about a circular and planar
-``guiding orbit'' \citep[][Section 3.2 and references
+``guiding orbit'' \citep[][Section~3.2 and references
 therein]{2008gady.book.....B}. We refer to the radius of this orbit (the
@@
 spiral arms). We also make the assumption of a perfectly flat circular
-velocity curve with velocity $v_c$ \citep[decent near the Solar circle,
-e.g.][]{2017MNRAS.465...76M}.
+velocity curve with velocity $v_c$, a good approximation near the Solar circle \citep[e.g.][]{2017MNRAS.465...76M}.
 
@@
 \eeq
-We can simplify this and write
+Using Equations~\eqref{eq:orbits_epi} and \eqref{eq:orbits_vel_epi}, we can
+simplify this and write
 \beq\label{eq:JR_Jz_epi_final}
@@
 \eeq
-where the last equality comes from the fact that $v_{R,\text{max}} = \kappa
-A_R$ and similarly for $z$.
+where the last equality in each line comes from the fact that $v_{R,\text{max}} = \kappa
+A_R$ and $v_{z,\text{max}} = \nu A_z$.
 
@@
 In the remainder of this section, we compare our analytic estimates of the
-effect on actions of a midplane offset against numerical calculations. A
+effect of a midplane offset on actions against numerical calculations. A
 numerical evaluation of the effect of velocity offsets on actions is deferred
@@
 \citep{2015ApJS..216...29B}. This potential includes a Hernquist bulge and
-nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
+nucleus \citep{1990ApJ...356..359H}, a Miyamoto--Nagai disk
 \citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
@@
 Galactocentric coordinates depends on the LSR). In this potential, we have
-that the circular velocity $v_{\text{circ}}$ is $231\,\kms$ on the Solar
+that the circular velocity $v_{\text{circ}}$ is $231\,\kms$ at the Solar
 circle.
@@
 
-We begin by considering the thick-disk orbit. Suppose an immortal observer who
+We begin with the thick-disk orbit. Consider an immortal observer who
 can measure the thick-disk star's phase-space position at many different times
-(and hence different orbital phases), but used a coordinate system in which
+(and hence different orbital phases). However, the immortal observer uses a coordinate system in which
 the midplane is systematically offset in height by $100\,\pc$ from its actual
@@
 the phase-space position to work with, in which case the determination of the
-orbital phase in $z$ or $R$ is degenerate with the degree of systematic offset
+orbital phase in $R$ or $z$ is degenerate with the degree of systematic offset
 in that coordinate (see Figure~\ref{fig:cartoon}).  In the following we
@@
 
-For a systematic offset in $x$ (or Galactic center distance), the
+For a systematic offset in $x$ (or distance to the Galactic center), the
 95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
@@
 values being within $\sim43\%$ of the true $J_z$. The $x$ offset induces
-uncertainties in $J_R$ and $J_{\phi}$ of $\sim21\%$ and $\sim3\%$,
-respectively. A $\sim3\%$ error in $J_z$ is also induced.}
+uncertainties in $J_R$, $J_{\phi}$, and $J_z$ of $\sim21\%$, $\sim3\%$, and $\sim3\%$.}
 \label{fig:one_orbit_wrong_ref}
@@
 corresponds to the distance from the true $J_z$ value to one of the modes of
-the error distribution.}
+the distribution of observed $J_z$ values.}
 \label{fig:Jz_hist}
@@
 about when the midplane error is $\gtrsim z_{\text{max}}$, where
-$z_{\text{max}}$ is the maximum height of the orbit (equivalent to $A_z$ in
+$z_{\text{max}}$ is the maximum height of the orbit, equivalent to $A_z$ in the epicyclic approximation (see
 Section~\ref{ssec:epi_action}).
@@
 include the estimation of the action errors derived under the epicyclic
-approximation from Equation~\eqref{eq:Ji_err_mosttime}. This Equation is
+approximation from Equation~\eqref{eq:Ji_err_mosttime}, with $\Delta v_{\phi}=0$, as dashed lines in the color of each orbit. This equation is
 relevant since during most of the orbit the star will be at maximum radial and
 vertical amplitude and we do not consider velocity offsets. Note that we
-consider an error in the $x$-coordinate $\Delta x$, which is not quite the
+consider an error in the $x$-coordinate $\Delta x$, which is not exactly the
 same as $\Delta R$. For observations of stars close to us, we have that
@@
 consider observations of the star throughout its entire orbit. This introduces
-a factor of $2/\pi$ when converting from $\Delta R$ to $\Delta x$, which we
+a factor of $2/\pi$ when converting from $\Delta x$ to $\Delta R$, which we
 derive in Appendix~\ref{app:deltax}.
@@
 different offsets in $z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
-(red, cyan, green respectively). As before, the error ($\Delta J_z$) is one
+(orange, teal, and green respectively). As before, the error ($\Delta J_z$) is one
 half the 95\uth{} minus 5\uth{} percentile of the distribution of observed
@@
 fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of $10, 50,
-100\,\pc$ (orange, cyan, and green, respectively) in
+100\,\pc$ (orange, teal, and green, respectively) in
 Figure~\ref{fig:dJz_fun_Jz}. For a thin disk-like orbit
@@
 Importantly, the location of \sgra{} may not be equivalent to
-the location of the ``dynamical Galactic center:'' the point in
+the location of the ``dynamical Galactic center,'' i.e., the point in
 three-dimensional space about which the stars in the solar neighborhood are
@@
 defining the coordinates of the Galactic center in B1950 coordinates as
-(17:42:26.6, -28:55:00) and the north Galactic pole as (12:49:00, +27:24:00).
+(17:42:26.6, -28:55:00) and the North Galactic pole as (12:49:00, +27:24:00).
 These two quantities, together with $R_0$, define the orientation of the
@@
 properties with observations of the Milky Way \citep{2016ApJ...827L..23W,
-2018MNRAS.481.4133G} The $\z = 0$ snapshots\footnote{In this work, to avoid
+2018MNRAS.481.4133G}. The $\z = 0$ snapshots\footnote{In this work, to avoid
 confusion with the vertical height $z$, we refer to cosmological redshift as
@@
 median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
-roughly solar positions (with cylindrical radii within $500\,\pc$ of
-$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi{},
+roughly solar positions (with galactocentric cylindrical radii within $0.5\,\kpc$ of
+$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4$, $12.5$, and $10.3\,\pc$ for \mi{},
 \mf{}, and \mm{}, respectively. These values are summarized in
@@
 dense, star-forming gas disk is thought to have a scale height of about
-$40\,\pc$ ($3\text{-}4$ times the cold gas softening length;
-\citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of
-about 300 pc (30 times the stellar softening length;
-\citealt{2008ApJ...673..864J}). We therefore expect that resolution effects
+$40\,\pc$, $\sim3\textup{--}4$ times the cold gas softening length
+\citep{2019ApJ...871..145A}. The thin stellar disk has a scale height of
+about $300\,\pc$, $\sim30$ times the stellar softening length
+\citep{2008ApJ...673..864J}. We therefore expect that resolution effects
 may still be affecting the scale heights of these components in the
@@
 Cosmological simulations of Milky Way-mass galaxies are not perfect
-representations of the true Milky Way in other ways as well\footnote{The failure of cosmological simulations to exactly reproduce the Milky Way is not entirely due to limitations of the numerical model.
-Candidate Milky Way-like galaxies are chosen solely on their mass and isolation, for which there are a wide variety of possible galaxies with qualitatively different properties.}, as discussed in
+representations of the true Milky Way in other ways as well,\footnote{The failure of cosmological simulations to exactly reproduce the Milky Way is not entirely due to limitations of the numerical model.
+Candidate Milky Way-like galaxies are chosen solely on their mass and isolation, for which there are a wide variety of possible galaxies with qualitatively different properties.} as discussed in
 \citet{2018arXiv180610564S}. For instance, the velocity structure of \mi{} is
@@
 Section 3 of \citet{2018arXiv180610564S} as our fiducial coordinate system for
-each galaxy. In short, the center of the galaxy is found iteratively, the
+each galaxy. In short, the center of the galaxy is found iteratively. The
 center of mass velocity is then determined by all star particles within
-$15\,\kpc$ of this center, and the galaxy is then rotated onto a
-principal-axis frame determined by stars younger than $1\,\Gyr$ inside of the
-fiducial solar radius $R_{\odot} = 8.2\,\kpc$, such that the disk plane is the
+$15\,\kpc$ of this center. The galaxy is then rotated onto a
+principal axis frame determined by stars younger than $1\,\Gyr$ inside of the
+fiducial solar radius $R_{0} = 8.2\,\kpc$, such that the disk plane is the
 $x$--$y$ plane.
@@
 convergence. The resulting median $z$ is taken to be what our observer would
-measure as the local Galactic midplane at each $\phi$. This procedure assumes
-perfect density estimation, and therefore perfect corrections for extinction,
-within the cylinder defining the ``solar neighborhood.'' Imperfect extinction
-correction is likely to increase the amplitude of the estimated fluctuations
-in $z$. To account for the effect of particle noise, we bootstrap-resample
+measure as the local Galactic midplane at each $\phi$.
+
+This procedure assumes perfect density estimation, and therefore perfect
+corrections for extinction within the cylinder defining the ``solar
+neighborhood.'' Imperfect extinction correction is likely to increase the
+amplitude of the estimated fluctuations in $z$.
+
+To account for the effect of particle noise, we bootstrap resample
 stars within a cylinder of stars with height $2\,\kpc$ and the same radius
@@
 each galaxy, using their local ``solar neighborhood'' (the cylinder defined
-above). The $1\,\sigma$ error from the bootstrap procedure is shown as the
+above).
+We consider $50$ equally spaced bins in azimuth, a sufficient number such that each cylinder shares no stars with its neighboring cylinder.
+The $1\,\sigma$ error from the bootstrap procedure is shown as the
 dashed-line error bars. The middle $90\%$ of midplane values across the Solar
@@
 has the thinnest ``thin disk'' of stars, but the largest stellar mass) the
-variation is more like $\pm 50\,\pc$.
+variation is closer to $\pm 50\,\pc$.
 
 We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
-but for the toy simulation of a Sagittarius encounter
+but for four succcessive timesteps of the toy simulation of a Sagittarius encounter
 \citep{2018MNRAS.481..286L}. Again we have subtracted a best fit curve of the
@@
 being $0.0013$, $0.0007$, $-0.63$, and $-1.00\,\text{rad}$ and $\text{C}$
-being $2.6$, $-7.8$, $-65$, and $-53\,\pc$ chronologically for each of the four timesteps shown. In the {\em upper} two panels, we see that the midplane is
+being $2.6$, $-7.8$, $-65$, and $-53\,\pc$ chronologically for each of the four timesteps shown.
+
+In the upper panels, we see that the midplane is
 relatively flat in the inner galaxy, but additional encounters drive strong
@@
 panels show the midplane as a function of azimuth before the first encounter
-near the Solar circle. As later encounters occur which bring the
+near the Solar circle at $t=2.0\,\Gyr$ and $t=4.0\,\Gyr$. As later encounters occur which bring the
 Sagittarius-like object closer to the Solar circle, we can see that stronger
@@
 the LSR is less pronounced than for the midplane, and since offsets in
-velocity only contribute to second order to $J_R$ and $J_z$ when a star is at
+velocity only contribute to second order to $\Delta J_R$ and $\Delta J_z$ when a star is at
 maximum amplitude in $R$ or $z$ (where the majority of the orbit is, see
@@
 by up to $\pm 100$ pc, as a natural consequence of the non-axisymmetry of the
-Galactic disk at small scales. While this is not in itself surprising or new,
-we also demonstrate that extending the coordinate system established by our
+Galactic disk at small scales.
+
+While this is not in itself surprising or new,
+we also demonstrated that extending the coordinate system established by our
 local midplane to a global axisymmetric coordinate system spanning the
@@
 the actions under this symmetry assumption, when starting from the present-day
-positions and velocities of stars as measured by, e.g., \emph{Gaia}. These
+positions and velocities of stars as measured by, e.g., \emph{Gaia}.
+
+These
 systematic errors are most important for stars on thin disk-like orbits, where
@@
 subdominant to the intrinsic variation of the midplane across large distances.
-On shorter scales, they are likely the dominant source of error.
+On shorter scales, they are likely the dominant source of error, and they are
+likely to be important for detailled modelling of kinematically cold stars.
 
@@
 \caption{The LSR as a function of azimuth at the
-fiducial Solar circle ($R_{\odot} = 8.2\,\kpc$). No best-fit subtraction is
+fiducial Solar circle ($R_0 = 8.2\,\kpc$). No best-fit subtraction is
 performed here as we did in the case of the midplane
-(Section~\ref{ssec:local_midplane}).}
+(Section~\ref{ssec:local_midplane}). Variations in $v_z$ are on the order of
+$\sim5\textup{--}10\,\kms$.}
 \label{fig:lsr_variations}

### 2019-05-04 paper/ms.tex
Remove statement about variations being less in sim
Clarify warp drive statement

@@
 punched through the galactic disk outside the solar circle and induced some
-warping. However, we expect that, as with the use of finite smoothing lengths,
-the degree of midplane variation in the simulations is likely a factor of a
-few lower than what could be expected in the Milky Way.
+warping.
 
@@
 $z$-coordinate is perpendicular to the disk plane, we place our imaginary
-observer at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth
-between $0<\phi<2\pi$.\footnote{The immortal observer also has warp drive.} At
+observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and vary the azimuth
+between $0<\phi<2\pi$.\footnote{The immortal observer also has warp drive, allowing them to travel around the galaxy much more quickly than its dynamical timescale.} At
 each value of $\phi$ we then compute the median $z$ for stars within a

### 2019-05-04 paper/ms.tex
Add sentence on link between FIRE and Sgr
Update orbits table headings

@@
 work. For the last two columns: $z_{\text{max}}$ is the maximum height of the
-orbit while $A_R$ is the magnitude of the radial excursions of the orbit,
-computed as one half the maximum radius minus the minimum
-radius.\label{tab:orbits}}
+orbit while $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of the radial excursions of the orbit.
+In the epicyclic approximation, $A_z = z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ \label{tab:orbits}}
 \tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
@@
 \colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
-\colhead{$A_R$}\\ \colhead{ } &
+\colhead{$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$}\\ \colhead{ } &
 \colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
@@
 \citet{2018MNRAS.481..286L}.
+The lower right panel, which shows the galaxy after some relaxation has occured is qualitatively similar to the midplane variations we saw in the FIRE simulations (Figure~\ref{fig:midplane}), evidence that they are at least partially driven by mergers.
 

### 2019-05-04 paper/ms.tex
Small changes throughout

@@
 standard of rest induces similar issues. The variation of the midplane must be
-taken into account when performing dynamical modelling across the large
+taken into account when performing dynamical modeling across the large
 regions of the disk accessible to {\em Gaia} and future missions.
@@
 \end{center}
-\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
+\caption{A histogram of the observed values in $J_z$ for the thick-disk (top panel) and thin-disk (bottom panel) orbits
 assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
@@
 \sgra{}, the distance from the Sun to
-\sgra{} can be precisely measured \citep{2009ApJ...692.1075G,
+\sgra{}, referred to as $R_0$, can be precisely measured \citep{2009ApJ...692.1075G,
 2018AA...615L..15G}, with a recent measurement using near-infrared
@@
 The azimuthal component ($V_{\odot}$) is more difficult to measure, but can be
-modelled using the asymmetric drift relation \citep{2008gady.book.....B}. The
+modeled using the asymmetric drift relation \citep{2008gady.book.....B}. The
 values of the components of the LSR are usually taken from
@@
 Considerable effort has been placed on each of these measurements, but
-uncertainties remain, and detailed dynamical modelling across the disk ---
+uncertainties remain, and detailed dynamical modeling across the disk ---
 particularly for dynamically cold stars --- may have to take them into
@@
 $6\times10^{10}\,\Msun$, a Hernquist bulge of mass $10^{10}\,\Msun$ and scale
-length $0.7\,\kpc$ \citep{1990ApJ...356..359H}. Sagittarius is modelled as a
+length $0.7\,\kpc$ \citep{1990ApJ...356..359H}. Sagittarius is modeled as a
 dark matter Hernquist sphere of mass \Gus{something} and scale length
@@
 
-\item Dynamical modelling across large regions of the disk, over which the
+\item Dynamical modeling across large regions of the disk, over which the
 midplane location varies by more than the limits discussed above, is

### 2019-05-04 paper/ms.tex
Add discussion section on coordinate system offsets

@@
 
-Importantly, the location of \sgra{} may not be equivalent to
-the location of the ``dynamical Galactic center,'' i.e., the point in
-three-dimensional space about which the stars in the solar neighborhood are
-orbiting. This assumption, although sensible and frequently made, has not yet been
-justified. It is tested in any construction of a  dynamical model where the
-distance from the Sun to the Galactic center, $R_0$, is a free parameter. For
-example, \citet{2015ApJ...803...80K} measured $R_0$ while modelling the
-dynamics of the stream Palomar 5. Many other dynamical measurements of $R_0$
-have been made (\citealt{2016ARAA..54..529B} summarize many pre-{\em Gaia}
-results) but none have yet achieved a precision comparable to that of the
-distance to \sgra{}.
-
 \subsection{Galactic Orientation}
@@
 
-Galactocentric coordinate system offsets can also be generated by errors in
-the parameters of the Galactocentric coordinate system. However, such
-parameters have been measured to high enough precision that they are likely
-subdominant to the intrinsic variation of the midplane across large distances.
-On shorter scales, they are likely the dominant source of error, and they are
-likely to be important for detailled modelling of kinematically cold stars.
-
 This finding has many implications for the study of the dynamics of stars in
@@
 we saw in both sets of simulations.
+
+\subsection{Coordinate System Offsets}\label{ssec:coord_off}
+In Section~\ref{sec:mes_err} we discussed the current observational efforts to
+measure the parameters of the Galactocentric coordinate system. We noted that
+the solar height is discrepant between gas-based and stellar-based
+determinations by $\sim5\,\pc$. While the effect is small and thus only likely
+to dominate the intrinsic midplane variations on small scales, it may be
+important for detailed modeling of kinematically cold stars where a $5\,\pc$
+offset is relevant.
+
+We also noted that the distance from the Sun to \sgra{} has been recently
+measured to a very low uncertainty of $\sim0.3\%$ or $\pm25\,\pc$
+\citep{2019arXiv190405721A}. However, the location of \sgra{} may not be
+equivalent to the location of the ``dynamical Galactic center,'' i.e., the
+point in three-dimensional space about which the stars in the solar
+neighborhood are orbiting. This assumption, although sensible and frequently
+made, has not yet been justified.
+
+If the dynamical Galactic center is offset from \sgra{} by $100\,\pc$, only a
+$1.2\%$ difference, then this induces a $21\%$ error in $J_R$ (see
+Section~\ref{ssec:quant}). The reason such a large error in $J_R$ can be
+generated by a small error in $R_0$ can be understood from the epicyclic
+approximation, which states that $\Delta J_R/J_R = 2\Delta R/A_R$. The fractional
+error in $J_R$ is related to the error in $R_0$ as a fraction of the {\em
+radial amplitude} of the orbit, which is much smaller than $R_0$
+($\sim1.2\,\kpc$ for the thin-disk and thick-disk orbits we considered). This
+also implies the very precise $0.3\%$ measurement of $R_0$ still translates to
+a $\sim4\%$ uncertainty in $J_R$.
+
+The assumption that the dynamical Galactic center and \sgra{} are colocated is
+tested in any construction of a dynamical model where $R_0$ is a free
+parameter. For example, \citet{2015ApJ...803...80K} measured $R_0$ while
+modeling the dynamics of the stream Palomar 5. Many other dynamical
+measurements of $R_0$ have been made (\citealt{2016ARAA..54..529B} summarize
+many pre-{\em Gaia} results), but none have yet achieved a precision
+comparable to that of the distance to \sgra{}.
+
+We did not consider in this text the effect of the {\em angular position} of
+the dynamical Galactic center being offset with \sgra{}.
 

### 2019-05-05 paper/ms.tex
Add conclusion bullet point on coordinate system offsets
Fix italics throughout

@@
 \title{The Galactic Midplane Is Not a Plane: Implications for Dynamical
-Analysis with {\em Gaia} Data and Beyond}
+Analysis with \textit{Gaia} Data and Beyond}
 
@@
 invariant and thus good labels of a star's orbit. Such analyses are often used
-to interpret data from the {\em Gaia} mission. However, inaccurate
+to interpret data from the \textit{Gaia} mission. However, inaccurate
 measurements of the phase space position and uncertainties in the assumed
@@
 taken into account when performing dynamical modeling across the large
-regions of the disk accessible to {\em Gaia} and future missions.
+regions of the disk accessible to \textit{Gaia} and future missions.
 
@@
 Our understanding of the Milky Way is currently undergoing a revolution as a
-result of {\em Gaia} Data Release 2 (DR2).  Recent major discoveries include
+result of \textit{Gaia} Data Release 2 (DR2).  Recent major discoveries include
 the remnants of a major merger \citep{2018ApJ...860L..11K,
@@
 With the advent of 6D phase-space measurements over a relatively large
-($\gtrsim 2$ kpc) volume from the {\em Gaia} satellite, the study of stars'
+($\gtrsim 2$ kpc) volume from the \textit{Gaia} satellite, the study of stars'
 actions has gained new popularity. One reason is dimensionality reduction---an
@@
 function of their actions and independent of the conjugate angles, allowing
-investigation of the relationship between {\em orbital} properties of stars
+investigation of the relationship between \emph{orbital} properties of stars
 and other intrinsic, and at least partially invariant, properties such as age
@@
 authors have used actions to study stellar scattering in the Milky Way disk
-using the improved astrometry of {\em Gaia} DR2 and various age catalogues
+using the improved astrometry of \textit{Gaia} DR2 and various age catalogues
 \citep{2018ApJ...867...31B, 2018arXiv180803278T}. Actions have also been used
@@
 axisymmetric; with the vast improvement in the quality of phase-space
-measurements due to {\em Gaia}, this assumption appears increasingly
+measurements due to \textit{Gaia}, this assumption appears increasingly
 inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this assumption were
@@
 We quantify how a systematic error in the Galactocentric coordinate system
-induces phase-dependence in the {\em observed} actions. We will consider three
+induces phase-dependence in the \emph{observed} actions. We will consider three
 orbits in the model potential described in Section~\ref{ssec:action_comp} that
@@
 We also perform the same procedure but assume a $100\,\pc$ offset in the $x$
-component (i.e. subtracting the vector $(100, 0, 0)\,\pc$) in the {\em lower}
+component (i.e. subtracting the vector $(100, 0, 0)\,\pc$) in the lower
 panels. This is equivalent to a systematic offset in the determination of the
@@
 The Solar height is usually taken to be $\sim 25\,\pc$
-\citep{2001ApJ...553..184C}, with a more recent measurement from {\em Gaia}
+\citep{2001ApJ...553..184C}, with a more recent measurement from \textit{Gaia}
 DR2 placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another
@@
 $5\,\kms$  on $\sim\,\kpc$ scales looking toward the Galactic anticenter. Work
-by the {\em Gaia} collaboration confirmed these preliminary results on the
+by the \textit{Gaia} collaboration confirmed these preliminary results on the
 velocity and spatial scales of these oscillation with clear spatial maps made
@@
 measurements of $R_0$ have been made (\citealt{2016ARAA..54..529B} summarize
-many pre-{\em Gaia} results), but none have yet achieved a precision
+many pre-\textit{Gaia} results), but none have yet achieved a precision
 comparable to that of the distance to \sgra{}.
 
-We did not consider in this text the effect of the {\em angular position} of
+We did not consider in this text the effect of the \emph{angular position} of
 the dynamical Galactic center being offset with \sgra{}.
@@
 coordinate system offset is very non-Gaussian. The distribution is bimodal
-with {\em neither mode at null}. As a result, error propagation of coordinate
+with \emph{neither mode at null}. As a result, error propagation of coordinate
 system offsets is complex when considering actions, and is likely to
@@
 susceptible to this type of systematic error, since the assumption that our
-local Galactic midplane is the global Galactic midplane is not true {\em a
-priori}.
+local Galactic midplane is the global Galactic midplane is not true a
+priori.
 
@@
 \item Assuming a vertical velocity variation of the Milky Way of
-$\sim5\,\kms$, consistent with recent results from {\em Gaia} DR2
+$\sim5\,\kms$, consistent with recent results from \textit{Gaia} DR2
 \citep{2018A&A...616A..11G, 2019arXiv190209569F} and our results from the FIRE
@@
 analysis with the vertical frequencies of disk-like orbits.
+
+\item While inaccuracies in the parameters of the currently adopted
+Galactocentric coordinate system are likely to be unimportant for most
+applications, we discussed the importance of testing the assumption that the
+dynamical Galactic center is colocated with \sgra{}, which we mentioned how to
+do.
 
@@
 Even without a high-precision definition of the global Galactic midplane,
-current observations from {\em Gaia} should soon permit a measurement of the
+current observations from \textit{Gaia} should soon permit a measurement of the
 real azimuthal dependence of the midplane location. For some applications,

### 2019-05-05 paper/ms.tex
Add info about orbits in Fig 5

@@
 orbit while $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of the radial excursions of the orbit.
-In the epicyclic approximation, $A_z = z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ \label{tab:orbits}}
+In the epicyclic approximation, $A_z = z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$.
+\label{tab:orbits}}
 \tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
@@
 \caption{The fractional error in $J_z$ as a function of $J_z$ for a few
-different offsets in $z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
-(orange, teal, and green respectively). As before, the error ($\Delta J_z$) is one
-half the 95\uth{} minus 5\uth{} percentile of the distribution of observed
+different offsets in $z$. All orbits have the same initial position of $(8, 0, 0)\,\kpc$ and velocity $(-190, 0, v_z)\,\kms$, where we vary $v_z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
+(orange, teal, and green respectively). As before, the error ($\Delta J_z$) is
+one half the 95\uth{} minus 5\uth{} percentile of the distribution of observed
 $J_z$ values over the course of the orbit. There are large errors for
 thin-disk like orbits ($J_z \sim 0.5\,\actunit$), even for a small midplane
-offset of $10\,\pc$. }
+offset of $10\,\pc$. As dashed lines in each color we also plot the prediction
+for $\Delta J_z/J_z$ from the epicyclic approximation
+(Equation~\eqref{eq:Ji_err_mosttime}), which shows excellent agreement with
+the numerically computed values.}
 \label{fig:dJz_fun_Jz}
@@
 100\,\pc$ (orange, teal, and green, respectively) in
-Figure~\ref{fig:dJz_fun_Jz}. For a thin disk-like orbit
+Figure~\ref{fig:dJz_fun_Jz}. For each orbit, we set the initial position to be $(8,0,0)\,\kpc$ and the initial velocity to be $(-190, 0, v_z)\,\kms$, where we vary $v_z$. For a thin disk-like orbit
 ($J_z\sim0.5\,\actunit$), even a $10\,\pc$ offset in $z$ is enough to induce a
@@
 We consider the variations of the LSR as a function of azimuth at the fiducial
-solar circle ($R_{\odot} = 8.2\,\kpc$) in Figure~\ref{fig:lsr_variations}. At
+solar circle ($R_{0} = 8.2\,\kpc$) in Figure~\ref{fig:lsr_variations}. At
 each azimuth, $\phi$, we take the median velocity in cylindrical coordinates

### 2019-05-05 paper/ms.tex
Move figures around in paper

@@
 
-In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
-system errors have on the measured actions. We then discuss two potential
-sources of these coordinate system offsets in
-Sections~\ref{sec:mes_err}~and~\ref{sec:local_fire}. In
-Section~\ref{sec:mes_err} we discuss the potential errors in measuring the
-parameters of the Galactocentric coordinate system. In
-Section~\ref{sec:local_fire} we examine the azimuthal variations of the
-midplane itself in two sets of simulations: cosmological, hydrodynamical,
-zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic
-Environments (FIRE)
-collaboration\footnote{\url{https://fire.northwestern.edu}}
-\citep{2014MNRAS.445..581H, 2016ApJ...827L..23W, 2018MNRAS.480..800H}, and a controlled N-body
-simulation of a Sagittarius encounter with a galaxy otherwise tailored to the
-stellar mass, scale length, and scale height of the Milky Way
-\citep{2018MNRAS.481..286L}.
-
-In Section~\ref{sec:discussion}, we discuss the implications of midplane
-variations, and the resulting systematic uncertainty in the vertical action,
-for action-space analyses. We also estimate the expected midplane variations of the Milky Way based on the observed velocity variations. We
-summarize our main results and conclude in Section~\ref{sec:conclusion}.
-
-\section{Motivation} \label{sec:ref_frame}
-We first demonstrate the significance  to action computations of a systematic
-offset in the determination of the Galactic midplane or distance to the
-Galactic center. We will find that such offsets are especially important for
-disk-like orbits. The consequences we explore here may also arise from various
-other systematic errors. For instance, the axisymmetric Galactic potential
-model used in many works to compute actions may not be a good description of
-the true potential --- or the parameters used may yield a potential that is
-systematically incorrect outside an original fitted region. In this work, we
-assume that the Galaxy is perfectly described by our model axisymmetric
-potential, and simply explore the consequences of offsets in the determination
-of the Galactocentric coordinate system.
-
-\subsection{Effect of Midplane Offset} \label{ssec:cartoon}
-We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
-inaccurate or erroneous determination of the midplane (and hence the
-zero-point of the $z$ coordinate) leads to a phase-dependence of the actions
-calculated from an orbit specified by a phase-space point and an assumed
-potential model. The $y$-axis corresponds to the vertical height of the orbit
-and the $x$-axis shows orbital phase. The solid gray curve indicates the
-``true'' orbit of the star; that is, its orbit as it oscillates around the
-true midplane. The dashed gray line is the midplane location in the potential
-model used by an observer to integrate the orbit of the star and hence estimate its actions,
-which is offset from the true midplane of the star's orbit. The model
-potential is otherwise identical to the one in which the star is actually
-moving.
-
-Now suppose this observer makes a measurement of the star's position and
-velocity at the teal point or the orange point (i.e. at two different orbital
-phases). Then the teal and orange curves correspond to the orbits that the
-observer would compute for each point based on the potential model with the
-offset midplane. In action space, this would correspond to a different value
-of $J_z$ for the teal and orange points. In this way, assuming the wrong
-coordinate system induces a phase-dependence in the actions estimated for the
-star, which in the correct potential (in this example, the one with the
-correct midplane) should be phase-independent.
-
-\begin{figure*}
+\begin{figure*}[ht!]
 \begin{center}
@@
 
+In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
+system errors have on the measured actions. We then discuss two potential
+sources of these coordinate system offsets in
+Sections~\ref{sec:mes_err}~and~\ref{sec:local_fire}. In
+Section~\ref{sec:mes_err} we discuss the potential errors in measuring the
+parameters of the Galactocentric coordinate system. In
+Section~\ref{sec:local_fire} we examine the azimuthal variations of the
+midplane itself in two sets of simulations: cosmological, hydrodynamical,
+zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic
+Environments (FIRE)
+collaboration\footnote{\url{https://fire.northwestern.edu}}
+\citep{2014MNRAS.445..581H, 2016ApJ...827L..23W, 2018MNRAS.480..800H}, and a controlled N-body
+simulation of a Sagittarius encounter with a galaxy otherwise tailored to the
+stellar mass, scale length, and scale height of the Milky Way
+\citep{2018MNRAS.481..286L}.
+
+In Section~\ref{sec:discussion}, we discuss the implications of midplane
+variations, and the resulting systematic uncertainty in the vertical action,
+for action-space analyses. We also estimate the expected midplane variations of the Milky Way based on the observed velocity variations. We
+summarize our main results and conclude in Section~\ref{sec:conclusion}.
+
+\section{Motivation} \label{sec:ref_frame}
+We first demonstrate the significance  to action computations of a systematic
+offset in the determination of the Galactic midplane or distance to the
+Galactic center. We will find that such offsets are especially important for
+disk-like orbits. The consequences we explore here may also arise from various
+other systematic errors. For instance, the axisymmetric Galactic potential
+model used in many works to compute actions may not be a good description of
+the true potential --- or the parameters used may yield a potential that is
+systematically incorrect outside an original fitted region. In this work, we
+assume that the Galaxy is perfectly described by our model axisymmetric
+potential, and simply explore the consequences of offsets in the determination
+of the Galactocentric coordinate system.
+
+\subsection{Effect of Midplane Offset} \label{ssec:cartoon}
+We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
+inaccurate or erroneous determination of the midplane (and hence the
+zero-point of the $z$ coordinate) leads to a phase-dependence of the actions
+calculated from an orbit specified by a phase-space point and an assumed
+potential model. The $y$-axis corresponds to the vertical height of the orbit
+and the $x$-axis shows orbital phase. The solid gray curve indicates the
+``true'' orbit of the star; that is, its orbit as it oscillates around the
+true midplane. The dashed gray line is the midplane location in the potential
+model used by an observer to integrate the orbit of the star and hence estimate its actions,
+which is offset from the true midplane of the star's orbit. The model
+potential is otherwise identical to the one in which the star is actually
+moving.
+
+Now suppose this observer makes a measurement of the star's position and
+velocity at the teal point or the orange point (i.e. at two different orbital
+phases). Then the teal and orange curves correspond to the orbits that the
+observer would compute for each point based on the potential model with the
+offset midplane. In action space, this would correspond to a different value
+of $J_z$ for the teal and orange points. In this way, assuming the wrong
+coordinate system induces a phase-dependence in the actions estimated for the
+star, which in the correct potential (in this example, the one with the
+correct midplane) should be phase-independent.
+
 \subsection{Epicyclic Approximation} \label{ssec:epi_action}
+
+\begin{deluxetable*}{cccccccc}
+\tablecaption{Description and names of the three orbits considered in this
+work. For the last two columns: $z_{\text{max}}$ is the maximum height of the
+orbit while $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of the radial excursions of the orbit.
+In the epicyclic approximation, $A_z = z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$.
+\label{tab:orbits}}
+\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
+velocity} & \colhead{$J_R$} &
+\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
+\colhead{$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$}\\ \colhead{ } &
+\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
+\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
+\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
+\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$}}
+\startdata 
+thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 & 1.29 \\
+thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 & 1.19 \\ 
+halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16 & 2.34
+\enddata
+\end{deluxetable*}
+
 Before turning to numerical methods, we first write down analytically the
@@
 \subsection{Quantification of the Midplane Effect} \label{ssec:quant}
-We quantify how a systematic error in the Galactocentric coordinate system
-induces phase-dependence in the \emph{observed} actions. We will consider three
-orbits in the model potential described in Section~\ref{ssec:action_comp} that
-are typical of stars in the stellar thin disk, stellar thick disk, and the
-stellar halo. We summarize their initial positions in phase space and the
-actions computed by integrating their orbits in the correct potential in
-Table~\ref{tab:orbits}. Each orbit, integrated without systematic coordinate
-errors, is plotted in Appendix~\ref{app:orbits}. We will refer to these orbits
-by their names (thin-disk, thick-disk, halo) henceforth.
-
-We begin with the thick-disk orbit. Consider an immortal observer who
-can measure the thick-disk star's phase-space position at many different times
-(and hence different orbital phases). However, the immortal observer uses a coordinate system in which
-the midplane is systematically offset in height by $100\,\pc$ from its actual
-location --- i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position
-in the orbit. This corresponds to the immortal observer physically located at
-e.g. the position $(8, 0, 0)\,\kpc$ in the coordinate system of the true
-potential, but erroneously thinking they are located at $(8, 0, 0.1)\,\kpc$.
-
-\begin{deluxetable*}{cccccccc}
-\tablecaption{Description and names of the three orbits considered in this
-work. For the last two columns: $z_{\text{max}}$ is the maximum height of the
-orbit while $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of the radial excursions of the orbit.
-In the epicyclic approximation, $A_z = z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$.
-\label{tab:orbits}}
-\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
-velocity} & \colhead{$J_R$} &
-\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
-\colhead{$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$}\\ \colhead{ } &
-\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
-\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$}}
-\startdata 
-thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 & 1.29 \\
-thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 & 1.19 \\ 
-halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16 & 2.34
-\enddata
-\end{deluxetable*}
-
-We consider that the immortal observer makes a measurement every $\Myr$.
-At each measurement, we compute the actions as outlined above using the true potential but by specifiying the star's starting position using the systematically offset coordinate system.
-Essentially we are shifting the
-integrated orbit and then reintegrating at each point along the shifted orbit,
-where the shifted orbit is the observed orbit. The erroneous computed actions
-for each phase-space starting point are shown for the first $\Gyr$ of the
-orbit in the upper panels
-Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
-scheme fails and very large actions are reported by \texttt{gala}---we perform
-a $5\sigma$ clip on each action to exclude such orbits, but this only excludes
-a total of $5$ orbits out of the $1000$ considered for
-Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
-vast majority of orbits are computed properly.}
-
-We also perform the same procedure but assume a $100\,\pc$ offset in the $x$
-component (i.e. subtracting the vector $(100, 0, 0)\,\pc$) in the lower
-panels. This is equivalent to a systematic offset in the determination of the
-distance from the Sun to the Galactic center.
-
-Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
-offset coordinate systems depend on the time (i.e. orbital phase) at which the
-star's phase-space position is observed. To the immortal observer this would
-appear as time-dependence in the actions, which appear to oscillate around
-their value using the correct coordinate system even though in this example
-the observer is using the correctly constructed best-fit static, axisymmetric
-potential. The relative size of the phase variation in each action depends on
-the direction of the systematic offset as well as the true values of the
-actions (i.e. the type of orbit). In reality we will have one measurement of
-the phase-space position to work with, in which case the determination of the
-orbital phase in $R$ or $z$ is degenerate with the degree of systematic offset
-in that coordinate (see Figure~\ref{fig:cartoon}).  In the following we
-therefore quote percentile ranges for the possible values computed for each
-action as a proxy for the effect of these systematic errors in the coordinate
-system.
-
-For a systematic offset in $z$ (upper panels), the
-95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
-$6.2\,\actunit$ for $J_R$ and $J_z$, respectively. These are fractional errors
-of $5.7\,\%$ and $86\,\%$, respectively. The error induced in $J_{\phi}$ is
-negligible, as expected since $J_{\phi}$ only depends on the $x$- and
-$y$-components of the position and velocity of the stars.\footnote{In practice,
-however, $J_{\phi}$ is computed as part of the torus-fitting method.} It is worth
-pointing out that a $100\,\pc$ error in an orbit with
-$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
---- induced an $86\%$ error in the computation of $J_z$.
-
-For a systematic offset in $x$ (or distance to the Galactic center), the
-95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
-$47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively.
-These are fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively ---
-despite only a $1.2\%$ error in the distance to the Galactic center.
-
-\begin{figure*}
+
+\begin{figure*}[ht!]
 \begin{center}
@@
 
+We quantify how a systematic error in the Galactocentric coordinate system
+induces phase-dependence in the \emph{observed} actions. We will consider three
+orbits in the model potential described in Section~\ref{ssec:action_comp} that
+are typical of stars in the stellar thin disk, stellar thick disk, and the
+stellar halo. We summarize their initial positions in phase space and the
+actions computed by integrating their orbits in the correct potential in
+Table~\ref{tab:orbits}. Each orbit, integrated without systematic coordinate
+errors, is plotted in Appendix~\ref{app:orbits}. We will refer to these orbits
+by their names (thin-disk, thick-disk, halo) henceforth.
+
+We begin with the thick-disk orbit. Consider an immortal observer who
+can measure the thick-disk star's phase-space position at many different times
+(and hence different orbital phases). However, the immortal observer uses a coordinate system in which
+the midplane is systematically offset in height by $100\,\pc$ from its actual
+location --- i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position
+in the orbit. This corresponds to the immortal observer physically located at
+e.g. the position $(8, 0, 0)\,\kpc$ in the coordinate system of the true
+potential, but erroneously thinking they are located at $(8, 0, 0.1)\,\kpc$.
+
+We consider that the immortal observer makes a measurement every $\Myr$.
+At each measurement, we compute the actions as outlined above using the true potential but by specifiying the star's starting position using the systematically offset coordinate system.
+Essentially we are shifting the
+integrated orbit and then reintegrating at each point along the shifted orbit,
+where the shifted orbit is the observed orbit. The erroneous computed actions
+for each phase-space starting point are shown for the first $\Gyr$ of the
+orbit in the upper panels
+Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
+scheme fails and very large actions are reported by \texttt{gala}---we perform
+a $5\sigma$ clip on each action to exclude such orbits, but this only excludes
+a total of $5$ orbits out of the $1000$ considered for
+Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
+vast majority of orbits are computed properly.}
+
+We also perform the same procedure but assume a $100\,\pc$ offset in the $x$
+component (i.e. subtracting the vector $(100, 0, 0)\,\pc$) in the lower
+panels. This is equivalent to a systematic offset in the determination of the
+distance from the Sun to the Galactic center.
+
+Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
+offset coordinate systems depend on the time (i.e. orbital phase) at which the
+star's phase-space position is observed. To the immortal observer this would
+appear as time-dependence in the actions, which appear to oscillate around
+their value using the correct coordinate system even though in this example
+the observer is using the correctly constructed best-fit static, axisymmetric
+potential. The relative size of the phase variation in each action depends on
+the direction of the systematic offset as well as the true values of the
+actions (i.e. the type of orbit). In reality we will have one measurement of
+the phase-space position to work with, in which case the determination of the
+orbital phase in $R$ or $z$ is degenerate with the degree of systematic offset
+in that coordinate (see Figure~\ref{fig:cartoon}).  In the following we
+therefore quote percentile ranges for the possible values computed for each
+action as a proxy for the effect of these systematic errors in the coordinate
+system.
+
+For a systematic offset in $z$ (upper panels), the
+95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
+$6.2\,\actunit$ for $J_R$ and $J_z$, respectively. These are fractional errors
+of $5.7\,\%$ and $86\,\%$, respectively. The error induced in $J_{\phi}$ is
+negligible, as expected since $J_{\phi}$ only depends on the $x$- and
+$y$-components of the position and velocity of the stars.\footnote{In practice,
+however, $J_{\phi}$ is computed as part of the torus-fitting method.} It is worth
+pointing out that a $100\,\pc$ error in an orbit with
+$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
+--- induced an $86\%$ error in the computation of $J_z$.
+
+For a systematic offset in $x$ (or distance to the Galactic center), the
+95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
+$47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively.
+These are fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively ---
+despite only a $1.2\%$ error in the distance to the Galactic center.
+
 In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
@@
 
-We now describe the intuition behind the shape of Figure~\ref{fig:Jz_hist}.
-Consider first the thick-disk orbit (top panel), where the offset in $z$
-is much less than $z_{\text{max}}$ of the orbit. The peaks in the distribution
-correspond to the turning points of the orbit (or points of maximum vertical
-amplitude), where $v_z \sim 0$ and where the star is on most of it orbit.
-This is why the distribution is peaked on these values. For the thin-disk
-orbit (bottom panel), the offset in $z$ is comparable to
-$z_{\text{max}}$. Now, there will be some points in the orbit where $v_z = 0$
-and $z=0$ (in the observed, erroneous coordinate system). At these points, the
-computed $J_z$ will vanish. The asymmetry and systematic offset then comes
-about because of the constraint that $J_z \geq 0$.\footnote{This is similar to
-arguments in cosmology for why gravity produces non-Gaussianity in the density
-field, since the density cannot become negative but it can grow arbitrarily
-large.}
-
-Empirically describing the error distribution shown in
-Figure~\ref{fig:Jz_hist} is difficult. Of course, Gaussian summary statistics
-are not adequate. We therefore elect to measure this error by computing one
-half the 95\uth{} percentile minus the 5\uth{} percentile of the distribution
-of observed action values. We refer to this quantity as $\Delta J_i$ for each
-action value. We plot this quantity in Figure~\ref{fig:Jz_hist} as a
-horizontal arrow anchored on the true action value. Because of the bimodality
-of the error distribution, this quantity roughly measures the distance from
-the true action value to the peak of one of the modes. Furthermore, this
-bimodality also implies that $\Delta J_i$ is not very sensitive to the exact
-percentiles used. This summary statistic does not reflect the bias induced
-when the midplane error is $\gtrsim z_{\text{max}}$.
-
-We now repeat the same procedure as in Figure~\ref{fig:one_orbit_wrong_ref}
-but for systematic offsets between $0$ and $500\,\pc$ in the $z$ and $x$
-components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report one half the
-95\uth{} minus 5\uth{} percentile divided by the true action value ($\Delta
-J_i/J_i$) for the three different fiducial orbits in Table~\ref{tab:orbits}.
-The thick-disk (\thickcolor) orbit is the one from
-Figure~\ref{fig:one_orbit_wrong_ref}, but we also now consider the effect on
-the action determined for the thin-disk (\thincolor) and halo (\halocolor)
-orbits.
-
-The upper row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
-in each action for an offset in the $z$-component (i.e. the Galactic
-midplane). In the lower row we consider offsets in the $x$ component
-(i.e. the Solar radius). The left, center, and right columns
-show the fractional spread in the values computed for $J_z$, $J_{\phi}$, and
-$J_R$, respectively. We compute the fractional spread $\Delta J_i/J_i$, where
-$\Delta J_i$ is defined in the previous paragraph and $J_i$ is the action
-value computed in the true coordinate system, over the course of the first
-$\Gyr$.
-
-In the upper middle panel of Figure~\ref{fig:many_orbit_wrong_ref},
-there is essentially no spread in the determination of $J_{\phi}$. This is
-expected since $J_{\phi}$ is independent of $z$ and is
-thus unaffected by offsets in $z$, as discussed earlier. Indeed, the result we found in
-Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
-orbit types. This is also a demonstration of the robustness of the orbit
-integration and action calculation methods we use.
-
-The upper right panel of Figure~\ref{fig:many_orbit_wrong_ref} shows
-that the fractional error in $J_z$ is more exaggerated for more planar
-(disk-like) orbits. For the thin-disk orbit, a systematic offset of
-$\sim15\,\pc$ in the $z$ coordinate already results in $25\%$ deviations in
-the actions. The required offset for $25\%$ deviation is $\sim120\,\pc$ for
-the thick-disk orbit. The halo orbit is relatively resistant to errors in the
-midplane, with only $\sim15\%$ error in $J_z$ out to a midplane offset of
-$500\,\pc$.
-
-For the offset in the Solar radius (lower panels), the error is largest
-for $J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
-deviations in $J_z$. In the lower middle and lower right panels
-all three lines nearly overlap.
-
-In each panel of Figure~\ref{fig:many_orbit_wrong_ref}, where relevant, we
-include the estimation of the action errors derived under the epicyclic
-approximation from Equation~\eqref{eq:Ji_err_mosttime}, with $\Delta v_{\phi}=0$, as dashed lines in the color of each orbit. This equation is
-relevant since during most of the orbit the star will be at maximum radial and
-vertical amplitude and we do not consider velocity offsets. Note that we
-consider an error in the $x$-coordinate $\Delta x$, which is not exactly the
-same as $\Delta R$. For observations of stars close to us, we have that
-$\Delta x \sim \Delta R$, but for the experiment performed in this Section we
-consider observations of the star throughout its entire orbit. This introduces
-a factor of $2/\pi$ when converting from $\Delta x$ to $\Delta R$, which we
-derive in Appendix~\ref{app:deltax}.
-
-The epicyclic approximation is a good predictor of $\Delta J_z$, even for halo
-orbits. It performs similarly for $\Delta J_R$, now underpredicting for halo orbits
-and slightly overpredicting for thin-disk orbits. Note that for the particular
-orbits we chose, the thin-disk orbit has slightly larger $A_R$ than the
-thick-disk orbit, and so we expect the epicyclic approximation to perform
-slightly worse. The epicyclic approximation underpredicts $\Delta J_{\phi}$ for all
-orbits.
-
-\begin{figure*}
+\begin{figure*}[htb!]
 \begin{center}
@@
 \end{figure*}
+
+We now describe the intuition behind the shape of Figure~\ref{fig:Jz_hist}.
+Consider first the thick-disk orbit (top panel), where the offset in $z$
+is much less than $z_{\text{max}}$ of the orbit. The peaks in the distribution
+correspond to the turning points of the orbit (or points of maximum vertical
+amplitude), where $v_z \sim 0$ and where the star is on most of it orbit.
+This is why the distribution is peaked on these values. For the thin-disk
+orbit (bottom panel), the offset in $z$ is comparable to
+$z_{\text{max}}$. Now, there will be some points in the orbit where $v_z = 0$
+and $z=0$ (in the observed, erroneous coordinate system). At these points, the
+computed $J_z$ will vanish. The asymmetry and systematic offset then comes
+about because of the constraint that $J_z \geq 0$.\footnote{This is similar to
+arguments in cosmology for why gravity produces non-Gaussianity in the density
+field, since the density cannot become negative but it can grow arbitrarily
+large.}
+
+Empirically describing the error distribution shown in
+Figure~\ref{fig:Jz_hist} is difficult. Of course, Gaussian summary statistics
+are not adequate. We therefore elect to measure this error by computing one
+half the 95\uth{} percentile minus the 5\uth{} percentile of the distribution
+of observed action values. We refer to this quantity as $\Delta J_i$ for each
+action value. We plot this quantity in Figure~\ref{fig:Jz_hist} as a
+horizontal arrow anchored on the true action value. Because of the bimodality
+of the error distribution, this quantity roughly measures the distance from
+the true action value to the peak of one of the modes. Furthermore, this
+bimodality also implies that $\Delta J_i$ is not very sensitive to the exact
+percentiles used. This summary statistic does not reflect the bias induced
+when the midplane error is $\gtrsim z_{\text{max}}$.
+
+We now repeat the same procedure as in Figure~\ref{fig:one_orbit_wrong_ref}
+but for systematic offsets between $0$ and $500\,\pc$ in the $z$ and $x$
+components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report one half the
+95\uth{} minus 5\uth{} percentile divided by the true action value ($\Delta
+J_i/J_i$) for the three different fiducial orbits in Table~\ref{tab:orbits}.
+The thick-disk (\thickcolor) orbit is the one from
+Figure~\ref{fig:one_orbit_wrong_ref}, but we also now consider the effect on
+the action determined for the thin-disk (\thincolor) and halo (\halocolor)
+orbits.
+
+The upper row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
+in each action for an offset in the $z$-component (i.e. the Galactic
+midplane). In the lower row we consider offsets in the $x$ component
+(i.e. the Solar radius). The left, center, and right columns
+show the fractional spread in the values computed for $J_z$, $J_{\phi}$, and
+$J_R$, respectively. We compute the fractional spread $\Delta J_i/J_i$, where
+$\Delta J_i$ is defined in the previous paragraph and $J_i$ is the action
+value computed in the true coordinate system, over the course of the first
+$\Gyr$.
+
+In the upper middle panel of Figure~\ref{fig:many_orbit_wrong_ref},
+there is essentially no spread in the determination of $J_{\phi}$. This is
+expected since $J_{\phi}$ is independent of $z$ and is
+thus unaffected by offsets in $z$, as discussed earlier. Indeed, the result we found in
+Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
+orbit types. This is also a demonstration of the robustness of the orbit
+integration and action calculation methods we use.
+
+The upper right panel of Figure~\ref{fig:many_orbit_wrong_ref} shows
+that the fractional error in $J_z$ is more exaggerated for more planar
+(disk-like) orbits. For the thin-disk orbit, a systematic offset of
+$\sim15\,\pc$ in the $z$ coordinate already results in $25\%$ deviations in
+the actions. The required offset for $25\%$ deviation is $\sim120\,\pc$ for
+the thick-disk orbit. The halo orbit is relatively resistant to errors in the
+midplane, with only $\sim15\%$ error in $J_z$ out to a midplane offset of
+$500\,\pc$.
+
+For the offset in the Solar radius (lower panels), the error is largest
+for $J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
+deviations in $J_z$. In the lower middle and lower right panels
+all three lines nearly overlap.
+
+In each panel of Figure~\ref{fig:many_orbit_wrong_ref}, where relevant, we
+include the estimation of the action errors derived under the epicyclic
+approximation from Equation~\eqref{eq:Ji_err_mosttime}, with $\Delta v_{\phi}=0$, as dashed lines in the color of each orbit. This equation is
+relevant since during most of the orbit the star will be at maximum radial and
+vertical amplitude and we do not consider velocity offsets. Note that we
+consider an error in the $x$-coordinate $\Delta x$, which is not exactly the
+same as $\Delta R$. For observations of stars close to us, we have that
+$\Delta x \sim \Delta R$, but for the experiment performed in this Section we
+consider observations of the star throughout its entire orbit. This introduces
+a factor of $2/\pi$ when converting from $\Delta x$ to $\Delta R$, which we
+derive in Appendix~\ref{app:deltax}.
+
+The epicyclic approximation is a good predictor of $\Delta J_z$, even for halo
+orbits. It performs similarly for $\Delta J_R$, now underpredicting for halo orbits
+and slightly overpredicting for thin-disk orbits. Note that for the particular
+orbits we chose, the thin-disk orbit has slightly larger $A_R$ than the
+thick-disk orbit, and so we expect the epicyclic approximation to perform
+slightly worse. The epicyclic approximation underpredicts $\Delta J_{\phi}$ for all
+orbits.
 
@@
 
-These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
-gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles
-of mass $\sim 5000 \text{--} 7000\,
-\Msun$, with the lower end
-coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
-for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
-respectively.\footnote{This is $2.8$ times the often-quoted
-Plummer-equivalent.} The gas softening length is adaptive, but at $\z=0$ the
-median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
-roughly solar positions (with galactocentric cylindrical radii within $0.5\,\kpc$ of
-$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4$, $12.5$, and $10.3\,\pc$ for \mi{},
-\mf{}, and \mm{}, respectively. These values are summarized in
-Table~\ref{tab:scale_height}, along with measurements of the stellar and gas
-disk scale heights.
-
-\begin{deluxetable*}{cccccc}
+\begin{deluxetable*}{cccccc}[htb!]
 \tablecaption{Stellar and gas disk scale heights of the real Milky Way and the
-simulated galaxies considered in this work.\label{tab:scale_height}}
+simulated galaxies considered in this work. For comparison, we also give the softening lengths for the simulated galaxies.\label{tab:scale_height}}
 \tablehead{\colhead{galaxy} & \colhead{cold\tablenotemark{a} gas disk} &
@@
 
+These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
+gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles
+of mass $\sim 5000 \text{--} 7000\,
+\Msun$, with the lower end
+coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
+for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
+respectively.\footnote{This is $2.8$ times the often-quoted
+Plummer-equivalent.} The gas softening length is adaptive, but at $\z=0$ the
+median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
+roughly solar positions (with galactocentric cylindrical radii within $0.5\,\kpc$ of
+$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4$, $12.5$, and $10.3\,\pc$ for \mi{},
+\mf{}, and \mm{}, respectively. These values are summarized in
+Table~\ref{tab:scale_height}, along with measurements of the stellar and gas
+disk scale heights.
+
 The softening lengths used in the simulations can affect the ability to
@@
 
-This procedure assumes perfect density estimation, and therefore perfect
-corrections for extinction within the cylinder defining the ``solar
-neighborhood.'' Imperfect extinction correction is likely to increase the
-amplitude of the estimated fluctuations in $z$.
-
-To account for the effect of particle noise, we bootstrap resample
-stars within a cylinder of stars with height $2\,\kpc$ and the same radius
-$1000$ times and determine the $1\,\sigma$ error bars by repeating the
-midplane determination with those reselection of stars.
-
-To allow for potential small inaccuracies in the determination of the original
-fiducial coordinate system, we also subtract the best fit $\text{A}
-\sin{\left(\phi + \text{B}\right)} + \text{C}$ curve from the midplane as a
-function of azimuth to account for an overall tilt of the midplane (a
-simplified version of the strategy described in
-\citealt{2019ApJ...871..145A}). For $\text{A}$ the values are $-165$, $45$,
-and $8.8\,\pc$, for $\text{B}$ the values are $0.67$, $-0.088$, and
-$0.031\,\text{rad}$ and for $\text{C}$ the values are $-69$, $19$, and
-$-18\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. For the assumed Solar
-radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
-from the values of $\text{A}$ --- we compute $1.15$, $0.31$, and $0.06\,\deg$
-for \mi{}, \mf{}, and \mm{}. These angle offsets are consistent with the
-values given in \citet{2018arXiv180610564S} for the difference between the
-$z$-axis as defined by the gas and stars.
-
-Figure~\ref{fig:midplane} shows the relative $z$ location of the inferred
-midplane our imaginary observer would determine as a function of azimuth for
-each galaxy, using their local ``solar neighborhood'' (the cylinder defined
-above).
-We consider $50$ equally spaced bins in azimuth, a sufficient number such that each cylinder shares no stars with its neighboring cylinder.
-The $1\,\sigma$ error from the bootstrap procedure is shown as the
-dashed-line error bars. The middle $90\%$ of midplane values across the Solar
-circle spans $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of
-the three cases the midplane therefore varies by more than $\pm 100\,\pc$
-depending on the azimuth along the Solar circle; in the third (\mm{}, which
-has the thinnest ``thin disk'' of stars, but the largest stellar mass) the
-variation is closer to $\pm 50\,\pc$.
-
-We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
-but for four succcessive timesteps of the toy simulation of a Sagittarius encounter
-\citep{2018MNRAS.481..286L}. Again we have subtracted a best fit curve of the
-form $\text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}$, with the values
-for $\text{A}$ being $9.5$, $2.5$, $-212$, and $-394\,\pc$, for $\text{B}$
-being $0.0013$, $0.0007$, $-0.63$, and $-1.00\,\text{rad}$ and $\text{C}$
-being $2.6$, $-7.8$, $-65$, and $-53\,\pc$ chronologically for each of the four timesteps shown.
-
-In the upper panels, we see that the midplane is
-relatively flat in the inner galaxy, but additional encounters drive strong
-midplane variation. In the lower left panel, we see a strong $m=2$ mode develop,
-consistent with the $R=8\,\kpc$ panel of Figure~17 in
-\citet{2018MNRAS.481..286L}.
-The lower right panel, which shows the galaxy after some relaxation has occured is qualitatively similar to the midplane variations we saw in the FIRE simulations (Figure~\ref{fig:midplane}), evidence that they are at least partially driven by mergers.
-
-\begin{figure*}
+\begin{figure*}[htb!]
 \begin{center}
@@
 
-\begin{figure*}
+\begin{figure*}[htb!]
 \begin{center}
@@
 
+This procedure assumes perfect density estimation, and therefore perfect
+corrections for extinction within the cylinder defining the ``solar
+neighborhood.'' Imperfect extinction correction is likely to increase the
+amplitude of the estimated fluctuations in $z$.
+
+To account for the effect of particle noise, we bootstrap resample
+stars within a cylinder of stars with height $2\,\kpc$ and the same radius
+$1000$ times and determine the $1\,\sigma$ error bars by repeating the
+midplane determination with those reselection of stars.
+
+To allow for potential small inaccuracies in the determination of the original
+fiducial coordinate system, we also subtract the best fit $\text{A}
+\sin{\left(\phi + \text{B}\right)} + \text{C}$ curve from the midplane as a
+function of azimuth to account for an overall tilt of the midplane (a
+simplified version of the strategy described in
+\citealt{2019ApJ...871..145A}). For $\text{A}$ the values are $-165$, $45$,
+and $8.8\,\pc$, for $\text{B}$ the values are $0.67$, $-0.088$, and
+$0.031\,\text{rad}$ and for $\text{C}$ the values are $-69$, $19$, and
+$-18\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. For the assumed Solar
+radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
+from the values of $\text{A}$ --- we compute $1.15$, $0.31$, and $0.06\,\deg$
+for \mi{}, \mf{}, and \mm{}. These angle offsets are consistent with the
+values given in \citet{2018arXiv180610564S} for the difference between the
+$z$-axis as defined by the gas and stars.
+
+Figure~\ref{fig:midplane} shows the relative $z$ location of the inferred
+midplane our imaginary observer would determine as a function of azimuth for
+each galaxy, using their local ``solar neighborhood'' (the cylinder defined
+above).
+We consider $50$ equally spaced bins in azimuth, a sufficient number such that each cylinder shares no stars with its neighboring cylinder.
+The $1\,\sigma$ error from the bootstrap procedure is shown as the
+dashed-line error bars. The middle $90\%$ of midplane values across the Solar
+circle spans $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of
+the three cases the midplane therefore varies by more than $\pm 100\,\pc$
+depending on the azimuth along the Solar circle; in the third (\mm{}, which
+has the thinnest ``thin disk'' of stars, but the largest stellar mass) the
+variation is closer to $\pm 50\,\pc$.
+
+We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
+but for four succcessive timesteps of the toy simulation of a Sagittarius encounter
+\citep{2018MNRAS.481..286L}. Again we have subtracted a best fit curve of the
+form $\text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}$, with the values
+for $\text{A}$ being $9.5$, $2.5$, $-212$, and $-394\,\pc$, for $\text{B}$
+being $0.0013$, $0.0007$, $-0.63$, and $-1.00\,\text{rad}$ and $\text{C}$
+being $2.6$, $-7.8$, $-65$, and $-53\,\pc$ chronologically for each of the four timesteps shown.
+
+In the upper panels, we see that the midplane is
+relatively flat in the inner galaxy, but additional encounters drive strong
+midplane variation. In the lower left panel, we see a strong $m=2$ mode develop,
+consistent with the $R=8\,\kpc$ panel of Figure~17 in
+\citet{2018MNRAS.481..286L}.
+The lower right panel, which shows the galaxy after some relaxation has occured is qualitatively similar to the midplane variations we saw in the FIRE simulations (Figure~\ref{fig:midplane}), evidence that they are at least partially driven by mergers.
+
 \subsection{Velocity Variations} \label{ssec:lsr_var}

### 2019-05-05 paper/ms.tex
Try new travis LaTeX

@@
 what is normally considered the regime of the epicyclic approximation:
-quasi-harmonic oscillations around a circular ``guiding center orbit.'' Here
-we consider those implications for the prospect of distribution function
-fitting.
+quasi-harmonic oscillations around a circular ``guiding center orbit.''
 

### 2019-05-05 paper/ms.tex
Expand on warp drive footnote

@@
 observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and vary the azimuth
-between $0<\phi<2\pi$.\footnote{The immortal observer also has warp drive, allowing them to travel around the galaxy much more quickly than its dynamical timescale.} At
+between $0<\phi<2\pi$.\footnote{The immortal observer also has warp drive, allowing them to travel around the galaxy much more quickly than its dynamical timescale (warp factor $\sim0.25$).
+Using warp drive allows them to sidestep time dilation effects.} At
 each value of $\phi$ we then compute the median $z$ for stars within a

### 2019-05-05 paper/ms.tex
Add other Ji distributions to appendix

@@
 corresponds to the distance from the true $J_z$ value to one of the modes of
-the distribution of observed $J_z$ values.}
+the distribution of observed $J_z$ values. Similar plots for $J_R$ induced by
+a $z$ offset and $J_R$ and $J_{\phi}$ induced by an $x$ offset are given in
+Appendix~\ref{app:hist}.}
 \label{fig:Jz_hist}
@@
 Section~\ref{ssec:epi_action}).
+
+In Appendix~\ref{app:hist} we plot the same histogram as in
+Figure~\ref{fig:Jz_hist}, but for the observed distributions of $J_R$ induced
+by a $z$ offset (upper left panel of Figure~\ref{fig:many_orbit_wrong_ref})
+and the observed distributions of $J_R$ and $J_{\phi}$ errors induced by an
+$x$ offset (lower left and lower center panels of
+Figure~\ref{fig:many_orbit_wrong_ref}, respectively). We find similar error
+distributions as in Figure~\ref{fig:Jz_hist}, with the exception that the
+observed $J_R$ distribution induced by an $x$ offset more closely resembles a
+Gaussian.
 
@@
 
+\section{$J_R$ and $J_{\phi}$ Observed Distributions} \label{app:hist}
+
+In Figure~\ref{fig:JR_zerr_hist} we plot the observed $J_R$ distribution
+induced by an offset in $z$, and in Figures~\ref{fig:JR_xerr_hist} and
+\ref{fig:Jphi_xerr_hist} we plot the observed $J_R$ and $J_{\phi}$
+distributions induced by an offset in $x$. We plot the distributions for the
+thick-disk (top panels) and thin-disk (bottom panels) orbits.
+
+We find that the observed $J_R$ distribution induced by an offset in $x$ more
+closely resembles a Gaussian, while the observed $J_R$ distribution induced by
+an offset in $z$ and the observed $J_{\phi}$ distribution induced by an offset
+in $x$ are both similar to the observed $J_z$ distribution induced by an
+offset in $z$ (see Figure~\ref{fig:Jz_hist}).
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.35224200913in]{fig/schmactions_JR_zerr_hist.pdf}
+\end{center}
+\caption{A histogram of the observed values in $J_R$ for the thick-disk (top
+panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
+$100\,\pc$. The induced error distribution more closely resembles a Gaussian
+centered on the null value.}
+\label{fig:JR_zerr_hist}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.35224200913in]{fig/schmactions_JR_xerr_hist.pdf}
+\end{center}
+\caption{A histogram of the observed values in $J_R$ for the thick-disk (top
+panel) and thin-disk (bottom panel) orbits assuming a $x$ offset of
+$100\,\pc$. The induced error distribution is non-Gaussian and bimodal, with
+neither of the modes centered on null.}
+\label{fig:JR_xerr_hist}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.35224200913in]{fig/schmactions_Jphi_xerr_hist.pdf}
+\end{center}
+\caption{A histogram of the observed values in $J_{\phi}$ for the thick-disk (top
+panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
+$100\,\pc$. The induced error distribution is non-Gaussian and bimodal, with
+neither of the modes centered on null.}
+\label{fig:Jphi_xerr_hist}
+\end{figure}
+
 \section{LSR Variations} \label{app:lsr}

### 2019-05-05 paper/ms.tex
Slightly update Ji hist plots

@@
 between $0<\phi<2\pi$.\footnote{The immortal observer also has warp drive, allowing them to travel around the galaxy much more quickly than its dynamical timescale (warp factor $\sim0.25$).
-Using warp drive allows them to sidestep time dilation effects.} At
+Using warp drive allows them to circumvent time dilation effects.} At
 each value of $\phi$ we then compute the median $z$ for stars within a

### 2019-05-05 schmactions/one_orbit/plot_one_orbit.py

@@ -134,7 +134,7 @@ ax[0].text(J0+dJRJR/3.5, 38.5, r'$\Delta J_R$', color='k')
 
 ax[0].axvline(x=J0, color='k', ls='dashed', lw=1)
 ax[0].set_ylabel(r'$\text{count}$')
-ax[0].text(27.25, 40, r'\text{thick-disk orbit}', color='k')
+ax[0].text(27.25, 42.25, r'\text{thick-disk orbit}', color='k')
 
 k0t, k1t, k2t = sclip(xact_thin)
 dJRJR_thin = 0.5*(np.percentile(xact_thin[:,0][k0t], 95) - np.percentile(xact_thin[:,0][k0t], 5))
@@ -216,7 +216,7 @@ ax[1].text(J1_thin+dJphiJphi_thin/3.5, 35.75, r'$\Delta J_{\phi}$', color='k')
 
 ax[1].axvline(x=J1_thin, color='k', ls='dashed', lw=1)
 ax[1].set_ylabel(r'$\text{count}$')
-ax[1].text(-1550, 44, r'\text{thin-disk orbit}', color='k')
+ax[1].text(-1550, 41, r'\text{thin-disk orbit}', color='k')
 
 ax[1].set_xlabel(r'$J_{\phi,\text{obs}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
 

### 2019-05-08 paper/ms.tex
Update ms.tex
Fix travis
Small updates

--- 2019-05-05 paper/ms.tex
@@
 \author{Chervin F. P. Laporte}
-\altaffiliation{Simons Fellow}
+\altaffiliation{CITA National Fellow}
 \affil{\victoria}
@@
 include the estimation of the action errors derived under the epicyclic
-approximation from Equation~\eqref{eq:Ji_err_mosttime}, with $\Delta v_{\phi}=0$, as dashed lines in the color of each orbit. This equation is
+approximation from Equation~\eqref{eq:Ji_err_mosttime}, with $\Delta
+v_{\phi}=0$, as dashed lines in the color of each orbit. This equation is
 relevant since during most of the orbit the star will be at maximum radial and
@@
 In addition to the cosmological zoom-ins, we will also briefly consider results
-from a toy Sagittarius encounter simulation. This simulation offers us the
+from a live N-body simulation of a Sagittarius-like encounter. This simulation offers us the
 ability to see how the midplane varies in a more controlled environment. The
@@
 radius of $3.5\,\kpc$, scale height $0.53\,\kpc$, and mass
-$6\times10^{10}\,\Msun$, a Hernquist bulge of mass $10^{10}\,\Msun$ and scale
-length $0.7\,\kpc$ \citep{1990ApJ...356..359H}. Sagittarius is modeled as a
-dark matter Hernquist sphere of mass \Gus{something} and scale length
-\Gus{something}, along with a stellar component modeled as a Hernquist sphere
+$6\times10^{10}\,\Msun$, and a Hernquist bulge of mass $10^{10}\,\Msun$ and scale
+length $0.7\,\kpc$ \citep{1990ApJ...356..359H}. We use the L2 model, for which Sagittarius is modeled as a
+dark matter Hernquist sphere of mass $8\times10^{10}\,\Msun$ and scale length
+$8\,\kpc$, along with a stellar component modeled as a Hernquist sphere
 of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$. All components
@@
 The mass resolution of the simulation is $2.6\times10^4$, $1.2\times10^4$, and
-$1.0\times10^4\,\Msun$ for the dark matter, disk, and bulge components, respectively.
-For the disk and bulge components, a softening length of $30\,\pc$ is used
-whereas for the halo a softening length of $60\,\pc$ is used. For Sagittarius,
-the softening length for the dark matter and the stars is $60$ and $40\,\pc$,
-respectively.
+$1.0\times10^4\,\Msun$ for the dark matter, disk, and bulge components,
+respectively. For the disk and bulge components, a softening length of
+$30\,\pc$ is used whereas for the halo a softening length of $60\,\pc$ is
+used. For Sagittarius, the softening length for the dark matter and the stars
+is $60$ and $40\,\pc$, respectively.
 
@@
 observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and vary the azimuth
-between $0<\phi<2\pi$.\footnote{The immortal observer also has warp drive, allowing them to travel around the galaxy much more quickly than its dynamical timescale (warp factor $\sim0.25$).
-Using warp drive allows them to sidestep time dilation effects.} At
+between $0<\phi<2\pi$. At
 each value of $\phi$ we then compute the median $z$ for stars within a
@@
 in part by the Roy \& Diana Vagelos Program in the Molecular Life Sciences and
-the Roy \& Diana Vagelos Challenge Award. K.V.J.'s work was performed in part
-during the Gaia19 workshop and the 2019 Santa Barbara Gaia Sprint (also
-supported by the Heising-Simons Foundation), both hosted by the Kavli
-Institute for Theoretical Physics at the University of California, Santa
-Barbara. K.V.J.'s contributions were supported in part by the National Science
-Foundation under grants NSF PHY-1748958 and NSF AST-1614743. M.-M.M.L. was
-partly supported by NSF grant AST18-15461. The work of D.A.-A., A.B., D.W.H.,
-C.F.P.L., R.S., K.V.J., M.-M.M.L., and M.K.N. is supported by the Simons
+the Roy \& Diana Vagelos Challenge Award. K.V.J.'s and C.F.P.L.'s work was
+performed in part during the Gaia19 workshop and the 2019 Santa Barbara Gaia
+Sprint (also supported by the Heising-Simons Foundation), both hosted by the
+Kavli Institute for Theoretical Physics at the University of California, Santa
+Barbara. This work used the Extreme Science and Engineering Discovery
+Environment (XSEDE), which is supported by National Science Foundation grant
+number OCI-1053575. K.V.J.'s contributions were supported in part by the
+National Science Foundation under grants NSF PHY-1748958 and NSF AST-1614743.
+M.-M.M.L. was partly supported by NSF grant AST18-15461. The work of D.A.-A.,
+A.B., D.W.H., R.S., K.V.J., M.-M.M.L., and M.K.N. is supported by the Simons
 Foundation.

### 2019-05-08 paper/ms.tex
Remove mentions of toy

@@
 
-The fiducial coordinate system for these toy simulations is the rest frame of
-the aligned toy galaxy at the beginning of the simulation.
+The fiducial coordinate system for these N-body simulations is the rest frame of
+the aligned host galaxy at the beginning of the simulation.
 
@@
 \caption{The local midplane determined at the fiducial Solar radius
-($8.2\,\kpc$) for 4 different time steps from toy simulations
+($8.2\,\kpc$) for 4 different time steps from live N-body simulations
 of a Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
@@
 We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
-but for four succcessive timesteps of the toy simulation of a Sagittarius encounter
+but for four succcessive timesteps of the live N-body simulation of a Sagittarius encounter
 \citep{2018MNRAS.481..286L}. Again we have subtracted a best fit curve of the

### 2019-05-08 paper/ms.tex
Fix line indentations throughout

@@
 induces a phase-dependence in the actions, which we interpret as a systematic
-error. An offset in the vertical
-position of $\sim15\,\pc$ for a thin disk orbit or $\sim120\,\pc$ for
-a thick disk orbit is sufficient to induce a $25\,\%$ systematic error in the
-vertical action $J_z$ computed under the assumption of an axisymmetric
-potential. The induced error distribution is non-Gaussian and bimodal, with
-neither of the modes peaking on the null value. Furthermore, we show that the local midplane varies by
-$\sim100\,\pc$ at the Solar circle in Milky Way-mass cosmological zoom-in
-simulations from the FIRE project.
-From observations that the mean vertical velocity of the Milky Way disk varies by $\sim5\,\kms$ with radius, we estimate that the true midplane variations are of order $\sim60\textup{--}90\,\pc$.
-Thus, current state-of-the-art action
-calculations --- which assume the global and local midplanes are the same ---
-are likely to include a systematic vertical offset depending on the volume considered. Variation in the local
-standard of rest induces similar issues. The variation of the midplane must be
-taken into account when performing dynamical modeling across the large
-regions of the disk accessible to \textit{Gaia} and future missions.
+error. An offset in the vertical position of $\sim15\,\pc$ for a thin disk
+orbit or $\sim120\,\pc$ for a thick disk orbit is sufficient to induce a
+$25\,\%$ systematic error in the vertical action $J_z$ computed under the
+assumption of an axisymmetric potential. The induced error distribution is
+non-Gaussian and bimodal, with neither of the modes peaking on the null value.
+Furthermore, we show that the local midplane varies by $\sim100\,\pc$ at the
+Solar circle in Milky Way-mass cosmological zoom-in simulations from the FIRE
+project. From observations that the mean vertical velocity of the Milky Way
+disk varies by $\sim5\,\kms$ with radius, we estimate that the true midplane
+variations are of order $\sim60\textup{--}90\,\pc$. Thus, current
+state-of-the-art action calculations --- which assume the global and local
+midplanes are the same --- are likely to include a systematic vertical offset
+depending on the volume considered. Variation in the local standard of rest
+induces similar issues. The variation of the midplane must be taken into
+account when performing dynamical modeling across the large regions of the
+disk accessible to \textit{Gaia} and future missions.
 
@@
 Our understanding of the Milky Way is currently undergoing a revolution as a
-result of \textit{Gaia} Data Release 2 (DR2).  Recent major discoveries include
-the remnants of a major merger \citep{2018ApJ...860L..11K,
+result of \textit{Gaia} Data Release 2 (DR2).  Recent major discoveries
+include the remnants of a major merger \citep{2018ApJ...860L..11K,
 2018Natur.563...85H, 2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space
@@
 the $p_i$ are the conjugate momenta. In a slowly-evolving axisymmetric
-potential, these actions are invariant and $J_{\phi} \equiv L_z$, where $L_z$ is the $z$-component of the angular momentum
+potential, these actions are invariant and $J_{\phi} \equiv L_z$, where $L_z$
+is the $z$-component of the angular momentum
 \citep{2008gady.book.....B,2014RvMP...86....1S}.
@@
 midplane. For stars relatively far from the Sun, their local midplane location
-will not agree with our local midplane extrapolated onto their position due
-to relatively small-scale variations in the density of the gas and stellar
+will not agree with our local midplane extrapolated onto their position due to
+relatively small-scale variations in the density of the gas and stellar
 distribution. Converting positions and velocities of more distant stars from a
@@
 mismatch will result in large systematic uncertainty, especially in the
-vertical action $J_z$. A similar
-argument applies to any remaining uncertainty in measurements of the
-Galactocentric radius of the Sun and to variations in the LSR.
+vertical action $J_z$. A similar argument applies to any remaining uncertainty
+in measurements of the Galactocentric radius of the Sun and to variations in
+the LSR.
 
@@
 dashed line. Suppose this observer measures the phase-space position of the
-star at two different orbital phases (teal and orange points). If the
-observer were then to integrate the star's orbit using a model potential with
-the erroneous midplane, they would obtain the teal and orange curves for the
+star at two different orbital phases (teal and orange points). If the observer
+were then to integrate the star's orbit using a model potential with the
+erroneous midplane, they would obtain the teal and orange curves for the
 star's orbit, respectively. The actions estimated from these two  erroneous
@@
 collaboration\footnote{\url{https://fire.northwestern.edu}}
-\citep{2014MNRAS.445..581H, 2016ApJ...827L..23W, 2018MNRAS.480..800H}, and a controlled N-body
-simulation of a Sagittarius encounter with a galaxy otherwise tailored to the
-stellar mass, scale length, and scale height of the Milky Way
-\citep{2018MNRAS.481..286L}.
+\citep{2014MNRAS.445..581H, 2016ApJ...827L..23W, 2018MNRAS.480..800H}, and a
+controlled N-body simulation of a Sagittarius encounter with a galaxy
+otherwise tailored to the stellar mass, scale length, and scale height of the
+Milky Way \citep{2018MNRAS.481..286L}.
 
@@
 variations, and the resulting systematic uncertainty in the vertical action,
-for action-space analyses. We also estimate the expected midplane variations of the Milky Way based on the observed velocity variations. We
-summarize our main results and conclude in Section~\ref{sec:conclusion}.
+for action-space analyses. We also estimate the expected midplane variations
+of the Milky Way based on the observed velocity variations. We summarize our
+main results and conclude in Section~\ref{sec:conclusion}.
 
@@
 true midplane. The dashed gray line is the midplane location in the potential
-model used by an observer to integrate the orbit of the star and hence estimate its actions,
-which is offset from the true midplane of the star's orbit. The model
-potential is otherwise identical to the one in which the star is actually
-moving.
+model used by an observer to integrate the orbit of the star and hence
+estimate its actions, which is offset from the true midplane of the star's
+orbit. The model potential is otherwise identical to the one in which the star
+is actually moving.
 
@@
 work. For the last two columns: $z_{\text{max}}$ is the maximum height of the
-orbit while $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of the radial excursions of the orbit.
-In the epicyclic approximation, $A_z = z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$.
+orbit while $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of
+the radial excursions of the orbit. In the epicyclic approximation, $A_z =
+z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$.
 \label{tab:orbits}}
@@
 spiral arms). We also make the assumption of a perfectly flat circular
-velocity curve with velocity $v_c$, a good approximation near the Solar circle \citep[e.g.][]{2017MNRAS.465...76M}.
+velocity curve with velocity $v_c$, a good approximation near the Solar circle
+\citep[e.g.][]{2017MNRAS.465...76M}.
 
@@
 \eeq
-where the last equality in each line comes from the fact that $v_{R,\text{max}} = \kappa
-A_R$ and $v_{z,\text{max}} = \nu A_z$.
+where the last equality in each line comes from the fact that
+$v_{R,\text{max}} = \kappa A_R$ and $v_{z,\text{max}} = \nu A_z$.
 
@@
 \Delta v_{\phi}, \Delta v_R, \Delta v_z)$. We can then apply the standard
-propagation of errors formula to Equation~\eqref{eq:JR_Jz_epi_energy}
-to determine the error in each of the actions. For $J_{\phi}$, the induced error is:
+propagation of errors formula to Equation~\eqref{eq:JR_Jz_epi_energy} to
+determine the error in each of the actions. For $J_{\phi}$, the induced error
+is:
 \beq\label{eq:induced_Jphi}
@@
 \begin{split}
-\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta R}{R_g} + \frac{\Delta v_{\phi}}{v_c} \\
+\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta R}{R_g} 
+                                    + \frac{\Delta v_{\phi}}{v_c} \\
 \frac{\Delta J_{R}}{J_{R}} &= \frac{2\Delta R}{A_R} \\
@@
 numerical computations of the actions for a range of orbits in a model
-Galactic potential. We compute actions as in
-\citet{2018ApJ...867...31B}, using the code \texttt{gala} v0.3 to perform
-orbit integrations and conversion to action space
-\citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute actions we use the
-torus-mapping technique first presented by \citet{1990MNRAS.244..634M} and
-adapted by \citet{2014MNRAS.441.3284S} to calculate actions for an orbital
-time-series starting from a phase-space position $(x, v)$ and integrated in a
-potential $\Phi$. We use the default \texttt{MWPotential} as our potential,
-which is based on the Milky Way potential available in \texttt{galpy}
-\citep{2015ApJS..216...29B}. This potential includes a Hernquist bulge and
-nucleus \citep{1990ApJ...356..359H}, a Miyamoto--Nagai disk
-\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
-is fit to empirically match some observations. We use the Dormand-Prince
-8(5,3) integration scheme \citep{Dormand80:integrator} with a timestep of
-$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
-Sun-like star.
+Galactic potential. We compute actions as in \citet{2018ApJ...867...31B},
+using the code \texttt{gala} v0.3 to perform orbit integrations and conversion
+to action space \citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute
+actions we use the torus-mapping technique first presented by
+\citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to
+calculate actions for an orbital time-series starting from a phase-space
+position $(x, v)$ and integrated in a potential $\Phi$. We use the default
+\texttt{MWPotential} as our potential, which is based on the Milky Way
+potential available in \texttt{galpy} \citep{2015ApJS..216...29B}. This
+potential includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
+a Miyamoto--Nagai disk \citep{1975PASJ...27..533M}, and an NFW halo
+\citep{1997ApJ...490..493N}, and is fit to empirically match some
+observations. We use the Dormand-Prince 8(5,3) integration scheme
+\citep{Dormand80:integrator} with a timestep of $1\,\Myr$ and integrate for
+$5\,\Gyr$, corresponding to $\sim 20$ orbits for a Sun-like star.
 
@@
 St\"ackel potential (with analytic actions) to approximate the Galactic
-potential \citep{1985MNRAS.216..273D,2012MNRAS.426.1324B}, was used in many
+potential \citep{1985MNRAS.216..273D, 2012MNRAS.426.1324B}, was used in many
 recent works exploring actions in the Galactic disk
-\citep[e.g.][]{2019MNRAS.484.3291T,2018MNRAS.481.4093S,2018arXiv180803278T}.
+\citep[e.g.][]{2019MNRAS.484.3291T, 2018MNRAS.481.4093S, 2018arXiv180803278T}.
 For disk-like orbits, existing implementations of the St\"ackel Fudge method
@@
 integrate the orbit according to the procedure laid out in
-Section~\ref{ssec:action_comp}, and which we plot in Appendix~\ref{app:orbits}. Then, we subtract $100\,\pc$ from the $z$
-value (upper panels) or the $x$ value (lower panels) of each
-position in the orbit, corresponding to an erroneous observer assuming a
-midplane (upper) or solar radius (lower) that is off by
-$100\,\pc$. We then allow an immortal observer to observe the orbit over
-$1\,\Gyr$ and perform the same orbit integration procedure at each timestep,
-and report the values of the actions. The computation of $J_{\phi}$ is
-pristine to errors in $z$, with only numerical artifacts remaining. Only small
-errors are induced in $J_R$, with the middle $90\%$ of values over the $\Gyr$
-being within $\sim8\%$ of the true $J_R$. As expected, large errors are
-induced in $J_z$ with a $100\,\pc$ offset in $z$, with the middle $90\%$ of
-values being within $\sim43\%$ of the true $J_z$. The $x$ offset induces
-uncertainties in $J_R$, $J_{\phi}$, and $J_z$ of $\sim21\%$, $\sim3\%$, and $\sim3\%$.}
+Section~\ref{ssec:action_comp}, and which we plot in
+Appendix~\ref{app:orbits}. Then, we subtract $100\,\pc$ from the $z$ value
+(upper panels) or the $x$ value (lower panels) of each position in the orbit,
+corresponding to an erroneous observer assuming a midplane (upper) or solar
+radius (lower) that is off by $100\,\pc$. We then allow an immortal observer
+to observe the orbit over $1\,\Gyr$ and perform the same orbit integration
+procedure at each timestep, and report the values of the actions. The
+computation of $J_{\phi}$ is pristine to errors in $z$, with only numerical
+artifacts remaining. Only small errors are induced in $J_R$, with the middle
+$90\%$ of values over the $\Gyr$ being within $\sim8\%$ of the true $J_R$. As
+expected, large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$,
+with the middle $90\%$ of values being within $\sim43\%$ of the true $J_z$.
+The $x$ offset induces uncertainties in $J_R$, $J_{\phi}$, and $J_z$ of
+$\sim21\%$, $\sim3\%$, and $\sim3\%$.}
 \label{fig:one_orbit_wrong_ref}
@@
 \end{center}
-\caption{A histogram of the observed values in $J_z$ for the thick-disk (top panel) and thin-disk (bottom panel) orbits
-assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
-values have a bias (from e.g. an incorrectly computed midplane), then the
-induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
-sort of error propagation must take this into account. Intuition for the shape
-of each panel is given in the text. We also plot one half the 95\uth{}
-percentile minus the 5\uth{} percentile of each distribution as a horizontal
-arrow anchored on the true $J_z$ value. We call this $\Delta J_z$ and will use
-it (along with the similarly defined $\Delta J_R$ and $\Delta J_{\phi}$) to
-empirically describe the error distribution. We see that $\Delta J_z$ roughly
-corresponds to the distance from the true $J_z$ value to one of the modes of
-the distribution of observed $J_z$ values. Similar plots for $J_R$ induced by
-a $z$ offset and $J_R$ and $J_{\phi}$ induced by an $x$ offset are given in
+\caption{A histogram of the observed values in $J_z$ for the thick-disk (top
+panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
+$100\,\pc$. One can see that if the observed $z$ values have a bias (from e.g.
+an incorrectly computed midplane), then the induced error distribution in
+$J_z$ is decidedly non-Gaussian. Therefore, any sort of error propagation must
+take this into account. Intuition for the shape of each panel is given in the
+text. We also plot one half the 95\uth{} percentile minus the 5\uth{}
+percentile of each distribution as a horizontal arrow anchored on the true
+$J_z$ value. We call this $\Delta J_z$ and will use it (along with the
+similarly defined $\Delta J_R$ and $\Delta J_{\phi}$) to empirically describe
+the error distribution. We see that $\Delta J_z$ roughly corresponds to the
+distance from the true $J_z$ value to one of the modes of the distribution of
+observed $J_z$ values. Similar plots for $J_R$ induced by a $z$ offset and
+$J_R$ and $J_{\phi}$ induced by an $x$ offset are given in
 Appendix~\ref{app:hist}.}
@@
 We quantify how a systematic error in the Galactocentric coordinate system
-induces phase-dependence in the \emph{observed} actions. We will consider three
-orbits in the model potential described in Section~\ref{ssec:action_comp} that
-are typical of stars in the stellar thin disk, stellar thick disk, and the
-stellar halo. We summarize their initial positions in phase space and the
-actions computed by integrating their orbits in the correct potential in
-Table~\ref{tab:orbits}. Each orbit, integrated without systematic coordinate
-errors, is plotted in Appendix~\ref{app:orbits}. We will refer to these orbits
-by their names (thin-disk, thick-disk, halo) henceforth.
-
-We begin with the thick-disk orbit. Consider an immortal observer who
-can measure the thick-disk star's phase-space position at many different times
-(and hence different orbital phases). However, the immortal observer uses a coordinate system in which
-the midplane is systematically offset in height by $100\,\pc$ from its actual
-location --- i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position
-in the orbit. This corresponds to the immortal observer physically located at
-e.g. the position $(8, 0, 0)\,\kpc$ in the coordinate system of the true
-potential, but erroneously thinking they are located at $(8, 0, 0.1)\,\kpc$.
-
-We consider that the immortal observer makes a measurement every $\Myr$.
-At each measurement, we compute the actions as outlined above using the true potential but by specifiying the star's starting position using the systematically offset coordinate system.
-Essentially we are shifting the
+induces phase-dependence in the \emph{observed} actions. We will consider
+three orbits in the model potential described in
+Section~\ref{ssec:action_comp} that are typical of stars in the stellar thin
+disk, stellar thick disk, and the stellar halo. We summarize their initial
+positions in phase space and the actions computed by integrating their orbits
+in the correct potential in Table~\ref{tab:orbits}. Each orbit, integrated
+without systematic coordinate errors, is plotted in Appendix~\ref{app:orbits}.
+We will refer to these orbits by their names (thin-disk, thick-disk, halo)
+henceforth.
+
+We begin with the thick-disk orbit. Consider an immortal observer who can
+measure the thick-disk star's phase-space position at many different times
+(and hence different orbital phases). However, the immortal observer uses a
+coordinate system in which the midplane is systematically offset in height by
+$100\,\pc$ from its actual location --- i.e. we subtract the vector $(0, 0,
+100)\,\pc$ from each position in the orbit. This corresponds to the immortal
+observer physically located at e.g. the position $(8, 0, 0)\,\kpc$ in the
+coordinate system of the true potential, but erroneously thinking they are
+located at $(8, 0, 0.1)\,\kpc$.
+
+We consider that the immortal observer makes a measurement every $\Myr$. At
+each measurement, we compute the actions as outlined above using the true
+potential but by specifiying the star's starting position using the
+systematically offset coordinate system. Essentially we are shifting the
 integrated orbit and then reintegrating at each point along the shifted orbit,
@@
 
-For a systematic offset in $z$ (upper panels), the
-95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
-$6.2\,\actunit$ for $J_R$ and $J_z$, respectively. These are fractional errors
-of $5.7\,\%$ and $86\,\%$, respectively. The error induced in $J_{\phi}$ is
-negligible, as expected since $J_{\phi}$ only depends on the $x$- and
-$y$-components of the position and velocity of the stars.\footnote{In practice,
-however, $J_{\phi}$ is computed as part of the torus-fitting method.} It is worth
-pointing out that a $100\,\pc$ error in an orbit with
-$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
---- induced an $86\%$ error in the computation of $J_z$.
+For a systematic offset in $z$ (upper panels), the 95\textsuperscript{th}
+minus 5\textsuperscript{th} percentiles are $2.2$ and $6.2\,\actunit$ for
+$J_R$ and $J_z$, respectively. These are fractional errors of $5.7\,\%$ and
+$86\,\%$, respectively. The error induced in $J_{\phi}$ is negligible, as
+expected since $J_{\phi}$ only depends on the $x$- and $y$-components of the
+position and velocity of the stars.\footnote{In practice, however, $J_{\phi}$
+is computed as part of the torus-fitting method.} It is worth pointing out
+that a $100\,\pc$ error in an orbit with $z_{\text{max}}=850\,\pc$ --- a
+$12\%$ error --- induced an $86\%$ error in the computation of $J_z$.
 
@@
 In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
-$J_z$ for the thick-disk orbit (top panel) and thin-disk orbit (bottom panel) assuming a $z$ offset of $100\,\pc$ (i.e. as in the upper
-right panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$
-is plotted as a vertical dashed line. Here we see that the errors in $J_z$
-induced by a systematic offset in $z$ are non-Gaussian and bimodal.
-Furthermore, neither of the modes lie at zero. In other words, neither of the
-modes of the distribution of observed $J_z$ values are centered on the true
-$J_z$ (i.e. the $J_z$ computed with no offset in $z$). In the case of the
-thin-disk orbit (bottom panel), we see that, in addition to the prior
-complications, the error distribution is not even centered on null. This comes
-about when the midplane error is $\gtrsim z_{\text{max}}$, where
-$z_{\text{max}}$ is the maximum height of the orbit, equivalent to $A_z$ in the epicyclic approximation (see
+$J_z$ for the thick-disk orbit (top panel) and thin-disk orbit (bottom panel)
+assuming a $z$ offset of $100\,\pc$ (i.e. as in the upper right panel of
+Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted as a
+vertical dashed line. Here we see that the errors in $J_z$ induced by a
+systematic offset in $z$ are non-Gaussian and bimodal. Furthermore, neither of
+the modes lie at zero. In other words, neither of the modes of the
+distribution of observed $J_z$ values are centered on the true $J_z$ (i.e. the
+$J_z$ computed with no offset in $z$). In the case of the thin-disk orbit
+(bottom panel), we see that, in addition to the prior complications, the error
+distribution is not even centered on null. This comes about when the midplane
+error is $\gtrsim z_{\text{max}}$, where $z_{\text{max}}$ is the maximum
+height of the orbit, equivalent to $A_z$ in the epicyclic approximation (see
 Section~\ref{ssec:epi_action}).
@@
 text for the justification in using this to measure the magnitude of the
-induced error. The left, center, and right panels show the
-result for $J_R$, $J_{\phi}$, and $J_z$, respectively. The upper panels
-consider an offset in $z$ and the lower panels consider an offset in $x$
-(equivalently, an offset in the Solar radius). In some panels we also plot as
-dashed lines the epicyclic prediction of the induced action error
-(Equation~\eqref{eq:Ji_err_mosttime}). In the epicyclic approximation, a $z$
-offset only induces an error in $J_z$ --- for all three orbits the epicyclic
-approximation is a good description of the $J_z$ error. An $x$ offset induces
-an error in $J_R$ and $J_{\phi}$. The error in $J_R$ is somewhat
-well-described for the thin-disk and thick-disk orbits, and a poor description
-for the halo orbit. For $J_{\phi}$, the epicyclic approximation is not a good
-description for any orbit.}
+induced error. The left, center, and right panels show the result for $J_R$,
+$J_{\phi}$, and $J_z$, respectively. The upper panels consider an offset in
+$z$ and the lower panels consider an offset in $x$ (equivalently, an offset in
+the Solar radius). In some panels we also plot as dashed lines the epicyclic
+prediction of the induced action error (Equation~\eqref{eq:Ji_err_mosttime}).
+In the epicyclic approximation, a $z$ offset only induces an error in $J_z$
+--- for all three orbits the epicyclic approximation is a good description of
+the $J_z$ error. An $x$ offset induces an error in $J_R$ and $J_{\phi}$. The
+error in $J_R$ is somewhat well-described for the thin-disk and thick-disk
+orbits, and a poor description for the halo orbit. For $J_{\phi}$, the
+epicyclic approximation is not a good description for any orbit.}
 \label{fig:many_orbit_wrong_ref}
@@
 We now describe the intuition behind the shape of Figure~\ref{fig:Jz_hist}.
-Consider first the thick-disk orbit (top panel), where the offset in $z$
-is much less than $z_{\text{max}}$ of the orbit. The peaks in the distribution
+Consider first the thick-disk orbit (top panel), where the offset in $z$ is
+much less than $z_{\text{max}}$ of the orbit. The peaks in the distribution
 correspond to the turning points of the orbit (or points of maximum vertical
-amplitude), where $v_z \sim 0$ and where the star is on most of it orbit.
-This is why the distribution is peaked on these values. For the thin-disk
-orbit (bottom panel), the offset in $z$ is comparable to
-$z_{\text{max}}$. Now, there will be some points in the orbit where $v_z = 0$
-and $z=0$ (in the observed, erroneous coordinate system). At these points, the
-computed $J_z$ will vanish. The asymmetry and systematic offset then comes
-about because of the constraint that $J_z \geq 0$.\footnote{This is similar to
-arguments in cosmology for why gravity produces non-Gaussianity in the density
-field, since the density cannot become negative but it can grow arbitrarily
-large.}
+amplitude), where $v_z \sim 0$ and where the star is on most of it orbit. This
+is why the distribution is peaked on these values. For the thin-disk orbit
+(bottom panel), the offset in $z$ is comparable to $z_{\text{max}}$. Now,
+there will be some points in the orbit where $v_z = 0$ and $z=0$ (in the
+observed, erroneous coordinate system). At these points, the computed $J_z$
+will vanish. The asymmetry and systematic offset then comes about because of
+the constraint that $J_z \geq 0$.\footnote{This is similar to arguments in
+cosmology for why gravity produces non-Gaussianity in the density field, since
+the density cannot become negative but it can grow arbitrarily large.}
 
@@
 
-The upper row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
-in each action for an offset in the $z$-component (i.e. the Galactic
-midplane). In the lower row we consider offsets in the $x$ component
-(i.e. the Solar radius). The left, center, and right columns
-show the fractional spread in the values computed for $J_z$, $J_{\phi}$, and
-$J_R$, respectively. We compute the fractional spread $\Delta J_i/J_i$, where
-$\Delta J_i$ is defined in the previous paragraph and $J_i$ is the action
-value computed in the true coordinate system, over the course of the first
-$\Gyr$.
-
-In the upper middle panel of Figure~\ref{fig:many_orbit_wrong_ref},
-there is essentially no spread in the determination of $J_{\phi}$. This is
-expected since $J_{\phi}$ is independent of $z$ and is
-thus unaffected by offsets in $z$, as discussed earlier. Indeed, the result we found in
+The upper row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread
+induced in each action for an offset in the $z$-component (i.e. the Galactic
+midplane). In the lower row we consider offsets in the $x$ component (i.e. the
+Solar radius). The left, center, and right columns show the fractional spread
+in the values computed for $J_z$, $J_{\phi}$, and $J_R$, respectively. We
+compute the fractional spread $\Delta J_i/J_i$, where $\Delta J_i$ is defined
+in the previous paragraph and $J_i$ is the action value computed in the true
+coordinate system, over the course of the first $\Gyr$.
+
+In the upper middle panel of Figure~\ref{fig:many_orbit_wrong_ref}, there is
+essentially no spread in the determination of $J_{\phi}$. This is expected
+since $J_{\phi}$ is independent of $z$ and is thus unaffected by offsets in
+$z$, as discussed earlier. Indeed, the result we found in
 Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
@@
 
-The upper right panel of Figure~\ref{fig:many_orbit_wrong_ref} shows
-that the fractional error in $J_z$ is more exaggerated for more planar
-(disk-like) orbits. For the thin-disk orbit, a systematic offset of
-$\sim15\,\pc$ in the $z$ coordinate already results in $25\%$ deviations in
-the actions. The required offset for $25\%$ deviation is $\sim120\,\pc$ for
-the thick-disk orbit. The halo orbit is relatively resistant to errors in the
-midplane, with only $\sim15\%$ error in $J_z$ out to a midplane offset of
-$500\,\pc$.
-
-For the offset in the Solar radius (lower panels), the error is largest
-for $J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
-deviations in $J_z$. In the lower middle and lower right panels
-all three lines nearly overlap.
+The upper right panel of Figure~\ref{fig:many_orbit_wrong_ref} shows that the
+fractional error in $J_z$ is more exaggerated for more planar (disk-like)
+orbits. For the thin-disk orbit, a systematic offset of $\sim15\,\pc$ in the
+$z$ coordinate already results in $25\%$ deviations in the actions. The
+required offset for $25\%$ deviation is $\sim120\,\pc$ for the thick-disk
+orbit. The halo orbit is relatively resistant to errors in the midplane, with
+only $\sim15\%$ error in $J_z$ out to a midplane offset of $500\,\pc$.
+
+For the offset in the Solar radius (lower panels), the error is largest for
+$J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
+deviations in $J_z$. In the lower middle and lower right panels all three
+lines nearly overlap.
 
@@
 The epicyclic approximation is a good predictor of $\Delta J_z$, even for halo
-orbits. It performs similarly for $\Delta J_R$, now underpredicting for halo orbits
-and slightly overpredicting for thin-disk orbits. Note that for the particular
-orbits we chose, the thin-disk orbit has slightly larger $A_R$ than the
-thick-disk orbit, and so we expect the epicyclic approximation to perform
-slightly worse. The epicyclic approximation underpredicts $\Delta J_{\phi}$ for all
-orbits.
+orbits. It performs similarly for $\Delta J_R$, now underpredicting for halo
+orbits and slightly overpredicting for thin-disk orbits. Note that for the
+particular orbits we chose, the thin-disk orbit has slightly larger $A_R$ than
+the thick-disk orbit, and so we expect the epicyclic approximation to perform
+slightly worse. The epicyclic approximation underpredicts $\Delta J_{\phi}$
+for all orbits.
 
@@
 \caption{The fractional error in $J_z$ as a function of $J_z$ for a few
-different offsets in $z$. All orbits have the same initial position of $(8, 0, 0)\,\kpc$ and velocity $(-190, 0, v_z)\,\kms$, where we vary $v_z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
-(orange, teal, and green respectively). As before, the error ($\Delta J_z$) is
-one half the 95\uth{} minus 5\uth{} percentile of the distribution of observed
-$J_z$ values over the course of the orbit. There are large errors for
-thin-disk like orbits ($J_z \sim 0.5\,\actunit$), even for a small midplane
-offset of $10\,\pc$. As dashed lines in each color we also plot the prediction
-for $\Delta J_z/J_z$ from the epicyclic approximation
-(Equation~\eqref{eq:Ji_err_mosttime}), which shows excellent agreement with
-the numerically computed values.}
+different offsets in $z$. All orbits have the same initial position of $(8, 0,
+0)\,\kpc$ and velocity $(-190, 0, v_z)\,\kms$, where we vary $v_z$. We show
+this for a $z$ offset of $10, 50, 100\,\pc$ (orange, teal, and green
+respectively). As before, the error ($\Delta J_z$) is one half the 95\uth{}
+minus 5\uth{} percentile of the distribution of observed $J_z$ values over the
+course of the orbit. There are large errors for thin-disk like orbits ($J_z
+\sim 0.5\,\actunit$), even for a small midplane offset of $10\,\pc$. As dashed
+lines in each color we also plot the prediction for $\Delta J_z/J_z$ from the
+epicyclic approximation (Equation~\eqref{eq:Ji_err_mosttime}), which shows
+excellent agreement with the numerically computed values.}
 \label{fig:dJz_fun_Jz}
@@
 100\,\pc$ (orange, teal, and green, respectively) in
-Figure~\ref{fig:dJz_fun_Jz}. For each orbit, we set the initial position to be $(8,0,0)\,\kpc$ and the initial velocity to be $(-190, 0, v_z)\,\kms$, where we vary $v_z$. For a thin disk-like orbit
-($J_z\sim0.5\,\actunit$), even a $10\,\pc$ offset in $z$ is enough to induce a
-$\sim25\%$ error in $J_z$. For larger values of $J_z$, the fractional errors
-are suppressed, but the induced error can still be large depending on how
-great the $z$ offset is. We also plot the epicylic prediction for $\Delta J_z
-/ J_z$ from Equation~\eqref{eq:Ji_err_mosttime} as dashed lines for each $z$
-offset. We find that the epicyclic approximation matches the numerical
-estimate quite well.
+Figure~\ref{fig:dJz_fun_Jz}. For each orbit, we set the initial position to be
+$(8,0,0)\,\kpc$ and the initial velocity to be $(-190, 0, v_z)\,\kms$, where
+we vary $v_z$. For a thin disk-like orbit ($J_z\sim0.5\,\actunit$), even a
+$10\,\pc$ offset in $z$ is enough to induce a $\sim25\%$ error in $J_z$. For
+larger values of $J_z$, the fractional errors are suppressed, but the induced
+error can still be large depending on how great the $z$ offset is. We also
+plot the epicylic prediction for $\Delta J_z / J_z$ from
+Equation~\eqref{eq:Ji_err_mosttime} as dashed lines for each $z$ offset. We
+find that the epicyclic approximation matches the numerical estimate quite
+well.
 
@@
 \citep[\sgra{}, e.g.][]{2004ApJ...616..872R}. From stellar motions near
-\sgra{}, the distance from the Sun to
-\sgra{}, referred to as $R_0$, can be precisely measured \citep{2009ApJ...692.1075G,
-2018AA...615L..15G}, with a recent measurement using near-infrared
-interferometry of $8.178 \pm 0.035$~kpc \citep{2019arXiv190405721A}.
+\sgra{}, the distance from the Sun to \sgra{}, referred to as $R_0$, can be
+precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a
+recent measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
+\citep{2019arXiv190405721A}.
 
@@
 \subsection{Local Standard of Rest}
-Finally, one must define the LSR, or mean velocity of
-stars near the Sun relative to the Galactic center (which is defined to have
-zero velocity), and the velocity of the Sun relative to the LSR. The radial
-($U_{\odot}$) and vertical ($W_{\odot}$) components are computed by taking the
-mean motions of different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}.
-The azimuthal component ($V_{\odot}$) is more difficult to measure, but can be
-modeled using the asymmetric drift relation \citep{2008gady.book.....B}. The
-values of the components of the LSR are usually taken from
-\citet{2010MNRAS.403.1829S}. The value of the circular velocity is taken to be
-$\sim 220\kms$ \citep[e.g.][]{2012ApJ...759..131B}.
+Finally, one must define the LSR, or mean velocity of stars near the Sun
+relative to the Galactic center (which is defined to have zero velocity), and
+the velocity of the Sun relative to the LSR. The radial ($U_{\odot}$) and
+vertical ($W_{\odot}$) components are computed by taking the mean motions of
+different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal
+component ($V_{\odot}$) is more difficult to measure, but can be modeled using
+the asymmetric drift relation \citep{2008gady.book.....B}. The values of the
+components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}. The
+value of the circular velocity is taken to be $\sim 220\kms$
+\citep[e.g.][]{2012ApJ...759..131B}.
 
@@
 \tablecaption{Stellar and gas disk scale heights of the real Milky Way and the
-simulated galaxies considered in this work. For comparison, we also give the softening lengths for the simulated galaxies.\label{tab:scale_height}}
+simulated galaxies considered in this work. For comparison, we also give the
+softening lengths for the simulated galaxies.\label{tab:scale_height}}
 \tablehead{\colhead{galaxy} & \colhead{cold\tablenotemark{a} gas disk} &
@@
 These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
-gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles
-of mass $\sim 5000 \text{--} 7000\,
-\Msun$, with the lower end
-coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
-for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
+gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles of
+mass $\sim 5000 \text{--} 7000\, \Msun$, with the lower end coming from
+stellar evolution \citep{2018arXiv180610564S}. Softening lengths for dark
+matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
 respectively.\footnote{This is $2.8$ times the often-quoted
@@
 median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
-roughly solar positions (with galactocentric cylindrical radii within $0.5\,\kpc$ of
-$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4$, $12.5$, and $10.3\,\pc$ for \mi{},
-\mf{}, and \mm{}, respectively. These values are summarized in
-Table~\ref{tab:scale_height}, along with measurements of the stellar and gas
-disk scale heights.
+roughly solar positions (with galactocentric cylindrical radii within
+$0.5\,\kpc$ of $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4$, $12.5$, and
+$10.3\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. These values are
+summarized in Table~\ref{tab:scale_height}, along with measurements of the
+stellar and gas disk scale heights.
 
@@
 $40\,\pc$, $\sim3\textup{--}4$ times the cold gas softening length
-\citep{2019ApJ...871..145A}. The thin stellar disk has a scale height of
-about $300\,\pc$, $\sim30$ times the stellar softening length
-\citep{2008ApJ...673..864J}. We therefore expect that resolution effects
-may still be affecting the scale heights of these components in the
-simulations, especially the cold gas. Indeed, the stellar scale heights of the
-simulated galaxies are equal to or larger than the Milky Way's while the gas scale
+\citep{2019ApJ...871..145A}. The thin stellar disk has a scale height of about
+$300\,\pc$, $\sim30$ times the stellar softening length
+\citep{2008ApJ...673..864J}. We therefore expect that resolution effects may
+still be affecting the scale heights of these components in the simulations,
+especially the cold gas. Indeed, the stellar scale heights of the simulated
+galaxies are equal to or larger than the Milky Way's while the gas scale
 heights are significantly larger (although the proper basis comparison is less
@@
 Cosmological simulations of Milky Way-mass galaxies are not perfect
-representations of the true Milky Way in other ways as well,\footnote{The failure of cosmological simulations to exactly reproduce the Milky Way is not entirely due to limitations of the numerical model.
-Candidate Milky Way-like galaxies are chosen solely on their mass and isolation, for which there are a wide variety of possible galaxies with qualitatively different properties.} as discussed in
-\citet{2018arXiv180610564S}. For instance, the velocity structure of \mi{} is
-closer to M31's than the Milky Way's (Loebman et al. in prep). However, in
-this work we are most interested in the global properties of the potential,
-and specifically in deviations from axisymmetry. From this perspective, the
-simulated galaxies are actually \emph{more} axisymmetric than we might expect
-of the Milky Way. While they have prominent spiral arms, none has as strong a
-bar as the Milky Way does at present day, and none has a nearby companion like the
-Large Magellanic Cloud. One of the three we consider (\mf) does have an
-ongoing interaction with a satellite galaxy similar to Sagittarius, which has
-punched through the galactic disk outside the solar circle and induced some
-warping.
+representations of the true Milky Way in other ways as well,\footnote{The
+failure of cosmological simulations to exactly reproduce the Milky Way is not
+entirely due to limitations of the numerical model. Candidate Milky Way-like
+galaxies are chosen solely on their mass and isolation, for which there are a
+wide variety of possible galaxies with qualitatively different properties.} as
+discussed in \citet{2018arXiv180610564S}. For instance, the velocity structure
+of \mi{} is closer to M31's than the Milky Way's (Loebman et al. in prep).
+However, in this work we are most interested in the global properties of the
+potential, and specifically in deviations from axisymmetry. From this
+perspective, the simulated galaxies are actually \emph{more} axisymmetric than
+we might expect of the Milky Way. While they have prominent spiral arms, none
+has as strong a bar as the Milky Way does at present day, and none has a
+nearby companion like the Large Magellanic Cloud. One of the three we consider
+(\mf) does have an ongoing interaction with a satellite galaxy similar to
+Sagittarius, which has punched through the galactic disk outside the solar
+circle and induced some warping.
 
@@
 center of mass velocity is then determined by all star particles within
-$15\,\kpc$ of this center. The galaxy is then rotated onto a
-principal axis frame determined by stars younger than $1\,\Gyr$ inside of the
-fiducial solar radius $R_{0} = 8.2\,\kpc$, such that the disk plane is the
-$x$--$y$ plane.
+$15\,\kpc$ of this center. The galaxy is then rotated onto a principal axis
+frame determined by stars younger than $1\,\Gyr$ inside of the fiducial solar
+radius $R_{0} = 8.2\,\kpc$, such that the disk plane is the $x$--$y$ plane.
 
@@
 \label{ssec:sag_sim}
-In addition to the cosmological zoom-ins, we will also briefly consider results
-from a live N-body simulation of a Sagittarius-like encounter. This simulation offers us the
-ability to see how the midplane varies in a more controlled environment. The
-details of the simulation are given in \citet{2018MNRAS.481..286L}, but we
-briefly summarize the most relevant details here. For the Milky Way, the dark
-halo is modeled as a Hernquist sphere of mass $10^{12}\,\Msun$ and scale
-length of $52\,\kpc$, the disk modeled as an exponential disk with a scale
-radius of $3.5\,\kpc$, scale height $0.53\,\kpc$, and mass
-$6\times10^{10}\,\Msun$, and a Hernquist bulge of mass $10^{10}\,\Msun$ and scale
-length $0.7\,\kpc$ \citep{1990ApJ...356..359H}. We use the L2 model, for which Sagittarius is modeled as a
-dark matter Hernquist sphere of mass $8\times10^{10}\,\Msun$ and scale length
-$8\,\kpc$, along with a stellar component modeled as a Hernquist sphere
-of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$. All components
-are realized with live distributions of N-body particles; the Milky Way and Sagittarius are
-each initialized to be in equilibrium in isolation \Gus{check!}.
+In addition to the cosmological zoom-ins, we will also briefly consider
+results from a live N-body simulation of a Sagittarius-like encounter. This
+simulation offers us the ability to see how the midplane varies in a more
+controlled environment. The details of the simulation are given in
+\citet{2018MNRAS.481..286L}, but we briefly summarize the most relevant
+details here. For the Milky Way, the dark halo is modeled as a Hernquist
+sphere of mass $10^{12}\,\Msun$ and scale length of $52\,\kpc$, the disk
+modeled as an exponential disk with a scale radius of $3.5\,\kpc$, scale
+height $0.53\,\kpc$, and mass $6\times10^{10}\,\Msun$, and a Hernquist bulge
+of mass $10^{10}\,\Msun$ and scale length $0.7\,\kpc$
+\citep{1990ApJ...356..359H}. We use the L2 model, for which Sagittarius is
+modeled as a dark matter Hernquist sphere of mass $8\times10^{10}\,\Msun$ and
+scale length $8\,\kpc$, along with a stellar component modeled as a Hernquist
+sphere of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$. All
+components are realized with live distributions of N-body particles; the Milky
+Way and Sagittarius are each initialized to be in equilibrium in isolation
+\Gus{check!}.
 
@@
 
-The fiducial coordinate system for these N-body simulations is the rest frame of
-the aligned host galaxy at the beginning of the simulation.
+The fiducial coordinate system for these N-body simulations is the rest frame
+of the aligned host galaxy at the beginning of the simulation.
 
@@
 $z$-coordinate is perpendicular to the disk plane, we place our imaginary
-observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and vary the azimuth
-between $0<\phi<2\pi$. At
-each value of $\phi$ we then compute the median $z$ for stars within a
-cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to the
-fiducial disk. We then re-define the new midplane of the cylinder to be the
-median $z$, re-select stars, and iterate until the median $z$ value converges.
-We find that only $10$ iterations of this procedure are necessary for
-convergence. The resulting median $z$ is taken to be what our observer would
-measure as the local Galactic midplane at each $\phi$.
+observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and
+vary the azimuth between $0<\phi<2\pi$. At each value of $\phi$ we then
+compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and
+height $1\,\kpc$ perpendicular to the fiducial disk. We then re-define the new
+midplane of the cylinder to be the median $z$, re-select stars, and iterate
+until the median $z$ value converges. We find that only $10$ iterations of
+this procedure are necessary for convergence. The resulting median $z$ is
+taken to be what our observer would measure as the local Galactic midplane at
+each $\phi$.
 
@@
 ($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} (left,
-center, and right panels) as a function of azimuthal angle. All calculations are done at
-cosmological redshift $\z =0$. The local midplane is determined at a position
-$\phi$ by taking the median height of all stars within $R=0.5\,\kpc$ and
-$z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
-using the new height $10$ times to converge on the local midplane height. In
-order to allow for the possibility that the fiducial galactocentric coordinate
-system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
-each panel --- this figure is reproduced with the original midplane
-determination (i.e. before subtracting the best fit sine curve) in
-Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
-within a $2\,\kpc$ height of the fiducial midplane to determine $1\,\sigma $
-error bars, which we report as dashed lines.}
+center, and right panels) as a function of azimuthal angle. All calculations
+are done at cosmological redshift $\z =0$. The local midplane is determined at
+a position $\phi$ by taking the median height of all stars within
+$R=0.5\,\kpc$ and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is
+performed again using the new height $10$ times to converge on the local
+midplane height. In order to allow for the possibility that the fiducial
+galactocentric coordinate system is incorrect, we subtract the best fit
+$A\sin{(\phi+B)}+C$ curve from each panel --- this figure is reproduced with
+the original midplane determination (i.e. before subtracting the best fit sine
+curve) in Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on
+all stars within a $2\,\kpc$ height of the fiducial midplane to determine
+$1\,\sigma $ error bars, which we report as dashed lines.}
 \label{fig:midplane}
@@
 \caption{The local midplane determined at the fiducial Solar radius
-($8.2\,\kpc$) for 4 different time steps from live N-body simulations
-of a Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
+($8.2\,\kpc$) for 4 different time steps from live N-body simulations of a
+Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
 before, we have subtracted the best fit $A\sin{(\phi+B)}+C$ curve to account
 for inaccuracies in the galactocentric coordinate system. Error bars are
-calculated as in Figure~\ref{fig:midplane}. (\Gus{Check:}) The upper
-panels show the midplane as a function of azimuth before the first encounter
-near the Solar circle at $t=2.0\,\Gyr$ and $t=4.0\,\Gyr$. As later encounters occur which bring the
-Sagittarius-like object closer to the Solar circle, we can see that stronger
-midplane variations result. In fact, one can see that an $m=2$ mode is excited
-in the lower left panel at $t=6.0\,\Gyr$ --- this is consistent with Figure~17 of
-\citet{2018MNRAS.481..286L}, which shows that an $m=2$ mode is excited (note
-that $m=0$ and $m=1$ modes are stronger, but these are removed in our
-sine-curve subtraction). The fact that the $t=6.9\,\Gyr$ panel looks
-qualitatively similar to the panels from the FIRE simulations
+calculated as in Figure~\ref{fig:midplane}. (\Gus{Check:}) The upper panels
+show the midplane as a function of azimuth before the first encounter near the
+Solar circle at $t=2.0\,\Gyr$ and $t=4.0\,\Gyr$. As later encounters occur
+which bring the Sagittarius-like object closer to the Solar circle, we can see
+that stronger midplane variations result. In fact, one can see that an $m=2$
+mode is excited in the lower left panel at $t=6.0\,\Gyr$ --- this is
+consistent with Figure~17 of \citet{2018MNRAS.481..286L}, which shows that an
+$m=2$ mode is excited (note that $m=0$ and $m=1$ modes are stronger, but these
+are removed in our sine-curve subtraction). The fact that the $t=6.9\,\Gyr$
+panel looks qualitatively similar to the panels from the FIRE simulations
 (Figure~\ref{fig:midplane}) is evidence that the midplane variations are
@@
 
-To account for the effect of particle noise, we bootstrap resample
-stars within a cylinder of stars with height $2\,\kpc$ and the same radius
-$1000$ times and determine the $1\,\sigma$ error bars by repeating the
-midplane determination with those reselection of stars.
+To account for the effect of particle noise, we bootstrap resample stars
+within a cylinder of stars with height $2\,\kpc$ and the same radius $1000$
+times and determine the $1\,\sigma$ error bars by repeating the midplane
+determination with those reselection of stars.
 
@@
 each galaxy, using their local ``solar neighborhood'' (the cylinder defined
-above).
-We consider $50$ equally spaced bins in azimuth, a sufficient number such that each cylinder shares no stars with its neighboring cylinder.
-The $1\,\sigma$ error from the bootstrap procedure is shown as the
-dashed-line error bars. The middle $90\%$ of midplane values across the Solar
-circle spans $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of
-the three cases the midplane therefore varies by more than $\pm 100\,\pc$
-depending on the azimuth along the Solar circle; in the third (\mm{}, which
-has the thinnest ``thin disk'' of stars, but the largest stellar mass) the
-variation is closer to $\pm 50\,\pc$.
+above). We consider $50$ equally spaced bins in azimuth, a sufficient number
+such that each cylinder shares no stars with its neighboring cylinder. The
+$1\,\sigma$ error from the bootstrap procedure is shown as the dashed-line
+error bars. The middle $90\%$ of midplane values across the Solar circle spans
+$185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the three cases
+the midplane therefore varies by more than $\pm 100\,\pc$ depending on the
+azimuth along the Solar circle; in the third (\mm{}, which has the thinnest
+``thin disk'' of stars, but the largest stellar mass) the variation is closer
+to $\pm 50\,\pc$.
 
 We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
-but for four succcessive timesteps of the live N-body simulation of a Sagittarius encounter
-\citep{2018MNRAS.481..286L}. Again we have subtracted a best fit curve of the
-form $\text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}$, with the values
-for $\text{A}$ being $9.5$, $2.5$, $-212$, and $-394\,\pc$, for $\text{B}$
-being $0.0013$, $0.0007$, $-0.63$, and $-1.00\,\text{rad}$ and $\text{C}$
-being $2.6$, $-7.8$, $-65$, and $-53\,\pc$ chronologically for each of the four timesteps shown.
-
-In the upper panels, we see that the midplane is
-relatively flat in the inner galaxy, but additional encounters drive strong
-midplane variation. In the lower left panel, we see a strong $m=2$ mode develop,
-consistent with the $R=8\,\kpc$ panel of Figure~17 in
-\citet{2018MNRAS.481..286L}.
-The lower right panel, which shows the galaxy after some relaxation has occured is qualitatively similar to the midplane variations we saw in the FIRE simulations (Figure~\ref{fig:midplane}), evidence that they are at least partially driven by mergers.
+but for four succcessive timesteps of the live N-body simulation of a
+Sagittarius encounter \citep{2018MNRAS.481..286L}. Again we have subtracted a
+best fit curve of the form $\text{A} \sin{\left(\phi + \text{B}\right)} +
+\text{C}$, with the values for $\text{A}$ being $9.5$, $2.5$, $-212$, and
+$-394\,\pc$, for $\text{B}$ being $0.0013$, $0.0007$, $-0.63$, and
+$-1.00\,\text{rad}$ and $\text{C}$ being $2.6$, $-7.8$, $-65$, and $-53\,\pc$
+chronologically for each of the four timesteps shown.
+
+In the upper panels, we see that the midplane is relatively flat in the inner
+galaxy, but additional encounters drive strong midplane variation. In the
+lower left panel, we see a strong $m=2$ mode develop, consistent with the
+$R=8\,\kpc$ panel of Figure~17 in \citet{2018MNRAS.481..286L}. The lower right
+panel, which shows the galaxy after some relaxation has occured is
+qualitatively similar to the midplane variations we saw in the FIRE
+simulations (Figure~\ref{fig:midplane}), evidence that they are at least
+partially driven by mergers.
 
 \subsection{Velocity Variations} \label{ssec:lsr_var}
-We also expect that the LSR should vary as a function
-of azimuth. We perform this calculation in Appendix~\ref{app:lsr} to estimate
-the components of the LSR as a function of azimuth, but performing a best-fit
-subtraction to correct for misalignment of the original coordinate system (as
-in the previous section) is more involved. Since we find that the variation in
-the LSR is less pronounced than for the midplane, and since offsets in
-velocity only contribute to second order to $\Delta J_R$ and $\Delta J_z$ when a star is at
+We also expect that the LSR should vary as a function of azimuth. We perform
+this calculation in Appendix~\ref{app:lsr} to estimate the components of the
+LSR as a function of azimuth, but performing a best-fit subtraction to correct
+for misalignment of the original coordinate system (as in the previous
+section) is more involved. Since we find that the variation in the LSR is less
+pronounced than for the midplane, and since offsets in velocity only
+contribute to second order to $\Delta J_R$ and $\Delta J_z$ when a star is at
 maximum amplitude in $R$ or $z$ (where the majority of the orbit is, see
@@
 
-While this is not in itself surprising or new,
-we also demonstrated that extending the coordinate system established by our
-local midplane to a global axisymmetric coordinate system spanning the
-entirety of the Galactic disk introduces systematic error in computations of
-the actions under this symmetry assumption, when starting from the present-day
-positions and velocities of stars as measured by, e.g., \emph{Gaia}.
-
-These
-systematic errors are most important for stars on thin disk-like orbits, where
-they can be large enough to yield actions representative of orbits in the
-thick disk. This effect is entirely due to the extension of a local to a
+While this is not in itself surprising or new, we also demonstrated that
+extending the coordinate system established by our local midplane to a global
+axisymmetric coordinate system spanning the entirety of the Galactic disk
+introduces systematic error in computations of the actions under this symmetry
+assumption, when starting from the present-day positions and velocities of
+stars as measured by, e.g., \emph{Gaia}.
+
+These systematic errors are most important for stars on thin disk-like orbits,
+where they can be large enough to yield actions representative of orbits in
+the thick disk. This effect is entirely due to the extension of a local to a
 global coordinate system, and is separate from real diffusion in stellar
@@
 generated by a small error in $R_0$ can be understood from the epicyclic
-approximation, which states that $\Delta J_R/J_R = 2\Delta R/A_R$. The fractional
-error in $J_R$ is related to the error in $R_0$ as a fraction of the {\em
-radial amplitude} of the orbit, which is much smaller than $R_0$
+approximation, which states that $\Delta J_R/J_R = 2\Delta R/A_R$. The
+fractional error in $J_R$ is related to the error in $R_0$ as a fraction of
+the {\em radial amplitude} of the orbit, which is much smaller than $R_0$
 ($\sim1.2\,\kpc$ for the thin-disk and thick-disk orbits we considered). This
@@
 \caption{The three orbits presented in Table~\ref{tab:orbits} and considered
-throughout the work. We plot the thin-disk, thick-disk, and halo orbits in the left, center, and right columns, respectively. The upper
-row shows a plot of $x$ vs. $y$ while the lower row shows $R$ vs. $z$.}
+throughout the work. We plot the thin-disk, thick-disk, and halo orbits in the
+left, center, and right columns, respectively. The upper row shows a plot of
+$x$ vs. $y$ while the lower row shows $R$ vs. $z$.}
 \label{fig:plot_orbits}
@@
 \end{center}
-\caption{A histogram of the observed values in $J_{\phi}$ for the thick-disk (top
-panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
+\caption{A histogram of the observed values in $J_{\phi}$ for the thick-disk
+(top panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
 $100\,\pc$. The induced error distribution is non-Gaussian and bimodal, with

### 2019-05-08 paper/ms.tex
Add start of GC table

@@
 \usepackage{mathtools}
+\usepackage{makecell}
 
@@
 \section{Coordinate System Offsets I. Measurement Errors} \label{sec:mes_err}
+
+\begin{deluxetable*}{cccccccccc}
+\tablecaption{The highest-precision measurement of the parameters that define the Galactocentric coordinate system, and the error in each action that results for the thin-disk and thick-disk orbits we considered.
+\label{tab:orbits}}
+\tablehead{\colhead{name} & \colhead{description} & \colhead{value} & \colhead{uncertainty}
+& \multicolumn{3}{c}{thin-disk orbit} & \multicolumn{3}{c}{thick-disk orbit} \\
+ \colhead{ } & \colhead{ } & \colhead{ } & \colhead{ } & \colhead{$J_R$} & \colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$J_R$} & \colhead{$J_{\phi}$} & \colhead{$J_z$} }
+\startdata 
+$R_0$ & \makecell{distance to \\ Galactic center} & $8.178\,\kpc$ & $0.035\,\kpc$ ($0.3\%$) & $5.4\%$ & $0.3\%$ & \nodata & $5.9\%$ & $0.3\%$ & \nodata \\
+$z_{\odot}$ & \makecell{solar height \\ above midplane} & $20.8\,\pc$ & $0.3\,\pc$ ($1.4\%$) & \nodata & \nodata & $0.5\%$ & \nodata & \nodata & $0.04\%$
+\enddata
+\end{deluxetable*}
+
 We consider the first potential source of coordinate system offsets which

### 2019-05-08 paper/ms.tex
Remove bar buckling

@@
 ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly
-indicating either local substructure infall \citep{2018MNRAS.481.1501B,
-2018arXiv180800451L} or bar buckling \citep{2018arXiv181109205K}, and a gap
+indicating local substructure infall \citep{2018MNRAS.481.1501B,
+2018arXiv180800451L}, and a gap
 suggestive of perturbation by a dark matter subtructure in the tidal stream

### 2019-05-08 paper/ms.tex
Small edits throughout

@@
 \citep[e.g.][]{1915MNRAS..76...70J, 1985ApJ...295..388V, 2017ApJ...839...61T},
-or for associating stars with similar dynamical properties (e.g. to determine
-membership in moving groups).
+or for associating stars with similar dynamical properties, e.g., to
+potentially determine membership in moving groups.
 
@@
 the system's symmetry. However, the Galactic potential is not strictly
-axisymmetric; with the vast improvement in the quality of phase-space
-measurements due to \textit{Gaia}, this assumption appears increasingly
-inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this assumption were
-close enough for many purposes, the parameters used in axisymmetric models may
+axisymmetric. This has been known for some time \citep[e.g.][]{1957AJ.....62...93K,2009MNRAS.396L..56M, 2012ApJ...750L..41W}, but with the vast improvement in the quality of phase-space
+measurements due to \textit{Gaia} this assumption is increasingly
+inadequate \citep[e.g.][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. Even if this assumption were
+close enough for many purposes, the parameters used in axisymmetric models of the Galactic potential may
 be inadequately constrained by current observations.
@@
 actions, especially for stars in the thin disk. Our main point is that the
-``local'' Galactic midplane is not the same as the ``global'' Galactic
-midplane. For stars relatively far from the Sun, their local midplane location
+``local'' and ``global'' Galactic
+midplanes are not the same. For stars relatively far from the Sun, their local midplane
 will not agree with our local midplane extrapolated onto their position due to
@@
 distribution. Converting positions and velocities of more distant stars from a
-heliocentric to a Galactocentric coordinate system introduces a systematic
-bias in the $z$ coordinate, and hence the actions computed for these stars.
+heliocentric to a Galactocentric coordinate system thus introduces a systematic
+bias in the $z$ coordinate, and consequently the actions computed for these stars.
 The further from the solar neighborhood the target star, the more likely the
@@
 star at two different orbital phases (teal and orange points). If the observer
-were then to integrate the star's orbit using a model potential with the
+were to then integrate the star's orbit using a model potential with the
 erroneous midplane, they would obtain the teal and orange curves for the
 star's orbit, respectively. The actions estimated from these two erroneous
-orbits would subsequently differ, both from each other and from the actions
-estimated for the true orbit (in the potential with the correct midplane).
+measurements would subsequently differ, both from each other and from the true measurement (in the potential with the correct coordinate system).
 Hence an incorrect midplane in the potential model assumed will induce
@@
 We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
-inaccurate or erroneous determination of the midplane (and hence the
-zero-point of the $z$ coordinate) leads to a phase-dependence of the actions
+inaccurate or erroneous determination of the midplane leads to a phase-dependence of the actions
 calculated from an orbit specified by a phase-space point and an assumed
@@
 of thin disk-like orbits and a good description of thick disk-like orbits in
-an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and
-spiral arms). We also make the assumption of a perfectly flat circular
+an axisymmetric potential, i.e., ignoring the influence of the Galactic bar and
+spiral arms. We also make the assumption of a perfectly flat circular
 velocity curve with velocity $v_c$, a good approximation near the Solar circle
@@
 three orbits in the model potential described in
-Section~\ref{ssec:action_comp} that are typical of stars in the stellar thin
-disk, stellar thick disk, and the stellar halo. We summarize their initial
+Section~\ref{ssec:action_comp} that are typical of stars in the thin
+disk, thick disk, and halo. We summarize their initial
 positions in phase space and the actions computed by integrating their orbits
@@
 
-We consider that the immortal observer makes a measurement every $\Myr$. At
+We consider that the observer makes a measurement every $\Myr$. At
 each measurement, we compute the actions as outlined above using the true
@@
 offset coordinate systems depend on the time (i.e. orbital phase) at which the
-star's phase-space position is observed. To the immortal observer this would
-appear as time-dependence in the actions, which appear to oscillate around
-their value using the correct coordinate system even though in this example
+star's phase-space position is observed. To the observer this would
+appear as time-dependence in the actions, oscillating around
+their value using the correct coordinate system. This time dependence comes even though
 the observer is using the correctly constructed best-fit static, axisymmetric
@@
 the direction of the systematic offset as well as the true values of the
-actions (i.e. the type of orbit). In reality we will have one measurement of
+actions (i.e. the type of orbit). In reality, we will have one measurement of
 the phase-space position to work with, in which case the determination of the
@@
 half the 95\uth{} percentile minus the 5\uth{} percentile of the distribution
-of observed action values. We refer to this quantity as $\Delta J_i$ for each
-action value. We plot this quantity in Figure~\ref{fig:Jz_hist} as a
+of observed action values. We plot this quantity, which we refer to as $\Delta J_i$ for each action, in Figure~\ref{fig:Jz_hist} as a
 horizontal arrow anchored on the true action value. Because of the bimodality

### 2019-05-08 paper/ms.tex
Move section 3 around, add couple sentences

@@
 
-\section{Coordinate System Offsets I. Measurement Errors} \label{sec:mes_err}
-We consider the first potential source of coordinate system offsets which
-would generate the action errors described in Section~\ref{sec:ref_frame}:
-uncertainties in measurements of the position and velocity of the Sun relative
-to the Galactic center. Converting from heliocentric to Galactocentric
-coordinates relies on these measurements. Therefore, errors in their values
-will induce a systematic offset in the Galactocentric phase-space position of
-any observed star.
-
-\subsection{Galactic Center}
-First, one must define the center of the Galaxy. This is usually taken to be
-the location of the central supermassive black hole, Sagittarius~A\textsuperscript{*}
-\citep[\sgra{}, e.g.][]{2004ApJ...616..872R}. From stellar motions near
-\sgra{}, the distance from the Sun to \sgra{}, referred to as $R_0$, can be
-precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a
-recent measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
-\citep{2019arXiv190405721A}.
-
-\subsection{Galactic Orientation}
-Second, one must define the angular orientation of the Galaxy. This was
-defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
-defining the coordinates of the Galactic center in B1950 coordinates as
-(17:42:26.6, -28:55:00) and the North Galactic pole as (12:49:00, +27:24:00).
-These two quantities, together with $R_0$, define the orientation of the
-Galactic plane. However, there is growing evidence that the stellar midplane
-is tilted relative to this coordinate system \citep{2014ApJ...797...53G,
-2016ARAA..54..529B}, though not the H~\textsc{ii} midplane
-\citep{2019ApJ...871..145A}.
-
-\subsection{Solar Height}
-Third, one must define the Sun's vertical distance from the Galactic midplane,
-which can be determined by identifying where the stellar density and
-velocities reach a maximum (effectively the median height of all disk stars).
-The Solar height is usually taken to be $\sim 25\,\pc$
-\citep{2001ApJ...553..184C}, with a more recent measurement from \textit{Gaia}
-DR2 placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another
-strategy is to use the cold gas or H~\textsc{ii} regions in the disk to define
-the Galactic midplane, leading to slightly different values (by $\sim 5\,\pc$)
-for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em
-Gaia} review of these measurements is given by \citet{2016ARAA..54..529B}.
-
-\subsection{Local Standard of Rest}
-Finally, one must define the LSR, or mean velocity of stars near the Sun
-relative to the Galactic center (which is defined to have zero velocity), and
-the velocity of the Sun relative to the LSR. The radial ($U_{\odot}$) and
-vertical ($W_{\odot}$) components are computed by taking the mean motions of
-different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal
-component ($V_{\odot}$) is more difficult to measure, but can be modeled using
-the asymmetric drift relation \citep{2008gady.book.....B}. The values of the
-components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}. The
-value of the circular velocity is taken to be $\sim 220\kms$
-\citep[e.g.][]{2012ApJ...759..131B}.
-
-\vspace{6pt}
-
-Considerable effort has been placed on each of these measurements, but
-uncertainties remain, and detailed dynamical modeling across the disk ---
-particularly for dynamically cold stars --- may have to take them into
-account. For instance, the discrepancy of $\sim 5\,\pc$ between the midplane
-defined by stars and by gas will induce a $\sim10\%$ error in $J_z$ for an
-orbit with $z_{\text{max}}\sim100\,\pc$ (see Section~\ref{sec:ref_frame}).
-
-\section{Coordinate System Offsets II. Azimuthal Midplane Variations in
+\section{Coordinate System Offsets I. Azimuthal Midplane Variations in
 Simulations}
@@
 
-We now consider a second source of systematic offsets in the Galactocentric
+We first consider one source systematic offsets in the Galactocentric
 coordinate system: intrinsic variations of the midplane across the disk. In
@@
 
+\section{Coordinate System Offsets II. Measurement Errors} \label{sec:mes_err}
+We consider the second potential source of coordinate system offsets which
+would generate the action errors described in Section~\ref{sec:ref_frame}:
+uncertainties in measurements of the position and velocity of the Sun relative
+to the Galactic center. Converting from heliocentric to Galactocentric
+coordinates relies on these measurements. Therefore, errors in their values
+will induce a systematic offset in the Galactocentric phase-space position of
+any observed star.
+
+\subsection{Galactic Center}
+First, one must define the center of the Galaxy. This is usually taken to be
+the location of the central supermassive black hole, Sagittarius~A\textsuperscript{*}
+\citep[\sgra{}, e.g.][]{2004ApJ...616..872R}. From stellar motions near
+\sgra{}, the distance from the Sun to \sgra{}, referred to as $R_0$, can be
+precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a
+recent measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
+\citep{2019arXiv190405721A}.
+
+In the epicyclic approximation (Section~\ref{ssec:epi_action}), the $0.3\%$
+error in $R_0$ induces a $5.4\%$ and $5.9\%$ error in the thin-disk and
+thick-disk orbits, respectively. For $J_{\phi}$, the error is $0.3\%$ for both
+orbits.
+
+\subsection{Galactic Orientation}
+Second, one must define the angular orientation of the Galaxy. This was
+defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
+defining the coordinates of the Galactic center in B1950 coordinates as
+(17:42:26.6, -28:55:00) and the North Galactic pole as (12:49:00, +27:24:00).
+These two quantities, together with $R_0$, define the orientation of the
+Galactic plane. However, there is growing evidence that the stellar midplane
+is tilted relative to this coordinate system \citep{2014ApJ...797...53G,
+2016ARAA..54..529B}, though not the H~\textsc{ii} midplane
+\citep{2019ApJ...871..145A}.
+
+\subsection{Solar Height}
+Third, one must define the Sun's vertical distance from the Galactic midplane, $z_{\odot}$.
+It can be determined by identifying where the stellar density and
+velocities reach a maximum (effectively the median height of all disk stars).
+The Solar height is usually taken to be $\sim 25\,\pc$
+\citep{2001ApJ...553..184C}, with a more recent measurement from \textit{Gaia}
+DR2 placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another
+strategy is to use the cold gas or H~\textsc{ii} regions in the disk to define
+the Galactic midplane, leading to slightly different values (by $\sim 5\,\pc$)
+for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em
+Gaia} review of these measurements is given by \citet{2016ARAA..54..529B}.
+
+Under the epicyclic approximation, the highest precision measurement of
+$z_{\odot}$ of $1.4\%$ from \citet{2019MNRAS.482.1417B} induces a $0.5\%$ and
+$0.04\%$ error in $J_z$ for the thick-disk and thin-disk orbits, respectively.
+The discrepancy between the stellar and gas midplanes of $\sim5\,\pc$ induces
+an $8.3\%$ and $1.2\%$ error in $J_z$ for the two orbits.
+
+\subsection{Local Standard of Rest}
+Finally, one must define the LSR, or mean velocity of stars near the Sun
+relative to the Galactic center (which is defined to have zero velocity), and
+the velocity of the Sun relative to the LSR. The radial ($U_{\odot}$) and
+vertical ($W_{\odot}$) components are computed by taking the mean motions of
+different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal
+component ($V_{\odot}$) is more difficult to measure, but can be modeled using
+the asymmetric drift relation \citep{2008gady.book.....B}. The values of the
+components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}. The
+value of the circular velocity is taken to be $\sim 220\kms$
+\citep[e.g.][]{2012ApJ...759..131B}.
+
+\vspace{6pt}
+
+Considerable effort has been placed on each of these measurements, but
+uncertainties remain, and detailed dynamical modeling across the disk ---
+particularly for dynamically cold stars --- may have to take them into
+account. For instance, the discrepancy of $\sim 5\,\pc$ between the midplane
+defined by stars and by gas will induce a $\sim10\%$ error in $J_z$ for an
+orbit with $z_{\text{max}}\sim100\,\pc$ (see Section~\ref{sec:ref_frame}).
+
 \section{Discussion} \label{sec:discussion} 

### 2019-05-09 paper/ms.tex
changes from last telecon

--- 2019-05-08 paper/ms.tex
@@
 \usepackage{mathtools}
-\usepackage{makecell}
 
@@
 \newcommand{\mm}{\texttt{m12m}}
-
-\newcommand{\sgra}{Sgr~A\textsuperscript{*}}
 
@@
 \title{The Galactic Midplane Is Not a Plane: Implications for Dynamical
-Analysis with \textit{Gaia} Data and Beyond}
-
-% \correspondingauthor{Angus Beane}
+Analysis with {\em Gaia} Data and Beyond}
+
+\correspondingauthor{Angus Beane}
 \email{abeane@sas.upenn.edu}
 
-\author[0000-0002-8658-1453]{Angus Beane}
+\author{Angus Beane}
 \affil{\cca}
@@
 
-\author[0000-0003-3939-3297]{Robyn Sanderson}
+\author{Robyn Sanderson}
 \affil{\cca}
@@
 
-\author[0000-0001-8275-9181]{Douglas Grion Filho}
-\affil{\columbia}
-
-\author[0000-0003-0064-4060]{Mordecai-Mark Mac Low}
+\author{Mordecai-Mark Mac Low}
 \affil{\cca}
@@
 
-\author[0000-0003-2866-9403]{David W. Hogg}
+\author{Douglas Filho}
+\affil{\columbia}
+
+\author{David W. Hogg}
+\affil{\cca}
 \affil{\nyuphys}
 \affil{\nyucds}
-\affil{\cca}
 \affil{\mpia}
@@
 \author{Chervin F. P. Laporte}
-\altaffiliation{CITA National Fellow}
+\altaffiliation{Simons Fellow}
 \affil{\victoria}
@@
 
-Stellar actions, computed for stars from their six-dimensional phase space
-measurements and an assumed Galactic potential, are used to label and
-distinguish orbits. In principle, actions have the attractive quality of being
-invariant and thus good labels of a star's orbit. Such analyses are often used
-to interpret data from the \textit{Gaia} mission. However, inaccurate
-measurements of the phase space position and uncertainties in the assumed
-Galactic potential induce an error in the computed action. We show that, in
-addition to these complications, a systematic bias in the phase space position
-induces a phase-dependence in the actions, which we interpret as a systematic
-error. An offset in the vertical position of $\sim15\,\pc$ for a thin disk
-orbit or $\sim120\,\pc$ for a thick disk orbit is sufficient to induce a
-$25\,\%$ systematic error in the vertical action $J_z$ computed under the
-assumption of an axisymmetric potential. The induced error distribution is
-non-Gaussian and bimodal, with neither of the modes peaking on the null value.
-Furthermore, we show that the local midplane varies by $\sim100\,\pc$ at the
-Solar circle in Milky Way-mass cosmological zoom-in simulations from the FIRE
-project. From observations that the mean vertical velocity of the Milky Way
-disk varies by $\sim5\,\kms$ with radius, we estimate that the true midplane
-variations are of order $\sim60\textup{--}90\,\pc$. Thus, current
-state-of-the-art action calculations --- which assume the global and local
-midplanes are the same --- are likely to include a systematic vertical offset
-depending on the volume considered. Variation in the local standard of rest
-induces similar issues. The variation of the midplane must be taken into
-account when performing dynamical modeling across the large regions of the
-disk accessible to \textit{Gaia} and future missions.
+Stellar actions, computed for stars from their six-dimensional phase space measurements and
+an assumed Galactic potential, are used to label and distinguish orbits. In
+principle, actions have the attractive quality of being invariant and thus
+good labels of a star's orbit. 
+%mm Of course,
+    Such analyses are often used to interpret data from the {\em Gaia}
+    mission. However, 
+inaccurate measurements of the 
+phase space position and uncertainties in the assumed Galactic potential
+induce an error in the computed action. We show that, in addition to these
+complications, a systematic bias in the phase space position induces a
+phase-dependence in the actions, which we interpret as 
+%mm an 
+    a systematic
+error. The induced error distribution is
+non-Gaussian and bimodal, 
+with neither of the modes peaking on
+%mm  the true action value. [the error in a Gaussian peaks on zero] 
+    the null value.
+A variation in the vertical position midplane of $\sim15\,\pc$ for a thin disk orbit or
+$\sim120\,\pc$ for a thick disk orbit is 
+%mm all that is necessary 
+      sufficient
+to induce a $25\,\%$ systematic error in the vertical action $J_z$ computed under the assumption of an axisymmetric potential.
+%mm [fragment?] orbits, respectively. 
+Furthermore, we show that the local midplane varies by
+$\sim100\,\pc$ at the Solar circle in Milky Way-mass cosmological zoom-in
+simulations from the FIRE collaboration. Thus, current state-of-the-art action
+calculations --- which assume the global and local midplanes are the same ---
+include a systematic 
+%mm offset in $z$.
+   vertical offset variation. 
+Variation in the local standard of rest
+%mm will induce 
+     induces
+similar issues. The variation of the midplane must be taken into
+account when performing dynamical modelling across
+the large regions of the disk
+accessible 
+%mm with the {\em Gaia} mission. [moved first mention up to top of
+%abstract, given that this is in the title.]
+    to {\em Gaia}.
 
@@
 
-\keywords{Galaxy: disk -- Galaxy: evolution -- Galaxy: kinematics and dynamics
--- Galaxy: structure -- solar neighborhood -- stars: kinematics and dynamics}
+\keywords{Galaxy:~disk -- solar~neighborhood --
+open~clusters~and~associations:~general -- Galaxy:~structure --
+Galaxy:~evolution}
 
@@
 Our understanding of the Milky Way is currently undergoing a revolution as a
-result of \textit{Gaia} Data Release 2 (DR2).  Recent major discoveries
-include the remnants of a major merger \citep{2018ApJ...860L..11K,
+result of {\em Gaia} Data Release 2 (DR2).  Recent major discoveries include
+the remnants of a major merger \citep{2018ApJ...860L..11K,
 2018Natur.563...85H, 2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space
 ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly
-indicating local substructure infall \citep{2018MNRAS.481.1501B,
-2018arXiv180800451L}, and a gap
+indicating either local substructure infall \citep{2018MNRAS.481.1501B,
+2018arXiv180800451L} or bar buckling \citep{2018arXiv181109205K}, and a gap
 suggestive of perturbation by a dark matter subtructure in the tidal stream
@@
 of the Galactic plane, the height of the Sun above the Galactic midplane, and
-the local standard of rest (LSR). We review the observational efforts to
-measure these parameters in Section~\ref{sec:mes_err}.
+the local standard of rest (LSR). We review the observational efforts to measure these parameters in Section~\ref{sec:mes_err}.
 
@@
 coordinate system. They are the cyclical integral of the canonical momentum
-over its conjugate position:
+over its conjugate position,
 \beq\label{eq:actions}
 J_i \equiv
-\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{.}
+\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,.}
 \eeq
@@
 and azimuthal coordinates respectively in a cylindrical coordinate system and
-the $p_i$ are the conjugate momenta. In a slowly-evolving axisymmetric
-potential, these actions are invariant and $J_{\phi} \equiv L_z$, where $L_z$
-is the $z$-component of the angular momentum
+the $p_i$ are the conjugate momenta. In a slowly-evolving axisymmetric potential, these
+actions are invariant and $J_{\phi} \equiv L_z$
 \citep{2008gady.book.....B,2014RvMP...86....1S}.
@@
 With the advent of 6D phase-space measurements over a relatively large
-($\gtrsim 2$ kpc) volume from the \textit{Gaia} satellite, the study of stars'
+($\gtrsim 2$ kpc) volume from the {\em Gaia} satellite, the study of stars'
 actions has gained new popularity. One reason is dimensionality reduction---an
@@
 function of their actions and independent of the conjugate angles, allowing
-investigation of the relationship between \emph{orbital} properties of stars
+investigation of the relationship between {\em orbital} properties of stars
 and other intrinsic, and at least partially invariant, properties such as age
@@
 authors have used actions to study stellar scattering in the Milky Way disk
-using the improved astrometry of \textit{Gaia} DR2 and various age catalogues
+using the improved astrometry of {\em Gaia} DR2 and various age catalogues
 \citep{2018ApJ...867...31B, 2018arXiv180803278T}. Actions have also been used
@@
 axisymmetric; with the vast improvement in the quality of phase-space
-measurements due to \textit{Gaia}, this assumption appears increasingly
+measurements due to {\em Gaia}, this assumption appears increasingly
 inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this assumption were
@@
 ``local'' Galactic midplane is not the same as the ``global'' Galactic
-midplane. For stars relatively far from the Sun, their local midplane location
-will not agree with our local midplane extrapolated onto their position due to
-relatively small-scale variations in the density of the gas and stellar
-distribution. Converting positions and velocities of more distant stars from a
-heliocentric to a Galactocentric coordinate system introduces a systematic
-bias in the $z$ coordinate, and hence the actions computed for these stars.
-The further from the solar neighborhood the target star, the more likely the
-mismatch will result in large systematic uncertainty, especially in the
-vertical action $J_z$. A similar argument applies to any remaining uncertainty
-in measurements of the Galactocentric radius of the Sun and to variations in
-the LSR.
-
-\begin{figure*}[ht!]
-\begin{center}
-\includegraphics[width=\textwidth]{fig/cartoon.pdf}
-\end{center}
-\caption{Cartoon approximation showing the effect an error in the
-determination of the coordinate midplane can have on orbit integration and
-action estimation. The $x$-axis shows the orbital phase and the $y$-axis the
-vertical height. The top gray curve depicts an example ``true'' orbit
-oscillating about the true midplane (horizontal solid gray line). Consider an
-observer who erroneously assumes the midplane is located at the horizontal
-dashed line. Suppose this observer measures the phase-space position of the
-star at two different orbital phases (teal and orange points). If the observer
-were then to integrate the star's orbit using a model potential with the
-erroneous midplane, they would obtain the teal and orange curves for the
-star's orbit, respectively. The actions estimated from these two  erroneous
-orbits would subsequently differ, both from each other and from the actions
-estimated for the true orbit (in the potential with the correct midplane).
-Hence an incorrect midplane in the potential model assumed will induce
-phase-dependence in the actions estimated for a given star in that potential.}
-\label{fig:cartoon}
-\end{figure*}
+midplane. For stars relatively far from the Sun, the local midplane location
+will not agree with our local midplane extrapolated onto their position,
+due to relatively small-scale variations in the density of the gas and
+stellar distribution. Converting positions and velocities of more distant
+stars from a heliocentric to a Galactocentric coordinate system introduces a
+systematic bias in the $z$ coordinate, and hence the actions computed for
+these stars. The further from the solar neighborhood the target star, the more
+likely the mismatch will result in large systematic uncertainty, especially in
+the vertical action $J_z$. Depending on the type of orbit being studied, this
+sets a limit on the ``dynamical solar neighborhood'' over which the local
+coordinate frame can be usefully extended for action-space analysis. A similar
+argument applies to any remaining uncertainty in measurements of the
+Galactocentric radius of the Sun and to variations in the local
+standard of rest (LSR).
 
@@
 collaboration\footnote{\url{https://fire.northwestern.edu}}
-\citep{2014MNRAS.445..581H, 2016ApJ...827L..23W, 2018MNRAS.480..800H}, and a
-controlled N-body simulation of a Sagittarius encounter with a galaxy
-otherwise tailored to the stellar mass, scale length, and scale height of the
-Milky Way \citep{2018MNRAS.481..286L}.
-
-In Section~\ref{sec:discussion}, we discuss the implications of midplane
+\citep{2016ApJ...827L..23W, 2017arXiv170206148H}, and a controlled N-body
+simulation of a Sagittarius encounter with a galaxy otherwise tailored to the stellar mass, scale length, and scale height of the Milky Way \citep{2018MNRAS.481..286L}.
+
+In Section~\ref{sec:discussion} we discuss the implications of midplane
 variations, and the resulting systematic uncertainty in the vertical action,
-for action-space analyses. We also estimate the expected midplane variations
-of the Milky Way based on the observed velocity variations. We summarize our
-main results and conclude in Section~\ref{sec:conclusion}.
+for action-space analyses beyond the dynamical solar neighborhood. In light of our findings, we considered the
+orbits of the nine open clusters within $250\,\pc$ as reported by the {\em
+Gaia} collaboration \citep{2018AA...616A..10G}. In
+Table~\ref{tab:real_clusters}, we report the distance to each cluster, the
+three actions, and the maximum vertical extent of their orbits
+($z_{\text{max}}$). In Section~\ref{sssec:reconstruction} we estimate and
+discuss the size of the dynamical solar neighborhood for each cluster based on
+its mass, to show that action-space tagging is likely still a viable method if
+contamination from foreground or background stars is unimportant. We summarize
+and conclude in Section~\ref{sec:conclusion}.
+
+\begin{deluxetable*}{cCCCCCCCC}
+\tablecaption{Relevant dynamical parameters for the $9$ open clusters within
+$250\,\pc$ of the Sun, using the relevant initial positions and velocities
+from \citet{2018AA...616A..10G}.
+Actions are computed using the method outlined in Section~\ref{ssec:action_comp}, with typical values for the Galactocentric coordinate system parameters.
+These are only shown to illustrate the general types of orbits open clusters are on.
+The neighborhood column is explained in more
+detail in Section~\ref{sssec:reconstruction}.\label{tab:real_clusters}}
+\tablehead{\colhead{cluster} & \colhead{$N_{\text{memb}}$} & \colhead{distance} & \colhead{$J_R$} &
+\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} & \colhead{mass} \\ 
+\colhead{ } & \colhead{ } &
+\colhead{$\mathrm{pc}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
+\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
+\colhead{$\mathrm{pc}$} & \colhead{$M_{\odot}$}}
+\startdata
+alphaPer & 740 & 170 & 20 & -1700 & 0.005 & 11 & 352\tablenotemark{b} \\
+Blanco1 & 489 & 240 & 2 & -1800 & 2 & 210 & 410\tablenotemark{c} \\
+ComaBer & 153 & 90 & 2 & -1900 & 0.7 & 140 & 112\tablenotemark{d} \\
+Hyades & 515 & 50 & 20 & -1800 & 0.3 & 80 & 400\tablenotemark{e} \\
+IC2391 & 325 & 150 & 6 & -1800 & 0.02 & 20 & \nodata \\
+IC2602 & 492 & 150 & 10 & -1700 & 0.3 & 89 & \nodata \\
+NGC2451A\tablenotemark{a} & 400 & 190 & 6 & -1800 & 0.2 & 73 & \nodata \\
+Pleiades & 1326 & 140 & 20 & -1700 & 0.4 & 96 & 800\tablenotemark{f} \\
+Praesepe & 938 & 190 & 20 & -1800 & 0.6 & 120 & 550\tablenotemark{d}
+\enddata
+
+\tablenotetext{a}{This cluster is labeled NGC2451, which we interpret as
+NGC2451A since NGC2451B lies further than $250\,\pc$ away.}
+\tablenotetext{b}{\citet{2016MNRAS.457.1028S}}
+\tablenotetext{c}{\citet{2007AA...471..499M}}
+\tablenotetext{d}{\citet{2007AJ....134.2340K}}
+\tablenotetext{e}{\citet{1998AA...331...81P}}
+\tablenotetext{f}{\citet{2001AJ....121.2053A}}
+
+\tablerefs{\citet{2018AA...616A..10G}}
+\end{deluxetable*}
 
 \section{Motivation} \label{sec:ref_frame}
-We first demonstrate the significance  to action computations of a systematic
-offset in the determination of the Galactic midplane or distance to the
-Galactic center. We will find that such offsets are especially important for
-disk-like orbits. The consequences we explore here may also arise from various
-other systematic errors. For instance, the axisymmetric Galactic potential
-model used in many works to compute actions may not be a good description of
-the true potential --- or the parameters used may yield a potential that is
-systematically incorrect outside an original fitted region. In this work, we
-assume that the Galaxy is perfectly described by our model axisymmetric
-potential, and simply explore the consequences of offsets in the determination
-of the Galactocentric coordinate system.
+We first demonstrate the significance  to
+action computations of a systematic offset in the
+determination of the Galactic midplane or distance to the Galactic center. We will find that such offsets are especially important
+for disk-like orbits. The consequences we explore here may also arise from
+various other systematic errors. For instance, the axisymmetric Galactic
+potential model used in many works to compute actions may not be a good
+description of the true potential --- or the parameters used may yield a
+potential that is systematically incorrect outside an original fitted region.
+In this work, we assume that the Galaxy is perfectly described by our model
+axisymmetric potential, and simply explore the consequences of offsets in the determination of the Galactocentric coordinate system.
 
@@
 We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
-inaccurate or erroneous determination of the midplane (and hence the
-zero-point of the $z$ coordinate) leads to a phase-dependence of the actions
-calculated from an orbit specified by a phase-space point and an assumed
-potential model. The $y$-axis corresponds to the vertical height of the orbit
-and the $x$-axis shows orbital phase. The solid gray curve indicates the
+inaccurate or erroneous determination of the midplane (and hence the zero-point of the $z$ coordinate) leads to a phase-dependence of the
+actions calculated from an orbit specified by a phase-space point and an
+assumed potential model. The $y$-axis corresponds to the vertical height of
+the orbit and the $x$-axis shows orbital phase. The solid gray curve indicates the
 ``true'' orbit of the star; that is, its orbit as it oscillates around the
-true midplane. The dashed gray line is the midplane location in the potential
-model used by an observer to integrate the orbit of the star and hence
-estimate its actions, which is offset from the true midplane of the star's
-orbit. The model potential is otherwise identical to the one in which the star
-is actually moving.
-
-Now suppose this observer makes a measurement of the star's position and
-velocity at the teal point or the orange point (i.e. at two different orbital
-phases). Then the teal and orange curves correspond to the orbits that the
-observer would compute for each point based on the potential model with the
-offset midplane. In action space, this would correspond to a different value
-of $J_z$ for the teal and orange points. In this way, assuming the wrong
-coordinate system induces a phase-dependence in the actions estimated for the
-star, which in the correct potential (in this example, the one with the
-correct midplane) should be phase-independent.
+true midplane. The dashed gray line is the midplane location in the
+potential model used to integrate the orbit of the star and
+hence estimate its actions, which is offset from the true midplane of the star's orbit. The model potential is otherwise identical to the one in which the star is actually moving.
+
+Now suppose this observer makes a measurement of the star's position and velocity at the purple
+point or the green point (i.e. at two different orbital phases). Then the
+purple and green curves correspond to the orbits that the observer would
+compute for each point based on the potential model with the offset
+midplane. In action space, this would correspond to a different value of $J_z$
+for the purple and green points. In this way, assuming the wrong coordinate
+system induces a phase-dependence in the actions estimated for the star, which
+in the correct potential (in this example, the one with the correct midplane)
+should be phase-independent.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{fig/cartoon.pdf}
+\end{center}
+\caption{Cartoon approximation showing the effect an error in the determination of the
+coordinate midplane can have on orbit integration and action estimation. The
+$x$-axis shows the orbital phase and the $y$-axis the vertical height. The top
+gray curve depicts an example ``true'' orbit oscillating about the true
+midplane (horizontal solid gray line). Consider an observer who erroneously
+assumes the midplane is located at the horizontal dashed line. Suppose this
+observer measures the phase-space position of the star at two different
+orbital phases (purple and green points). If the observer were then to
+integrate the star's orbit using a model potential with the erroneous
+midplane, they would obtain the purple and green curves for the star's orbit,
+respectively. The actions estimated from these two  erroneous orbits would
+subsequently differ, both from each other and from the actions estimated for
+the true orbit (in the potential with the correct midplane). Hence an
+incorrect midplane in the potential model assumed will induce phase-dependence
+in the actions estimated for a given star in that potential.}
+\label{fig:cartoon}
+\end{figure*}
 
 \subsection{Epicyclic Approximation} \label{ssec:epi_action}
+Before turning to numerical methods, we first write down analytically the systematic error in the actions induced from either a midplane offset or a Galactocentric radius
+offset. We use the epicyclic approximation, which assumes that the motion in
+the $z$ and $R$ components of the orbit are decoupled and follow simple
+harmonic motion about a circular and planar ``guiding orbit'' \citep[][Section 3.2 and references therein]{2008gady.book.....B}.
+We refer to the radius of this orbit (the ``guiding radius'') as $R_g$.
+This approximation is an excellent description of
+thin disk-like orbits and a good description of thick disk-like orbits in an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and spiral arms). We
+also make the assumption of a perfectly flat circular velocity curve with
+velocity $v_c$ \citep[decent near the Solar circle, e.g.][]{2017MNRAS.465...76M}.
+
+Under this approximation, we can write down the cylindrical components
+of the orbits as,
+\beq\label{eq:orbits_epi}
+\begin{split}
+\phi(t) &= \Omega_c t \\
+R(t) &= R_g + A_R \sin{(\nu t + \beta)} \\
+z(t) &= A_z \sin{(\kappa t + \alpha)}
+\text{,}
+\end{split}
+\eeq
+where $\kappa$ and $\nu$ are the vertical and radial (or epicyclic)
+frequencies, $\Omega_c \equiv v_c/R_g$ the rotation frequency, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
+vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
+constants. Similarly, the velocities of the orbit are given by
+\beq\label{eq:orbits_vel_epi}
+\begin{split}
+v_{\phi}(t) &= v_c \\
+v_R(t) &= A_R \nu \cos{(\nu t + \beta)} \\
+v_z(t) &= A_z \kappa \cos{(\kappa t + \alpha)}
+\text{.}
+\end{split}
+\eeq
+
+In this case, we have that for $J_{\phi}$ \citep[][Section~3.5.3b]{2008gady.book.....B}, 
+\beq\label{eq:Jphi_epi}
+J_{\phi} = R_g v_c\text{,}
+\eeq
+and for $J_R$ and $J_z$,
+\beq\label{eq:JR_Jz_epi}
+\begin{split}
+J_R &= \frac{E_R}{\kappa} \\
+J_z &= \frac{E_z}{\nu} \text{,}
+\end{split}
+\eeq
+where $E_R$ and $E_z$ are the energy per unit mass in the radial and vertical
+coordinates, respectively. Since $a_R = -\kappa^2(R-R_g)$, and $a_z = - \nu^2
+z$, we have that,
+\beq\label{eq:JR_Jz_epi_energy}
+\begin{split}
+J_R &= \frac{\frac{1}{2}v_R^2 + \kappa^2 (R-R_g)^2}{\kappa} \\
+J_z &= \frac{\frac{1}{2}v_z^2 + \nu^2 z^2}{\nu}
+\end{split}
+\eeq
+
+While the formula for $J_{\phi}$ is phase independent, $J_R$ and $J_z$ are
+both phase dependent via Eq. \eqref{eq:orbits_epi}. We consider two cases. First, at the point of maximum
+amplitude or when $(v_R, v_z) = (0, 0)$ and second, at the point of maximum
+velocity $(R, z) = (R_g, 0)$. For $J_R$ we have that,
+\beq\label{eq:JR_epi_final}
+    \begin{aligned}
+        J_R &= \kappa A_R^2 && \text{if $v_R = 0$} \\
+    J_R &= \frac{v_{R,\text{max}}^2}{2\kappa^2} && \text{if $R=R_g$} \text{.}
+    \end{aligned}
+\eeq
+For $J_z$, we have that,
+\beq\label{eq:Jz_epi_final}
+    \begin{aligned}
+        J_z &= \kappa A_z^2 && \text{if $v_z = 0$} \\
+        J_z &= \frac{v_{z,\text{max}}^2}{2\kappa^2} && \text{if $z=0$} \text{.}
+    \end{aligned}
+\eeq
+
+Now assume that the coordinates $(R, z, v_{\phi}, v_R, v_z)$ are offset by
+$(\Delta R, \Delta z, \Delta v_{\phi}, \Delta v_R, \Delta v_z)$. Then, the
+induced error in $J_{\phi}$ is, regardless of phase,
+\beq\label{eq:induced_Jphi}
+\frac{\Delta J_{\phi}}{J_{\phi}} = \frac{\Delta R}{R_g}\text{.}
+\eeq
+For $J_R$, the induced error is,
+\beq\label{eq:induced_JR}
+\begin{aligned}
+    \frac{\Delta J_R}{J_R} &= \frac{2\Delta R}{A_R} + 
+        \left(\frac{\Delta v_R}{v_{R,\text{max}}}\right)^2 && \text{if $v_R = 0$} \\
+    \frac{\Delta J_R}{J_R} &= \frac{2\Delta v_R}{v_{R,\text{max}}} + 
+        \left(\frac{\Delta R}{A_R}\right)^2  && \text{if $R=R_g$} \text{.}
+\end{aligned}
+\eeq
+For $J_{z}$, the induced error is,
+\beq\label{eq:induced_Jz}
+\begin{aligned}
+    \frac{\Delta J_z}{J_z} &= \frac{2\Delta z}{A_z} + 
+        \left(\frac{\Delta v_z}{v_{z,\text{max}}}\right)^2 && \text{if $v_z = 0$} \\
+    \frac{\Delta J_z}{J_z} &= \frac{2\Delta v_z}{v_{z,\text{max}}} + 
+        \left(\frac{\Delta z}{A_z}\right)^2  && \text{if $z=0$} \text{.}
+\end{aligned}
+\eeq
+
+Since most of the time stars will be at maximum amplitude (i.e. turnaround) in both $R$ and $z$, we can approximate the order of magnitude of the systematic error in the actions by
+\beq\label{eq:Ji_err_mosttime}
+\begin{split}
+\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta R}{R_g} \\
+\frac{\Delta J_{R}}{J_{R}} &= \frac{2\Delta R}{A_R} \\
+\frac{\Delta J_{z}}{J_{z}} &= \frac{2\Delta z}{A_z} \text{,}
+\end{split}
+\eeq
+where we have also ignored second order terms.
+
+In the remainder of this section, we compare our analytic estimates of the effect on actions of a midplane offset against numerical calculations.
+A numerical evaluation of the effect of velocity offsets on actions is deferred to future work, as we discuss in Section~\ref{ssec:lsr_var}.
+
+\subsection{Numerical Methods} \label{ssec:action_comp}
+We now quantify the argument made in Section~\ref{ssec:cartoon} using numerical computations of the actions for a range of orbits in a model Galactic potential.
+We compute actions as in our previous work \citep{2018ApJ...867...31B}, using
+the code \texttt{gala} v0.3 to perform orbit integrations and conversion to
+action space \citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute actions
+we use the torus-mapping technique first presented by
+\citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to
+calculate actions for an orbital time-series starting from a phase-space
+position $(x, v)$ and integrated in a potential $\Phi$. We use the default
+\texttt{MWPotential} as our potential, which is based on the Milky Way
+potential available in \texttt{galpy} \citep{2015ApJS..216...29B}. This
+potential includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
+a Miyamoto-Nagai disk \citep{1975PASJ...27..533M}, and an NFW halo
+\citep{1997ApJ...490..493N}, and is fit to empirically match some
+observations. We use the Dormand-Prince 8(5,3) integration scheme
+\citep{Dormand80:integrator} with a timestep of $1\,\Myr$ and integrate for
+$5\,\Gyr$, corresponding to $\sim 20$ orbits for a Sun-like star.
+
+We assume the Sun is located at $(8.2, 0, 0)\,\kpc$. None of our orbit
+calculations depend on the value of the LSR in this toy potential (though this
+is important when using real data, since the conversion from heliocentric to
+Galactocentric coordinates depends on the LSR). In this potential, we have
+that the circular velocity $v_{\text{circ}}$ is $231\,\kms$ on the Solar
+circle.
+
+Other methods for computing actions are used in the literature. For example,
+the St\"ackel Fudge method \citep{2016MNRAS.457.2107S}, which uses a single
+St\"ackel potential (with analytic actions) to approximate the Galactic
+potential \citep{1985MNRAS.216..273D,2012MNRAS.426.1324B}, was used in many
+recent works exploring actions in the Galactic disk
+\citep[e.g.][]{2019MNRAS.484.3291T,2018MNRAS.481.4093S,2018arXiv180803278T}.
+For disk-like orbits, existing implementations of the St\"ackel Fudge method
+are of acceptable accuracy, but since we also consider halo-like orbits in
+this work (where the St\"ackel Fudge method is inaccurate) we choose to use
+orbit integration and torus mapping throughout \citep{2016MNRAS.457.2107S}.
+
+\subsection{Quantification of the Midplane Effect} \label{ssec:quant}
+We quantify how a systematic error in the Galactocentric coordinate system
+induces phase-dependence in the {\em observed} actions. We will consider three
+orbits in the model potential described in Section~\ref{ssec:action_comp} that
+are typical of stars in the stellar thin disk, stellar thick disk, and the
+stellar halo. We summarize their initial positions in phase space and the
+actions computed by integrating their orbits in the correct potential in
+Table~\ref{tab:orbits}. Each orbit, integrated without systematic coordinate
+errors, is plotted in Appendix~\ref{app:orbits}. We will refer to these orbits
+by their names (thin-disk, thick-disk, halo) henceforth.
+
+We begin by considering the thick-disk orbit. Suppose an immortal observer who can observe
+the thick-disk star's phase-space position at many different times (and hence
+different orbital phases), but used a coordinate system in which the midplane
+is systematically offset in height by $100\,\pc$ from its actual location ---
+i.e. we subtract the vector $(0, 0, 100)\,\pc$ from each position in the
+orbit. This corresponds to the immortal observer physically located at e.g. the position
+$(8, 0, 0)\,\kpc$ in the coordinate system of the true potential, but
+erroneously thinking they are located at $(8, 0, 0.1)\,\kpc$.
 
@@
 \tablecaption{Description and names of the three orbits considered in this
-work. For the last two columns: $z_{\text{max}}$ is the maximum height of the
-orbit while $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of
-the radial excursions of the orbit. In the epicyclic approximation, $A_z =
-z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$.
-\label{tab:orbits}}
+work. For the last two columns: $z_{\text{max}}$ the maximum height of the
+orbit while $A_R$ is the magnitude of the radial excursions of the orbit,
+computed as one half the maximum radius minus the minimum
+radius.\label{tab:orbits}}
 \tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
@@
 \colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
-\colhead{$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$}\\ \colhead{ } &
+\colhead{$A_R$}\\ \colhead{ } &
 \colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
@@
 
-Before turning to numerical methods, we first write down analytically the
-systematic error in the actions induced from either a midplane offset or a
-Galactocentric radius offset. We use the epicyclic approximation, which
-assumes that the motion in the $z$ and $R$ components of the orbit are
-decoupled and follow simple harmonic motion about a circular and planar
-``guiding orbit'' \citep[][Section~3.2 and references
-therein]{2008gady.book.....B}. We refer to the radius of this orbit (the
-``guiding radius'') as $R_g$. This approximation is an excellent description
-of thin disk-like orbits and a good description of thick disk-like orbits in
-an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and
-spiral arms). We also make the assumption of a perfectly flat circular
-velocity curve with velocity $v_c$, a good approximation near the Solar circle
-\citep[e.g.][]{2017MNRAS.465...76M}.
-
-Under this approximation, we can write down the cylindrical components
-of the orbits as
-\beq\label{eq:orbits_epi}
-\begin{split}
-\phi(t) &= \Omega_c t \\
-R(t) &= R_g + A_R \sin{(\kappa t + \alpha)} \\
-z(t) &= A_z \sin{(\nu t + \beta)}
-\text{,}
-\end{split}
-\eeq
-where $\kappa$ and $\nu$ are the radial and vertical frequencies, $\Omega_c
-\equiv v_c/R_g$ is the rotation frequency, $A_z$ and $A_R$ are the amplitudes of
-the stellar motion in the vertical and radial coordinates, and $\alpha$ and
-$\beta$ are arbitrary constants. Similarly, the velocities of the orbit are
-given by:
-\beq\label{eq:orbits_vel_epi}
-\begin{split}
-v_{\phi}(t) &= v_c \\
-v_R(t) &= \kappa A_R \cos{(\kappa t + \alpha)} \\
-v_z(t) &= \nu A_z \cos{(\nu t + \beta)}
-\text{.}
-\end{split}
-\eeq
-
-In this case, we have that for $J_{\phi}$
-\citep[][Section~3.5.3b]{2008gady.book.....B}:
-\beq\label{eq:Jphi_epi}
-J_{\phi} = R_g v_c\text{,}
-\eeq
-and for $J_R$ and $J_z$:
-\beq\label{eq:JR_Jz_epi}
-\begin{split}
-J_R &= \frac{E_R}{\kappa} \\
-J_z &= \frac{E_z}{\nu} \text{,}
-\end{split}
-\eeq
-where $E_R$ and $E_z$ are the energy per unit mass in the radial and vertical
-coordinates, respectively. Therefore,
-\beq\label{eq:JR_Jz_epi_energy}
-\begin{split}
-J_R &= \frac{v_R^2 + \kappa^2 (R-R_g)^2}{2\kappa} \\
-J_z &= \frac{v_z^2 + \nu^2 z^2}{2\nu}\text{.}
-\end{split}
-\eeq
-Using Equations~\eqref{eq:orbits_epi} and \eqref{eq:orbits_vel_epi}, we can
-simplify this and write
-\beq\label{eq:JR_Jz_epi_final}
-\begin{split}
-J_R &= \frac{\kappa A_R^2}{2} = \frac{v_{R,\text{max}}^2}{2\kappa} \\
-J_z &= \frac{\nu A_z^2}{2} = \frac{v_{z,\text{max}}^2}{2\nu}\text{,}
-\end{split}
-\eeq
-where the last equality in each line comes from the fact that
-$v_{R,\text{max}} = \kappa A_R$ and $v_{z,\text{max}} = \nu A_z$.
-
-Notice that while the value for each of $J_{\phi}$, $J_R$, and $J_z$ is phase
-independent, the contribution from the kinetic and potential terms in
-Equation~\eqref{eq:JR_Jz_epi_energy} is phase dependent. Now assume that the
-coordinates $(R, z, v_{\phi}, v_R, v_z)$ are offset by $(\Delta R, \Delta z,
-\Delta v_{\phi}, \Delta v_R, \Delta v_z)$. We can then apply the standard
-propagation of errors formula to Equation~\eqref{eq:JR_Jz_epi_energy} to
-determine the error in each of the actions. For $J_{\phi}$, the induced error
-is:
-\beq\label{eq:induced_Jphi}
-\frac{\Delta J_{\phi}}{J_{\phi}} = \frac{\Delta R}{R_g} 
-                                    + \frac{\Delta v_{\phi}}{v_c}\text{.}
-\eeq
-For $J_R$, the induced error is:
-\beq\label{eq:induced_JR}
-\frac{\Delta J_R}{J_R} = \frac{2(R-R_g)}{A_R^2}\Delta R
-                         + \frac{2v_R}{v_{R,\text{max}}^2} \Delta v_R \text{.}
-\eeq
-For $J_z$, the induced error is:
-\beq\label{eq:induced_Jz}
-\frac{\Delta J_z}{J_z} = \frac{2z}{A_z^2}\Delta z
-                         + \frac{2v_z}{v_{z,\text{max}}^2} \Delta v_z \text{.}
-\eeq
-We have ignored second order contributions.
-
-Since most of the time stars will be at maximum amplitude (i.e. turnaround) in
-both $R$ and $z$, we can approximate the order of magnitude of the systematic
-error in the actions by
-\beq\label{eq:Ji_err_mosttime}
-\begin{split}
-\frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta R}{R_g} 
-                                    + \frac{\Delta v_{\phi}}{v_c} \\
-\frac{\Delta J_{R}}{J_{R}} &= \frac{2\Delta R}{A_R} \\
-\frac{\Delta J_{z}}{J_{z}} &= \frac{2\Delta z}{A_z} \text{,}
-\end{split}
-\eeq
-where we have also ignored second order terms.
-
-In the remainder of this section, we compare our analytic estimates of the
-effect of a midplane offset on actions against numerical calculations. A
-numerical evaluation of the effect of velocity offsets on actions is deferred
-to future work, as we discuss in Section~\ref{ssec:lsr_var}.
-
-\subsection{Numerical Methods} \label{ssec:action_comp}
-We now quantify the argument made in Section~\ref{ssec:cartoon} using
-numerical computations of the actions for a range of orbits in a model
-Galactic potential. We compute actions as in \citet{2018ApJ...867...31B},
-using the code \texttt{gala} v0.3 to perform orbit integrations and conversion
-to action space \citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute
-actions we use the torus-mapping technique first presented by
-\citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to
-calculate actions for an orbital time-series starting from a phase-space
-position $(x, v)$ and integrated in a potential $\Phi$. We use the default
-\texttt{MWPotential} as our potential, which is based on the Milky Way
-potential available in \texttt{galpy} \citep{2015ApJS..216...29B}. This
-potential includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
-a Miyamoto--Nagai disk \citep{1975PASJ...27..533M}, and an NFW halo
-\citep{1997ApJ...490..493N}, and is fit to empirically match some
-observations. We use the Dormand-Prince 8(5,3) integration scheme
-\citep{Dormand80:integrator} with a timestep of $1\,\Myr$ and integrate for
-$5\,\Gyr$, corresponding to $\sim 20$ orbits for a Sun-like star.
-
-We assume the Sun is located at $(8.2, 0, 0)\,\kpc$. None of our orbit
-calculations depend on the value of the LSR in this toy potential (though this
-is important when using real data, since the conversion from heliocentric to
-Galactocentric coordinates depends on the LSR). In this potential, we have
-that the circular velocity $v_{\text{circ}}$ is $231\,\kms$ at the Solar
-circle.
-
-Other methods for computing actions are used in the literature. For example,
-the St\"ackel Fudge method \citep{2016MNRAS.457.2107S}, which uses a single
-St\"ackel potential (with analytic actions) to approximate the Galactic
-potential \citep{1985MNRAS.216..273D, 2012MNRAS.426.1324B}, was used in many
-recent works exploring actions in the Galactic disk
-\citep[e.g.][]{2019MNRAS.484.3291T, 2018MNRAS.481.4093S, 2018arXiv180803278T}.
-For disk-like orbits, existing implementations of the St\"ackel Fudge method
-are of acceptable accuracy, but since we also consider halo-like orbits in
-this work (where the St\"ackel Fudge method is inaccurate) we choose to use
-orbit integration and torus mapping throughout \citep{2016MNRAS.457.2107S}.
-
-\subsection{Quantification of the Midplane Effect} \label{ssec:quant}
-
-\begin{figure*}[ht!]
+Starting from each observation of the star's phase-space position that such an
+immortal observer makes every $\Myr$, we then compute the
+actions as outlined above using an integration from that starting point in
+phase space carried out using the true potential but specifying the star's
+present-day position using the systematically offset coordinate system.
+Essentially we are shifting the integrated orbit and then reintegrating at
+each point along the shifted orbit, where the shifted orbit is the observed
+orbit. The erroneous computed actions for each phase-space starting point are
+shown for the first $\Gyr$ of the orbit in the {\em upper} panels
+Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
+scheme fails and very large actions are reported by \texttt{gala}---we perform
+a $5\sigma$ clip on each action to exclude such orbits, but this only excludes
+a total of $5$ orbits out of the $1000$ considered for
+Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
+vast majority of orbits are computed properly.}
+
+We also perform the same procedure but assume a $100\,\pc$ offset in the $x$
+component (i.e. subtracting the vector $(100, 0, 0)\,\pc$) in the {\em lower}
+panels. This is equivalent to a systematic offset in the determination of the
+distance from the Sun to the Galactic center.
+
+Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
+offset coordinate systems depend on the time (i.e. orbital phase) at which the
+star's phase-space position is observed. To the immortal observer
+this would appear as time-dependence in the actions, which appear to oscillate
+around their value using the correct coordinate system even though in this
+example the observer is using the correctly constructed best-fit static,
+axisymmetric potential. The relative size of the phase variation in each
+action depends on the direction of the systematic offset as well as the true
+values of the actions (i.e. the type of orbit). In reality we will have one
+measurement of the phase-space position to work with, in which case the
+determination of the orbital phase in $z$ or $R$ is degenerate with the degree
+of systematic offset in that coordinate (see Figure~\ref{fig:cartoon}).
+Additionally, in searching for migrated members of known open clusters, we
+would compute actions for stars with a wide range of orbital phases and
+locations in the Galaxy. In the following we therefore quote percentile ranges
+for the possible values computed for each action as a proxy for the effect of
+these systematic errors in the coordinate system.
+
+For a systematic offset in $z$ ({\em upper} panels), the
+95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $2.2$ and
+$6.2\,\actunit$ for $J_R$ and $J_z$, respectively. These are fractional errors
+of $5.7\,\%$ and $85.7\,\%$, respectively. The error induced in $J_{\phi}$ is
+negligible, as expected since $J_{\phi}$ is independent of $z$. It is worth
+pointing out that a $100\,\pc$ error in an orbit with
+$z_{\text{max}}=850\,\pc$ --- a $12\%$ error
+--- induced an $86\%$ error in the computation of $J_z$.
+
+For a systematic offset in $x$ (or Galactic center distance), the
+95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
+$47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively.
+These are fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively ---
+despite only a $1.2\%$ error in the distance to the Galactic center.
+
+\begin{figure*}
 \begin{center}
@@
 error in the Galactocentric coordinate system. We consider here the thick-disk
-orbit, which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520,
-7.0)\,\actunit$ and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We
-integrate the orbit according to the procedure laid out in
-Section~\ref{ssec:action_comp}, and which we plot in
-Appendix~\ref{app:orbits}. Then, we subtract $100\,\pc$ from the $z$ value
-(upper panels) or the $x$ value (lower panels) of each position in the orbit,
-corresponding to an erroneous observer assuming a midplane (upper) or solar
-radius (lower) that is off by $100\,\pc$. We then allow an immortal observer
-to observe the orbit over $1\,\Gyr$ and perform the same orbit integration
+orbit, which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520, 7.0)\,\actunit$
+and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We integrate the
+orbit according to the procedure laid out in Section~\ref{ssec:action_comp}.
+Then, we subtract $100\,\pc$ from the $z$ value ({\em upper panels}) or the
+$x$ value ({\em lower panels}) of each position in the orbit, corresponding to
+an erroneous observer assuming a midplane ({\em upper}) or solar radius ({\em
+lower}) that is off by $100\,\pc$. We then allow our (immortal) observer to
+observe the orbit over $1\,\Gyr$ and perform the same orbit integration
 procedure at each timestep, and report the values of the actions. The
@@
 with the middle $90\%$ of values being within $\sim43\%$ of the true $J_z$.
-The $x$ offset induces uncertainties in $J_R$, $J_{\phi}$, and $J_z$ of
-$\sim21\%$, $\sim3\%$, and $\sim3\%$.}
+The $x$ offset induces uncertainties in $J_R$ and $J_{\phi}$ of $\sim21\%$ and
+$\sim3\%$, respectively. A $\sim3\%$ error in $J_z$ is also induced.}
 \label{fig:one_orbit_wrong_ref}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{fig/schmactions_Jz_zerr_hist.pdf}
+\includegraphics[width=\columnwidth]{fig/schmactions_Jz_hist.pdf}
 \end{center}
-\caption{A histogram of the observed values in $J_z$ for the thick-disk (top
-panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
-$100\,\pc$. One can see that if the observed $z$ values have a bias (from e.g.
-an incorrectly computed midplane), then the induced error distribution in
-$J_z$ is decidedly non-Gaussian. Therefore, any sort of error propagation must
-take this into account. Intuition for the shape of each panel is given in the
-text. We also plot one half the 95\uth{} percentile minus the 5\uth{}
-percentile of each distribution as a horizontal arrow anchored on the true
-$J_z$ value. We call this $\Delta J_z$ and will use it (along with the
-similarly defined $\Delta J_R$ and $\Delta J_{\phi}$) to empirically describe
-the error distribution. We see that $\Delta J_z$ roughly corresponds to the
-distance from the true $J_z$ value to one of the modes of the distribution of
-observed $J_z$ values. Similar plots for $J_R$ induced by a $z$ offset and
-$J_R$ and $J_{\phi}$ induced by an $x$ offset are given in
-Appendix~\ref{app:hist}.}
+\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
+assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
+values have a bias (from e.g. an incorrectly computed midplane), then the
+induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
+sort of error propagation must take this into account. Intuition for the shape
+of each panel is given in the text. We also plot one half the 95\uth{}
+percentile minus the 5\uth{} percentile of each distribution as a horizontal
+arrow anchored on the true $J_z$ value. We call this $\Delta J_z$ and will use
+it (along with the similarly defined $\Delta J_R$ and $\Delta J_{\phi}$) to
+empirically describe the error distribution. We see that $\Delta J_z$ roughly
+corresponds to the distance from the true $J_z$ value to one of the modes of
+the error distribution.}
 \label{fig:Jz_hist}
@@
 
-We quantify how a systematic error in the Galactocentric coordinate system
-induces phase-dependence in the \emph{observed} actions. We will consider
-three orbits in the model potential described in
-Section~\ref{ssec:action_comp} that are typical of stars in the stellar thin
-disk, stellar thick disk, and the stellar halo. We summarize their initial
-positions in phase space and the actions computed by integrating their orbits
-in the correct potential in Table~\ref{tab:orbits}. Each orbit, integrated
-without systematic coordinate errors, is plotted in Appendix~\ref{app:orbits}.
-We will refer to these orbits by their names (thin-disk, thick-disk, halo)
-henceforth.
-
-We begin with the thick-disk orbit. Consider an immortal observer who can
-measure the thick-disk star's phase-space position at many different times
-(and hence different orbital phases). However, the immortal observer uses a
-coordinate system in which the midplane is systematically offset in height by
-$100\,\pc$ from its actual location --- i.e. we subtract the vector $(0, 0,
-100)\,\pc$ from each position in the orbit. This corresponds to the immortal
-observer physically located at e.g. the position $(8, 0, 0)\,\kpc$ in the
-coordinate system of the true potential, but erroneously thinking they are
-located at $(8, 0, 0.1)\,\kpc$.
-
-We consider that the immortal observer makes a measurement every $\Myr$. At
-each measurement, we compute the actions as outlined above using the true
-potential but by specifiying the star's starting position using the
-systematically offset coordinate system. Essentially we are shifting the
-integrated orbit and then reintegrating at each point along the shifted orbit,
-where the shifted orbit is the observed orbit. The erroneous computed actions
-for each phase-space starting point are shown for the first $\Gyr$ of the
-orbit in the upper panels
-Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
-scheme fails and very large actions are reported by \texttt{gala}---we perform
-a $5\sigma$ clip on each action to exclude such orbits, but this only excludes
-a total of $5$ orbits out of the $1000$ considered for
-Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
-vast majority of orbits are computed properly.}
-
-We also perform the same procedure but assume a $100\,\pc$ offset in the $x$
-component (i.e. subtracting the vector $(100, 0, 0)\,\pc$) in the lower
-panels. This is equivalent to a systematic offset in the determination of the
-distance from the Sun to the Galactic center.
-
-Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
-offset coordinate systems depend on the time (i.e. orbital phase) at which the
-star's phase-space position is observed. To the immortal observer this would
-appear as time-dependence in the actions, which appear to oscillate around
-their value using the correct coordinate system even though in this example
-the observer is using the correctly constructed best-fit static, axisymmetric
-potential. The relative size of the phase variation in each action depends on
-the direction of the systematic offset as well as the true values of the
-actions (i.e. the type of orbit). In reality we will have one measurement of
-the phase-space position to work with, in which case the determination of the
-orbital phase in $R$ or $z$ is degenerate with the degree of systematic offset
-in that coordinate (see Figure~\ref{fig:cartoon}).  In the following we
-therefore quote percentile ranges for the possible values computed for each
-action as a proxy for the effect of these systematic errors in the coordinate
-system.
-
-For a systematic offset in $z$ (upper panels), the 95\textsuperscript{th}
-minus 5\textsuperscript{th} percentiles are $2.2$ and $6.2\,\actunit$ for
-$J_R$ and $J_z$, respectively. These are fractional errors of $5.7\,\%$ and
-$86\,\%$, respectively. The error induced in $J_{\phi}$ is negligible, as
-expected since $J_{\phi}$ only depends on the $x$- and $y$-components of the
-position and velocity of the stars.\footnote{In practice, however, $J_{\phi}$
-is computed as part of the torus-fitting method.} It is worth pointing out
-that a $100\,\pc$ error in an orbit with $z_{\text{max}}=850\,\pc$ --- a
-$12\%$ error --- induced an $86\%$ error in the computation of $J_z$.
-
-For a systematic offset in $x$ (or distance to the Galactic center), the
-95\textsuperscript{th} minus 5\textsuperscript{th} percentiles are $6.9$,
-$47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively.
-These are fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively ---
-despite only a $1.2\%$ error in the distance to the Galactic center.
-
 In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
-$J_z$ for the thick-disk orbit (top panel) and thin-disk orbit (bottom panel)
-assuming a $z$ offset of $100\,\pc$ (i.e. as in the upper right panel of
-Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted as a
-vertical dashed line. Here we see that the errors in $J_z$ induced by a
+$J_z$ for the thick-disk orbit ({\em upper panel}) and thin-disk orbit ({\em
+lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. the {\em upper right} panel
+of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted
+as a vertical dashed line. Here we see that the errors in $J_z$ induced by a
 systematic offset in $z$ are non-Gaussian and bimodal. Furthermore, neither of
@@
 $J_z$ computed with no offset in $z$). In the case of the thin-disk orbit
-(bottom panel), we see that, in addition to the prior complications, the error
-distribution is not even centered on null. This comes about when the midplane
-error is $\gtrsim z_{\text{max}}$, where $z_{\text{max}}$ is the maximum
-height of the orbit, equivalent to $A_z$ in the epicyclic approximation (see
-Section~\ref{ssec:epi_action}).
-
-In Appendix~\ref{app:hist} we plot the same histogram as in
-Figure~\ref{fig:Jz_hist}, but for the observed distributions of $J_R$ induced
-by a $z$ offset (upper left panel of Figure~\ref{fig:many_orbit_wrong_ref})
-and the observed distributions of $J_R$ and $J_{\phi}$ errors induced by an
-$x$ offset (lower left and lower center panels of
-Figure~\ref{fig:many_orbit_wrong_ref}, respectively). We find similar error
-distributions as in Figure~\ref{fig:Jz_hist}, with the exception that the
-observed $J_R$ distribution induced by an $x$ offset more closely resembles a
-Gaussian.
-
-\begin{figure*}[htb!]
-\begin{center}
-\includegraphics[width=\textwidth]{fig/schmactions_many_orbits.pdf}
-\end{center}
-\caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error
-in the measured action ($\Delta J_i$) from coordinate system errors for the
-thin, thick, and halo orbits (Table~\ref{tab:orbits}). See discussion in the
-text for the justification in using this to measure the magnitude of the
-induced error. The left, center, and right panels show the result for $J_R$,
-$J_{\phi}$, and $J_z$, respectively. The upper panels consider an offset in
-$z$ and the lower panels consider an offset in $x$ (equivalently, an offset in
-the Solar radius). In some panels we also plot as dashed lines the epicyclic
-prediction of the induced action error (Equation~\eqref{eq:Ji_err_mosttime}).
-In the epicyclic approximation, a $z$ offset only induces an error in $J_z$
---- for all three orbits the epicyclic approximation is a good description of
-the $J_z$ error. An $x$ offset induces an error in $J_R$ and $J_{\phi}$. The
-error in $J_R$ is somewhat well-described for the thin-disk and thick-disk
-orbits, and a poor description for the halo orbit. For $J_{\phi}$, the
-epicyclic approximation is not a good description for any orbit.}
-\label{fig:many_orbit_wrong_ref}
-\end{figure*}
+({\em lower panel}), we see that, in addition to the prior complications, the
+error distribution is not even centered on null. This comes about when the
+midplane error is $\gtrsim z_{\text{max}}$.
 
 We now describe the intuition behind the shape of Figure~\ref{fig:Jz_hist}.
-Consider first the thick-disk orbit (top panel), where the offset in $z$ is
-much less than $z_{\text{max}}$ of the orbit. The peaks in the distribution
-correspond to the turning points of the orbit (or points of maximum vertical
-amplitude), where $v_z \sim 0$ and where the star is on most of it orbit. This
-is why the distribution is peaked on these values. For the thin-disk orbit
-(bottom panel), the offset in $z$ is comparable to $z_{\text{max}}$. Now,
-there will be some points in the orbit where $v_z = 0$ and $z=0$ (in the
-observed, erroneous coordinate system). At these points, the computed $J_z$
-will vanish. The asymmetry and systematic offset then comes about because of
-the constraint that $J_z \geq 0$.\footnote{This is similar to arguments in
-cosmology for why gravity produces non-Gaussianity in the density field, since
-the density cannot become negative but it can grow arbitrarily large.}
+Consider first the thick-disk orbit ({\em top} panel), where the
+offset in $z$ is much less than $z_{\text{max}}$ of the orbit. The peaks in
+the distribution correspond to the turning points of the orbit (or points of
+maximum vertical amplitude), where $v_z \sim 0$ and where the star spends
+most of it orbit. This is why the distribution is peaked on these values. For
+the thin-disk orbit ({\em bottom} panel), the offset in $z$ is comparable to
+$z_{\text{max}}$. Now, there will be some points in the orbit where $v_z = 0$
+and $z=0$ (in the observed, erroneous coordinate system). At these points, the
+computed $J_z$ will vanish. The asymmetry and systematic offset then comes
+about because of the constraint that $J_z \geq 0$.\footnote{This is similar to
+arguments in cosmology for why gravity produces non-Gaussianity in the density
+field, since the density cannot become negative but it can grow arbitrarily
+large.}
 
@@
 
-The upper row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread
-induced in each action for an offset in the $z$-component (i.e. the Galactic
-midplane). In the lower row we consider offsets in the $x$ component (i.e. the
-Solar radius). The left, center, and right columns show the fractional spread
-in the values computed for $J_z$, $J_{\phi}$, and $J_R$, respectively. We
-compute the fractional spread $\Delta J_i/J_i$, where $\Delta J_i$ is defined
-in the previous paragraph and $J_i$ is the action value computed in the true
-coordinate system, over the course of the first $\Gyr$.
-
-In the upper middle panel of Figure~\ref{fig:many_orbit_wrong_ref}, there is
-essentially no spread in the determination of $J_{\phi}$. This is expected
-since $J_{\phi}$ is independent of $z$ and is thus unaffected by offsets in
-$z$, as discussed earlier. Indeed, the result we found in
+The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
+in each action for an offset in the $z$-component (i.e. the Galactic
+midplane). In the {\em bottom row} we consider offsets in the $x$ component
+(i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
+show the fractional spread in the values computed for $J_z$, $J_{\phi}$, and $J_R$,
+respectively. We compute the fractional spread $\Delta J_i/J_i$ as one half
+the 95\uth{} minus 5\uth{} percentile of the action computed using the offset
+coordinate system over the course of the first $\Gyr$, divided by the true
+action (computed in the correct coordinate system).
+
+In the {\em upper middle} panel of Figure~\ref{fig:many_orbit_wrong_ref},
+there is essentially no spread in the determination of $J_{\phi}$. This is expected
+since, in principle, $J_{\phi}$ only depends on the $x$- and $y$-components of the
+position and velocity of the stars\footnote{In practice, however, $J_{\phi}$ is
+computed as part of the torus-fitting method.}, and is thus unaffected by
+offsets in $z$. Indeed, the result we found in
 Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
-orbit types. This is also a demonstration of the robustness of the orbit
-integration and action calculation methods we use.
-
-The upper right panel of Figure~\ref{fig:many_orbit_wrong_ref} shows that the
-fractional error in $J_z$ is more exaggerated for more planar (disk-like)
-orbits. For the thin-disk orbit, a systematic offset of $\sim15\,\pc$ in the
-$z$ coordinate already results in $25\%$ deviations in the actions. The
-required offset for $25\%$ deviation is $\sim120\,\pc$ for the thick-disk
-orbit. The halo orbit is relatively resistant to errors in the midplane, with
-only $\sim15\%$ error in $J_z$ out to a midplane offset of $500\,\pc$.
-
-For the offset in the Solar radius (lower panels), the error is largest for
-$J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
-deviations in $J_z$. In the lower middle and lower right panels all three
-lines nearly overlap.
+orbit types.
+
+The {\em upper right} panel of Figure~\ref{fig:many_orbit_wrong_ref} shows
+that the fractional error in $J_z$ is more exaggerated for more planar
+(disk-like) orbits. For the thin-disk orbit, a systematic offset of
+$\sim15\,\pc$ in the $z$ coordinate already results in $25\%$ deviations in
+the actions. The required offset for $25\%$ deviation is $\sim120\,\pc$ for
+the thick-disk orbit. The halo orbit is relatively resistant to errors in the
+midplane, with only $\sim15\%$ error in $J_z$ out to a midplane offset of
+$500\,\pc$.
+
+For the offset in the Solar radius ({\em bottom panels}), the error is largest
+for $J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
+deviations in $J_z$. In the {\em bottom middle} and {\em bottom right} panels
+all three lines nearly overlap.
 
@@
 include the estimation of the action errors derived under the epicyclic
-approximation from Equation~\eqref{eq:Ji_err_mosttime}, with $\Delta
-v_{\phi}=0$, as dashed lines in the color of each orbit. This equation is
-relevant since during most of the orbit the star will be at maximum radial and
-vertical amplitude and we do not consider velocity offsets. Note that we
-consider an error in the $x$-coordinate $\Delta x$, which is not exactly the
-same as $\Delta R$. For observations of stars close to us, we have that
-$\Delta x \sim \Delta R$, but for the experiment performed in this Section we
-consider observations of the star throughout its entire orbit. This introduces
-a factor of $2/\pi$ when converting from $\Delta x$ to $\Delta R$, which we
-derive in Appendix~\ref{app:deltax}.
-
-The epicyclic approximation is a good predictor of $\Delta J_z$, even for halo
-orbits. It performs similarly for $\Delta J_R$, now underpredicting for halo
-orbits and slightly overpredicting for thin-disk orbits. Note that for the
-particular orbits we chose, the thin-disk orbit has slightly larger $A_R$ than
-the thick-disk orbit, and so we expect the epicyclic approximation to perform
-slightly worse. The epicyclic approximation underpredicts $\Delta J_{\phi}$
-for all orbits.
+approximation from Equation~\ref{eq:Ji_err_mosttime}.
+This Equation is relevant since during most of the orbit the star will be at maximum radial and vertical amplitude and we do not consider velocity offsets.
+Note that we consider an error
+in the $x$-coordinate $\Delta x$, which is not quite the same as $\Delta R$.
+For observations of stars close to us, we have that $\Delta x \sim \Delta R$,
+but for the experiment performed in this Section we consider observations of
+the star throughout its entire orbit. This introduces a factor of $2/\pi$ when
+converting from $\Delta R$ to $\Delta x$, which we derive in
+Appendix~\ref{app:deltax}.
+
+The epicyclic approximation is a good predictor of $J_z$, even for halo
+orbits. It performs similarly for $J_R$, now underpredicting for halo orbits
+and slightly overpredicting for thin-disk orbits. Note that for the particular
+orbits we chose, the thin-disk orbit has slightly larger $A_R$ than the
+thick-disk orbit, and so we expect the epicyclic approximation to perform
+slightly worse. The epicyclic approximation underpredicts $J_{\phi}$ for all
+orbits.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{fig/schmactions_many_orbits.pdf}
+\end{center}
+\caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error
+in the measured action ($\Delta J_i$) from coordinate system errors for the
+thin, thick, and halo orbits (Table~\ref{tab:orbits}). See discussion in the
+text for the justification in using this to measure the magnitude of the
+induced error. The {\em left}, {\em center}, and {\em right} panels show the
+result for $J_R$, $J_{\phi}$, and $J_z$, respectively. The {\em upper} panels
+consider an offset in $z$ and the {\em lower} panels consider an offset in $x$
+(equivalently, an offset in the Solar radius). In some panels we also plot as
+dashed lines the epicyclic prediction of the induced action error
+(Equations~\ref{eq:actions_epi_error}~and~\ref{eq:Jr_epi_dx}). In the
+epicyclic approximation, a $z$ offset only induces an error in $J_z$ --- for
+all three orbits the epicyclic approximation is a good description of the
+$J_z$ error. An $x$ offset induces an error in $J_R$ and $J_{\phi}$. The error in
+$J_R$ is somewhat well-described for the thin-disk and thick-disk orbits, and
+a poor description for the halo orbit. For $J_{\phi}$, the epicyclic approximation
+is not a good description for any orbit.}
+\label{fig:many_orbit_wrong_ref}
+\end{figure*}
 
@@
 \caption{The fractional error in $J_z$ as a function of $J_z$ for a few
-different offsets in $z$. All orbits have the same initial position of $(8, 0,
-0)\,\kpc$ and velocity $(-190, 0, v_z)\,\kms$, where we vary $v_z$. We show
-this for a $z$ offset of $10, 50, 100\,\pc$ (orange, teal, and green
-respectively). As before, the error ($\Delta J_z$) is one half the 95\uth{}
-minus 5\uth{} percentile of the distribution of observed $J_z$ values over the
-course of the orbit. There are large errors for thin-disk like orbits ($J_z
-\sim 0.5\,\actunit$), even for a small midplane offset of $10\,\pc$. As dashed
-lines in each color we also plot the prediction for $\Delta J_z/J_z$ from the
-epicyclic approximation (Equation~\eqref{eq:Ji_err_mosttime}), which shows
-excellent agreement with the numerically computed values.}
+different offsets in $z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
+(red, cyan, green respectively). As before, the error ($\Delta J_z$) is one
+half the 95\uth{} minus 5\uth{} percentile of the distribution of observed
+$J_z$ values over the course of the orbit. There are large errors for
+thin-disk like orbits ($J_z \sim 0.5\,\actunit$), even for a small midplane
+offset of $10\,\pc$. }
 \label{fig:dJz_fun_Jz}
@@
 fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of $10, 50,
-100\,\pc$ (orange, teal, and green, respectively) in
-Figure~\ref{fig:dJz_fun_Jz}. For each orbit, we set the initial position to be
-$(8,0,0)\,\kpc$ and the initial velocity to be $(-190, 0, v_z)\,\kms$, where
-we vary $v_z$. For a thin disk-like orbit ($J_z\sim0.5\,\actunit$), even a
-$10\,\pc$ offset in $z$ is enough to induce a $\sim25\%$ error in $J_z$. For
-larger values of $J_z$, the fractional errors are suppressed, but the induced
-error can still be large depending on how great the $z$ offset is. We also
-plot the epicylic prediction for $\Delta J_z / J_z$ from
-Equation~\eqref{eq:Ji_err_mosttime} as dashed lines for each $z$ offset. We
-find that the epicyclic approximation matches the numerical estimate quite
-well.
+100\,\pc$ (orange, cyan, and green, respectively) in
+Figure~\ref{fig:dJz_fun_Jz}. For a thin disk-like orbit
+($J_z\sim0.5\,\actunit$), even a $10\,\pc$ offset in $z$ is enough to induce a
+$\sim25\%$ error in $J_z$. For larger values of $J_z$, the fractional errors
+are suppressed, but the induced error can still be large depending on how
+great the $z$ offset is. We also plot the epicylic prediction for $\Delta J_z
+/ J_z$ from Equation~\ref{eq:Ji_err_mosttime} as dashed lines for each $z$
+offset. We find that the epicyclic approximation matches the numerical
+estimate quite well.
 
@@
 would generate the action errors described in Section~\ref{sec:ref_frame}:
-uncertainties in measurements of the position and velocity of the Sun relative
-to the Galactic center. Converting from heliocentric to Galactocentric
-coordinates relies on these measurements. Therefore, errors in their values
-will induce a systematic offset in the Galactocentric phase-space position of
-any observed star.
+uncertainties in measurements of the position and velocity of the Sun relative to the Galactic center. Converting from
+heliocentric to Galactocentric coordinates relies on these measurements.
+Therefore, errors in their values will induce a systematic offset in the
+Galactocentric phase-space position of any observed star.
 
@@
 First, one must define the center of the Galaxy. This is usually taken to be
-the location of the central supermassive black hole, Sagittarius~A\textsuperscript{*}
-\citep[\sgra{}, e.g.][]{2004ApJ...616..872R}. From stellar motions near
-\sgra{}, the distance from the Sun to \sgra{}, referred to as $R_0$, can be
-precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a
-recent measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
+the location of the central supermassive black hole, Sgr~A\textsuperscript{*}
+\citep[e.g.][]{2004ApJ...616..872R}. From stellar motions near
+Sgr~A\textsuperscript{*}, the distance from the Sun to Sgr~A\textsuperscript{*} can
+be precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a recent
+measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
 \citep{2019arXiv190405721A}.
 
+Importantly, the location of Sgr~A\textsuperscript{*} may not be equivalent to the location of the ``dynamical Galactic center:'' the point in three-dimensional
+space about which the stars in the solar neighborhood are orbiting.
+This assumption, although frequently made, has not yet been justified.
+It is tested in any construction of a  dynamical model where the distance from the Sun to
+the Galactic center, $R_0$, is a free parameter.
+For example, \citet{2015ApJ...803...80K} measured $R_0$ while
+modelling the dynamics of the stream Palomar 5. Many other dynamical measurements
+of $R_0$ have been made (\citealt{2016ARAA..54..529B} summarize many pre-{\em Gaia} results) but none have yet achieved a precision comparable to that of the distance to Sgr~A\textsuperscript{*}.
+% \citep{1981gask.book.....M, 2011PASJ...63..867S, 2012MNRAS.427..274S,2013AstL...39...95B, 2013IAUS..289..444Z}.
+
 \subsection{Galactic Orientation}
-Second, one must define the angular orientation of the Galaxy. This was
-defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
+Second, one must define the angular orientation of the Galaxy. This
+was defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
 defining the coordinates of the Galactic center in B1950 coordinates as
-(17:42:26.6, -28:55:00) and the North Galactic pole as (12:49:00, +27:24:00).
-These two quantities, together with $R_0$, define the orientation of the
-Galactic plane. However, there is growing evidence that the stellar midplane
-is tilted relative to this coordinate system \citep{2014ApJ...797...53G,
-2016ARAA..54..529B}, though not the H~\textsc{ii} midplane
-\citep{2019ApJ...871..145A}.
+(17:42:26.6, -28:55:00) and the north Galactic pole as (12:49:00, +27:24:00). These two quantities, together with $R_0$, define the orientation of the Galactic plane.
+However, there is growing evidence that the stellar midplane is tilted relative to this coordinate system \citep{2014ApJ...797...53G, 2016ARAA..54..529B}, though not
+the H~\textsc{ii} midplane \citep{2019ApJ...871..145A}.
 
 \subsection{Solar Height}
-Third, one must define the Sun's vertical distance from the Galactic midplane,
-which can be determined by identifying where the stellar density and
-velocities reach a maximum (effectively the median height of all disk stars).
-The Solar height is usually taken to be $\sim 25\,\pc$
-\citep{2001ApJ...553..184C}, with a more recent measurement from \textit{Gaia}
-DR2 placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another
-strategy is to use the cold gas or H~\textsc{ii} regions in the disk to define
-the Galactic midplane, leading to slightly different values (by $\sim 5\,\pc$)
-for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em
-Gaia} review of these measurements is given by \citet{2016ARAA..54..529B}.
+Third, one must define the Sun's vertical distance from the Galactic
+midplane, which can be determined by identifying where the stellar density
+and velocities reach a maximum (effectively the median height of all disk
+stars). The Solar height is usually taken to be $\sim 25\,\pc$
+\citep{2001ApJ...553..184C}, with a more recent measurement from {\em Gaia} DR2
+placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another strategy is
+to use the cold gas or H~\textsc{ii} regions in the disk to define the Galactic
+midplane, leading to slightly different values ($\sim 5\,\pc$) for the Sun's
+relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review
+of these measurements is given by \citet{2016ARAA..54..529B}.
 
 \subsection{Local Standard of Rest}
-Finally, one must define the LSR, or mean velocity of stars near the Sun
-relative to the Galactic center (which is defined to have zero velocity), and
-the velocity of the Sun relative to the LSR. The radial ($U_{\odot}$) and
-vertical ($W_{\odot}$) components are computed by taking the mean motions of
-different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal
-component ($V_{\odot}$) is more difficult to measure, but can be modeled using
-the asymmetric drift relation \citep{2008gady.book.....B}. The values of the
+Finally, one must define the local standard of rest (LSR), or mean velocity of stars near the Sun relative to the Galactic center (which is defined to have zero velocity), and the velocity of the Sun relative to the LSR.
+The radial
+($U_{\odot}$) and vertical ($W_{\odot}$) components are computed by
+taking the mean motions of different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal component
+($V_{\odot}$) is more difficult to measure, but can be modelled using the
+asymmetric drift relation \citep{2008gady.book.....B}. The values of the
 components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}. The
@@
 
-Considerable effort has been placed on each of these measurements, but
-uncertainties remain, and detailed dynamical modeling across the disk ---
-particularly for dynamically cold stars --- may have to take them into
-account. For instance, the discrepancy of $\sim 5\,\pc$ between the midplane
-defined by stars and by gas will induce a $\sim10\%$ error in $J_z$ for an
-orbit with $z_{\text{max}}\sim100\,\pc$ (see Section~\ref{sec:ref_frame}).
+Considerable effort has been placed on each of these
+measurements, but uncertainties remain, and detailed dynamical modelling
+across the disk --- particularly for dynamically cold stars --- may have to
+take them into account. For instance, the discrepancy of $\sim 5\,\pc$ between
+the midplane defined by stars and by gas will induce a $\sim10\%$ error in
+$J_z$ for an orbit with $z_{\text{max}}\sim100\,\pc$ (see
+Section~\ref{sec:ref_frame}).
 
@@
 We now consider a second source of systematic offsets in the Galactocentric
-coordinate system: intrinsic variations of the midplane across the disk. In
-this work we specifically consider azimuthal variations in the midplane as
-defined by the stellar mass density. We use two sets of simulations of Milky
-Way-mass galaxies to estimate the size of this effect.
+coordinate system: intrinsic variations of the midplane across the disk. We
+use two sets of simulations of Milky Way-mass galaxies to estimate the size of this effect. We will also
+briefly discuss a rough estimate of the midplane variation derived from the
+observed velocity distribution of the Milky Way, but defer a detailed
+treatment to future work.
 
 \subsection{Description of FIRE Simulations} \label{ssec:cosmozoom}
-The FIRE cosmological hydrodynamic simulations
-\citep{2014MNRAS.445..581H,2018MNRAS.480..800H} use the zoom-in technique
-\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O} to model the formation
-of a small group of galaxies at high resolution in a full cosmological
-context.  Feedback from supernovae, stellar winds, and radiation from massive
-stars is implemented at the scale of star forming regions following stellar
-population synthesis models, generating galactic winds self-consistently
-\citep{2015MNRAS.454.2691M, 2017MNRAS.470.4698A} while reproducing many
-observed galaxy properties, including stellar masses, star formation
-histories, metallicities, and morphologies and kinematics of thin and thick
-disks \citep{2014MNRAS.445..581H, 2016MNRAS.456.2140M, 2017MNRAS.467.2430M,
-2016ApJ...827L..23W, 2018MNRAS.481.4133G, 2018MNRAS.480..800H}. For this work,
-we focus on the three Milky Way-mass zoom-ins considered in
-\citet{2018arXiv180610564S}, which were simulated as part of the
-\textit{Latte} suite and show broad agreement of many of their global
-properties with observations of the Milky Way \citep{2016ApJ...827L..23W,
-2018MNRAS.481.4133G}. The $\z = 0$ snapshots\footnote{In this work, to avoid
-confusion with the vertical height $z$, we refer to cosmological redshift as
-$\z$.} of these three simulations, named \mi{}, \mf{}, and \mm{} for
-shorthand, are publicly available alongside associated mock \textit{Gaia} DR2
+We briefly describe the cosmological-hydrodynamical zoom-in simulations used
+in this work. Cosmological zoom-ins allow one to simulate a selected region at
+high resolution embedded in a low-resolution cosmological background
+\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added
+a hydrodynamical simulation of the baryonic component with recipes for star
+formation and feedback. The FIRE collaboration has simulated a number of Milky
+Way-mass galaxies using this technique as part of the {\em Latte} suite of
+FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. For this
+work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S},
+which show broad agreement of many of their global properties with
+observations of the Milky Way. The $\z=0$ snapshots\footnote{In this work, to
+avoid confusion with the vertical height $z$, we refer to cosmological
+redshift as $\z$.} of these three simulations, called \mi{}, \mf{}, and \mm{}
+for shorthand, are publicly available alongside associated mock {\em Gaia} DR2
 catalogues generated from them.\footnote{\url{http://ananke.hub.yt}}
 
-\begin{deluxetable*}{cccccc}[htb!]
-\tablecaption{Stellar and gas disk scale heights of the real Milky Way and the
-simulated galaxies considered in this work. For comparison, we also give the
-softening lengths for the simulated galaxies.\label{tab:scale_height}}
+These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
+gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles
+of mass $\sim 5000 \text{--} 7000\,
+\Msun$,\footnote{In fact, the simulation \mi{} contained a gas
+splitting bug which causes a small number of gas and star particles to have
+larger than typical masses. These constitute only $\sim0.2\%$ of the star
+particles in the $\z=0$ snapshots, and are only a factor of a few more massive
+than typical star particles. Since our results are broadly consistent between
+the three simulations, we ignore this minor complication.} with the lower end
+coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
+for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
+respectively.\footnote{This is $2.8$ times the often-quoted
+Plummer-equivalent.} The gas softening length is adaptive, but at $\z=0$ the
+median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
+roughly solar positions (with cylindrical radii within $500\,\pc$ of
+$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi{},
+\mf{}, and \mm{}, respectively. These values are summarized in
+Table~\ref{tab:scale_height}.
+
+\begin{deluxetable*}{cccccc}
+\tablecaption{Stellar disk scale heights of the real Milky Way and the
+simulated galaxies considered in this work.\label{tab:scale_height}}
 \tablehead{\colhead{galaxy} & \colhead{cold\tablenotemark{a} gas disk} &
@@
 softening length} & \colhead{stellar softening length} \\
-\colhead{} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)}
-& \colhead{(pc)} }
+\colhead{} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} } 
 \startdata
-Milky Way\tablenotemark{b} & 40 & 300 & 900 & \nodata & \nodata \\
+MW\tablenotemark{b} & 40 & 300 & 900 & \nodata & \nodata \\
 \mi\tablenotemark{c} & 800\tablenotemark{d} & 480 & 2000 & 12.4 & 11.2 \\
@@
 
-These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
-gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles of
-mass $\sim 5000 \text{--} 7000\, \Msun$, with the lower end coming from
-stellar evolution \citep{2018arXiv180610564S}. Softening lengths for dark
-matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
-respectively.\footnote{This is $2.8$ times the often-quoted
-Plummer-equivalent.} The gas softening length is adaptive, but at $\z=0$ the
-median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
-roughly solar positions (with galactocentric cylindrical radii within
-$0.5\,\kpc$ of $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4$, $12.5$, and
-$10.3\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. These values are
-summarized in Table~\ref{tab:scale_height}, along with measurements of the
-stellar and gas disk scale heights.
-
 The softening lengths used in the simulations can affect the ability to
@@
 dense, star-forming gas disk is thought to have a scale height of about
-$40\,\pc$, $\sim3\textup{--}4$ times the cold gas softening length
-\citep{2019ApJ...871..145A}. The thin stellar disk has a scale height of about
-$300\,\pc$, $\sim30$ times the stellar softening length
-\citep{2008ApJ...673..864J}. We therefore expect that resolution effects may
-still be affecting the scale heights of these components in the simulations,
-especially the cold gas. Indeed, the stellar scale heights of the simulated
-galaxies are equal to or larger than the Milky Way's while the gas scale
+$40\,\pc$ ($3\text{-}4$ times the cold gas softening length;
+\citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of
+about 300 pc (30 times the stellar softening length;
+\citealt{2008ApJ...673..864J}). We therefore expect that resolution effects
+may still be affecting the scale heights of these components in the
+simulations, especially the cold gas. Indeed, the stellar scale heights of the
+simulated galaxies are equal to or larger than the MW's while the gas scale
 heights are significantly larger (although the proper basis comparison is less
@@
 precluding extending the precision of the gas midplane definition to the
-stellar component.
+stellar component. In this work we specifically consider azimuthal variations
+in the midplane \emph{as defined by the stellar mass density}, which will
+subsequently be lower in the simulations than in the MW: within 200 pc of the
+Sun the measured gas and stellar volume densities in the MW are roughly
+comparable at $\sim 0.04\ \Msun\
+\unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
+Table 3 of \citealt{2018arXiv180610564S}). We thus expect that the degree of
+midplane variation we find in the simulated systems is likely to be
+\emph{less} than in the real MW, since on the smallest scales the simulations
+are \emph{smoother} than the real Galaxy.
 
 Cosmological simulations of Milky Way-mass galaxies are not perfect
-representations of the true Milky Way in other ways as well,\footnote{The
-failure of cosmological simulations to exactly reproduce the Milky Way is not
-entirely due to limitations of the numerical model. Candidate Milky Way-like
-galaxies are chosen solely on their mass and isolation, for which there are a
-wide variety of possible galaxies with qualitatively different properties.} as
-discussed in \citet{2018arXiv180610564S}. For instance, the velocity structure
-of \mi{} is closer to M31's than the Milky Way's (Loebman et al. in prep).
-However, in this work we are most interested in the global properties of the
-potential, and specifically in deviations from axisymmetry. From this
-perspective, the simulated galaxies are actually \emph{more} axisymmetric than
-we might expect of the Milky Way. While they have prominent spiral arms, none
-has as strong a bar as the Milky Way does at present day, and none has a
-nearby companion like the Large Magellanic Cloud. One of the three we consider
-(\mf) does have an ongoing interaction with a satellite galaxy similar to
-Sagittarius, which has punched through the galactic disk outside the solar
-circle and induced some warping.
+representations of the true Milky Way in other ways as well, as discussed in
+\citet{2018arXiv180610564S}. For instance, the velocity structure of \mi{} is
+closer to M31's than the Milky Way's (Loebman et al. in prep). However, in
+this work we are most interested in the global properties of the potential,
+and specifically in deviations from axisymmetry. From this perspective, the
+simulated galaxies are actually \emph{more} axisymmetric than we might expect
+of the Milky Way. While they have prominent spiral arms, none has as strong a
+bar as the MW does at present day, and none has a nearby companion like the
+Large Magellanic Cloud. One of the three we consider (\mf) does have an
+ongoing interaction with a satellite galaxy similar to Sagittarius, which has
+punched through the galactic disk outside the solar circle and induced some
+warping. However, we expect that, as with the use of finite smoothing lengths,
+the degree of midplane variation in the simulations is likely a factor of a
+few lower than what could be expected in the MW.
 
@@
 Section 3 of \citet{2018arXiv180610564S} as our fiducial coordinate system for
-each galaxy. In short, the center of the galaxy is found iteratively. The
+each galaxy. In short, the center of the galaxy is found iteratively, the
 center of mass velocity is then determined by all star particles within
-$15\,\kpc$ of this center. The galaxy is then rotated onto a principal axis
-frame determined by stars younger than $1\,\Gyr$ inside of the fiducial solar
-radius $R_{0} = 8.2\,\kpc$, such that the disk plane is the $x$--$y$ plane.
-
-\subsection{Description of Milky Way-Sagittarius Interaction Simulation} 
-\label{ssec:sag_sim}
-In addition to the cosmological zoom-ins, we will also briefly consider
-results from a live N-body simulation of a Sagittarius-like encounter. This
-simulation offers us the ability to see how the midplane varies in a more
-controlled environment. The details of the simulation are given in
-\citet{2018MNRAS.481..286L}, but we briefly summarize the most relevant
-details here. For the Milky Way, the dark halo is modeled as a Hernquist
-sphere of mass $10^{12}\,\Msun$ and scale length of $52\,\kpc$, the disk
-modeled as an exponential disk with a scale radius of $3.5\,\kpc$, scale
-height $0.53\,\kpc$, and mass $6\times10^{10}\,\Msun$, and a Hernquist bulge
-of mass $10^{10}\,\Msun$ and scale length $0.7\,\kpc$
-\citep{1990ApJ...356..359H}. We use the L2 model, for which Sagittarius is
-modeled as a dark matter Hernquist sphere of mass $8\times10^{10}\,\Msun$ and
-scale length $8\,\kpc$, along with a stellar component modeled as a Hernquist
-sphere of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$. All
-components are realized with live distributions of N-body particles; the Milky
-Way and Sagittarius are each initialized to be in equilibrium in isolation
-\Gus{check!}.
+$15\,\kpc$ of this center, and the galaxy is then rotated onto a
+principal-axis frame determined by stars younger than $1\,\Gyr$ inside of the
+fiducial solar radius $R_{\odot} = 8.2\,\kpc$, such that the disk plane is the
+$x$--$y$ plane.
+
+\subsection{Description of MW-Sagittarius Interaction Simulation} \label{ssec:sag_sim}
+In addition to the cosmological zoom-in, we will also briefly consider results
+from a toy Sagittarius encounter simulation. This simulation offers us the
+ability to see how the midplane varies in a more controlled environment. The
+details of the simulation are given in \citet{2018MNRAS.481..286L}, but we
+briefly summarize the most relevant details here. For the Milky Way, the dark
+halo is modeled as a Hernquist sphere of mass $10^{12}\,\Msun$ and scale
+length of $52\,\kpc$, the disk modeled as an exponential disk with a scale
+radius of $3.5\,\kpc$, scale height $0.53\,\kpc$, and mass
+$6\times10^{10}\,\Msun$, a Hernquist bulge of mass $10^{10}\,\Msun$ and scale
+length $0.7\,\kpc$ \citep{1990ApJ...356..359H}. Sagittarius is modelled as a
+dark matter Hernquist sphere of mass \Gus{something} and scale length
+\Gus{something}, along with a stellar component modeled as a Hernquist sphere
+of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$.
+All components are realized with live distributions of N-body particles; the MW and Sgr are each initialized to be in equilibrium in isolation \Gus{check!}.
 
 The mass resolution of the simulation is $2.6\times10^4$, $1.2\times10^4$, and
-$1.0\times10^4\,\Msun$ for the dark matter, disk, and bulge components,
-respectively. For the disk and bulge components, a softening length of
-$30\,\pc$ is used whereas for the halo a softening length of $60\,\pc$ is
-used. For Sagittarius, the softening length for the dark matter and the stars
-is $60$ and $40\,\pc$, respectively.
-
-The fiducial coordinate system for these N-body simulations is the rest frame
-of the aligned host galaxy at the beginning of the simulation.
+$1.0\times10^4$ for the dark matter, disk, and bulge components, respectively.
+For the disk and bulge components, a softening length of $30\,\pc$ is used
+whereas for the halo a softening length of $60\,\pc$ is used. For Sagittarius,
+the softening length for the dark matter and the stars is $60$ and $40\,\pc$,
+respectively.
 
 \subsection{The Local Midplane} \label{ssec:local_midplane}
-Using the two sets of simulations, we determine the local midplane that an
-observer might measure if they were situated in each of these galaxies as a
-function of azimuth at the solar circle. Starting from the coordinate system
-described in the previous section, which is approximately aligned so that the
+Using the two sets of simulations, we determine the local midplane that an observer
+might measure if they were situated in each of these galaxies as a function of
+azimuth at the solar circle. Starting from the coordinate system described in
+the previous section, which is approximately aligned so that the
 $z$-coordinate is perpendicular to the disk plane, we place our imaginary
-observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and
-vary the azimuth between $0<\phi<2\pi$. At each value of $\phi$ we then
-compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and
-height $1\,\kpc$ perpendicular to the fiducial disk. We then re-define the new
-midplane of the cylinder to be the median $z$, re-select stars, and iterate
-until the median $z$ value converges. We find that only $10$ iterations of
-this procedure are necessary for convergence. The resulting median $z$ is
-taken to be what our observer would measure as the local Galactic midplane at
-each $\phi$.
-
-\begin{figure*}[htb!]
-\begin{center}
-\includegraphics[width=\textwidth]{fig/midplane_fit.pdf}
-\end{center}
-\caption{The local midplane determined at the fiducial Solar radius
-($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} (left,
-center, and right panels) as a function of azimuthal angle. All calculations
-are done at cosmological redshift $\z =0$. The local midplane is determined at
-a position $\phi$ by taking the median height of all stars within
-$R=0.5\,\kpc$ and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is
-performed again using the new height $10$ times to converge on the local
-midplane height. In order to allow for the possibility that the fiducial
-galactocentric coordinate system is incorrect, we subtract the best fit
-$A\sin{(\phi+B)}+C$ curve from each panel --- this figure is reproduced with
-the original midplane determination (i.e. before subtracting the best fit sine
-curve) in Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on
-all stars within a $2\,\kpc$ height of the fiducial midplane to determine
-$1\,\sigma $ error bars, which we report as dashed lines.}
-\label{fig:midplane}
-\end{figure*}
-
-\begin{figure*}[htb!]
-\begin{center}
-\includegraphics[width=342.078286667pt]{fig/midplane_fit_chervinsim.pdf}
-\end{center}
-\caption{The local midplane determined at the fiducial Solar radius
-($8.2\,\kpc$) for 4 different time steps from live N-body simulations of a
-Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
-before, we have subtracted the best fit $A\sin{(\phi+B)}+C$ curve to account
-for inaccuracies in the galactocentric coordinate system. Error bars are
-calculated as in Figure~\ref{fig:midplane}. (\Gus{Check:}) The upper panels
-show the midplane as a function of azimuth before the first encounter near the
-Solar circle at $t=2.0\,\Gyr$ and $t=4.0\,\Gyr$. As later encounters occur
-which bring the Sagittarius-like object closer to the Solar circle, we can see
-that stronger midplane variations result. In fact, one can see that an $m=2$
-mode is excited in the lower left panel at $t=6.0\,\Gyr$ --- this is
-consistent with Figure~17 of \citet{2018MNRAS.481..286L}, which shows that an
-$m=2$ mode is excited (note that $m=0$ and $m=1$ modes are stronger, but these
-are removed in our sine-curve subtraction). The fact that the $t=6.9\,\Gyr$
-panel looks qualitatively similar to the panels from the FIRE simulations
-(Figure~\ref{fig:midplane}) is evidence that the midplane variations are
-produced in part by mergers.}
-\label{fig:midplane_chervin}
-\end{figure*}
-
-This procedure assumes perfect density estimation, and therefore perfect
-corrections for extinction within the cylinder defining the ``solar
-neighborhood.'' Imperfect extinction correction is likely to increase the
-amplitude of the estimated fluctuations in $z$.
-
-To account for the effect of particle noise, we bootstrap resample stars
-within a cylinder of stars with height $2\,\kpc$ and the same radius $1000$
-times and determine the $1\,\sigma$ error bars by repeating the midplane
-determination with those reselection of stars.
+observer at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth
+between $0<\phi<2\pi$.\footnote{The immortal observer also has warp
+drive.} At each value of $\phi$ we then compute the median $z$ for stars
+within a cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to
+the fiducial disk. We then re-define the new midplane of the cylinder to be the median $z$, re-select stars, and iterate until the median $z$ value
+converges. We find that only $10$ iterations of this procedure are necessary
+for convergence. The resulting median $z$ is taken to be what our observer
+would measure as the local Galactic midplane at each $\phi$. This procedure
+assumes perfect density estimation, and therefore perfect corrections for
+extinction, within the cylinder defining the ``solar neighborhood.'' Imperfect
+extinction correction is likely to increase the amplitude of the estimated
+fluctuations in $z$. To account for the effect of particle noise, we
+bootstrap-resample stars within a cylinder of stars with height $2\,\kpc$ and
+the same radius $1000$ times and determine the $1\,\sigma$ error bars by repeating the midplane determination with those reselection of stars.
 
@@
 each galaxy, using their local ``solar neighborhood'' (the cylinder defined
-above). We consider $50$ equally spaced bins in azimuth, a sufficient number
-such that each cylinder shares no stars with its neighboring cylinder. The
-$1\,\sigma$ error from the bootstrap procedure is shown as the dashed-line
-error bars. The middle $90\%$ of midplane values across the Solar circle spans
-$185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the three cases
-the midplane therefore varies by more than $\pm 100\,\pc$ depending on the
-azimuth along the Solar circle; in the third (\mm{}, which has the thinnest
-``thin disk'' of stars, but the largest stellar mass) the variation is closer
-to $\pm 50\,\pc$.
+above). The $1\,\sigma$ error from the bootstrap procedure is shown as the
+dashed-line error bars. The middle $90\%$ of midplane values across the Solar
+circle spans $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of
+the three cases the midplane therefore varies by more than $\pm 100\,\pc$
+depending on the azimuth along the Solar circle; in the third (\mm{}, which
+has the thinnest ``thin disk'' of stars, but the largest stellar mass) the
+variation is more like $\pm 50\,\pc$.
 
 We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
-but for four succcessive timesteps of the live N-body simulation of a
-Sagittarius encounter \citep{2018MNRAS.481..286L}. Again we have subtracted a
-best fit curve of the form $\text{A} \sin{\left(\phi + \text{B}\right)} +
-\text{C}$, with the values for $\text{A}$ being $9.5$, $2.5$, $-212$, and
-$-394\,\pc$, for $\text{B}$ being $0.0013$, $0.0007$, $-0.63$, and
-$-1.00\,\text{rad}$ and $\text{C}$ being $2.6$, $-7.8$, $-65$, and $-53\,\pc$
-chronologically for each of the four timesteps shown.
-
-In the upper panels, we see that the midplane is relatively flat in the inner
-galaxy, but additional encounters drive strong midplane variation. In the
-lower left panel, we see a strong $m=2$ mode develop, consistent with the
-$R=8\,\kpc$ panel of Figure~17 in \citet{2018MNRAS.481..286L}. The lower right
-panel, which shows the galaxy after some relaxation has occured is
-qualitatively similar to the midplane variations we saw in the FIRE
-simulations (Figure~\ref{fig:midplane}), evidence that they are at least
-partially driven by mergers.
+but for the toy simulation of a Sagittarius encounter
+\citep{2018MNRAS.481..286L}. Again we have subtracted a best fit curve of the
+form $\text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}$, with the values
+for $\text{A}$ being $9.5$, $2.5$, $-212$, and $-394\,\pc$, for $\text{B}$ being
+$0.0013$, $0.0007$, $-0.63$, and $-1.00\,\text{rad}$ and $\text{C}$ being $2.6$, $-7.8$,
+$-65$, and $-53\,\pc$. Each of these values are for each successive timestep. In the
+first two panels, we see that the midplane is relatively flat in the inner
+galaxy. But additional encounters drive strong midplane variation. In the
+third panel, we see a strong $m=2$ mode develop, consistent with the
+$R=8\,\kpc$ panel of Figure~17 in \citet{2018MNRAS.481..286L}.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{fig/midplane_fit.pdf}
+\end{center}
+\caption{The local midplane determined at the fiducial Solar radius
+($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} ({\em left},
+{\em center}, and {\em right} panels). All calculations are done at cosmological redshift $\z =0$. The local midplane is determined at a
+position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
+and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
+using the new height $10$ times to converge on the local midplane height. In
+order to allow for the possibility that the fiducial galactocentric coordinate
+system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
+each panel --- this figure is reproduced with the original midplane
+determination (i.e. before subtracting the best fit sine curve) in
+Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on all stars
+within a $2\,\kpc$ height of the fiducial midplane to determine $1\,\sigma
+$ error bars, which we report as dashed lines.}
+\label{fig:midplane}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=342.078286667pt]{fig/midplane_fit_chervinsim.pdf}
+\end{center}
+\caption{The local midplane determined at the fiducial Solar radius
+($8.2\,\kpc$) for 4 different time steps from previously run toy simulations
+of a Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
+before, we have subtracted the best fit $A\sin{(\phi+B)}+C$ curve to account
+for inaccuracies in the galactocentric coordinate system. Error bars are
+calculated as in Figure~\ref{fig:midplane}. (\Gus{Guessing here, need to
+figure out what the panels actually show:}) The left two panels show the
+midplane as a function of azimuth before the first encounter. As later
+encounters occur, we can see that stronger midplane variations result. In
+fact, one can see that an $m=2$ mode is excited in the third panel --- this is
+consistent with Figure~17 of \citet{2018MNRAS.481..286L}, which shows that an
+$m=2$ mode is excited (note that $m=0$ and $m=1$ modes are stronger, but these
+are removed in our sine-curve subtraction).}
+\label{fig:midplane_chervin}
+\end{figure*}
 
 \subsection{Velocity Variations} \label{ssec:lsr_var}
-We also expect that the LSR should vary as a function of azimuth. We perform
-this calculation in Appendix~\ref{app:lsr} to estimate the components of the
-LSR as a function of azimuth, but performing a best-fit subtraction to correct
-for misalignment of the original coordinate system (as in the previous
-section) is more involved. Since we find that the variation in the LSR is less
-pronounced than for the midplane, and since offsets in velocity only
-contribute to second order to $\Delta J_R$ and $\Delta J_z$ when a star is at
-maximum amplitude in $R$ or $z$ (where the majority of the orbit is, see
-Section~\ref{ssec:epi_action}), we defer this calculation to future work.
+We also expect that the local standard of rest (LSR) should vary as a function
+of azimuth. We perform this calculation in Appendix~\ref{app:lsr} to estimate
+the components of the LSR as a function of azimuth, but performing a best-fit
+subtraction to correct for misalignment of the original coordinate system (as
+in the previous section) is more involved. Since we find that the variation in
+the LSR is less pronounced than for the midplane, and since the dominant
+effect is limited to $J_R$ (Figure~\ref{fig:many_orbit_wrong_ref}), we defer
+this calculation to future work.
+
+An interesting, qualitative calculation can be made based on the velocity
+variations seen in e.g. \citet{2019arXiv190209569F}. The velocity variations
+in this work (admittedly found with respect to $J_{\phi}$ or radius) are
+$\sim5\text{--}10\,\kms$. The vertical frequency of the thin-disk and
+thick-disk orbits are $\sim 0.09\,\Myr^{-1}$ and $\sim0.06\,\Myr^{-1}$,
+respectively. By dimensional analysis, we therefore expect the midplane
+offsets to be of order $\sim 57\text{--}170\,\pc$. We stress that this is a
+rough calculation, but it shows that already we see evidence in the data from
+velocities for midplane offsets on the order of what we see in both sets of
+simulations.
 
@@
 by up to $\pm 100$ pc, as a natural consequence of the non-axisymmetry of the
-Galactic disk at small scales.
-
-While this is not in itself surprising or new, we also demonstrated that
-extending the coordinate system established by our local midplane to a global
-axisymmetric coordinate system spanning the entirety of the Galactic disk
-introduces systematic error in computations of the actions under this symmetry
-assumption, when starting from the present-day positions and velocities of
-stars as measured by, e.g., \emph{Gaia}.
-
-These systematic errors are most important for stars on thin disk-like orbits,
-where they can be large enough to yield actions representative of orbits in
-the thick disk. This effect is entirely due to the extension of a local to a
+Galactic disk at small scales. While this is not in itself surprising or new,
+we also demonstrate that extending the coordinate system established by our
+local midplane to a global axisymmetric coordinate system spanning the
+entirety of the Galactic disk introduces systematic error in computations of
+the actions under this symmetry assumption, when starting from the present-day
+positions and velocities of stars as measured by, e.g., \emph{Gaia}. These
+systematic errors are most important for stars on thin disk-like orbits, where
+they can be large enough to yield actions representative of orbits in the
+thick disk. This effect is entirely due to the extension of a local to a
 global coordinate system, and is separate from real diffusion in stellar
@@
 
+Galactocentric coordinate system offsets can also be generated by errors in
+the parameters of the Galactocentric coordinate system. However, such
+parameters have been measured to high enough precision that they are likely
+subdominant to the intrinsic variation of the midplane across large distances.
+On shorter scales, they are likely the dominant source of error.
+
 This finding has many implications for the study of the dynamics of stars in
 what is normally considered the regime of the epicyclic approximation:
-quasi-harmonic oscillations around a circular ``guiding center orbit.''
-
-\subsection{Estimate of Milky Way Midplane Offsets from Current Data}
-\label{ssec:mw_data_midplane}
-
-Systematic variations in $v_z$ were first noted as asymmetries in the local
-velocity distribution towards the North and South Galactic Caps from the
-radial velocity surveys of the Sloan Digital Sky Surveys
-\citep{2012ApJ...750L..41W} and RAdial Velocity Experiment
-\citep{2013MNRAS.436..101W}. Subsequently, \citet{2013ApJ...777L...5C} pointed
-out suggestions of an oscillation in average vertical velocities of order
-$5\,\kms$  on $\sim\,\kpc$ scales looking toward the Galactic anticenter. Work
-by the \textit{Gaia} collaboration confirmed these preliminary results on the
-velocity and spatial scales of these oscillation with clear spatial maps made
-using DR2 data of median $v_z$ over a significant Galactic volume
-\citep{2018A&A...616A..11G, 2019arXiv190209569F}. We also see in the FIRE
-simulation that the vertical velocity variation as a function of azimuth is
-$\sim5\text{--}10\,\kms$ (Figure~\ref{fig:lsr_variations}), consistent with
-these observations.
-
-The vertical frequency of the thin-disk and thick-disk orbits are $\sim
-0.09\,\Myr^{-1}$ and $\sim0.06\,\Myr^{-1}$, respectively. By dimensional
-analysis, and assuming a vertical velocity variation of $5\,\kms$, we
-therefore expect the midplane offsets to be $\sim 57\textup{--}85\,\pc$. We
-stress that this is a rough calculation, but it shows that already we see
-evidence in the data from velocities for midplane offsets on the order of what
-we saw in both sets of simulations.
-
-\subsection{Coordinate System Offsets}\label{ssec:coord_off}
-In Section~\ref{sec:mes_err} we discussed the current observational efforts to
-measure the parameters of the Galactocentric coordinate system. We noted that
-the solar height is discrepant between gas-based and stellar-based
-determinations by $\sim5\,\pc$. While the effect is small and thus only likely
-to dominate the intrinsic midplane variations on small scales, it may be
-important for detailed modeling of kinematically cold stars where a $5\,\pc$
-offset is relevant.
-
-We also noted that the distance from the Sun to \sgra{} has been recently
-measured to a very low uncertainty of $\sim0.3\%$ or $\pm25\,\pc$
-\citep{2019arXiv190405721A}. However, the location of \sgra{} may not be
-equivalent to the location of the ``dynamical Galactic center,'' i.e., the
-point in three-dimensional space about which the stars in the solar
-neighborhood are orbiting. This assumption, although sensible and frequently
-made, has not yet been justified.
-
-If the dynamical Galactic center is offset from \sgra{} by $100\,\pc$, only a
-$1.2\%$ difference, then this induces a $21\%$ error in $J_R$ (see
-Section~\ref{ssec:quant}). The reason such a large error in $J_R$ can be
-generated by a small error in $R_0$ can be understood from the epicyclic
-approximation, which states that $\Delta J_R/J_R = 2\Delta R/A_R$. The
-fractional error in $J_R$ is related to the error in $R_0$ as a fraction of
-the {\em radial amplitude} of the orbit, which is much smaller than $R_0$
-($\sim1.2\,\kpc$ for the thin-disk and thick-disk orbits we considered). This
-also implies the very precise $0.3\%$ measurement of $R_0$ still translates to
-a $\sim4\%$ uncertainty in $J_R$.
-
-The assumption that the dynamical Galactic center and \sgra{} are colocated is
-tested in any construction of a dynamical model where $R_0$ is a free
-parameter. For example, \citet{2015ApJ...803...80K} measured $R_0$ while
-modeling the dynamics of the stream Palomar 5. Many other dynamical
-measurements of $R_0$ have been made (\citealt{2016ARAA..54..529B} summarize
-many pre-\textit{Gaia} results), but none have yet achieved a precision
-comparable to that of the distance to \sgra{}.
-
-We did not consider in this text the effect of the \emph{angular position} of
-the dynamical Galactic center being offset with \sgra{}.
+quasi-harmonic oscillations around a circular ``guiding center orbit.'' Here
+we consider a few of those implications, starting with an estimate of the
+region around the Sun where we expect an extension of the local axisymmetric
+coordinate system, and therefore the epicyclic approximation, to still be
+valid. This region is also where it should be feasible to perform ``dynamical
+tagging'' in the thin disk; that is, using the computed orbits or integrals of
+motion of field stars to match them with known open clusters.
+
+\subsection{Disrupted Cluster Reconstruction}
+\label{sssec:reconstruction}
+
+We now consider the implications for attempts to reconstruct disrupting open
+clusters by selecting ejected members based on their actions.
+
+Suppose an open cluster of mass $m_c$ on an orbit with actions $\bm{J}_c$ is
+disrupted. We would like to determine which stars with complete 6D phase-space
+positions, for example in the {\em Gaia} dataset, are likely to be ejected members from that cluster.\footnote{This is likely to
+be more successful with DR3/4 than DR2, since upcoming data releases will
+contain a far greater number of stars with radial velocities --- though
+targeted, follow-up radial velocity surveys may allow this program to be
+performed sooner.} One
+could generate an initial catalogue of candidate members by integrating the
+cluster's orbit and selecting stars close to that orbit, accounting for the
+fact that the stream of ejected stars will not exactly follow the orbit of the
+cluster \citep[e.g.][]{2011MNRAS.413.1852E}.
+
+One could then make a further selection by requiring that the action of each
+field star $i$ be close to the cluster's actions to within some bound, i.e.
+$\bm{J}_i \in V_e$, where $V_e$ is a volume centered on $\bm{J}_c$ and has size
+comparable to the intrinsic action-space dispersion of the disrupting
+cluster. This dispersion can be estimated as \citep[Section~8.3.3][]{2008gady.book.....B},
+\beq \label{eq:action_disp}
+\frac{\Delta J_i}{J_i} \sim \left(\frac{m_c}{M}\right)^{1/3}\text{,}
+\eeq
+where $m_c$ is the mass of the cluster and $M$ is the mass enclosed by the
+cluster's orbit around the Galaxy, in this case assumed to be the mass inside
+the Solar circle ($\sim 10^{11}\Msun$). Stars with actions within $V_e$ would
+then be considered cluster members.
+
+Two issues are immediately apparent with this approach. The first is that it
+neglects the overall smooth distribution \emph{in action space} which will
+contribute some stars to the selection by chance. For dissolving open clusters
+in the thin disk, this background is likely to be significant, but we will
+assume for the sake of argument that these foreground stars can be modeled
+within the ``dynamical solar neighborhood'' where it is safe to extend the
+local coordinate system. Since the effect of extending the coordinate frame
+beyond this region is generally to inflate the actions computed for thin-disk
+stars, using too large a volume is likely to under-estimate the density of the
+smooth thin-disk-like component in action space, while scattering more
+thin-disk stars onto thick-disk-like orbits and over-estimating the density of
+the smooth thick-disk-like component.
+
+The second issue, which is related to the point of this paper, is that based
+on the results in our previous sections computing $\bm{J}_i$ for a field star
+and directly comparing it to  $\bm{J}_c$ will have significant systematic
+error if the cluster and field star are far enough apart to have significantly
+different local midplanes (see Figures \ref{fig:midplane} and \ref{fig:midplane_chervin}). Since our criterion for cluster membership (distance in action space) scales with the
+cluster mass, the region where the search will succeed varies with cluster
+mass as well. To illustrate this effect, we assume a cluster mass $m_c$ and
+estimate the expected action-space dispersion from
+Equation~\ref{eq:action_disp}. We then consult
+Figure~\ref{fig:many_orbit_wrong_ref} to determine the midplane offset
+necessary in order for the induced error in the vertical action to be the same
+as the expected action-space dispersion. We perform this analysis for the
+thin-disk-like and thick-disk-like orbits considered in
+Section~\ref{ssec:quant}, assuming that the halo-like orbit is unlikely for
+members of known open clusters (see Table~\ref{tab:real_clusters}). The result
+for the allowable midplane offset for these two orbits in shown in
+Figure~\ref{fig:cluster_offset} as a function of cluster mass.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{fig/cluster_offset.pdf}
+\end{center}
+\caption{The $z$ offset necessary in order for the action error induced by our
+midplane effect to be comparable to the intrinsic action dispersion from a
+cluster of mass $m_c$. We compute the intrinsic dispersion from
+Equation~\ref{eq:action_disp} and convert the dispersion to a $z$ offset using
+the result from Figure~\ref{fig:many_orbit_wrong_ref}. We consider the result
+for the thin and thick orbit (Table~\ref{tab:orbits}). Nearby open clusters
+(Table~\ref{tab:real_clusters}) have orbits closest to the thin orbit.}
+\label{fig:cluster_offset}
+\end{figure}
 
@@
 \item Inaccuracy in the Galactocentric coordinate
-system induces phase-dependence in the actions calculated from the observed
-positions and velocities of stars
-(Figures~\ref{fig:cartoon}~and~\ref{fig:one_orbit_wrong_ref}). Since stars'
-instantaneous phase-space positions are measured without prior knowledge of
-their orbital phases, this results in systematic error in the computed actions
+system induces phase-dependence in the actions calculated from the observed positions and velocities of stars
+(Figures~\ref{fig:cartoon}~and~\ref{fig:one_orbit_wrong_ref}).
+Since stars' instantaneous phase-space positions are measured without prior knowledge of their orbital phases, this results in systematic error in the computed actions
 (Figure~\ref{fig:many_orbit_wrong_ref}).
 
-\item Inaccuracy in the midplane location most severely affects computation of
-the vertical action $J_z$.  A midplane offset of $\sim6\,\pc$ ($\sim50\,\pc$)
-for a typical thin (thick) disk orbit results in a  $10\%$ error in $J_z$ (as
-defined by one half the middle $90\uth$ percentile). The fractional error is
-significantly less for halo orbits.
+\item Inaccuracy in the midplane location most severely affects computation of the vertical action $J_z$.
+A midplane offset of $\sim6\,\pc$ ($\sim50\,\pc$) for a typical thin (thick) disk orbit results in a  $10\%$ error in $J_z$ (as defined by one half the middle
+$90\uth$ percentile). The fractional error is significantly less for halo orbits.
 
 \item The distribution of systematic errors in the actions induced by a
-coordinate system offset is very non-Gaussian. The distribution is bimodal
-with \emph{neither mode at null}. As a result, error propagation of coordinate
-system offsets is complex when considering actions, and is likely to
-significantly deform the action-space distribution function.
-
-\item Dynamical modeling across large regions of the disk, over which the
-midplane location varies by more than the limits discussed above, is
-susceptible to this type of systematic error, since the assumption that our
-local Galactic midplane is the global Galactic midplane is not true a
-priori.
+coordinate system offset is very non-Gaussian. The distribution is bimodal with
+{\em neither mode at null}. 
+As a result, error propagation of coordinate system offsets is complex when considering actions, and is likely to significantly deform the action-space distribution function.
+
+\item Dynamical modelling across large regions of the disk, over which the midplane location varies by more than the limits discussed above, is
+susceptible to this type of systematic error, since the assumption that our local
+Galactic midplane is the global Galactic midplane is not true {\em a priori}.
 
@@
 simulations of Milky Way mass galaxies from the FIRE collaboration. We found
-that the midplane varies as a function of azimuth at the Solar circle by
-$\sim185, 162, 84\,\pc$ (middle $90\%$) for the three galaxies we considered
-(\mi{}, \mf{}, and \mm{}, respectively).
-
-\item Assuming a vertical velocity variation of the Milky Way of
-$\sim5\,\kms$, consistent with recent results from \textit{Gaia} DR2
-\citep{2018A&A...616A..11G, 2019arXiv190209569F} and our results from the FIRE
-simulations (Figure~\ref{fig:lsr_variations}), we estimated that the
-corresponding midplane offsets are $\sim60\textup{--}90\,\pc$ by dimensional
-analysis with the vertical frequencies of disk-like orbits.
-
-\item While inaccuracies in the parameters of the currently adopted
-Galactocentric coordinate system are likely to be unimportant for most
-applications, we discussed the importance of testing the assumption that the
-dynamical Galactic center is colocated with \sgra{}, which we mentioned how to
-do.
+that the midplane varies as a function of azimuth by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
+three galaxies we considered (\mi{}, \mf{}, and \mm{}, respectively).
 
@@
 
-\item While in this work we have focused on systematic errors in action
-computation, all of our conclusions also extend to studies of stars that
-simply rely on orbit integration, since the computation of actions and orbit
-integrations are mathematically equivalent.
+\item While in this work we have focused on systematic errors in action computation, all of our conclusions also extend to studies of stars that simply rely on orbit
+integration, since the computation of actions and orbit integrations are
+mathematically equivalent.
 
@@
 Our main point is that the local midplane varies between different points in
-the Galaxy, and that this variation can lead to significant systematic errors
-in the computation of actions under the assumption of a global axisymmetric
-potential. We do not propose here what a modeler should use for the ``true''
-global midplane, since it depends on the particular problem being studied.
-Even without a high-precision definition of the global Galactic midplane,
-current observations from \textit{Gaia} should soon permit a measurement of the
-real azimuthal dependence of the midplane location. For some applications,
-such as those using actions as labels to group stars on similar orbits, using
-such a measurement to shift stars to a consistent midplane height as a
-function of azimuth before using a global axisymmetric approximation to the
-potential may be sufficient, although this ignores the \emph{dynamical}
-implications of shifts in the midplane height (which result from fluctuations
-in the local density). However, for other applications, such as the study of
-action diffusion which motivated this paper in the first place, a more
-extensive perturbative approach is likely to be needed. We plan to explore the
-mitigation of these effects in future work.
+the Galaxy, and that this variation can lead to significant systematic errors in the computation of actions under the assumption of a global axisymmetric potential. We do
+not propose here what a modeler should use for the ``true'' global midplane,
+since it depends on the particular problem being studied. Even without a
+high-precision definition of the global Galactic midplane, current
+observations from {\em Gaia} should soon permit a measurement of the real
+azimuthal dependence of the midplane location.
+For some applications, such as those using actions as labels to group stars on similar orbits, using such a measurement to shift stars to a consistent midplane height as a function of azimuth before using a global axisymmetric approximation to the potential may be sufficient, although this ignores the \emph{dynamical} implications of shifts in the midplane height (which result from fluctuations in the local density).
+However, for other applications, such as the study of action diffusion which motivated this paper in the first place, a more extensive perturbative approach is likely to be needed.
+We plan to explore the mitigation of these effects in future work.
 
 \acknowledgments
-We would like to thank Megan Bedell, Robert A. Benjamin, Tobias Buck, Elena
-D'Onghia, Benoit Famaey, and Adrian Price-Whelan for helpful discussions. A.B.
-would like to thank Todd Phillips for helpful discussions. A.B. was supported
-in part by the Roy \& Diana Vagelos Program in the Molecular Life Sciences and
-the Roy \& Diana Vagelos Challenge Award. K.V.J.'s and C.F.P.L.'s work was
-performed in part during the Gaia19 workshop and the 2019 Santa Barbara Gaia
-Sprint (also supported by the Heising-Simons Foundation), both hosted by the
-Kavli Institute for Theoretical Physics at the University of California, Santa
-Barbara. This work used the Extreme Science and Engineering Discovery
-Environment (XSEDE), which is supported by National Science Foundation grant
-number OCI-1053575. K.V.J.'s contributions were supported in part by the
-National Science Foundation under grants NSF PHY-1748958 and NSF AST-1614743.
-M.-M.M.L. was partly supported by NSF grant AST18-15461. The work of D.A.-A.,
-A.B., D.W.H., R.S., K.V.J., M.-M.M.L., and M.K.N. is supported by the Simons
-Foundation.
+We would like to thank Megan Bedell, Tobias Buck, Elena D'Onghia, Benoit
+Famaey, and Adrian Price-Whelan for helpful discussions. A.B. would like to
+thank Todd Phillips for helpful discussions. A.B. was supported in part by the
+Roy \& Diana Vagelos Program in the Molecular Life Sciences and the Roy \&
+Diana Vagelos Challenge Award. M.-M.M.L. was partly supported by NSF grant
+AST18-15461. The work of D.A.-A., A.B., D.W.H., C.F.P.L., R.S., K.V.J.,
+M.-M.M.L., and M.K.N. is supported by the Simons Foundation.
 
 \appendix
-\section{Orbits} \label{app:orbits}
-We plot the three orbits considered throughout the work
-(Table~\ref{tab:orbits}) in Figure~\ref{fig:plot_orbits}.
-
-\begin{figure*}[htb!]
-\begin{center}
-\includegraphics[width=\textwidth]{fig/orbits.pdf}
-\end{center}
-\caption{The three orbits presented in Table~\ref{tab:orbits} and considered
-throughout the work. We plot the thin-disk, thick-disk, and halo orbits in the
-left, center, and right columns, respectively. The upper row shows a plot of
-$x$ vs. $y$ while the lower row shows $R$ vs. $z$.}
-\label{fig:plot_orbits}
-\end{figure*}
-
 \section{$\Delta R$-$\Delta x$ Relation} \label{app:deltax}
@@
 R_{\text{err}}^2 &= R^2 - 2 R \cos{\phi} \Delta x \\
-\implies \Delta R &\equiv \abs{R_{\text{err}} - R} 
-        = \abs{\cos{\phi}} \Delta x\text{.}
+\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\phi}} \Delta x\text{.}
 \end{split}
@@
 
-\section{$J_R$ and $J_{\phi}$ Observed Distributions} \label{app:hist}
-
-In Figure~\ref{fig:JR_zerr_hist} we plot the observed $J_R$ distribution
-induced by an offset in $z$, and in Figures~\ref{fig:JR_xerr_hist} and
-\ref{fig:Jphi_xerr_hist} we plot the observed $J_R$ and $J_{\phi}$
-distributions induced by an offset in $x$. We plot the distributions for the
-thick-disk (top panels) and thin-disk (bottom panels) orbits.
-
-We find that the observed $J_R$ distribution induced by an offset in $x$ more
-closely resembles a Gaussian, while the observed $J_R$ distribution induced by
-an offset in $z$ and the observed $J_{\phi}$ distribution induced by an offset
-in $x$ are both similar to the observed $J_z$ distribution induced by an
-offset in $z$ (see Figure~\ref{fig:Jz_hist}).
-
-\begin{figure}
+\section{Orbits} \label{app:orbits}
+We plot the three orbits considered throughout the work
+(Table~\ref{tab:orbits}) in Figure~\ref{fig:plot_orbits}.
+
+\begin{figure*}
 \begin{center}
-\includegraphics[width=3.35224200913in]{fig/schmactions_JR_zerr_hist.pdf}
+\includegraphics[width=\textwidth]{fig/orbits.pdf}
 \end{center}
-\caption{A histogram of the observed values in $J_R$ for the thick-disk (top
-panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
-$100\,\pc$. The induced error distribution more closely resembles a Gaussian
-centered on the null value.}
-\label{fig:JR_zerr_hist}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.35224200913in]{fig/schmactions_JR_xerr_hist.pdf}
-\end{center}
-\caption{A histogram of the observed values in $J_R$ for the thick-disk (top
-panel) and thin-disk (bottom panel) orbits assuming a $x$ offset of
-$100\,\pc$. The induced error distribution is non-Gaussian and bimodal, with
-neither of the modes centered on null.}
-\label{fig:JR_xerr_hist}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.35224200913in]{fig/schmactions_Jphi_xerr_hist.pdf}
-\end{center}
-\caption{A histogram of the observed values in $J_{\phi}$ for the thick-disk
-(top panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
-$100\,\pc$. The induced error distribution is non-Gaussian and bimodal, with
-neither of the modes centered on null.}
-\label{fig:Jphi_xerr_hist}
-\end{figure}
+\caption{The three orbits presented in Table~\ref{tab:orbits} and considered
+throughout the work. We plot the thin, thick, and halo orbits in the {\em
+left}, {\em center}, and {\em right} columns, respectively. The {\em upper}
+row shows a plot of $x$ vs. $y$ while the {\em lower} row shows $R$ vs. $z$.}
+\label{fig:plot_orbits}
+\end{figure*}
 
@@
 We consider the variations of the LSR as a function of azimuth at the fiducial
-solar circle ($R_{0} = 8.2\,\kpc$) in Figure~\ref{fig:lsr_variations}. At
-each azimuth, $\phi$, we take the median velocity in cylindrical coordinates
-of all stars within $200\,\pc$ of the position, following
-\citet{2018arXiv180610564S}. No best-fit subtraction was performed as in
-Figure~\ref{fig:midplane}.
-
-\begin{figure*}[htb!]
+solar circle ($R_{\odot} = 8.2\,\kpc$). At each azimuth, $\phi$, we take the
+median velocity in cylindrical coordinates of all stars within $200\,\pc$ of
+the position, following \citet{2018arXiv180610564S}. No best-fit subtraction
+was performed as in Figure~\ref{fig:midplane}.
+
+\begin{figure*}
 \begin{center}
@@
 \end{center}
-\caption{The LSR as a function of azimuth at the
-fiducial Solar circle ($R_0 = 8.2\,\kpc$). No best-fit subtraction is
+\caption{The local standard of rest (LSR) as a function of azimuth at the
+fiducial Solar circle ($R_{\odot} = 8.2\,\kpc$). No best-fit subtraction is
 performed here as we did in the case of the midplane
-(Section~\ref{ssec:local_midplane}). Variations in $v_z$ are on the order of
-$\sim5\textup{--}10\,\kms$.}
-\label{fig:lsr_variations}
+(Section~\ref{ssec:local_midplane}).}
 \end{figure*}

### 2019-05-09 paper/ms.tex
minor wording changes, plus additions to the intro of s4 that will hopefully help put context of the different sims up front and satisfy some of Chervin's comments

--- 2019-05-08 paper/ms.tex
@@
 ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly
-indicating local substructure infall \citep{2018MNRAS.481.1501B,
-2018arXiv180800451L}, and a gap
+indicating either local substructure infall \citep{2018MNRAS.481.1501B,
+2018arXiv180800451L} or bar buckling \citep{2018arXiv181109205K}, and a gap
 suggestive of perturbation by a dark matter subtructure in the tidal stream
@@
 \citep[e.g.][]{1915MNRAS..76...70J, 1985ApJ...295..388V, 2017ApJ...839...61T},
-or for associating stars with similar dynamical properties, e.g., to
-potentially determine membership in moving groups.
+or for associating stars with similar dynamical properties (e.g. to determine
+membership in moving groups).
 
@@
 the system's symmetry. However, the Galactic potential is not strictly
-axisymmetric. This has been known for some time \citep[e.g.][]{1957AJ.....62...93K,2009MNRAS.396L..56M, 2012ApJ...750L..41W}, but with the vast improvement in the quality of phase-space
-measurements due to \textit{Gaia} this assumption is increasingly
-inadequate \citep[e.g.][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. Even if this assumption were
-close enough for many purposes, the parameters used in axisymmetric models of the Galactic potential may
+axisymmetric; with the vast improvement in the quality of phase-space
+measurements due to \textit{Gaia}, this assumption appears increasingly
+inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this assumption were
+close enough for many purposes, the parameters used in axisymmetric models may
 be inadequately constrained by current observations.
@@
 actions, especially for stars in the thin disk. Our main point is that the
-``local'' and ``global'' Galactic
-midplanes are not the same. For stars relatively far from the Sun, their local midplane
+``local'' Galactic midplane is not the same as the ``global'' Galactic
+midplane. For stars relatively far from the Sun, their local midplane location
 will not agree with our local midplane extrapolated onto their position due to
@@
 distribution. Converting positions and velocities of more distant stars from a
-heliocentric to a Galactocentric coordinate system thus introduces a systematic
-bias in the $z$ coordinate, and consequently the actions computed for these stars.
+heliocentric to a Galactocentric coordinate system introduces a systematic
+bias in the $z$ coordinate, and hence the actions computed for these stars.
 The further from the solar neighborhood the target star, the more likely the
@@
 star at two different orbital phases (teal and orange points). If the observer
-were to then integrate the star's orbit using a model potential with the
+were then to integrate the star's orbit using a model potential with the
 erroneous midplane, they would obtain the teal and orange curves for the
 star's orbit, respectively. The actions estimated from these two  erroneous
-measurements would subsequently differ, both from each other and from the true measurement (in the potential with the correct coordinate system).
+orbits would subsequently differ, both from each other and from the actions
+estimated for the true orbit (in the potential with the correct midplane).
 Hence an incorrect midplane in the potential model assumed will induce
@@
 We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
-inaccurate or erroneous determination of the midplane leads to a phase-dependence of the actions
+inaccurate or erroneous determination of the midplane (and hence the
+zero-point of the $z$ coordinate) leads to a phase-dependence of the actions
 calculated from an orbit specified by a phase-space point and an assumed
@@
 correct midplane) should be phase-independent.
+This example uses an offset in $z$, but one would expect an analogous effect from an offset in another coordinate, such as the distance to the Galactic center or the local standard of rest.
 
@@
 
-Before turning to numerical methods, we first write down analytically the
-systematic error in the actions induced from either a midplane offset or a
-Galactocentric radius offset. We use the epicyclic approximation, which
+Before turning to numerical methods, we derive analytic expressions for the
+systematic error in the actions induced from offsets in the position (e.g. the midplane or
+Galactic center distance) or velocity (e.g. the local standard of rest) of the zero-point in the assumed coordinate system. We use the epicyclic approximation, which
 assumes that the motion in the $z$ and $R$ components of the orbit are
@@
 of thin disk-like orbits and a good description of thick disk-like orbits in
-an axisymmetric potential, i.e., ignoring the influence of the Galactic bar and
-spiral arms. We also make the assumption of a perfectly flat circular
+an axisymmetric potential (i.e. ignoring the influence of the Galactic bar and
+spiral arms). We also make the assumption of a perfectly flat circular
 velocity curve with velocity $v_c$, a good approximation near the Solar circle
@@
 where $\kappa$ and $\nu$ are the radial and vertical frequencies, $\Omega_c
-\equiv v_c/R_g$ is the rotation frequency, $A_z$ and $A_R$ are the amplitudes of
+\equiv v_c/R_g$ is the orbital frequency of the guiding center, $A_z$ and $A_R$ are the amplitudes of
 the stellar motion in the vertical and radial coordinates, and $\alpha$ and
-$\beta$ are arbitrary constants. Similarly, the velocities of the orbit are
+$\beta$ are arbitrary phases. Similarly, the velocities of the orbit are
 given by:
@@
 \end{center}
-\caption{A histogram of the observed values in $J_z$ for the thick-disk (top
+\caption{A histogram of the computed values of $J_z$ for the thick-disk (top
 panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
-$100\,\pc$. One can see that if the observed $z$ values have a bias (from e.g.
+$100\,\pc$, at different orbital phases. One can see that if the observed $z$ values have a bias (from e.g.
 an incorrectly computed midplane), then the induced error distribution in
@@
 distance from the true $J_z$ value to one of the modes of the distribution of
-observed $J_z$ values. Similar plots for $J_R$ induced by a $z$ offset and
+computed $J_z$ values. Similar plots for $J_R$ induced by a $z$ offset and
 $J_R$ and $J_{\phi}$ induced by an $x$ offset are given in
@@
 
-We quantify how a systematic error in the Galactocentric coordinate system
-induces phase-dependence in the \emph{observed} actions. We will consider
+We now quantify how a systematic error in the Galactocentric coordinate system
+induces phase-dependence in the actions calculated from the observed position and velocity of a star. We consider
 three orbits in the model potential described in
-Section~\ref{ssec:action_comp} that are typical of stars in the thin
-disk, thick disk, and halo. We summarize their initial
+Section~\ref{ssec:action_comp} that are typical of stars in the stellar thin
+disk, stellar thick disk, and the stellar halo. We summarize their initial
 positions in phase space and the actions computed by integrating their orbits
@@
 
-We begin with the thick-disk orbit. Consider an immortal observer who can
+We begin with the thick-disk orbit. Consider an observer who can
 measure the thick-disk star's phase-space position at many different times
-(and hence different orbital phases). However, the immortal observer uses a
+(and hence different orbital phases), but does so using a
 coordinate system in which the midplane is systematically offset in height by
-$100\,\pc$ from its actual location --- i.e. we subtract the vector $(0, 0,
-100)\,\pc$ from each position in the orbit. This corresponds to the immortal
+$100\,\pc$ from its actual location. To model this we subtract the vector $(0, 0,
+100)\,\pc$ from each position in the orbit. This corresponds to an
 observer physically located at e.g. the position $(8, 0, 0)\,\kpc$ in the
@@
 systematically offset coordinate system. Essentially we are shifting the
-integrated orbit and then reintegrating at each point along the shifted orbit,
-where the shifted orbit is the observed orbit. The erroneous computed actions
+integrated orbit and then reintegrating at each point along the shifted orbit. The actions computed using the offset coordinate system
 for each phase-space starting point are shown for the first $\Gyr$ of the
@@
 star's phase-space position is observed. To the observer this would
-appear as time-dependence in the actions, oscillating around
-their value using the correct coordinate system. This time dependence comes even though
+appear as time-dependence in the actions, which appear to oscillate around
+their value using the correct coordinate system even though in this example
 the observer is using the correctly constructed best-fit static, axisymmetric
@@
 the direction of the systematic offset as well as the true values of the
-actions (i.e. the type of orbit). In reality, we will have one measurement of
+actions (i.e. the type of orbit). In reality we will have one measurement of
 the phase-space position to work with, in which case the determination of the
@@
 that a $100\,\pc$ error in an orbit with $z_{\text{max}}=850\,\pc$ --- a
-$12\%$ error --- induced an $86\%$ error in the computation of $J_z$.
+$12\%$ error---induced an $86\%$ error in the computation of $J_z$.
 
@@
 $47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively.
-These are fractional errors of $21\%$, $3.1\%$ and $3.1\%$, respectively ---
-despite only a $1.2\%$ error in the distance to the Galactic center.
-
-In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
-$J_z$ for the thick-disk orbit (top panel) and thin-disk orbit (bottom panel)
+These are fractional errors of $21\%$, $3.1\%$ and $3.1\%$ in these actions, respectively, despite only a $1.2\%$ error in the distance to the Galactic center.
+
+In Figure~\ref{fig:Jz_hist} we plot a histogram of the values of
+$J_z$ computed for the thick-disk orbit (top panel) and thin-disk orbit (bottom panel)
 assuming a $z$ offset of $100\,\pc$ (i.e. as in the upper right panel of
-Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted as a
-vertical dashed line. Here we see that the errors in $J_z$ induced by a
-systematic offset in $z$ are non-Gaussian and bimodal. Furthermore, neither of
-the modes lie at zero. In other words, neither of the modes of the
-distribution of observed $J_z$ values are centered on the true $J_z$ (i.e. the
-$J_z$ computed with no offset in $z$). In the case of the thin-disk orbit
-(bottom panel), we see that, in addition to the prior complications, the error
-distribution is not even centered on null. This comes about when the midplane
+Figure~\ref{fig:one_orbit_wrong_ref}), as a function of orbital phase. The true value of $J_z$ is plotted as a
+vertical dashed line. The systematic error in $J_z$ induced by a
+systematic offset in $z$ is non-Gaussian and bimodal; neither of the modes is centered on the true $J_z$ (i.e. the $J_z$ computed with no offset in $z$). In the case of the thin-disk orbit
+(bottom panel), we see that, in addition to the prior complications, the
+distribution is not even centered on the true value. This comes about when the midplane
 error is $\gtrsim z_{\text{max}}$, where $z_{\text{max}}$ is the maximum
@@
 In Appendix~\ref{app:hist} we plot the same histogram as in
-Figure~\ref{fig:Jz_hist}, but for the observed distributions of $J_R$ induced
+Figure~\ref{fig:Jz_hist}, but for the distributions of $J_R$ induced
 by a $z$ offset (upper left panel of Figure~\ref{fig:many_orbit_wrong_ref})
-and the observed distributions of $J_R$ and $J_{\phi}$ errors induced by an
+and the distributions of $J_R$ and $J_{\phi}$ errors induced by an
 $x$ offset (lower left and lower center panels of
@@
 amplitude), where $v_z \sim 0$ and where the star is on most of it orbit. This
-is why the distribution is peaked on these values. For the thin-disk orbit
+is why the distribution, which is calculated at evenly spaced time intervals, peaks at these values. For the thin-disk orbit
 (bottom panel), the offset in $z$ is comparable to $z_{\text{max}}$. Now,
 there will be some points in the orbit where $v_z = 0$ and $z=0$ (in the
-observed, erroneous coordinate system). At these points, the computed $J_z$
+erroneous coordinate system). At these points, the computed $J_z$
 will vanish. The asymmetry and systematic offset then comes about because of
@@
 
-Empirically describing the error distribution shown in
-Figure~\ref{fig:Jz_hist} is difficult. Of course, Gaussian summary statistics
-are not adequate. We therefore elect to measure this error by computing one
+Gaussian summary statistics are clearly insufficient to describe the distribution shown in
+Figure~\ref{fig:Jz_hist}. We therefore elect to measure this error by computing one
 half the 95\uth{} percentile minus the 5\uth{} percentile of the distribution
-of observed action values. We plot this quantity, which we refer to as $\Delta J_i$ for each action, in Figure~\ref{fig:Jz_hist} as a
+of action values. We refer to this quantity as $\Delta J_i$ for each
+action value. We plot this quantity in Figure~\ref{fig:Jz_hist} as a
 horizontal arrow anchored on the true action value. Because of the bimodality
@@
 respectively). As before, the error ($\Delta J_z$) is one half the 95\uth{}
-minus 5\uth{} percentile of the distribution of observed $J_z$ values over the
+minus 5\uth{} percentile of the distribution of $J_z$ values over the
 course of the orbit. There are large errors for thin-disk like orbits ($J_z
@@
 
-\section{Coordinate System Offsets II. Azimuthal Midplane Variations in
-Simulations}
+\section{Coordinate System Offsets from Azimuthal Midplane Variations}
 \label{sec:local_fire}
 
-We now consider a second source of systematic offsets in the Galactocentric
-coordinate system: intrinsic variations of the midplane across the disk. In
-this work we specifically consider azimuthal variations in the midplane as
-defined by the stellar mass density. We use two sets of simulations of Milky
-Way-mass galaxies to estimate the size of this effect.
+Aside from measurement uncertainties, intrinsic variations of the midplane position across the disk of the galaxy can contribute to systematic errors in the actions computed assuming an axisymmetric potential model.
+The midplane position can vary as a function of location due to small local variations in the density of stars and gas in the Galaxy.
+Hints of this variation as a function of Galactocentric radius have been noted through their impact on the stellar velocity distribution by \citet{2012ApJ...750L..41W},\citet{2013ApJ...777L...5C}, and \citet{2013MNRAS.436..101W} pre-Gaia and recently by \citet{2019arXiv190209569F} post-Gaia.
+As pointed out by e.g.
+\citet{2014ApJ...797...53G} and \citet{2019ApJ...871..145A}, among many others, the gas distribution in the Galaxy also shows significant density variation across the disk.
+In this work, to stay close to the assumptions of our analytic calculations in \S \ref{ssec:epi_action}, we specifically consider \emph{azimuthal} variations in the midplane at the Solar circle, as defined by the stellar mass density.
+Since to our knowledge there are no direct empirical measurements of these variations for the Milky Way, we use two sets of simulations to estimate the size of this effect.
+One set are three zoomed, cosmological-hydrodynamical simulations of Milky Way-mass galaxies from the FIRE collaboration, described briefly in \S \ref{ssec:cosmozoom}, which include stars, gas, and dark matter in a fully cosmological setting but are not tailored to specific properties of the MW (such as the scale height or scale length, or the details of the accretion history).
+We use these simulations to span the range of possibilities for azimuthal midplane variations.
+The other simulations are isolated N-body simulations of interactions between the MW and a Sagittarius-like dwarf galaxy, described briefly in \S \ref{ssec:sag_sim}; these include dark matter and stars and are tailored to existing measurements of the structure of the MW's disk and of the orbit and properties of the Sagittarius (Sgr) dwarf galaxy.
+Comparing the azimuthal midplane variations in these simulations before and after the interaction with Sgr gives an idea of the effect of one minor merger whose properties are relatively well measured.
 
@@
 
-\section{$J_R$ and $J_{\phi}$ Observed Distributions} \label{app:hist}
-
-In Figure~\ref{fig:JR_zerr_hist} we plot the observed $J_R$ distribution
-induced by an offset in $z$, and in Figures~\ref{fig:JR_xerr_hist} and
-\ref{fig:Jphi_xerr_hist} we plot the observed $J_R$ and $J_{\phi}$
+\section{$J_R$ and $J_{\phi}$ Distributions} \label{app:hist}
+
+In Figure~\ref{fig:JR_zerr_hist} we plot the distribution of $J_R$ as a function of orbital phase induced by an offset in $z$, and in Figures~\ref{fig:JR_xerr_hist} and
+\ref{fig:Jphi_xerr_hist} we plot the $J_R$ and $J_{\phi}$
 distributions induced by an offset in $x$. We plot the distributions for the
@@
 
-We find that the observed $J_R$ distribution induced by an offset in $x$ more
-closely resembles a Gaussian, while the observed $J_R$ distribution induced by
-an offset in $z$ and the observed $J_{\phi}$ distribution induced by an offset
-in $x$ are both similar to the observed $J_z$ distribution induced by an
+We find that the $J_R$ distribution induced by an offset in $x$ more
+closely resembles a Gaussian, while the $J_R$ distribution induced by
+an offset in $z$ and the $J_{\phi}$ distribution induced by an offset
+in $x$ are both similar to the $J_z$ distribution induced by an
 offset in $z$ (see Figure~\ref{fig:Jz_hist}).
@@
 \end{center}
-\caption{A histogram of the observed values in $J_R$ for the thick-disk (top
+\caption{A histogram of the values computed for $J_R$ for the thick-disk (top
 panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
@@
 \end{center}
-\caption{A histogram of the observed values in $J_R$ for the thick-disk (top
+\caption{A histogram of the values computed for $J_R$ for the thick-disk (top
 panel) and thin-disk (bottom panel) orbits assuming a $x$ offset of
@@
 \end{center}
-\caption{A histogram of the observed values in $J_{\phi}$ for the thick-disk
+\caption{A histogram of the values computed for $J_{\phi}$ for the thick-disk
 (top panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of

### 2019-05-09 paper/ms.tex
moved discussion of measurement uncertainties to discussion, and added some to conclusions list

@@
 
-\section{Coordinate System Offsets I. Measurement Errors} \label{sec:mes_err}
-We consider the first potential source of coordinate system offsets which
-would generate the action errors described in Section~\ref{sec:ref_frame}:
-uncertainties in measurements of the position and velocity of the Sun relative
-to the Galactic center. Converting from heliocentric to Galactocentric
-coordinates relies on these measurements. Therefore, errors in their values
-will induce a systematic offset in the Galactocentric phase-space position of
-any observed star.
-
-\subsection{Galactic Center}
-First, one must define the center of the Galaxy. This is usually taken to be
-the location of the central supermassive black hole, Sagittarius~A\textsuperscript{*}
-\citep[\sgra{}, e.g.][]{2004ApJ...616..872R}. From stellar motions near
-\sgra{}, the distance from the Sun to \sgra{}, referred to as $R_0$, can be
-precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a
-recent measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
-\citep{2019arXiv190405721A}.
-
-\subsection{Galactic Orientation}
-Second, one must define the angular orientation of the Galaxy. This was
-defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
-defining the coordinates of the Galactic center in B1950 coordinates as
-(17:42:26.6, -28:55:00) and the North Galactic pole as (12:49:00, +27:24:00).
-These two quantities, together with $R_0$, define the orientation of the
-Galactic plane. However, there is growing evidence that the stellar midplane
-is tilted relative to this coordinate system \citep{2014ApJ...797...53G,
-2016ARAA..54..529B}, though not the H~\textsc{ii} midplane
-\citep{2019ApJ...871..145A}.
-
-\subsection{Solar Height}
-Third, one must define the Sun's vertical distance from the Galactic midplane,
-which can be determined by identifying where the stellar density and
-velocities reach a maximum (effectively the median height of all disk stars).
-The Solar height is usually taken to be $\sim 25\,\pc$
-\citep{2001ApJ...553..184C}, with a more recent measurement from \textit{Gaia}
-DR2 placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another
-strategy is to use the cold gas or H~\textsc{ii} regions in the disk to define
-the Galactic midplane, leading to slightly different values (by $\sim 5\,\pc$)
-for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em
-Gaia} review of these measurements is given by \citet{2016ARAA..54..529B}.
-
-\subsection{Local Standard of Rest}
-Finally, one must define the LSR, or mean velocity of stars near the Sun
-relative to the Galactic center (which is defined to have zero velocity), and
-the velocity of the Sun relative to the LSR. The radial ($U_{\odot}$) and
-vertical ($W_{\odot}$) components are computed by taking the mean motions of
-different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal
-component ($V_{\odot}$) is more difficult to measure, but can be modeled using
-the asymmetric drift relation \citep{2008gady.book.....B}. The values of the
-components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}. The
-value of the circular velocity is taken to be $\sim 220\kms$
-\citep[e.g.][]{2012ApJ...759..131B}.
-
-\vspace{6pt}
-
-Considerable effort has been placed on each of these measurements, but
-uncertainties remain, and detailed dynamical modeling across the disk ---
-particularly for dynamically cold stars --- may have to take them into
-account. For instance, the discrepancy of $\sim 5\,\pc$ between the midplane
-defined by stars and by gas will induce a $\sim10\%$ error in $J_z$ for an
-orbit with $z_{\text{max}}\sim100\,\pc$ (see Section~\ref{sec:ref_frame}).
-
 \section{Coordinate System Offsets from Azimuthal Midplane Variations}
@@
 
-Aside from measurement uncertainties, intrinsic variations of the midplane position across the disk of the galaxy can contribute to systematic errors in the actions computed assuming an axisymmetric potential model.
+Intrinsic variations of the midplane position across the disk of the galaxy can contribute to systematic errors in the actions computed assuming an axisymmetric potential model.
 The midplane position can vary as a function of location due to small local variations in the density of stars and gas in the Galaxy.
@@
 nearby companion like the Large Magellanic Cloud. One of the three we consider
-(\mf) does have an ongoing interaction with a satellite galaxy similar to
+(\mf) does have an ongoing interaction with a satellite galaxy similar in mass to
 Sagittarius, which has punched through the galactic disk outside the solar
-circle and induced some warping.
+circle, leaving behind some of its stars and inducing warping in the disk.
 
@@
 \section{Discussion} \label{sec:discussion} 
-We have used high-resolution cosmological simulations to illustrate that we
+We have used high-resolution simulations to illustrate that we
 expect the ``local midplane'' defined by stellar density to vary with azimuth
@@
 Galactic disk at small scales.
-
-While this is not in itself surprising or new, we also demonstrated that
-extending the coordinate system established by our local midplane to a global
-axisymmetric coordinate system spanning the entirety of the Galactic disk
+While this is not in itself surprising or new, we also demonstrated that extending the coordinate system established by our local midplane to a global axisymmetric coordinate system spanning the entirety of the Galactic disk
 introduces systematic error in computations of the actions under this symmetry
@@
 
-\subsection{Estimate of Milky Way Midplane Offsets from Current Data}
+\subsection{Estimates of Milky Way Midplane Offsets from Current Data}
 \label{ssec:mw_data_midplane}
@@
 
-\subsection{Coordinate System Offsets}\label{ssec:coord_off}
-In Section~\ref{sec:mes_err} we discussed the current observational efforts to
-measure the parameters of the Galactocentric coordinate system. We noted that
-the solar height is discrepant between gas-based and stellar-based
-determinations by $\sim5\,\pc$. While the effect is small and thus only likely
-to dominate the intrinsic midplane variations on small scales, it may be
-important for detailed modeling of kinematically cold stars where a $5\,\pc$
-offset is relevant.
-
-We also noted that the distance from the Sun to \sgra{} has been recently
-measured to a very low uncertainty of $\sim0.3\%$ or $\pm25\,\pc$
-\citep{2019arXiv190405721A}. However, the location of \sgra{} may not be
-equivalent to the location of the ``dynamical Galactic center,'' i.e., the
+\subsection{Uncertainties in the Solar position and velocity}\label{ssec:coord_off}
+
+Uncertainties in measurements of the position and velocity of the Sun relative
+to the Galactic center can also contribute to systematic error in the actions, since converting from heliocentric to Galactocentric
+coordinates relies on these measurements. Therefore, errors in their values
+will induce a systematic offset in the Galactocentric phase-space position of
+any observed star.
+Considerable effort has been placed on each of these measurements, but uncertainties remain, and detailed dynamical modeling across the disk---particularly for dynamically cold stars---may have to take them into
+account. 
+Here we briefly review the current measurements and their uncertainties, and the implications for the calculation of actions.
+
+\subsubsection{Galactic Center position}
+First, one must define the center of the Galaxy. This is usually taken to be
+the location of the central supermassive black hole, Sagittarius~A\textsuperscript{*}
+\citep[\sgra{}, e.g.][]{2004ApJ...616..872R}. From stellar motions near
+\sgra{}, the distance from the Sun to \sgra{}, referred to as $R_0$, can be
+precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a
+recent measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
+\citep{2019arXiv190405721A}, or $\sim0.3\%$. However, the location of \sgra{} may not be equivalent to the location of the ``dynamical Galactic center,'' i.e., the
 point in three-dimensional space about which the stars in the solar
@@
 
+\subsubsection{Galactic Orientation}
+Second, one must define the angular orientation of the Galaxy. This was
+defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
+defining the coordinates of the Galactic center in B1950 coordinates as
+(17:42:26.6, -28:55:00) and the North Galactic pole as (12:49:00, +27:24:00).
+These two quantities, together with $R_0$, define the orientation of the
+Galactic plane. However, there is growing evidence that the stellar midplane
+is tilted relative to this coordinate system \citep{2014ApJ...797...53G,
+2016ARAA..54..529B}, though not the H~\textsc{ii} midplane
+\citep{2019ApJ...871..145A}.  This tilt will contribute a linear dependence in Galactocentric radius to the systematic errors in actions induced by midplane offsets.
+
+\subsubsection{Solar Height}
+Third, one must define the Sun's vertical distance from the Galactic midplane,
+which can be determined by identifying where the stellar density and
+velocities reach a maximum (effectively the median height of all disk stars).
+The Solar height is usually taken to be $\sim 25\,\pc$
+\citep{2001ApJ...553..184C}, with a more recent measurement from \textit{Gaia}
+DR2 placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another
+strategy is to use the cold gas or H~\textsc{ii} regions in the disk to define
+the Galactic midplane, leading to slightly different values (by $\sim 5\,\pc$)
+for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em
+Gaia} review of these measurements is given by \citet{2016ARAA..54..529B}.
+The discrepancy between gas-based and stellar-based determinations of the solar height is small, and thus only likely to dominate over intrinsic midplane variations on small scales, but will be relevant for detailed modeling of young and kinematically cold stars.
+For instance, it will induce a $\sim10\%$ error in $J_z$ for an
+orbit with $z_{\text{max}}\sim100\,\pc$ (see Section~\ref{sec:ref_frame}).
+
+\subsubsection{Local Standard of Rest}
+Finally, one must define the LSR, or mean velocity of stars near the Sun
+relative to the Galactic center (which is defined to have zero velocity), and
+the velocity of the Sun relative to the LSR. The radial ($U_{\odot}$) and
+vertical ($W_{\odot}$) components are computed by taking the mean motions of
+different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal
+component ($V_{\odot}$) is more difficult to measure, but can be modeled using
+the asymmetric drift relation \citep{2008gady.book.....B}. The values of the
+components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}.
+Their uncertainties should also lead to systematic errors in the actions, as given in Eqs.
+\ref{eq:induced_Jphi}--\ref{eq:induced_Jz}.
+For example, the
+value of the circular velocity is taken to be $\sim 220\kms$
+\citep[e.g.][]{2012ApJ...759..131B} with roughly 10\% uncertainty.
+Based on the analytic estimates in \S \ref{ssec:epi_action}, we expect this to translate to at least a 10\% systematic error in $J_{\phi}$.
+
+\vspace{6pt}
+
 \section{Conclusions}\label{sec:conclusion}
@@
 midplane along the Solar circle in three different high-resolution, zoom-in
-simulations of Milky Way mass galaxies from the FIRE collaboration. We found
+simulations of Milky Way mass galaxies from the FIRE collaboration, as well as a controlled simulation of the interaction of the MW with Sagittarius. We found
 that the midplane varies as a function of azimuth at the Solar circle by
-$\sim185, 162, 84\,\pc$ (middle $90\%$) for the three galaxies we considered
-(\mi{}, \mf{}, and \mm{}, respectively).
+$\sim185, 162, 84\,\pc$ (middle $90\%$) for the three cosmological simulations we considered
+(\mi{}, \mf{}, and \mm{}, respectively) and by $\sim30$ and $\sim100$ pc before and after the interaction in the controlled simulation.
+Comparison of the corresponding velocity variations with recent measurements in the MW suggests that these values are consistent with expectations for midplane variations in the MW itself, which have not yet been measured as a function of azimuth.
 
@@
 
-\item While inaccuracies in the parameters of the currently adopted
-Galactocentric coordinate system are likely to be unimportant for most
-applications, we discussed the importance of testing the assumption that the
-dynamical Galactic center is colocated with \sgra{}, which we mentioned how to
-do.
+\item Some inaccuracies in the parameters of the currently adopted
+Galactocentric coordinate system are likely important for some
+applications. In particular, it is imperative to test the assumption that the
+dynamical Galactic center is colocated with \sgra{}. We discuss how to
+do this in \S \ref{ssec:coord_off}.
 

### 2019-05-09 paper/ms.tex
small changes based on other Chervin comments

@@
 representations of the true Milky Way in other ways as well,\footnote{The
-failure of cosmological simulations to exactly reproduce the Milky Way is not
-entirely due to limitations of the numerical model. Candidate Milky Way-like
+failure of cosmological simulations to exactly reproduce the Milky Way is not necessarily due to limitations of the numerical model. Candidate Milky Way-like
 galaxies are chosen solely on their mass and isolation, for which there are a
@@
 components are realized with live distributions of N-body particles; the Milky
-Way and Sagittarius are each initialized to be in equilibrium in isolation
-\Gus{check!}.
+Way and Sagittarius are each initialized to be in equilibrium in isolation.
 
@@
 panel looks qualitatively similar to the panels from the FIRE simulations
-(Figure~\ref{fig:midplane}) is evidence that the midplane variations are
-produced in part by mergers.}
+(Figure~\ref{fig:midplane}) suggests that midplane variations are
+produced at least in part by mergers.}
 \label{fig:midplane_chervin}
@@
 that the midplane varies as a function of azimuth at the Solar circle by
-$\sim185, 162, 84\,\pc$ (middle $90\%$) for the three cosmological simulations we considered
-(\mi{}, \mf{}, and \mm{}, respectively) and by $\sim30$ and $\sim100$ pc before and after the interaction in the controlled simulation.
+$\sim$185, 162, 84 pc (middle $90\%$) for the three cosmological simulations we considered
+(\mi{}, \mf{}, and \mm{}, respectively) and by $\sim$30 and $\sim$100 pc before and after the interaction in the controlled simulation.
 Comparison of the corresponding velocity variations with recent measurements in the MW suggests that these values are consistent with expectations for midplane variations in the MW itself, which have not yet been measured as a function of azimuth.
@@
 
-\item Some inaccuracies in the parameters of the currently adopted
+\item Inaccuracies in the parameters of the currently adopted
 Galactocentric coordinate system are likely important for some
@@
 
-\item This work emphasizes the importance of combining chemistry and dynamics.
-Chemical tagging \citep{2002ARA&A..40..487F} and dynamical tagging must
-complement and confirm each other. The importance of combining chemistry and
-dynamics for Galactic archaeology was recently pointed out by
-\citet{2019arXiv190210719K}.
+\item This work underlines the importance of combining chemistry and dynamics.
+Since chemical tagging \citep{2002ARA&A..40..487F} is not subject to the same systematic errors discussed in this paper, it should be used to confirm dynamical associations and to offset the effect of these systematic errors on the action-space distribution function.
+%The importance of combining chemistry and dynamics for Galactic archaeology was recently pointed out by \citet{2019arXiv190210719K}.
 

### 2019-05-09 paper/ms.tex
small changes to acknowledgments and my name in the author list

@@
 
-\author[0000-0003-3939-3297]{Robyn Sanderson}
+\author[0000-0003-3939-3297]{Robyn E. Sanderson}
 \affil{\cca}
@@
 in part by the Roy \& Diana Vagelos Program in the Molecular Life Sciences and
-the Roy \& Diana Vagelos Challenge Award. K.V.J.'s and C.F.P.L.'s work was
+the Roy \& Diana Vagelos Challenge Award. R.E.S's, K.V.J.'s and C.F.P.L.'s work was
 performed in part during the Gaia19 workshop and the 2019 Santa Barbara Gaia
@@
 M.-M.M.L. was partly supported by NSF grant AST18-15461. The work of D.A.-A.,
-A.B., D.W.H., R.S., K.V.J., M.-M.M.L., and M.K.N. is supported by the Simons
+A.B., D.W.H., R.E.S., K.V.J., M.-M.M.L., and M.K.N. is supported by the Simons
 Foundation.

### 2019-05-10 paper/ms.tex

@@
 their value using the correct coordinate system even though in this example
+appear as time-dependence in the actions, oscillating around
+their value using the correct coordinate system. This time dependence comes even though
 the observer is using the correctly constructed best-fit static, axisymmetric
@@
 action value. We plot this quantity in Figure~\ref{fig:Jz_hist} as a
+of observed action values. We plot this quantity, which we refer to as $\Delta J_i$ for each action, in Figure~\ref{fig:Jz_hist} as a
 horizontal arrow anchored on the true action value. Because of the bimodality

### 2019-05-10 paper/ms.tex
Remove section that got merged

@@
 
-\section{Coordinate System Offsets II. Measurement Errors} \label{sec:mes_err}
-We consider the second potential source of coordinate system offsets which
-would generate the action errors described in Section~\ref{sec:ref_frame}:
-uncertainties in measurements of the position and velocity of the Sun relative
-to the Galactic center. Converting from heliocentric to Galactocentric
-coordinates relies on these measurements. Therefore, errors in their values
-will induce a systematic offset in the Galactocentric phase-space position of
-any observed star.
-
-\subsection{Galactic Center}
-First, one must define the center of the Galaxy. This is usually taken to be
-the location of the central supermassive black hole, Sagittarius~A\textsuperscript{*}
-\citep[\sgra{}, e.g.][]{2004ApJ...616..872R}. From stellar motions near
-\sgra{}, the distance from the Sun to \sgra{}, referred to as $R_0$, can be
-precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a
-recent measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
-\citep{2019arXiv190405721A}.
-
-In the epicyclic approximation (Section~\ref{ssec:epi_action}), the $0.3\%$
-error in $R_0$ induces a $5.4\%$ and $5.9\%$ error in the thin-disk and
-thick-disk orbits, respectively. For $J_{\phi}$, the error is $0.3\%$ for both
-orbits.
-
-\subsection{Galactic Orientation}
-Second, one must define the angular orientation of the Galaxy. This was
-defined in 1958 by the IAU subcomission 33b \citep{1960MNRAS.121..123B} by
-defining the coordinates of the Galactic center in B1950 coordinates as
-(17:42:26.6, -28:55:00) and the North Galactic pole as (12:49:00, +27:24:00).
-These two quantities, together with $R_0$, define the orientation of the
-Galactic plane. However, there is growing evidence that the stellar midplane
-is tilted relative to this coordinate system \citep{2014ApJ...797...53G,
-2016ARAA..54..529B}, though not the H~\textsc{ii} midplane
-\citep{2019ApJ...871..145A}.
-
-\subsection{Solar Height}
-Third, one must define the Sun's vertical distance from the Galactic midplane, $z_{\odot}$.
-It can be determined by identifying where the stellar density and
-velocities reach a maximum (effectively the median height of all disk stars).
-The Solar height is usually taken to be $\sim 25\,\pc$
-\citep{2001ApJ...553..184C}, with a more recent measurement from \textit{Gaia}
-DR2 placing it at $20.8 \pm 0.3\,\pc$ \citep{2019MNRAS.482.1417B}. Another
-strategy is to use the cold gas or H~\textsc{ii} regions in the disk to define
-the Galactic midplane, leading to slightly different values (by $\sim 5\,\pc$)
-for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em
-Gaia} review of these measurements is given by \citet{2016ARAA..54..529B}.
-
-Under the epicyclic approximation, the highest precision measurement of
-$z_{\odot}$ of $1.4\%$ from \citet{2019MNRAS.482.1417B} induces a $0.5\%$ and
-$0.04\%$ error in $J_z$ for the thick-disk and thin-disk orbits, respectively.
-The discrepancy between the stellar and gas midplanes of $\sim5\,\pc$ induces
-an $8.3\%$ and $1.2\%$ error in $J_z$ for the two orbits.
-
-\subsection{Local Standard of Rest}
-Finally, one must define the LSR, or mean velocity of stars near the Sun
-relative to the Galactic center (which is defined to have zero velocity), and
-the velocity of the Sun relative to the LSR. The radial ($U_{\odot}$) and
-vertical ($W_{\odot}$) components are computed by taking the mean motions of
-different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal
-component ($V_{\odot}$) is more difficult to measure, but can be modeled using
-the asymmetric drift relation \citep{2008gady.book.....B}. The values of the
-components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}. The
-value of the circular velocity is taken to be $\sim 220\kms$
-\citep[e.g.][]{2012ApJ...759..131B}.
-
-\vspace{6pt}
-
-Considerable effort has been placed on each of these measurements, but
-uncertainties remain, and detailed dynamical modeling across the disk ---
-particularly for dynamically cold stars --- may have to take them into
-account. For instance, the discrepancy of $\sim 5\,\pc$ between the midplane
-defined by stars and by gas will induce a $\sim10\%$ error in $J_z$ for an
-orbit with $z_{\text{max}}\sim100\,\pc$ (see Section~\ref{sec:ref_frame}).
-
 \section{Discussion} \label{sec:discussion} 

### 2019-05-10 paper/ms.tex
Reorganize intro to midplane section

@@
 
-Intrinsic variations of the midplane position across the disk of the galaxy can contribute to systematic errors in the actions computed assuming an axisymmetric potential model.
-The midplane position can vary as a function of location due to small local variations in the density of stars and gas in the Galaxy.
-Hints of this variation as a function of Galactocentric radius have been noted through their impact on the stellar velocity distribution by \citet{2012ApJ...750L..41W},\citet{2013ApJ...777L...5C}, and \citet{2013MNRAS.436..101W} pre-Gaia and recently by \citet{2019arXiv190209569F} post-Gaia.
-As pointed out by e.g.
-\citet{2014ApJ...797...53G} and \citet{2019ApJ...871..145A}, among many others, the gas distribution in the Galaxy also shows significant density variation across the disk.
-In this work, to stay close to the assumptions of our analytic calculations in \S \ref{ssec:epi_action}, we specifically consider \emph{azimuthal} variations in the midplane at the Solar circle, as defined by the stellar mass density.
-Since to our knowledge there are no direct empirical measurements of these variations for the Milky Way, we use two sets of simulations to estimate the size of this effect.
-One set are three zoomed, cosmological-hydrodynamical simulations of Milky Way-mass galaxies from the FIRE collaboration, described briefly in \S \ref{ssec:cosmozoom}, which include stars, gas, and dark matter in a fully cosmological setting but are not tailored to specific properties of the MW (such as the scale height or scale length, or the details of the accretion history).
+The stellar midplane of the Galaxy should vary as a function of azimuth and
+Galactocentric radius due to small, local variations in the stellar density.
+Hints of this variation as a function of Galactocentric radius have been noted
+through their impact on the stellar velocity distribution pre-\emph{Gaia} by
+\citet{2012ApJ...750L..41W}, \citet{2013ApJ...777L...5C}, and
+\citet{2013MNRAS.436..101W} and recently post-\emph{Gaia} by
+\citet{2019arXiv190209569F}. As pointed out by e.g.
+\citet{2014ApJ...797...53G} and \citet{2019ApJ...871..145A}, among many
+others, the gas distribution in the Galaxy also shows significant density
+variation across the disk.
+
+The local Galactocentric coordinate system is defined based on the location of
+the Sun relative to the midplane. Extending this coordinate system to a global
+one therefore introduces systematic errors in the $z$ components of stellar
+positions. As discussed in Section~\ref{sec:ref_frame}, this systematic error
+introduces errors in integrating orbits and computing actions.
+
+We specifically consider \emph{azimuthal} variations in the midplane at the Solar circle, as defined by the stellar mass density.
+Since, to our knowledge, there are no direct empirical measurements of these variations in the Milky Way, we use example simulations from two classes of simulations to estimate the size of this effect.
+
+One set are three zoom-in, cosmological-hydrodynamical simulations of Milky Way-mass galaxies from the FIRE collaboration, described briefly in \S \ref{ssec:cosmozoom}.
+These include stars, gas, and dark matter in a fully cosmological setting but are not tailored to specific properties of the Milky Way (such as the scale height or scale length, or the details of the accretion history).
 We use these simulations to span the range of possibilities for azimuthal midplane variations.
-The other simulations are isolated N-body simulations of interactions between the MW and a Sagittarius-like dwarf galaxy, described briefly in \S \ref{ssec:sag_sim}; these include dark matter and stars and are tailored to existing measurements of the structure of the MW's disk and of the orbit and properties of the Sagittarius (Sgr) dwarf galaxy.
-Comparing the azimuthal midplane variations in these simulations before and after the interaction with Sgr gives an idea of the effect of one minor merger whose properties are relatively well measured.
+
+The other set of simulations are isolated N-body simulations of interactions between the Milky Way and a Sagittarius-like dwarf galaxy, described briefly in \S \ref{ssec:sag_sim}.
+These include dark matter and stars and are tailored to existing measurements of the structure of the Milky Way's disk and of the orbit and properties of the Sagittarius dwarf galaxy.
+Comparing the azimuthal midplane variations in the host galaxy of these simulations before and after the interaction with the Sagittarious-like object gives an idea of the effect of one minor merger whose properties are relatively well measured.
 

### 2019-05-10 paper/ms.tex
Small changes throughout

@@
 \end{center}
-\caption{A histogram of the computed values of $J_z$ for the thick-disk (top
+\caption{A histogram of the computed values of $J_z$ at different orbital phases for the thick-disk (top
 panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
-$100\,\pc$, at different orbital phases. One can see that if the observed $z$ values have a bias (from e.g.
+$100\,\pc$. One can see that if the observed $z$ values have a bias (from e.g.
 an incorrectly computed midplane), then the induced error distribution in
@@
 In Figure~\ref{fig:Jz_hist} we plot a histogram of the values of
-$J_z$ computed for the thick-disk orbit (top panel) and thin-disk orbit (bottom panel)
-assuming a $z$ offset of $100\,\pc$ (i.e. as in the upper right panel of
-Figure~\ref{fig:one_orbit_wrong_ref}), as a function of orbital phase. The true value of $J_z$ is plotted as a
+$J_z$ computed at different orbital phases for the thick-disk orbit (top panel) and thin-disk orbit (bottom panel)
+assuming a $z$ offset of $100\,\pc$ (as in the upper right panel of
+Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted as a
 vertical dashed line. The systematic error in $J_z$ induced by a
@@
 different offsets in $z$. All orbits have the same initial position of $(8, 0,
-0)\,\kpc$ and velocity $(-190, 0, v_z)\,\kms$, where we vary $v_z$. We show
+0)\,\kpc$ and velocity $(0, -190, v_z)\,\kms$, where we vary $v_z$. We show
 this for a $z$ offset of $10, 50, 100\,\pc$ (orange, teal, and green
@@
 Figure~\ref{fig:dJz_fun_Jz}. For each orbit, we set the initial position to be
-$(8,0,0)\,\kpc$ and the initial velocity to be $(-190, 0, v_z)\,\kms$, where
+$(8,0,0)\,\kpc$ and the initial velocity to be $(0, -190, v_z)\,\kms$, where
 we vary $v_z$. For a thin disk-like orbit ($J_z\sim0.5\,\actunit$), even a
@@
 for inaccuracies in the galactocentric coordinate system. Error bars are
-calculated as in Figure~\ref{fig:midplane}. (\Gus{Check:}) The upper panels
+calculated as in Figure~\ref{fig:midplane}. The upper panels
 show the midplane as a function of azimuth before the first encounter near the
@@
 consistent with Figure~17 of \citet{2018MNRAS.481..286L}, which shows that an
-$m=2$ mode is excited (note that $m=0$ and $m=1$ modes are stronger, but these
+$m=2$ mode is excited ($m=0$ and $m=1$ modes are stronger, but these
 are removed in our sine-curve subtraction). The fact that the $t=6.9\,\Gyr$
@@
 
-\subsubsection{Galactic Center position}
+\subsubsection{Galactic Center Position}
 First, one must define the center of the Galaxy. This is usually taken to be
@@
 
-We did not consider in this text the effect of the \emph{angular position} of
+We did not consider in this work the effect of the \emph{angular position} of
 the dynamical Galactic center being offset with \sgra{}.
@@
 
-\vspace{6pt}
-
 \section{Conclusions}\label{sec:conclusion}
@@
 midplane along the Solar circle in three different high-resolution, zoom-in
-simulations of Milky Way mass galaxies from the FIRE collaboration, as well as a controlled simulation of the interaction of the MW with Sagittarius. We found
+simulations of Milky Way mass galaxies from the FIRE collaboration, as well as a controlled simulation of the interaction of the Milky Way with Sagittarius. We found
 that the midplane varies as a function of azimuth at the Solar circle by
@@
 (\mi{}, \mf{}, and \mm{}, respectively) and by $\sim$30 and $\sim$100 pc before and after the interaction in the controlled simulation.
-Comparison of the corresponding velocity variations with recent measurements in the MW suggests that these values are consistent with expectations for midplane variations in the MW itself, which have not yet been measured as a function of azimuth.
+Comparison of the corresponding velocity variations with recent measurements in the Milky Way suggests that these values are consistent with expectations for midplane variations in the Milky Way itself, which have not yet been measured as a function of azimuth.
 

### 2019-05-10 paper/ms.tex
Removed observed actions language

@@
 \end{center}
-\caption{The artificial phase-dependence in the observed actions induced by an
+\caption{The artificial phase-dependence in the computed actions with an
 error in the Galactocentric coordinate system. We consider here the thick-disk
@@
 distributions as in Figure~\ref{fig:Jz_hist}, with the exception that the
-observed $J_R$ distribution induced by an $x$ offset more closely resembles a
+computed $J_R$ distribution induced by an $x$ offset more closely resembles a
 Gaussian.
@@
 half the 95\uth{} percentile minus the 5\uth{} percentile of the distribution
-of action values. We refer to this quantity as $\Delta J_i$ for each
-action value. We plot this quantity in Figure~\ref{fig:Jz_hist} as a
-of observed action values. We plot this quantity, which we refer to as $\Delta J_i$ for each action, in Figure~\ref{fig:Jz_hist} as a
+of action values. We plot this quantity, which we refer to as $\Delta J_i$ for each action, in Figure~\ref{fig:Jz_hist} as a
 horizontal arrow anchored on the true action value. Because of the bimodality

### 2019-05-10 paper/ms.tex
Add citation to Buist paper

@@
 correct midplane) should be phase-independent.
-This example uses an offset in $z$, but one would expect an analogous effect from an offset in another coordinate, such as the distance to the Galactic center or the local standard of rest.
+
+This example uses an offset in $z$, but one would expect an analogous effect
+from an offset in another coordinate, such as the distance to the Galactic
+center or the local standard of rest. A similar effect in which actions gain
+time-dependence due to a time-varying potential was pointed out by
+\citet{2015A&A...584A.120B}.
 
@@
 that the midplane varies as a function of azimuth at the Solar circle by
-$\sim$185, 162, 84 pc (middle $90\%$) for the three cosmological simulations we considered
-(\mi{}, \mf{}, and \mm{}, respectively) and by $\sim$30 and $\sim$100 pc before and after the interaction in the controlled simulation.
+$\sim185$, $162$, $84\,\pc$ (middle $90\%$) for the three cosmological simulations we considered
+(\mi{}, \mf{}, and \mm{}, respectively) and by $\sim30$ and $\sim100\,\pc$ before and after the interaction in the controlled simulation.
 Comparison of the corresponding velocity variations with recent measurements in the Milky Way suggests that these values are consistent with expectations for midplane variations in the Milky Way itself, which have not yet been measured as a function of azimuth.
@@
 Since chemical tagging \citep{2002ARA&A..40..487F} is not subject to the same systematic errors discussed in this paper, it should be used to confirm dynamical associations and to offset the effect of these systematic errors on the action-space distribution function.
-%The importance of combining chemistry and dynamics for Galactic archaeology was recently pointed out by \citet{2019arXiv190210719K}.
 
@@
 in part by the Roy \& Diana Vagelos Program in the Molecular Life Sciences and
-the Roy \& Diana Vagelos Challenge Award. R.E.S's, K.V.J.'s and C.F.P.L.'s work was
+the Roy \& Diana Vagelos Challenge Award. K.V.J.'s, C.F.P.L.'s, and R.E.S.'s work was
 performed in part during the Gaia19 workshop and the 2019 Santa Barbara Gaia

### 2019-05-10 paper/ms.tex
Clarify softening lengths, reformat table

@@
 simulated galaxies considered in this work. For comparison, we also give the
-softening lengths for the simulated galaxies.\label{tab:scale_height}}
-\tablehead{\colhead{galaxy} & \colhead{cold\tablenotemark{a} gas disk} &
-\colhead{stellar thin disk} & \colhead{stellar thick disk} & \colhead{cold gas
-softening length} & \colhead{stellar softening length} \\
+softening lengths for the simulated galaxies, computed for cold gas ($T<1000\,\text{K}$) and stars with $\abs{R-R_0} < 0.5\,\kpc$ and $\abs{z} < 1\,\kpc$.
+We have assumed that $R_0=8.2\,\kpc$.\label{tab:scale_height}}
+\tablehead{\colhead{galaxy} & \colhead{\makecell{cold\tablenotemark{a} gas disk \\ scale height}} &
+\colhead{\makecell{thin disk \\ scale height}} & \colhead{\makecell{thick disk \\ scale height}} & \colhead{\makecell{cold\tablenotemark{a} gas \\
+softening length}} & \colhead{\makecell{stellar \\ softening length}} \\
 \colhead{} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)}

### 2019-05-10 short_info/median_gas_smoothing_length.py

@@ -4,18 +4,16 @@ import numpy as np
 Rsun = 8.2
 dR = 0.5
 dZ = 1.0
-maxT = 1E4
+maxT = 1E3
 
 def read_snap(gal):
     # takes in galaxy (string = m12i, m12f, m12m, etc.)
     # reads in and sets fiducial coordinates
     # returns snap
     gal_info = 'fiducial_coord/' + gal + '_res7100_center.txt'
-    sim_directory = '/mnt/ceph/users/firesims/fire2/metaldiff/'+gal+'_res7100'
-    snap = gizmo.io.Read.read_snapshots(['star', 'gas', 'dark'], 'index', 600,
-                                        properties=['position', 'id',
-                                                    'mass', 'velocity',
-                                                    'form.scalefactor',
+    sim_directory = '/Users/abeane/scratch/actions_systematic/data/fire2/metaldiff/'+gal+'_res7100'
+    snap = gizmo.io.Read.read_snapshots(['gas'], 'index', 600,
+                                        properties=['position',
                                                     'smooth.length', 'temperature'],
                                         assign_center=False,
                                         simulation_directory=sim_directory)

### 2019-05-10 paper/ms.tex
Add exact numbers for midplane variation of Sgr sims

@@
 $-1.00\,\text{rad}$ and $\text{C}$ being $2.6$, $-7.8$, $-65$, and $-53\,\pc$
-chronologically for each of the four timesteps shown.
+chronologically for each of the four timesteps shown. The middle $90\%$ of midplane values across the Solar circle spans $49$, $62$, $137$, and $119\,\pc$, for each timestep.
 
@@
 $\sim185$, $162$, $84\,\pc$ (middle $90\%$) for the three cosmological simulations we considered
-(\mi{}, \mf{}, and \mm{}, respectively) and by $\sim30$ and $\sim100\,\pc$ before and after the interaction in the controlled simulation.
+(\mi{}, \mf{}, and \mm{}, respectively) and by $\sim60$ and $\sim120\,\pc$ before and after the interaction in the controlled simulation.
 Comparison of the corresponding velocity variations with recent measurements in the Milky Way suggests that these values are consistent with expectations for midplane variations in the Milky Way itself, which have not yet been measured as a function of azimuth.

### 2019-05-10 paper/ms.tex
Add correspondence btw Fig 5 and orbits

@@
 $(8,0,0)\,\kpc$ and the initial velocity to be $(0, -190, v_z)\,\kms$, where
-we vary $v_z$. For a thin disk-like orbit ($J_z\sim0.5\,\actunit$), even a
+we vary $v_z$.\footnote{The thin-disk, thick-disk, and halo orbits correspond to setting $v_z=10$, $50$, and $190$, respectively (see Table~\ref{tab:orbits}).} For a thin disk-like orbit ($J_z\sim0.5\,\actunit$), even a
 $10\,\pc$ offset in $z$ is enough to induce a $\sim25\%$ error in $J_z$. For

### 2019-05-10 paper/ms.tex
Update coordinate system discussion

@@
 2016ARAA..54..529B}, though not the H~\textsc{ii} midplane
-\citep{2019ApJ...871..145A}.  This tilt will contribute a linear dependence in Galactocentric radius to the systematic errors in actions induced by midplane offsets.
+\citep{2019ApJ...871..145A}.
+
+This tilt will contribute a systematic offset in $z$, with the exact magnitude
+depending on the position of the observed star. For instance,
+\citet{2014ApJ...797...53G} quote a $\sim0.4\degree$ tilt at $3.1\,\kpc$,
+corresponding to a vertical height of $\sim22\,\pc$. This corresponds to a
+$37\%$ ($5\%$) error in $J_z$ for the thin-disk (thick-disk) orbit.
 
@@
 For instance, it will induce a $\sim10\%$ error in $J_z$ for an
-orbit with $z_{\text{max}}\sim100\,\pc$ (see Section~\ref{sec:ref_frame}).
+orbit with $z_{\text{max}}\sim100\,\pc$.
 
@@
 value of the circular velocity is taken to be $\sim 220\kms$
-\citep[e.g.][]{2012ApJ...759..131B} with roughly 10\% uncertainty.
-Based on the analytic estimates in \S \ref{ssec:epi_action}, we expect this to translate to at least a 10\% systematic error in $J_{\phi}$.
+\citep[e.g.][]{2012ApJ...759..131B} with roughly $10\%$ uncertainty. We expect this to translate to at least a $10\%$ systematic error in $J_{\phi}$.
 

### 2019-05-10 paper/ms.tex
Update major merger citations

@@
 result of \textit{Gaia} Data Release 2 (DR2).  Recent major discoveries
-include the remnants of a major merger \citep{2018ApJ...860L..11K,
-2018Natur.563...85H, 2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space
+include the affirmation of remnants of a major merger \citep{2018ApJ...860L..11K, 2018MNRAS.478..611B, 2018Natur.563...85H, 2019MNRAS.486..378L, 2019MNRAS.482.3426M} hinted at in pre-\textit{Gaia} work \citep[e.g.][]{2005MNRAS.359...93M, 2011MNRAS.412.1203N}, a phase-space
 ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly

### 2019-05-10 paper/ms.tex
Update end of intro

@@
 the local standard of rest (LSR). We review the observational efforts to
-measure these parameters in Section~\ref{sec:mes_err}.
+measure these parameters in Section~\ref{ssec:coord_off}.
 
@@
 In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
-system errors have on the measured actions. We then discuss two potential
-sources of these coordinate system offsets in
-Sections~\ref{sec:mes_err}~and~\ref{sec:local_fire}. In
-Section~\ref{sec:mes_err} we discuss the potential errors in measuring the
-parameters of the Galactocentric coordinate system. In
+system errors have on the measured actions. In
 Section~\ref{sec:local_fire} we examine the azimuthal variations of the

### 2019-05-10 paper/ms.tex
Add dust map, use 5-10 km/s

@@
 project. From observations that the mean vertical velocity of the Milky Way
-disk varies by $\sim5\,\kms$ with radius, we estimate that the true midplane
-variations are of order $\sim60\textup{--}90\,\pc$. Thus, current
+disk varies by $\sim5\textup{--}10\,\kms$ with radius, we estimate that the true midplane
+variations are of order $\sim60\textup{--}170\,\pc$. Similar offsets are seen in three-dimensional dust maps. Thus, current
 state-of-the-art action calculations --- which assume the global and local
@@
 0.09\,\Myr^{-1}$ and $\sim0.06\,\Myr^{-1}$, respectively. By dimensional
-analysis, and assuming a vertical velocity variation of $5\,\kms$, we
-therefore expect the midplane offsets to be $\sim 57\textup{--}85\,\pc$. We
-stress that this is a rough calculation, but it shows that already we see
-evidence in the data from velocities for midplane offsets on the order of what
-we saw in both sets of simulations.
+analysis, and assuming a vertical velocity variation of $5\textup{--}10\,\kms$, we
+therefore expect the midplane offsets to be $\sim 57\textup{--}170\,\pc$. We
+stress that this is a rough calculation.
+
+Three-dimensional dust maps also offer a view into the expected variation of
+the stellar disk, since dust should trace high-star formation regions.
+Figure~9 of \citet{2019MNRAS.483.4277C}, Figure~1 of
+\citet{2019arXiv190105971L}, and Figure~2 of \citet{2019arXiv190502734G} all
+show that the midplane varies by $\sim10\degree$ at a distance of
+$\sim0.75\,\kpc$, corresponding to a physical height of $\sim130\,\pc$.
+
+Already we see evidence in the data from velocities and dust maps for midplane
+offsets on the order of what we saw in both sets of simulations.
 
@@
 \item Assuming a vertical velocity variation of the Milky Way of
-$\sim5\,\kms$, consistent with recent results from \textit{Gaia} DR2
+$\sim5\textup{--}10\,\kms$, consistent with recent results from \textit{Gaia} DR2
 \citep{2018A&A...616A..11G, 2019arXiv190209569F} and our results from the FIRE
 simulations (Figure~\ref{fig:lsr_variations}), we estimated that the
-corresponding midplane offsets are $\sim60\textup{--}90\,\pc$ by dimensional
-analysis with the vertical frequencies of disk-like orbits.
+corresponding midplane offsets are $\sim60\textup{--}170\,\pc$ by dimensional
+analysis with the vertical frequencies of disk-like orbits. Similar offsets are seen in three-dimensional dust maps.
 

### 2019-05-10 paper/ms.tex
Add orbit integration to discussion section

@@
 
+\subsection{Orbit Integration}\label{ssec:orbit_integrate}
+We have mainly been concerned with actions, since they provide a convenient
+way to quantify different types of orbits. However, all of our conclusions
+apply also to studies which simply rely on orbit integrations, since the two
+are equivalent. For instance, computing orbital properties of open or globular
+clusters \citep[e.g.][]{2016A&A...588A.120C, 2018A&A...615A..49C,
+2018A&A...616A..12G} should ideally take the midplane variation into account.
+Orbit integrations of nearby systems over short time periods
+\citep[e.g.][]{2014MNRAS.445.2169M} are unlikely to be impacted. It should
+also be unimportant for applications to halo-like orbits, e.g., in modeling of
+stellar streams \citep[e.g.][]{2014ApJ...795...95B} or the substructure
+potentially responsible for the gap in GD1 \citep{2018arXiv181103631B}.
+
 \section{Conclusions}\label{sec:conclusion}

### 2019-05-10 paper/ms.tex
Small changes throughout

@@
 \subsection{Effect of Midplane Offset} \label{ssec:cartoon}
-We present a cartoon approximation in Figure~\ref{fig:cartoon} to show how an
+We present a cartoon approximation of an orbit in Figure~\ref{fig:cartoon} to show how an
 inaccurate or erroneous determination of the midplane leads to a phase-dependence of the actions
-calculated from an orbit specified by a phase-space point and an assumed
-potential model. The $y$-axis corresponds to the vertical height of the orbit
-and the $x$-axis shows orbital phase. The solid gray curve indicates the
-``true'' orbit of the star; that is, its orbit as it oscillates around the
-true midplane. The dashed gray line is the midplane location in the potential
-model used by an observer to integrate the orbit of the star and hence
-estimate its actions, which is offset from the true midplane of the star's
-orbit. The model potential is otherwise identical to the one in which the star
-is actually moving.
+calculated from a phase-space point and an assumed
+potential model. The $x$-axis corresponds to orbital phase
+and the $y$-axis to vertical height. The solid gray curve indicates the
+``true'' orbit of the star; that is, as it oscillates around the
+true midplane.
+
+The dashed gray line, offset from the true midplane, is the midplane location
+used by an observer to integrate the orbit of the star and estimate its
+actions. The model potential is otherwise identical to the one in which the
+star is actually moving.
 
@@
 ``guiding radius'') as $R_g$. This approximation is an excellent description
-of thin disk-like orbits and a good description of thick disk-like orbits in
+of the thin disk and a good description of the thick disk in
 an axisymmetric potential, i.e., ignoring the influence of the Galactic bar and
@@
 calculate actions for an orbital time-series starting from a phase-space
-position $(x, v)$ and integrated in a potential $\Phi$. We use the default
-\texttt{MWPotential} as our potential, which is based on the Milky Way
-potential available in \texttt{galpy} \citep{2015ApJS..216...29B}. This
-potential includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
+position $(x, v)$ and integrated in a potential $\Phi$. For our Galactic potential we use \texttt{MWPotential}, based on the Milky Way
+potential available in \texttt{galpy} \citep{2015ApJS..216...29B}, which includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
 a Miyamoto--Nagai disk \citep{1975PASJ...27..533M}, and an NFW halo
 \citep{1997ApJ...490..493N}, and is fit to empirically match some
-observations. We use the Dormand-Prince 8(5,3) integration scheme
+observations.
+We use the Dormand-Prince 8(5,3) integration scheme
 \citep{Dormand80:integrator} with a timestep of $1\,\Myr$ and integrate for
@@
 potential but by specifiying the star's starting position using the
-systematically offset coordinate system. Essentially we are shifting the
-integrated orbit and then reintegrating at each point along the shifted orbit. The actions computed using the offset coordinate system
+systematically offset coordinate system. Essentially we are shifting and then reintegrating at each point along the original orbit.
+The actions computed using the offset coordinate system
 for each phase-space starting point are shown for the first $\Gyr$ of the
@@
 
-We also perform the same procedure but assume a $100\,\pc$ offset in the $x$
-component (i.e. subtracting the vector $(100, 0, 0)\,\pc$) in the lower
-panels. This is equivalent to a systematic offset in the determination of the
-distance from the Sun to the Galactic center.
+We also perform the same procedure in the lower panels but assume an $x$
+component offset of $100\,\pc$, i.e., subtracting the vector $(100, 0,
+0)\,\pc$. This is equivalent to a measurement error in the distance from the
+Sun to the Galactic center.
 
 Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
-offset coordinate systems depend on the time (i.e. orbital phase) at which the
+offset coordinate systems depend on the time/orbital phase at which the
 star's phase-space position is observed. To the observer this would
@@
 assuming a $z$ offset of $100\,\pc$ (as in the upper right panel of
-Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted as a
+Figure~\ref{fig:one_orbit_wrong_ref}). The true value is plotted as a
 vertical dashed line. The systematic error in $J_z$ induced by a
-systematic offset in $z$ is non-Gaussian and bimodal; neither of the modes is centered on the true $J_z$ (i.e. the $J_z$ computed with no offset in $z$). In the case of the thin-disk orbit
+systematic offset in $z$ is non-Gaussian and bimodal; neither of the modes is centered on the true value (computed with no offset in $z$). In the case of the thin-disk orbit
 (bottom panel), we see that, in addition to the prior complications, the
@@
 The upper row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread
-induced in each action for an offset in the $z$-component (i.e. the Galactic
-midplane). In the lower row we consider offsets in the $x$ component (i.e. the
+induced in each action for an offset in the $z$-component. In the lower row we consider offsets in the $x$ component (i.e. the
 Solar radius). The left, center, and right columns show the fractional spread
@@
 Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
-orbit types. This is also a demonstration of the robustness of the orbit
+orbit types. This is also a demonstration of the robustness of the
 integration and action calculation methods we use.
@@
 fractional error in $J_z$ is more exaggerated for more planar (disk-like)
-orbits. For the thin-disk orbit, a systematic offset of $\sim15\,\pc$ in the
-$z$ coordinate already results in $25\%$ deviations in the actions. The
-required offset for $25\%$ deviation is $\sim120\,\pc$ for the thick-disk
-orbit. The halo orbit is relatively resistant to errors in the midplane, with
-only $\sim15\%$ error in $J_z$ out to a midplane offset of $500\,\pc$.
+orbits. For the thin-disk (thick-disk) orbit, a systematic offset of $15\,\pc$ ($120\,\pc$) in the $z$-coordinate results in a $25\%$ deviation in $J_z$.
+The halo orbit is relatively resistant to errors in the midplane, with
+only $\sim15\%$ error in $J_z$ out to an offset of $500\,\pc$.
 
@@
 (\mf) does have an ongoing interaction with a satellite galaxy similar in mass to
-Sagittarius, which has punched through the galactic disk outside the solar
+Sagittarius, which has punched through the Galactic disk outside the solar
 circle, leaving behind some of its stars and inducing warping in the disk.
@@
 \citet{2018MNRAS.481..286L}, but we briefly summarize the most relevant
-details here. For the Milky Way, the dark halo is modeled as a Hernquist
+details here.
+
+For the Milky Way, the dark halo is modeled as a Hernquist
 sphere of mass $10^{12}\,\Msun$ and scale length of $52\,\kpc$, the disk
@@
 ($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} (left,
-center, and right panels) as a function of azimuthal angle. All calculations
-are done at cosmological redshift $\z =0$. The local midplane is determined at
+center, and right panels) as a function of azimuthal angle, at cosmological redshift $\z =0$. The local midplane is determined at
 a position $\phi$ by taking the median height of all stars within
@@
 galactocentric coordinate system is incorrect, we subtract the best fit
-$A\sin{(\phi+B)}+C$ curve from each panel --- this figure is reproduced with
-the original midplane determination (i.e. before subtracting the best fit sine
-curve) in Appendix~\ref{app:lsr}. We then bootstrap resample $1000$ times on
+$A\sin{(\phi+B)}+C$ curve from each panel. We then bootstrap resample $1000$ times on
 all stars within a $2\,\kpc$ height of the fiducial midplane to determine
@@
 show the midplane as a function of azimuth before the first encounter near the
-Solar circle at $t=2.0\,\Gyr$ and $t=4.0\,\Gyr$. As later encounters occur
-which bring the Sagittarius-like object closer to the Solar circle, we can see
-that stronger midplane variations result. In fact, one can see that an $m=2$
-mode is excited in the lower left panel at $t=6.0\,\Gyr$ --- this is
-consistent with Figure~17 of \citet{2018MNRAS.481..286L}, which shows that an
-$m=2$ mode is excited ($m=0$ and $m=1$ modes are stronger, but these
-are removed in our sine-curve subtraction). The fact that the $t=6.9\,\Gyr$
-panel looks qualitatively similar to the panels from the FIRE simulations
-(Figure~\ref{fig:midplane}) suggests that midplane variations are
-produced at least in part by mergers.}
+Solar circle at $t=2.0\,\Gyr$ and $t=4.0\,\Gyr$, with an encounter happening close to the Solar circle near $t=6.0\,\Gyr$.
+The fact that the $t=6.9\,\Gyr$ panel, which shows the midplane variation after some relaxation,
+looks qualitatively similar to the panels from the FIRE simulations
+(Figure~\ref{fig:midplane}) is evidence that midplane variations are
+generated, in part, by mergers.}
 \label{fig:midplane_chervin}
@@
 To account for the effect of particle noise, we bootstrap resample stars
-within a cylinder of stars with height $2\,\kpc$ and the same radius $1000$
+within a cylinder of height $2\,\kpc$ and the same radius $1000$
 times and determine the $1\,\sigma$ error bars by repeating the midplane
@@
 Figure~\ref{fig:midplane} shows the relative $z$ location of the inferred
-midplane our imaginary observer would determine as a function of azimuth for
+midplane an observer would determine as a function of azimuth for
 each galaxy, using their local ``solar neighborhood'' (the cylinder defined
@@
 lower left panel, we see a strong $m=2$ mode develop, consistent with the
-$R=8\,\kpc$ panel of Figure~17 in \citet{2018MNRAS.481..286L}. The lower right
-panel, which shows the galaxy after some relaxation has occured is
+$R=8\,\kpc$ panel of Figure~17 in \citet{2018MNRAS.481..286L} ($m=0$ and $m=1$ modes are stronger, but these
+are removed in our sine-curve subtraction). The lower right
+panel, which shows the galaxy at $t=6.9\,\Gyr$ when some relaxation has occured, is
 qualitatively similar to the midplane variations we saw in the FIRE
@@
 
-Systematic variations in $v_z$ were first noted as asymmetries in the local
+Systematic variations in $v_z$ and number were first noted as asymmetries in the local
 velocity distribution towards the North and South Galactic Caps from the
@@
 out suggestions of an oscillation in average vertical velocities of order
-$5\,\kms$  on $\sim\,\kpc$ scales looking toward the Galactic anticenter. Work
+$5\,\kms$  on $\sim\,\kpc$ scales looking toward the Galactic anticenter.
+
+Work
 by the \textit{Gaia} collaboration confirmed these preliminary results on the
@@
 using DR2 data of median $v_z$ over a significant Galactic volume
-\citep{2018A&A...616A..11G, 2019arXiv190209569F}. We also see in the FIRE
+\citep{2018A&A...616A..11G, 2019arXiv190209569F}, which can be explained with models of Sagittarius-like encounters \citep{2013MNRAS.429..159G,2018MNRAS.481..286L,2018arXiv180800451L}.
+We also see in the FIRE
 simulation that the vertical velocity variation as a function of azimuth is

### 2019-05-10 paper/ms.tex
Replace 3 JR Jphi figures with 1 figure

@@
 
-In Figure~\ref{fig:JR_zerr_hist} we plot the distribution of $J_R$ as a function of orbital phase induced by an offset in $z$, and in Figures~\ref{fig:JR_xerr_hist} and
-\ref{fig:Jphi_xerr_hist} we plot the $J_R$ and $J_{\phi}$
-distributions induced by an offset in $x$. We plot the distributions for the
-thick-disk (top panels) and thin-disk (bottom panels) orbits.
+In Figure~\ref{fig:Jphi_JR_hist} we plot the distribution of $J_R$ as a
+function of orbital phase induced by an offset in $x$ and $z$ and the
+distribution of $J_{\phi}$ for an offset in $x$. We plot the distributions for
+the thick-disk (upper panels) and thin-disk (lower panels) orbits.
 
@@
 
-\begin{figure}
+\begin{figure*}
 \begin{center}
-\includegraphics[width=3.35224200913in]{fig/schmactions_JR_zerr_hist.pdf}
+\includegraphics[width=7.10000594991in]{fig/schmactions_Jphi_JR_hist.pdf}
 \end{center}
-\caption{A histogram of the values computed for $J_R$ for the thick-disk (top
-panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
-$100\,\pc$. The induced error distribution more closely resembles a Gaussian
-centered on the null value.}
-\label{fig:JR_zerr_hist}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.35224200913in]{fig/schmactions_JR_xerr_hist.pdf}
-\end{center}
-\caption{A histogram of the values computed for $J_R$ for the thick-disk (top
-panel) and thin-disk (bottom panel) orbits assuming a $x$ offset of
-$100\,\pc$. The induced error distribution is non-Gaussian and bimodal, with
-neither of the modes centered on null.}
-\label{fig:JR_xerr_hist}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.35224200913in]{fig/schmactions_Jphi_xerr_hist.pdf}
-\end{center}
-\caption{A histogram of the values computed for $J_{\phi}$ for the thick-disk
-(top panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
-$100\,\pc$. The induced error distribution is non-Gaussian and bimodal, with
-neither of the modes centered on null.}
-\label{fig:Jphi_xerr_hist}
-\end{figure}
+\caption{A histogram of the values computed for $J_R$ and $J_{\phi}$ for the
+thick-disk (upper panels) and thin-disk (lower panels). For $J_R$ we assume an
+$x$ offset (left) and $z$ offset (center) of $100\,\pc$, while for $J_{\phi}$
+we consider only an $x$ offset (right). In each panel the true value is given
+by a vertical dashed line. The induced error distribution in $J_R$ for an $x$
+offset more closely resembles a Gaussian centered on the null value, but not
+for the other two offsets considered.}
+\label{fig:Jphi_JR_hist}
+\end{figure*}
 

### 2019-05-10 schmactions/one_orbit/plot_one_orbit.py

@@ -122,106 +122,104 @@ plt.close()
 # JR (x error) histogram
 # # #
 
-fig, ax = plt.subplots(2,1, figsize=(columnwidth,6.5))
+fig, ax = plt.subplots(2,3, figsize=(textwidth,5))
 
-k0, k1, k2 = sclip(zact)
-dJRJR = 0.5*(np.percentile(xact[:,0][k0], 95) - np.percentile(xact[:,0][k0], 5))
+if True: # so I can indent
+    k0, k1, k2 = sclip(zact)
+    dJRJR = 0.5*(np.percentile(xact[:,0][k0], 95) - np.percentile(xact[:,0][k0], 5))
+    
+    ax[0][0].hist(xact[:,0][k0], bins=np.linspace(J0 - 5, J0 + 5, 60),
+            edgecolor=tb_c[0], fc='none', histtype='stepfilled', linewidth=1.5)
+    ax[0][0].arrow(J0, 38, dJRJR, 0, head_width=1, head_length=0.4, length_includes_head=True, color='k')
+    ax[0][0].text(J0+dJRJR/3.5, 38.5, r'$\Delta J_R$', color='k')
+    
+    ax[0][0].axvline(x=J0, color='k', ls='dashed', lw=1)
+    ax[0][0].set_ylabel(r'$\text{count}$')
+    ax[0][0].text(27.25, 42.75, r'\text{thick-disk}', color='k')
 
-ax[0].hist(xact[:,0][k0], bins=np.linspace(J0 - 5, J0 + 5, 60),
-        edgecolor=tb_c[0], fc='none', histtype='stepfilled', linewidth=1.5)
-ax[0].arrow(J0, 38, dJRJR, 0, head_width=1, head_length=0.4, length_includes_head=True, color='k')
-ax[0].text(J0+dJRJR/3.5, 38.5, r'$\Delta J_R$', color='k')
+    ax[0][0].set_title(r'$x\,\text{offset}=100\,\text{pc}$')
+    
+    k0t, k1t, k2t = sclip(xact_thin)
+    dJRJR_thin = 0.5*(np.percentile(xact_thin[:,0][k0t], 95) - np.percentile(xact_thin[:,0][k0t], 5))
+    
+    ax[1][0].hist(xact_thin[:,0][k0t], bins=np.linspace(J0_thin-5, J0_thin+5, 60),
[…]

### 2019-05-10 paper/ms.tex
Update citations, add Bailer-Jones

@@
 indicating local substructure infall \citep{2018MNRAS.481.1501B,
-2018arXiv180800451L}, and a gap
+2019MNRAS.485.3134L}, and a gap
 suggestive of perturbation by a dark matter subtructure in the tidal stream
@@
 using DR2 data of median $v_z$ over a significant Galactic volume
-\citep{2018A&A...616A..11G, 2019arXiv190209569F}, which can be explained with models of Sagittarius-like encounters \citep{2013MNRAS.429..159G,2018MNRAS.481..286L,2018arXiv180800451L}.
+\citep{2018A&A...616A..11G, 2019arXiv190209569F}, which can be explained with models of Sagittarius-like encounters \citep{2013MNRAS.429..159G,2018MNRAS.481..286L,2019MNRAS.485.3134L}.
 We also see in the FIRE
@@
 Orbit integrations of nearby systems over short time periods
-\citep[e.g.][]{2014MNRAS.445.2169M} are unlikely to be impacted. It should
-also be unimportant for applications to halo-like orbits, e.g., in modeling of
+\citep[e.g.][]{2014MNRAS.445.2169M, 2018A&A...616A..37B} are unlikely to be impacted. It should
+also be unimportant for halo applications, e.g., in modeling of
 stellar streams \citep[e.g.][]{2014ApJ...795...95B} or the substructure

### 2019-05-10 paper/ms.tex
Update orbits table, other small edits

@@
 
-\begin{deluxetable*}{cccccccc}
+\begin{deluxetable*}{cCCCCCCCCC}
 \tablecaption{Description and names of the three orbits considered in this
-work. For the last two columns: $z_{\text{max}}$ is the maximum height of the
-orbit while $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of
-the radial excursions of the orbit. In the epicyclic approximation, $A_z =
+work. For the last four columns: $z_{\text{max}}$ is the maximum height of the
+orbit, $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of
+the radial excursions of the orbit, and $\kappa$ and $\nu$ are the radial/epicyclic and vertical frequencies of the orbit. In the epicyclic approximation, $A_z =
 z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$.
@@
 \colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
-\colhead{$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$}\\ \colhead{ } &
-\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
+\colhead{$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$} & \colhead{$\kappa$} & \colhead{$\nu$} \\ \colhead{ } &
+\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km\,s}^{-1}$} &
 \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
 \colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$}}
+\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{Myr}^{-1}$} & \colhead{$\mathrm{Myr}^{-1}$}}
 \startdata 
-thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 & 1.29 \\
-thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 & 1.19 \\ 
-halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16 & 2.34
+thin-disk & (8, 0, 0) & (0, -190, 10) & 40.3 & -1520 & 0.69 & 0.12 & 1.29 & 0.049 & 0.093 \\
+thick-disk & (8, 0, 0) & (0, -190, 50) & 32.5 & -1520 & 23.0 & 0.85 & 1.19 & 0.048 & 0.061 \\ 
+halo & (8, 0, 0) & (0, -190, 190) & 32.8 & -1520 & 529.1 & 6.16 & 2.34 & 0.033 & 0.025
 \enddata
@@
 distribution of $J_{\phi}$ for an offset in $x$. We plot the distributions for
-the thick-disk (upper panels) and thin-disk (lower panels) orbits.
-
-We find that the $J_R$ distribution induced by an offset in $x$ more
+the thick-disk (upper panels) and thin-disk (lower panels) orbits. We find that the $J_R$ distribution induced by an offset in $x$ more
 closely resembles a Gaussian, while the $J_R$ distribution induced by

### 2019-05-10 schmactions/orbits_plot/plot_orbits.py

@@ -30,6 +30,8 @@ for init_vel, x in zip(init_vel_list, np.transpose(ax)):
     s = schmactions(init_pos, init_vel, save_orbit=True)
     pos = np.transpose(s.orbit.xyz.to_value(u.kpc))[:2000] # only grab the first 2 Gyr
 
+    print(s.res['freqs'])
+
     R = np.linalg.norm(pos[:,:2], axis=1)
     phi = np.mod(np.arctan2(pos[:,1], pos[:,2]), 2.*np.pi)
 

### 2019-05-10 paper/ms.tex
Add discussion of Gomez azimuthal variations

@@
 Comparing the azimuthal midplane variations in the host galaxy of these simulations before and after the interaction with the Sagittarious-like object gives an idea of the effect of one minor merger whose properties are relatively well measured.
+Azimuthal variations of the mean vertical height of stars has been explicitly pointed out in a different simulation of a Sagittarius-like encounter by \citet{2013MNRAS.429..159G}.
 
@@
 
+These values for the midplane variation are consistent with the azimuthal
+midplane variations seen by \citet{2013MNRAS.429..159G}. However, they only
+saw significant variations in their Heavy but not their Light Sagittarius model (virial masses of $10^{11}\,\Msun$ and
+$\sim3\times10^{10}\,\Msun$, respectively). The model we used (L2 from
+\citet{2018MNRAS.481..286L}) has a virial mass between these two models of
+$6\times10^{10}\,\Msun$.
+
 In the upper panels, we see that the midplane is relatively flat in the inner

### 2019-05-10 paper/ms.tex
Fix instances of e.g., italics throughout

@@
 result of \textit{Gaia} Data Release 2 (DR2).  Recent major discoveries
-include the affirmation of remnants of a major merger \citep{2018ApJ...860L..11K, 2018MNRAS.478..611B, 2018Natur.563...85H, 2019MNRAS.486..378L, 2019MNRAS.482.3426M} hinted at in pre-\textit{Gaia} work \citep[e.g.][]{2005MNRAS.359...93M, 2011MNRAS.412.1203N}, a phase-space
+include the affirmation of remnants of a major merger \citep{2018ApJ...860L..11K, 2018MNRAS.478..611B, 2018Natur.563...85H, 2019MNRAS.486..378L, 2019MNRAS.482.3426M} hinted at in pre-\textit{Gaia} work \citep[e.g.,][]{2005MNRAS.359...93M, 2011MNRAS.412.1203N}, a phase-space
 ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly
@@
 models of the stellar distribution function
-\citep[e.g.][]{1915MNRAS..76...70J, 1985ApJ...295..388V, 2017ApJ...839...61T},
+\citep[e.g.,][]{1915MNRAS..76...70J, 1985ApJ...295..388V, 2017ApJ...839...61T},
 or for associating stars with similar dynamical properties, e.g., to
@@
 the system's symmetry. However, the Galactic potential is not strictly
-axisymmetric. This has been known for some time \citep[e.g.][]{1957AJ.....62...93K,2009MNRAS.396L..56M, 2012ApJ...750L..41W}, but with the vast improvement in the quality of phase-space
+axisymmetric. This has been known for some time \citep[e.g.,][]{1957AJ.....62...93K,2009MNRAS.396L..56M, 2012ApJ...750L..41W}, but with the vast improvement in the quality of phase-space
 measurements due to \textit{Gaia} this assumption is increasingly
-inadequate \citep[e.g.][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. Even if this assumption were
+inadequate \citep[e.g.,][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. Even if this assumption were
 close enough for many purposes, the parameters used in axisymmetric models of the Galactic potential may
@@
 Before turning to numerical methods, we derive analytic expressions for the
-systematic error in the actions induced from offsets in the position (e.g. the midplane or
-Galactic center distance) or velocity (e.g. the local standard of rest) of the zero-point in the assumed coordinate system. We use the epicyclic approximation, which
+systematic error in the actions induced from offsets in the position (e.g., the midplane or
+Galactic center distance) or velocity (e.g., the local standard of rest) of the zero-point in the assumed coordinate system. We use the epicyclic approximation, which
 assumes that the motion in the $z$ and $R$ components of the orbit are
@@
 velocity curve with velocity $v_c$, a good approximation near the Solar circle
-\citep[e.g.][]{2017MNRAS.465...76M}.
+\citep[e.g.,][]{2017MNRAS.465...76M}.
 
@@
 recent works exploring actions in the Galactic disk
-\citep[e.g.][]{2019MNRAS.484.3291T, 2018MNRAS.481.4093S, 2018arXiv180803278T}.
+\citep[e.g.,][]{2019MNRAS.484.3291T, 2018MNRAS.481.4093S, 2018arXiv180803278T}.
 For disk-like orbits, existing implementations of the St\"ackel Fudge method
@@
 panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
-$100\,\pc$. One can see that if the observed $z$ values have a bias (from e.g.
+$100\,\pc$. One can see that if the observed $z$ values have a bias (from, e.g.,
 an incorrectly computed midplane), then the induced error distribution in
@@
 100)\,\pc$ from each position in the orbit. This corresponds to an
-observer physically located at e.g. the position $(8, 0, 0)\,\kpc$ in the
+observer physically located at, e.g., the position $(8, 0, 0)\,\kpc$ in the
 coordinate system of the true potential, but erroneously thinking they are
@@
 Hints of this variation as a function of Galactocentric radius have been noted
-through their impact on the stellar velocity distribution pre-\emph{Gaia} by
+through their impact on the stellar velocity distribution pre-\textit{Gaia} by
 \citet{2012ApJ...750L..41W}, \citet{2013ApJ...777L...5C}, and
-\citet{2013MNRAS.436..101W} and recently post-\emph{Gaia} by
-\citet{2019arXiv190209569F}. As pointed out by e.g.
+\citet{2013MNRAS.436..101W} and recently post-\textit{Gaia} by
+\citet{2019arXiv190209569F}. As pointed out by, e.g.,
 \citet{2014ApJ...797...53G} and \citet{2019ApJ...871..145A}, among many
@@
 \citep{2014MNRAS.445..581H,2018MNRAS.480..800H} use the zoom-in technique
-\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O} to model the formation
+\citep[e.g.,][]{1993ApJ...412..455K,2014MNRAS.437.1894O} to model the formation
 of a small group of galaxies at high resolution in a full cosmological
@@
 assumption, when starting from the present-day positions and velocities of
-stars as measured by, e.g., \emph{Gaia}.
+stars as measured by, e.g., \textit{Gaia}.
 
@@
 the location of the central supermassive black hole, Sagittarius~A\textsuperscript{*}
-\citep[\sgra{}, e.g.][]{2004ApJ...616..872R}. From stellar motions near
+\citep[\sgra{}, e.g.,][]{2004ApJ...616..872R}. From stellar motions near
 \sgra{}, the distance from the Sun to \sgra{}, referred to as $R_0$, can be
@@
 fractional error in $J_R$ is related to the error in $R_0$ as a fraction of
-the {\em radial amplitude} of the orbit, which is much smaller than $R_0$
+the \emph{radial amplitude} of the orbit, which is much smaller than $R_0$
 ($\sim1.2\,\kpc$ for the thin-disk and thick-disk orbits we considered). This
@@
 the Galactic midplane, leading to slightly different values (by $\sim 5\,\pc$)
-for the Sun's relative height \citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em
+for the Sun's relative height \citep[e.g.,][]{2019ApJ...871..145A}. A pre-\textit{
 Gaia} review of these measurements is given by \citet{2016ARAA..54..529B}.
@@
 vertical ($W_{\odot}$) components are computed by taking the mean motions of
-different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}. The azimuthal
+different stellar groups \citep[e.g.,][]{2012MNRAS.427..274S}. The azimuthal
 component ($V_{\odot}$) is more difficult to measure, but can be modeled using
@@
 value of the circular velocity is taken to be $\sim 220\kms$
-\citep[e.g.][]{2012ApJ...759..131B} with roughly $10\%$ uncertainty. We expect this to translate to at least a $10\%$ systematic error in $J_{\phi}$.
+\citep[e.g.,][]{2012ApJ...759..131B} with roughly $10\%$ uncertainty. We expect this to translate to at least a $10\%$ systematic error in $J_{\phi}$.
 
@@
 are equivalent. For instance, computing orbital properties of open or globular
-clusters \citep[e.g.][]{2016A&A...588A.120C, 2018A&A...615A..49C,
+clusters \citep[e.g.,][]{2016A&A...588A.120C, 2018A&A...615A..49C,
 2018A&A...616A..12G} should ideally take the midplane variation into account.
 Orbit integrations of nearby systems over short time periods
-\citep[e.g.][]{2014MNRAS.445.2169M, 2018A&A...616A..37B} are unlikely to be impacted. It should
+\citep[e.g.,][]{2014MNRAS.445.2169M, 2018A&A...616A..37B} are unlikely to be impacted. It should
 also be unimportant for halo applications, e.g., in modeling of
-stellar streams \citep[e.g.][]{2014ApJ...795...95B} or the substructure
+stellar streams \citep[e.g.,][]{2014ApJ...795...95B} or the substructure
 potentially responsible for the gap in GD1 \citep{2018arXiv181103631B}.

### 2019-05-10 paper/ms.tex
Other small changes throughout

@@
 Furthermore, we show that the local midplane varies by $\sim100\,\pc$ at the
-Solar circle in Milky Way-mass cosmological zoom-in simulations from the FIRE
+solar circle in Milky Way-mass cosmological zoom-in simulations from the FIRE
 project. From observations that the mean vertical velocity of the Milky Way
@@
 spiral arms. We also make the assumption of a perfectly flat circular
-velocity curve with velocity $v_c$, a good approximation near the Solar circle
+velocity curve with velocity $v_c$, a good approximation near the solar circle
 \citep[e.g.,][]{2017MNRAS.465...76M}.
@@
 
-We specifically consider \emph{azimuthal} variations in the midplane at the Solar circle, as defined by the stellar mass density.
+We specifically consider \emph{azimuthal} variations in the midplane at the solar circle, as defined by the stellar mass density.
 Since, to our knowledge, there are no direct empirical measurements of these variations in the Milky Way, we use example simulations from two classes of simulations to estimate the size of this effect.
 
-One set are three zoom-in, cosmological-hydrodynamical simulations of Milky Way-mass galaxies from the FIRE collaboration, described briefly in \S \ref{ssec:cosmozoom}.
+One set are three zoom-in, cosmological-hydrodynamical simulations of Milky Way-mass galaxies from the FIRE collaboration, described briefly in Section~\ref{ssec:cosmozoom}.
 These include stars, gas, and dark matter in a fully cosmological setting but are not tailored to specific properties of the Milky Way (such as the scale height or scale length, or the details of the accretion history).
@@
 
-The other set of simulations are isolated N-body simulations of interactions between the Milky Way and a Sagittarius-like dwarf galaxy, described briefly in \S \ref{ssec:sag_sim}.
+The other set of simulations are isolated N-body simulations of interactions between the Milky Way and a Sagittarius-like dwarf galaxy, described briefly in Section~\ref{ssec:sag_sim}.
 These include dark matter and stars and are tailored to existing measurements of the structure of the Milky Way's disk and of the orbit and properties of the Sagittarius dwarf galaxy.
@@
 show the midplane as a function of azimuth before the first encounter near the
-Solar circle at $t=2.0\,\Gyr$ and $t=4.0\,\Gyr$, with an encounter happening close to the Solar circle near $t=6.0\,\Gyr$.
+solar circle at $t=2.0\,\Gyr$ and $t=4.0\,\Gyr$, with an encounter happening close to the solar circle near $t=6.0\,\Gyr$.
 The fact that the $t=6.9\,\Gyr$ panel, which shows the midplane variation after some relaxation,
@@
 $1\,\sigma$ error from the bootstrap procedure is shown as the dashed-line
-error bars. The middle $90\%$ of midplane values across the Solar circle spans
+error bars. The middle $90\%$ of midplane values across the solar circle spans
 $185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the three cases
 the midplane therefore varies by more than $\pm 100\,\pc$ depending on the
-azimuth along the Solar circle; in the third (\mm{}, which has the thinnest
+azimuth along the solar circle; in the third (\mm{}, which has the thinnest
 ``thin disk'' of stars, but the largest stellar mass) the variation is closer
@@
 $-1.00\,\text{rad}$ and $\text{C}$ being $2.6$, $-7.8$, $-65$, and $-53\,\pc$
-chronologically for each of the four timesteps shown. The middle $90\%$ of midplane values across the Solar circle spans $49$, $62$, $137$, and $119\,\pc$, for each timestep.
+chronologically for each of the four timesteps shown. The middle $90\%$ of midplane values across the solar circle spans $49$, $62$, $137$, and $119\,\pc$, for each timestep.
 
@@
 components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}.
-Their uncertainties should also lead to systematic errors in the actions, as given in Eqs.
-\ref{eq:induced_Jphi}--\ref{eq:induced_Jz}.
+Their uncertainties should also lead to systematic errors in the actions, as given in Equations~\eqref{eq:induced_Jphi}--\eqref{eq:induced_Jz}.
 For example, the
@@
 \item To demonstrate the previous point, we measured the local galactic
-midplane along the Solar circle in three different high-resolution, zoom-in
+midplane along the solar circle in three different high-resolution, zoom-in
 simulations of Milky Way mass galaxies from the FIRE collaboration, as well as a controlled simulation of the interaction of the Milky Way with Sagittarius. We found
-that the midplane varies as a function of azimuth at the Solar circle by
+that the midplane varies as a function of azimuth at the solar circle by
 $\sim185$, $162$, $84\,\pc$ (middle $90\%$) for the three cosmological simulations we considered
@@
 dynamical Galactic center is colocated with \sgra{}. We discuss how to
-do this in \S \ref{ssec:coord_off}.
+do this in Section~\ref{ssec:coord_off}.
 
@@
 \caption{The LSR as a function of azimuth at the
-fiducial Solar circle ($R_0 = 8.2\,\kpc$). No best-fit subtraction is
+fiducial solar circle ($R_0 = 8.2\,\kpc$). No best-fit subtraction is
 performed here as we did in the case of the midplane

### 2019-05-10 paper/ms.tex
Wording changes throughout

@@
 
-Stellar actions, computed for stars from their six-dimensional phase space
-measurements and an assumed Galactic potential, are used to label and
-distinguish orbits. In principle, actions have the attractive quality of being
-invariant and thus good labels of a star's orbit. Such analyses are often used
-to interpret data from the \textit{Gaia} mission. However, inaccurate
-measurements of the phase space position and uncertainties in the assumed
-Galactic potential induce an error in the computed action. We show that, in
-addition to these complications, a systematic bias in the phase space position
+Orbital properties of stars, computed from their six-dimensional phase space
+measurements and an assumed Galactic potential, are used to understand the
+structure and evolution of the Galaxy. Stellar actions, computed from orbits,
+have the attractive quality of being invariant and thus good quantitative
+labels of a star's orbit. A systematic bias in the phase space position
 induces a phase-dependence in the actions, which we interpret as a systematic
-error. An offset in the vertical position of $\sim15\,\pc$ for a thin disk
-orbit or $\sim120\,\pc$ for a thick disk orbit is sufficient to induce a
-$25\,\%$ systematic error in the vertical action $J_z$ computed under the
-assumption of an axisymmetric potential. The induced error distribution is
+error distinct from typical observational errors and Galactic potential uncertainties.
+An offset in the vertical position of $\sim15\,\pc$ ($\sim120\,\pc$) for a thin (thick) disk orbit is sufficient to induce a
+$25\,\%$ systematic error in the vertical action $J_z$. The induced error distribution is
 non-Gaussian and bimodal, with neither of the modes peaking on the null value.
@@
 disk varies by $\sim5\textup{--}10\,\kms$ with radius, we estimate that the true midplane
-variations are of order $\sim60\textup{--}170\,\pc$. Similar offsets are seen in three-dimensional dust maps. Thus, current
-state-of-the-art action calculations --- which assume the global and local
-midplanes are the same --- are likely to include a systematic vertical offset
+variations are of order $\sim60\textup{--}170\,\pc$, with similar offsets seen in recent three-dimensional dust maps. Thus, current
+state-of-the-art action calculations, which assume the global and local
+midplanes are identical, are likely to include a systematic vertical offset
 depending on the volume considered. Variation in the local standard of rest
 induces similar issues. The variation of the midplane must be taken into
-account when performing dynamical modeling across the large regions of the
+account when performing dynamical analysis across the large regions of the
 disk accessible to \textit{Gaia} and future missions.
@@
 \keywords{Galaxy: disk -- Galaxy: evolution -- Galaxy: kinematics and dynamics
--- Galaxy: structure -- solar neighborhood -- stars: kinematics and dynamics}
+-- Galaxy: structure -- stars: kinematics and dynamics}
 
@@
 
-Underlying the quantitative analysis of many of the mechanisms at work that
-give rise to these signatures is the assumption of a global, axisymmetric
-Galactocentric coordinate system \citep{2008gady.book.....B} in which the
-Sun's position (and therefore the origin of the coordinate system) can be
-defined and measured both precisely and accurately. This involves measuring
+The assumption of a global, axisymmetric Galactocentric coordinate system \citep{2008gady.book.....B} underlies much of the quantitative analysis of the mechanisms that give rise to these signautres.
+In order to do construct such a system, the
+Sun's relative position must be
+defined and measured both precisely and accurately. This involves determining
 the angular position of and distance to the Galactic center, the orientation
 of the Galactic plane, the height of the Sun above the Galactic midplane, and
-the local standard of rest (LSR). We review the observational efforts to
+the local standard of rest (LSR). We review and discuss the observational efforts to
 measure these parameters in Section~\ref{ssec:coord_off}.
@@
 J_i \equiv
-\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{.}
+\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
 \eeq
+where $p_i$ are the conjugate momenta.
 Under the assumption of axisymmetry, $i=R,z,\phi$ are the radial, vertical,
-and azimuthal coordinates respectively in a cylindrical coordinate system and
-the $p_i$ are the conjugate momenta. In a slowly-evolving axisymmetric
+and azimuthal coordinates respectively in a cylindrical coordinate system. In a slowly-evolving axisymmetric
 potential, these actions are invariant and $J_{\phi} \equiv L_z$, where $L_z$
@@
 With the advent of 6D phase-space measurements over a relatively large
-($\gtrsim 2$ kpc) volume from the \textit{Gaia} satellite, the study of stars'
+($\gtrsim 2$ kpc) volume from the \textit{Gaia} satellite, the study of stellar
 actions has gained new popularity. One reason is dimensionality reduction---an
@@
 system, the dynamical properties of a population of stars should be uniquely a
-function of their actions and independent of the conjugate angles, allowing
-investigation of the relationship between \emph{orbital} properties of stars
-and other intrinsic, and at least partially invariant, properties such as age
+function of their actions and independent of the conjugate angles. This allows one to use actions to study the relationship between \emph{orbital} properties of stars
+and other intrinsic, and, at least partially, invariant properties such as age
 or metallicity \citep{2018ApJ...867...31B, 2018arXiv180803278T,
@@
 case of orbits near resonances with spiral arms, bar(s), and other large scale
-perturbations \citep{2014RvMP...86....1S}. For this reason, we and other
-authors have used actions to study stellar scattering in the Milky Way disk
+perturbations \citep{2014RvMP...86....1S}. For this reason, actions have been used to study stellar scattering in the Milky Way disk
 using the improved astrometry of \textit{Gaia} DR2 and various age catalogues
@@
 
-The transformation to action space is nonlinear in both the phase-space
-position and the potential, and requires relatively strong assumptions about
-the system's symmetry. However, the Galactic potential is not strictly
+The true Galactic potential is not strictly
 axisymmetric. This has been known for some time \citep[e.g.,][]{1957AJ.....62...93K,2009MNRAS.396L..56M, 2012ApJ...750L..41W}, but with the vast improvement in the quality of phase-space
@@
 
-In this work we discuss how the imposition of a global axisymmetric coordinate
-system based on a local measurement of the midplane location and
-Galactocentric distance introduces systematic error in the computation of
-actions, especially for stars in the thin disk. Our main point is that the
-``local'' and ``global'' Galactic
-midplanes are not the same. For stars relatively far from the Sun, their local midplane
-will not agree with our local midplane extrapolated onto their position due to
-relatively small-scale variations in the density of the gas and stellar
-distribution. Converting positions and velocities of more distant stars from a
+The main point of this paper is as follows. Due to relatively small-scale variations in the density of gas and stars, the
+local midplane should vary as a function of radius and azimuth. For stars
+relatively far from the Sun, their local midplane will differ from our
+local midplane extrapolated onto their position.
+Converting positions and velocities of more distant stars from a
 heliocentric to a Galactocentric coordinate system thus introduces a systematic
-bias in the $z$ coordinate, and consequently the actions computed for these stars.
-The further from the solar neighborhood the target star, the more likely the
+bias in the $z$ coordinate. We show that this bias induces non-Gaussian errors in the actions computed for these stars.
+The further from the solar neighborhood the target star is, the more likely the
 mismatch will result in large systematic uncertainty, especially in the
@@
 Section~\ref{sec:local_fire} we examine the azimuthal variations of the
-midplane itself in two sets of simulations: cosmological, hydrodynamical,
-zoom-in simulations of Milky Way-mass galaxies from the Feedback in Realistic
+midplane itself in examples from two classes of simulations: cosmological, hydrodynamical,
+zoom-in simulations of isolated Milky Way-mass galaxies from the Feedback in Realistic
 Environments (FIRE)
@@
 otherwise tailored to the stellar mass, scale length, and scale height of the
-Milky Way \citep{2018MNRAS.481..286L}.
-
-In Section~\ref{sec:discussion}, we discuss the implications of midplane
+Milky Way \citep{2018MNRAS.481..286L}. In Section~\ref{sec:discussion}, we discuss the implications of midplane
 variations, and the resulting systematic uncertainty in the vertical action,
 for action-space analyses. We also estimate the expected midplane variations
-of the Milky Way based on the observed velocity variations. We summarize our
+of the Milky Way based on the observed velocity variations and three-dimensional dust maps. We summarize our
 main results and conclude in Section~\ref{sec:conclusion}.
@@
 We first demonstrate the significance to action computations of a systematic
-offset in the determination of the Galactic midplane or distance to the
-Galactic center. We will find that such offsets are especially important for
+offset in the determination of the Galactic midplane, distance to the
+Galactic center, or LSR. We will find that such offsets are especially important for
 disk-like orbits. The consequences we explore here may also arise from various
@@
 assume that the Galaxy is perfectly described by our model axisymmetric
-potential, and simply explore the consequences of offsets in the determination
-of the Galactocentric coordinate system.
+potential, and simply explore the consequences of offsets in the Galactocentric coordinate system.
 
@@
 from an offset in another coordinate, such as the distance to the Galactic
-center or the local standard of rest. A similar effect in which actions gain
+center or the LSR. A similar effect in which actions gain
 time-dependence due to a time-varying potential was pointed out by
@@
 systematic error in the actions induced from offsets in the position (e.g., the midplane or
-Galactic center distance) or velocity (e.g., the local standard of rest) of the zero-point in the assumed coordinate system. We use the epicyclic approximation, which
+Galactic center distance) or velocity (e.g., the local standard of rest) of the assumed coordinate system's origin. We use the epicyclic approximation, which
 assumes that the motion in the $z$ and $R$ components of the orbit are
@@
 spiral arms. We also make the assumption of a perfectly flat circular
-velocity curve with velocity $v_c$, a good approximation near the solar circle
+velocity curve with $v_c(R) = v_c$, a good approximation near the solar circle
 \citep[e.g.,][]{2017MNRAS.465...76M}.
@@
 \eeq
-where $\kappa$ and $\nu$ are the radial and vertical frequencies, $\Omega_c
-\equiv v_c/R_g$ is the orbital frequency of the guiding center, $A_z$ and $A_R$ are the amplitudes of
-the stellar motion in the vertical and radial coordinates, and $\alpha$ and
+where $\kappa$ and $\nu$ are the radial/epicyclic and vertical frequencies, $\Omega_c
+\equiv v_c/R_g$ is the orbital frequency of the guiding center, $A_z$ and $A_R$ are the amplitudes in the vertical and radial coordinates, and $\alpha$ and
 $\beta$ are arbitrary phases. Similarly, the velocities of the orbit are
@@
 Using Equations~\eqref{eq:orbits_epi} and \eqref{eq:orbits_vel_epi}, we can
-simplify this and write
+simplify this:
 \beq\label{eq:JR_Jz_epi_final}
@@
 \eeq
-where we have also ignored second order terms.
+where we have again ignored second order terms.
 
@@
 We assume the Sun is located at $(8.2, 0, 0)\,\kpc$. None of our orbit
-calculations depend on the value of the LSR in this toy potential (though this
+integrations depend on the value of the LSR in this toy potential (though this
 is important when using real data, since the conversion from heliocentric to
@@
 disk, thick disk, and halo. We summarize their initial
-positions in phase space and the actions computed by integrating their orbits
-in the correct potential in Table~\ref{tab:orbits}. Each orbit, integrated
+positions in phase space, the actions computed by integrating their orbits
+in the correct potential, and other properties in Table~\ref{tab:orbits}. Each orbit, integrated
 without systematic coordinate errors, is plotted in Appendix~\ref{app:orbits}.
@@
 
-We consider that the observer makes a measurement every $\Myr$. At
-each measurement, we compute the actions as outlined above using the true
-potential but by specifiying the star's starting position using the
-systematically offset coordinate system. Essentially we are shifting and then reintegrating at each point along the original orbit.
+We consider the observer making a measurement, integrating an orbit, and
+computing actions every $\Myr$ using the prescription above. However, we
+specify the star's starting position using the systematically offset
+coordinate system.
+Essentially we are shifting and then reintegrating at each point along the original orbit.
 The actions computed using the offset coordinate system
 for each phase-space starting point are shown for the first $\Gyr$ of the
-orbit in the upper panels
+orbit in the upper panels of
 Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
@@
 
-We also perform the same procedure in the lower panels but assume an $x$
+We also perform the same procedure in the lower panels but assuming an $x$
 component offset of $100\,\pc$, i.e., subtracting the vector $(100, 0,
@@
 Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
-offset coordinate systems depend on the time/orbital phase at which the
-star's phase-space position is observed. To the observer this would
-appear as time-dependence in the actions, which appear to oscillate around
-their value using the correct coordinate system even though in this example
-appear as time-dependence in the actions, oscillating around
-their value using the correct coordinate system. This time dependence comes even though
+offset coordinate systems oscillates as a function of the time/orbital phase at which the
+star's phase-space position is observed. 
+This time dependence comes even though
 the observer is using the correctly constructed best-fit static, axisymmetric
@@
 is computed as part of the torus-fitting method.} It is worth pointing out
-that a $100\,\pc$ error in an orbit with $z_{\text{max}}=850\,\pc$ --- a
-$12\%$ error---induced an $86\%$ error in the computation of $J_z$.
+that a $100\,\pc$ error in an orbit with $z_{\text{max}}=850\,\pc$\,---\,a
+$12\%$ error\,---\,induced an $86\%$ error in the computation of $J_z$.
 
@@
 vertical dashed line. The systematic error in $J_z$ induced by a
-systematic offset in $z$ is non-Gaussian and bimodal; neither of the modes is centered on the true value (computed with no offset in $z$). In the case of the thin-disk orbit
+systematic offset in $z$ is non-Gaussian and bimodal; neither of the modes is centered on the null value. In the case of the thin-disk orbit
 (bottom panel), we see that, in addition to the prior complications, the
 distribution is not even centered on the true value. This comes about when the midplane
-error is $\gtrsim z_{\text{max}}$, where $z_{\text{max}}$ is the maximum
-height of the orbit, equivalent to $A_z$ in the epicyclic approximation (see
+error is~$\gtrsim z_{\text{max}}$, where $z_{\text{max}}$ is the maximum
+height of the orbit (equivalent to $A_z$ in the epicyclic approximation, see
 Section~\ref{ssec:epi_action}).
@@
 computed $J_R$ distribution induced by an $x$ offset more closely resembles a
-Gaussian.
+Gaussian distribution.
 
@@
 will vanish. The asymmetry and systematic offset then comes about because of
-the constraint that $J_z \geq 0$.\footnote{This is similar to arguments in
+the constraint that $J_z \geq 0$.\footnote{This argument is similar to ones given in
 cosmology for why gravity produces non-Gaussianity in the density field, since
@@
 percentiles used. This summary statistic does not reflect the bias induced
-when the midplane error is $\gtrsim z_{\text{max}}$.
+when the midplane error is~$\gtrsim z_{\text{max}}$.
 
@@
 but for systematic offsets between $0$ and $500\,\pc$ in the $z$ and $x$
-components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report one half the
-95\uth{} minus 5\uth{} percentile divided by the true action value ($\Delta
-J_i/J_i$) for the three different fiducial orbits in Table~\ref{tab:orbits}.
+components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report $\Delta
+J_i/J_i$ for the three different fiducial orbits in Table~\ref{tab:orbits}.
 The thick-disk (\thickcolor) orbit is the one from
@@
 
-The upper row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread
-induced in each action for an offset in the $z$-component. In the lower row we consider offsets in the $x$ component (i.e. the
+The upper panels of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread
+induced in each action for an offset in the $z$-component. In the lower panels we consider offsets in the $x$ component (i.e. the
 Solar radius). The left, center, and right columns show the fractional spread
-in the values computed for $J_z$, $J_{\phi}$, and $J_R$, respectively. We
-compute the fractional spread $\Delta J_i/J_i$, where $\Delta J_i$ is defined
-in the previous paragraph and $J_i$ is the action value computed in the true
-coordinate system, over the course of the first $\Gyr$.
+in the values computed for $J_z$, $J_{\phi}$, and $J_R$, respectively.
 
@@
 relevant since during most of the orbit the star will be at maximum radial and
-vertical amplitude and we do not consider velocity offsets. Note that we
+vertical amplitude. Note that we
 consider an error in the $x$-coordinate $\Delta x$, which is not exactly the
 same as $\Delta R$. For observations of stars close to us, we have that
-$\Delta x \sim \Delta R$, but for the experiment performed in this Section we
+$\Delta x \sim \Delta R$, but for the experiment performed in this section we
 consider observations of the star throughout its entire orbit. This introduces
@@
 
-One set are three zoom-in, cosmological-hydrodynamical simulations of Milky Way-mass galaxies from the FIRE collaboration, described briefly in Section~\ref{ssec:cosmozoom}.
+One set are three zoom-in, cosmological hydrodynamical simulations of isolated Milky Way-mass galaxies from the FIRE collaboration, described briefly in Section~\ref{ssec:cosmozoom}.
 These include stars, gas, and dark matter in a fully cosmological setting but are not tailored to specific properties of the Milky Way (such as the scale height or scale length, or the details of the accretion history).
@@
 These include dark matter and stars and are tailored to existing measurements of the structure of the Milky Way's disk and of the orbit and properties of the Sagittarius dwarf galaxy.
-Comparing the azimuthal midplane variations in the host galaxy of these simulations before and after the interaction with the Sagittarious-like object gives an idea of the effect of one minor merger whose properties are relatively well measured.
+Comparing the azimuthal midplane variations in the host galaxy of these simulations before and after the interaction with the Sagittarius-like object gives an idea of the effect of one minor merger whose properties are relatively well measured.
 Azimuthal variations of the mean vertical height of stars has been explicitly pointed out in a different simulation of a Sagittarius-like encounter by \citet{2013MNRAS.429..159G}.
@@
 disks \citep{2014MNRAS.445..581H, 2016MNRAS.456.2140M, 2017MNRAS.467.2430M,
-2016ApJ...827L..23W, 2018MNRAS.481.4133G, 2018MNRAS.480..800H}. For this work,
+2016ApJ...827L..23W, 2018MNRAS.481.4133G, 2018MNRAS.480..800H}.
+
+For this work,
 we focus on the three Milky Way-mass zoom-ins considered in
@@
 \tablecaption{Stellar and gas disk scale heights of the real Milky Way and the
-simulated galaxies considered in this work. For comparison, we also give the
+simulated galaxies considered in this work. For comparison, we also give the median
 softening lengths for the simulated galaxies, computed for cold gas ($T<1000\,\text{K}$) and stars with $\abs{R-R_0} < 0.5\,\kpc$ and $\abs{z} < 1\,\kpc$.
@@
 These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
-gas particles of mass $\sim 7000$ to $20,000\,\Msun$, and star particles of
-mass $\sim 5000 \text{--} 7000\, \Msun$, with the lower end coming from
+gas particles of mass $\sim7000$ to $20,000\,\Msun$, and star particles of
+mass $\sim5000$ to $7000\,\Msun$, with the lower end coming from
 stellar evolution \citep{2018arXiv180610564S}. Softening lengths for dark
@@
 discussed in \citet{2018arXiv180610564S}. For instance, the velocity structure
-of \mi{} is closer to M31's than the Milky Way's (Loebman et al. in prep).
+of \mi{} is closer to M31's than the Milky Way's (S. Loebman et al., in preparation).
 However, in this work we are most interested in the global properties of the
@@
 \citep{1990ApJ...356..359H}. We use the L2 model, for which Sagittarius is
-modeled as a dark matter Hernquist sphere of mass $8\times10^{10}\,\Msun$ and
-scale length $8\,\kpc$, along with a stellar component modeled as a Hernquist
-sphere of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$. All
+modeled with two components: a dark matter Hernquist sphere of mass $8\times10^{10}\,\Msun$ and
+scale length $8\,\kpc$, and a stellar component modeled as a Hernquist
+sphere of mass $6.4\times10^8\,\Msun$ and scale length $0.85\,\kpc$. All
 components are realized with live distributions of N-body particles; the Milky
@@
 observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and
-vary the azimuth between $0<\phi<2\pi$. At each value of $\phi$ we then
+vary the azimuth between $0<\phi<2\pi$. We consider $50$ equally spaced bins in azimuth, a sufficient number
+such that each cylinder shares no stars with its neighboring cylinder. At each value of $\phi$ we then
 compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and
@@
 times and determine the $1\,\sigma$ error bars by repeating the midplane
-determination with those reselection of stars.
+determination with that reselection.
 
@@
 simplified version of the strategy described in
-\citealt{2019ApJ...871..145A}). For $\text{A}$ the values are $-165$, $45$,
+\citealt{2019ApJ...871..145A}). For $\text{A}$, the values are $-165$, $45$,
 and $8.8\,\pc$, for $\text{B}$ the values are $0.67$, $-0.088$, and
@@
 midplane an observer would determine as a function of azimuth for
-each galaxy, using their local ``solar neighborhood'' (the cylinder defined
-above). We consider $50$ equally spaced bins in azimuth, a sufficient number
-such that each cylinder shares no stars with its neighboring cylinder. The
+each galaxy, using their local solar neighborhood (the cylinder defined
+above). The
 $1\,\sigma$ error from the bootstrap procedure is shown as the dashed-line
@@
 ``thin disk'' of stars, but the largest stellar mass) the variation is closer
-to $\pm 50\,\pc$.
+to~$\pm 50\,\pc$.
 
@@
 $\sim3\times10^{10}\,\Msun$, respectively). The model we used (L2 from
-\citet{2018MNRAS.481..286L}) has a virial mass between these two models of
-$6\times10^{10}\,\Msun$.
+\citet{2018MNRAS.481..286L}) has a virial mass of $6\times10^{10}\,\Msun$, intermediate between their two models.
 
@@
 \section{Discussion} \label{sec:discussion} 
-We have used high-resolution simulations to illustrate that we
-expect the ``local midplane'' defined by stellar density to vary with azimuth
-by up to $\pm 100$ pc, as a natural consequence of the non-axisymmetry of the
-Galactic disk at small scales.
-While this is not in itself surprising or new, we also demonstrated that extending the coordinate system established by our local midplane to a global axisymmetric coordinate system spanning the entirety of the Galactic disk
-introduces systematic error in computations of the actions under this symmetry
-assumption, when starting from the present-day positions and velocities of
-stars as measured by, e.g., \textit{Gaia}.
+We have used high-resolution simulations to illustrate that we expect the
+local midplane defined by stellar density to vary with azimuth by up to $\pm
+100$ pc, as a natural consequence of the non-axisymmetry of the Galactic disk
+at small scales. While this is not in itself surprising or new, we have also
+discussed that the discrepancy between our local midplane and that of distant
+stars introduces a systematic error in the $z$ component when converting from
+heliocentric to Galactocentric coordinates. This systematic error introduces a
+non-Gaussian error in the vertical action, $J_z$, when starting from the
+present-day positions and velocities of stars as measured by, e.g.,
+\textit{Gaia}.
 
@@
 axisymmetry, such as resonant perturbation by spiral arms or scattering from
-molecular clouds.
-
-This finding has many implications for the study of the dynamics of stars in
-what is normally considered the regime of the epicyclic approximation:
-quasi-harmonic oscillations around a circular ``guiding center orbit.''
+molecular clouds \citep{2014RvMP...86....1S}.
 
@@
 The vertical frequency of the thin-disk and thick-disk orbits are $\sim
-0.09\,\Myr^{-1}$ and $\sim0.06\,\Myr^{-1}$, respectively. By dimensional
+0.09\,\Myr^{-1}$ and $\sim0.06\,\Myr^{-1}$, respectively (Table~\ref{tab:orbits}). By dimensional
 analysis, and assuming a vertical velocity variation of $5\textup{--}10\,\kms$, we
@@
 
-\subsection{Uncertainties in the Solar position and velocity}\label{ssec:coord_off}
+\subsection{Uncertainties in the Solar Position and Velocity}\label{ssec:coord_off}
 
@@
 any observed star.
-Considerable effort has been placed on each of these measurements, but uncertainties remain, and detailed dynamical modeling across the disk---particularly for dynamically cold stars---may have to take them into
+Considerable effort has been placed on each of these measurements, but uncertainties remain, and detailed modeling across the disk\,---\,particularly for dynamically cold stars\,---\,may have to take them into
 account. 
-Here we briefly review the current measurements and their uncertainties, and the implications for the calculation of actions.
+Here we briefly review the current measurements, their uncertainties, and the implications for the calculation of actions.
 
@@
 \citep[\sgra{}, e.g.,][]{2004ApJ...616..872R}. From stellar motions near
-\sgra{}, the distance from the Sun to \sgra{}, referred to as $R_0$, can be
-precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}, with a
-recent measurement using near-infrared interferometry of $8.178 \pm 0.035$~kpc
-\citep{2019arXiv190405721A}, or $\sim0.3\%$. However, the location of \sgra{} may not be equivalent to the location of the ``dynamical Galactic center,'' i.e., the
+\sgra{}, the distance from the Sun to \sgra{}, $R_0$, can be
+precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}. A
+recent measurement using near-infrared interferometry places $R_0$ at $8.178 \pm 0.035$~kpc
+\citep{2019arXiv190405721A}, or a $\sim0.3\%$ uncertainty.
+
+However, the location of \sgra{} may not be equivalent to the location of the ``dynamical Galactic center,'' i.e., the
 point in three-dimensional space about which the stars in the solar
@@
 If the dynamical Galactic center is offset from \sgra{} by $100\,\pc$, only a
-$1.2\%$ difference, then this induces a $21\%$ error in $J_R$ (see
+$1.2\%$ difference, then this induces a $\sim20\%$ error in $J_R$ for the disk-like orbits we considered (see
 Section~\ref{ssec:quant}). The reason such a large error in $J_R$ can be
 generated by a small error in $R_0$ can be understood from the epicyclic
-approximation, which states that $\Delta J_R/J_R = 2\Delta R/A_R$. The
+approximation (Equation~\eqref{eq:Ji_err_mosttime}, which states that $\Delta J_R/J_R = 2\Delta R/A_R$. The
 fractional error in $J_R$ is related to the error in $R_0$ as a fraction of
@@
 \section{Conclusions}\label{sec:conclusion}
-Actions have promise as excellent orbit labels. If the Galaxy can be
+Determining the orbital properties of stars is important for understanding the structure and evolution of the Galaxy. Actions have promise as excellent orbit labels. If the Galaxy can be
 approximated as axisymmetric and 6D phase space positions can be measured
@@
 local Galactic midplane is the global Galactic midplane is not true a
-priori.
-
-\item To demonstrate the previous point, we measured the local galactic
+priori. A violation of this assumption (by, e.g., intrinsic midplane variations) leads to a systematic error in $z$ which generates the issues raised in the last three points.
+
+\item To demonstrate the severity of the previous point, we measured the local galactic
 midplane along the solar circle in three different high-resolution, zoom-in
@@
 simply rely on orbit integration, since the computation of actions and orbit
-integrations are mathematically equivalent.
+integrations are equivalent.
 

### 2019-05-11 paper/ms.tex
Change to typewriter for orbits, fix softening lengths

@@
 
+\newcommand{\thin}{\texttt{thin-disk}}
+\newcommand{\thick}{\texttt{thick-disk}}
+\newcommand{\halo}{\texttt{halo}}
 \newcommand{\thincolor}{pink}
@@
 \startdata 
-thin-disk & (8, 0, 0) & (0, -190, 10) & 40.3 & -1520 & 0.69 & 0.12 & 1.29 & 0.049 & 0.093 \\
-thick-disk & (8, 0, 0) & (0, -190, 50) & 32.5 & -1520 & 23.0 & 0.85 & 1.19 & 0.048 & 0.061 \\ 
-halo & (8, 0, 0) & (0, -190, 190) & 32.8 & -1520 & 529.1 & 6.16 & 2.34 & 0.033 & 0.025
+\thin{} & (8, 0, 0) & (0, -190, 10) & 40.3 & -1520 & 0.69 & 0.12 & 1.29 & 0.049 & 0.093 \\
+\thick{} & (8, 0, 0) & (0, -190, 50) & 32.5 & -1520 & 23.0 & 0.85 & 1.19 & 0.048 & 0.061 \\ 
+\halo{} & (8, 0, 0) & (0, -190, 190) & 32.8 & -1520 & 529.1 & 6.16 & 2.34 & 0.033 & 0.025
 \enddata
@@
 \caption{The artificial phase-dependence in the computed actions with an
-error in the Galactocentric coordinate system. We consider here the thick-disk
-orbit, which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520,
+error in the Galactocentric coordinate system. We consider here \thick{},
+which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520,
 7.0)\,\actunit$ and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We
@@
 \end{center}
-\caption{A histogram of the computed values of $J_z$ at different orbital phases for the thick-disk (top
-panel) and thin-disk (bottom panel) orbits assuming a $z$ offset of
+\caption{A histogram of the computed values of $J_z$ at different orbital phases for \thick{} (top
+panel) and \thin{} (bottom panel) orbits assuming a $z$ offset of
 $100\,\pc$. One can see that if the observed $z$ values have a bias (from, e.g.,
@@
 without systematic coordinate errors, is plotted in Appendix~\ref{app:orbits}.
-We will refer to these orbits by their names (thin-disk, thick-disk, halo)
+We will refer to these orbits by their names (\thin{}, \thick{}, \halo{})
 henceforth.
 
-We begin with the thick-disk orbit. Consider an observer who can
-measure the thick-disk star's phase-space position at many different times
+We begin with \thick{}. Consider an observer who can
+measure the orbit's phase-space position at many different times
 (and hence different orbital phases), but does so using a
@@
 
-In Figure~\ref{fig:Jz_hist} we plot a histogram of the values of
-$J_z$ computed at different orbital phases for the thick-disk orbit (top panel) and thin-disk orbit (bottom panel)
+In Figure~\ref{fig:Jz_hist}, we plot a histogram of the values of
+$J_z$ computed at different orbital phases for \thick{} (top panel) and \thin{} (bottom panel),
 assuming a $z$ offset of $100\,\pc$ (as in the upper right panel of
@@
 vertical dashed line. The systematic error in $J_z$ induced by a
-systematic offset in $z$ is non-Gaussian and bimodal; neither of the modes is centered on the null value. In the case of the thin-disk orbit
+systematic offset in $z$ is non-Gaussian and bimodal; neither of the modes is centered on the null value. In the case of \thin{}
 (bottom panel), we see that, in addition to the prior complications, the
@@
 \caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error
-in the measured action ($\Delta J_i$) from coordinate system errors for the
-thin, thick, and halo orbits (Table~\ref{tab:orbits}). See discussion in the
+in the measured action ($\Delta J_i$) from coordinate system errors for
+\thin{}, \thick{}, and \halo{} orbits (Table~\ref{tab:orbits}). See discussion in the
 text for the justification in using this to measure the magnitude of the
@@
 the $J_z$ error. An $x$ offset induces an error in $J_R$ and $J_{\phi}$. The
-error in $J_R$ is somewhat well-described for the thin-disk and thick-disk
-orbits, and a poor description for the halo orbit. For $J_{\phi}$, the
+error in $J_R$ is somewhat well-described for \thin{} and \thick{}, 
+and a poor description for \halo{}. For $J_{\phi}$, the
 epicyclic approximation is not a good description for any orbit.}
@@
 We now describe the intuition behind the shape of Figure~\ref{fig:Jz_hist}.
-Consider first the thick-disk orbit (top panel), where the offset in $z$ is
+Consider first the \thick{} (top panel), where the offset in $z$ is
 much less than $z_{\text{max}}$ of the orbit. The peaks in the distribution
@@
 amplitude), where $v_z \sim 0$ and where the star is on most of it orbit. This
-is why the distribution, which is calculated at evenly spaced time intervals, peaks at these values. For the thin-disk orbit
+is why the distribution, which is calculated at evenly spaced time intervals, peaks at these values. For the \thin{}
 (bottom panel), the offset in $z$ is comparable to $z_{\text{max}}$. Now,
@@
 J_i/J_i$ for the three different fiducial orbits in Table~\ref{tab:orbits}.
-The thick-disk (\thickcolor) orbit is the one from
-Figure~\ref{fig:one_orbit_wrong_ref}, but we also now consider the effect on
-the action determined for the thin-disk (\thincolor) and halo (\halocolor)
-orbits.
+Figure~\ref{fig:one_orbit_wrong_ref} considered \thick{} (\thickcolor), but we also now consider the effect on
+the action determined for \thin{} (\thincolor) and \halo{} (\halocolor).
 
@@
 $z$, as discussed earlier. Indeed, the result we found in
-Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all
+Figure~\ref{fig:one_orbit_wrong_ref} for \thick{} holds for all
 orbit types. This is also a demonstration of the robustness of the
@@
 fractional error in $J_z$ is more exaggerated for more planar (disk-like)
-orbits. For the thin-disk (thick-disk) orbit, a systematic offset of $15\,\pc$ ($120\,\pc$) in the $z$-coordinate results in a $25\%$ deviation in $J_z$.
-The halo orbit is relatively resistant to errors in the midplane, with
+orbits. For \thin{} (\thick{}), a systematic offset of $15\,\pc$ ($120\,\pc$) in the $z$-coordinate results in a $25\%$ deviation in $J_z$.
+We find that \halo{} is relatively resistant to errors in the midplane, with
 only $\sim15\%$ error in $J_z$ out to an offset of $500\,\pc$.
@@
 
-The epicyclic approximation is a good predictor of $\Delta J_z$, even for halo
-orbits. It performs similarly for $\Delta J_R$, now underpredicting for halo
-orbits and slightly overpredicting for thin-disk orbits. Note that for the
-particular orbits we chose, the thin-disk orbit has slightly larger $A_R$ than
-the thick-disk orbit, and so we expect the epicyclic approximation to perform
-slightly worse. The epicyclic approximation underpredicts $\Delta J_{\phi}$
+The epicyclic approximation is a good predictor of $\Delta J_z$, even for \halo{}.
+It performs similarly for $\Delta J_R$, now underpredicting for \halo{}
+and slightly overpredicting for \thin{}. Note that for the
+particular orbits we chose, \thin{} has slightly larger $A_R$ than
+\thick{}, and so we actually expect the epicyclic approximation to perform
+slightly worse for \thin{} in this case. The epicyclic approximation underpredicts $\Delta J_{\phi}$
 for all orbits.
@@
 minus 5\uth{} percentile of the distribution of $J_z$ values over the
-course of the orbit. There are large errors for thin-disk like orbits ($J_z
+course of the orbit. There are large errors for \thin{} like orbits ($J_z
 \sim 0.5\,\actunit$), even for a small midplane offset of $10\,\pc$. As dashed
@@
 $(8,0,0)\,\kpc$ and the initial velocity to be $(0, -190, v_z)\,\kms$, where
-we vary $v_z$.\footnote{The thin-disk, thick-disk, and halo orbits correspond to setting $v_z=10$, $50$, and $190$, respectively (see Table~\ref{tab:orbits}).} For a thin disk-like orbit ($J_z\sim0.5\,\actunit$), even a
+we vary $v_z$.\footnote{One can recover \thin{}, \thick{}, and \halo{} by setting $v_z=10$, $50$, and $190$, respectively (see Table~\ref{tab:orbits}).} For a \thin{}-like orbit ($J_z\sim0.5\,\actunit$), even a
 $10\,\pc$ offset in $z$ is enough to induce a $\sim25\%$ error in $J_z$. For
@@
 Milky Way\tablenotemark{b} & 40 & 300 & 900 & \nodata & \nodata \\
-\mi\tablenotemark{c} & 800\tablenotemark{d} & 480 & 2000 & 12.4 & 11.2 \\
-\mf\tablenotemark{c} & 360 & 440 & 1280 & 12.5 & 11.2 \\
-\mm\tablenotemark{c} & 250 & 290 & 1030 &10.3 & 11.2 \\
+\mi\tablenotemark{c} & 800\tablenotemark{d} & 480 & 2000 & 53.4 & 11.2 \\
+\mf\tablenotemark{c} & 360 & 440 & 1280 & 57.2 & 11.2 \\
+\mm\tablenotemark{c} & 250 & 290 & 1030 & 60.1 & 11.2 \\
 \enddata
 % \tablerefs{\citet{2016ARAA..54..529B,2018arXiv180610564S,2008ApJ...673..864J}}
-\tablenotetext{a}{$T<1000$ K}
+\tablenotetext{a}{$T<100\,\text{K}$}
 \tablenotetext{b}{\citet{2008ApJ...673..864J,2016ARAA..54..529B}}
@@
 Plummer-equivalent.} The gas softening length is adaptive, but at $\z=0$ the
-median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
+median softening length for cold ($T < 100\,\text{K}$) gas particles around
 roughly solar positions (with galactocentric cylindrical radii within
-$0.5\,\kpc$ of $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4$, $12.5$, and
-$10.3\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. These values are
+$0.5\,\kpc$ of $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $53.4$, $57.2$, and
+$60.1\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. These values are
 summarized in Table~\ref{tab:scale_height}, along with measurements of the
@@
 dense, star-forming gas disk is thought to have a scale height of about
-$40\,\pc$, $\sim3\textup{--}4$ times the cold gas softening length
+$40\,\pc$, on the order of the cold gas softening length
 \citep{2019ApJ...871..145A}. The thin stellar disk has a scale height of about
 $300\,\pc$, $\sim30$ times the stellar softening length
-\citep{2008ApJ...673..864J}. We therefore expect that resolution effects may
-still be affecting the scale heights of these components in the simulations,
+\citep{2008ApJ...673..864J}. We therefore expect that resolution effects are
+still affecting the scale heights of these components in the simulations,
 especially the cold gas. Indeed, the stellar scale heights of the simulated
@@
 
-The vertical frequency of the thin-disk and thick-disk orbits are $\sim
+The vertical frequency of \thin{} and \thick{} are $\sim
 0.09\,\Myr^{-1}$ and $\sim0.06\,\Myr^{-1}$, respectively (Table~\ref{tab:orbits}). By dimensional
@@
 the \emph{radial amplitude} of the orbit, which is much smaller than $R_0$
-($\sim1.2\,\kpc$ for the thin-disk and thick-disk orbits we considered). This
+($\sim1.2\,\kpc$ for \thin{} and \thick{}). This
 also implies the very precise $0.3\%$ measurement of $R_0$ still translates to
@@
 corresponding to a vertical height of $\sim22\,\pc$. This corresponds to a
-$37\%$ ($5\%$) error in $J_z$ for the thin-disk (thick-disk) orbit.
+$37\%$ ($5\%$) error in $J_z$ for \thin{} (\thick{}).
 
@@
 \caption{The three orbits presented in Table~\ref{tab:orbits} and considered
-throughout the work. We plot the thin-disk, thick-disk, and halo orbits in the
+throughout the work. We plot \thin{}, \thick{}, and \halo{} in the
 left, center, and right columns, respectively. The upper row shows a plot of
@@
 distribution of $J_{\phi}$ for an offset in $x$. We plot the distributions for
-the thick-disk (upper panels) and thin-disk (lower panels) orbits. We find that the $J_R$ distribution induced by an offset in $x$ more
+\thick{} (upper panels) and \thin{} (lower panels). We find that the $J_R$ distribution induced by an offset in $x$ more
 closely resembles a Gaussian, while the $J_R$ distribution induced by
@@
 \end{center}
-\caption{A histogram of the values computed for $J_R$ and $J_{\phi}$ for the
-thick-disk (upper panels) and thin-disk (lower panels). For $J_R$ we assume an
+\caption{A histogram of the values computed for $J_R$ and $J_{\phi}$ for
+\thick{} (upper panels) and \thin{} (lower panels). For $J_R$ we assume an
 $x$ offset (left) and $z$ offset (center) of $100\,\pc$, while for $J_{\phi}$

### 2019-05-11 schmactions/many_orbits/plot_many_orbits.py

@@ -52,7 +52,7 @@ def sclip(a, s=4):
     k2 = np.where(a2bool)[0]
     return k0, k1, k2
 
-name_list = ['thin-disk', 'thick-disk', 'halo']
+name_list = [r'\texttt{thin-disk}', r'\texttt{thick-disk}', r'\texttt{halo}']
 fname_list = ['thin', 'thick', 'halo']
 Az_list = np.array([120, 850, 6160]) # pc
 AR_list = np.array([1290, 1190, 2340]) # pc

### 2019-05-11 schmactions/one_orbit/plot_one_orbit.py

@@ -105,11 +105,11 @@ for x in ax[1]:
     x.set_xticks(np.arange(0,1000,100), minor=True)
 
 for x in ax[:,0]:
-    x.set_ylabel(r'$J_{R,\text{obs}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+    x.set_ylabel(r'$J_{R,\text{comp}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
 for x in ax[:,1]:
-    x.set_ylabel(r'$J_{\phi,\text{obs}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+    x.set_ylabel(r'$J_{\phi,\text{comp}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
 for x in ax[:,2]:
-    x.set_ylabel(r'$J_{z,\text{obs}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+    x.set_ylabel(r'$J_{z,\text{comp}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
 
 ax[0][1].set_title(r'$z\,\text{offset}=100\,\text{pc}$')
 ax[1][1].set_title(r'$x\,\text{offset}=100\,\text{pc}$')
@@ -135,7 +135,7 @@ if True: # so I can indent
     
     ax[0][0].axvline(x=J0, color='k', ls='dashed', lw=1)
     ax[0][0].set_ylabel(r'$\text{count}$')
-    ax[0][0].text(27.25, 42.75, r'\text{thick-disk}', color='k')
+    ax[0][0].text(27.25, 42.75, r'\texttt{thick-disk}', color='k')
 
     ax[0][0].set_title(r'$x\,\text{offset}=100\,\text{pc}$')
     
@@ -149,9 +149,9 @@ if True: # so I can indent
  
     ax[1][0].axvline(x=J0_thin, color='k', ls='dashed', lw=1)
     ax[1][0].set_ylabel(r'$\text{count}$')
-    ax[1][0].text(35.1, 32.5, r'\text{thin-disk}', color='k')
+    ax[1][0].text(35.1, 32.5, r'\texttt{thin-disk}', color='k')
  
-    ax[1][0].set_xlabel(r'$J_{R,\text{obs}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
[…]

### 2019-05-11 schmactions/orbits_plot/plot_orbits.py

@@ -49,9 +49,9 @@ ax[0][0].set_ylabel(r'$y\,[\,\text{kpc}\,]$')
 for x in ax[0]:
     x.set_xlabel(r'$x\,[\,\text{kpc}\,]$')
 
-ax[0][0].set_title(r'thin-disk')
-ax[0][1].set_title(r'thick-disk')
-ax[0][2].set_title(r'halo')
+ax[0][0].set_title(r'\texttt{thin-disk}')
+ax[0][1].set_title(r'\texttt{thick-disk}')
+ax[0][2].set_title(r'\texttt{halo}')
 
 ax[1][0].set_ylabel(r'$z\,[\,\text{kpc}\,]$')
 for x in ax[1]:

### 2019-05-11 short_info/median_gas_smoothing_length.py

@@ -4,14 +4,14 @@ import numpy as np
 Rsun = 8.2
 dR = 0.5
 dZ = 1.0
-maxT = 1E3
+maxT = 100
 
 def read_snap(gal):
     # takes in galaxy (string = m12i, m12f, m12m, etc.)
     # reads in and sets fiducial coordinates
     # returns snap
     gal_info = 'fiducial_coord/' + gal + '_res7100_center.txt'
-    sim_directory = '/Users/abeane/scratch/actions_systematic/data/fire2/metaldiff/'+gal+'_res7100'
+    sim_directory = '/mnt/ceph/users/firesims/fire2/metaldiff/'+gal+'_res7100'
     snap = gizmo.io.Read.read_snapshots(['gas'], 'index', 600,
                                         properties=['position',
                                                     'smooth.length', 'temperature'],
@@ -38,5 +38,5 @@ for gal in ['m12i', 'm12f', 'm12m']:
     Tbool = snap['gas']['temperature'] < maxT
 
     keys = np.where(np.logical_and(np.logical_and(rbool, zbool), Tbool))[0]
-    print(gal, 2.8 * np.median( snap['gas']['smooth.length'] ) /1000, ' kpc' )
+    print(gal, 2.8 * np.median( snap['gas']['smooth.length'][keys] ) /1000, ' kpc' )
 

### 2019-05-11 paper/ms.tex
Small edits

@@
 \usepackage{makecell}
+\usepackage{scrextend}
 
@@
 corresponding to an erroneous observer assuming a midplane (upper) or solar
-radius (lower) that is off by $100\,\pc$. We then allow an immortal observer
-to observe the orbit over $1\,\Gyr$ and perform the same orbit integration
+radius (lower) that is off by $100\,\pc$. We then allow an observer
+to measure the orbit over $1\,\Gyr$ and perform the same orbit integration
 procedure at each timestep, and report the values of the actions. The
@@
 \caption{A histogram of the computed values of $J_z$ at different orbital phases for \thick{} (top
-panel) and \thin{} (bottom panel) orbits assuming a $z$ offset of
+panel) and \thin{} (bottom panel) assuming a $z$ offset of
 $100\,\pc$. One can see that if the observed $z$ values have a bias (from, e.g.,
@@
 \thin{}, \thick{}, and \halo{} orbits (Table~\ref{tab:orbits}). See discussion in the
-text for the justification in using this to measure the magnitude of the
+text and Figure~\ref{fig:Jz_hist} for the justification in using this to measure the magnitude of the
 induced error. The left, center, and right panels show the result for $J_R$,
@@
 different offsets in $z$. All orbits have the same initial position of $(8, 0,
-0)\,\kpc$ and velocity $(0, -190, v_z)\,\kms$, where we vary $v_z$. We show
+0)\,\kpc$ and velocity $(0, -190, v_z)\,\kms$, where we vary $v_z$.\footref{note:vz_orbits} We show
 this for a $z$ offset of $10, 50, 100\,\pc$ (orange, teal, and green
@@
 $(8,0,0)\,\kpc$ and the initial velocity to be $(0, -190, v_z)\,\kms$, where
-we vary $v_z$.\footnote{One can recover \thin{}, \thick{}, and \halo{} by setting $v_z=10$, $50$, and $190$, respectively (see Table~\ref{tab:orbits}).} For a \thin{}-like orbit ($J_z\sim0.5\,\actunit$), even a
+we vary $v_z$.\footnote{\label{note:vz_orbits}One can recover \thin{}, \thick{}, and \halo{} by setting $v_z=10$, $50$, and $190$, respectively (see Table~\ref{tab:orbits}).} For a \thin{}-like orbit ($J_z\sim0.5\,\actunit$), even a
 $10\,\pc$ offset in $z$ is enough to induce a $\sim25\%$ error in $J_z$. For

### 2019-05-11 paper/ms.tex
Small changes throughout

@@
 \newcommand{\z}{z_r}
-\newcommand{\uth}{\textsuperscript{th}}
 \newcommand{\n}{\text{n}}
@@
 An offset in the vertical position of $\sim15\,\pc$ ($\sim120\,\pc$) for a thin (thick) disk orbit is sufficient to induce a
-$25\,\%$ systematic error in the vertical action $J_z$. The induced error distribution is
+$25\%$ systematic error in the vertical action $J_z$. The induced error distribution is
 non-Gaussian and bimodal, with neither of the modes peaking on the null value.
@@
 take this into account. Intuition for the shape of each panel is given in the
-text. We also plot one half the 95\uth{} percentile minus the 5\uth{}
+text. We also plot one half the 95th percentile minus the 5th
 percentile of each distribution as a horizontal arrow anchored on the true
@@
 minus 5\textsuperscript{th} percentiles are $2.2$ and $6.2\,\actunit$ for
-$J_R$ and $J_z$, respectively. These are fractional errors of $5.7\,\%$ and
-$86\,\%$, respectively. The error induced in $J_{\phi}$ is negligible, as
+$J_R$ and $J_z$, respectively. These are fractional errors of $5.7\%$ and
+$86\%$, respectively. The error induced in $J_{\phi}$ is negligible, as
 expected since $J_{\phi}$ only depends on the $x$- and $y$-components of the
@@
 \end{center}
-\caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error
+\caption{We report one half the 95th minus 5th percentile of the error
 in the measured action ($\Delta J_i$) from coordinate system errors for
@@
 Figure~\ref{fig:Jz_hist}. We therefore elect to measure this error by computing one
-half the 95\uth{} percentile minus the 5\uth{} percentile of the distribution
+half the 95th percentile minus the 5th percentile of the distribution
 of action values. We plot this quantity, which we refer to as $\Delta J_i$ for each action, in Figure~\ref{fig:Jz_hist} as a
@@
 0)\,\kpc$ and velocity $(0, -190, v_z)\,\kms$, where we vary $v_z$.\footref{note:vz_orbits} We show
-this for a $z$ offset of $10, 50, 100\,\pc$ (orange, teal, and green
-respectively). As before, the error ($\Delta J_z$) is one half the 95\uth{}
-minus 5\uth{} percentile of the distribution of $J_z$ values over the
-course of the orbit. There are large errors for \thin{} like orbits ($J_z
+this for a $z$ offset of $10, 50, 100\,\pc$ (orange, teal, and green,
+respectively). As before, the error ($\Delta J_z$) is one half the 95th
+minus 5th percentile of the distribution of $J_z$ values over the
+course of the orbit. There are large errors for \thin{}-like orbits ($J_z
 \sim 0.5\,\actunit$), even for a small midplane offset of $10\,\pc$. As dashed
@@
 
-\begin{deluxetable*}{cccccc}[htb!]
-\tablecaption{Stellar and gas disk scale heights of the real Milky Way and the
-simulated galaxies considered in this work. For comparison, we also give the median
-softening lengths for the simulated galaxies, computed for cold gas ($T<1000\,\text{K}$) and stars with $\abs{R-R_0} < 0.5\,\kpc$ and $\abs{z} < 1\,\kpc$.
+\begin{deluxetable*}{cCCCCC}[htb!]
+\tablecaption{Stellar and gas disk scale heights of the Milky Way and the
+FIRE galaxies considered in this work (described in Section~\ref{ssec:cosmozoom}). For comparison, we also give the median
+softening lengths for the FIRE galaxies, computed for cold gas ($T<1000\,\text{K}$) and stars with $\abs{R-R_0} < 0.5\,\kpc$ and $\abs{z} < 1\,\kpc$.
 We have assumed that $R_0=8.2\,\kpc$.\label{tab:scale_height}}
@@
 \end{center}
-\caption{The local midplane determined at the fiducial Solar radius
-($8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} (left,
+\caption{The local midplane determined at the fiducial solar circle
+($R_0 = 8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} (left,
 center, and right panels) as a function of azimuthal angle, at cosmological redshift $\z =0$. The local midplane is determined at
@@
 galactocentric coordinate system is incorrect, we subtract the best fit
-$A\sin{(\phi+B)}+C$ curve from each panel. We then bootstrap resample $1000$ times on
-all stars within a $2\,\kpc$ height of the fiducial midplane to determine
+$A\sin{(\phi+B)}+C$ curve from each panel. We then bootstrap resample $1000$ times to determine
 $1\,\sigma $ error bars, which we report as dashed lines.}
@@
 \end{center}
-\caption{The local midplane determined at the fiducial Solar radius
-($8.2\,\kpc$) for 4 different time steps from live N-body simulations of a
+\caption{The local midplane determined at the fiducial solar circle
+($R_0 = 8.2\,\kpc$) for four different time steps from live N-body simulations of a
 Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
@@
 for a typical thin (thick) disk orbit results in a $10\%$ error in $J_z$ (as
-defined by one half the middle $90\uth$ percentile). The fractional error is
+defined by one half the middle 90th percentile). The fractional error is
 significantly less for halo orbits.
@@
 left, center, and right columns, respectively. The upper row shows a plot of
-$x$ vs. $y$ while the lower row shows $R$ vs. $z$.}
+$x$~vs.~$y$ while the lower row shows $R$~vs.~$z$.}
 \label{fig:plot_orbits}

### 2019-05-11 paper/ms.tex
Remove runs directory

@@
 structure and evolution of the Galaxy. Stellar actions, computed from orbits,
-have the attractive quality of being invariant and thus good quantitative
-labels of a star's orbit. A systematic bias in the phase space position
-induces a phase-dependence in the actions, which we interpret as a systematic
-error distinct from typical observational errors and Galactic potential uncertainties.
-An offset in the vertical position of $\sim15\,\pc$ ($\sim120\,\pc$) for a thin (thick) disk orbit is sufficient to induce a
-$25\%$ systematic error in the vertical action $J_z$. The induced error distribution is
-non-Gaussian and bimodal, with neither of the modes peaking on the null value.
-Furthermore, we show that the local midplane varies by $\sim100\,\pc$ at the
-solar circle in Milky Way-mass cosmological zoom-in simulations from the FIRE
-project. From observations that the mean vertical velocity of the Milky Way
-disk varies by $\sim5\textup{--}10\,\kms$ with radius, we estimate that the true midplane
-variations are of order $\sim60\textup{--}170\,\pc$, with similar offsets seen in recent three-dimensional dust maps. Thus, current
-state-of-the-art action calculations, which assume the global and local
-midplanes are identical, are likely to include a systematic vertical offset
-depending on the volume considered. Variation in the local standard of rest
-induces similar issues. The variation of the midplane must be taken into
-account when performing dynamical analysis across the large regions of the
-disk accessible to \textit{Gaia} and future missions.
+have the attractive quality of being invariant and thus good quantitative labels
+of a star's orbit. A systematic bias in the phase space position induces a
+phase-dependence in the actions, which we interpret as a systematic error
+distinct from typical observational errors and Galactic potential uncertainties.
+An offset in the vertical position of $\sim15\,\pc$ ($\sim120\,\pc$) for a thin
+(thick) disk orbit is sufficient to induce a $25\%$ systematic error in the
+vertical action $J_z$. The induced error distribution is non-Gaussian and
+bimodal, with neither of the modes peaking on the null value. Furthermore, we
+show that the local midplane varies by $\sim100\,\pc$ at the solar circle in
+Milky Way-mass cosmological zoom-in simulations from the FIRE project. From
+observations that the mean vertical velocity of the Milky Way disk varies by
+$\sim5\textup{--}10\,\kms$ with radius, we estimate that the true midplane
+variations are of order $\sim60\textup{--}170\,\pc$, with similar offsets seen
+in recent three-dimensional dust maps. Thus, current state-of-the-art action
+calculations, which assume the global and local midplanes are identical, are
+likely to include a systematic vertical offset depending on the volume
+considered. Variation in the local standard of rest induces similar issues. The
+variation of the midplane must be taken into account when performing dynamical
+analysis across the large regions of the disk accessible to \textit{Gaia} and
+future missions.
 
@@
 
-\section{Introduction} \label{sec:intro}
-Our understanding of the Milky Way is currently undergoing a revolution as a
-result of \textit{Gaia} Data Release 2 (DR2). Recent major discoveries
-include the affirmation of remnants of a major merger \citep{2018ApJ...860L..11K, 2018MNRAS.478..611B, 2018Natur.563...85H, 2019MNRAS.486..378L, 2019MNRAS.482.3426M} hinted at in pre-\textit{Gaia} work \citep[e.g.,][]{2005MNRAS.359...93M, 2011MNRAS.412.1203N}, a phase-space
+\section{Introduction} \label{sec:intro} Our understanding of the Milky Way is
+currently undergoing a revolution as a result of \textit{Gaia} Data Release 2
+(DR2). Recent major discoveries include the affirmation of remnants of a major
+merger \citep{2018ApJ...860L..11K, 2018MNRAS.478..611B, 2018Natur.563...85H,
+2019MNRAS.486..378L, 2019MNRAS.482.3426M} hinted at in pre-\textit{Gaia} work
+\citep[e.g.,][]{2005MNRAS.359...93M, 2011MNRAS.412.1203N}, a phase-space
 ``spiral'' in the solar neighborhood \citep{2018Natur.561..360A} possibly
 indicating local substructure infall \citep{2018MNRAS.481.1501B,
-2019MNRAS.485.3134L}, and a gap
-suggestive of perturbation by a dark matter subtructure in the tidal stream
-GD1 \citep{2018ApJ...863L..20P, 2018arXiv181103631B}. These discoveries all
-indicate that the Milky Way's stellar distribution, which demonstrably departs
-significantly from axisymmetry, is undergoing phase mixing and dynamical
-interactions across a range of spatial and temporal scales.
-
-The assumption of a global, axisymmetric Galactocentric coordinate system \citep{2008gady.book.....B} underlies much of the quantitative analysis of the mechanisms that give rise to these signautres.
-In order to do construct such a system, the
-Sun's relative position must be
-defined and measured both precisely and accurately. This involves determining
-the angular position of and distance to the Galactic center, the orientation
-of the Galactic plane, the height of the Sun above the Galactic midplane, and
-the local standard of rest (LSR). We review and discuss the observational efforts to
-measure these parameters in Section~\ref{ssec:coord_off}.
+2019MNRAS.485.3134L}, and a gap suggestive of perturbation by a dark matter
+subtructure in the tidal stream GD1 \citep{2018ApJ...863L..20P,
+2018arXiv181103631B}. These discoveries all indicate that the Milky Way's
+stellar distribution, which demonstrably departs significantly from axisymmetry,
+is undergoing phase mixing and dynamical interactions across a range of spatial
+and temporal scales.
+
+The assumption of a global, axisymmetric Galactocentric coordinate system
+\citep{2008gady.book.....B} underlies much of the quantitative analysis of the
+mechanisms that give rise to these signautres. In order to do construct such a
+system, the Sun's relative position must be defined and measured both precisely
+and accurately. This involves determining the angular position of and distance
+to the Galactic center, the orientation of the Galactic plane, the height of the
+Sun above the Galactic midplane, and the local standard of rest (LSR). We review
+and discuss the observational efforts to measure these parameters in
+Section~\ref{ssec:coord_off}.
 
@@
 often desirable to convert this measurement into action space to concisely
-summarize its projected orbit, model the stellar distribution function, or
-find stars with similar dynamical properties. Actions are conserved quantities
-that describe the orbit of a star under the assumption of regular, bound
-orbits in a system where the equations of motion are separable in a particular
-coordinate system. They are the cyclical integral of the canonical momentum
-over its conjugate position:
-\beq\label{eq:actions}
-J_i \equiv
-\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
-\eeq
-where $p_i$ are the conjugate momenta.
-Under the assumption of axisymmetry, $i=R,z,\phi$ are the radial, vertical,
-and azimuthal coordinates respectively in a cylindrical coordinate system. In a slowly-evolving axisymmetric
-potential, these actions are invariant and $J_{\phi} \equiv L_z$, where $L_z$
-is the $z$-component of the angular momentum
-\citep{2008gady.book.....B,2014RvMP...86....1S}.
-
-With the advent of 6D phase-space measurements over a relatively large
-($\gtrsim 2$ kpc) volume from the \textit{Gaia} satellite, the study of stellar
-actions has gained new popularity. One reason is dimensionality reduction---an
-individual stellar orbit is concisely described by three actions, as opposed
-to six phase space coordinates. Second, under the assumption of a phase-mixed
+summarize its projected orbit, model the stellar distribution function, or find
+stars with similar dynamical properties. Actions are conserved quantities that
+describe the orbit of a star under the assumption of regular, bound orbits in a
+system where the equations of motion are separable in a particular coordinate
+system. They are the cyclical integral of the canonical momentum over its
+conjugate position: \beq\label{eq:actions} J_i \equiv \frac{1}{2\pi}
+\oint_{\text{orbit}}p_i\,\text{d}x_i\text{,} \eeq where $p_i$ are the conjugate
+momenta. Under the assumption of axisymmetry, $i=R,z,\phi$ are the radial,
+vertical, and azimuthal coordinates respectively in a cylindrical coordinate
+system. In a slowly-evolving axisymmetric potential, these actions are invariant
+and $J_{\phi} \equiv L_z$, where $L_z$ is the $z$-component of the angular
+momentum \citep{2008gady.book.....B,2014RvMP...86....1S}.
+
+With the advent of 6D phase-space measurements over a relatively large ($\gtrsim
+2$ kpc) volume from the \textit{Gaia} satellite, the study of stellar actions
+has gained new popularity. One reason is dimensionality reduction---an
+individual stellar orbit is concisely described by three actions, as opposed to
+six phase space coordinates. Second, under the assumption of a phase-mixed
 system, the dynamical properties of a population of stars should be uniquely a
-function of their actions and independent of the conjugate angles. This allows one to use actions to study the relationship between \emph{orbital} properties of stars
-and other intrinsic, and, at least partially, invariant properties such as age
-or metallicity \citep{2018ApJ...867...31B, 2018arXiv180803278T,
+function of their actions and independent of the conjugate angles. This allows
+one to use actions to study the relationship between \emph{orbital} properties
+of stars and other intrinsic, and, at least partially, invariant properties such
+as age or metallicity \citep{2018ApJ...867...31B, 2018arXiv180803278T,
 2018MNRAS.481.4093S, 2019arXiv190304030G, 2019arXiv190309320D,
 2019MNRAS.486.1167B}. Actions also provide a convenient basis for constructing
-models of the stellar distribution function
-\citep[e.g.,][]{1915MNRAS..76...70J, 1985ApJ...295..388V, 2017ApJ...839...61T},
-or for associating stars with similar dynamical properties, e.g., to
-potentially determine membership in moving groups.
-
-If the system being considered departs from axisymmetry in a significant
-and/or non-adiabatic way, the actions computed using an axisymmetric
-approximation to the true potential can exhibit cyclic dependence on the
-orbital phase (or time at which the star's position and velocity are
-observed), large-scale migration, or diffusion from their initial values. In
-the Milky Way, stellar actions are expected to diffuse on short time scales
-due to scattering with gas clouds and to evolve on longer time scales in the
-case of orbits near resonances with spiral arms, bar(s), and other large scale
-perturbations \citep{2014RvMP...86....1S}. For this reason, actions have been used to study stellar scattering in the Milky Way disk
-using the improved astrometry of \textit{Gaia} DR2 and various age catalogues
-\citep{2018ApJ...867...31B, 2018arXiv180803278T}. Actions have also been used
-to study different models of spiral structure in the Milky Way
-\citep{2019MNRAS.tmp..155S}. Characteristics of the distribution of stars in
-the extended solar neighborhood in action space are discussed in
-\citet{2019MNRAS.484.3291T}.
-
-The true Galactic potential is not strictly
-axisymmetric. This has been known for some time \citep[e.g.,][]{1957AJ.....62...93K,2009MNRAS.396L..56M, 2012ApJ...750L..41W}, but with the vast improvement in the quality of phase-space
-measurements due to \textit{Gaia} this assumption is increasingly
-inadequate \citep[e.g.,][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. Even if this assumption were
-close enough for many purposes, the parameters used in axisymmetric models of the Galactic potential may
-be inadequately constrained by current observations.
-
-The main point of this paper is as follows. Due to relatively small-scale variations in the density of gas and stars, the
-local midplane should vary as a function of radius and azimuth. For stars
-relatively far from the Sun, their local midplane will differ from our
-local midplane extrapolated onto their position.
-Converting positions and velocities of more distant stars from a
+models of the stellar distribution function \citep[e.g.,][]{1915MNRAS..76...70J,
+1985ApJ...295..388V, 2017ApJ...839...61T}, or for associating stars with similar
+dynamical properties, e.g., to potentially determine membership in moving
+groups.
+
+If the system being considered departs from axisymmetry in a significant and/or
+non-adiabatic way, the actions computed using an axisymmetric approximation to
+the true potential can exhibit cyclic dependence on the orbital phase (or time
+at which the star's position and velocity are observed), large-scale migration,
+or diffusion from their initial values. In the Milky Way, stellar actions are
+expected to diffuse on short time scales due to scattering with gas clouds and
+to evolve on longer time scales in the case of orbits near resonances with
+spiral arms, bar(s), and other large scale perturbations
+\citep{2014RvMP...86....1S}. For this reason, actions have been used to study
+stellar scattering in the Milky Way disk using the improved astrometry of
+\textit{Gaia} DR2 and various age catalogues \citep{2018ApJ...867...31B,
+2018arXiv180803278T}. Actions have also been used to study different models of
+spiral structure in the Milky Way \citep{2019MNRAS.tmp..155S}. Characteristics
+of the distribution of stars in the extended solar neighborhood in action space
+are discussed in \citet{2019MNRAS.484.3291T}.
+
+The true Galactic potential is not strictly axisymmetric. This has been known
+for some time \citep[e.g.,][]{1957AJ.....62...93K,2009MNRAS.396L..56M,
+2012ApJ...750L..41W}, but with the vast improvement in the quality of
+phase-space measurements due to \textit{Gaia} this assumption is increasingly
+inadequate \citep[e.g.,][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. Even if
+this assumption were close enough for many purposes, the parameters used in
+axisymmetric models of the Galactic potential may be inadequately constrained by
+current observations.
+
+The main point of this paper is as follows. Due to relatively small-scale
+variations in the density of gas and stars, the local midplane should vary as a
+function of radius and azimuth. For stars relatively far from the Sun, their
+local midplane will differ from our local midplane extrapolated onto their
+position. Converting positions and velocities of more distant stars from a
 heliocentric to a Galactocentric coordinate system thus introduces a systematic
-bias in the $z$ coordinate. We show that this bias induces non-Gaussian errors in the actions computed for these stars.
-The further from the solar neighborhood the target star is, the more likely the
-mismatch will result in large systematic uncertainty, especially in the
-vertical action $J_z$. A similar argument applies to any remaining uncertainty
-in measurements of the Galactocentric radius of the Sun and to variations in
-the LSR.
+bias in the $z$ coordinate. We show that this bias induces non-Gaussian errors
+in the actions computed for these stars. The further from the solar neighborhood
+the target star is, the more likely the mismatch will result in large systematic
+uncertainty, especially in the vertical action $J_z$. A similar argument applies
+to any remaining uncertainty in measurements of the Galactocentric radius of the
+Sun and to variations in the LSR.
 
@@
 \end{center}
-\caption{Cartoon approximation showing the effect an error in the
-determination of the coordinate midplane can have on orbit integration and
-action estimation. The $x$-axis shows the orbital phase and the $y$-axis the
-vertical height. The top gray curve depicts an example ``true'' orbit
-oscillating about the true midplane (horizontal solid gray line). Consider an
-observer who erroneously assumes the midplane is located at the horizontal
-dashed line. Suppose this observer measures the phase-space position of the
-star at two different orbital phases (teal and orange points). If the observer
-were to then integrate the star's orbit using a model potential with the
-erroneous midplane, they would obtain the teal and orange curves for the
-star's orbit, respectively. The actions estimated from these two erroneous
-measurements would subsequently differ, both from each other and from the true measurement (in the potential with the correct coordinate system).
-Hence an incorrect midplane in the potential model assumed will induce
-phase-dependence in the actions estimated for a given star in that potential.}
-\label{fig:cartoon}
-\end{figure*}
-
-In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
-system errors have on the measured actions. In
-Section~\ref{sec:local_fire} we examine the azimuthal variations of the
-midplane itself in examples from two classes of simulations: cosmological, hydrodynamical,
-zoom-in simulations of isolated Milky Way-mass galaxies from the Feedback in Realistic
-Environments (FIRE)
+\caption{Cartoon approximation showing the effect an error in the determination
+of the coordinate midplane can have on orbit integration and action estimation.
+The $x$-axis shows the orbital phase and the $y$-axis the vertical height. The
+top gray curve depicts an example ``true'' orbit oscillating about the true
+midplane (horizontal solid gray line). Consider an observer who erroneously
+assumes the midplane is located at the horizontal dashed line. Suppose this
+observer measures the phase-space position of the star at two different orbital
+phases (teal and orange points). If the observer were to then integrate the
+star's orbit using a model potential with the erroneous midplane, they would
+obtain the teal and orange curves for the star's orbit, respectively. The
+actions estimated from these two erroneous measurements would subsequently
+differ, both from each other and from the true measurement (in the potential
+with the correct coordinate system). Hence an incorrect midplane in the
+potential model assumed will induce phase-dependence in the actions estimated
+for a given star in that potential.} \label{fig:cartoon} \end{figure*}
+
+In Section~\ref{sec:ref_frame}, we describe the general impact coordinate system
+errors have on the measured actions. In Section~\ref{sec:local_fire} we examine
+the azimuthal variations of the midplane itself in examples from two classes of
+simulations: cosmological, hydrodynamical, zoom-in simulations of isolated Milky
+Way-mass galaxies from the Feedback in Realistic Environments (FIRE)
 collaboration\footnote{\url{https://fire.northwestern.edu}}
 \citep{2014MNRAS.445..581H, 2016ApJ...827L..23W, 2018MNRAS.480..800H}, and a
-controlled N-body simulation of a Sagittarius encounter with a galaxy
-otherwise tailored to the stellar mass, scale length, and scale height of the
-Milky Way \citep{2018MNRAS.481..286L}. In Section~\ref{sec:discussion}, we discuss the implications of midplane
-variations, and the resulting systematic uncertainty in the vertical action,
-for action-space analyses. We also estimate the expected midplane variations
-of the Milky Way based on the observed velocity variations and three-dimensional dust maps. We summarize our
-main results and conclude in Section~\ref{sec:conclusion}.
+controlled N-body simulation of a Sagittarius encounter with a galaxy otherwise
+tailored to the stellar mass, scale length, and scale height of the Milky Way
+\citep{2018MNRAS.481..286L}. In Section~\ref{sec:discussion}, we discuss the
+implications of midplane variations, and the resulting systematic uncertainty in
+the vertical action, for action-space analyses. We also estimate the expected
+midplane variations of the Milky Way based on the observed velocity variations
+and three-dimensional dust maps. We summarize our main results and conclude in
+Section~\ref{sec:conclusion}.
 
@@
 We first demonstrate the significance to action computations of a systematic
-offset in the determination of the Galactic midplane, distance to the
-Galactic center, or LSR. We will find that such offsets are especially important for
+offset in the determination of the Galactic midplane, distance to the Galactic
+center, or LSR. We will find that such offsets are especially important for
 disk-like orbits. The consequences we explore here may also arise from various
-other systematic errors. For instance, the axisymmetric Galactic potential
-model used in many works to compute actions may not be a good description of
-the true potential --- or the parameters used may yield a potential that is
+other systematic errors. For instance, the axisymmetric Galactic potential model
+used in many works to compute actions may not be a good description of the true
+potential --- or the parameters used may yield a potential that is
 systematically incorrect outside an original fitted region. In this work, we
 assume that the Galaxy is perfectly described by our model axisymmetric
-potential, and simply explore the consequences of offsets in the Galactocentric coordinate system.
+potential, and simply explore the consequences of offsets in the Galactocentric
+coordinate system.
 
 \subsection{Effect of Midplane Offset} \label{ssec:cartoon}
-We present a cartoon approximation of an orbit in Figure~\ref{fig:cartoon} to show how an
-inaccurate or erroneous determination of the midplane leads to a phase-dependence of the actions
-calculated from a phase-space point and an assumed
-potential model. The $x$-axis corresponds to orbital phase
-and the $y$-axis to vertical height. The solid gray curve indicates the
-``true'' orbit of the star; that is, as it oscillates around the
-true midplane.
+We present a cartoon approximation of an orbit in Figure~\ref{fig:cartoon} to
+show how an inaccurate or erroneous determination of the midplane leads to a
+phase-dependence of the actions calculated from a phase-space point and an
+assumed potential model. The $x$-axis corresponds to orbital phase and the
+$y$-axis to vertical height. The solid gray curve indicates the ``true'' orbit
+of the star; that is, as it oscillates around the true midplane.
 
 The dashed gray line, offset from the true midplane, is the midplane location
-used by an observer to integrate the orbit of the star and estimate its
-actions. The model potential is otherwise identical to the one in which the
-star is actually moving.
+used by an observer to integrate the orbit of the star and estimate its actions.
+The model potential is otherwise identical to the one in which the star is
+actually moving.
 
@@
 observer would compute for each point based on the potential model with the
-offset midplane. In action space, this would correspond to a different value
-of $J_z$ for the teal and orange points. In this way, assuming the wrong
-coordinate system induces a phase-dependence in the actions estimated for the
-star, which in the correct potential (in this example, the one with the
-correct midplane) should be phase-independent.
+offset midplane. In action space, this would correspond to a different value of
+$J_z$ for the teal and orange points. In this way, assuming the wrong coordinate
+system induces a phase-dependence in the actions estimated for the star, which
+in the correct potential (in this example, the one with the correct midplane)
+should be phase-independent.
 
@@
 from an offset in another coordinate, such as the distance to the Galactic
-center or the LSR. A similar effect in which actions gain
-time-dependence due to a time-varying potential was pointed out by
-\citet{2015A&A...584A.120B}.
+center or the LSR. A similar effect in which actions gain time-dependence due to
+a time-varying potential was pointed out by \citet{2015A&A...584A.120B}.
 
@@
 \begin{deluxetable*}{cCCCCCCCCC}
-\tablecaption{Description and names of the three orbits considered in this
-work. For the last four columns: $z_{\text{max}}$ is the maximum height of the
-orbit, $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of
-the radial excursions of the orbit, and $\kappa$ and $\nu$ are the radial/epicyclic and vertical frequencies of the orbit. In the epicyclic approximation, $A_z =
+
+\tablecaption{Description and names of the three orbits considered in this work.
+For the last four columns: $z_{\text{max}}$ is the maximum height of the orbit,
+$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of the radial
+excursions of the orbit, and $\kappa$ and $\nu$ are the radial/epicyclic and
+vertical frequencies of the orbit. In the epicyclic approximation, $A_z =
 z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$.
-\label{tab:orbits}}
-\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
-velocity} & \colhead{$J_R$} &
-\colhead{$J_{\phi}$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
-\colhead{$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$} & \colhead{$\kappa$} & \colhead{$\nu$} \\ \colhead{ } &
-\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km\,s}^{-1}$} &
-\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
+\label{tab:orbits}} \tablehead{\colhead{name} & \colhead{initial position} &
+\colhead{initial velocity} & \colhead{$J_R$} & \colhead{$J_{\phi}$} &
+\colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
+\colhead{$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$} & \colhead{$\kappa$} &
+\colhead{$\nu$} \\ \colhead{ } & \colhead{$\mathrm{kpc}$} &
+\colhead{$\mathrm{km\,s}^{-1}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
 \colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{Myr}^{-1}$} & \colhead{$\mathrm{Myr}^{-1}$}}
+\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$} &
+\colhead{$\mathrm{Myr}^{-1}$} & \colhead{$\mathrm{Myr}^{-1}$}}
 \startdata
-\thin{} & (8, 0, 0) & (0, -190, 10) & 40.3 & -1520 & 0.69 & 0.12 & 1.29 & 0.049 & 0.093 \\
-\thick{} & (8, 0, 0) & (0, -190, 50) & 32.5 & -1520 & 23.0 & 0.85 & 1.19 & 0.048 & 0.061 \\ 
-\halo{} & (8, 0, 0) & (0, -190, 190) & 32.8 & -1520 & 529.1 & 6.16 & 2.34 & 0.033 & 0.025
+\thin{} & (8, 0, 0) & (0, -190, 10) & 40.3 & -1520 & 0.69 & 0.12 & 1.29 & 0.049
+& 0.093 \\
+\thick{} & (8, 0, 0) & (0, -190, 50) & 32.5 & -1520 & 23.0 & 0.85 &
+1.19 & 0.048 & 0.061 \\
+\halo{} & (8, 0, 0) & (0, -190, 190) & 32.8 & -1520 &
+529.1 & 6.16 & 2.34 & 0.033 & 0.025
 \enddata

### 2019-05-11 runs/axisymmetric/actions/compute_actions.py

@@ -1,9 +1,0 @@
-from oceanic.analysis import snapshot_action_calculator
-from oceanic.options import options_reader
-import dill
-
-opt = options_reader('../options_m12i')
-#ac = snapshot_action_calculator(opt)
-ac = snapshot_action_calculator(opt, snapshot_file='cluster_snapshots.p')
-
-ac.all_actions(fileout='cluster_snapshots_actions.p')

### 2019-05-11 runs/axisymmetric/actions/run_actions.sh

@@ -1,34 +1,0 @@
-#!/bin/bash
-#SBATCH --job-name=fix
-#SBATCH --mail-type=END,FAIL
-#SBATCH --mail-user=8044322869@vtext.com
-#SBATCH -p cca 
-#SBATCH --mem=0
-#SBATCH -N1
-#SBATCH --exclusive
-#SBATCH --time=7-00:00:00
-#SBATCH --output=fix_%j.log
-pwd; hostname; date
-
-source /mnt/home/abeane/.bash_profile
-
-module load gcc
-module load cuda
-module load cudnn
-module load openmpi
-
-source /mnt/home/abeane/amuse_env/bin/activate
-
-echo $CUDA_VISIBLE_DEVICES
-
-export dir=`pwd`
-cd ../../
-source init.sh
-cd $dir 
-
-export OMP_NUM_THREADS=1 #this prevents conflicts in parallel libraries
-
-##mpirun -np 4 python axisymmetric_milkyway.py 
-python -u compute_actions.py 
-
-date

### 2019-05-11 runs/axisymmetric/analysis/gen_movie_acc_map.py

@@ -1,87 +1,0 @@
-from oceanic.gizmo_interface import gizmo_interface
-from oceanic.options import options_reader
-from oceanic.analysis import cluster_animator 
-from oceanic.analysis import acceleration_heatmap
-from oceanic.oceanic_io import load_interface
-import sys
-import dill
-
-options_file = sys.argv[1]
-cluster_file = sys.argv[2]
-
-cluster = dill.load(open(cluster_file, 'rb'))
-
-xcenter = 0.0
-ycenter = 0.0
-xmin = -0.1
-xmax = 0.1
-ymin = -0.1
-ymax = 0.1
-nres = 360
-zval = 0.0
-
-xaxis='x'
-yaxis='y'
-acc='x'
-
-start=None
-end=None
-
-cmap = 'bwr_r'
-cmin = -0.5
-cmax = 0.5
-
-# pLz_bound = 4.0
-# pJr_bound = 2.0
-# pJz_bound = 1.5
-pLz_bound = 20.0
-pJr_bound = 5.0
-pJz_bound = 5.0
-normalize=False
-
-color_by_dist = True
-log_distance = True
-dist_vmin = -1
-dist_vmax = 2
-
-axisymmetric = True
-
-plot_panel=True
-acc_map=False
-
-opt = options_reader(options_file)
-
-#interface = load_interface('../interface', skinny=True)
-interface = None
-
-def run():
-    an = cluster_animator(cluster, xaxis=xaxis, yaxis=yaxis, xmin=xmin, xmax=xmax,
-                ymin=ymin, ymax=ymax, start=start, end=end, acc_map=acc_map,
-                nres=nres, acc=acc, cmap=cmap, cmin=cmin, cmax=cmax, interface=interface,
-                options=options_file, plot_panel=plot_panel, pLz_bound=pLz_bound,
-                pJr_bound=pJr_bound, pJz_bound=pJz_bound, normalize=normalize,
-                color_by_dist=color_by_dist, log_distance=log_distance,
-                dist_vmin=dist_vmin, dist_vmax=dist_vmax, axisymmetric=axisymmetric)
-    an()
-
-if len(sys.argv) == 4:
-    key = int(sys.argv[3])
-    time = cluster[key]['time']
-    ac = acceleration_heatmap(options_file, interface)
-    ac(time, index=key, cache=True, return_heatmap=True,
[…]

### 2019-05-11 runs/axisymmetric/analysis/run_acc_map_cache.sh

@@ -1,29 +1,0 @@
-#!/bin/bash
-#SBATCH --job-name=acc
-#SBATCH --mail-type=END,FAIL
-#SBATCH --mail-user=8044322869@vtext.com
-#SBATCH -p ib   
-#SBATCH --oversubscribe
-#SBATCH --ntasks=1
-#SBATCH --cpus-per-task=1
-#SBATCH --mem=52000
-#SBATCH --time=02:00:00
-#SBATCH --output=log/acc_%j.log
-#SBATCH --array=0-1000
-pwd; hostname; date
-
-source /mnt/home/abeane/.bash_profile
-
-export dir=`pwd`
-cd ../
-source init.sh
-cd $dir
-
-export OMP_NUM_THREADS=1 #this prevents conflicts in parallel libraries
-
-##mpirun -np 4 python axisymmetric_milkyway.py 
-echo $SLURM_ARRAY_TASK_ID
-# /bin/time -f "-----\nreal %es\nuser %Us\nsys %Ss\nMaxRSS %Mk" python -u gen_movie_acc_map.py ../options ../cluster_snapshots.p $SLURM_ARRAY_TASK_ID
-python -u gen_movie_acc_map.py ../options ../cluster_snapshots.p $SLURM_ARRAY_TASK_ID
-
-date

### 2019-05-11 runs/axisymmetric/run_m12i.sh

@@ -1,27 +1,0 @@
-#!/bin/bash
-#SBATCH --job-name=amuse
-#SBATCH --mail-type=END,FAIL
-#SBATCH --mail-user=8044322869@vtext.com
-#SBATCH -p gpu
-#SBATCH --gres=gpu:1 -n 8 
-#SBATCH --ntasks-per-node=8
-#SBATCH --mem=150g
-#SBATCH --time=7-00:00:00
-#SBATCH --output=log/asmw_%j.log
-#SBATCH --constraint=v100
-pwd; hostname; date
-
-source /mnt/home/abeane/.bash_profile
-
-echo $CUDA_VISIBLE_DEVICES
-
-export dir=`pwd`
-cd ../
-source init.sh
-cd $dir
-
-export OMP_NUM_THREADS=1 #this prevents conflicts in parallel libraries
-
-python -u oc_nbody.py options_m12i 
-
-date

### 2019-05-11 runs/first_rotate/actions/compute_actions.py

@@ -1,15 +1,0 @@
-from oceanic.analysis import snapshot_action_calculator
-from oceanic.options import options_reader
-import dill
-
-opt = options_reader('../options_m12i')
-ac = snapshot_action_calculator(opt, snapshot_file='cluster_snapshots_m12i.p')
-ac.all_actions(fileout='cluster_snapshots_m12i_actions.p')
-
-opt = options_reader('../options_m12i_slow')
-ac = snapshot_action_calculator(opt, snapshot_file='cluster_snapshots_m12i_slow.p')
-ac.all_actions(fileout='cluster_snapshots_m12i_slow_actions.p')
-
-opt = options_reader('../options_m12i_fast')
-ac = snapshot_action_calculator(opt, snapshot_file='cluster_snapshots_m12i_fast.p')
-ac.all_actions(fileout='cluster_snapshots_m12i_fast_actions.p')

### 2019-05-11 runs/first_rotate/analysis/run_movie.sh

@@ -1,29 +1,0 @@
-#!/bin/bash
-#SBATCH --job-name=acc
-#SBATCH --mail-type=END,FAIL
-#SBATCH --mail-user=8044322869@vtext.com
-#SBATCH -p cca  
-#SBATCH -N1
-#SBATCH --exclusive
-#SBATCH --mem=0
-#SBATCH --time=04:00:00
-#SBATCH --output=log/movie_%j.log
-###SBATCH --array=0-1000
-pwd; hostname; date
-
-source /mnt/home/abeane/.bash_profile
-
-export dir=`pwd`
-cd ../../
-source init.sh
-cd $dir
-
-export OMP_NUM_THREADS=1 #this prevents conflicts in parallel libraries
-
-##mpirun -np 4 python axisymmetric_milkyway.py 
-echo $SLURM_ARRAY_TASK_ID
-# /bin/time -f "-----\nreal %es\nuser %Us\nsys %Ss\nMaxRSS %Mk" python -u gen_movie_acc_map.py ../options ../cluster_snapshots.p $SLURM_ARRAY_TASK_ID
-python -u gen_movie_acc_map.py ../options cluster_snapshots_actions.p $SLURM_ARRAY_TASK_ID
-#python -u gen_movie_acc_map.py ../options cluster_snapshots_fixedchosen_actions.p $SLURM_ARRAY_TASK_ID
-
-date

### 2019-05-11 runs/first_rotate/oc_nbody.py

@@ -1,1 +1,0 @@
-../oceanic/oc_nbody.py

### 2019-05-11 runs/first_rotate/pa_rotate/rotate_pa.py

@@ -1,7 +1,0 @@
-import numpy as np
-from scipy.optimize import minimize
-from tqdm import tqdm
-
-import utilities as ut
-
-gal_info = np.genfromtxt('../

### 2019-05-11 runs/init.sh

@@ -1,4 +1,0 @@
-module load gcc openmpi lib/x264/stable-aaa9aa8 ffmpeg
-module load cuda cudnn
-source ~/amuse_env/bin/activate
-export PYTHONPATH='/mnt/ceph/users/abeane/actions_systematic/runs/include':'/mnt/ceph/users/abeane/actions_systematic/runs'

### 2019-05-11 runs/oceanic/analysis.py

@@ -1,698 +1,0 @@
-import matplotlib as mpl; mpl.use('agg')
-
-import matplotlib.pyplot as plt
-from matplotlib import rc
-from matplotlib.animation import FuncAnimation
-from amuse.units import units
-
-import agama
-import gizmo_analysis as gizmo
-import numpy as np
-from tqdm import tqdm
-
-import sys
-import dill
-
-#from oceanic.gizmo_interface import gizmo_interface
-from oceanic.options import options_reader
-from amuse.units import units
-from tqdm import tqdm
-
-class agama_wrapper(object):
-    def __init__(self, opt):
-        opt.set_options(self)
-        agama.setUnits(mass=1, length=1, velocity=1)
-
-    def update_index(self, index, ss_id=None, snap=None):
-        agama.setUnits(mass=1, length=1, velocity=1)
-        self.current_index = index
-        self.ss_id = ss_id
-
-        if snap is not None:
-            self.snap = snap
-        else:
-            self.snap = \
-                gizmo.io.Read.read_snapshots(['star', 'gas', 'dark'],
-                                             'index', index,
-                                             properties=['id', 'position',
-                                                         'velocity', 'mass',
-                                                         'form.scalefactor'],
-                                             simulation_directory=
-                                             self.simulation_directory,
-                                             assign_center=False)
-        
-        if self.gal_info is not None:
-            try:
-                gal = np.genfromtxt(self.gal_info, comments='#', delimiter=',')
-            except:
-                gal = np.genfromtxt('../'+self.gal_info, comments='#', delimiter=',')
-            cen = gal[0]
-            vel = gal[1]
-            pa = gal[2:]
-            self.snap.center_position = cen
-            self.snap.center_velocity = vel
-            self.snap.principal_axes_vectors = pa
-            for k in self.snap.keys():
-                self.snap[k].center_position = cen
-                self.snap[k].center_velocity = vel
-                self.snap[k].principal_axes_vectors = pa
-        else:
-            gizmo.io.Read.assign_center(self.snap)
-            gizmo.io.Read.assign_principal_axes(self.snap, distance_max=20, age_percent=100)
-
-        if self.sim_name is None:
[…]

### 2019-05-11 runs/oceanic/oc_code.py

@@ -1,71 +1,0 @@
-import numpy as np
-from amuse.units import units, nbody_system
-from amuse.ic.kingmodel import new_king_model
-from oceanic.options import options_reader
-from amuse.lab import new_kroupa_mass_distribution
-
-
-class oc_code(object):
-    def __init__(self, options_reader):
-
-        options_reader.set_options(self)
-        # self.N, self.W0, self.Mcluster, self.Rcluster, self.ncpu
-        # self.gpu_enabled, self.ngpu, self.softening
-
-        if self.use_kroupa is True:
-            m = new_kroupa_mass_distribution(self.N,
-                                             self.kroupa_max | units.MSun)
-            tot_m = np.sum(m.value_in(units.MSun))
-            self.converter = \
-                nbody_system.nbody_to_si(tot_m | units.MSun,
-                                         self.Rcluster | units.parsec)
-
-            self.bodies = new_king_model(self.N, self.W0,
-                                         convert_nbody=self.converter)
-            self.bodies.mass = m
-            self.bodies.scale_to_standard(convert_nbody=self.converter)
-            self.bodies.move_to_center()
-        else:
-            self.converter = \
-                nbody_system.nbody_to_si(self.Mcluster | units.MSun,
-                                         self.Rcluster | units.parsec)
-
-            self.bodies = new_king_model(self.N, self.W0,
-                                         convert_nbody=self.converter)
-
-        if self.gpu_enabled:
-            self.code = self.nbodycode(self.converter,
-                                       number_of_workers=self.ngpu, mode='gpu')
-        else:
-            self.code = self.nbodycode(self.converter,
-                                       number_of_workers=self.ncpu)
-
-        parameters = {'epsilon_squared': (self.softening | units.parsec)**2}
-
-        for name, value in parameters.items():
-            setattr(self.code.parameters, name, value)
-        self.code.particles.add_particles(self.bodies)
-
-    def clean_ejections(self, system):
-        cluster_positions =\
-            np.transpose([system.particles.x.value_in(units.parsec),
-                          system.particles.y.value_in(units.parsec),
-                          system.particles.z.value_in(units.parsec)])
-
-        median_position = np.median(cluster_positions, axis=0)
-        dist = np.linalg.norm(np.subtract(cluster_positions, median_position),
-                              axis=1)
-        keys = np.where(dist > self.eject_cut)[0]
-        if len(keys > 1):
[…]

### 2019-05-11 runs/oceanic_schmactions/identify_ss.py

@@ -1,87 +1,0 @@
-import gizmo_analysis as gizmo
-import numpy as np 
-import agama
-
-# first read in snapshot
-
-np.random.seed(162)
-agama.setUnits(mass=1, length=1, velocity=1)
-
-snap_idx=600
-sim_dir = '/mnt/ceph/users/firesims/fire2/metaldiff/m12i_res7100/'
-gal_info = 'm12i_info.txt'
-
-Jr_min_list = [10.0, 40.0]
-Jr_max_list = [20.0, 60.0]
-Jz_min_list = [10.0, 40.0]
-Jz_max_list = [20.0, 60.0]
-
-Rmin = 7.7
-Rmax = 8.7
-zmax = 0.5
-
-age_min = 0.25
-age_max = 0.75
-
-snap = gizmo.io.Read.read_snapshots(['star', 'gas', 'dark'],
-                                     'index', snap_idx,
-                                     simulation_directory=sim_dir,
-                                     assign_center=False)
-
-ref = np.genfromtxt(gal_info, comments='#', delimiter=',')
-snap.center_position = ref[0]
-snap.center_velocity = ref[1]
-snap.principal_axes_vectors = ref[2:]
-for k in snap.keys():
-    for attr in ['center_position', 'center_velocity', 
-                 'principal_axes_vectors']:
-        setattr(snap[k], attr, getattr(snap, attr))
-
-star_fid_pos = snap['star'].prop('host.distance.principal')
-gas_fid_pos = snap['gas'].prop('host.distance.principal')
-dark_fid_pos = snap['dark'].prop('host.distance.principal')
-
-star_fid_vel = snap['star'].prop('host.velocity.principal')
-
-star_mass = snap['star']['mass']
-gas_mass = snap['gas']['mass']
-dark_mass = snap['dark']['mass']
-
-star_age = snap['star'].prop('age')
-
-position = np.concatenate((star_fid_pos, gas_fid_pos))
-mass = np.concatenate((star_mass, gas_mass))
-
-pdark = agama.Potential(type="Multipole",
-                        particles=(dark_fid_pos, dark_mass),
-                        symmetry='a', gridsizeR=20, lmax=2)
-pbar = agama.Potential(type="CylSpline",
-                       particles=(position, mass),
-                       gridsizer=20, gridsizez=20,
[…]

### 2019-05-11 runs/oceanic_schmactions/plot_actions.py

@@ -1,46 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-import matplotlib as mpl
-from matplotlib import rc
-rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
-rc('text', usetex=True)
-mpl.rcParams['text.latex.preamble'] = [r'\usepackage{amsmath}']
-
-tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
-        '#edc948', '#b07aa1', '#ff9da7', '#9c755f', '#bab0ac']
-
-tend = 2000 # Myr
-dt = 0.1 # Myr
-tlist = np.arange(0 , tend, dt)
-
-cadence = 20
-tlist = tlist[::cadence]
-
-tmax = 2000
-
-#idx_list = ['0', '1']
-idx_list = ['0']
-for idx in idx_list:
-    fig, ax = plt.subplots(1, 3, figsize=(7, 2.25))
-    fiducial_actions = np.load('fiducial_actions_'+idx+'.npy')
-    oa_actions = np.load('oa_actions_'+idx+'.npy')
-    
-    ax[0].plot(tlist, fiducial_actions[:,0], c=tb_c[-1], ls='dashed')
-    ax[1].plot(tlist, fiducial_actions[:,2], c=tb_c[-1], ls='dashed')
-    ax[2].plot(tlist, fiducial_actions[:,1], c=tb_c[-1], ls='dashed')
-    ax[0].plot(tlist, oa_actions[:,0], c=tb_c[0])
-    ax[1].plot(tlist, oa_actions[:,2], c=tb_c[0])
-    ax[2].plot(tlist, oa_actions[:,1], c=tb_c[0])
-    
-    for x in ax:
-        x.set_xlim(0, tmax)
-        x.set_xlabel(r'$t\,[\,\text{Myr}\,]$')
-    
-    ax[0].set_ylabel(r'$J_r\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
-    ax[1].set_ylabel(r'$L_z\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
-    ax[2].set_ylabel(r'$J_z\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
-
-    fig.tight_layout()
-    fig.savefig('actions_time_'+idx+'.pdf')
-

### 2019-05-11 runs/oceanic_schmactions/rotate_pa.py

@@ -1,151 +1,0 @@
-import gizmo_analysis as gizmo
-import numpy as np
-import agama
-import sys
-import utilities as ut
-from scipy.optimize import minimize
-
-np.random.seed(162)
-agama.setUnits(mass=1, length=1, velocity=1)
-
-snap_idx=600
-sim_dir = '/mnt/ceph/users/firesims/fire2/metaldiff/m12i_res7100/'
-gal_info = 'm12i_info.txt'
-
-cadence_list = [50, 20]
-#cadence_list = [50]
-
-tend = 2000 # Myr
-dt = 0.1 # Myr
-tlist = np.arange(0 , tend, dt)
-
-snap = gizmo.io.Read.read_snapshots(['star', 'gas', 'dark'],
-                                     'index', snap_idx,
-                                     simulation_directory=sim_dir,
-                                     assign_center=False)
-
-ref = np.genfromtxt(gal_info, comments='#', delimiter=',')
-snap.center_position = ref[0]
-snap.center_velocity = ref[1]
-snap.principal_axes_vectors = ref[2:]
-for k in snap.keys():
-    for attr in ['center_position', 'center_velocity', 
-                 'principal_axes_vectors']:
-        setattr(snap[k], attr, getattr(snap, attr))
-
-fiducial_center = ref[0]
-fiducial_velocity = ref[1]
-fiducial_pa = ref[2:]
-
-def get_position(snap, pos, center, pa=None):
-    values = ut.coordinate.get_distances(
-                        pos, center,
-                        snap.info['box.length'], snap.snapshot['scalefactor'])
-    if pa is not None:
-        values = ut.coordinate.get_coordinates_rotated(
-                    values, pa)
-    return values
-
-def get_velocity(snap, pos, vel, center, center_velocity, pa=None):
-    values = ut.coordinate.get_velocity_differences(
-                        vel, center_velocity,
-                        pos, center,
-                        snap.info['box.length'], snap.snapshot['scalefactor'],
-                        snap.snapshot['time.hubble'])
-    if pa is not None:
-        values = ut.coordinate.get_coordinates_rotated(
-                    values, pa)
-    return values
-
-def euler_matrix(theta) :
-    theta = np.array(theta)
-    if len(theta)==2:
-        theta = np.hstack((theta, 0))
[…]

### 2019-05-11 runs/oceanic_schmactions/run_amuse.py

@@ -1,205 +1,0 @@
-import gizmo_analysis as gizmo
-import numpy as np 
-import agama
-import utilities as ut
-import sys
-from pykdgrav import ConstructKDTree, GetAccelParallel
-from astropy.constants import G as G_astropy
-
-from amuse.units import units, nbody_system
-from amuse.lab import ph4, new_kroupa_mass_distribution
-from amuse.ic.kingmodel import new_king_model
-from amuse.couple.bridge import Bridge
-
-from tqdm import tqdm
-
-from astropy.constants import G as G_astropy
-import astropy.units as u
-
-np.random.seed(162)
-agama.setUnits(mass=1, length=1, velocity=1)
-
-snap_idx=600
-sim_dir = '/mnt/ceph/users/firesims/fire2/metaldiff/m12i_res7100/'
-gal_info = 'm12i_info.txt'
-
-star_char_mass = 0.048
-dark_softening_in_pc = 112.0
-Rmax = 50.0
-
-tend = 2000 # Myr
-dt = 0.1 # Myr
-tlist = np.arange(0 , tend, dt)
-
-ss_ids = [23693026,
-          17012804]
-ss_id = ss_ids[int(sys.argv[1])]
-
-G = G_astropy.to_value(u.kpc**2 * u.km / (u.s * u.Myr * u.Msun))
-theta = 0.5
-
-# cluster parameter
-N = 256
-kroupa_max = 5 | units.MSun
-Rcluster = 0.8 | units.parsec
-W0 = 3
-softening = 0.01 | units.parsec
-
-
-def get_position(snap, pos, center, pa=None):
-    values = ut.coordinate.get_distances(
-                        pos, center,
-                        snap.info['box.length'], snap.snapshot['scalefactor'])
-    if pa is not None:
-        values = ut.coordinate.get_coordinates_rotated(
-                    values, pa)
-    return values
-
-def get_velocity(snap, pos, vel, center, center_velocity, pa=None):
-    values = ut.coordinate.get_velocity_differences(
-                        vel, center_velocity,
-                        pos, center,
-                        snap.info['box.length'], snap.snapshot['scalefactor'],
-                        snap.snapshot['time.hubble'])
-    if pa is not None:
-        values = ut.coordinate.get_coordinates_rotated(
-                    values, pa)
-    return values
-
-class pykdgrav(object):
-    def __init__(self, r, m, soft, theta, G):
-        print('constructing kdtree')
[…]

### 2019-05-11 runs/oceanic_schmactions/run_amuse.sh

@@ -1,27 +1,0 @@
-#!/bin/bash
-#SBATCH --job-name=amuse
-#SBATCH --mail-type=END,FAIL
-#SBATCH --mail-user=8044322869@vtext.com
-#SBATCH -p gpu
-#SBATCH --gres=gpu:1 -n 8 
-#SBATCH --ntasks-per-node=8
-#SBATCH --mem=150g
-#SBATCH --time=7-00:00:00
-#SBATCH --output=log/asmw_%j.log
-#SBATCH --constraint=v100
-pwd; hostname; date
-
-source /mnt/home/abeane/.bash_profile
-
-echo $CUDA_VISIBLE_DEVICES
-
-export dir=`pwd`
-cd ../
-source init.sh
-cd $dir
-
-export OMP_NUM_THREADS=1 #this prevents conflicts in parallel libraries
-
-python -u run_amuse.py 0
-
-date

### 2019-05-11 runs/schmactions/pa_rotate/rotate_pa.py

@@ -1,81 +1,0 @@
-import numpy as np
-from scipy.optimize import minimize
-from tqdm import tqdm
-
-from oceanic.analysis import snapshot_action_calculator
-from oceanic.options import options_reader
-
-import dill
-import utilities as ut
-
-from scipy.optimize import minimize
-
-gal_info = np.genfromtxt('../m12i_info.txt', comments='#', delimiter=',')
-
-cen = gal_info[0]
-vel = gal_info[1]
-pa = gal_info[2:]
-
-def euler_matrix(theta) :
-    theta = np.array(theta)
-    if len(theta)==2:
-        theta = np.hstack((theta, 0))
-    R_x = np.array([[1,         0,                  0  ] ,                
-                    [0,         np.cos(theta[0]), -np.sin(theta[0]) ],
-                    [0,         np.sin(theta[0]), np.cos(theta[0])  ]
-                    ])
-    R_y = np.array([[np.cos(theta[1]),    0,      np.sin(theta[1])  ],
-                    [0,                     1,      0 ] ,                
-                    [-np.sin(theta[1]),   0,      np.cos(theta[1])  ]
-                    ])
-    R_z = np.array([[np.cos(theta[2]),    -np.sin(theta[2]),    0],
-                    [np.sin(theta[2]),    np.cos(theta[2]),     0],
-                    [0,                     0,                      1]
-                    ])
-    R = np.dot(R_x, np.dot(R_y, R_z))
-    
-    # this is the real one
-    # R = np.dot(R_z, np.dot( R_y, R_x ))
-    return R
-
-def euler_rotate(theta, vec):
-    mat = euler_matrix(theta)
-    return np.transpose(np.tensordot(mat, np.transpose(vec), axes=1))
-
-opt = options_reader('../options_m12i')
-ac = snapshot_action_calculator(opt, snapshot_file='cluster_snapshots_m12i.p')
-
-cluster = dill.load(open('cluster_snapshots_m12i.p', 'rb'))
-npart = len(cluster[0]['position'])
-nsnap = len(cluster)
-
-pos = np.array([cl['position'] for cl in cluster])
-vel = np.array([cl['velocity'] for cl in cluster])
-time = np.array([cl['time'] for cl in cluster])
-
-pos_flat = np.reshape(pos, (npart*nsnap, 3))
-vel_flat = np.reshape(vel, (npart*nsnap, 3))
[…]

### 2019-05-11 runs/schmactions/run_m12i.sh

@@ -1,33 +1,0 @@
-#!/bin/bash
-#SBATCH --job-name=amuse
-#SBATCH --mail-type=END,FAIL
-#SBATCH --mail-user=8044322869@vtext.com
-#SBATCH -p gpu
-#SBATCH --gres=gpu:1 -n 8 
-#SBATCH --ntasks-per-node=8
-#SBATCH --mem=150g
-
-###SBATCH -p cca
-###SBATCH -N1
-###SBATCH --ntasks-per-node=40
-###SBATCH --mem=0
-###SBATCH --constraint=skylake
-
-#SBATCH --time=7-00:00:00
-#SBATCH --output=log/asmw_%j.log
-pwd; hostname; date
-
-source /mnt/home/abeane/.bash_profile
-
-echo $CUDA_VISIBLE_DEVICES
-
-export dir=`pwd`
-cd ../
-source init.sh
-cd $dir
-
-export OMP_NUM_THREADS=1 #this prevents conflicts in parallel libraries
-
-python -u oc_nbody.py options_m12i 
-
-date

### 2019-05-11 runs/schmactions/run_m12i_fast.sh

@@ -1,27 +1,0 @@
-#!/bin/bash
-#SBATCH --job-name=amuse
-#SBATCH --mail-type=END,FAIL
-#SBATCH --mail-user=8044322869@vtext.com
-#SBATCH -p gpu
-#SBATCH --gres=gpu:1 -n 8 
-#SBATCH --ntasks-per-node=8
-#SBATCH --mem=150g
-#SBATCH --time=7-00:00:00
-#SBATCH --output=log/asmw_%j.log
-#SBATCH --constraint=v100
-pwd; hostname; date
-
-source /mnt/home/abeane/.bash_profile
-
-echo $CUDA_VISIBLE_DEVICES
-
-export dir=`pwd`
-cd ../
-source init.sh
-cd $dir
-
-export OMP_NUM_THREADS=1 #this prevents conflicts in parallel libraries
-
-python -u oc_nbody.py options_m12i_fast
-
-date

### 2019-05-11 runs/schmactions/run_m12i_long.sh

@@ -1,33 +1,0 @@
-#!/bin/bash
-#SBATCH --job-name=amuse
-#SBATCH --mail-type=END,FAIL
-#SBATCH --mail-user=8044322869@vtext.com
-#SBATCH -p gpu
-#SBATCH --gres=gpu:1 -n 8 
-#SBATCH --ntasks-per-node=8
-#SBATCH --mem=150g
-
-###SBATCH -p cca
-###SBATCH -N1
-###SBATCH --ntasks-per-node=40
-###SBATCH --mem=0
-###SBATCH --constraint=skylake
-
-#SBATCH --time=7-00:00:00
-#SBATCH --output=log/asmw_%j.log
-pwd; hostname; date
-
-source /mnt/home/abeane/.bash_profile
-
-echo $CUDA_VISIBLE_DEVICES
-
-export dir=`pwd`
-cd ../
-source init.sh
-cd $dir
-
-export OMP_NUM_THREADS=1 #this prevents conflicts in parallel libraries
-
-python -u oc_nbody.py options_m12i_long
-
-date

### 2019-05-11 runs/schmactions/run_m12i_slow.sh

@@ -1,27 +1,0 @@
-#!/bin/bash
-#SBATCH --job-name=amuse
-#SBATCH --mail-type=END,FAIL
-#SBATCH --mail-user=8044322869@vtext.com
-#SBATCH -p gpu
-#SBATCH --gres=gpu:1 -n 8 
-#SBATCH --ntasks-per-node=8
-#SBATCH --mem=150g
-#SBATCH --time=7-00:00:00
-#SBATCH --output=log/asmw_%j.log
-#SBATCH --constraint=v100
-pwd; hostname; date
-
-source /mnt/home/abeane/.bash_profile
-
-echo $CUDA_VISIBLE_DEVICES
-
-export dir=`pwd`
-cd ../
-source init.sh
-cd $dir
-
-export OMP_NUM_THREADS=1 #this prevents conflicts in parallel libraries
-
-python -u oc_nbody.py options_m12i_slow 
-
-date

### 2019-05-12 paper/ms.tex
Add CCA affil.
Improve Fig 2

@@
 \affil{\columbia}
+\affil{\cca}
 
@@
 to measure the orbit over $1\,\Gyr$ and perform the same orbit integration
-procedure at each timestep, and report the values of the actions. The
+procedure at each timestep, and report the values of the actions, with the true values given as horizontal dashed lines. The
 computation of $J_{\phi}$ is pristine to errors in $z$, with only numerical

### 2019-05-12 schmactions/one_orbit/plot_one_orbit.py

@@ -25,13 +25,6 @@ tb_c = ['#4e79a7', '#f28e2b', '#e15759', '#76b7b2', '#59a14f',
 xmin = 0
 xmax = 1000
 
-y0min = 20
-y0max = 40
-y1min = -1800
-y1max = -1000
-y2min = 10
-y2max = 30
-
 histmin = 15
 histmax = 30
 
@@ -64,6 +57,13 @@ init_pos = [8, 0, 0] * u.kpc
 init_vel = [0, -190, 50] * u.km/u.s
 init_vel_thin = [0, -190, 10] * u.km/u.s
 
+y0min = J0 - 9
+y0max = J0 + 9
+y1min = J1 - 50
+y1max = J1 + 50
+y2min = J2 - 9
+y2max = J2 + 9
+
 s = schmactions(init_pos, init_vel)
 sthin = schmactions(init_pos, init_vel_thin)
 
@@ -82,15 +82,18 @@ xtime_thin = sthin.extract_time(xout_thin)
 fig, ax = plt.subplots(2, 3, figsize=(textwidth, 4), sharex=True)
 
 for x,t,a in zip(ax, (ztime, xtime), (zact, xact)):
-    # get keys corresponding to 4 sigmaclip
-    k0, k1, k2 = sclip(a)
-    x[0].plot(t[k0], a[:,0][k0], c=tb_c[0])
-    x[1].plot(t[k1], a[:,1][k1], c=tb_c[0])
-    x[2].plot(t[k2], a[:,2][k2], c=tb_c[0])
+    # get keys corresponding to 2 sigmaclip
+    k0, k1, k2 = sclip(a, s=4)
 
-    x[0].plot(t, np.full(len(t), J0), c=tb_c[0], ls='dashed')
-    x[1].plot(t, np.full(len(t), J1), c=tb_c[0], ls='dashed')
-    x[2].plot(t, np.full(len(t), J2), c=tb_c[0], ls='dashed')
+    print('num excluded = ', len(t) - len(k0), len(t) - len(k1), len(t) - len(k2))
+
+    x[0].plot(t, np.full(len(t), J0), c=tb_c[-1], ls='dashed')
+    x[1].plot(t, np.full(len(t), J1), c=tb_c[-1], ls='dashed')
+    x[2].plot(t, np.full(len(t), J2), c=tb_c[-1], ls='dashed')
+
[…]

### 2019-05-13 paper/ms.tex
Swapped Melissa's affil
revisions to abstract, intro (Sect 1)

@@
 \author[0000-0003-0064-4060]{Mordecai-Mark Mac Low}
+\affil{\amnh}
 \affil{\cca}
-\affil{\amnh}
 
@@
 have the attractive quality of being invariant and thus good quantitative labels
-of a star's orbit. A systematic bias in the phase space position induces a
-phase-dependence in the actions, which we interpret as a systematic error
+of a star's orbit. A systematic bias in the phase space position
+induces 
+%mm a phase-dependence [to better distinguish the two different
+%meanings of phase in this sentence!]
+   an orbital phase dependence
+in the actions, which we interpret as a systematic error
 distinct from typical observational errors and Galactic potential uncertainties.
-An offset in the vertical position of $\sim15\,\pc$ ($\sim120\,\pc$) for a thin
-(thick) disk orbit is sufficient to induce a $25\%$ systematic error in the
+An offset in the vertical position of 
+%mm
+     the Galactic midplane of
+$\sim15\,\pc$ 
+%mm [this formulation isn't worth the extra difficulty reading]
+%($\sim120\,\pc$) for a thin (thick) disk 
+  for a thin disk orbit or $\sim 120\,\pc$ for a thick disk
+orbit is sufficient to induce a $25\%$ systematic error in the
 vertical action $J_z$. The induced error distribution is non-Gaussian and
@@
 \citep{2008gady.book.....B} underlies much of the quantitative analysis of the
-mechanisms that give rise to these signautres. In order to do construct such a
+mechanisms that give rise to these 
+%mm signautres.
+    signatures. 
+In order to 
+%mm do 
+construct such a
 system, the Sun's relative position must be defined and measured both precisely
 and accurately. This involves determining the angular position of and distance
-to the Galactic center, the orientation of the Galactic plane, the height of the
-Sun above the Galactic midplane, and the local standard of rest (LSR). We review
+to the Galactic center, the orientation of the Galactic 
+%mm plane, the height of the Sun above the Galactic midplane, 
+   midplane, the distance of the Sun from it,
+and the local standard of rest (LSR). We review
 and discuss the observational efforts to measure these parameters in
@@
 2012ApJ...750L..41W}, but with the vast improvement in the quality of
-phase-space measurements due to \textit{Gaia} this assumption is increasingly
-inadequate \citep[e.g.,][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. Even if
+phase-space measurements due to \textit{Gaia} 
+%mm this assumption 
+     the assumption of axisymmetry
+is increasingly inadequate \citep[e.g.,][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. Even if
 this assumption were close enough for many purposes, the parameters used in
@@
 
-The main point of this paper is as follows. Due to relatively small-scale
-variations in the density of gas and stars, the local midplane should vary as a
-function of radius and azimuth. For stars relatively far from the Sun, their
-local midplane will differ from our local midplane extrapolated onto their
-position. Converting positions and velocities of more distant stars from a
+%mm The main point of this paper is as follows. Due to relatively
+%small-scale variations in the density of gas and stars, the local
+%midplane should vary as a
+      Small-scale variations in the density of gas and stars cause the
+      local midplane position to vary as a
+function of radius and azimuth. 
+%mm For stars relatively far from the Sun, their local midplane will
+%differ 
+    Stars far from the Sun have a local midplane that differs
+from our local midplane extrapolated onto their position. Converting
+positions and velocities of more distant stars from a 
 heliocentric to a Galactocentric coordinate system thus introduces a systematic

### 2019-05-14 paper/ms.tex
work on sect 2 into 2.3

@@
 
+%mm [the x-axis of this figure should be labeled orbital phase for
+%consistency with the text]
 \begin{figure*}[ht!]
@@
 show how an inaccurate or erroneous determination of the midplane leads to a
-phase-dependence of the actions calculated from a phase-space point and an
+%mm phase-dependence of the actions calculated from a phase-space
+%point and an
+   dependence on orbital phase of the value of the actions calculated
+   from a point in phase space and an
 assumed potential model. The $x$-axis corresponds to orbital phase and the
-$y$-axis to vertical height. The solid gray curve indicates the ``true'' orbit
-of the star; that is, as it oscillates around the true midplane.
+$y$-axis to vertical height. The solid gray curve indicates the 
+%mm ``true'' orbit of the star; that is, as it oscillates around the
+%true midplane.
+     true orbit of the star as it oscillates around the true midplane.
 
@@
 
-This example uses an offset in $z$, but one would expect an analogous effect
-from an offset in another coordinate, such as the distance to the Galactic
+This example uses an offset in $z$, but one 
+%mm would expect an analogous effect from an offset in another coordinate, 
+    analagous effects occur from offsets in other coordinates, 
+such as the distance to the Galactic
 center or the LSR. A similar effect in which actions gain time-dependence due to
@@
 systematic error in the actions induced from offsets in the position (e.g., the midplane or
-Galactic center distance) or velocity (e.g., the local standard of rest) of the assumed coordinate system's origin. We use the epicyclic approximation, which
+Galactic center distance) or velocity (e.g., the 
+%mm local standard of rest)
+     LSR) 
+of the assumed coordinate system's origin. We use the epicyclic approximation, which
 assumes that the motion in the $z$ and $R$ components of the orbit are
 decoupled and follow simple harmonic motion about a circular and planar
-``guiding orbit'' \citep[][Section~3.2 and references
-therein]{2008gady.book.....B}. We refer to the radius of this orbit (the
-``guiding radius'') as $R_g$. This approximation is an excellent description
+%mm ``guiding orbit''
+     guiding orbit 
+\citep[][Section~3.2 and references
+therein]{2008gady.book.....B}. We refer to the radius of this orbit
+%mm  (the``guiding radius'') as $R_g$. 
+     as the guiding radius $R_g$.
+This approximation is an excellent description
 of the thin disk and a good description of the thick disk in
-an axisymmetric potential, i.e., ignoring the influence of the Galactic bar and
+an axisymmetric potential
+%mm , i.e., ignoring 
+    that ignores
+the influence of the Galactic bar and
 spiral arms. We also make the assumption of a perfectly flat circular
@@
 \equiv v_c/R_g$ is the orbital frequency of the guiding center, $A_z$ and $A_R$ are the amplitudes in the vertical and radial coordinates, and $\alpha$ and
-$\beta$ are arbitrary phases. Similarly, the velocities of the orbit are
-given by:
+$\beta$ are 
+%mm arbitrary phases
+    the initial orbital phases at $t = 0$.
+Similarly, the velocities of the orbit are given by:
 \beq\label{eq:orbits_vel_epi}
@@
 
-In this case, we have that for $J_{\phi}$
+In this case, 
+%mm we have that for $J_{\phi}$
+    the azimuthal action is
 \citep[][Section~3.5.3b]{2008gady.book.....B}:
@@
 \eeq
-and for $J_R$ and $J_z$:
+and
+%mm for $J_R$ and $J_z$:
+    the radial and vertical actions are
 \beq\label{eq:JR_Jz_epi}
 \begin{split}
-J_R &= \frac{E_R}{\kappa} \\
-J_z &= \frac{E_z}{\nu} \text{,}
+J_R &= %mm \frac{E_R}{\kappa} \\
+            E_R/\kappa \\
+J_z &= %mm \frac{E_z}{\nu}  \text{,}
+            E_z / \nu, 
 \end{split}
@@
 potential available in \texttt{galpy} \citep{2015ApJS..216...29B}, which includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
-a Miyamoto--Nagai disk \citep{1975PASJ...27..533M}, and an NFW halo
-\citep{1997ApJ...490..493N}, and is fit to empirically match some
+a Miyamoto--Nagai disk \citep{1975PASJ...27..533M}, and 
+%mm an NFW halo \citep{1997ApJ...490..493N},
+    a Navarro, Frenk, \& White \citeyear{1997ApJ...490..493N} halo,
+and is fit to empirically match some
 observations.

### 2019-05-14 paper/ms.tex
up to Fig 4 discussion

@@
 \newcommand{\Gyr}{\text{Gyr}}
-\newcommand{\kms}{\text{km}/\text{s}}
+%mm  \newcommand{\kms}{\text{km}/\text{s}}
+         \newcommand{\kms}{\text{km} \text{s}^{-1}}
 \newcommand{\actunit}{\text{kpc}\,\kms}
@@
 $J_z$ is decidedly non-Gaussian. Therefore, any sort of error propagation must
-take this into account. Intuition for the shape of each panel is given in the
+take this into account. 
+%mm Intuition 
+    A heuristic explanation
+for the shape of each panel is given in the
 text. We also plot one half the 95th percentile minus the 5th
@@
 We consider the observer making a measurement, integrating an orbit, and
-computing actions every $\Myr$ using the prescription above. However, we
+computing actions every 
+%mm $\Myr$
+    megayear 
+using the prescription above. However, we
 specify the star's starting position using the systematically offset
@@
 The actions computed using the offset coordinate system
-for each phase-space starting point are shown for the first $\Gyr$ of the
-orbit in the upper panels of
+for each phase-space starting point are shown for the first 
+%mm $\Gyr$
+  gigayear 
+of the orbit in the upper panels of
 Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
@@
 Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
-offset coordinate systems oscillates as a function of the time/orbital phase at which the
+offset coordinate systems oscillate
+%mm s 
+as a function of the time/orbital phase at which the
 star's phase-space position is observed.
 This time dependence comes even though
-the observer is using the correctly constructed best-fit static, axisymmetric
+the observer is using the correctly constructed, best-fit, static, axisymmetric
 potential. The relative size of the phase variation in each action depends on
@@
 
-We now describe the intuition behind the shape of Figure~\ref{fig:Jz_hist}.
+We now 
+%mm describe the intuition behind 
+    suggest a heuristic explanation for
+the shape of Figure~\ref{fig:Jz_hist}.
 Consider first the \thick{} (top panel), where the offset in $z$ is
@@
 half the 95th percentile minus the 5th percentile of the distribution
-of action values. We plot this quantity, which we refer to as $\Delta J_i$ for each action, in Figure~\ref{fig:Jz_hist} as a
-horizontal arrow anchored on the true action value. Because of the bimodality
+of action values. We plot this quantity
+%mm
+     in Figure~\ref{fig:Jz_hist}, 
+which we refer to as $\Delta J_i$ for each action, 
+%mm in Figure~\ref{fig:Jz_hist} 
+as a horizontal arrow anchored on the true action value. Because of the bimodality
 of the error distribution, this quantity roughly measures the distance from

### 2019-05-15 paper/ms.tex
finished section 2

@@
 \newcommand{\Gus}[1]{\textcolor{red}{#1}}
+\newcommand{\MM}[1]{\textcolor{blue}{#1}}   %mm  Mordecai
 
@@
 fractional error in $J_z$ is more exaggerated for more planar (disk-like)
-orbits. For \thin{} (\thick{}), a systematic offset of $15\,\pc$ ($120\,\pc$) in the $z$-coordinate results in a $25\%$ deviation in $J_z$.
+orbits. For \thin{}, 
+%mm (\thick{}), 
+a systematic offset of $15\,\pc$
+%mm ($120\,\pc$)
+in the $z$-coordinate results in a $25\%$ deviation in $J_z$,
+%mm
+     while a $120\,\pc$ offset results in the same deviation for \thick{}.
 We find that \halo{} is relatively resistant to errors in the midplane, with
@@
 v_{\phi}=0$, as dashed lines in the color of each orbit. This equation is
-relevant since during most of the orbit the star will be at maximum radial and
+relevant since during most of the orbit the star will be 
+%mm at 
+    close to 
+maximum radial and
 vertical amplitude. Note that we
@@
 To further understand the effect of the midplane error, we also plot the
-fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of $10, 50,
+fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of 
+%mm $10, 50,
+     10, 50, and $
 100\,\pc$ (orange, teal, and green, respectively) in
-Figure~\ref{fig:dJz_fun_Jz}. For each orbit, we set the initial position to be
+Figure~\ref{fig:dJz_fun_Jz}. 
+For each orbit, we set the initial position to be
 $(8,0,0)\,\kpc$ and the initial velocity to be $(0, -190, v_z)\,\kms$, where
-we vary $v_z$.\footnote{\label{note:vz_orbits}One can recover \thin{}, \thick{}, and \halo{} by setting $v_z=10$, $50$, and $190$, respectively (see Table~\ref{tab:orbits}).} For a \thin{}-like orbit ($J_z\sim0.5\,\actunit$), even a
-$10\,\pc$ offset in $z$ is enough to induce a $\sim25\%$ error in $J_z$. For
+we vary $v_z$.\footnote{\label{note:vz_orbits}One can recover \thin{},
+  \thick{}, and \halo{} by setting $v_z=10$, $50$, and $190
+%mm
+     \kms$,
+ respectively, 
+%mm [one significant figure values]
+      giving $J_z \simeq 0.7, 20,$ and $500\,\actunit$  
+(see Table~\ref{tab:orbits}).} For a \thin{}-like orbit ($J_z\sim
+%mm 0.5
+         0.7%
+\,\actunit$), 
+even a
+$10\,\pc$ offset in $z$ is enough to induce a 
+%mm $\sim25\%$ 
+         $\sim20\%$ 
+error in $J_z$. For
 larger values of $J_z$, the fractional errors are suppressed, but the induced

### 2019-05-15 paper/ms.tex
Revisions up to 4.1

@@
 %mm  \newcommand{\kms}{\text{km}/\text{s}}
-         \newcommand{\kms}{\text{km} \text{s}^{-1}}
+         \newcommand{\kms}{\text{km}\,\text{s}^{-1}}
 \newcommand{\actunit}{\text{kpc}\,\kms}
@@
 
-\section{Coordinate System Offsets from Azimuthal Midplane Variations}
+\section{
+%mm Coordinate System Offsets from Azimuthal 
+Midplane Variations}
 \label{sec:local_fire}
@@
 properties with observations of the Milky Way \citep{2016ApJ...827L..23W,
-2018MNRAS.481.4133G}. The $\z = 0$ snapshots\footnote{In this work, to avoid
-confusion with the vertical height $z$, we refer to cosmological redshift as
-$\z$.} of these three simulations, named \mi{}, \mf{}, and \mm{} for
-shorthand, are publicly available alongside associated mock \textit{Gaia} DR2
+2018MNRAS.481.4133G}. The 
+%mm $\z = 0$ snapshots\footnote{In this work, to avoid
+%confusion with the vertical height $z$, we refer to cosmological redshift as
+%$\z$.} of these three simulations, 
+      The snapshots of these three simulations at cosmological redshift $\z = 0$,
+named \mi{}, \mf{}, and \mm{},
+%mm for shorthand, 
+are publicly available alongside associated mock \textit{Gaia} DR2
 catalogues generated from them.\footnote{\url{http://ananke.hub.yt}}
@@
 clear in the case of the gas; the quoted value for the Milky Way comes from
-studies of high-mass star-forming regions). As noted above, the midplanes
-defined by gas and stars can be tilted with respect to one another as well,
+studies of high-mass star-forming regions). 
+%mm As noted above, the  [not sure it has been?]
+     The 
+midplanes defined by gas and stars can be tilted with respect to one another as well,
 precluding extending the precision of the gas midplane definition to the
@@
 Cosmological simulations of Milky Way-mass galaxies are not perfect
-representations of the true Milky Way in other ways as well,\footnote{The
-failure of cosmological simulations to exactly reproduce the Milky Way is not necessarily due to limitations of the numerical model. Candidate Milky Way-like
-galaxies are chosen solely on their mass and isolation, for which there are a
-wide variety of possible galaxies with qualitatively different properties.} as
-discussed in \citet{2018arXiv180610564S}. For instance, the velocity structure
-of \mi{} is closer to M31's than the Milky Way's (S. Loebman et al., in preparation).
+representations of the true Milky Way in other ways as well,
+%mm [This looks like text, not footnote] \footnote{
+       as discussed in \citet{2018arXiv180610564S}.
+The failure of cosmological simulations to exactly reproduce the Milky
+Way is not necessarily due to limitations of the numerical
+model. Candidate Milky Way-like galaxies are chosen solely on their
+mass and isolation, for which there are a wide variety of possible galaxies with qualitatively different
+properties. 
+%mm }  as discussed in \citet{2018arXiv180610564S}. For instance, 
+       For example,
+the velocity structure of \mi{} is closer to M31's than the Milky
+Way's (S. Loebman et al., in preparation).
+
+%mm [inserted para break]
 However, in this work we are most interested in the global properties of the
@@
 simulation offers us the ability to see how the midplane varies in a more
-controlled environment. The details of the simulation are given in
+controlled environment. The 
+%mm details of the simulation are given in
+     simulation is described by
 \citet{2018MNRAS.481..286L}, but we briefly summarize the most relevant
@@
 For the Milky Way, the dark halo is modeled as a Hernquist
-sphere of mass $10^{12}\,\Msun$ and scale length of $52\,\kpc$, the disk
+sphere of mass $10^{12}\,\Msun$ and scale length of $52\,\kpc$, the
+disk
+%mm
+    is
 modeled as an exponential disk with a scale radius of $3.5\,\kpc$, scale
@@
 of mass $10^{10}\,\Msun$ and scale length $0.7\,\kpc$
-\citep{1990ApJ...356..359H}. We use the L2 model, for which Sagittarius is
+\citep{1990ApJ...356..359H}
+%mm
+     is included.
+%mm We use the L2 model, for which Sagittarius is
+     The Sagittarius dwarf is
 modeled with two components: a dark matter Hernquist sphere of mass $8\times10^{10}\,\Msun$ and
@@
 sphere of mass $6.4\times10^8\,\Msun$ and scale length $0.85\,\kpc$. All
-components are realized with live distributions of N-body particles; the Milky
-Way and Sagittarius are each initialized to be in equilibrium in isolation.
-
-The mass resolution of the simulation is $2.6\times10^4$, $1.2\times10^4$, and
-$1.0\times10^4\,\Msun$ for the dark matter, disk, and bulge components,
+components are realized with 
+%mm live distributions of 
+    distributions of live
+N-body particles; the Milky Way and Sagittarius are each initialized
+to be in equilibrium in isolation.
+
+The mass resolution of the simulation is 
+%mm $2.6\times10^4$, $1.2\times10^4$, 
+       $2.6\times10^4\,\Msun$, $1.2\times10^4\,\Msun$, 
+and $1.0\times10^4\,\Msun$ for the dark matter, disk, and bulge components,
 respectively. For the disk and bulge components, a softening length of
 $30\,\pc$ is used whereas for the halo a softening length of $60\,\pc$ is
-used. For Sagittarius, the softening length for the dark matter and the stars
-is $60$ and $40\,\pc$, respectively.
+used. For Sagittarius, the softening length for the dark matter and
+the stars is 
+%mm $60$
+     $60\,\pc$ 
+and $40\,\pc$, respectively.
 
@@
 \subsection{The Local Midplane} \label{ssec:local_midplane}
-Using the two sets of simulations, we determine the local midplane that an
-observer might measure if they were situated in each of these galaxies as a
-function of azimuth at the solar circle. Starting from the coordinate system
-described in the previous section, which is approximately aligned so that the
+Using the two sets of simulations, we determine the local midplane
+%mm 
+       as a function of azimuth at the solar circle
+that an observer might measure if they were situated in each of these
+galaxies% 
+%mm as a function of azimuth at the solar circle
+. Starting from the coordinate system
+described in the previous section, which is 
+approximately 
+%mm
+    \MM{[isn't this by definition, at least on average, not
+      approximate?]} 
+aligned so that the
 $z$-coordinate is perpendicular to the disk plane, we place our imaginary
 observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and
-vary the azimuth between $0<\phi<2\pi$. We consider $50$ equally spaced bins in azimuth, a sufficient number
-such that each cylinder shares no stars with its neighboring cylinder. At each value of $\phi$ we then
-compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and
-height $1\,\kpc$ perpendicular to the fiducial disk. We then re-define the new
+vary the azimuth between $0<\phi<2\pi$. 
+%mm [this was written backwards.  Reordered.] 
+%We consider $50$ equally spaced bins in azimuth, a sufficient number
+%such that each cylinder shares no stars with its neighboring
+%cylinder. 
+At 
+%mm each value of 
+    equally spaced values of 
+$\phi$ we then compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and
+height $1\,\kpc$ perpendicular to the fiducial disk  
+%mm .
+     and centered on it. We choose to use 50 values, sufficiently few that no cylinder shares stars
+     with its neighbors.
+We then re-define the new
 midplane of the cylinder to be the median $z$, re-select stars, and iterate
@@
 a position $\phi$ by taking the median height of all stars within
-$R=0.5\,\kpc$ and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is
-performed again using the new height $10$ times to converge on the local
-midplane height. In order to allow for the possibility that the fiducial
+$R=0.5\,\kpc$ and $z=1\,\kpc$ (in cylindrical coordinates). 
+%mm [redundant with text] The procedure is performed again using the new height $10$ times
+%to converge on the local midplane height. 
+In order to allow for the possibility that the fiducial
 galactocentric coordinate system is incorrect, we subtract the best fit
-$A\sin{(\phi+B)}+C$ curve from each panel. We then bootstrap resample $1000$ times to determine
+%mm $A\sin{(\phi+B)}+C$ 
+     sine
+curve from each panel. We then bootstrap resample $1000$ times to determine
 $1\,\sigma $ error bars, which we report as dashed lines.}
@@
 Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
-before, we have subtracted the best fit $A\sin{(\phi+B)}+C$ curve to account
+before, we have subtracted the best fit 
+%mm $A\sin{(\phi+B)}+C$ 
+     sine
+curve to account
 for inaccuracies in the galactocentric coordinate system. Error bars are
@@
 To allow for potential small inaccuracies in the determination of the original
-fiducial coordinate system, we also subtract the best fit $\text{A}
-\sin{\left(\phi + \text{B}\right)} + \text{C}$ curve from the midplane as a
+fiducial coordinate system, we also subtract the best fit 
+%mm \text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}\eeq curve
+    curve of the form 
+    \beq A \sin{\left(\phi + B\right)} + C \label{eq:fit-curve}\eeq 
+from the midplane as a
 function of azimuth to account for an overall tilt of the midplane (a
 simplified version of the strategy described in
-\citealt{2019ApJ...871..145A}). For $\text{A}$, the values are $-165$, $45$,
-and $8.8\,\pc$, for $\text{B}$ the values are $0.67$, $-0.088$, and
-$0.031\,\text{rad}$ and for $\text{C}$ the values are $-69$, $19$, and
-$-18\,\pc$ for \mi{}, \mf{}, and \mm{}, respectively. For the assumed Solar
-radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
-from the values of $\text{A}$ --- we compute $1.15$, $0.31$, and $0.06\,\deg$
-for \mi{}, \mf{}, and \mm{}. These angle offsets are consistent with the
+\citealt{2019ApJ...871..145A}). 
+%mm [perhaps this would be better presented as a table] For
+%$\text{A}$, the values are $-165$, $45$, and $8.8\,\pc$, for
+%$\text{B}$ the values are $0.67$, $-0.088$, and $0.031\,\text{rad}$
+%and for $\text{C}$ the values are $-69$, $19$, and $-18\,\pc$ for
+%\mi{}, \mf{}, and \mm{}, respectively. 
+    For models (\mi{}, \mf{}, \mm{}) the best fit values are $A =
+    (-165, 45, 8.8)\,\pc$, $B = (0.67, -0.088, 0.031)\,\text{rad}$,
+    and $C = (-69, 19, -18)\,\pc$.
+For the assumed Solar radius of $8.2\,\kpc$, we can approximate the
+angle offset 
+%mm 
+    $\Delta \theta$ 
+for the $z$-axis from the values of 
+%mm [change to rad for consistency with B.   Note that you did not
+%give two significant figures of accuracy for the last value.
+%Strictly speaking should be corrected, though this is pretty minor.]
+%$\text{A}$ --- we compute $1.15$, $0.31$, and $0.06\,\deg$ for \mi{},
+%\mf{}, and \mm{}.  
+      $A$. We find for the same models $\Delta \theta = (0.020, 5.4
+      \times 10^{-3}, 1.0 \times 10^{-3})\,\text{rad}$.
+These angle offsets are consistent with the
 values given in \citet{2018arXiv180610564S} for the difference between the
@@
 error bars. The middle $90\%$ of midplane values across the solar circle spans
-$185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. In two of the three cases
+%$185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. 
+    $(1.9 \times 10^2, 1.6 \times 10^2, 84)\,\pc$ for these models.
+In two of the three cases
 the midplane therefore varies by more than $\pm 100\,\pc$ depending on the
@@
 Sagittarius encounter \citep{2018MNRAS.481..286L}. Again we have subtracted a
-best fit curve of the form $\text{A} \sin{\left(\phi + \text{B}\right)} +
-\text{C}$, with the values for $\text{A}$ being $9.5$, $2.5$, $-212$, and
-$-394\,\pc$, for $\text{B}$ being $0.0013$, $0.0007$, $-0.63$, and
-$-1.00\,\text{rad}$ and $\text{C}$ being $2.6$, $-7.8$, $-65$, and $-53\,\pc$
-chronologically for each of the four timesteps shown. The middle $90\%$ of midplane values across the solar circle spans $49$, $62$, $137$, and $119\,\pc$, for each timestep.
+best fit curve of the form 
+%mm [again, you are missing significant digits in B: always report the
+%same number. When you gave 3 I cut back to two, but I can't fill in when
+%you only give one.] $\text{A} \sin{\left(\phi + \text{B}\right)} +
+%\text{C}$, with the values for $\text{A}$ being $9.5$, $2.5$, $-212$, and
+%$-394\,\pc$, for $\text{B}$ being $0.0013$, $0.0007$, $-0.63$, and
+%$-1.00\,\text{rad}$ and $\text{C}$ being $2.6$, $-7.8$, $-65$, and
+%$-53\,\pc$ chronologically for each of the four timesteps shown. 
+    given in Equation~(\ref{eq:fit-curve}), with the values at times
+    $(2, 4, 6, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -2.1 \times 10^2,
+    -3.9 \times 10^2)\,\pc$, $B
+    = (1.3 \times 10^{-3}, 7 \times 10^{-4}, -0.63,
+    -1.0)\,\text{rad}$, and $C = (2.6,-7.8, -65, -53)\,\pc$.
+The middle $90\%$ of midplane values across the solar circle spans 
+%mm $49$, $62$, $137$, and $119\,\pc$, for each timestep.
+       $(49, 62, 1.4 \times 10^2, 1.2 \times 10^2)\,\pc$.
 
@@
 \section{Discussion} \label{sec:discussion}
-We have used high-resolution simulations to illustrate that we expect the
-local midplane defined by stellar density to vary with azimuth by up to $\pm
+We have used high-resolution simulations to illustrate 
+%mm that 
+    why
+we expect the local midplane defined by stellar density to vary with azimuth by up to $\pm
 100$ pc, as a natural consequence of the non-axisymmetry of the Galactic disk
 at small scales. While this is not in itself surprising or new, we have also
-discussed that the discrepancy between our local midplane and that of distant
+%mm discussed [shown?]
+    argued
+that the discrepancy between our local midplane and that of distant
 stars introduces a systematic error in the $z$ component when converting from

### 2019-05-15 paper/ms.tex
Update affiliation order

--- 2019-05-14 paper/ms.tex
@@
 \newcommand{\Gyr}{\text{Gyr}}
-%mm  \newcommand{\kms}{\text{km}/\text{s}}
-         \newcommand{\kms}{\text{km} \text{s}^{-1}}
+\newcommand{\kms}{\text{km}/\text{s}}
 \newcommand{\actunit}{\text{kpc}\,\kms}
@@
 
-% \correspondingauthor{Angus Beane}
+\correspondingauthor{Angus Beane}
 \email{abeane@sas.upenn.edu}
@@
 \author[0000-0003-3939-3297]{Robyn E. Sanderson}
+\affil{\penn}
 \affil{\cca}
-\affil{\penn}
 
 \author{Melissa K. Ness}
+\affil{\columbia}
 \affil{\cca}
-\affil{\columbia}
 
@@
 \author[0000-0003-0064-4060]{Mordecai-Mark Mac Low}
+\affil{\cca}
 \affil{\amnh}
-\affil{\cca}
 
@@
 have the attractive quality of being invariant and thus good quantitative labels
-of a star's orbit. A systematic bias in the phase space position
-induces 
-%mm a phase-dependence [to better distinguish the two different
-%meanings of phase in this sentence!]
-   an orbital phase dependence
-in the actions, which we interpret as a systematic error
+of a star's orbit. A systematic bias in the phase space position induces a
+phase-dependence in the actions, which we interpret as a systematic error
 distinct from typical observational errors and Galactic potential uncertainties.
-An offset in the vertical position of 
-%mm
-     the Galactic midplane of
-$\sim15\,\pc$ 
-%mm [this formulation isn't worth the extra difficulty reading]
-%($\sim120\,\pc$) for a thin (thick) disk 
-  for a thin disk orbit or $\sim 120\,\pc$ for a thick disk
-orbit is sufficient to induce a $25\%$ systematic error in the
+An offset in the vertical position of $\sim15\,\pc$ ($\sim120\,\pc$) for a thin
+(thick) disk orbit is sufficient to induce a $25\%$ systematic error in the
 vertical action $J_z$. The induced error distribution is non-Gaussian and
@@
 \citep{2008gady.book.....B} underlies much of the quantitative analysis of the
-mechanisms that give rise to these 
-%mm signautres.
-    signatures. 
-In order to 
-%mm do 
-construct such a
+mechanisms that give rise to these signautres. In order to do construct such a
 system, the Sun's relative position must be defined and measured both precisely
 and accurately. This involves determining the angular position of and distance
-to the Galactic center, the orientation of the Galactic 
-%mm plane, the height of the Sun above the Galactic midplane, 
-   midplane, the distance of the Sun from it,
-and the local standard of rest (LSR). We review
+to the Galactic center, the orientation of the Galactic plane, the height of the
+Sun above the Galactic midplane, and the local standard of rest (LSR). We review
 and discuss the observational efforts to measure these parameters in
@@
 2012ApJ...750L..41W}, but with the vast improvement in the quality of
-phase-space measurements due to \textit{Gaia} 
-%mm this assumption 
-     the assumption of axisymmetry
-is increasingly inadequate \citep[e.g.,][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. Even if
+phase-space measurements due to \textit{Gaia} this assumption is increasingly
+inadequate \citep[e.g.,][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. Even if
 this assumption were close enough for many purposes, the parameters used in
@@
 
-%mm The main point of this paper is as follows. Due to relatively
-%small-scale variations in the density of gas and stars, the local
-%midplane should vary as a
-      Small-scale variations in the density of gas and stars cause the
-      local midplane position to vary as a
-function of radius and azimuth. 
-%mm For stars relatively far from the Sun, their local midplane will
-%differ 
-    Stars far from the Sun have a local midplane that differs
-from our local midplane extrapolated onto their position. Converting
-positions and velocities of more distant stars from a 
+The main point of this paper is as follows. Due to relatively small-scale
+variations in the density of gas and stars, the local midplane should vary as a
+function of radius and azimuth. For stars relatively far from the Sun, their
+local midplane will differ from our local midplane extrapolated onto their
+position. Converting positions and velocities of more distant stars from a
 heliocentric to a Galactocentric coordinate system thus introduces a systematic
@@
 
-%mm [the x-axis of this figure should be labeled orbital phase for
-%consistency with the text]
 \begin{figure*}[ht!]
@@
 show how an inaccurate or erroneous determination of the midplane leads to a
-%mm phase-dependence of the actions calculated from a phase-space
-%point and an
-   dependence on orbital phase of the value of the actions calculated
-   from a point in phase space and an
+phase-dependence of the actions calculated from a phase-space point and an
 assumed potential model. The $x$-axis corresponds to orbital phase and the
-$y$-axis to vertical height. The solid gray curve indicates the 
-%mm ``true'' orbit of the star; that is, as it oscillates around the
-%true midplane.
-     true orbit of the star as it oscillates around the true midplane.
+$y$-axis to vertical height. The solid gray curve indicates the ``true'' orbit
+of the star; that is, as it oscillates around the true midplane.
 
@@
 
-This example uses an offset in $z$, but one 
-%mm would expect an analogous effect from an offset in another coordinate, 
-    analagous effects occur from offsets in other coordinates, 
-such as the distance to the Galactic
+This example uses an offset in $z$, but one would expect an analogous effect
+from an offset in another coordinate, such as the distance to the Galactic
 center or the LSR. A similar effect in which actions gain time-dependence due to
@@
 systematic error in the actions induced from offsets in the position (e.g., the midplane or
-Galactic center distance) or velocity (e.g., the 
-%mm local standard of rest)
-     LSR) 
-of the assumed coordinate system's origin. We use the epicyclic approximation, which
+Galactic center distance) or velocity (e.g., the local standard of rest) of the assumed coordinate system's origin. We use the epicyclic approximation, which
 assumes that the motion in the $z$ and $R$ components of the orbit are
 decoupled and follow simple harmonic motion about a circular and planar
-%mm ``guiding orbit''
-     guiding orbit 
-\citep[][Section~3.2 and references
-therein]{2008gady.book.....B}. We refer to the radius of this orbit
-%mm  (the``guiding radius'') as $R_g$. 
-     as the guiding radius $R_g$.
-This approximation is an excellent description
+``guiding orbit'' \citep[][Section~3.2 and references
+therein]{2008gady.book.....B}. We refer to the radius of this orbit (the
+``guiding radius'') as $R_g$. This approximation is an excellent description
 of the thin disk and a good description of the thick disk in
-an axisymmetric potential
-%mm , i.e., ignoring 
-    that ignores
-the influence of the Galactic bar and
+an axisymmetric potential, i.e., ignoring the influence of the Galactic bar and
 spiral arms. We also make the assumption of a perfectly flat circular
@@
 \equiv v_c/R_g$ is the orbital frequency of the guiding center, $A_z$ and $A_R$ are the amplitudes in the vertical and radial coordinates, and $\alpha$ and
-$\beta$ are 
-%mm arbitrary phases
-    the initial orbital phases at $t = 0$.
-Similarly, the velocities of the orbit are given by:
+$\beta$ are arbitrary phases. Similarly, the velocities of the orbit are
+given by:
 \beq\label{eq:orbits_vel_epi}
@@
 
-In this case, 
-%mm we have that for $J_{\phi}$
-    the azimuthal action is
+In this case, we have that for $J_{\phi}$
 \citep[][Section~3.5.3b]{2008gady.book.....B}:
@@
 \eeq
-and
-%mm for $J_R$ and $J_z$:
-    the radial and vertical actions are
+and for $J_R$ and $J_z$:
 \beq\label{eq:JR_Jz_epi}
 \begin{split}
-J_R &= %mm \frac{E_R}{\kappa} \\
-            E_R/\kappa \\
-J_z &= %mm \frac{E_z}{\nu}  \text{,}
-            E_z / \nu, 
+J_R &= \frac{E_R}{\kappa} \\
+J_z &= \frac{E_z}{\nu} \text{,}
 \end{split}
@@
 potential available in \texttt{galpy} \citep{2015ApJS..216...29B}, which includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
-a Miyamoto--Nagai disk \citep{1975PASJ...27..533M}, and 
-%mm an NFW halo \citep{1997ApJ...490..493N},
-    a Navarro, Frenk, \& White \citeyear{1997ApJ...490..493N} halo,
-and is fit to empirically match some
+a Miyamoto--Nagai disk \citep{1975PASJ...27..533M}, and an NFW halo
+\citep{1997ApJ...490..493N}, and is fit to empirically match some
 observations.
@@
 $J_z$ is decidedly non-Gaussian. Therefore, any sort of error propagation must
-take this into account. 
-%mm Intuition 
-    A heuristic explanation
-for the shape of each panel is given in the
+take this into account. Intuition for the shape of each panel is given in the
 text. We also plot one half the 95th percentile minus the 5th
@@
 We consider the observer making a measurement, integrating an orbit, and
-computing actions every 
-%mm $\Myr$
-    megayear 
-using the prescription above. However, we
+computing actions every $\Myr$ using the prescription above. However, we
 specify the star's starting position using the systematically offset
@@
 The actions computed using the offset coordinate system
-for each phase-space starting point are shown for the first 
-%mm $\Gyr$
-  gigayear 
-of the orbit in the upper panels of
+for each phase-space starting point are shown for the first $\Gyr$ of the
+orbit in the upper panels of
 Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
@@
 Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
-offset coordinate systems oscillate
-%mm s 
-as a function of the time/orbital phase at which the
+offset coordinate systems oscillates as a function of the time/orbital phase at which the
 star's phase-space position is observed.
 This time dependence comes even though
-the observer is using the correctly constructed, best-fit, static, axisymmetric
+the observer is using the correctly constructed best-fit static, axisymmetric
 potential. The relative size of the phase variation in each action depends on
@@
 
-We now 
-%mm describe the intuition behind 
-    suggest a heuristic explanation for
-the shape of Figure~\ref{fig:Jz_hist}.
+We now describe the intuition behind the shape of Figure~\ref{fig:Jz_hist}.
 Consider first the \thick{} (top panel), where the offset in $z$ is
@@
 half the 95th percentile minus the 5th percentile of the distribution
-of action values. We plot this quantity
-%mm
-     in Figure~\ref{fig:Jz_hist}, 
-which we refer to as $\Delta J_i$ for each action, 
-%mm in Figure~\ref{fig:Jz_hist} 
-as a horizontal arrow anchored on the true action value. Because of the bimodality
+of action values. We plot this quantity, which we refer to as $\Delta J_i$ for each action, in Figure~\ref{fig:Jz_hist} as a
+horizontal arrow anchored on the true action value. Because of the bimodality
 of the error distribution, this quantity roughly measures the distance from

### 2019-05-16 paper/ms.tex
final revision

--- 2019-05-15 paper/ms.tex
@@
 
-\subsection{Estimates of Milky Way Midplane Offsets from Current Data}
+\subsection{Estimates of Milky Way Midplane Offsets 
+%mm from Current Data
+      }
+
 \label{ssec:mw_data_midplane}
@@
 out suggestions of an oscillation in average vertical velocities of order
-$5\,\kms$ on $\sim\,\kpc$ scales looking toward the Galactic anticenter.
+$5\,\kms$ on 
+%mm $\sim\,\kpc$
+    roughly kiloparsec 
+scales looking toward the Galactic anticenter.
 
@@
 by the \textit{Gaia} collaboration confirmed these preliminary results on the
-velocity and spatial scales of these oscillation with clear spatial maps made
+velocity and spatial scales of 
+%mm these 
+oscillation with clear spatial maps made
 using DR2 data of median $v_z$ over a significant Galactic volume
@@
 Three-dimensional dust maps also offer a view into the expected variation of
-the stellar disk, since dust should trace high-star formation regions.
+the stellar disk, since dust should trace 
+%mm high-star formation regions. 
+    regions of massive star formation.
 Figure~9 of \citet{2019MNRAS.483.4277C}, Figure~1 of
@@
 show that the midplane varies by $\sim10\degree$ at a distance of
-$\sim0.75\,\kpc$, corresponding to a physical height of $\sim130\,\pc$.
+$\sim0.75\,\kpc$, corresponding to a physical 
+%mm height
+        vertical variation 
+of $\sim130\,\pc$.
 
@@
 account.
-Here we briefly review the current measurements, their uncertainties, and the implications for the calculation of actions.
+Here we briefly review the current measurements
+%mm
+   of the four relevant quantities, 
+their uncertainties, and the implications for the calculation of actions.
 
@@
 recent measurement using near-infrared interferometry places $R_0$ at $8.178 \pm 0.035$~kpc
-\citep{2019arXiv190405721A}, or a $\sim0.3\%$ uncertainty.
-
-However, the location of \sgra{} may not be equivalent to the location of the ``dynamical Galactic center,'' i.e., the
+\citep{2019arXiv190405721A}, or a 
+%mm $\sim0.3\%$ [0.035/8.178 = 0.00428]
+     0.4\% 
+uncertainty.
+
+However, the location of \sgra{} may not be equivalent to the location
+of the 
+%mm ``dynamical Galactic center,'' i.e., the
+     dynamical Galactic center, the
 point in three-dimensional space about which the stars in the solar
@@
 generated by a small error in $R_0$ can be understood from the epicyclic
-approximation (Equation~\eqref{eq:Ji_err_mosttime}, which states that $\Delta J_R/J_R = 2\Delta R/A_R$. The
+approximation 
+%mm (Equation~\eqref{eq:Ji_err_mosttime}, [eqref adds unneeded
+%parentheses here] 
+         (Eq.~\ref{eq:Ji_err_mosttime}),
+which states that $\Delta J_R/J_R = 2\Delta R/A_R$. The
 fractional error in $J_R$ is related to the error in $R_0$ as a fraction of
@@
 We did not consider in this work the effect of the \emph{angular position} of
-the dynamical Galactic center being offset with \sgra{}.
+the dynamical Galactic center being offset 
+%mm with 
+   from
+\sgra{}.
 
@@
 corresponding to a vertical height of $\sim22\,\pc$. This corresponds to a
-$37\%$ ($5\%$) error in $J_z$ for \thin{} (\thick{}).
+$37\%$ 
+%mm ($5\%$) 
+error in $J_z$ for \thin{} 
+%mm (\thick{}).
+    and a 5\% error for \thick{}.
 
@@
 the Galactic midplane, leading to slightly different values (by $\sim 5\,\pc$)
-for the Sun's relative height \citep[e.g.,][]{2019ApJ...871..145A}. A pre-\textit{
-Gaia} review of these measurements is given by \citet{2016ARAA..54..529B}.
-The discrepancy between gas-based and stellar-based determinations of the solar height is small, and thus only likely to dominate over intrinsic midplane variations on small scales, but will be relevant for detailed modeling of young and kinematically cold stars.
-For instance, it will induce a $\sim10\%$ error in $J_z$ for an
-orbit with $z_{\text{max}}\sim100\,\pc$.
+for the Sun's relative height \citep[e.g.,][]{2019ApJ...871..145A}. A
+pre-\textit{%mm [adding an EOL comment to eliminate spurious space]
+Gaia} review of these measurements is given by
+\citet{2016ARAA..54..529B}. The discrepancy between gas-based and
+stellar-based determinations of the solar height is small, and thus
+only likely to dominate over intrinsic midplane variations on small
+scales, but will be relevant for detailed modeling of young and
+kinematically cold stars. For instance, it will induce a $\sim10\%$
+error in $J_z$ for an orbit with $z_{\text{max}}\sim100\,\pc$.
 
@@
 For example, the
-value of the circular velocity is taken to be $\sim 220\kms$
+value of the circular velocity is taken to be 
+%mm $\sim 220\kms$
+         $\sim 220\,\kms$
 \citep[e.g.,][]{2012ApJ...759..131B} with roughly $10\%$ uncertainty. We expect this to translate to at least a $10\%$ systematic error in $J_{\phi}$.
@@
 way to quantify different types of orbits. However, all of our conclusions
-apply also to studies which simply rely on orbit integrations, since the two
+%mm apply also to studies which
+         also apply to studies that 
+simply rely on orbit integrations, since the two
 are equivalent. For instance, computing orbital properties of open or globular
@@
 \section{Conclusions}\label{sec:conclusion}
-Determining the orbital properties of stars is important for understanding the structure and evolution of the Galaxy. Actions have promise as excellent orbit labels. If the Galaxy can be
+Determining the orbital properties of stars is important for
+understanding the structure and evolution of the Galaxy. Actions have
+%mm promise as
+    been argued to be 
+excellent orbit labels. If the Galaxy can be 
+%mm
+    well
 approximated as axisymmetric and 6D phase space positions can be measured
@@
 is not constant across the disk presents a significant complication to
-computed actions actually being invariant. Our main conclusions are as
-follows:
+computed actions actually being invariant. Our main conclusions are% 
+%mm as follows:
+:
 
@@
 \item Inaccuracy in the Galactocentric coordinate
-system induces phase-dependence in the actions calculated from the observed
+system induces
+%mm phase-
+     orbital phase 
+dependence in the actions calculated from the observed
 positions and velocities of stars
@@
 \item Inaccuracy in the midplane location most severely affects computation of
-the vertical action $J_z$. A midplane offset of $\sim6\,\pc$ ($\sim50\,\pc$)
-for a typical thin (thick) disk orbit results in a $10\%$ error in $J_z$ (as
-defined by one half the middle 90th percentile). The fractional error is
+the vertical action $J_z$. A midplane offset of $\sim6\,\pc$ 
+%mm ($\sim50\,\pc$)
+for a typical thin
+%mm (thick) 
+disk orbit results in a $10\%$ error in $J_z$, 
+%mm (as defined by one half the middle 90th percentile). 
+       and even for a thick disk orbit a $50\,\pc$ offset will result
+       in the same size error.
+The fractional error is
 significantly less for halo orbits.
@@
 \item The distribution of systematic errors in the actions induced by a
-coordinate system offset is very non-Gaussian. The distribution is bimodal
+coordinate system offset is 
+%mm very 
+   highly
+non-Gaussian. The distribution is bimodal
 with \emph{neither mode at null}. As a result, error propagation of coordinate
@@
 local Galactic midplane is the global Galactic midplane is not true a
-priori. A violation of this assumption (by, e.g., intrinsic midplane variations) leads to a systematic error in $z$ which generates the issues raised in the last three points.
-
-\item To demonstrate the severity of the previous point, we measured the local galactic
+priori. A violation of this assumption (by, e.g., intrinsic midplane
+variations) leads to a systematic error in $z$ which generates the 
+%mm issues raised in the last three points.
+   large errors in actions summarized above.
+
+\item 
+%mm To demonstrate the severity of the previous point, we measured
+    We show that such midplane variation is likely by measuring
+the local galactic
 midplane along the solar circle in three different high-resolution, zoom-in
@@
 that the midplane varies as a function of azimuth at the solar circle by
-$\sim185$, $162$, $84\,\pc$ (middle $90\%$) for the three cosmological simulations we considered
-(\mi{}, \mf{}, and \mm{}, respectively) and by $\sim60$ and $\sim120\,\pc$ before and after the interaction in the controlled simulation.
-Comparison of the corresponding velocity variations with recent measurements in the Milky Way suggests that these values are consistent with expectations for midplane variations in the Milky Way itself, which have not yet been measured as a function of azimuth.
+%mm [too much detail] $\sim185$, $162$, $84\,\pc$ (middle $90\%$) for
+%the three cosmological simulations we considered (\mi{}, \mf{}, and
+%\mm{}, respectively) and by $\sim60$ and $\sim120\,\pc$ before and
+%after the interaction in the controlled simulation. 
+   60--185~pc in these simulations.
+%mm [this seems redundant with the next point.] Comparison of the corresponding velocity variations with recent measurements in the Milky Way suggests that these values are consistent with expectations for midplane variations in the Milky Way itself, which have not yet been measured as a function of azimuth.
 
 \item Assuming a vertical velocity variation of the Milky Way of
-$\sim5\textup{--}10\,\kms$, consistent with recent results from \textit{Gaia} DR2
-\citep{2018A&A...616A..11G, 2019arXiv190209569F} and our results from the FIRE
+$\sim5\textup{--}10\,\kms$, consistent with recent results from
+\textit{Gaia} 
+%mm DR2 \citep{2018A&A...616A..11G, 2019arXiv190209569F} 
+and our results from the FIRE
 simulations (Figure~\ref{fig:lsr_variations}), we estimated that the
 corresponding midplane offsets are $\sim60\textup{--}170\,\pc$ by dimensional
-analysis with the vertical frequencies of disk-like orbits. Similar offsets are seen in three-dimensional dust maps.
+analysis 
+%mm with 
+   using 
+the vertical frequencies of disk-like orbits. 
+%mm 
+    This range of values is consistent with the variations seen in the simulations.
+Similar offsets are seen in three-dimensional dust maps.
 
@@
 in the computation of actions under the assumption of a global axisymmetric
-potential. We do not propose here what a modeler should use for the ``true''
-global midplane, since it depends on the particular problem being studied.
+potential. 
+%mm [the "true" midplane is spatially and time dependent, so I'm not
+%sure this makes sense as something that could be proposed.] 
+%We do not propose here what a modeler should use for the ``true''
+%global midplane, since it depends on the particular problem being studied.
 Even without a high-precision definition of the global Galactic midplane,
@@
 in the local density). However, for other applications, such as the study of
-action diffusion which motivated this paper in the first place, a more
-extensive perturbative approach is likely to be needed. We plan to explore the
+action diffusion,
+%mm which motivated this paper in the first place, 
+a more extensive perturbative approach is likely to be needed. We plan to explore the
 mitigation of these effects in future work.
@@
 We would like to thank Megan Bedell, Robert A. Benjamin, Tobias Buck, Elena
-D'Onghia, Benoit Famaey, and Adrian Price-Whelan for helpful discussions. A.B.
-would like to thank Todd Phillips for helpful discussions. A.B. was supported
+D'Onghia, Benoit Famaey, 
+%mm and Adrian Price-Whelan
+    Adrian Price-Whelan, and Todd Phillips
+for helpful
+discussions. 
+%mm A.B. would like to thank Todd Phillips for helpful discussions. 
+A.B. was supported
 in part by the Roy \& Diana Vagelos Program in the Molecular Life Sciences and
@@
 Barbara. This work used the Extreme Science and Engineering Discovery
-Environment (XSEDE), which is supported by National Science Foundation grant
-number OCI-1053575. K.V.J.'s contributions were supported in part by the
-National Science Foundation under grants NSF PHY-1748958 and NSF AST-1614743.
-M.-M.M.L. was partly supported by NSF grant AST18-15461. The work of D.A.-A.,
-A.B., D.W.H., R.E.S., K.V.J., M.-M.M.L., and M.K.N. is supported by the Simons
-Foundation.
+Environment (XSEDE), 
+%mm
+    \MM{[Does anyone want their XSEDE grant number here?]}
+which is supported by 
+%mm National Science Foundation 
+     NSF
+grant 
+%mm number 
+OCI-1053575. K.V.J.'s contributions were supported in part by the
+%mm National Science Foundation 
+     NSF
+under grants 
+%mm NSF 
+PHY-1748958 and 
+%mm NSF 
+AST-1614743.
+M.-M.M.L. was partly supported by NSF grant 
+%mm AST18-15461. [for consistency with above]
+        AST-1815461. 
+%mm The work of D.A.-A., A.B., D.W.H., R.E.S., K.V.J., M.-M.M.L., and
+%M.K.N. is supported by the Simons Foundation. [This is redundant with
+%our Flatiron Institute affiliations, unless there is additional,
+%separate, support for any of us.]
 

### 2019-05-17 midplane_test/FIRE/midplane.py
Make midplane.py faster

@@ -4,15 +4,11 @@ from tqdm import tqdm
 import matplotlib.pyplot as plt
 from scipy.optimize import minimize
 
-from pykdgrav import ConstructKDTree, GetAccelParallel
 from astropy.constants import G as G_astropy
-from scipy.optimize import root_scalar
 import astropy.units as u
 
 import sys
-import itertools
 from joblib import Parallel, delayed
-import multiprocessing
 
 from matplotlib import rc
 import matplotlib as mpl
@@ -42,12 +38,12 @@ def read_snap(gal):
     # reads in and sets fiducial coordinates
     # returns snap
     gal_info = 'fiducial_coord/' + gal + '_res7100_center.txt'
-    sim_directory = '/mnt/ceph/users/firesims/fire2/metaldiff/'+gal+'_res7100'
-    snap = gizmo.io.Read.read_snapshots(['star', 'gas', 'dark'], 'index', 600,
+    root = '/mnt/ceph/users/firesims/fire2/metaldiff/'
+    #root = '/Users/abeane/scratch/actions_systematic/data/fire2/metaldiff/'
+    sim_directory = root + gal + '_res7100'
+    snap = gizmo.io.Read.read_snapshots(['star'], 'index', 600,
                                         properties=['position', 'id',
-                                                    'mass', 'velocity',
-                                                    'form.scalefactor',
-                                                    'smooth.length'],
+                                                    'mass', 'velocity'],
                                         assign_center=False,
                                         simulation_directory=sim_directory)
     ref = np.genfromtxt(gal_info, comments='#', delimiter=',')
@@ -155,7 +151,8 @@ def main(gal):
     np.save('output/out_'+gal+'.npy', out)
 
 if __name__ == '__main__':
-    glist = ['m12i', 'm12f', 'm12m']
+    # glist = ['m12i', 'm12f', 'm12m']
+    glist = ['m12m']
     for gal in glist:
         main(gal)
 

### 2019-05-17 paper/ms.tex
Small changes throughout

@@
 
-This example uses an offset in $z$, but one 
+This example uses an offset in $z$, but 
 %mm would expect an analogous effect from an offset in another coordinate, 
@@
 %mm arbitrary phases
-    the initial orbital phases at $t = 0$.
+    the initial orbital phases.
 Similarly, the velocities of the orbit are given by:
@@
 half the 95th percentile minus the 5th percentile of the distribution
-of action values. We plot this quantity
-%mm
-     in Figure~\ref{fig:Jz_hist}, 
-which we refer to as $\Delta J_i$ for each action, 
-%mm in Figure~\ref{fig:Jz_hist} 
-as a horizontal arrow anchored on the true action value. Because of the bimodality
+of action values. We refer to this quantity as $\Delta J_i$ for each action and plot it in Figure~\ref{fig:Jz_hist} as a horizontal arrow anchored on the true action value.
+ Because of the bimodality
 of the error distribution, this quantity roughly measures the distance from
@@
 0)\,\kpc$ and velocity $(0, -190, v_z)\,\kms$, where we vary $v_z$.\footref{note:vz_orbits} We show
-this for a $z$ offset of $10, 50, 100\,\pc$ (orange, teal, and green,
+this for a $z$ offset of $10$, $50$, and $100\,\pc$ (orange, teal, and green,
 respectively). As before, the error ($\Delta J_z$) is one half the 95th
@@
 %mm $10, 50,
-     10, 50, and $
+     $10$, $50$, and $
 100\,\pc$ (orange, teal, and green, respectively) in
@@
 we vary $v_z$.\footnote{\label{note:vz_orbits}One can recover \thin{},
-  \thick{}, and \halo{} by setting $v_z=10$, $50$, and $190
-%mm
-     \kms$,
+  \thick{}, and \halo{} by setting $v_z=10$, $50$, and $190$,
  respectively, 
@@
 %mm Coordinate System Offsets from Azimuthal 
-Midplane Variations}
+Azimuthal Midplane Variations}
 \label{sec:local_fire}
@@
 2018MNRAS.481.4133G}. The 
-%mm $\z = 0$ snapshots\footnote{In this work, to avoid
-%confusion with the vertical height $z$, we refer to cosmological redshift as
-%$\z$.} of these three simulations, 
+$\z = 0$ snapshots\footnote{In this work, to avoid
+confusion with the vertical height $z$, we refer to cosmological redshift as
+$\z$.} of these three simulations, 
       The snapshots of these three simulations at cosmological redshift $\z = 0$,
@@
 %mm $2.6\times10^4$, $1.2\times10^4$, 
-       $2.6\times10^4\,\Msun$, $1.2\times10^4\,\Msun$, 
+       $2.6\times10^4$, $1.2\times10^4$, 
 and $1.0\times10^4\,\Msun$ for the dark matter, disk, and bulge components,
@@
 the stars is 
-%mm $60$
-     $60\,\pc$ 
+     $60$ 
 and $40\,\pc$, respectively.
@@
 approximately 
-%mm
-    \MM{[isn't this by definition, at least on average, not
-      approximate?]} 
 aligned so that the
-$z$-coordinate is perpendicular to the disk plane, we place our imaginary
+$z$-coordinate is perpendicular to the disk plane at the solar circle, we place our imaginary
 observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and
@@
 %\mf{}, and \mm{}.  
-      $A$. We find for the same models $\Delta \theta = (0.020, 5.4
-      \times 10^{-3}, 1.0 \times 10^{-3})\,\text{rad}$.
+      $A$. We find for the same models $\Delta \theta = (1.15, 0.31,
+      0.062)\,\text{deg}$.
 These angle offsets are consistent with the
@@
 D'Onghia, Benoit Famaey, 
-%mm and Adrian Price-Whelan
-    Adrian Price-Whelan, and Todd Phillips
+and Adrian Price-Whelan
+    % Adrian Price-Whelan, and Todd Phillips
 for helpful
 discussions. 
-%mm A.B. would like to thank Todd Phillips for helpful discussions. 
+A.B. would like to thank Todd Phillips for helpful discussions. 
 A.B. was supported

### 2019-05-17 paper/ms.tex
Small edits, remove comments

@@
 \newcommand{\Gyr}{\text{Gyr}}
-%mm  \newcommand{\kms}{\text{km}/\text{s}}
 \newcommand{\kms}{\text{km}\,\text{s}^{-1}}
@@
 induces 
-%mm a phase-dependence [to better distinguish the two different
-%meanings of phase in this sentence!]
    an orbital phase dependence
@@
 An offset in the vertical position of 
-%mm
      the Galactic midplane of
 $\sim15\,\pc$ 
-%mm [this formulation isn't worth the extra difficulty reading]
-%($\sim120\,\pc$) for a thin (thick) disk 
   for a thin disk orbit or $\sim 120\,\pc$ for a thick disk
@@
 mechanisms that give rise to these 
-%mm signautres.
     signatures. 
 In order to 
-%mm do 
 construct such a
@@
 and accurately. This involves determining the angular position of and distance
-to the Galactic center, the orientation of the Galactic 
-%mm plane, the height of the Sun above the Galactic midplane, 
-   midplane, the distance of the Sun from it,
+to the Galactic center, the orientation of and distance to the Galactic 
+   midplane,
 and the local standard of rest (LSR). We review
@@
 phase-space measurements due to \textit{Gaia} 
-%mm this assumption 
      the assumption of axisymmetry
@@
 
-%mm The main point of this paper is as follows. Due to relatively
-%small-scale variations in the density of gas and stars, the local
-%midplane should vary as a
       Small-scale variations in the density of gas and stars cause the
@@
 function of radius and azimuth. 
-%mm For stars relatively far from the Sun, their local midplane will
-%differ 
     Stars far from the Sun have a local midplane that differs
@@
 show how an inaccurate or erroneous determination of the midplane leads to a
-%mm phase-dependence of the actions calculated from a phase-space
-%point and an
    dependence on orbital phase of the value of the actions calculated
@@
 $y$-axis to vertical height. The solid gray curve indicates the 
-%mm ``true'' orbit of the star; that is, as it oscillates around the
-%true midplane.
      true orbit of the star as it oscillates around the true midplane.
@@
 This example uses an offset in $z$, but 
-%mm would expect an analogous effect from an offset in another coordinate, 
     analagous effects occur from offsets in other coordinates, 
@@
 Galactic center distance) or velocity (e.g., the 
-%mm local standard of rest)
      LSR) 
@@
 decoupled and follow simple harmonic motion about a circular and planar
-%mm ``guiding orbit''
      guiding orbit 
@@
 therein]{2008gady.book.....B}. We refer to the radius of this orbit
-%mm  (the``guiding radius'') as $R_g$. 
      as the guiding radius $R_g$.
@@
 an axisymmetric potential
-%mm , i.e., ignoring 
     that ignores
@@
 $\beta$ are 
-%mm arbitrary phases
     the initial orbital phases.
@@
 In this case, 
-%mm we have that for $J_{\phi}$
     the azimuthal action is
@@
 and
-%mm for $J_R$ and $J_z$:
     the radial and vertical actions are
@@
 \begin{split}
-J_R &= %mm \frac{E_R}{\kappa} \\
-            E_R/\kappa \\
-J_z &= %mm \frac{E_z}{\nu}  \text{,}
-            E_z / \nu, 
+J_R &=  E_R/\kappa \\
+J_z &=  E_z / \nu\text{,} 
 \end{split}
@@
 a Miyamoto--Nagai disk \citep{1975PASJ...27..533M}, and 
-%mm an NFW halo \citep{1997ApJ...490..493N},
     a Navarro, Frenk, \& White \citeyear{1997ApJ...490..493N} halo,
@@
 take this into account. 
-%mm Intuition 
     A heuristic explanation
@@
 computing actions every 
-%mm $\Myr$
     megayear 
@@
 for each phase-space starting point are shown for the first 
-%mm $\Gyr$
   gigayear 
@@
 offset coordinate systems oscillate
-%mm s 
 as a function of the time/orbital phase at which the
@@
 We now 
-%mm describe the intuition behind 
     suggest a heuristic explanation for
@@
 orbits. For \thin{}, 
-%mm (\thick{}), 
 a systematic offset of $15\,\pc$
-%mm ($120\,\pc$)
 in the $z$-coordinate results in a $25\%$ deviation in $J_z$,
-%mm
      while a $120\,\pc$ offset results in the same deviation for \thick{}.
@@
 relevant since during most of the orbit the star will be 
-%mm at 
     close to 
@@
 fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of 
-%mm $10, 50,
      $10$, $50$, and $
@@
  respectively, 
-%mm [one significant figure values]
       giving $J_z \simeq 0.7, 20,$ and $500\,\actunit$  
-(see Table~\ref{tab:orbits}).} For a \thin{}-like orbit ($J_z\sim
-%mm 0.5
-         0.7%
-\,\actunit$), 
+(see Table~\ref{tab:orbits}).} For a \thin{}-like orbit ($J_z\sim0.7\,\actunit$), 
 even a
 $10\,\pc$ offset in $z$ is enough to induce a 
-%mm $\sim25\%$ 
          $\sim20\%$ 
@@
 \section{
-%mm Coordinate System Offsets from Azimuthal 
 Azimuthal Midplane Variations}
@@
 named \mi{}, \mf{}, and \mm{},
-%mm for shorthand, 
 are publicly available alongside associated mock \textit{Gaia} DR2
@@
 \enddata
-% \tablerefs{\citet{2016ARAA..54..529B,2018arXiv180610564S,2008ApJ...673..864J}}
 \tablenotetext{a}{$T<100\,\text{K}$}
@@
 studies of high-mass star-forming regions). 
-%mm As noted above, the  [not sure it has been?]
      The 
@@
 representations of the true Milky Way in other ways as well,
-%mm [This looks like text, not footnote] \footnote{
        as discussed in \citet{2018arXiv180610564S}.
@@
 properties. 
-%mm }  as discussed in \citet{2018arXiv180610564S}. For instance, 
        For example,
@@
 
-%mm [inserted para break]
 However, in this work we are most interested in the global properties of the
@@
 controlled environment. The 
-%mm details of the simulation are given in
      simulation is described by
@@
 disk
-%mm
     is
@@
 \citep{1990ApJ...356..359H}
-%mm
      is included.
-%mm We use the L2 model, for which Sagittarius is
      The Sagittarius dwarf is
@@
 components are realized with 
-%mm live distributions of 
     distributions of live
@@
 The mass resolution of the simulation is 
-%mm $2.6\times10^4$, $1.2\times10^4$, 
        $2.6\times10^4$, $1.2\times10^4$, 
@@
 Using the two sets of simulations, we determine the local midplane
-%mm 
        as a function of azimuth at the solar circle
@@
 galaxies% 
-%mm as a function of azimuth at the solar circle
 . Starting from the coordinate system
@@
 vary the azimuth between $0<\phi<2\pi$. 
-%mm [this was written backwards.  Reordered.] 
-%We consider $50$ equally spaced bins in azimuth, a sufficient number
-%such that each cylinder shares no stars with its neighboring
-%cylinder. 
 At 
-%mm each value of 
     equally spaced values of 
@@
 height $1\,\kpc$ perpendicular to the fiducial disk  
-%mm .
      and centered on it. We choose to use 50 values, sufficiently few that no cylinder shares stars
@@
 $R=0.5\,\kpc$ and $z=1\,\kpc$ (in cylindrical coordinates). 
-%mm [redundant with text] The procedure is performed again using the new height $10$ times
-%to converge on the local midplane height. 
 In order to allow for the possibility that the fiducial
 galactocentric coordinate system is incorrect, we subtract the best fit
-%mm $A\sin{(\phi+B)}+C$ 
      sine
@@
 before, we have subtracted the best fit 
-%mm $A\sin{(\phi+B)}+C$ 
      sine
@@
 fiducial coordinate system, we also subtract the best fit 
-%mm \text{A} \sin{\left(\phi + \text{B}\right)} + \text{C}\eeq curve
     curve of the form 
@@
 angle offset 
-%mm 
     $\Delta \theta$ 
@@
 error bars. The middle $90\%$ of midplane values across the solar circle spans
-%$185$, $162$, $84\,\pc$ for \mi{}, \mf{}, and \mm{}. 
     $(1.9 \times 10^2, 1.6 \times 10^2, 84)\,\pc$ for these models.
@@
     given in Equation~(\ref{eq:fit-curve}), with the values at times
-    $(2, 4, 6, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -2.1 \times 10^2,
+    $(2.0, 4.0, 6.0, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -2.1 \times 10^2,
     -3.9 \times 10^2)\,\pc$, $B
@@
 %mm $49$, $62$, $137$, and $119\,\pc$, for each timestep.
-       $(49, 62, 1.4 \times 10^2, 1.2 \times 10^2)\,\pc$.
+       $(49, 62, 140, 120)\,\pc$.
 
@@
 We have used high-resolution simulations to illustrate 
-%mm that 
     why
@@
 at small scales. While this is not in itself surprising or new, we have also
-%mm discussed [shown?]
     argued
@@
 \subsection{Estimates of Milky Way Midplane Offsets 
-%mm from Current Data
       }
@@
 $5\,\kms$ on 
-%mm $\sim\,\kpc$
     roughly kiloparsec 
@@
 velocity and spatial scales of 
-%mm these 
 oscillation with clear spatial maps made
@@
 the stellar disk, since dust should trace 
-%mm high-star formation regions. 
     regions of massive star formation.
@@
 $\sim0.75\,\kpc$, corresponding to a physical 
-%mm height
         vertical variation 
@@
 Here we briefly review the current measurements
-%mm
    of the four relevant quantities, 
@@
 \citep{2019arXiv190405721A}, or a 
-%mm $\sim0.3\%$ [0.035/8.178 = 0.00428]
      0.4\% 
@@
 of the 
-%mm ``dynamical Galactic center,'' i.e., the
      dynamical Galactic center, the
@@
 approximation 
-%mm (Equation~\eqref{eq:Ji_err_mosttime}, [eqref adds unneeded
-%parentheses here] 
-         (Eq.~\ref{eq:Ji_err_mosttime}),
+         (Equation~\eqref{eq:Ji_err_mosttime}),
 which states that $\Delta J_R/J_R = 2\Delta R/A_R$. The
@@
 the dynamical Galactic center being offset 
-%mm with 
    from
@@
 $37\%$ 
-%mm ($5\%$) 
 error in $J_z$ for \thin{} 
-%mm (\thick{}).
     and a 5\% error for \thick{}.
@@
 for the Sun's relative height \citep[e.g.,][]{2019ApJ...871..145A}. A
-pre-\textit{%mm [adding an EOL comment to eliminate spurious space]
-Gaia} review of these measurements is given by
+pre-\textit{Gaia} review of these measurements is given by
 \citet{2016ARAA..54..529B}. The discrepancy between gas-based and
@@
 value of the circular velocity is taken to be 
-%mm $\sim 220\kms$
          $\sim 220\,\kms$
@@
 way to quantify different types of orbits. However, all of our conclusions
-%mm apply also to studies which
          also apply to studies that 
@@
 understanding the structure and evolution of the Galaxy. Actions have
-%mm promise as
     been argued to be 
 excellent orbit labels. If the Galaxy can be 
-%mm
     well
@@
 is not constant across the disk presents a significant complication to
-computed actions actually being invariant. Our main conclusions are% 
-%mm as follows:
-:
+computed actions actually being invariant. Our main conclusions are:
 
@@
 system induces
-%mm phase-
      orbital phase 
@@
 the vertical action $J_z$. A midplane offset of $\sim6\,\pc$ 
-%mm ($\sim50\,\pc$)
 for a typical thin
-%mm (thick) 
 disk orbit results in a $10\%$ error in $J_z$, 
-%mm (as defined by one half the middle 90th percentile). 
        and even for a thick disk orbit a $50\,\pc$ offset will result
@@
 coordinate system offset is 
-%mm very 
    highly
@@
 variations) leads to a systematic error in $z$ which generates the 
-%mm issues raised in the last three points.
    large errors in actions summarized above.
@@
 \item 
-%mm To demonstrate the severity of the previous point, we measured
     We show that such midplane variation is likely by measuring
@@
 \textit{Gaia} 
-%mm DR2 \citep{2018A&A...616A..11G, 2019arXiv190209569F} 
 and our results from the FIRE
@@
 analysis 
-%mm with 
    using 
 the vertical frequencies of disk-like orbits. 
-%mm 
     This range of values is consistent with the variations seen in the simulations.
@@
 action diffusion,
-%mm which motivated this paper in the first place, 
 a more extensive perturbative approach is likely to be needed. We plan to explore the
@@
 and Adrian Price-Whelan
-    % Adrian Price-Whelan, and Todd Phillips
 for helpful

### 2019-05-17 paper/ms.tex
Update orbits table
Update abstract

@@
 
+% Orbital properties of stars, computed from their six-dimensional phase space measurements and an assumed Galactic potential, are used to understand the structure and evolution of the Galaxy.
+Stellar actions, computed from orbits, are invoked as invariant and thus quantitative labels of a star’s orbit.
+
+%   Quantitatively, an offset in the vertical position of ∼ 15 pc (∼ 120 pc) for a thin (thick) disk orbit is sufficient to induce a 25\% systematic error in the vertical action Jz.
+Such an offset can be trivially generated by local midplane variations.
+In the FIRE simulations these are on the order of ~100pc at the solar circle.
+For the Milky Way, from the mean vertical velocity variation of the Milky Way disk of ∼ 5–10 km/s with radius, combined with the dust-map density distribution, we estimate that the Milky Way midplane variations are of order ∼ 60–170 pc.
+Action calculations which assume the global and local midplanes are identical, are likely to include the induced errors we report, depending on the volume considered.
+Variation in the local standard of rest causes similar issues.
+The variation of the midplane must be taken into account when performing dynamical analysis across the large regions of the disk accessible to Gaia and future missions.
+
 Orbital properties of stars, computed from their six-dimensional phase space
@@
 structure and evolution of the Galaxy. Stellar actions, computed from orbits,
-have the attractive quality of being invariant and thus good quantitative labels
-of a star's orbit. A systematic bias in the phase space position
-induces 
-   an orbital phase dependence
-in the actions, which we interpret as a systematic error
-distinct from typical observational errors and Galactic potential uncertainties.
-An offset in the vertical position of 
-     the Galactic midplane of
-$\sim15\,\pc$ 
-  for a thin disk orbit or $\sim 120\,\pc$ for a thick disk
-orbit is sufficient to induce a $25\%$ systematic error in the
-vertical action $J_z$. The induced error distribution is non-Gaussian and
-bimodal, with neither of the modes peaking on the null value. Furthermore, we
-show that the local midplane varies by $\sim100\,\pc$ at the solar circle in
-Milky Way-mass cosmological zoom-in simulations from the FIRE project. From
-observations that the mean vertical velocity of the Milky Way disk varies by
-$\sim5\textup{--}10\,\kms$ with radius, we estimate that the true midplane
-variations are of order $\sim60\textup{--}170\,\pc$, with similar offsets seen
-in recent three-dimensional dust maps. Thus, current state-of-the-art action
-calculations, which assume the global and local midplanes are identical, are
-likely to include a systematic vertical offset depending on the volume
-considered. Variation in the local standard of rest induces similar issues. The
-variation of the midplane must be taken into account when performing dynamical
-analysis across the large regions of the disk accessible to \textit{Gaia} and
-future missions.
+have the attractive quality of being invariant under certain assumptions and
+are therefore used as quantitative labels of a star's orbit. We report a
+subtle but important systematic error that is induced in the actions as a
+consequence of local midplane variations expected for the Milky Way. This
+error is difficult to model because it is non-Gaussian and bimodal, with
+neither mode peaking on the null value. An offset in the vertical position of
+the Galactic midplane of $\sim15\,\pc$ for a thin disk-like orbit or $\sim
+120\,\pc$ for a thick disk-like orbit is sufficient to induce a $25\%$
+systematic error in the vertical action $J_z$. In FIRE simulations of Milky
+Way-mass galaxies, these variations are on the order of $\sim100\,\pc$ at the
+solar circle. From observations of the mean vertical velocity variation of
+$\sim5\textup{--}10\,\kms$ with radius, we estimate that the Milky Way
+midplane variations are of order $\sim60\textup{--}170\,\pc$\,--\,also
+consistent with recent three-dimensional dust maps. Action calculations and
+orbit integrations which assume the global and local midplanes are identical
+are likely to include the induced errors we report, depending on the volume
+considered. Variation in the local standard of rest or distance to the
+Galactic center causes similar issues. The variation of the midplane must be
+taken into account when performing dynamical analysis across the large regions
+of the disk accessible to \textit{Gaia} and future missions.
+
+%  Furthermore, we
+% show that the local midplane varies by $\sim100\,\pc$ at the solar circle in
+% Milky Way-mass cosmological zoom-in simulations from the FIRE project. From
+% observations that the mean vertical velocity of the Milky Way disk varies by
+% $\sim5\textup{--}10\,\kms$ with radius, we estimate that the true midplane
+% variations are of order $\sim60\textup{--}170\,\pc$, with similar offsets seen
+% in recent three-dimensional dust maps. Thus, current state-of-the-art action
+% calculations, which assume the global and local midplanes are identical, are
+% likely to include a systematic vertical offset depending on the volume
+% considered. Variation in the local standard of rest induces similar issues. The
+% variation of the midplane must be taken into account when performing dynamical
+% analysis across the large regions of the disk accessible to \textit{Gaia} and
+% future missions.
 
@@
 z_{\text{max}}$ and $A_R = \frac{1}{2}(R_{\text{max}} - R_{\text{min}})$.
-\label{tab:orbits}} \tablehead{\colhead{name} & \colhead{initial position} &
-\colhead{initial velocity} & \colhead{$J_R$} & \colhead{$J_{\phi}$} &
+\label{tab:orbits}} \tablehead{\colhead{name} & \colhead{\makecell{initial \\ position}} &
+\colhead{\makecell{initial \\ velocity}} & \colhead{$J_R$} & \colhead{$J_{\phi}$} &
 \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
 \colhead{$\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$} & \colhead{$\kappa$} &
-\colhead{$\nu$} \\ \colhead{ } & \colhead{$\mathrm{kpc}$} &
-\colhead{$\mathrm{km\,s}^{-1}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
-\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$} &
-\colhead{$\mathrm{Myr}^{-1}$} & \colhead{$\mathrm{Myr}^{-1}$}}
+\colhead{$\nu$} \\ \colhead{ } & \colhead{$(\mathrm{kpc})$} &
+\colhead{$(\mathrm{km\,s}^{-1})$} & \colhead{$(\mathrm{kpc\,km\,s^{-1}})$} &
+\colhead{$(\mathrm{kpc\,km\,s^{-1}})$} & \colhead{$(\mathrm{kpc\,km\,s^{-1}})$} &
+\colhead{$(\mathrm{kpc})$} & \colhead{$(\mathrm{kpc})$} &
+\colhead{$(\mathrm{Myr}^{-1})$} & \colhead{$(\mathrm{Myr}^{-1})$}}
 \startdata

### 2019-05-18 paper/ms.tex
Fix acknowledgements

@@
 We would like to thank Megan Bedell, Robert A. Benjamin, Tobias Buck, Elena
-D'Onghia, Benoit Famaey, 
-and Adrian Price-Whelan
-for helpful
-discussions. 
-A.B. would like to thank Todd Phillips for helpful discussions. 
-A.B. was supported
-in part by the Roy \& Diana Vagelos Program in the Molecular Life Sciences and
-the Roy \& Diana Vagelos Challenge Award. K.V.J.'s, C.F.P.L.'s, and R.E.S.'s work was
-performed in part during the Gaia19 workshop and the 2019 Santa Barbara Gaia
-Sprint (also supported by the Heising-Simons Foundation), both hosted by the
-Kavli Institute for Theoretical Physics at the University of California, Santa
-Barbara. This work used the Extreme Science and Engineering Discovery
-Environment (XSEDE), 
-%mm
-    \MM{[Does anyone want their XSEDE grant number here?]}
-which is supported by 
-%mm National Science Foundation 
-     NSF
-grant 
-%mm number 
-OCI-1053575. K.V.J.'s contributions were supported in part by the
-%mm National Science Foundation 
-     NSF
-under grants 
-%mm NSF 
-PHY-1748958 and 
-%mm NSF 
-AST-1614743.
-M.-M.M.L. was partly supported by NSF grant 
-%mm AST18-15461. [for consistency with above]
+D'Onghia, Benoit Famaey, and Adrian Price-Whelan for helpful discussions. A.B.
+would like to thank Todd Phillips for helpful discussions. This project was
+developed in part at the 2019 Santa Barbara Gaia Sprint, hosted by the Kavli
+Institute for Theoretical Physics (KITP) at the University of California,
+Santa Barbara. This research was supported in part at KITP by the
+Heising-Simons Foundation and the National Science Foundation (NSF) under
+Grant No. PHY-1748958. This work used the Extreme Science and Engineering
+Discovery Environment (XSEDE), which is supported by NSF Grant No.
+OCI-1053575. A.B. was supported in part by the Roy \& Diana Vagelos Program in
+the Molecular Life Sciences and the Roy \& Diana Vagelos Challenge Award.
+K.V.J.'s contributions were supported in part by the NSF under Grant No.
+AST-1614743. M.-M.M.L. was partly supported by the NSF under Grant No.
 AST-1815461.
-%mm The work of D.A.-A., A.B., D.W.H., R.E.S., K.V.J., M.-M.M.L., and
-%M.K.N. is supported by the Simons Foundation. [This is redundant with
-%our Flatiron Institute affiliations, unless there is additional,
-%separate, support for any of us.]
 

### 2019-05-19 paper/ms.tex
Small edits throughout

@@
 construct such a
-system, the Sun's relative position must be defined and measured both precisely
+system, the Sun's relative position and velocity must be defined and measured both precisely
 and accurately. This involves determining the angular position of and distance
@@
 2$ kpc) volume from the \textit{Gaia} satellite, the study of stellar actions
-has gained new popularity. One reason is dimensionality reduction---an
+has gained new popularity. One reason is dimensionality reduction\,---\,an
 individual stellar orbit is concisely described by three actions, as opposed to
@@
 \startdata
-\thin{} & (8, 0, 0) & (0, -190, 10) & 40.3 & -1520 & 0.69 & 0.12 & 1.29 & 0.049
+\thin{} & (8, 0, 0) & (0, -190, 10) & 40 & -1500 & 0.69 & 0.12 & 1.3 & 0.049
 & 0.093 \\
-\thick{} & (8, 0, 0) & (0, -190, 50) & 32.5 & -1520 & 23.0 & 0.85 &
-1.19 & 0.048 & 0.061 \\
-\halo{} & (8, 0, 0) & (0, -190, 190) & 32.8 & -1520 &
-529.1 & 6.16 & 2.34 & 0.033 & 0.025
+\thick{} & (8, 0, 0) & (0, -190, 50) & 33 & -1500 & 23 & 0.85 &
+1.2 & 0.048 & 0.061 \\
+\halo{} & (8, 0, 0) & (0, -190, 190) & 33 & -1500 &
+530 & 6.2 & 2.3 & 0.033 & 0.025
 \enddata
@@
 Galactocentric coordinates depends on the LSR). In this potential, we have
-that the circular velocity $v_{\text{circ}}$ is $231\,\kms$ at the Solar
+that the circular velocity $v_{\text{circ}}$ is $231\,\kms$ at the solar
 circle.
@@
 $z$ and the lower panels consider an offset in $x$ (equivalently, an offset in
-the Solar radius). In some panels we also plot as dashed lines the epicyclic
+the solar radius). In some panels we also plot as dashed lines the epicyclic
 prediction of the induced action error (Equation~\eqref{eq:Ji_err_mosttime}).
@@
 amplitude), where $v_z \sim 0$ and where the star is on most of it orbit. This
-is why the distribution, which is calculated at evenly spaced time intervals, peaks at these values. For the \thin{}
+is why the distribution, which is calculated at evenly spaced time intervals, peaks at these values. For \thin{}
 (bottom panel), the offset in $z$ is comparable to $z_{\text{max}}$. Now,
@@
 induced in each action for an offset in the $z$-component. In the lower panels we consider offsets in the $x$ component (i.e. the
-Solar radius). The left, center, and right columns show the fractional spread
+solar radius). The left, center, and right columns show the fractional spread
 in the values computed for $J_z$, $J_{\phi}$, and $J_R$, respectively.
@@
 
-For the offset in the Solar radius (lower panels), the error is largest for
+For the offset in the solar radius (lower panels), the error is largest for
 $J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
@@
 described in the previous section, which is 
-approximately 
 aligned so that the
-$z$-coordinate is perpendicular to the disk plane at the solar circle, we place our imaginary
+$z$-coordinate is approximately perpendicular to the disk plane at the solar circle, we place our imaginary
 observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and
@@
     For models (\mi{}, \mf{}, \mm{}) the best fit values are $A =
-    (-165, 45, 8.8)\,\pc$, $B = (0.67, -0.088, 0.031)\,\text{rad}$,
+    (-165, 45, 8.8)\,\pc$, $B = (38, -5.0, 1.8)\,\text{deg}$,
     and $C = (-69, 19, -18)\,\pc$.
-For the assumed Solar radius of $8.2\,\kpc$, we can approximate the
+For the assumed solar radius of $8.2\,\kpc$, we can approximate the
 angle offset 
@@
 for the $z$-axis from the values of 
-%mm [change to rad for consistency with B.   Note that you did not
-%give two significant figures of accuracy for the last value.
-%Strictly speaking should be corrected, though this is pretty minor.]
-%$\text{A}$ --- we compute $1.15$, $0.31$, and $0.06\,\deg$ for \mi{},
-%\mf{}, and \mm{}.  
       $A$. We find for the same models $\Delta \theta = (1.15, 0.31,
@@
 best fit curve of the form 
-%mm [again, you are missing significant digits in B: always report the
-%same number. When you gave 3 I cut back to two, but I can't fill in when
-%you only give one.] $\text{A} \sin{\left(\phi + \text{B}\right)} +
-%\text{C}$, with the values for $\text{A}$ being $9.5$, $2.5$, $-212$, and
-%$-394\,\pc$, for $\text{B}$ being $0.0013$, $0.0007$, $-0.63$, and
-%$-1.00\,\text{rad}$ and $\text{C}$ being $2.6$, $-7.8$, $-65$, and
-%$-53\,\pc$ chronologically for each of the four timesteps shown. 
     given in Equation~(\ref{eq:fit-curve}), with the values at times
@@
     -3.9 \times 10^2)\,\pc$, $B
-    = (1.3 \times 10^{-3}, 7 \times 10^{-4}, -0.63,
-    -1.0)\,\text{rad}$, and $C = (2.6,-7.8, -65, -53)\,\pc$.
+    = (0.074, 0.039, -36,
+    -57)\,\text{deg}$, and $C = (2.6,-7.8, -65, -53)\,\pc$.
 The middle $90\%$ of midplane values across the solar circle spans 
-%mm $49$, $62$, $137$, and $119\,\pc$, for each timestep.
        $(49, 62, 140, 120)\,\pc$.
@@
 velocities reach a maximum (effectively the median height of all disk stars).
-The Solar height is usually taken to be $\sim 25\,\pc$
+The solar height is usually taken to be $\sim 25\,\pc$
 \citep{2001ApJ...553..184C}, with a more recent measurement from \textit{Gaia}
@@
 potential. 
-%mm [the "true" midplane is spatially and time dependent, so I'm not
-%sure this makes sense as something that could be proposed.] 
-%We do not propose here what a modeler should use for the ``true''
-%global midplane, since it depends on the particular problem being studied.
-Even without a high-precision definition of the global Galactic midplane,
-current observations from \textit{Gaia} should soon permit a measurement of the
+Current observations from \textit{Gaia} should soon permit a measurement of the
 real azimuthal dependence of the midplane location. For some applications,

### 2019-05-20 schmactions/many_orbits/plot_many_orbits_Jzfun.py
Convert km/s to km s^-1

@@ -130,7 +130,7 @@ ax.set_xscale('log')
 ax.set_xlim([xmin, xmax])
 ax.set_ylim([ymin, ymax])
 
-ax.set_xlabel(r'$J_z\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+ax.set_xlabel(r'$J_z\,[\,\text{kpc}\,\text{km}\,\text{s}^{-1}\,]$')
 ax.set_ylabel(r'$\Delta J_z/J_z\,[\,\%\,]$')
 
 myh = Line2D([0], [0], color='k', linestyle='dashed')

### 2019-05-20 schmactions/one_orbit/plot_one_orbit.py

@@ -108,11 +108,11 @@ for x in ax[1]:
     x.set_xticks(np.arange(0,1000,100), minor=True)
 
 for x in ax[:,0]:
-    x.set_ylabel(r'$J_{R,\text{comp}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+    x.set_ylabel(r'$J_{R,\text{comp}}\,[\,\text{kpc}\,\text{km}\,\text{s}^{-1}\,]$')
 for x in ax[:,1]:
-    x.set_ylabel(r'$J_{\phi,\text{comp}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+    x.set_ylabel(r'$J_{\phi,\text{comp}}\,[\,\text{kpc}\,\text{km}\,\text{s}^{-1}\,]$')
 for x in ax[:,2]:
-    x.set_ylabel(r'$J_{z,\text{comp}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+    x.set_ylabel(r'$J_{z,\text{comp}}\,[\,\text{kpc}\,\text{km}\,\text{s}^{-1}\,]$')
 
 ax[0][1].set_title(r'$z\,\text{offset}=100\,\text{pc}$')
 ax[1][1].set_title(r'$x\,\text{offset}=100\,\text{pc}$')
@@ -154,7 +154,7 @@ if True: # so I can indent
     ax[1][0].set_ylabel(r'$\text{count}$')
     ax[1][0].text(35.1, 32.5, r'\texttt{thin-disk}', color='k')
  
-    ax[1][0].set_xlabel(r'$J_{R,\text{comp}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
+    ax[1][0].set_xlabel(r'$J_{R,\text{comp}}\,[\,\text{kpc}\,\text{km}\,\text{s}^{-1}\,]$')
     
     # # #
     # JR (z error) histogram
@@ -187,7 +187,7 @@ if True: # so I can indent
     ax[1][1].set_ylabel(r'$\text{count}$')
     ax[1][1].text(39.3, 59, r'\texttt{thin-disk}', color='k')
     
-    ax[1][1].set_xlabel(r'$J_{R,\text{comp}}\,[\,\text{kpc}\,\text{km}/\text{s}\,]$')
[…]

### 2019-05-20 schmactions/schmactions.py

@@ -10,7 +10,6 @@ from tqdm import tqdm
 import warnings
 
 from joblib import Parallel, delayed
-import multiprocessing
 
 class schmactions(object):
     def __init__(self, init_pos, init_vel, mw=None,

### 2019-05-20 paper/ms.tex
Add software citations
Add FI to ack
Small changes throughout

@@
 
-% \submitjournal{ApJ}
+\submitjournal{ApJ}
 
@@
 error in the Galactocentric coordinate system. We consider here \thick{},
-which has actions of $(J_R, J_{\phi}, J_z) = (37.9, -1520,
+which has actions of $(J_R, J_{\phi}, J_z) = (38, -1500,
 7.0)\,\actunit$ and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We
@@
 scheme fails and very large actions are reported by \texttt{gala}---we perform
-a $5\sigma$ clip on each action to exclude such orbits, but this only excludes
+a $4\sigma$ clip on each action to exclude such orbits, but this only excludes
 a total of $5$ orbits out of the $1000$ considered for
@@
 error bars. The middle $90\%$ of midplane values across the solar circle spans
-    $(1.9 \times 10^2, 1.6 \times 10^2, 84)\,\pc$ for these models.
+    $(190, 160, 84)\,\pc$ for these models.
 In two of the three cases
@@
     given in Equation~(\ref{eq:fit-curve}), with the values at times
-    $(2.0, 4.0, 6.0, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -2.1 \times 10^2,
-    -3.9 \times 10^2)\,\pc$, $B
+    $(2.0, 4.0, 6.0, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -210,
+    -390)\,\pc$, $B
     = (0.074, 0.039, -36,
@@
 that the midplane varies as a function of azimuth at the solar circle by
-%mm [too much detail] $\sim185$, $162$, $84\,\pc$ (middle $90\%$) for
-%the three cosmological simulations we considered (\mi{}, \mf{}, and
-%\mm{}, respectively) and by $\sim60$ and $\sim120\,\pc$ before and
-%after the interaction in the controlled simulation. 
-   60--185~pc in these simulations.
-%mm [this seems redundant with the next point.] Comparison of the corresponding velocity variations with recent measurements in the Milky Way suggests that these values are consistent with expectations for midplane variations in the Milky Way itself, which have not yet been measured as a function of azimuth.
+   $60\textup{--}185\,\pc$ in these simulations.
 
@@
 Discovery Environment (XSEDE), which is supported by NSF Grant No.
-OCI-1053575. A.B. was supported in part by the Roy \& Diana Vagelos Program in
+OCI-1053575. This work uses data hosted by the Flatiron Institute's FIRE data hub. A.B. was supported in part by the Roy \& Diana Vagelos Program in
 the Molecular Life Sciences and the Roy \& Diana Vagelos Challenge Award.
@@
 AST-1815461.
+
+\software{This work made use of the following software: \texttt{astropy} \citep{astropy:2013,astropy:2018}, \texttt{gala}
+\citep{gala, gala:zenodo}, \texttt{matplotlib} \citep{Hunter:2007}, \texttt{numpy} \citep{numpy:2011}, \texttt{scipy}
+\citep{scipy:2001}, and \texttt{tqdm} \citep{casper_o_da_costa_luis_2019_2800317}.}
 

### 2019-05-20 real_cluster/compute_real_cluster_actions.py
Remove real cluster folder

@@ -1,66 +1,0 @@
-import numpy as np 
-
-import gala.dynamics as gd
-import gala.potential as gp
-import gala.integrate as gi
-
-from pyia import GaiaData
-from astropy.table import Table 
-import astropy.units as u
-import astropy.coordinates as coord
-
-import warnings
-
-dt = 1 * u.Myr
-mw = gp.MilkyWayPotential()
-
-t = Table.read('J_A+A_616_A10_tablea3.dat.fits', format='fits')
-
-# rename stuff
-t['ra'] = t['RAdeg']
-t['dec'] = t['DEdeg']
-t['parallax'] = t['plx']
-t['pmra'] = t['pmRA']
-t['pmdec'] = t['pmDE']
-t['radial_velocity'] = t['RV']
-
-g = GaiaData(t)
-
-dist = 1/(g.parallax.to_value(u.mas)) * 1000 # pc
-name = t['Cluster'].tolist()
-m = np.c_[name, dist]
-
-rsun = 8.2 * u.kpc
-zsun = 0.025 * u.kpc
-vsun = [11.1, 232.24, 7.25] * u.km/u.s
-gc_frame = coord.Galactocentric(galcen_distance=rsun, galcen_v_sun=coord.CartesianDifferential(*vsun), z_sun=zsun)
-
-sc = g.skycoord
-scdyn = gd.PhaseSpacePosition(sc.transform_to(gc_frame).cartesian)
-
-def actions(star):
-    with warnings.catch_warnings(record=True):
-        warnings.simplefilter("ignore")
-        orbit = mw.integrate_orbit(star, dt=dt, t1=0*u.Gyr, t2=5*u.Gyr, Integrator=gi.DOPRI853Integrator)
-        res = gd.actionangle.find_actions(orbit, N_max=8)
-        ans = res['actions'].to(u.kpc * u.km / u.s).value
-        return np.append(ans,orbit.zmax(approximate=True).to(u.pc).value)
-
-result = [actions(s) for s in scdyn]
-result2 = np.concatenate((m, result), axis=1)
-r = Table(result2, names=('cluster', 'distance', 'Jr', 'Lz', 'Jz', 'zmax'))
-r['distance'].unit = u.pc
-r['Jr'].unit = u.kpc * u.km/u.s
-r['Lz'].unit = u.kpc * u.km/u.s
-r['Jz'].unit = u.kpc * u.km/u.s
-r['zmax'].unit = u.pc
-
[…]

### 2019-05-21 paper/ms.tex
Update affil

@@
 History, Central Park West at 79th St., New York, NY 10024, USA}
-\newcommand{\columbia}{Department of Astronomy, Columbia University, 550 W
+\newcommand{\columbia}{Department of Astronomy, Columbia University, 550 West
 120th St., New York, NY 10027, USA}
@@
 Physics, New York University, 726 Broadway, New York, NY 10003, USA}
-\newcommand{\nyucds}{Center for Data Science, New York University, 60 Fifth
+\newcommand{\nyucds}{Center for Data Science, New York University, 60 5th
 Ave., New York, NY 10011, USA}

### 2019-05-21 paper/ms.tex
Typographical changes

@@
 $\sim5\textup{--}10\,\kms$ with radius, we estimate that the Milky Way
-midplane variations are of order $\sim60\textup{--}170\,\pc$\,--\,also
+midplane variations are of order $\sim60\textup{--}170\,\pc$\,---\,also
 consistent with recent three-dimensional dust maps. Action calculations and
@@
 \citep{2008gady.book.....B} underlies much of the quantitative analysis of the
-mechanisms that give rise to these 
-    signatures. 
+mechanisms that give rise to these signatures. 
 In order to 
@@
 and accurately. This involves determining the angular position of and distance
-to the Galactic center, the orientation of and distance to the Galactic 
-   midplane,
+to the Galactic center, the orientation of and distance to the Galactic midplane,
 and the local standard of rest (LSR). We review
@@
 2012ApJ...750L..41W}, but with the vast improvement in the quality of
-phase-space measurements due to \textit{Gaia} 
-     the assumption of axisymmetry
+phase-space measurements due to \textit{Gaia} the assumption of axisymmetry
 is increasingly inadequate \citep[e.g.,][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. Even if
@@
 used in many works to compute actions may not be a good description of the true
-potential --- or the parameters used may yield a potential that is
+potential\,---\,or the parameters used may yield a potential that is
 systematically incorrect outside an original fitted region. In this work, we
@@
 \eeq
-and
-    the radial and vertical actions are
+and the radial and vertical actions are
 \beq\label{eq:JR_Jz_epi}
 \begin{split}
-J_R &=  E_R/\kappa \\
+J_R &= E_R / \kappa \\
 J_z &= E_z / \nu\text{,} 
@@
 The actions computed using the offset coordinate system
-for each phase-space starting point are shown for the first 
-  gigayear 
+for each phase-space starting point are shown for the first gigayear 
 of the orbit in the upper panels of
 Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
-scheme fails and very large actions are reported by \texttt{gala}---we perform
+scheme fails and very large actions are reported by \texttt{gala}\,---\,we perform
 a $4\sigma$ clip on each action to exclude such orbits, but this only excludes
@@
 prediction of the induced action error (Equation~\eqref{eq:Ji_err_mosttime}).
-In the epicyclic approximation, a $z$ offset only induces an error in $J_z$
---- for all three orbits the epicyclic approximation is a good description of
+In the epicyclic approximation, a $z$ offset only induces an error in $J_z$\,---\,for all three orbits the epicyclic approximation is a good description of
 the $J_z$ error. An $x$ offset induces an error in $J_R$ and $J_{\phi}$. The
@@
 
-We now 
-    suggest a heuristic explanation for
+We now suggest a heuristic explanation for
 the shape of Figure~\ref{fig:Jz_hist}.
@@
 v_{\phi}=0$, as dashed lines in the color of each orbit. This equation is
-relevant since during most of the orbit the star will be 
-    close to 
+relevant since during most of the orbit the star will be close to 
 maximum radial and
@@
 To further understand the effect of the midplane error, we also plot the
-fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of 
-     $10$, $50$, and $
+fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of $10$, $50$, and $
 100\,\pc$ (orange, teal, and green, respectively) in
@@
 simulation offers us the ability to see how the midplane varies in a more
-controlled environment. The 
-     simulation is described by
+controlled environment. The simulation is described by
 \citet{2018MNRAS.481..286L}, but we briefly summarize the most relevant
@@
 sphere of mass $10^{12}\,\Msun$ and scale length of $52\,\kpc$, the
-disk
-    is
+disk is
 modeled as an exponential disk with a scale radius of $3.5\,\kpc$, scale
@@
 of mass $10^{10}\,\Msun$ and scale length $0.7\,\kpc$
-\citep{1990ApJ...356..359H}
-     is included.
-     The Sagittarius dwarf is
+\citep{1990ApJ...356..359H} is included. The Sagittarius dwarf is
 modeled with two components: a dark matter Hernquist sphere of mass $8\times10^{10}\,\Msun$ and
@@
 sphere of mass $6.4\times10^8\,\Msun$ and scale length $0.85\,\kpc$. All
-components are realized with 
-    distributions of live
+components are realized with distributions of live
 N-body particles; the Milky Way and Sagittarius are each initialized
@@
 
-The mass resolution of the simulation is 
-       $2.6\times10^4$, $1.2\times10^4$, 
+The mass resolution of the simulation is $2.6\times10^4$, $1.2\times10^4$, 
 and $1.0\times10^4\,\Msun$ for the dark matter, disk, and bulge components,
@@
 used. For Sagittarius, the softening length for the dark matter and
-the stars is 
-     $60$ 
+the stars is $60$ 
 and $40\,\pc$, respectively.
@@
 \subsection{The Local Midplane} \label{ssec:local_midplane}
-Using the two sets of simulations, we determine the local midplane
-       as a function of azimuth at the solar circle
+Using the two sets of simulations, we determine the local midplane as a function of azimuth at the solar circle
 that an observer might measure if they were situated in each of these
@@
 vary the azimuth between $0<\phi<2\pi$. 
-At 
-    equally spaced values of 
+At equally spaced values of 
 $\phi$ we then compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and
-height $1\,\kpc$ perpendicular to the fiducial disk  
-     and centered on it. We choose to use 50 values, sufficiently few that no cylinder shares stars
-     with its neighbors.
+height $1\,\kpc$ perpendicular to the fiducial disk and centered on it. We choose to use 50 values, sufficiently few that no cylinder shares stars with its neighbors.
 We then re-define the new
@@
 In order to allow for the possibility that the fiducial
-galactocentric coordinate system is incorrect, we subtract the best fit
-     sine
+galactocentric coordinate system is incorrect, we subtract the best fit sine
 curve from each panel. We then bootstrap resample $1000$ times to determine
@@
 Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
-before, we have subtracted the best fit 
-     sine
+before, we have subtracted the best fit sine
 curve to account
@@
 To allow for potential small inaccuracies in the determination of the original
-fiducial coordinate system, we also subtract the best fit 
-    curve of the form 
-    \beq A \sin{\left(\phi + B\right)} + C \label{eq:fit-curve}\eeq 
+fiducial coordinate system, we also subtract the best fit curve of the form 
+\beq \label{eq:fit-curve}
+A \sin{\left(\phi + B\right)} + C 
+\eeq 
 from the midplane as a
@@
 for the $z$-axis from the values of 
-      $A$. We find for the same models $\Delta \theta = (1.15, 0.31,
-      0.062)\,\text{deg}$.
+$A$. We find for the same models $\Delta \theta = (1.15, 0.31, 0.062)\,\text{deg}$.
 These angle offsets are consistent with the
@@
 $1\,\sigma$ error from the bootstrap procedure is shown as the dashed-line
-error bars. The middle $90\%$ of midplane values across the solar circle spans
-    $(190, 160, 84)\,\pc$ for these models.
+error bars. The middle $90\%$ of midplane values across the solar circle spans $(190, 160, 84)\,\pc$ for these models.
 In two of the three cases
@@
 Sagittarius encounter \citep{2018MNRAS.481..286L}. Again we have subtracted a
-best fit curve of the form 
-    given in Equation~(\ref{eq:fit-curve}), with the values at times
+best fit curve of the form given in Equation~(\ref{eq:fit-curve}), with the values at times
 $(2.0, 4.0, 6.0, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -210,
@@
 
-\subsection{Estimates of Milky Way Midplane Offsets 
-      }
+\subsection{Estimates of Milky Way Midplane Offsets}
 
@@
 out suggestions of an oscillation in average vertical velocities of order
-$5\,\kms$ on 
-    roughly kiloparsec 
+$5\,\kms$ on roughly kiloparsec 
 scales looking toward the Galactic anticenter.
@@
 simulation that the vertical velocity variation as a function of azimuth is
-$\sim5\text{--}10\,\kms$ (Figure~\ref{fig:lsr_variations}), consistent with
+$\sim5\textup{--}10\,\kms$ (Figure~\ref{fig:lsr_variations}), consistent with
 these observations.
@@
 Three-dimensional dust maps also offer a view into the expected variation of
-the stellar disk, since dust should trace 
-    regions of massive star formation.
+the stellar disk, since dust should trace regions of massive star formation.
 Figure~9 of \citet{2019MNRAS.483.4277C}, Figure~1 of
@@
 show that the midplane varies by $\sim10\degree$ at a distance of
-$\sim0.75\,\kpc$, corresponding to a physical 
-        vertical variation 
+$\sim0.75\,\kpc$, corresponding to a physical vertical variation 
 of $\sim130\,\pc$.
@@
 account.
-Here we briefly review the current measurements
-   of the four relevant quantities, 
+Here we briefly review the current measurements of the four relevant quantities, 
 their uncertainties, and the implications for the calculation of actions.
@@
 recent measurement using near-infrared interferometry places $R_0$ at $8.178 \pm 0.035$~kpc
-\citep{2019arXiv190405721A}, or a 
-     0.4\% 
+\citep{2019arXiv190405721A}, or a $0.4\%$ 
 uncertainty.
@@
 However, the location of \sgra{} may not be equivalent to the location
-of the 
-     dynamical Galactic center, the
+of the dynamical Galactic center, the
 point in three-dimensional space about which the stars in the solar
@@
 generated by a small error in $R_0$ can be understood from the epicyclic
-approximation 
-         (Equation~\eqref{eq:Ji_err_mosttime}),
+approximation (Equation~\eqref{eq:Ji_err_mosttime}),
 which states that $\Delta J_R/J_R = 2\Delta R/A_R$. The
@@
 We did not consider in this work the effect of the \emph{angular position} of
-the dynamical Galactic center being offset 
-   from
+the dynamical Galactic center being offset from
 \sgra{}.
@@
 $37\%$ 
-error in $J_z$ for \thin{} 
-    and a 5\% error for \thick{}.
+error in $J_z$ for \thin{} and a 5\% error for \thick{}.
 
@@
 For example, the
-value of the circular velocity is taken to be 
-         $\sim 220\,\kms$
+value of the circular velocity is taken to be $\sim 220\,\kms$
 \citep[e.g.,][]{2012ApJ...759..131B} with roughly $10\%$ uncertainty. We expect this to translate to at least a $10\%$ systematic error in $J_{\phi}$.
@@
 We have mainly been concerned with actions, since they provide a convenient
-way to quantify different types of orbits. However, all of our conclusions
-         also apply to studies that 
+way to quantify different types of orbits. However, all of our conclusions also apply to studies that 
 simply rely on orbit integrations, since the two
@@
 Determining the orbital properties of stars is important for
-understanding the structure and evolution of the Galaxy. Actions have
-    been argued to be 
-excellent orbit labels. If the Galaxy can be 
-    well
+understanding the structure and evolution of the Galaxy. Actions have been argued to be 
+excellent orbit labels. If the Galaxy can be well
 approximated as axisymmetric and 6D phase space positions can be measured
@@
 \item Inaccuracy in the Galactocentric coordinate
-system induces
-     orbital phase 
+system induces orbital phase 
 dependence in the actions calculated from the observed
@@
 \item The distribution of systematic errors in the actions induced by a
-coordinate system offset is 
-   highly
+coordinate system offset is highly
 non-Gaussian. The distribution is bimodal

### 2019-05-21 cartoon/cartoon.py
Change orbital angle to orbital phase in cartoon

@@ -28,7 +28,7 @@ y = np.cos(x)
 y2 = 0.25*np.cos(x + np.pi) - 1.25
 y3 = 2.25*np.cos(x) - 1.25
 
-ax.set_xlabel(r'$\text{orbital angle}$')
+ax.set_xlabel(r'$\text{orbital phase}$')
 ax.set_ylabel(r'$z$')
 
 ax.set_xticklabels([])

### 2019-05-21 paper/ms.tex
Update keywords
Small changes throughout

@@
 
-\keywords{Galaxy: disk -- Galaxy: evolution -- Galaxy: kinematics and dynamics
--- Galaxy: structure -- stars: kinematics and dynamics}
+\keywords{Milky Way disk -- Milky Way evolution -- Milky Way dynamics -- Galaxy structure -- Orbits -- Stellar dynamics}
 
@@
 \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,} \eeq where $p_i$ are the conjugate
-momenta. Under the assumption of axisymmetry, $i=R,z,\phi$ are the radial,
-vertical, and azimuthal coordinates respectively in a cylindrical coordinate
+momenta. Under the assumption of axisymmetry, $i=R,\phi,z$ are the radial,
+azimuthal, and vertical coordinates respectively in a cylindrical coordinate
 system. In a slowly-evolving axisymmetric potential, these actions are invariant
@@
 to any remaining uncertainty in measurements of the Galactocentric radius of the
-Sun and to variations in the LSR.
+Sun, and to variations in the LSR.
 
@@
 \end{center}
-\caption{Cartoon approximation showing the effect an error in the determination
+\caption{Illustration showing the effect an error in the determination
 of the coordinate midplane can have on orbit integration and action estimation.
@@
 with the correct coordinate system). Hence an incorrect midplane in the
-potential model assumed will induce phase-dependence in the actions estimated
+potential model assumed will induce phase dependence in the actions estimated
 for a given star in that potential.} \label{fig:cartoon} \end{figure*}
@@
 In Section~\ref{sec:ref_frame}, we describe the general impact coordinate system
-errors have on the measured actions. In Section~\ref{sec:local_fire} we examine
+errors have on the measured actions. In Section~\ref{sec:local_fire}, we examine
 the azimuthal variations of the midplane itself in examples from two classes of
@@
 \subsection{Effect of Midplane Offset} \label{ssec:cartoon}
-We present a cartoon approximation of an orbit in Figure~\ref{fig:cartoon} to
+We present an illustration of an orbit in Figure~\ref{fig:cartoon} to
 show how an inaccurate or erroneous determination of the midplane leads to a
@@
 $y$-axis to vertical height. The solid gray curve indicates the 
-true orbit of the star as it oscillates around the true midplane.
-
-The dashed gray line, offset from the true midplane, is the midplane location
+true orbit of the star as it oscillates around the true midplane. The dashed gray line, offset from the true midplane, is the midplane location
 used by an observer to integrate the orbit of the star and estimate its actions.
@@
 Now suppose this observer makes a measurement of the star's position and
-velocity at the teal point or the orange point (i.e. at two different orbital
-phases). Then the teal and orange curves correspond to the orbits that the
+velocity at the teal point or the orange point (i.e., at two different orbital
+phases). Then, the teal and orange curves correspond to the orbits that the
 observer would compute for each point based on the potential model with the
@@
 in the correct potential (in this example, the one with the correct midplane)
-should be phase-independent.
+should be phase independent.
 
@@
 
-\tablecaption{Description and names of the three orbits considered in this work.
-For the last four columns: $z_{\text{max}}$ is the maximum height of the orbit,
+\tablecaption{Names and properties of the three orbits considered in this work, where
+$z_{\text{max}}$ is the maximum height of the orbit,
 $\frac{1}{2}(R_{\text{max}} - R_{\text{min}})$ is the magnitude of the radial
@@
 LSR) 
-of the assumed coordinate system's origin. We use the epicyclic approximation, which
+of the assumed coordinate system's origin.
+
+We use the epicyclic approximation, which
 assumes that the motion in the $z$ and $R$ components of the orbit are
@@
 \citep[][Section~3.2 and references
-therein]{2008gady.book.....B}. We refer to the radius of this orbit
+therein]{2008gady.book.....B}. The radius of this orbit is referred to
 as the guiding radius $R_g$.
@@
 \begin{split}
+R(t) &= R_g + A_R \sin{(\kappa t + \alpha)} \\
 \phi(t) &= \Omega_c t \\
-R(t) &= R_g + A_R \sin{(\kappa t + \alpha)} \\
 z(t) &= A_z \sin{(\nu t + \beta)}
@@
 where $\kappa$ and $\nu$ are the radial/epicyclic and vertical frequencies, $\Omega_c
-\equiv v_c/R_g$ is the orbital frequency of the guiding center, $A_z$ and $A_R$ are the amplitudes in the vertical and radial coordinates, and $\alpha$ and
+\equiv v_c/R_g$ is the orbital frequency of the guiding center, $A_R$ and $A_z$ are the amplitudes in the radial and vertical coordinates, and $\alpha$ and
 $\beta$ are 
@@
 \begin{split}
+v_R(t) &= \kappa A_R \cos{(\kappa t + \alpha)} \\
 v_{\phi}(t) &= v_c \\
-v_R(t) &= \kappa A_R \cos{(\kappa t + \alpha)} \\
 v_z(t) &= \nu A_z \cos{(\nu t + \beta)}
@@
 propagation of errors formula to Equation~\eqref{eq:JR_Jz_epi_energy} to
-determine the error in each of the actions. For $J_{\phi}$, the induced error
+determine the error in each of the actions. 
+For $J_R$, the induced error is:
+\beq\label{eq:induced_JR}
+\frac{\Delta J_R}{J_R} = \frac{2(R-R_g)}{A_R^2}\Delta R
+                         + \frac{2v_R}{v_{R,\text{max}}^2} \Delta v_R \text{.}
+\eeq
+For $J_{\phi}$, the induced error
 is:
@@
                                     + \frac{\Delta v_{\phi}}{v_c}\text{.}
-\eeq
-For $J_R$, the induced error is:
-\beq\label{eq:induced_JR}
-\frac{\Delta J_R}{J_R} = \frac{2(R-R_g)}{A_R^2}\Delta R
-                         + \frac{2v_R}{v_{R,\text{max}}^2} \Delta v_R \text{.}
 \eeq
@@
 
-Since most of the time stars will be at maximum amplitude (i.e. turnaround) in
+Since most of the time stars will be at maximum amplitude (i.e., turnaround) in
 both $R$ and $z$, we can approximate the order of magnitude of the systematic
@@
 \begin{split}
+\frac{\Delta J_{R}}{J_{R}} &= \frac{2\Delta R}{A_R} \\
 \frac{\Delta J_{\phi}}{J_{\phi}} &= \frac{\Delta R}{R_g}
                                     + \frac{\Delta v_{\phi}}{v_c} \\
-\frac{\Delta J_{R}}{J_{R}} &= \frac{2\Delta R}{A_R} \\
 \frac{\Delta J_{z}}{J_{z}} &= \frac{2\Delta z}{A_z} \text{,}
@@
 Galactic potential. We compute actions as in \citet{2018ApJ...867...31B},
-using the code \texttt{gala} v0.3 to perform orbit integrations and conversion
-to action space \citep{2017JOSS....2..388P,Price-Whelan:2018}. To compute
+using the code \texttt{gala}~v0.3 to perform orbit integrations and conversion
+to action space \citep{gala, gala:zenodo}. To compute
 actions we use the torus-mapping technique first presented by
@@
 minus 5\textsuperscript{th} percentiles are $2.2$ and $6.2\,\actunit$ for
-$J_R$ and $J_z$, respectively. These are fractional errors of $5.7\%$ and
-$86\%$, respectively. The error induced in $J_{\phi}$ is negligible, as
+$J_R$ and $J_z$, respectively. As a fraction of the true values, these are $5.7\%$ and
+$86\%$, respectively. The spread induced in $J_{\phi}$ is negligible, as
 expected since $J_{\phi}$ only depends on the $x$- and $y$-components of the
@@
 is computed as part of the torus-fitting method.} It is worth pointing out
-that a $100\,\pc$ error in an orbit with $z_{\text{max}}=850\,\pc$\,---\,a
-$12\%$ error\,---\,induced an $86\%$ error in the computation of $J_z$.
+that a $100\,\pc$ offset in an orbit with $z_{\text{max}}=850\,\pc$\,---\,a
+$12\%$ error\,---\,induced an $86\%$ spread in the computation of $J_z$.
 
@@
 $47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively.
-These are fractional errors of $21\%$, $3.1\%$ and $3.1\%$ in these actions, respectively, despite only a $1.2\%$ error in the distance to the Galactic center.
+These are fractionally $21\%$, $3.1\%$ and $3.1\%$ in these actions, respectively, despite only a $1.2\%$ error in the distance to the Galactic center.
 
@@
 \end{center}
-\caption{We report one half the 95th minus 5th percentile of the error
-in the measured action ($\Delta J_i$) from coordinate system errors for
-\thin{}, \thick{}, and \halo{} orbits (Table~\ref{tab:orbits}). See discussion in the
+\caption{We report the fractional error in each action $J_i$ induced by coordinate system offsets for \thin{}, \thick{}, and \halo{} (Table~\ref{tab:orbits}).
+The error ($\Delta J_i$) is computed as one half the 95th minus 5th percentile of the distribution of computed action values.
+See discussion in the
 text and Figure~\ref{fig:Jz_hist} for the justification in using this to measure the magnitude of the
@@
 $z$ and the lower panels consider an offset in $x$ (equivalently, an offset in
-the solar radius). In some panels we also plot as dashed lines the epicyclic
+the solar radius). In some panels, we also plot as dashed lines the epicyclic
 prediction of the induced action error (Equation~\eqref{eq:Ji_err_mosttime}).
@@
 Consider first the \thick{} (top panel), where the offset in $z$ is
-much less than $z_{\text{max}}$ of the orbit. The peaks in the distribution
+much less than the $z_{\text{max}}$ of the orbit. The peaks in the distribution
 correspond to the turning points of the orbit (or points of maximum vertical
@@
 J_i/J_i$ for the three different fiducial orbits in Table~\ref{tab:orbits}.
-Figure~\ref{fig:one_orbit_wrong_ref} considered \thick{} (\thickcolor), but we also now consider the effect on
-the action determined for \thin{} (\thincolor) and \halo{} (\halocolor).
-
 The upper panels of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread
@@
 $J_R$, with some deviations resulting in $J_{\phi}$ and relatively small
-deviations in $J_z$. In the lower middle and lower right panels all three
+deviations in $J_z$. In the lower center and lower right panels all three
 lines nearly overlap.
@@
 respectively). As before, the error ($\Delta J_z$) is one half the 95th
-minus 5th percentile of the distribution of $J_z$ values over the
+minus 5th percentile of the distribution of computed $J_z$ values over the
 course of the orbit. There are large errors for \thin{}-like orbits ($J_z
-\sim 0.5\,\actunit$), even for a small midplane offset of $10\,\pc$. As dashed
+\sim 0.7\,\actunit$), even for a small midplane offset of $10\,\pc$. As dashed
 lines in each color we also plot the prediction for $\Delta J_z/J_z$ from the
@@
 For the Milky Way, the dark halo is modeled as a Hernquist
-sphere of mass $10^{12}\,\Msun$ and scale length of $52\,\kpc$, the
+sphere of mass $10^{12}\,\Msun$ and scale length of $52\,\kpc$ \citep{1990ApJ...356..359H}, the
 disk is
 modeled as an exponential disk with a scale radius of $3.5\,\kpc$, scale
-height $0.53\,\kpc$, and mass $6\times10^{10}\,\Msun$, and a Hernquist bulge
-of mass $10^{10}\,\Msun$ and scale length $0.7\,\kpc$
-\citep{1990ApJ...356..359H} is included. The Sagittarius dwarf is
+height $0.53\,\kpc$, and mass $6\times10^{10}\,\Msun$, and the bulge as a Hernquist sphere
+of mass $10^{10}\,\Msun$ and scale length $0.7\,\kpc$. The Sagittarius dwarf is
 modeled with two components: a dark matter Hernquist sphere of mass $8\times10^{10}\,\Msun$ and
@@
 $\phi$ we then compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and
-height $1\,\kpc$ perpendicular to the fiducial disk and centered on it. We choose to use 50 values, sufficiently few that no cylinder shares stars with its neighbors.
+height $1\,\kpc$ perpendicular to the fiducial disk and centered on it. We choose to use 50 bins in azimuth, sufficiently few that no cylinder shares stars with its neighbors.
 We then re-define the new
@@
 this procedure are necessary for convergence. The resulting median $z$ is
-taken to be what our observer would measure as the local Galactic midplane at
+taken to be what our observer would measure as the local galactic midplane at
 each $\phi$.
@@
 \caption{The local midplane determined at the fiducial solar circle
-($R_0 = 8.2\,\kpc$) for four different time steps from live N-body simulations of a
+($R_0 = 8.2\,\kpc$) for four different time steps from a live N-body simulation of a
 Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
@@
 \citealt{2019ApJ...871..145A}). 
-%mm [perhaps this would be better presented as a table] For
-%$\text{A}$, the values are $-165$, $45$, and $8.8\,\pc$, for
-%$\text{B}$ the values are $0.67$, $-0.088$, and $0.031\,\text{rad}$
-%and for $\text{C}$ the values are $-69$, $19$, and $-18\,\pc$ for
-%\mi{}, \mf{}, and \mm{}, respectively. 
-For models (\mi{}, \mf{}, \mm{}) the best fit values are $A =
-(-165, 45, 8.8)\,\pc$, $B = (38, -5.0, 1.8)\,\text{deg}$,
+For simulations (\mi{}, \mf{}, \mm{}), the best fit values are $A =
+(-170, 45, 8.8)\,\pc$, $B = (38, -5.0, 1.8)\,\text{deg}$,
 and $C = (-69, 19, -18)\,\pc$.
@@
 for the $z$-axis from the values of 
-$A$. We find for the same models $\Delta \theta = (1.15, 0.31, 0.062)\,\text{deg}$.
+$A$. We find for the same simulations $\Delta \theta = (1.15, 0.31, 0.062)\,\text{deg}$.
 These angle offsets are consistent with the
@@
 $1\,\sigma$ error from the bootstrap procedure is shown as the dashed-line
-error bars. The middle $90\%$ of midplane values across the solar circle spans $(190, 160, 84)\,\pc$ for these models.
+error bars. The middle $90\%$ of midplane values across the solar circle spans $(190, 160, 84)\,\pc$ for these simulations.
 In two of the three cases
@@
 we expect the local midplane defined by stellar density to vary with azimuth by up to $\pm
-100$ pc, as a natural consequence of the non-axisymmetry of the Galactic disk
+100\,\pc$ as a natural consequence of the non-axisymmetry of the Galactic disk
 at small scales. While this is not in itself surprising or new, we have also
@@
 
-Systematic variations in $v_z$ and number were first noted as asymmetries in the local
+Systematic variations in $v_z$ and number density were first noted as asymmetries in the local
 velocity distribution towards the North and South Galactic Caps from the
@@
 If the dynamical Galactic center is offset from \sgra{} by $100\,\pc$, only a
-$1.2\%$ difference, then this induces a $\sim20\%$ error in $J_R$ for the disk-like orbits we considered (see
+$1.2\%$ difference, then this induces a $\sim15\%$ error in $J_R$ for the disk-like orbits we considered (see
 Section~\ref{ssec:quant}). The reason such a large error in $J_R$ can be
@@
 ($\sim1.2\,\kpc$ for \thin{} and \thick{}). This
-also implies the very precise $0.3\%$ measurement of $R_0$ still translates to
-a $\sim4\%$ uncertainty in $J_R$.
+also implies the very precise $0.4\%$ measurement of $R_0$ still translates to
+a $\sim6\%$ uncertainty in $J_R$.
 
@@
 
-We did not consider in this work the effect of the \emph{angular position} of
+We did not consider in this work the effect of the angular position of
 the dynamical Galactic center being offset from
@@
 \item Inaccuracy in the midplane location most severely affects computation of
-the vertical action $J_z$. A midplane offset of $\sim6\,\pc$ 
+the vertical action $J_z$. A midplane offset of $\sim15\,\pc$ 
 for a typical thin
-disk orbit results in a $10\%$ error in $J_z$, 
-and even for a thick disk orbit a $50\,\pc$ offset will result
+disk orbit results in a $25\%$ error in $J_z$, 
+and even for a thick disk orbit a $120\,\pc$ offset will result
 in the same size error.
@@
 \item This work underlines the importance of combining chemistry and dynamics.
-Since chemical tagging \citep{2002ARA&A..40..487F} is not subject to the same systematic errors discussed in this paper, it should be used to confirm dynamical associations and to offset the effect of these systematic errors on the action-space distribution function.
+Since chemical tagging \citep{2002ARA&A..40..487F} is not subject to the same systematic errors discussed in this work, it should be used to confirm dynamical associations and to offset the effect of these systematic errors on the action-space distribution function.
 
@@
 action diffusion,
-a more extensive perturbative approach is likely to be needed. We plan to explore the
+a more extensive perturbative approach is likely needed. We plan to explore the
 mitigation of these effects in future work.
@@
 D'Onghia, Benoit Famaey, and Adrian Price-Whelan for helpful discussions. A.B.
-would like to thank Todd Phillips for helpful discussions. This project was
+would like to thank Todd Phillips for helpful discussions. This work uses data hosted by the Flatiron Institute's FIRE data hub. This project was
 developed in part at the 2019 Santa Barbara Gaia Sprint, hosted by the Kavli
@@
 Discovery Environment (XSEDE), which is supported by NSF Grant No.
-OCI-1053575. This work uses data hosted by the Flatiron Institute's FIRE data hub. A.B. was supported in part by the Roy \& Diana Vagelos Program in
+OCI-1053575. A.B. was supported in part by the Roy \& Diana Vagelos Program in
 the Molecular Life Sciences and the Roy \& Diana Vagelos Challenge Award.
@@
 \thick{} (upper panels) and \thin{} (lower panels). We find that the $J_R$ distribution induced by an offset in $x$ more
-closely resembles a Gaussian, while the $J_R$ distribution induced by
+closely resembles a Gaussian distribution, while the $J_R$ distribution induced by
 an offset in $z$ and the $J_{\phi}$ distribution induced by an offset

### 2019-05-21 paper/ms.tex
Fix line indentation throughout

@@
 \usepackage{scrextend}
-
-\newcommand{\Gus}[1]{\textcolor{red}{#1}}
-\newcommand{\MM}[1]{\textcolor{blue}{#1}}   %mm  Mordecai
 
@@
 
-% affiliations
 \newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute,
@@
 \begin{abstract}
-
-% Orbital properties of stars, computed from their six-dimensional phase space measurements and an assumed Galactic potential, are used to understand the structure and evolution of the Galaxy.
-Stellar actions, computed from orbits, are invoked as invariant and thus quantitative labels of a star’s orbit.
-
-%   Quantitatively, an offset in the vertical position of ∼ 15 pc (∼ 120 pc) for a thin (thick) disk orbit is sufficient to induce a 25\% systematic error in the vertical action Jz.
-Such an offset can be trivially generated by local midplane variations.
-In the FIRE simulations these are on the order of ~100pc at the solar circle.
-For the Milky Way, from the mean vertical velocity variation of the Milky Way disk of ∼ 5–10 km/s with radius, combined with the dust-map density distribution, we estimate that the Milky Way midplane variations are of order ∼ 60–170 pc.
-Action calculations which assume the global and local midplanes are identical, are likely to include the induced errors we report, depending on the volume considered.
-Variation in the local standard of rest causes similar issues.
-The variation of the midplane must be taken into account when performing dynamical analysis across the large regions of the disk accessible to Gaia and future missions.
 
@@
 
-%  Furthermore, we
-% show that the local midplane varies by $\sim100\,\pc$ at the solar circle in
-% Milky Way-mass cosmological zoom-in simulations from the FIRE project. From
-% observations that the mean vertical velocity of the Milky Way disk varies by
-% $\sim5\textup{--}10\,\kms$ with radius, we estimate that the true midplane
-% variations are of order $\sim60\textup{--}170\,\pc$, with similar offsets seen
-% in recent three-dimensional dust maps. Thus, current state-of-the-art action
-% calculations, which assume the global and local midplanes are identical, are
-% likely to include a systematic vertical offset depending on the volume
-% considered. Variation in the local standard of rest induces similar issues. The
-% variation of the midplane must be taken into account when performing dynamical
-% analysis across the large regions of the disk accessible to \textit{Gaia} and
-% future missions.
-
 \end{abstract}
 
-\keywords{Milky Way disk -- Milky Way evolution -- Milky Way dynamics -- Galaxy structure -- Orbits -- Stellar dynamics}
-
-\section{Introduction} \label{sec:intro} Our understanding of the Milky Way is
-currently undergoing a revolution as a result of \textit{Gaia} Data Release 2
-(DR2). Recent major discoveries include the affirmation of remnants of a major
-merger \citep{2018ApJ...860L..11K, 2018MNRAS.478..611B, 2018Natur.563...85H,
-2019MNRAS.486..378L, 2019MNRAS.482.3426M} hinted at in pre-\textit{Gaia} work
+\keywords{Milky Way disk -- Milky Way evolution -- Milky Way dynamics --
+Galaxy structure -- Orbits -- Stellar dynamics}
+
+\section{Introduction} \label{sec:intro} 
+Our understanding of the Milky Way is currently undergoing a revolution as a
+result of \textit{Gaia} Data Release 2 (DR2). Recent major discoveries include
+the affirmation of remnants of a major merger \citep{2018ApJ...860L..11K,
+2018MNRAS.478..611B, 2018Natur.563...85H, 2019MNRAS.486..378L,
+2019MNRAS.482.3426M} hinted at in pre-\textit{Gaia} work
 \citep[e.g.,][]{2005MNRAS.359...93M, 2011MNRAS.412.1203N}, a phase-space
@@
 2018arXiv181103631B}. These discoveries all indicate that the Milky Way's
-stellar distribution, which demonstrably departs significantly from axisymmetry,
-is undergoing phase mixing and dynamical interactions across a range of spatial
-and temporal scales.
+stellar distribution, which demonstrably departs significantly from
+axisymmetry, is undergoing phase mixing and dynamical interactions across a
+range of spatial and temporal scales.
 
@@
 \citep{2008gady.book.....B} underlies much of the quantitative analysis of the
-mechanisms that give rise to these signatures. 
-In order to 
-construct such a
-system, the Sun's relative position and velocity must be defined and measured both precisely
-and accurately. This involves determining the angular position of and distance
-to the Galactic center, the orientation of and distance to the Galactic midplane,
-and the local standard of rest (LSR). We review
-and discuss the observational efforts to measure these parameters in
+mechanisms that give rise to these signatures. In order to construct such a
+system, the Sun's relative position and velocity must be defined and measured
+both precisely and accurately. This involves determining the angular position
+of and distance to the Galactic center, the orientation of and distance to the
+Galactic midplane, and the local standard of rest (LSR). We review and discuss
+the observational efforts to measure these parameters in
 Section~\ref{ssec:coord_off}.
@@
 often desirable to convert this measurement into action space to concisely
-summarize its projected orbit, model the stellar distribution function, or find
-stars with similar dynamical properties. Actions are conserved quantities that
-describe the orbit of a star under the assumption of regular, bound orbits in a
-system where the equations of motion are separable in a particular coordinate
-system. They are the cyclical integral of the canonical momentum over its
-conjugate position: \beq\label{eq:actions} J_i \equiv \frac{1}{2\pi}
-\oint_{\text{orbit}}p_i\,\text{d}x_i\text{,} \eeq where $p_i$ are the conjugate
-momenta. Under the assumption of axisymmetry, $i=R,\phi,z$ are the radial,
-azimuthal, and vertical coordinates respectively in a cylindrical coordinate
-system. In a slowly-evolving axisymmetric potential, these actions are invariant
-and $J_{\phi} \equiv L_z$, where $L_z$ is the $z$-component of the angular
-momentum \citep{2008gady.book.....B,2014RvMP...86....1S}.
-
-With the advent of 6D phase-space measurements over a relatively large ($\gtrsim
-2$ kpc) volume from the \textit{Gaia} satellite, the study of stellar actions
-has gained new popularity. One reason is dimensionality reduction\,---\,an
-individual stellar orbit is concisely described by three actions, as opposed to
-six phase space coordinates. Second, under the assumption of a phase-mixed
-system, the dynamical properties of a population of stars should be uniquely a
-function of their actions and independent of the conjugate angles. This allows
-one to use actions to study the relationship between \emph{orbital} properties
-of stars and other intrinsic, and, at least partially, invariant properties such
-as age or metallicity \citep{2018ApJ...867...31B, 2018arXiv180803278T,
-2018MNRAS.481.4093S, 2019arXiv190304030G, 2019arXiv190309320D,
-2019MNRAS.486.1167B}. Actions also provide a convenient basis for constructing
-models of the stellar distribution function \citep[e.g.,][]{1915MNRAS..76...70J,
-1985ApJ...295..388V, 2017ApJ...839...61T}, or for associating stars with similar
-dynamical properties, e.g., to potentially determine membership in moving
-groups.
-
-If the system being considered departs from axisymmetry in a significant and/or
-non-adiabatic way, the actions computed using an axisymmetric approximation to
-the true potential can exhibit cyclic dependence on the orbital phase (or time
-at which the star's position and velocity are observed), large-scale migration,
-or diffusion from their initial values. In the Milky Way, stellar actions are
-expected to diffuse on short time scales due to scattering with gas clouds and
-to evolve on longer time scales in the case of orbits near resonances with
-spiral arms, bar(s), and other large scale perturbations
-\citep{2014RvMP...86....1S}. For this reason, actions have been used to study
-stellar scattering in the Milky Way disk using the improved astrometry of
-\textit{Gaia} DR2 and various age catalogues \citep{2018ApJ...867...31B,
-2018arXiv180803278T}. Actions have also been used to study different models of
-spiral structure in the Milky Way \citep{2019MNRAS.tmp..155S}. Characteristics
-of the distribution of stars in the extended solar neighborhood in action space
-are discussed in \citet{2019MNRAS.484.3291T}.
+summarize its projected orbit, model the stellar distribution function, or
+find stars with similar dynamical properties. Actions are conserved quantities
+that describe the orbit of a star under the assumption of regular, bound
+orbits in a system where the equations of motion are separable in a particular
+coordinate system. They are the cyclical integral of the canonical momentum
+over its conjugate position: 
+\beq\label{eq:actions} 
+J_i \equiv \frac{1}{2\pi}
+\oint_{\text{orbit}}p_i\,\text{d}x_i\text{,} 
+\eeq 
+where $p_i$ are the conjugate momenta. Under the assumption of axisymmetry,
+$i=R,\phi,z$ are the radial, azimuthal, and vertical coordinates respectively
+in a cylindrical coordinate system. In a slowly-evolving axisymmetric
+potential, these actions are invariant and $J_{\phi} \equiv L_z$, where $L_z$
+is the $z$-component of the angular momentum
+\citep{2008gady.book.....B,2014RvMP...86....1S}.
+
+With the advent of 6D phase-space measurements over a relatively large
+($\gtrsim 2$ kpc) volume from the \textit{Gaia} satellite, the study of
+stellar actions has gained new popularity. One reason is dimensionality
+reduction\,---\,an individual stellar orbit is concisely described by three
+actions, as opposed to six phase space coordinates. Second, under the
+assumption of a phase-mixed system, the dynamical properties of a population
+of stars should be uniquely a function of their actions and independent of the
+conjugate angles. This allows one to use actions to study the relationship
+between \emph{orbital} properties of stars and other intrinsic, and, at least
+partially, invariant properties such as age or metallicity
+\citep{2018ApJ...867...31B, 2018arXiv180803278T, 2018MNRAS.481.4093S,
+2019arXiv190304030G, 2019arXiv190309320D, 2019MNRAS.486.1167B}. Actions also
+provide a convenient basis for constructing models of the stellar distribution
+function \citep[e.g.,][]{1915MNRAS..76...70J, 1985ApJ...295..388V,
+2017ApJ...839...61T}, or for associating stars with similar dynamical
+properties, e.g., to potentially determine membership in moving groups.
+
+If the system being considered departs from axisymmetry in a significant
+and/or non-adiabatic way, the actions computed using an axisymmetric
+approximation to the true potential can exhibit cyclic dependence on the
+orbital phase (or time at which the star's position and velocity are
+observed), large-scale migration, or diffusion from their initial values. In
+the Milky Way, stellar actions are expected to diffuse on short time scales
+due to scattering with gas clouds and to evolve on longer time scales in the
+case of orbits near resonances with spiral arms, bar(s), and other large scale
+perturbations \citep{2014RvMP...86....1S}. For this reason, actions have been
+used to study stellar scattering in the Milky Way disk using the improved
+astrometry of \textit{Gaia} DR2 and various age catalogues
+\citep{2018ApJ...867...31B, 2018arXiv180803278T}. Actions have also been used
+to study different models of spiral structure in the Milky Way
+\citep{2019MNRAS.tmp..155S}. Characteristics of the distribution of stars in
+the extended solar neighborhood in action space are discussed in
+\citet{2019MNRAS.484.3291T}.
 
@@
 2012ApJ...750L..41W}, but with the vast improvement in the quality of
-phase-space measurements due to \textit{Gaia} the assumption of axisymmetry
-is increasingly inadequate \citep[e.g.,][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. Even if
-this assumption were close enough for many purposes, the parameters used in
-axisymmetric models of the Galactic potential may be inadequately constrained by
-current observations.
-
-Small-scale variations in the density of gas and stars cause the
-local midplane position to vary as a
-function of radius and azimuth. 
-Stars far from the Sun have a local midplane that differs
-from our local midplane extrapolated onto their position. Converting
-positions and velocities of more distant stars from a 
-heliocentric to a Galactocentric coordinate system thus introduces a systematic
-bias in the $z$ coordinate. We show that this bias induces non-Gaussian errors
-in the actions computed for these stars. The further from the solar neighborhood
-the target star is, the more likely the mismatch will result in large systematic
-uncertainty, especially in the vertical action $J_z$. A similar argument applies
-to any remaining uncertainty in measurements of the Galactocentric radius of the
-Sun, and to variations in the LSR.
+phase-space measurements due to \textit{Gaia} the assumption of axisymmetry is
+increasingly inadequate \citep[e.g.,][]{2018Natur.561..360A,
+2019MNRAS.485.3134L}. Even if this assumption were close enough for many
+purposes, the parameters used in axisymmetric models of the Galactic potential
+may be inadequately constrained by current observations.
+
+Small-scale variations in the density of gas and stars cause the local
+midplane position to vary as a function of radius and azimuth. Stars far from
+the Sun have a local midplane that differs from our local midplane
+extrapolated onto their position. Converting positions and velocities of more
+distant stars from a heliocentric to a Galactocentric coordinate system thus
+introduces a systematic bias in the $z$ coordinate. We show that this bias
+induces non-Gaussian errors in the actions computed for these stars. The
+further from the solar neighborhood the target star is, the more likely the
+mismatch will result in large systematic uncertainty, especially in the
+vertical action $J_z$. A similar argument applies to any remaining uncertainty
+in measurements of the Galactocentric radius of the Sun, and to variations in
+the LSR.
 
@@
 \caption{Illustration showing the effect an error in the determination
-of the coordinate midplane can have on orbit integration and action estimation.
-The $x$-axis shows the orbital phase and the $y$-axis the vertical height. The
-top gray curve depicts an example ``true'' orbit oscillating about the true
-midplane (horizontal solid gray line). Consider an observer who erroneously
-assumes the midplane is located at the horizontal dashed line. Suppose this
-observer measures the phase-space position of the star at two different orbital
-phases (teal and orange points). If the observer were to then integrate the
-star's orbit using a model potential with the erroneous midplane, they would
-obtain the teal and orange curves for the star's orbit, respectively. The
-actions estimated from these two erroneous measurements would subsequently
-differ, both from each other and from the true measurement (in the potential
-with the correct coordinate system). Hence an incorrect midplane in the
-potential model assumed will induce phase dependence in the actions estimated
-for a given star in that potential.} \label{fig:cartoon} \end{figure*}
-
-In Section~\ref{sec:ref_frame}, we describe the general impact coordinate system
-errors have on the measured actions. In Section~\ref{sec:local_fire}, we examine
-the azimuthal variations of the midplane itself in examples from two classes of
-simulations: cosmological, hydrodynamical, zoom-in simulations of isolated Milky
-Way-mass galaxies from the Feedback in Realistic Environments (FIRE)
+of the coordinate midplane can have on orbit integration and action
+estimation. The $x$-axis shows the orbital phase and the $y$-axis the vertical
+height. The top gray curve depicts an example ``true'' orbit oscillating about
+the true midplane (horizontal solid gray line). Consider an observer who
+erroneously assumes the midplane is located at the horizontal dashed line.
+Suppose this observer measures the phase-space position of the star at two
+different orbital phases (teal and orange points). If the observer were to
+then integrate the star's orbit using a model potential with the erroneous
+midplane, they would obtain the teal and orange curves for the star's orbit,
+respectively. The actions estimated from these two erroneous measurements
+would subsequently differ, both from each other and from the true measurement
+(in the potential with the correct coordinate system). Hence an incorrect
+midplane in the potential model assumed will induce phase dependence in the
+actions estimated for a given star in that potential.} \label{fig:cartoon}
+\end{figure*}
+
+In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
+system errors have on the measured actions. In Section~\ref{sec:local_fire},
+we examine the azimuthal variations of the midplane itself in examples from
+two classes of simulations: cosmological, hydrodynamical, zoom-in simulations
+of isolated Milky Way-mass galaxies from the Feedback in Realistic
+Environments (FIRE)
 collaboration\footnote{\url{https://fire.northwestern.edu}}
 \citep{2014MNRAS.445..581H, 2016ApJ...827L..23W, 2018MNRAS.480..800H}, and a
-controlled N-body simulation of a Sagittarius encounter with a galaxy otherwise
-tailored to the stellar mass, scale length, and scale height of the Milky Way
-\citep{2018MNRAS.481..286L}. In Section~\ref{sec:discussion}, we discuss the
-implications of midplane variations, and the resulting systematic uncertainty in
-the vertical action, for action-space analyses. We also estimate the expected
-midplane variations of the Milky Way based on the observed velocity variations
-and three-dimensional dust maps. We summarize our main results and conclude in
-Section~\ref{sec:conclusion}.
+controlled N-body simulation of a Sagittarius encounter with a galaxy
+otherwise tailored to the stellar mass, scale length, and scale height of the
+Milky Way \citep{2018MNRAS.481..286L}. In Section~\ref{sec:discussion}, we
+discuss the implications of midplane variations, and the resulting systematic
+uncertainty in the vertical action, for action-space analyses. We also
+estimate the expected midplane variations of the Milky Way based on the
+observed velocity variations and three-dimensional dust maps. We summarize our
+main results and conclude in Section~\ref{sec:conclusion}.
 
@@
 disk-like orbits. The consequences we explore here may also arise from various
-other systematic errors. For instance, the axisymmetric Galactic potential model
-used in many works to compute actions may not be a good description of the true
-potential\,---\,or the parameters used may yield a potential that is
+other systematic errors. For instance, the axisymmetric Galactic potential
+model used in many works to compute actions may not be a good description of
+the true potential\,---\,or the parameters used may yield a potential that is
 systematically incorrect outside an original fitted region. In this work, we
 assume that the Galaxy is perfectly described by our model axisymmetric
-potential, and simply explore the consequences of offsets in the Galactocentric
-coordinate system.
+potential, and simply explore the consequences of offsets in the
+Galactocentric coordinate system.
 
 \subsection{Effect of Midplane Offset} \label{ssec:cartoon}
-We present an illustration of an orbit in Figure~\ref{fig:cartoon} to
-show how an inaccurate or erroneous determination of the midplane leads to a
-dependence on orbital phase of the value of the actions calculated
-from a point in phase space and an
-assumed potential model. The $x$-axis corresponds to orbital phase and the
-$y$-axis to vertical height. The solid gray curve indicates the 
-true orbit of the star as it oscillates around the true midplane. The dashed gray line, offset from the true midplane, is the midplane location
-used by an observer to integrate the orbit of the star and estimate its actions.
-The model potential is otherwise identical to the one in which the star is
+We present an illustration of an orbit in Figure~\ref{fig:cartoon} to show how
+an inaccurate or erroneous determination of the midplane leads to a dependence
+on orbital phase of the value of the actions calculated from a point in phase
+space and an assumed potential model. The $x$-axis corresponds to orbital
+phase and the $y$-axis to vertical height. The solid gray curve indicates the
+true orbit of the star as it oscillates around the true midplane. The dashed
+gray line, offset from the true midplane, is the midplane location used by an
+observer to integrate the orbit of the star and estimate its actions. The
+model potential is otherwise identical to the one in which the star is
 actually moving.
@@
 observer would compute for each point based on the potential model with the
-offset midplane. In action space, this would correspond to a different value of
-$J_z$ for the teal and orange points. In this way, assuming the wrong coordinate
-system induces a phase-dependence in the actions estimated for the star, which
-in the correct potential (in this example, the one with the correct midplane)
-should be phase independent.
-
-This example uses an offset in $z$, but 
-analagous effects occur from offsets in other coordinates, 
-such as the distance to the Galactic
-center or the LSR. A similar effect in which actions gain time-dependence due to
-a time-varying potential was pointed out by \citet{2015A&A...584A.120B}.
+offset midplane. In action space, this would correspond to a different value
+of $J_z$ for the teal and orange points. In this way, assuming the wrong
+coordinate system induces a phase-dependence in the actions estimated for the
+star, which in the correct potential (in this example, the one with the
+correct midplane) should be phase independent.
+
+This example uses an offset in $z$, but analagous effects occur from offsets
+in other coordinates, such as the distance to the Galactic center or the LSR.
+A similar effect in which actions gain time-dependence due to a time-varying
+potential was pointed out by \citet{2015A&A...584A.120B}.
 
@@
 Before turning to numerical methods, we derive analytic expressions for the
-systematic error in the actions induced from offsets in the position (e.g., the midplane or
-Galactic center distance) or velocity (e.g., the 
-LSR) 
-of the assumed coordinate system's origin.
-
-We use the epicyclic approximation, which
-assumes that the motion in the $z$ and $R$ components of the orbit are
-decoupled and follow simple harmonic motion about a circular and planar
-guiding orbit 
-\citep[][Section~3.2 and references
-therein]{2008gady.book.....B}. The radius of this orbit is referred to
-as the guiding radius $R_g$.
-This approximation is an excellent description
-of the thin disk and a good description of the thick disk in
-an axisymmetric potential
-that ignores
-the influence of the Galactic bar and
-spiral arms. We also make the assumption of a perfectly flat circular
-velocity curve with $v_c(R) = v_c$, a good approximation near the solar circle
+systematic error in the actions induced from offsets in the position (e.g.,
+the midplane or Galactic center distance) or velocity (e.g., the LSR) of the
+assumed coordinate system's origin.
+
+We use the epicyclic approximation, which assumes that the motion in the $z$
+and $R$ components of the orbit are decoupled and follow simple harmonic
+motion about a circular and planar guiding orbit \citep[][Section~3.2 and
+references therein]{2008gady.book.....B}. The radius of this orbit is referred
+to as the guiding radius $R_g$. This approximation is an excellent description
+of the thin disk and a good description of the thick disk in an axisymmetric
+potential that ignores the influence of the Galactic bar and spiral arms. We
+also make the assumption of a perfectly flat circular velocity curve with
+$v_c(R) = v_c$, a good approximation near the solar circle
 \citep[e.g.,][]{2017MNRAS.465...76M}.
@@
 where $\kappa$ and $\nu$ are the radial/epicyclic and vertical frequencies, $\Omega_c
-\equiv v_c/R_g$ is the orbital frequency of the guiding center, $A_R$ and $A_z$ are the amplitudes in the radial and vertical coordinates, and $\alpha$ and
-$\beta$ are 
-the initial orbital phases.
-Similarly, the velocities of the orbit are given by:
+\equiv v_c/R_g$ is the orbital frequency of the guiding center, $A_R$ and
+$A_z$ are the amplitudes in the radial and vertical coordinates, and $\alpha$
+and $\beta$ are the initial orbital phases. Similarly, the velocities of the
+orbit are given by:
 \beq\label{eq:orbits_vel_epi}
@@
 
-Since most of the time stars will be at maximum amplitude (i.e., turnaround) in
-both $R$ and $z$, we can approximate the order of magnitude of the systematic
-error in the actions by
+Since most of the time stars will be at maximum amplitude (i.e., turnaround)
+in both $R$ and $z$, we can approximate the order of magnitude of the
+systematic error in the actions by
 \beq\label{eq:Ji_err_mosttime}
@@
 using the code \texttt{gala}~v0.3 to perform orbit integrations and conversion
-to action space \citep{gala, gala:zenodo}. To compute
-actions we use the torus-mapping technique first presented by
-\citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to
-calculate actions for an orbital time-series starting from a phase-space
-position $(x, v)$ and integrated in a potential $\Phi$. For our Galactic potential we use \texttt{MWPotential}, based on the Milky Way
-potential available in \texttt{galpy} \citep{2015ApJS..216...29B}, which includes a Hernquist bulge and nucleus \citep{1990ApJ...356..359H},
-a Miyamoto--Nagai disk \citep{1975PASJ...27..533M}, and 
-a Navarro, Frenk, \& White \citeyear{1997ApJ...490..493N} halo,
-and is fit to empirically match some
-observations.
-We use the Dormand-Prince 8(5,3) integration scheme
+to action space \citep{gala, gala:zenodo}. To compute actions we use the
+torus-mapping technique first presented by \citet{1990MNRAS.244..634M} and
+adapted by \citet{2014MNRAS.441.3284S} to calculate actions for an orbital
+time-series starting from a phase-space position $(x, v)$ and integrated in a
+potential $\Phi$. For our Galactic potential we use \texttt{MWPotential},
+based on the Milky Way potential available in \texttt{galpy}
+\citep{2015ApJS..216...29B}, which includes a Hernquist bulge and nucleus
+\citep{1990ApJ...356..359H}, a Miyamoto--Nagai disk
+\citep{1975PASJ...27..533M}, and a Navarro, Frenk, \& White
+\citeyear{1997ApJ...490..493N} halo, and is fit to empirically match some
+observations. We use the Dormand-Prince 8(5,3) integration scheme
 \citep{Dormand80:integrator} with a timestep of $1\,\Myr$ and integrate for
@@
 recent works exploring actions in the Galactic disk
-\citep[e.g.,][]{2019MNRAS.484.3291T, 2018MNRAS.481.4093S, 2018arXiv180803278T}.
-For disk-like orbits, existing implementations of the St\"ackel Fudge method
-are of acceptable accuracy, but since we also consider halo-like orbits in
-this work (where the St\"ackel Fudge method is inaccurate) we choose to use
-orbit integration and torus mapping throughout \citep{2016MNRAS.457.2107S}.
+\citep[e.g.,][]{2019MNRAS.484.3291T, 2018MNRAS.481.4093S,
+2018arXiv180803278T}. For disk-like orbits, existing implementations of the
+St\"ackel Fudge method are of acceptable accuracy, but since we also consider
+halo-like orbits in this work (where the St\"ackel Fudge method is inaccurate)
+we choose to use orbit integration and torus mapping throughout
+\citep{2016MNRAS.457.2107S}.
 
@@
 error in the Galactocentric coordinate system. We consider here \thick{},
-which has actions of $(J_R, J_{\phi}, J_z) = (38, -1500,
-7.0)\,\actunit$ and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We
-integrate the orbit according to the procedure laid out in
-Section~\ref{ssec:action_comp}, and which we plot in
-Appendix~\ref{app:orbits}. Then, we subtract $100\,\pc$ from the $z$ value
-(upper panels) or the $x$ value (lower panels) of each position in the orbit,
-corresponding to an erroneous observer assuming a midplane (upper) or solar
-radius (lower) that is off by $100\,\pc$. We then allow an observer
-to measure the orbit over $1\,\Gyr$ and perform the same orbit integration
-procedure at each timestep, and report the values of the actions, with the true values given as horizontal dashed lines. The
-computation of $J_{\phi}$ is pristine to errors in $z$, with only numerical
-artifacts remaining. Only small errors are induced in $J_R$, with the middle
-$90\%$ of values over the $\Gyr$ being within $\sim8\%$ of the true $J_R$. As
-expected, large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$,
-with the middle $90\%$ of values being within $\sim43\%$ of the true $J_z$.
-The $x$ offset induces uncertainties in $J_R$, $J_{\phi}$, and $J_z$ of
-$\sim21\%$, $\sim3\%$, and $\sim3\%$.}
+which has actions of $(J_R, J_{\phi}, J_z) = (38, -1500, 7.0)\,\actunit$ and
+$z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We integrate the orbit
+according to the procedure laid out in Section~\ref{ssec:action_comp}, and
+which we plot in Appendix~\ref{app:orbits}. Then, we subtract $100\,\pc$ from
+the $z$ value (upper panels) or the $x$ value (lower panels) of each position
+in the orbit, corresponding to an erroneous observer assuming a midplane
+(upper) or solar radius (lower) that is off by $100\,\pc$. We then allow an
+observer to measure the orbit over $1\,\Gyr$ and perform the same orbit
+integration procedure at each timestep, and report the values of the actions,
+with the true values given as horizontal dashed lines. The computation of
+$J_{\phi}$ is pristine to errors in $z$, with only numerical artifacts
+remaining. Only small errors are induced in $J_R$, with the middle $90\%$ of
+values over the $\Gyr$ being within $\sim8\%$ of the true $J_R$. As expected,
+large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$, with the
+middle $90\%$ of values being within $\sim43\%$ of the true $J_z$. The $x$
+offset induces uncertainties in $J_R$, $J_{\phi}$, and $J_z$ of $\sim21\%$,
+$\sim3\%$, and $\sim3\%$.}
 \label{fig:one_orbit_wrong_ref}
@@
 \end{center}
-\caption{A histogram of the computed values of $J_z$ at different orbital phases for \thick{} (top
-panel) and \thin{} (bottom panel) assuming a $z$ offset of
-$100\,\pc$. One can see that if the observed $z$ values have a bias (from, e.g.,
-an incorrectly computed midplane), then the induced error distribution in
-$J_z$ is decidedly non-Gaussian. Therefore, any sort of error propagation must
-take this into account. 
-A heuristic explanation
-for the shape of each panel is given in the
-text. We also plot one half the 95th percentile minus the 5th
-percentile of each distribution as a horizontal arrow anchored on the true
-$J_z$ value. We call this $\Delta J_z$ and will use it (along with the
-similarly defined $\Delta J_R$ and $\Delta J_{\phi}$) to empirically describe
-the error distribution. We see that $\Delta J_z$ roughly corresponds to the
-distance from the true $J_z$ value to one of the modes of the distribution of
-computed $J_z$ values. Similar plots for $J_R$ induced by a $z$ offset and
-$J_R$ and $J_{\phi}$ induced by an $x$ offset are given in
+\caption{A histogram of the computed values of $J_z$ at different orbital
+phases for \thick{} (top panel) and \thin{} (bottom panel) assuming a $z$
+offset of $100\,\pc$. One can see that if the observed $z$ values have a bias
+(from, e.g., an incorrectly computed midplane), then the induced error
+distribution in $J_z$ is decidedly non-Gaussian. Therefore, any sort of error
+propagation must take this into account. A heuristic explanation for the shape
+of each panel is given in the text. We also plot one half the 95th percentile
+minus the 5th percentile of each distribution as a horizontal arrow anchored
+on the true $J_z$ value. We call this $\Delta J_z$ and will use it (along with
+the similarly defined $\Delta J_R$ and $\Delta J_{\phi}$) to empirically
+describe the error distribution. We see that $\Delta J_z$ roughly corresponds
+to the distance from the true $J_z$ value to one of the modes of the
+distribution of computed $J_z$ values. Similar plots for $J_R$ induced by a
+$z$ offset and $J_R$ and $J_{\phi}$ induced by an $x$ offset are given in
 Appendix~\ref{app:hist}.}
@@
 We now quantify how a systematic error in the Galactocentric coordinate system
-induces phase-dependence in the actions calculated from the observed position and velocity of a star. We consider
-three orbits in the model potential described in
-Section~\ref{ssec:action_comp} that are typical of stars in the thin
-disk, thick disk, and halo. We summarize their initial
-positions in phase space, the actions computed by integrating their orbits
-in the correct potential, and other properties in Table~\ref{tab:orbits}. Each orbit, integrated
-without systematic coordinate errors, is plotted in Appendix~\ref{app:orbits}.
-We will refer to these orbits by their names (\thin{}, \thick{}, \halo{})
-henceforth.
-
-We begin with \thick{}. Consider an observer who can
-measure the orbit's phase-space position at many different times
-(and hence different orbital phases), but does so using a
-coordinate system in which the midplane is systematically offset in height by
-$100\,\pc$ from its actual location. To model this we subtract the vector $(0, 0,
-100)\,\pc$ from each position in the orbit. This corresponds to an
-observer physically located at, e.g., the position $(8, 0, 0)\,\kpc$ in the
-coordinate system of the true potential, but erroneously thinking they are
-located at $(8, 0, 0.1)\,\kpc$.
+induces phase-dependence in the actions calculated from the observed position
+and velocity of a star. We consider three orbits in the model potential
+described in Section~\ref{ssec:action_comp} that are typical of stars in the
+thin disk, thick disk, and halo. We summarize their initial positions in phase
+space, the actions computed by integrating their orbits in the correct
+potential, and other properties in Table~\ref{tab:orbits}. Each orbit,
+integrated without systematic coordinate errors, is plotted in
+Appendix~\ref{app:orbits}. We will refer to these orbits by their names
+(\thin{}, \thick{}, \halo{}) henceforth.
+
+We begin with \thick{}. Consider an observer who can measure the orbit's
+phase-space position at many different times (and hence different orbital
+phases), but does so using a coordinate system in which the midplane is
+systematically offset in height by $100\,\pc$ from its actual location. To
+model this we subtract the vector $(0, 0, 100)\,\pc$ from each position in the
+orbit. This corresponds to an observer physically located at, e.g., the
+position $(8, 0, 0)\,\kpc$ in the coordinate system of the true potential, but
+erroneously thinking they are located at $(8, 0, 0.1)\,\kpc$.
 
 We consider the observer making a measurement, integrating an orbit, and
-computing actions every 
-megayear 
-using the prescription above. However, we
+computing actions every megayear using the prescription above. However, we
 specify the star's starting position using the systematically offset
-coordinate system.
-Essentially we are shifting and then reintegrating at each point along the original orbit.
-The actions computed using the offset coordinate system
-for each phase-space starting point are shown for the first gigayear 
-of the orbit in the upper panels of
+coordinate system. Essentially we are shifting and then reintegrating at each
+point along the original orbit. The actions computed using the offset
+coordinate system for each phase-space starting point are shown for the first
+gigayear of the orbit in the upper panels of
 Figure~\ref{fig:one_orbit_wrong_ref}.\footnote{Occasionally the numerical
-scheme fails and very large actions are reported by \texttt{gala}\,---\,we perform
-a $4\sigma$ clip on each action to exclude such orbits, but this only excludes
-a total of $5$ orbits out of the $1000$ considered for
+scheme fails and very large actions are reported by \texttt{gala}\,---\,we
+perform a $4\sigma$ clip on each action to exclude such orbits, but this only
+excludes a total of $5$ orbits out of the $1000$ considered for
 Figure~\ref{fig:one_orbit_wrong_ref}. Some numerical artifacts remain, but the
@@
 Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the
-offset coordinate systems oscillate
-as a function of the time/orbital phase at which the
-star's phase-space position is observed.
-This time dependence comes even though
-the observer is using the correctly constructed, best-fit, static, axisymmetric
-potential. The relative size of the phase variation in each action depends on
-the direction of the systematic offset as well as the true values of the
-actions (i.e. the type of orbit). In reality, we will have one measurement of
-the phase-space position to work with, in which case the determination of the
-orbital phase in $R$ or $z$ is degenerate with the degree of systematic offset
-in that coordinate (see Figure~\ref{fig:cartoon}). In the following we
-therefore quote percentile ranges for the possible values computed for each
-action as a proxy for the effect of these systematic errors in the coordinate
-system.
+offset coordinate systems oscillate as a function of the time/orbital phase at
+which the star's phase-space position is observed. This time dependence comes
+even though the observer is using the correctly constructed, best-fit, static,
+axisymmetric potential. The relative size of the phase variation in each
+action depends on the direction of the systematic offset as well as the true
+values of the actions (i.e. the type of orbit). In reality, we will have one
+measurement of the phase-space position to work with, in which case the
+determination of the orbital phase in $R$ or $z$ is degenerate with the degree
+of systematic offset in that coordinate (see Figure~\ref{fig:cartoon}). In the
+following we therefore quote percentile ranges for the possible values
+computed for each action as a proxy for the effect of these systematic errors
+in the coordinate system.
 
@@
 minus 5\textsuperscript{th} percentiles are $2.2$ and $6.2\,\actunit$ for
-$J_R$ and $J_z$, respectively. As a fraction of the true values, these are $5.7\%$ and
-$86\%$, respectively. The spread induced in $J_{\phi}$ is negligible, as
-expected since $J_{\phi}$ only depends on the $x$- and $y$-components of the
-position and velocity of the stars.\footnote{In practice, however, $J_{\phi}$
-is computed as part of the torus-fitting method.} It is worth pointing out
-that a $100\,\pc$ offset in an orbit with $z_{\text{max}}=850\,\pc$\,---\,a
-$12\%$ error\,---\,induced an $86\%$ spread in the computation of $J_z$.
+$J_R$ and $J_z$, respectively. As a fraction of the true values, these are
+$5.7\%$ and $86\%$, respectively. The spread induced in $J_{\phi}$ is
+negligible, as expected since $J_{\phi}$ only depends on the $x$- and
+$y$-components of the position and velocity of the stars.\footnote{In
+practice, however, $J_{\phi}$ is computed as part of the torus-fitting
+method.} It is worth pointing out that a $100\,\pc$ offset in an orbit with
+$z_{\text{max}}=850\,\pc$\,---\,a $12\%$ error\,---\,induced an $86\%$ spread
+in the computation of $J_z$.
 
@@
 $47$, and $0.71\,\actunit$ for $J_R$, $J_{\phi}$ and $J_z$, respectively.
-These are fractionally $21\%$, $3.1\%$ and $3.1\%$ in these actions, respectively, despite only a $1.2\%$ error in the distance to the Galactic center.
-
-In Figure~\ref{fig:Jz_hist}, we plot a histogram of the values of
-$J_z$ computed at different orbital phases for \thick{} (top panel) and \thin{} (bottom panel),
-assuming a $z$ offset of $100\,\pc$ (as in the upper right panel of
-Figure~\ref{fig:one_orbit_wrong_ref}). The true value is plotted as a
-vertical dashed line. The systematic error in $J_z$ induced by a
-systematic offset in $z$ is non-Gaussian and bimodal; neither of the modes is centered on the null value. In the case of \thin{}
-(bottom panel), we see that, in addition to the prior complications, the
-distribution is not even centered on the true value. This comes about when the midplane
-error is~$\gtrsim z_{\text{max}}$, where $z_{\text{max}}$ is the maximum
-height of the orbit (equivalent to $A_z$ in the epicyclic approximation, see
+These are fractionally $21\%$, $3.1\%$ and $3.1\%$ in these actions,
+respectively, despite only a $1.2\%$ error in the distance to the Galactic
+center.
+
+In Figure~\ref{fig:Jz_hist}, we plot a histogram of the values of $J_z$
+computed at different orbital phases for \thick{} (top panel) and \thin{}
+(bottom panel), assuming a $z$ offset of $100\,\pc$ (as in the upper right
+panel of Figure~\ref{fig:one_orbit_wrong_ref}). The true value is plotted as a
+vertical dashed line. The systematic error in $J_z$ induced by a systematic
+offset in $z$ is non-Gaussian and bimodal; neither of the modes is centered on
+the null value. In the case of \thin{} (bottom panel), we see that, in
+addition to the prior complications, the distribution is not even centered on
+the true value. This comes about when the midplane error is~$\gtrsim
+z_{\text{max}}$, where $z_{\text{max}}$ is the maximum height of the orbit
+(equivalent to $A_z$ in the epicyclic approximation, see
 Section~\ref{ssec:epi_action}).
@@
 In Appendix~\ref{app:hist} we plot the same histogram as in
-Figure~\ref{fig:Jz_hist}, but for the distributions of $J_R$ induced
-by a $z$ offset (upper left panel of Figure~\ref{fig:many_orbit_wrong_ref})
-and the distributions of $J_R$ and $J_{\phi}$ errors induced by an
-$x$ offset (lower left and lower center panels of
-Figure~\ref{fig:many_orbit_wrong_ref}, respectively). We find similar error
-distributions as in Figure~\ref{fig:Jz_hist}, with the exception that the
-computed $J_R$ distribution induced by an $x$ offset more closely resembles a
-Gaussian distribution.
+Figure~\ref{fig:Jz_hist}, but for the distributions of $J_R$ induced by a $z$
+offset (upper left panel of Figure~\ref{fig:many_orbit_wrong_ref}) and the
+distributions of $J_R$ and $J_{\phi}$ errors induced by an $x$ offset (lower
+left and lower center panels of Figure~\ref{fig:many_orbit_wrong_ref},
+respectively). We find similar error distributions as in
+Figure~\ref{fig:Jz_hist}, with the exception that the computed $J_R$
+distribution induced by an $x$ offset more closely resembles a Gaussian
+distribution.
 
@@
 \end{center}
-\caption{We report the fractional error in each action $J_i$ induced by coordinate system offsets for \thin{}, \thick{}, and \halo{} (Table~\ref{tab:orbits}).
-The error ($\Delta J_i$) is computed as one half the 95th minus 5th percentile of the distribution of computed action values.
-See discussion in the
-text and Figure~\ref{fig:Jz_hist} for the justification in using this to measure the magnitude of the
-induced error. The left, center, and right panels show the result for $J_R$,
-$J_{\phi}$, and $J_z$, respectively. The upper panels consider an offset in
-$z$ and the lower panels consider an offset in $x$ (equivalently, an offset in
-the solar radius). In some panels, we also plot as dashed lines the epicyclic
-prediction of the induced action error (Equation~\eqref{eq:Ji_err_mosttime}).
-In the epicyclic approximation, a $z$ offset only induces an error in $J_z$\,---\,for all three orbits the epicyclic approximation is a good description of
-the $J_z$ error. An $x$ offset induces an error in $J_R$ and $J_{\phi}$. The
-error in $J_R$ is somewhat well-described for \thin{} and \thick{},
-and a poor description for \halo{}. For $J_{\phi}$, the
-epicyclic approximation is not a good description for any orbit.}
+\caption{We report the fractional error in each action $J_i$ induced by
+coordinate system offsets for \thin{}, \thick{}, and \halo{}
+(Table~\ref{tab:orbits}). The error ($\Delta J_i$) is computed as one half the
+95th minus 5th percentile of the distribution of computed action values. See
+discussion in the text and Figure~\ref{fig:Jz_hist} for the justification in
+using this to measure the magnitude of the induced error. The left, center,
+and right panels show the result for $J_R$, $J_{\phi}$, and $J_z$,
+respectively. The upper panels consider an offset in $z$ and the lower panels
+consider an offset in $x$ (equivalently, an offset in the solar radius). In
+some panels, we also plot as dashed lines the epicyclic prediction of the
+induced action error (Equation~\eqref{eq:Ji_err_mosttime}). In the epicyclic
+approximation, a $z$ offset only induces an error in $J_z$\,---\,for all three
+orbits the epicyclic approximation is a good description of the $J_z$ error.
+An $x$ offset induces an error in $J_R$ and $J_{\phi}$. The error in $J_R$ is
+somewhat well-described for \thin{} and \thick{}, and a poor description for
+\halo{}. For $J_{\phi}$, the epicyclic approximation is not a good description
+for any orbit.}
 \label{fig:many_orbit_wrong_ref}
@@
 
-We now suggest a heuristic explanation for
-the shape of Figure~\ref{fig:Jz_hist}.
-Consider first the \thick{} (top panel), where the offset in $z$ is
-much less than the $z_{\text{max}}$ of the orbit. The peaks in the distribution
-correspond to the turning points of the orbit (or points of maximum vertical
-amplitude), where $v_z \sim 0$ and where the star is on most of it orbit. This
-is why the distribution, which is calculated at evenly spaced time intervals, peaks at these values. For \thin{}
-(bottom panel), the offset in $z$ is comparable to $z_{\text{max}}$. Now,
-there will be some points in the orbit where $v_z = 0$ and $z=0$ (in the
-erroneous coordinate system). At these points, the computed $J_z$
-will vanish. The asymmetry and systematic offset then comes about because of
-the constraint that $J_z \geq 0$.\footnote{This argument is similar to ones given in
-cosmology for why gravity produces non-Gaussianity in the density field, since
-the density cannot become negative but it can grow arbitrarily large.}
-
-Gaussian summary statistics are clearly insufficient to describe the distribution shown in
-Figure~\ref{fig:Jz_hist}. We therefore elect to measure this error by computing one
-half the 95th percentile minus the 5th percentile of the distribution
-of action values. We refer to this quantity as $\Delta J_i$ for each action and plot it in Figure~\ref{fig:Jz_hist} as a horizontal arrow anchored on the true action value.
- Because of the bimodality
-of the error distribution, this quantity roughly measures the distance from
-the true action value to the peak of one of the modes. Furthermore, this
-bimodality also implies that $\Delta J_i$ is not very sensitive to the exact
-percentiles used. This summary statistic does not reflect the bias induced
-when the midplane error is~$\gtrsim z_{\text{max}}$.
+We now suggest a heuristic explanation for the shape of
+Figure~\ref{fig:Jz_hist}. Consider first the \thick{} (top panel), where the
+offset in $z$ is much less than the $z_{\text{max}}$ of the orbit. The peaks
+in the distribution correspond to the turning points of the orbit (or points
+of maximum vertical amplitude), where $v_z \sim 0$ and where the star is on
+most of it orbit. This is why the distribution, which is calculated at evenly
+spaced time intervals, peaks at these values. For \thin{} (bottom panel), the
+offset in $z$ is comparable to $z_{\text{max}}$. Now, there will be some
+points in the orbit where $v_z = 0$ and $z=0$ (in the erroneous coordinate
+system). At these points, the computed $J_z$ will vanish. The asymmetry and
+systematic offset then comes about because of the constraint that $J_z \geq
+0$.\footnote{This argument is similar to ones given in cosmology for why
+gravity produces non-Gaussianity in the density field, since the density
+cannot become negative but it can grow arbitrarily large.}
+
+Gaussian summary statistics are clearly insufficient to describe the
+distribution shown in Figure~\ref{fig:Jz_hist}. We therefore elect to measure
+this error by computing one half the 95th percentile minus the 5th percentile
+of the distribution of action values. We refer to this quantity as $\Delta
+J_i$ for each action and plot it in Figure~\ref{fig:Jz_hist} as a horizontal
+arrow anchored on the true action value. Because of the bimodality of the
+error distribution, this quantity roughly measures the distance from the true
+action value to the peak of one of the modes. Furthermore, this bimodality
+also implies that $\Delta J_i$ is not very sensitive to the exact percentiles
+used. This summary statistic does not reflect the bias induced when the
+midplane error is~$\gtrsim z_{\text{max}}$.
 
@@
 The upper panels of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread
-induced in each action for an offset in the $z$-component. In the lower panels we consider offsets in the $x$ component (i.e. the
-solar radius). The left, center, and right columns show the fractional spread
-in the values computed for $J_z$, $J_{\phi}$, and $J_R$, respectively.
+induced in each action for an offset in the $z$-component. In the lower panels
+we consider offsets in the $x$ component (i.e. the solar radius). The left,
+center, and right columns show the fractional spread in the values computed
+for $J_z$, $J_{\phi}$, and $J_R$, respectively.
 
@@
 $z$, as discussed earlier. Indeed, the result we found in
-Figure~\ref{fig:one_orbit_wrong_ref} for \thick{} holds for all
-orbit types. This is also a demonstration of the robustness of the
-integration and action calculation methods we use.
+Figure~\ref{fig:one_orbit_wrong_ref} for \thick{} holds for all orbit types.
+This is also a demonstration of the robustness of the integration and action
+calculation methods we use.
 
@@
 fractional error in $J_z$ is more exaggerated for more planar (disk-like)
-orbits. For \thin{}, 
-a systematic offset of $15\,\pc$
-in the $z$-coordinate results in a $25\%$ deviation in $J_z$,
-while a $120\,\pc$ offset results in the same deviation for \thick{}.
-We find that \halo{} is relatively resistant to errors in the midplane, with
-only $\sim15\%$ error in $J_z$ out to an offset of $500\,\pc$.
+orbits. For \thin{}, a systematic offset of $15\,\pc$ in the $z$-coordinate
+results in a $25\%$ deviation in $J_z$, while a $120\,\pc$ offset results in
+the same deviation for \thick{}. We find that \halo{} is relatively resistant
+to errors in the midplane, with only $\sim15\%$ error in $J_z$ out to an
+offset of $500\,\pc$.
 
@@
 v_{\phi}=0$, as dashed lines in the color of each orbit. This equation is
-relevant since during most of the orbit the star will be close to 
-maximum radial and
-vertical amplitude. Note that we
-consider an error in the $x$-coordinate $\Delta x$, which is not exactly the
-same as $\Delta R$. For observations of stars close to us, we have that
-$\Delta x \sim \Delta R$, but for the experiment performed in this section we
-consider observations of the star throughout its entire orbit. This introduces
-a factor of $2/\pi$ when converting from $\Delta x$ to $\Delta R$, which we
-derive in Appendix~\ref{app:deltax}.
+relevant since during most of the orbit the star will be close to maximum
+radial and vertical amplitude. Note that we consider an error in the
+$x$-coordinate $\Delta x$, which is not exactly the same as $\Delta R$. For
+observations of stars close to us, we have that $\Delta x \sim \Delta R$, but
+for the experiment performed in this section we consider observations of the
+star throughout its entire orbit. This introduces a factor of $2/\pi$ when
+converting from $\Delta x$ to $\Delta R$, which we derive in
+Appendix~\ref{app:deltax}.
 
@@
 \thick{}, and so we actually expect the epicyclic approximation to perform
-slightly worse for \thin{} in this case. The epicyclic approximation underpredicts $\Delta J_{\phi}$
-for all orbits.
+slightly worse for \thin{} in this case. The epicyclic approximation
+underpredicts $\Delta J_{\phi}$ for all orbits.
 
@@
 different offsets in $z$. All orbits have the same initial position of $(8, 0,
-0)\,\kpc$ and velocity $(0, -190, v_z)\,\kms$, where we vary $v_z$.\footref{note:vz_orbits} We show
-this for a $z$ offset of $10$, $50$, and $100\,\pc$ (orange, teal, and green,
-respectively). As before, the error ($\Delta J_z$) is one half the 95th
-minus 5th percentile of the distribution of computed $J_z$ values over the
-course of the orbit. There are large errors for \thin{}-like orbits ($J_z
-\sim 0.7\,\actunit$), even for a small midplane offset of $10\,\pc$. As dashed
-lines in each color we also plot the prediction for $\Delta J_z/J_z$ from the
-epicyclic approximation (Equation~\eqref{eq:Ji_err_mosttime}), which shows
-excellent agreement with the numerically computed values.}
+0)\,\kpc$ and velocity $(0, -190, v_z)\,\kms$, where we vary
+$v_z$.\footref{note:vz_orbits} We show this for a $z$ offset of $10$, $50$,
+and $100\,\pc$ (orange, teal, and green, respectively). As before, the error
+($\Delta J_z$) is one half the 95th minus 5th percentile of the distribution
+of computed $J_z$ values over the course of the orbit. There are large errors
+for \thin{}-like orbits ($J_z \sim 0.7\,\actunit$), even for a small midplane
+offset of $10\,\pc$. As dashed lines in each color we also plot the prediction
+for $\Delta J_z/J_z$ from the epicyclic approximation
+(Equation~\eqref{eq:Ji_err_mosttime}), which shows excellent agreement with
+the numerically computed values.}
 \label{fig:dJz_fun_Jz}
@@
 To further understand the effect of the midplane error, we also plot the
-fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of $10$, $50$, and $
-100\,\pc$ (orange, teal, and green, respectively) in
-Figure~\ref{fig:dJz_fun_Jz}. 
-For each orbit, we set the initial position to be
+fractional error in $J_z$ as a function of $J_z$ for $z$ offsets of $10$,
+$50$, and $ 100\,\pc$ (orange, teal, and green, respectively) in
+Figure~\ref{fig:dJz_fun_Jz}. For each orbit, we set the initial position to be
 $(8,0,0)\,\kpc$ and the initial velocity to be $(0, -190, v_z)\,\kms$, where
 we vary $v_z$.\footnote{\label{note:vz_orbits}One can recover \thin{},
-\thick{}, and \halo{} by setting $v_z=10$, $50$, and $190$,
- respectively, 
-giving $J_z \simeq 0.7, 20,$ and $500\,\actunit$
-(see Table~\ref{tab:orbits}).} For a \thin{}-like orbit ($J_z\sim0.7\,\actunit$), 
-even a
-$10\,\pc$ offset in $z$ is enough to induce a 
-$\sim20\%$ 
-error in $J_z$. For
-larger values of $J_z$, the fractional errors are suppressed, but the induced
-error can still be large depending on how great the $z$ offset is. We also
-plot the epicylic prediction for $\Delta J_z / J_z$ from
+\thick{}, and \halo{} by setting $v_z=10$, $50$, and $190$, respectively,
+giving $J_z \simeq 0.7, 20,$ and $500\,\actunit$ (see
+Table~\ref{tab:orbits}).} For a \thin{}-like orbit ($J_z\sim0.7\,\actunit$),
+even a $10\,\pc$ offset in $z$ is enough to induce a $\sim20\%$ error in
+$J_z$. For larger values of $J_z$, the fractional errors are suppressed, but
+the induced error can still be large depending on how great the $z$ offset is.
+We also plot the epicylic prediction for $\Delta J_z / J_z$ from
 Equation~\eqref{eq:Ji_err_mosttime} as dashed lines for each $z$ offset. We
@@
 
-\section{
-Azimuthal Midplane Variations}
+\section{Azimuthal Midplane Variations}
 \label{sec:local_fire}
@@
 
-We specifically consider \emph{azimuthal} variations in the midplane at the solar circle, as defined by the stellar mass density.
-Since, to our knowledge, there are no direct empirical measurements of these variations in the Milky Way, we use example simulations from two classes of simulations to estimate the size of this effect.
-
-One set are three zoom-in, cosmological hydrodynamical simulations of isolated Milky Way-mass galaxies from the FIRE collaboration, described briefly in Section~\ref{ssec:cosmozoom}.
-These include stars, gas, and dark matter in a fully cosmological setting but are not tailored to specific properties of the Milky Way (such as the scale height or scale length, or the details of the accretion history).
-We use these simulations to span the range of possibilities for azimuthal midplane variations.
-
-The other set of simulations are isolated N-body simulations of interactions between the Milky Way and a Sagittarius-like dwarf galaxy, described briefly in Section~\ref{ssec:sag_sim}.
-These include dark matter and stars and are tailored to existing measurements of the structure of the Milky Way's disk and of the orbit and properties of the Sagittarius dwarf galaxy.
-Comparing the azimuthal midplane variations in the host galaxy of these simulations before and after the interaction with the Sagittarius-like object gives an idea of the effect of one minor merger whose properties are relatively well measured.
-Azimuthal variations of the mean vertical height of stars has been explicitly pointed out in a different simulation of a Sagittarius-like encounter by \citet{2013MNRAS.429..159G}.
+We specifically consider \emph{azimuthal} variations in the midplane at the
+solar circle, as defined by the stellar mass density. Since, to our knowledge,
+there are no direct empirical measurements of these variations in the Milky
+Way, we use example simulations from two classes of simulations to estimate
+the size of this effect.
+
+One set are three zoom-in, cosmological hydrodynamical simulations of isolated
+Milky Way-mass galaxies from the FIRE collaboration, described briefly in
+Section~\ref{ssec:cosmozoom}. These include stars, gas, and dark matter in a
+fully cosmological setting but are not tailored to specific properties of the
+Milky Way (such as the scale height or scale length, or the details of the
+accretion history). We use these simulations to span the range of
+possibilities for azimuthal midplane variations.
+
+The other set of simulations are isolated N-body simulations of interactions
+between the Milky Way and a Sagittarius-like dwarf galaxy, described briefly
+in Section~\ref{ssec:sag_sim}. These include dark matter and stars and are
+tailored to existing measurements of the structure of the Milky Way's disk and
+of the orbit and properties of the Sagittarius dwarf galaxy. Comparing the
+azimuthal midplane variations in the host galaxy of these simulations before
+and after the interaction with the Sagittarius-like object gives an idea of
+the effect of one minor merger whose properties are relatively well measured.
+Azimuthal variations of the mean vertical height of stars has been explicitly
+pointed out in a different simulation of a Sagittarius-like encounter by
+\citet{2013MNRAS.429..159G}.
 
@@
 \citep{2014MNRAS.445..581H,2018MNRAS.480..800H} use the zoom-in technique
-\citep[e.g.,][]{1993ApJ...412..455K,2014MNRAS.437.1894O} to model the formation
-of a small group of galaxies at high resolution in a full cosmological
-context. Feedback from supernovae, stellar winds, and radiation from massive
-stars is implemented at the scale of star forming regions following stellar
-population synthesis models, generating galactic winds self-consistently
-\citep{2015MNRAS.454.2691M, 2017MNRAS.470.4698A} while reproducing many
-observed galaxy properties, including stellar masses, star formation
-histories, metallicities, and morphologies and kinematics of thin and thick
-disks \citep{2014MNRAS.445..581H, 2016MNRAS.456.2140M, 2017MNRAS.467.2430M,
-2016ApJ...827L..23W, 2018MNRAS.481.4133G, 2018MNRAS.480..800H}.
-
-For this work,
-we focus on the three Milky Way-mass zoom-ins considered in
+\citep[e.g.,][]{1993ApJ...412..455K,2014MNRAS.437.1894O} to model the
+formation of a small group of galaxies at high resolution in a full
+cosmological context. Feedback from supernovae, stellar winds, and radiation
+from massive stars is implemented at the scale of star forming regions
+following stellar population synthesis models, generating galactic winds
+self-consistently \citep{2015MNRAS.454.2691M, 2017MNRAS.470.4698A} while
+reproducing many observed galaxy properties, including stellar masses, star
+formation histories, metallicities, and morphologies and kinematics of thin
+and thick disks \citep{2014MNRAS.445..581H, 2016MNRAS.456.2140M,
+2017MNRAS.467.2430M, 2016ApJ...827L..23W, 2018MNRAS.481.4133G,
+2018MNRAS.480..800H}.
+
+For this work, we focus on the three Milky Way-mass zoom-ins considered in
 \citet{2018arXiv180610564S}, which were simulated as part of the
@@
 properties with observations of the Milky Way \citep{2016ApJ...827L..23W,
-2018MNRAS.481.4133G}. The 
-$\z = 0$ snapshots\footnote{In this work, to avoid
+2018MNRAS.481.4133G}. The $\z = 0$ snapshots\footnote{In this work, to avoid
 confusion with the vertical height $z$, we refer to cosmological redshift as
-$\z$.} of these three simulations, 
-The snapshots of these three simulations at cosmological redshift $\z = 0$,
-named \mi{}, \mf{}, and \mm{},
-are publicly available alongside associated mock \textit{Gaia} DR2
-catalogues generated from them.\footnote{\url{http://ananke.hub.yt}}
+$\z$.} of these three simulations, The snapshots of these three simulations at
+cosmological redshift $\z = 0$, named \mi{}, \mf{}, and \mm{}, are publicly
+available alongside associated mock \textit{Gaia} DR2 catalogues generated
+from them.\footnote{\url{http://ananke.hub.yt}}
 
@@
 \tablecaption{Stellar and gas disk scale heights of the Milky Way and the
-FIRE galaxies considered in this work (described in Section~\ref{ssec:cosmozoom}). For comparison, we also give the median
-softening lengths for the FIRE galaxies, computed for cold gas ($T<1000\,\text{K}$) and stars with $\abs{R-R_0} < 0.5\,\kpc$ and $\abs{z} < 1\,\kpc$.
-We have assumed that $R_0=8.2\,\kpc$.\label{tab:scale_height}}
+FIRE galaxies considered in this work (described in
+Section~\ref{ssec:cosmozoom}). For comparison, we also give the median
+softening lengths for the FIRE galaxies, computed for cold gas
+($T<1000\,\text{K}$) and stars with $\abs{R-R_0} < 0.5\,\kpc$ and $\abs{z} <
+1\,\kpc$. We have assumed that $R_0=8.2\,\kpc$.\label{tab:scale_height}}
 \tablehead{\colhead{galaxy} & \colhead{\makecell{cold\tablenotemark{a} gas disk \\ scale height}} &
@@
 gas particles of mass $\sim7000$ to $20,000\,\Msun$, and star particles of
-mass $\sim5000$ to $7000\,\Msun$, with the lower end coming from
-stellar evolution \citep{2018arXiv180610564S}. Softening lengths for dark
-matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
+mass $\sim5000$ to $7000\,\Msun$, with the lower end coming from stellar
+evolution \citep{2018arXiv180610564S}. Softening lengths for dark matter and
+star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
 respectively.\footnote{This is $2.8$ times the often-quoted
@@
 clear in the case of the gas; the quoted value for the Milky Way comes from
-studies of high-mass star-forming regions). 
-The 
-midplanes defined by gas and stars can be tilted with respect to one another as well,
-precluding extending the precision of the gas midplane definition to the
-stellar component.
+studies of high-mass star-forming regions). The midplanes defined by gas and
+stars can be tilted with respect to one another as well, precluding extending
+the precision of the gas midplane definition to the stellar component.
 
 Cosmological simulations of Milky Way-mass galaxies are not perfect
-representations of the true Milky Way in other ways as well,
-as discussed in \citet{2018arXiv180610564S}.
-The failure of cosmological simulations to exactly reproduce the Milky
-Way is not necessarily due to limitations of the numerical
-model. Candidate Milky Way-like galaxies are chosen solely on their
-mass and isolation, for which there are a wide variety of possible galaxies with qualitatively different
-properties. 
-For example,
-the velocity structure of \mi{} is closer to M31's than the Milky
-Way's (S. Loebman et al., in preparation).
+representations of the true Milky Way in other ways as well, as discussed in
+\citet{2018arXiv180610564S}. The failure of cosmological simulations to
+exactly reproduce the Milky Way is not necessarily due to limitations of the
+numerical model. Candidate Milky Way-like galaxies are chosen solely on their
+mass and isolation, for which there are a wide variety of possible galaxies
+with qualitatively different properties. For example, the velocity structure
+of \mi{} is closer to M31's than the Milky Way's (S. Loebman et al., in
+preparation).
 
@@
 nearby companion like the Large Magellanic Cloud. One of the three we consider
-(\mf) does have an ongoing interaction with a satellite galaxy similar in mass to
-Sagittarius, which has punched through the Galactic disk outside the solar
+(\mf) does have an ongoing interaction with a satellite galaxy similar in mass
+to Sagittarius, which has punched through the Galactic disk outside the solar
 circle, leaving behind some of its stars and inducing warping in the disk.
@@
 
-For the Milky Way, the dark halo is modeled as a Hernquist
-sphere of mass $10^{12}\,\Msun$ and scale length of $52\,\kpc$ \citep{1990ApJ...356..359H}, the
-disk is
-modeled as an exponential disk with a scale radius of $3.5\,\kpc$, scale
-height $0.53\,\kpc$, and mass $6\times10^{10}\,\Msun$, and the bulge as a Hernquist sphere
-of mass $10^{10}\,\Msun$ and scale length $0.7\,\kpc$. The Sagittarius dwarf is
-modeled with two components: a dark matter Hernquist sphere of mass $8\times10^{10}\,\Msun$ and
-scale length $8\,\kpc$, and a stellar component modeled as a Hernquist
-sphere of mass $6.4\times10^8\,\Msun$ and scale length $0.85\,\kpc$. All
-components are realized with distributions of live
-N-body particles; the Milky Way and Sagittarius are each initialized
+For the Milky Way, the dark halo is modeled as a Hernquist sphere of mass
+$10^{12}\,\Msun$ and scale length of $52\,\kpc$ \citep{1990ApJ...356..359H},
+the disk is modeled as an exponential disk with a scale radius of $3.5\,\kpc$,
+scale height $0.53\,\kpc$, and mass $6\times10^{10}\,\Msun$, and the bulge as
+a Hernquist sphere of mass $10^{10}\,\Msun$ and scale length $0.7\,\kpc$. The
+Sagittarius dwarf is modeled with two components: a dark matter Hernquist
+sphere of mass $8\times10^{10}\,\Msun$ and scale length $8\,\kpc$, and a
+stellar component modeled as a Hernquist sphere of mass $6.4\times10^8\,\Msun$
+and scale length $0.85\,\kpc$. All components are realized with distributions
+of live N-body particles; the Milky Way and Sagittarius are each initialized
 to be in equilibrium in isolation.
 
-The mass resolution of the simulation is $2.6\times10^4$, $1.2\times10^4$, 
-and $1.0\times10^4\,\Msun$ for the dark matter, disk, and bulge components,
+The mass resolution of the simulation is $2.6\times10^4$, $1.2\times10^4$, and
+$1.0\times10^4\,\Msun$ for the dark matter, disk, and bulge components,
 respectively. For the disk and bulge components, a softening length of
 $30\,\pc$ is used whereas for the halo a softening length of $60\,\pc$ is
-used. For Sagittarius, the softening length for the dark matter and
-the stars is $60$ 
-and $40\,\pc$, respectively.
+used. For Sagittarius, the softening length for the dark matter and the stars
+is $60$ and $40\,\pc$, respectively.
 
@@
 \subsection{The Local Midplane} \label{ssec:local_midplane}
-Using the two sets of simulations, we determine the local midplane as a function of azimuth at the solar circle
-that an observer might measure if they were situated in each of these
-galaxies% 
-. Starting from the coordinate system
-described in the previous section, which is 
-aligned so that the
-$z$-coordinate is approximately perpendicular to the disk plane at the solar circle, we place our imaginary
-observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and
-vary the azimuth between $0<\phi<2\pi$. 
-At equally spaced values of 
-$\phi$ we then compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and
-height $1\,\kpc$ perpendicular to the fiducial disk and centered on it. We choose to use 50 bins in azimuth, sufficiently few that no cylinder shares stars with its neighbors.
-We then re-define the new
-midplane of the cylinder to be the median $z$, re-select stars, and iterate
-until the median $z$ value converges. We find that only $10$ iterations of
-this procedure are necessary for convergence. The resulting median $z$ is
-taken to be what our observer would measure as the local galactic midplane at
-each $\phi$.
+Using the two sets of simulations, we determine the local midplane as a
+function of azimuth at the solar circle that an observer might measure if they
+were situated in each of these galaxies. Starting from the coordinate system
+described in the previous section, which is aligned so that the $z$-coordinate
+is approximately perpendicular to the disk plane at the solar circle, we place
+our imaginary observer at $z=0$ and a galactocentric cylindrical radius of
+$8.2\,\kpc$ and vary the azimuth between $0<\phi<2\pi$. At equally spaced
+values of $\phi$ we then compute the median $z$ for stars within a cylinder of
+radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to the fiducial disk and
+centered on it. We choose to use 50 bins in azimuth, sufficiently few that no
+cylinder shares stars with its neighbors. We then re-define the new midplane
+of the cylinder to be the median $z$, re-select stars, and iterate until the
+median $z$ value converges. We find that only $10$ iterations of this
+procedure are necessary for convergence. The resulting median $z$ is taken to
+be what our observer would measure as the local galactic midplane at each
+$\phi$.
 
@@
 ($R_0 = 8.2\,\kpc$) for the three FIRE galaxies \mi{}, \mf{}, and \mm{} (left,
-center, and right panels) as a function of azimuthal angle, at cosmological redshift $\z =0$. The local midplane is determined at
-a position $\phi$ by taking the median height of all stars within
-$R=0.5\,\kpc$ and $z=1\,\kpc$ (in cylindrical coordinates). 
-In order to allow for the possibility that the fiducial
-galactocentric coordinate system is incorrect, we subtract the best fit sine
-curve from each panel. We then bootstrap resample $1000$ times to determine
-$1\,\sigma $ error bars, which we report as dashed lines.}
+center, and right panels) as a function of azimuthal angle, at cosmological
+redshift $\z =0$. The local midplane is determined at a position $\phi$ by
+taking the median height of all stars within $R=0.5\,\kpc$ and $z=1\,\kpc$ (in
+cylindrical coordinates). In order to allow for the possibility that the
+fiducial galactocentric coordinate system is incorrect, we subtract the best
+fit sine curve from each panel. We then bootstrap resample $1000$ times to
+determine $1\,\sigma $ error bars, which we report as dashed lines.}
 \label{fig:midplane}
@@
 \caption{The local midplane determined at the fiducial solar circle
-($R_0 = 8.2\,\kpc$) for four different time steps from a live N-body simulation of a
-Sagittarius encounter with the Milky Way \citep{2018MNRAS.481..286L}. As
-before, we have subtracted the best fit sine
-curve to account
-for inaccuracies in the galactocentric coordinate system. Error bars are
-calculated as in Figure~\ref{fig:midplane}. The upper panels
+($R_0 = 8.2\,\kpc$) for four different time steps from a live N-body
+simulation of a Sagittarius encounter with the Milky Way
+\citep{2018MNRAS.481..286L}. As before, we have subtracted the best fit sine
+curve to account for inaccuracies in the galactocentric coordinate system.
+Error bars are calculated as in Figure~\ref{fig:midplane}. The upper panels
 show the midplane as a function of azimuth before the first encounter near the
-solar circle at $t=2.0\,\Gyr$ and $t=4.0\,\Gyr$, with an encounter happening close to the solar circle near $t=6.0\,\Gyr$.
-The fact that the $t=6.9\,\Gyr$ panel, which shows the midplane variation after some relaxation,
-looks qualitatively similar to the panels from the FIRE simulations
+solar circle at $t=2.0\,\Gyr$ and $t=4.0\,\Gyr$, with an encounter happening
+close to the solar circle near $t=6.0\,\Gyr$. The fact that the $t=6.9\,\Gyr$
+panel, which shows the midplane variation after some relaxation, looks
+qualitatively similar to the panels from the FIRE simulations
 (Figure~\ref{fig:midplane}) is evidence that midplane variations are
@@
 To account for the effect of particle noise, we bootstrap resample stars
-within a cylinder of height $2\,\kpc$ and the same radius $1000$
-times and determine the $1\,\sigma$ error bars by repeating the midplane
-determination with that reselection.
+within a cylinder of height $2\,\kpc$ and the same radius $1000$ times and
+determine the $1\,\sigma$ error bars by repeating the midplane determination
+with that reselection.
 
@@
 \eeq 
-from the midplane as a
-function of azimuth to account for an overall tilt of the midplane (a
-simplified version of the strategy described in
-\citealt{2019ApJ...871..145A}). 
-For simulations (\mi{}, \mf{}, \mm{}), the best fit values are $A =
-(-170, 45, 8.8)\,\pc$, $B = (38, -5.0, 1.8)\,\text{deg}$,
-and $C = (-69, 19, -18)\,\pc$.
-For the assumed solar radius of $8.2\,\kpc$, we can approximate the
-angle offset 
-$\Delta \theta$ 
-for the $z$-axis from the values of 
-$A$. We find for the same simulations $\Delta \theta = (1.15, 0.31, 0.062)\,\text{deg}$.
-These angle offsets are consistent with the
-values given in \citet{2018arXiv180610564S} for the difference between the
-$z$-axis as defined by the gas and stars.
+from the midplane as a function of azimuth to account for an overall tilt of
+the midplane (a simplified version of the strategy described in
+\citealt{2019ApJ...871..145A}). For simulations (\mi{}, \mf{}, \mm{}), the
+best fit values are $A = (-170, 45, 8.8)\,\pc$, $B = (38, -5.0,
+1.8)\,\text{deg}$, and $C = (-69, 19, -18)\,\pc$. For the assumed solar radius
+of $8.2\,\kpc$, we can approximate the angle offset $\Delta \theta$ for the
+$z$-axis from the values of $A$. We find for the same simulations $\Delta
+\theta = (1.15, 0.31, 0.062)\,\text{deg}$. These angle offsets are consistent
+with the values given in \citet{2018arXiv180610564S} for the difference
+between the $z$-axis as defined by the gas and stars.
 
 Figure~\ref{fig:midplane} shows the relative $z$ location of the inferred
-midplane an observer would determine as a function of azimuth for
-each galaxy, using their local solar neighborhood (the cylinder defined
-above). The
+midplane an observer would determine as a function of azimuth for each galaxy,
+using their local solar neighborhood (the cylinder defined above). The
 $1\,\sigma$ error from the bootstrap procedure is shown as the dashed-line
-error bars. The middle $90\%$ of midplane values across the solar circle spans $(190, 160, 84)\,\pc$ for these simulations.
-In two of the three cases
-the midplane therefore varies by more than $\pm 100\,\pc$ depending on the
-azimuth along the solar circle; in the third (\mm{}, which has the thinnest
-``thin disk'' of stars, but the largest stellar mass) the variation is closer
-to~$\pm 50\,\pc$.
+error bars. The middle $90\%$ of midplane values across the solar circle spans
+$(190, 160, 84)\,\pc$ for these simulations. In two of the three cases the
+midplane therefore varies by more than $\pm 100\,\pc$ depending on the azimuth
+along the solar circle; in the third (\mm{}, which has the thinnest ``thin
+disk'' of stars, but the largest stellar mass) the variation is closer to~$\pm
+50\,\pc$.
 
@@
 Sagittarius encounter \citep{2018MNRAS.481..286L}. Again we have subtracted a
-best fit curve of the form given in Equation~(\ref{eq:fit-curve}), with the values at times
-$(2.0, 4.0, 6.0, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -210,
--390)\,\pc$, $B
-= (0.074, 0.039, -36,
--57)\,\text{deg}$, and $C = (2.6,-7.8, -65, -53)\,\pc$.
-The middle $90\%$ of midplane values across the solar circle spans 
-$(49, 62, 140, 120)\,\pc$.
+best fit curve of the form given in Equation~(\ref{eq:fit-curve}), with the
+values at times $(2.0, 4.0, 6.0, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -210,
+-390)\,\pc$, $B = (0.074, 0.039, -36, -57)\,\text{deg}$, and $C = (2.6,-7.8,
+-65, -53)\,\pc$. The middle $90\%$ of midplane values across the solar circle
+spans $(49, 62, 140, 120)\,\pc$.
 
@@
 midplane variations seen by \citet{2013MNRAS.429..159G}. However, they only
-saw significant variations in their Heavy but not their Light Sagittarius model (virial masses of $10^{11}\,\Msun$ and
-$\sim3\times10^{10}\,\Msun$, respectively). The model we used (L2 from
-\citet{2018MNRAS.481..286L}) has a virial mass of $6\times10^{10}\,\Msun$, intermediate between their two models.
+saw significant variations in their Heavy but not their Light Sagittarius
+model (virial masses of $10^{11}\,\Msun$ and $\sim3\times10^{10}\,\Msun$,
+respectively). The model we used (L2 from \citet{2018MNRAS.481..286L}) has a
+virial mass of $6\times10^{10}\,\Msun$, intermediate between their two models.
 
@@
 lower left panel, we see a strong $m=2$ mode develop, consistent with the
-$R=8\,\kpc$ panel of Figure~17 in \citet{2018MNRAS.481..286L} ($m=0$ and $m=1$ modes are stronger, but these
-are removed in our sine-curve subtraction). The lower right
-panel, which shows the galaxy at $t=6.9\,\Gyr$ when some relaxation has occured, is
-qualitatively similar to the midplane variations we saw in the FIRE
-simulations (Figure~\ref{fig:midplane}), evidence that they are at least
-partially driven by mergers.
+$R=8\,\kpc$ panel of Figure~17 in \citet{2018MNRAS.481..286L} ($m=0$ and $m=1$
+modes are stronger, but these are removed in our sine-curve subtraction). The
+lower right panel, which shows the galaxy at $t=6.9\,\Gyr$ when some
+relaxation has occured, is qualitatively similar to the midplane variations we
+saw in the FIRE simulations (Figure~\ref{fig:midplane}), evidence that they
+are at least partially driven by mergers.
 
@@
 \section{Discussion} \label{sec:discussion}
-We have used high-resolution simulations to illustrate 
-why
-we expect the local midplane defined by stellar density to vary with azimuth by up to $\pm
+We have used high-resolution simulations to illustrate why we expect the local
+midplane defined by stellar density to vary with azimuth by up to $\pm
 100\,\pc$ as a natural consequence of the non-axisymmetry of the Galactic disk
 at small scales. While this is not in itself surprising or new, we have also
-argued
-that the discrepancy between our local midplane and that of distant
+argued that the discrepancy between our local midplane and that of distant
 stars introduces a systematic error in the $z$ component when converting from
@@
 
-Systematic variations in $v_z$ and number density were first noted as asymmetries in the local
-velocity distribution towards the North and South Galactic Caps from the
-radial velocity surveys of the Sloan Digital Sky Surveys
-\citep{2012ApJ...750L..41W} and RAdial Velocity Experiment
+Systematic variations in $v_z$ and number density were first noted as
+asymmetries in the local velocity distribution towards the North and South
+Galactic Caps from the radial velocity surveys of the Sloan Digital Sky
+Surveys \citep{2012ApJ...750L..41W} and RAdial Velocity Experiment
 \citep{2013MNRAS.436..101W}. Subsequently, \citet{2013ApJ...777L...5C} pointed
 out suggestions of an oscillation in average vertical velocities of order
-$5\,\kms$ on roughly kiloparsec 
-scales looking toward the Galactic anticenter.
-
-Work
-by the \textit{Gaia} collaboration confirmed these preliminary results on the
-velocity and spatial scales of 
-oscillation with clear spatial maps made
+$5\,\kms$ on roughly kiloparsec scales looking toward the Galactic anticenter.
+
+Work by the \textit{Gaia} collaboration confirmed these preliminary results on
+the velocity and spatial scales of oscillation with clear spatial maps made
 using DR2 data of median $v_z$ over a significant Galactic volume
-\citep{2018A&A...616A..11G, 2019arXiv190209569F}, which can be explained with models of Sagittarius-like encounters \citep{2013MNRAS.429..159G,2018MNRAS.481..286L,2019MNRAS.485.3134L}.
-We also see in the FIRE
-simulation that the vertical velocity variation as a function of azimuth is
-$\sim5\textup{--}10\,\kms$ (Figure~\ref{fig:lsr_variations}), consistent with
-these observations.
-
-The vertical frequency of \thin{} and \thick{} are $\sim
-0.09\,\Myr^{-1}$ and $\sim0.06\,\Myr^{-1}$, respectively (Table~\ref{tab:orbits}). By dimensional
-analysis, and assuming a vertical velocity variation of $5\textup{--}10\,\kms$, we
-therefore expect the midplane offsets to be $\sim 57\textup{--}170\,\pc$. We
-stress that this is a rough calculation.
+\citep{2018A&A...616A..11G, 2019arXiv190209569F}, which can be explained with
+models of Sagittarius-like encounters
+\citep{2013MNRAS.429..159G,2018MNRAS.481..286L,2019MNRAS.485.3134L}. We also
+see in the FIRE simulation that the vertical velocity variation as a function
+of azimuth is $\sim5\textup{--}10\,\kms$ (Figure~\ref{fig:lsr_variations}),
+consistent with these observations.
+
+The vertical frequency of \thin{} and \thick{} are $\sim 0.09\,\Myr^{-1}$ and
+$\sim0.06\,\Myr^{-1}$, respectively (Table~\ref{tab:orbits}). By dimensional
+analysis, and assuming a vertical velocity variation of
+$5\textup{--}10\,\kms$, we therefore expect the midplane offsets to be $\sim
+57\textup{--}170\,\pc$. We stress that this is a rough calculation.
 
@@
 show that the midplane varies by $\sim10\degree$ at a distance of
-$\sim0.75\,\kpc$, corresponding to a physical vertical variation 
-of $\sim130\,\pc$.
+$\sim0.75\,\kpc$, corresponding to a physical vertical variation of
+$\sim130\,\pc$.
 
@@
 Uncertainties in measurements of the position and velocity of the Sun relative
-to the Galactic center can also contribute to systematic error in the actions, since converting from heliocentric to Galactocentric
-coordinates relies on these measurements. Therefore, errors in their values
-will induce a systematic offset in the Galactocentric phase-space position of
-any observed star.
-Considerable effort has been placed on each of these measurements, but uncertainties remain, and detailed modeling across the disk\,---\,particularly for dynamically cold stars\,---\,may have to take them into
-account.
-Here we briefly review the current measurements of the four relevant quantities, 
-their uncertainties, and the implications for the calculation of actions.
+to the Galactic center can also contribute to systematic error in the actions,
+since converting from heliocentric to Galactocentric coordinates relies on
+these measurements. Therefore, errors in their values will induce a systematic
+offset in the Galactocentric phase-space position of any observed star.
+Considerable effort has been placed on each of these measurements, but
+uncertainties remain, and detailed modeling across the disk\,---\,particularly
+for dynamically cold stars\,---\,may have to take them into account. Here we
+briefly review the current measurements of the four relevant quantities, their
+uncertainties, and the implications for the calculation of actions.
 
@@
 First, one must define the center of the Galaxy. This is usually taken to be
-the location of the central supermassive black hole, Sagittarius~A\textsuperscript{*}
-\citep[\sgra{}, e.g.,][]{2004ApJ...616..872R}. From stellar motions near
+the location of the central supermassive black hole,
+Sagittarius~A\textsuperscript{*} \citep[\sgra{},
+e.g.,][]{2004ApJ...616..872R}. From stellar motions near
 \sgra{}, the distance from the Sun to \sgra{}, $R_0$, can be
-precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}. A
-recent measurement using near-infrared interferometry places $R_0$ at $8.178 \pm 0.035$~kpc
-\citep{2019arXiv190405721A}, or a $0.4\%$ 
-uncertainty.
-
-However, the location of \sgra{} may not be equivalent to the location
-of the dynamical Galactic center, the
-point in three-dimensional space about which the stars in the solar
-neighborhood are orbiting. This assumption, although sensible and frequently
-made, has not yet been justified.
+precisely measured \citep{2009ApJ...692.1075G, 2018AA...615L..15G}. A recent
+measurement using near-infrared interferometry places $R_0$ at $8.178 \pm
+0.035$~kpc \citep{2019arXiv190405721A}, or a $0.4\%$ uncertainty.
+
+However, the location of \sgra{} may not be equivalent to the location of the
+dynamical Galactic center, the point in three-dimensional space about which
+the stars in the solar neighborhood are orbiting. This assumption, although
+sensible and frequently made, has not yet been justified.
 
 If the dynamical Galactic center is offset from \sgra{} by $100\,\pc$, only a
-$1.2\%$ difference, then this induces a $\sim15\%$ error in $J_R$ for the disk-like orbits we considered (see
-Section~\ref{ssec:quant}). The reason such a large error in $J_R$ can be
-generated by a small error in $R_0$ can be understood from the epicyclic
-approximation (Equation~\eqref{eq:Ji_err_mosttime}),
-which states that $\Delta J_R/J_R = 2\Delta R/A_R$. The
-fractional error in $J_R$ is related to the error in $R_0$ as a fraction of
-the \emph{radial amplitude} of the orbit, which is much smaller than $R_0$
-($\sim1.2\,\kpc$ for \thin{} and \thick{}). This
-also implies the very precise $0.4\%$ measurement of $R_0$ still translates to
-a $\sim6\%$ uncertainty in $J_R$.
+$1.2\%$ difference, then this induces a $\sim15\%$ error in $J_R$ for the
+disk-like orbits we considered (see Section~\ref{ssec:quant}). The reason such
+a large error in $J_R$ can be generated by a small error in $R_0$ can be
+understood from the epicyclic approximation
+(Equation~\eqref{eq:Ji_err_mosttime}), which states that $\Delta J_R/J_R =
+2\Delta R/A_R$. The fractional error in $J_R$ is related to the error in $R_0$
+as a fraction of the \emph{radial amplitude} of the orbit, which is much
+smaller than $R_0$ ($\sim1.2\,\kpc$ for \thin{} and \thick{}). This also
+implies the very precise $0.4\%$ measurement of $R_0$ still translates to a
+$\sim6\%$ uncertainty in $J_R$.
 
@@
 
-We did not consider in this work the effect of the angular position of
-the dynamical Galactic center being offset from
-\sgra{}.
+We did not consider in this work the effect of the angular position of the
+dynamical Galactic center being offset from \sgra{}.
 
@@
 corresponding to a vertical height of $\sim22\,\pc$. This corresponds to a
-$37\%$ 
-error in $J_z$ for \thin{} and a 5\% error for \thick{}.
+$37\%$ error in $J_z$ for \thin{} and a 5\% error for \thick{}.
 
@@
 \citet{2016ARAA..54..529B}. The discrepancy between gas-based and
-stellar-based determinations of the solar height is small, and thus
-only likely to dominate over intrinsic midplane variations on small
-scales, but will be relevant for detailed modeling of young and
-kinematically cold stars. For instance, it will induce a $\sim10\%$
-error in $J_z$ for an orbit with $z_{\text{max}}\sim100\,\pc$.
+stellar-based determinations of the solar height is small, and thus only
+likely to dominate over intrinsic midplane variations on small scales, but
+will be relevant for detailed modeling of young and kinematically cold stars.
+For instance, it will induce a $\sim10\%$ error in $J_z$ for an orbit with
+$z_{\text{max}}\sim100\,\pc$.
 
@@
 components of the LSR are usually taken from \citet{2010MNRAS.403.1829S}.
-Their uncertainties should also lead to systematic errors in the actions, as given in Equations~\eqref{eq:induced_Jphi}--\eqref{eq:induced_Jz}.
-For example, the
-value of the circular velocity is taken to be $\sim 220\,\kms$
-\citep[e.g.,][]{2012ApJ...759..131B} with roughly $10\%$ uncertainty. We expect this to translate to at least a $10\%$ systematic error in $J_{\phi}$.
+Their uncertainties should also lead to systematic errors in the actions, as
+given in Equations~\eqref{eq:induced_Jphi}--\eqref{eq:induced_Jz}. For
+example, the value of the circular velocity is taken to be $\sim 220\,\kms$
+\citep[e.g.,][]{2012ApJ...759..131B} with roughly $10\%$ uncertainty. We
+expect this to translate to at least a $10\%$ systematic error in $J_{\phi}$.
 
@@
 We have mainly been concerned with actions, since they provide a convenient
-way to quantify different types of orbits. However, all of our conclusions also apply to studies that 
-simply rely on orbit integrations, since the two
+way to quantify different types of orbits. However, all of our conclusions
+also apply to studies that simply rely on orbit integrations, since the two
 are equivalent. For instance, computing orbital properties of open or globular
@@
 Orbit integrations of nearby systems over short time periods
-\citep[e.g.,][]{2014MNRAS.445.2169M, 2018A&A...616A..37B} are unlikely to be impacted. It should
-also be unimportant for halo applications, e.g., in modeling of
-stellar streams \citep[e.g.,][]{2014ApJ...795...95B} or the substructure
-potentially responsible for the gap in GD1 \citep{2018arXiv181103631B}.
+\citep[e.g.,][]{2014MNRAS.445.2169M, 2018A&A...616A..37B} are unlikely to be
+impacted. It should also be unimportant for halo applications, e.g., in
+modeling of stellar streams \citep[e.g.,][]{2014ApJ...795...95B} or the
+substructure potentially responsible for the gap in GD1
+\citep{2018arXiv181103631B}.
 
 \section{Conclusions}\label{sec:conclusion}
-Determining the orbital properties of stars is important for
-understanding the structure and evolution of the Galaxy. Actions have been argued to be 
-excellent orbit labels. If the Galaxy can be well
-approximated as axisymmetric and 6D phase space positions can be measured
-accurately and precisely enough, then the computed actions are invariant with
-orbital phase. However, we have shown that the fact that the Galactic midplane
-is not constant across the disk presents a significant complication to
-computed actions actually being invariant. Our main conclusions are:
+Determining the orbital properties of stars is important for understanding the
+structure and evolution of the Galaxy. Actions have been argued to be
+excellent orbit labels. If the Galaxy can be well approximated as axisymmetric
+and 6D phase space positions can be measured accurately and precisely enough,
+then the computed actions are invariant with orbital phase. However, we have
+shown that the fact that the Galactic midplane is not constant across the disk
+presents a significant complication to computed actions actually being
+invariant. Our main conclusions are:
 
@@
 \item Inaccuracy in the Galactocentric coordinate
-system induces orbital phase 
-dependence in the actions calculated from the observed
-positions and velocities of stars
+system induces orbital phase dependence in the actions calculated from the
+observed positions and velocities of stars
 (Figures~\ref{fig:cartoon}~and~\ref{fig:one_orbit_wrong_ref}). Since stars'
@@
 \item Inaccuracy in the midplane location most severely affects computation of
-the vertical action $J_z$. A midplane offset of $\sim15\,\pc$ 
-for a typical thin
-disk orbit results in a $25\%$ error in $J_z$, 
-and even for a thick disk orbit a $120\,\pc$ offset will result
-in the same size error.
-The fractional error is
-significantly less for halo orbits.
+the vertical action $J_z$. A midplane offset of $\sim15\,\pc$ for a typical
+thin disk orbit results in a $25\%$ error in $J_z$, and even for a thick disk
+orbit a $120\,\pc$ offset will result in the same size error. The fractional
+error is significantly less for halo orbits.
 
 \item The distribution of systematic errors in the actions induced by a
-coordinate system offset is highly
-non-Gaussian. The distribution is bimodal
+coordinate system offset is highly non-Gaussian. The distribution is bimodal
 with \emph{neither mode at null}. As a result, error propagation of coordinate
@@
 susceptible to this type of systematic error, since the assumption that our
-local Galactic midplane is the global Galactic midplane is not true a
-priori. A violation of this assumption (by, e.g., intrinsic midplane
-variations) leads to a systematic error in $z$ which generates the 
-large errors in actions summarized above.
+local Galactic midplane is the global Galactic midplane is not true a priori.
+A violation of this assumption (by, e.g., intrinsic midplane variations) leads
+to a systematic error in $z$ which generates the large errors in actions
+summarized above.
 
 \item 
-We show that such midplane variation is likely by measuring
-the local galactic
+We show that such midplane variation is likely by measuring the local galactic
 midplane along the solar circle in three different high-resolution, zoom-in
-simulations of Milky Way mass galaxies from the FIRE collaboration, as well as a controlled simulation of the interaction of the Milky Way with Sagittarius. We found
-that the midplane varies as a function of azimuth at the solar circle by
-$60\textup{--}185\,\pc$ in these simulations.
+simulations of Milky Way mass galaxies from the FIRE collaboration, as well as
+a controlled simulation of the interaction of the Milky Way with Sagittarius.
+We found that the midplane varies as a function of azimuth at the solar circle
+by $60\textup{--}185\,\pc$ in these simulations.
 
 \item Assuming a vertical velocity variation of the Milky Way of
-$\sim5\textup{--}10\,\kms$, consistent with recent results from
-\textit{Gaia} 
-and our results from the FIRE
-simulations (Figure~\ref{fig:lsr_variations}), we estimated that the
-corresponding midplane offsets are $\sim60\textup{--}170\,\pc$ by dimensional
-analysis 
-using 
-the vertical frequencies of disk-like orbits. 
-This range of values is consistent with the variations seen in the simulations.
-Similar offsets are seen in three-dimensional dust maps.
+$\sim5\textup{--}10\,\kms$, consistent with recent results from \textit{Gaia}
+and our results from the FIRE simulations (Figure~\ref{fig:lsr_variations}),
+we estimated that the corresponding midplane offsets are
+$\sim60\textup{--}170\,\pc$ by dimensional analysis using the vertical
+frequencies of disk-like orbits. This range of values is consistent with the
+variations seen in the simulations. Similar offsets are seen in
+three-dimensional dust maps.
 
 \item Inaccuracies in the parameters of the currently adopted
-Galactocentric coordinate system are likely important for some
-applications. In particular, it is imperative to test the assumption that the
-dynamical Galactic center is colocated with \sgra{}. We discuss how to
-do this in Section~\ref{ssec:coord_off}.
+Galactocentric coordinate system are likely important for some applications.
+In particular, it is imperative to test the assumption that the dynamical
+Galactic center is colocated with \sgra{}. We discuss how to do this in
+Section~\ref{ssec:coord_off}.
 
 \item This work underlines the importance of combining chemistry and dynamics.
-Since chemical tagging \citep{2002ARA&A..40..487F} is not subject to the same systematic errors discussed in this work, it should be used to confirm dynamical associations and to offset the effect of these systematic errors on the action-space distribution function.
+Since chemical tagging \citep{2002ARA&A..40..487F} is not subject to the same
+systematic errors discussed in this work, it should be used to confirm
+dynamical associations and to offset the effect of these systematic errors on
+the action-space distribution function.
 
@@
 in the computation of actions under the assumption of a global axisymmetric
-potential. 
-Current observations from \textit{Gaia} should soon permit a measurement of the
-real azimuthal dependence of the midplane location. For some applications,
-such as those using actions as labels to group stars on similar orbits, using
-such a measurement to shift stars to a consistent midplane height as a
-function of azimuth before using a global axisymmetric approximation to the
-potential may be sufficient, although this ignores the \emph{dynamical}
-implications of shifts in the midplane height (which result from fluctuations
-in the local density). However, for other applications, such as the study of
-action diffusion,
-a more extensive perturbative approach is likely needed. We plan to explore the
-mitigation of these effects in future work.
+potential. Current observations from \textit{Gaia} should soon permit a
+measurement of the real azimuthal dependence of the midplane location. For
+some applications, such as those using actions as labels to group stars on
+similar orbits, using such a measurement to shift stars to a consistent
+midplane height as a function of azimuth before using a global axisymmetric
+approximation to the potential may be sufficient, although this ignores the
+\emph{dynamical} implications of shifts in the midplane height (which result
+from fluctuations in the local density). However, for other applications, such
+as the study of action diffusion, a more extensive perturbative approach is
+likely needed. We plan to explore the mitigation of these effects in future
+work.
 
@@
 D'Onghia, Benoit Famaey, and Adrian Price-Whelan for helpful discussions. A.B.
-would like to thank Todd Phillips for helpful discussions. This work uses data hosted by the Flatiron Institute's FIRE data hub. This project was
-developed in part at the 2019 Santa Barbara Gaia Sprint, hosted by the Kavli
-Institute for Theoretical Physics (KITP) at the University of California,
-Santa Barbara. This research was supported in part at KITP by the
-Heising-Simons Foundation and the National Science Foundation (NSF) under
-Grant No. PHY-1748958. This work used the Extreme Science and Engineering
-Discovery Environment (XSEDE), which is supported by NSF Grant No.
-OCI-1053575. A.B. was supported in part by the Roy \& Diana Vagelos Program in
-the Molecular Life Sciences and the Roy \& Diana Vagelos Challenge Award.
-K.V.J.'s contributions were supported in part by the NSF under Grant No.
-AST-1614743. M.-M.M.L. was partly supported by the NSF under Grant No.
-AST-1815461.
-
-\software{This work made use of the following software: \texttt{astropy} \citep{astropy:2013,astropy:2018}, \texttt{gala}
-\citep{gala, gala:zenodo}, \texttt{matplotlib} \citep{Hunter:2007}, \texttt{numpy} \citep{numpy:2011}, \texttt{scipy}
-\citep{scipy:2001}, and \texttt{tqdm} \citep{casper_o_da_costa_luis_2019_2800317}.}
+would like to thank Todd Phillips for helpful discussions. This work uses data
+hosted by the Flatiron Institute's FIRE data hub. This project was developed
+in part at the 2019 Santa Barbara Gaia Sprint, hosted by the Kavli Institute
+for Theoretical Physics (KITP) at the University of California, Santa Barbara.
+This research was supported in part at KITP by the Heising-Simons Foundation
+and the National Science Foundation (NSF) under Grant No. PHY-1748958. This
+work used the Extreme Science and Engineering Discovery Environment (XSEDE),
+which is supported by NSF Grant No. OCI-1053575. A.B. was supported in part by
+the Roy \& Diana Vagelos Program in the Molecular Life Sciences and the Roy \&
+Diana Vagelos Challenge Award. K.V.J.'s contributions were supported in part
+by the NSF under Grant No. AST-1614743. M.-M.M.L. was partly supported by the
+NSF under Grant No. AST-1815461.
+
+\software{This work made use of the following software: \texttt{astropy}
+\citep{astropy:2013,astropy:2018}, \texttt{gala} \citep{gala, gala:zenodo},
+\texttt{matplotlib} \citep{Hunter:2007}, \texttt{numpy} \citep{numpy:2011},
+\texttt{scipy} \citep{scipy:2001}, and \texttt{tqdm}
+\citep{casper_o_da_costa_luis_2019_2800317}.}
 
@@
 \caption{The three orbits presented in Table~\ref{tab:orbits} and considered
-throughout the work. We plot \thin{}, \thick{}, and \halo{} in the
-left, center, and right columns, respectively. The upper row shows a plot of
+throughout the work. We plot \thin{}, \thick{}, and \halo{} in the left,
+center, and right columns, respectively. The upper row shows a plot of
 $x$~vs.~$y$ while the lower row shows $R$~vs.~$z$.}
@@
 distribution of $J_{\phi}$ for an offset in $x$. We plot the distributions for
-\thick{} (upper panels) and \thin{} (lower panels). We find that the $J_R$ distribution induced by an offset in $x$ more
-closely resembles a Gaussian distribution, while the $J_R$ distribution induced by
-an offset in $z$ and the $J_{\phi}$ distribution induced by an offset
-in $x$ are both similar to the $J_z$ distribution induced by an
-offset in $z$ (see Figure~\ref{fig:Jz_hist}).
+\thick{} (upper panels) and \thin{} (lower panels). We find that the $J_R$
+distribution induced by an offset in $x$ more closely resembles a Gaussian
+distribution, while the $J_R$ distribution induced by an offset in $z$ and the
+$J_{\phi}$ distribution induced by an offset in $x$ are both similar to the
+$J_z$ distribution induced by an offset in $z$ (see Figure~\ref{fig:Jz_hist}).
 
@@
 We consider the variations of the LSR as a function of azimuth at the fiducial
-solar circle ($R_{0} = 8.2\,\kpc$) in Figure~\ref{fig:lsr_variations}. At
-each azimuth, $\phi$, we take the median velocity in cylindrical coordinates
-of all stars within $200\,\pc$ of the position, following
+solar circle ($R_{0} = 8.2\,\kpc$) in Figure~\ref{fig:lsr_variations}. At each
+azimuth, $\phi$, we take the median velocity in cylindrical coordinates of all
+stars within $200\,\pc$ of the position, following
 \citet{2018arXiv180610564S}. No best-fit subtraction was performed as in
@@
 \end{center}
-\caption{The LSR as a function of azimuth at the
-fiducial solar circle ($R_0 = 8.2\,\kpc$). No best-fit subtraction is
-performed here as we did in the case of the midplane
-(Section~\ref{ssec:local_midplane}). Variations in $v_z$ are on the order of
-$\sim5\textup{--}10\,\kms$.}
+\caption{The LSR as a function of azimuth at the fiducial solar circle ($R_0 =
+8.2\,\kpc$). No best-fit subtraction is performed here as we did in the case
+of the midplane (Section~\ref{ssec:local_midplane}). Variations in $v_z$ are
+on the order of $\sim5\textup{--}10\,\kms$.}
 \label{fig:lsr_variations}

### 2019-05-21 paper/ms.tex
Update keywords order
Add link to repo

@@
 
-\keywords{Milky Way disk -- Milky Way evolution -- Milky Way dynamics --
+\keywords{Milky Way disk -- Milky Way dynamics -- Milky Way evolution --
 Galaxy structure -- Orbits -- Stellar dynamics}
@@
 
+\subsection*{Supplementary Material}
+All code used in this work is available at:
+\url{https://github.com/gusbeane/actions_systematic}.
+
 \acknowledgments

### 2019-05-21 paper/ms.tex
Shorten abstract

@@
 the Galactic midplane of $\sim15\,\pc$ for a thin disk-like orbit or $\sim
-120\,\pc$ for a thick disk-like orbit is sufficient to induce a $25\%$
-systematic error in the vertical action $J_z$. In FIRE simulations of Milky
-Way-mass galaxies, these variations are on the order of $\sim100\,\pc$ at the
-solar circle. From observations of the mean vertical velocity variation of
+120\,\pc$ for a thick disk-like orbit induces a $25\%$ systematic error in the
+vertical action $J_z$. In FIRE simulations of Milky Way-mass galaxies, these
+variations are on the order of $\sim100\,\pc$ at the solar circle. From
+observations of the mean vertical velocity variation of
 $\sim5\textup{--}10\,\kms$ with radius, we estimate that the Milky Way
-midplane variations are of order $\sim60\textup{--}170\,\pc$\,---\,also
-consistent with recent three-dimensional dust maps. Action calculations and
-orbit integrations which assume the global and local midplanes are identical
-are likely to include the induced errors we report, depending on the volume
-considered. Variation in the local standard of rest or distance to the
-Galactic center causes similar issues. The variation of the midplane must be
-taken into account when performing dynamical analysis across the large regions
-of the disk accessible to \textit{Gaia} and future missions.
+midplane variations are $\sim60\textup{--}170\,\pc$, consistent with
+three-dimensional dust maps. Action calculations and orbit integrations, which
+assume the global and local midplanes are identical, are likely to include
+this induced error, depending on the volume considered. Variation in the local
+standard of rest or distance to the Galactic center causes similar issues. The
+variation of the midplane must be taken into account when performing dynamical
+analysis across the large regions of the disk accessible to \textit{Gaia} and
+future missions.
 

### 2019-06-25 paper/ms.tex
Fix couple typos suggested by ref

@@
 induced in each action for an offset in the $z$-component. In the lower panels
-we consider offsets in the $x$ component (i.e. the solar radius). The left,
+we consider offsets in the $x$ component (i.e. the solar radius). \textbf{The left,
 center, and right columns show the fractional spread in the values computed
-for $J_z$, $J_{\phi}$, and $J_R$, respectively.
+for $J_R$, $J_{\phi}$, and $J_z$, respectively.}
 
@@
 results in a $25\%$ deviation in $J_z$, while a $120\,\pc$ offset results in
-the same deviation for \thick{}. We find that \halo{} is relatively resistant
-to errors in the midplane, with only $\sim15\%$ error in $J_z$ out to an
-offset of $500\,\pc$.
+the same deviation for \thick{}. We find that \halo{} is relatively
+\textbf{insensitive} to errors in the midplane, with only $\sim15\%$ error in
+$J_z$ out to an offset of $500\,\pc$.
 

### 2019-06-30 paper/ms.tex
Update title and short title
Tweak abstract

@@
 \shortauthors{Beane et al.}
-\shorttitle{The Galactic Midplane is Not a Plane}
+\shorttitle{Implications of Disk Oscillations}
 
@@
 
-\title{The Galactic Midplane Is Not a Plane: Implications for Dynamical
-Analysis with \textit{Gaia} Data and Beyond}
+\title{The Implications of Disk Oscillations for Dynamical Analysis in the
+\textit{Gaia} Era}
 
@@
 subtle but important systematic error that is induced in the actions as a
-consequence of local midplane variations expected for the Milky Way. This
+consequence of \textbf{disk oscillations} or local midplane variations expected for the Milky Way. This
 error is difficult to model because it is non-Gaussian and bimodal, with

### 2019-06-30 paper/ms.tex
Update discussion of pre-Gaia results

@@
 
-The true Galactic potential is not strictly axisymmetric. This has been known
-for some time \citep[e.g.,][]{1957AJ.....62...93K,2009MNRAS.396L..56M,
-2012ApJ...750L..41W}, but with the vast improvement in the quality of
-phase-space measurements due to \textit{Gaia} the assumption of axisymmetry is
-increasingly inadequate \citep[e.g.,][]{2018Natur.561..360A,
-2019MNRAS.485.3134L}. Even if this assumption were close enough for many
-purposes, the parameters used in axisymmetric models of the Galactic potential
-may be inadequately constrained by current observations.
+\textbf{The true Galactic potential is not strictly axisymmetric, beyond even
+the typically-quoted gas clouds, spiral arms, and bar(s). The presence of disk
+oscillations has been known for some time, observed in 1957 for H~\textsc{i}
+in the outer disk \citep{1957AJ.....62...93K}. The presence of a North-South
+asymmetry in stellar number density and velocity in the solar neighborhood has
+been found in SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the
+LAMOST survey \citep{2013ApJ...777L...5C}, and RAVE data
+\citep{2013MNRAS.436..101W}. Beyond the solar neighborhood, low-latitude
+overdensities have been observed, e.g., the Monoceros Ring
+\citep{2002ApJ...569..245N, 2003MNRAS.340L..21I, 2003ApJ...594L.119C,
+2014ApJ...791....9S, 2016ApJ...825..140M}, an overdensity in the direction of
+the Triangulum and Andromeda galaxies \citep{2007ApJ...668L.123M,
+2014ApJ...793...62S, 2015ApJ...801..105X, 2015MNRAS.452..676P} and A13
+\citep{2010ApJ...722..750S, 2017ApJ...844...74L}, and in the direction of the
+Galactic Center \citep{2014Natur.509..342F}. There is evidence that these
+stellar populations originated from the disk \citep{2018ApJ...854...47S,
+2018Natur.555..334B}. See, e.g., \citet{2018MNRAS.481..286L} for a discussion
+of these observations.}
+
+\textbf{With the vast improvement in the quality of phase-space measurements
+due to \textit{Gaia} the assumption of axisymmetry is increasingly inadequate
+\citep[e.g.,][]{2018Natur.561..360A, 2019MNRAS.485.3134L}. The presence of the
+phase-space spiral itself is evidence of disk oscillations
+\citep{2018Natur.561..360A}, but the fact that its strength changes across the
+disk is further evidence \citep{2019MNRAS.486.1167B}. Even if the axisymmetric
+assumption were close enough for many purposes, the parameters used in
+axisymmetric models of the Galactic potential may be inadequately constrained
+by current observations.}
 

### 2019-06-30 paper/ms.tex
Add calc of total energy error

@@
 
+\textbf{Because the energies in the radial and vertical coordinates are related to the
+radial and vertical actions by a constant
+(Equation~\eqref{eq:JR_Jz_epi_energy}), the fractional error on these two
+energies in the as in the actions. The energy in the azimuthal coordinate can
+be readily computed, and we therefore conclude that for the given offsets the
+error on the total computed energy is
+\beq\label{eq:error_total_energy}
+\frac{\Delta E}{E} = \frac{\Delta J_R}{J_R} + \frac{\Delta J_{\phi}}{J_{\phi}}
+                     + \frac{\Delta J_z}{J_z}\text{.}
+\eeq}
+
 In the remainder of this section, we compare our analytic estimates of the

### 2019-07-23 paper/ms.tex
Update title

@@
 
-\title{The Implications of Disk Oscillations for Dynamical Analysis in the
-\textit{Gaia} Era}
+\title{The Implications of Local Fluctuations in the Galactic Midplane for
+Dynamical Analysis in the \textit{Gaia} Era}
 
@@
 subtle but important systematic error that is induced in the actions as a
-consequence of \textbf{disk oscillations} or local midplane variations expected for the Milky Way. This
+consequence of local midplane variations expected for the Milky Way. This
 error is difficult to model because it is non-Gaussian and bimodal, with

### 2019-07-23 paper/ms.tex
Add paragraph on single measurement point

@@
 correct midplane) should be phase independent.
+
+\textbf{We would like to emphasize that in reality one only ever measures the
+position and velocity of a star at a single instance in time, and not over any
+significant portion of its orbit. However, if more than one star on the same
+orbit can be identified, then it might be possible to constrain the disk
+potential in analogy to how stellar streams are used to constrain the Galactic
+potential \citep[e.g.][]{1999ApJ...512L.109J, 2001ApJ...551..294I,
+2004ApJ...610L..97H, 2015ApJ...801...98S, 2015ApJ...803...80K,
+2019MNRAS.483.1427L}. However, identifying the remnants of a disrupted open
+cluster (the inner disk equivalent of a halo stellar stream) is a difficult
+challenge. One would want to use the same constants of motion that we discuss
+are suscept to significant errors. On the other hand, one could use chemical
+tagging to identify a disrupted open cluster \citep{2002ARA&A..40..487F}, but
+it may be difficult due to the large number of stars that are chemically
+indistinguishable but are presumably born in different clusters
+\citep{2018ApJ...853..198N}. An alternative technique that would deliver some
+of the benefits of finding a disk stream may lie in measuring the
+accelerations of stars \citep{2018arXiv181207581S}.}
 

### 2019-07-23 paper/ms.tex
Update ack

@@
 \acknowledgments
-We would like to thank Megan Bedell, Robert A. Benjamin, Tobias Buck, Elena
-D'Onghia, Benoit Famaey, and Adrian Price-Whelan for helpful discussions. A.B.
-would like to thank Todd Phillips for helpful discussions. This work uses data
-hosted by the Flatiron Institute's FIRE data hub. This project was developed
-in part at the 2019 Santa Barbara Gaia Sprint, hosted by the Kavli Institute
-for Theoretical Physics (KITP) at the University of California, Santa Barbara.
-This research was supported in part at KITP by the Heising-Simons Foundation
-and the National Science Foundation (NSF) under Grant No. PHY-1748958. This
-work used the Extreme Science and Engineering Discovery Environment (XSEDE),
-which is supported by NSF Grant No. OCI-1053575. A.B. was supported in part by
-the Roy \& Diana Vagelos Program in the Molecular Life Sciences and the Roy \&
-Diana Vagelos Challenge Award. K.V.J.'s contributions were supported in part
-by the NSF under Grant No. AST-1614743. M.-M.M.L. was partly supported by the
-NSF under Grant No. AST-1815461.
+We thank the anonymous referee for providing helpful comments. We would like
+to thank Megan Bedell, Robert A. Benjamin, Maria Bergemann, Joss
+Bland-Hawthorn, Tobias Buck, Elena D'Onghia, Benoit Famaey, and Adrian
+Price-Whelan for helpful discussions. A.B. would like to thank Todd Phillips
+for helpful discussions. This work uses data hosted by the Flatiron
+Institute's FIRE data hub. This project was developed in part at the 2019
+Santa Barbara Gaia Sprint, hosted by the Kavli Institute for Theoretical
+Physics (KITP) at the University of California, Santa Barbara. This research
+was supported in part at KITP by the Heising-Simons Foundation and the
+National Science Foundation (NSF) under Grant No. PHY-1748958. This work used
+the Extreme Science and Engineering Discovery Environment (XSEDE), which is
+supported by NSF Grant No. OCI-1053575. A.B. was supported in part by the Roy
+\& Diana Vagelos Program in the Molecular Life Sciences and the Roy \& Diana
+Vagelos Challenge Award. K.V.J.'s contributions were supported in part by the
+NSF under Grant No. AST-1614743. M.-M.M.L. was partly supported by the NSF
+under Grant No. AST-1815461.
 

### 2019-07-26 paper/ms.tex
fix formatting problem
rest of my comments

@@
 the typically-quoted gas clouds, spiral arms, and bar(s). The presence of disk
-oscillations has been known for some time, observed in 1957 for H~\textsc{i}
-in the outer disk \citep{1957AJ.....62...93K}. The presence of a North-South
+oscillations has been known
+%mm for some time, observed in 1957 for
+% H~% \textsc{i} in the outer disk \citep{1957AJ.....62...93K}.
+    since its observation in H~{sc i} by \citet{1957AJ.....62...93K}.
+The presence of a North-South
 asymmetry in stellar number density and velocity in the solar neighborhood has
-been found in SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the
+been found in
+% mm [should these multiple acronyms be expanded?]
+SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the
 LAMOST survey \citep{2013ApJ...777L...5C}, and RAVE data
@@
 
-Small-scale variations in the density of gas and stars cause the local
+% mm Small-scale
+     Such
+variations in the density of gas and stars cause the local
 midplane position to vary as a function of radius and azimuth. Stars far from
@@
 
-\textbf{We would like to emphasize that in reality one only ever measures the
-position and velocity of a star at a single instance in time, and not over any
-significant portion of its orbit. However, if more than one star on the same
-orbit can be identified, then it might be possible to constrain the disk
-potential in analogy to how stellar streams are used to constrain the Galactic
-potential \citep[e.g.][]{1999ApJ...512L.109J, 2001ApJ...551..294I,
-2004ApJ...610L..97H, 2015ApJ...801...98S, 2015ApJ...803...80K,
-2019MNRAS.483.1427L}. However, identifying the remnants of a disrupted open
-cluster (the inner disk equivalent of a halo stellar stream) is a difficult
-challenge. One would want to use the same constants of motion that we discuss
-are suscept to significant errors. On the other hand, one could use chemical
-tagging to identify a disrupted open cluster \citep{2002ARA&A..40..487F}, but
-it may be difficult due to the large number of stars that are chemically
-indistinguishable but are presumably born in different clusters
-\citep{2018ApJ...853..198N}. An alternative technique that would deliver some
-of the benefits of finding a disk stream may lie in measuring the
-accelerations of stars \citep{2018arXiv181207581S}.}
-
+%mm [moved para up as it directly follows previous para]
 This example uses an offset in $z$, but analagous effects occur from offsets
@@
 potential was pointed out by \citet{2015A&A...584A.120B}.
+
+\textbf{We would like to emphasize that in reality one only ever measures the
+position and velocity of a star at a single instance in time, and not over any
+significant portion of its orbit. However, if more than one star on the same
+orbit
+%mm can be identified, then it might be possible 
+     could be identified, then it would be possible
+to constrain the disk potential in analogy to how stellar streams are
+used to constrain the Galactic  potential
+\citep[e.g.][]{1999ApJ...512L.109J, 2001ApJ...551..294I, 
+2004ApJ...610L..97H, 2015ApJ...801...98S, 2015ApJ...803...80K,
+2019MNRAS.483.1427L}. However, identifying the remnants of a disrupted open
+cluster (the inner disk equivalent of a halo stellar stream) is a difficult
+challenge. One
+%mm would want to use the same constants of motion that we discuss are
+%suscept to significant errors. On the other hand, one could use
+    possibility would be to use the actions, but as we discuss in this
+    work they are subject to substantial errors.  Another possibility would
+    be to use
+chemical tagging to identify a disrupted open cluster \citep{2002ARA&A..40..487F}, but
+%mm it
+    this
+may be difficult due to the large number of stars that are chemically
+indistinguishable
+%mm but are
+    though
+presumably born in different clusters
+\citep{2018ApJ...853..198N}. An alternative technique that would deliver some
+of the benefits of finding a disk stream may lie in measuring the
+accelerations of stars \citep{2018arXiv181207581S}.}
+
+%mm This example uses an offset in $z$, but analagous effects occur from offsets
+%in other coordinates, such as the distance to the Galactic center or the LSR.
+%A similar effect in which actions gain time-dependence due to a time-varying
+%potential was pointed out by \citet{2015A&A...584A.120B}.
 
@@
 radial and vertical actions by a constant
-(Equation~\eqref{eq:JR_Jz_epi_energy}), the fractional error on these two
-energies in the as in the actions. The energy in the azimuthal coordinate can
-be readily computed, and we therefore conclude that for the given offsets the
-error on the total computed energy is
+%mm (Equation~\eqref{eq:JR_Jz_epi_energy}),
+    (Eq.~[\ref{eq:JR_Jz_epi_energy}])
+the fractional error
+%mm on
+     in
+these two
+energies
+%mm in the as in the actions. The
+%[I am guessing as to what this means, so **** check ****]
+      evolves in the same way as the actions.  We can see this because the
+ energy in the azimuthal coordinate can
+be readily computed,
+%mm [maybe give it here to ensure that everyone is on the same page?]
+%mm and we therefore conclude that
+    so
+for the given offsets the error
+%mm on
+    in
+the total computed energy is
 \beq\label{eq:error_total_energy}
@@
 best fit curve of the form given in Equation~(\ref{eq:fit-curve}), with the
-values at times $(2.0, 4.0, 6.0, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -210,
--390)\,\pc$, $B = (0.074, 0.039, -36, -57)\,\text{deg}$, and $C = (2.6,-7.8,
--65, -53)\,\pc$. The middle $90\%$ of midplane values across the solar circle
-spans $(49, 62, 140, 120)\,\pc$.
+values at times
+% mm  [line breaks are needed]
+%$(2.0, 4.0, 6.0, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -210,
+%-390)\,\pc$, $B = (0.074, 0.039, -36, -57)\,\text{deg}$, and $C = (2.6,-7.8,
+%-65, -53)\,\pc$. The middle $90\%$ of midplane values across the solar circle
+%spans mm $(49, 62, 140, 120)\,\pc$.
+    (2.0, 4.0, 6.0, 6.9)~Gyr being $A$ = (9.5, 2.5, -210, -390)~pc,
+    $B$ = (0.074, 0.039, -36, -57)~deg, and $C$ = (2.6,-7.8, -65,
+    -53)~pc. The middle $90\%$ of midplane values across the solar
+    circle spans (49, 62, 140, 120)~pc.
 

### 2019-07-26 paper/ms.tex
Update ackowledgements
Clean up/fix discussion of energy error

@@
 
-\textbf{Because the energies in the radial and vertical coordinates are related to the
-radial and vertical actions by a constant
-%mm (Equation~\eqref{eq:JR_Jz_epi_energy}),
-    (Eq.~[\ref{eq:JR_Jz_epi_energy}])
-the fractional error
-%mm on
-     in
-these two
-energies
-%mm in the as in the actions. The
-%[I am guessing as to what this means, so **** check ****]
-      evolves in the same way as the actions.  We can see this because the
- energy in the azimuthal coordinate can
-be readily computed,
-%mm [maybe give it here to ensure that everyone is on the same page?]
-%mm and we therefore conclude that
-    so
-for the given offsets the error
-%mm on
-    in
-the total computed energy is
+\textbf{Because the energies in the radial and vertical coordinates are
+related to the radial and vertical actions by a constant
+(Equation~\eqref{eq:JR_Jz_epi}), the fractional error in these two energies is
+the same as the actions. The energy in the azimuthal coordinate can be readily
+computed ($E_{\phi} = \frac{1}{2}v_{\phi}^2$), and we therefore conclude that
+the error induced in the total energy is:
 \beq\label{eq:error_total_energy}
-\frac{\Delta E}{E} = \frac{\Delta J_R}{J_R} + \frac{\Delta J_{\phi}}{J_{\phi}}
+\frac{\Delta E}{E} = \frac{\Delta J_R}{J_R} + 2\frac{\Delta J_{\phi}}{J_{\phi}}
                      + \frac{\Delta J_z}{J_z}\text{.}
@@
 for helpful discussions. This work uses data hosted by the Flatiron
-Institute's FIRE data hub. This project was developed in part at the 2019
+Institute's FIRE data hub. The Flatiron Institute is supported by the Simons Foundation. This project was developed in part at the 2019
 Santa Barbara Gaia Sprint, hosted by the Kavli Institute for Theoretical

### 2019-07-26 paper/ms.tex
Add some loose references

@@
 SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the
-LAMOST survey \citep{2013ApJ...777L...5C}, and RAVE data
+LAMOST survey \citep{2013ApJ...777L...5C, 2015ApJ...801..105X}, and RAVE data
 \citep{2013MNRAS.436..101W}. Beyond the solar neighborhood, low-latitude
@@
 Galactic Center \citep{2014Natur.509..342F}. There is evidence that these
-stellar populations originated from the disk \citep{2018ApJ...854...47S,
+stellar populations originated from the disk \citep{2015MNRAS.452..676P, 2018ApJ...854...47S,
 2018Natur.555..334B}. See, e.g., \citet{2018MNRAS.481..286L} for a discussion
@@
 \citep{2018Natur.561..360A}, but the fact that its strength changes across the
-disk is further evidence \citep{2019MNRAS.486.1167B}. Even if the axisymmetric
+disk is further evidence \citep{2019MNRAS.485.3134L, 2019MNRAS.486.1167B}. Even if the axisymmetric
 assumption were close enough for many purposes, the parameters used in

### 2019-07-26 paper/ms.tex
Clean up comments throughout

@@
 oscillations has been known
-%mm for some time, observed in 1957 for
-% H~% \textsc{i} in the outer disk \citep{1957AJ.....62...93K}.
-    since its observation in H~{sc i} by \citet{1957AJ.....62...93K}.
+    since its observation in H~{\sc i} by \citet{1957AJ.....62...93K}.
 The presence of a North-South
@@
 been found in
-% mm [should these multiple acronyms be expanded?]
 SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the
@@
 
-% mm Small-scale
 Such
@@
 
-%mm [moved para up as it directly follows previous para]
 This example uses an offset in $z$, but analagous effects occur from offsets
@@
 orbit
-%mm can be identified, then it might be possible 
      could be identified, then it would be possible
@@
 challenge. One
-%mm would want to use the same constants of motion that we discuss are
-%suscept to significant errors. On the other hand, one could use
-    possibility would be to use the actions, but as we discuss in this
+    possibility would be to use the stellar actions or other conserved quantities, but as we discuss in this
     work they are subject to substantial errors.  Another possibility would
@@
 chemical tagging to identify a disrupted open cluster \citep{2002ARA&A..40..487F}, but
-%mm it
     this
@@
 indistinguishable
-%mm but are
     though
@@
 accelerations of stars \citep{2018arXiv181207581S}.}
-
-%mm This example uses an offset in $z$, but analagous effects occur from offsets
-%in other coordinates, such as the distance to the Galactic center or the LSR.
-%A similar effect in which actions gain time-dependence due to a time-varying
-%potential was pointed out by \citet{2015A&A...584A.120B}.
 
@@
 values at times
-% mm  [line breaks are needed]
-%$(2.0, 4.0, 6.0, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -210,
-%-390)\,\pc$, $B = (0.074, 0.039, -36, -57)\,\text{deg}$, and $C = (2.6,-7.8,
-%-65, -53)\,\pc$. The middle $90\%$ of midplane values across the solar circle
-%spans mm $(49, 62, 140, 120)\,\pc$.
     (2.0, 4.0, 6.0, 6.9)~Gyr being $A$ = (9.5, 2.5, -210, -390)~pc,

### 2019-07-26 paper/ms.tex
Add changes from Robyn

@@
 assumption were close enough for many purposes, the parameters used in
-axisymmetric models of the Galactic potential may be inadequately constrained
-by current observations.}
+axisymmetric models of the Galactic potential may be biased by the non-axisymmetries in the disk.}
 
@@
 
-\textbf{We would like to emphasize that in reality one only ever measures the
-position and velocity of a star at a single instance in time, and not over any
-significant portion of its orbit. However, if more than one star on the same
-orbit
-     could be identified, then it would be possible
+\textbf{In reality one only ever measures the
+position and velocity of a star at a single instance in time.
+However, there are at least two possibilities for going beyond a single temporal snapshot of the phase-space distribution.
+If more than one star on the same
+or similar orbit
+     could be identified (e.g., by identifying the remnants of a disrupted open cluster), then it would be possible
 to constrain the disk potential in analogy to how stellar streams are
@@
 2004ApJ...610L..97H, 2015ApJ...801...98S, 2015ApJ...803...80K,
-2019MNRAS.483.1427L}. However, identifying the remnants of a disrupted open
-cluster (the inner disk equivalent of a halo stellar stream) is a difficult
-challenge. One
-    possibility would be to use the stellar actions or other conserved quantities, but as we discuss in this
+2019MNRAS.483.1427L}. This would be challenging since, in principle, one would want to use the stellar actions or other conserved quantities, but as we discuss in this
     work they are subject to substantial errors.  Another possibility would
@@
 presumably born in different clusters
-\citep{2018ApJ...853..198N}. An alternative technique that would deliver some
-of the benefits of finding a disk stream may lie in measuring the
-accelerations of stars \citep{2018arXiv181207581S}.}
+\citep{2018ApJ...853..198N}. A second way to improve on a single temporal snapshot is to measure directly the
+accelerations of stars \citep{2018arXiv181207581S}.
+While this would allow a direct measurement of the aliasing effect in Figure~\ref{fig:cartoon}, the observational challenges will likely limit this type of measurement to a small number of local stars for the near future.}
 

### 2019-07-26 paper/ms.tex
Update short title
Clean up energy argument

@@
 \shortauthors{Beane et al.}
-\shorttitle{Implications of Disk Oscillations}
+\shorttitle{Galactic Dynamical Analysis and Midplane Fluctuations}
 
@@
 (Equation~\eqref{eq:JR_Jz_epi}), the fractional error in these two energies is
-the same as the actions. The energy in the azimuthal coordinate can be readily
-computed ($E_{\phi} = \frac{1}{2}v_{\phi}^2$), and we therefore conclude that
-the error induced in the total energy is:
+the same as the actions.\footnote{This ignores any potential error in the epicyclic and vertical frequencies.} The energy in the azimuthal coordinate can be readily
+computed ($E_{\phi} = \frac{1}{2}v_{\phi}^2$).
+Using standard propagation of errors, we see that $\Delta E_{\phi}/E_{\phi}$ and $\Delta J_{\phi}/J_{\phi}$ (Equation~\eqref{eq:induced_Jphi}) differ by a $\Delta R/R_g$ term.
+Since $R_g\approx 8\,\kpc$, we expect this term to be negligible and we therefore conclude that:
 \beq\label{eq:error_total_energy}
-\frac{\Delta E}{E} = \frac{\Delta J_R}{J_R} + 2\frac{\Delta J_{\phi}}{J_{\phi}}
+\frac{\Delta E}{E} \approx \frac{\Delta J_R}{J_R} + 2\frac{\Delta J_{\phi}}{J_{\phi}}
                      + \frac{\Delta J_z}{J_z}\text{.}

### 2019-08-08 paper/ms.tex
Fix line indents

@@
 the typically-quoted gas clouds, spiral arms, and bar(s). The presence of disk
-oscillations has been known
-    since its observation in H~{\sc i} by \citet{1957AJ.....62...93K}.
-The presence of a North-South
-asymmetry in stellar number density and velocity in the solar neighborhood has
-been found in
-SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the
-LAMOST survey \citep{2013ApJ...777L...5C, 2015ApJ...801..105X}, and RAVE data
+oscillations has been known since its observation in H~{\sc i} by
+\citet{1957AJ.....62...93K}. The presence of a North-South asymmetry in
+stellar number density and velocity in the solar neighborhood has been found
+in SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the LAMOST survey
+\citep{2013ApJ...777L...5C, 2015ApJ...801..105X}, and RAVE data
 \citep{2013MNRAS.436..101W}. Beyond the solar neighborhood, low-latitude
@@
 Galactic Center \citep{2014Natur.509..342F}. There is evidence that these
-stellar populations originated from the disk \citep{2015MNRAS.452..676P, 2018ApJ...854...47S,
-2018Natur.555..334B}. See, e.g., \citet{2018MNRAS.481..286L} for a discussion
-of these observations.}
+stellar populations originated from the disk \citep{2015MNRAS.452..676P,
+2018ApJ...854...47S, 2018Natur.555..334B}. See, e.g.,
+\citet{2018MNRAS.481..286L} for a discussion of these observations.}
 
@@
 \citep{2018Natur.561..360A}, but the fact that its strength changes across the
-disk is further evidence \citep{2019MNRAS.485.3134L, 2019MNRAS.486.1167B}. Even if the axisymmetric
-assumption were close enough for many purposes, the parameters used in
-axisymmetric models of the Galactic potential may be biased by the non-axisymmetries in the disk.}
-
-Such
-variations in the density of gas and stars cause the local
-midplane position to vary as a function of radius and azimuth. Stars far from
-the Sun have a local midplane that differs from our local midplane
-extrapolated onto their position. Converting positions and velocities of more
-distant stars from a heliocentric to a Galactocentric coordinate system thus
-introduces a systematic bias in the $z$ coordinate. We show that this bias
-induces non-Gaussian errors in the actions computed for these stars. The
-further from the solar neighborhood the target star is, the more likely the
-mismatch will result in large systematic uncertainty, especially in the
-vertical action $J_z$. A similar argument applies to any remaining uncertainty
-in measurements of the Galactocentric radius of the Sun, and to variations in
-the LSR.
+disk is further evidence \citep{2019MNRAS.485.3134L, 2019MNRAS.486.1167B}.
+Even if the axisymmetric assumption were close enough for many purposes, the
+parameters used in axisymmetric models of the Galactic potential may be biased
+by the non-axisymmetries in the disk.}
+
+Such variations in the density of gas and stars cause the local midplane
+position to vary as a function of radius and azimuth. Stars far from the Sun
+have a local midplane that differs from our local midplane extrapolated onto
+their position. Converting positions and velocities of more distant stars from
+a heliocentric to a Galactocentric coordinate system thus introduces a
+systematic bias in the $z$ coordinate. We show that this bias induces
+non-Gaussian errors in the actions computed for these stars. The further from
+the solar neighborhood the target star is, the more likely the mismatch will
+result in large systematic uncertainty, especially in the vertical action
+$J_z$. A similar argument applies to any remaining uncertainty in measurements
+of the Galactocentric radius of the Sun, and to variations in the LSR.
 
@@
 
-\textbf{In reality one only ever measures the
-position and velocity of a star at a single instance in time.
-However, there are at least two possibilities for going beyond a single temporal snapshot of the phase-space distribution.
-If more than one star on the same
-or similar orbit
-     could be identified (e.g., by identifying the remnants of a disrupted open cluster), then it would be possible
-to constrain the disk potential in analogy to how stellar streams are
-used to constrain the Galactic  potential
-\citep[e.g.][]{1999ApJ...512L.109J, 2001ApJ...551..294I, 
-2004ApJ...610L..97H, 2015ApJ...801...98S, 2015ApJ...803...80K,
-2019MNRAS.483.1427L}. This would be challenging since, in principle, one would want to use the stellar actions or other conserved quantities, but as we discuss in this
-    work they are subject to substantial errors.  Another possibility would
-    be to use
-chemical tagging to identify a disrupted open cluster \citep{2002ARA&A..40..487F}, but
-    this
-may be difficult due to the large number of stars that are chemically
-indistinguishable
-    though
-presumably born in different clusters
-\citep{2018ApJ...853..198N}. A second way to improve on a single temporal snapshot is to measure directly the
-accelerations of stars \citep{2018arXiv181207581S}.
-While this would allow a direct measurement of the aliasing effect in Figure~\ref{fig:cartoon}, the observational challenges will likely limit this type of measurement to a small number of local stars for the near future.}
+\textbf{In reality one only ever measures the position and velocity of a star
+at a single instance in time. However, there are at least two possibilities
+for going beyond a single temporal snapshot of the phase-space distribution.
+If more than one star on the same or similar orbit could be identified (e.g.,
+by identifying the remnants of a disrupted open cluster), then it would be
+possible to constrain the disk potential in analogy to how stellar streams are
+used to constrain the Galactic  potential \citep[e.g.][]{1999ApJ...512L.109J,
+2001ApJ...551..294I, 2004ApJ...610L..97H, 2015ApJ...801...98S,
+2015ApJ...803...80K, 2019MNRAS.483.1427L}. This would be challenging since, in
+principle, one would want to use the stellar actions or other conserved
+quantities, but as we discuss in this work they are subject to substantial
+errors.  Another possibility would be to use chemical tagging to identify a
+disrupted open cluster \citep{2002ARA&A..40..487F}, but this may be difficult
+due to the large number of stars that are chemically indistinguishable though
+presumably born in different clusters \citep{2018ApJ...853..198N}. A second
+way to improve on a single temporal snapshot is to measure directly the
+accelerations of stars \citep{2018arXiv181207581S}. While this would allow a
+direct measurement of the aliasing effect in Figure~\ref{fig:cartoon}, the
+observational challenges will likely limit this type of measurement to a small
+number of local stars for the near future.}
 
@@
 (Equation~\eqref{eq:JR_Jz_epi}), the fractional error in these two energies is
-the same as the actions.\footnote{This ignores any potential error in the epicyclic and vertical frequencies.} The energy in the azimuthal coordinate can be readily
-computed ($E_{\phi} = \frac{1}{2}v_{\phi}^2$).
-Using standard propagation of errors, we see that $\Delta E_{\phi}/E_{\phi}$ and $\Delta J_{\phi}/J_{\phi}$ (Equation~\eqref{eq:induced_Jphi}) differ by a $\Delta R/R_g$ term.
-Since $R_g\approx 8\,\kpc$, we expect this term to be negligible and we therefore conclude that:
+the same as the actions.\footnote{This ignores any potential error in the
+epicyclic and vertical frequencies.} The energy in the azimuthal coordinate
+can be readily computed ($E_{\phi} = \frac{1}{2}v_{\phi}^2$). Using standard
+propagation of errors, we see that $\Delta E_{\phi}/E_{\phi}$ and $\Delta
+J_{\phi}/J_{\phi}$ (Equation~\eqref{eq:induced_Jphi}) differ by a $\Delta
+R/R_g$ term. Since $R_g\approx 8\,\kpc$, we expect this term to be negligible
+and we therefore conclude that:
 \beq\label{eq:error_total_energy}
@@
 induced in each action for an offset in the $z$-component. In the lower panels
-we consider offsets in the $x$ component (i.e. the solar radius). \textbf{The left,
-center, and right columns show the fractional spread in the values computed
-for $J_R$, $J_{\phi}$, and $J_z$, respectively.}
+we consider offsets in the $x$ component (i.e. the solar radius). \textbf{The
+left, center, and right columns show the fractional spread in the values
+computed for $J_R$, $J_{\phi}$, and $J_z$, respectively.}
 
@@
 for helpful discussions. This work uses data hosted by the Flatiron
-Institute's FIRE data hub. The Flatiron Institute is supported by the Simons Foundation. This project was developed in part at the 2019
-Santa Barbara Gaia Sprint, hosted by the Kavli Institute for Theoretical
-Physics (KITP) at the University of California, Santa Barbara. This research
-was supported in part at KITP by the Heising-Simons Foundation and the
-National Science Foundation (NSF) under Grant No. PHY-1748958. This work used
-the Extreme Science and Engineering Discovery Environment (XSEDE), which is
-supported by NSF Grant No. OCI-1053575. A.B. was supported in part by the Roy
-\& Diana Vagelos Program in the Molecular Life Sciences and the Roy \& Diana
-Vagelos Challenge Award. K.V.J.'s contributions were supported in part by the
-NSF under Grant No. AST-1614743. M.-M.M.L. was partly supported by the NSF
-under Grant No. AST-1815461.
+Institute's FIRE data hub. The Flatiron Institute is supported by the Simons
+Foundation. This project was developed in part at the 2019 Santa Barbara Gaia
+Sprint, hosted by the Kavli Institute for Theoretical Physics (KITP) at the
+University of California, Santa Barbara. This research was supported in part
+at KITP by the Heising-Simons Foundation and the National Science Foundation
+(NSF) under Grant No. PHY-1748958. This work used the Extreme Science and
+Engineering Discovery Environment (XSEDE), which is supported by NSF Grant No.
+OCI-1053575. A.B. was supported in part by the Roy \& Diana Vagelos Program in
+the Molecular Life Sciences and the Roy \& Diana Vagelos Challenge Award.
+K.V.J.'s contributions were supported in part by the NSF under Grant No.
+AST-1614743. M.-M.M.L. was partly supported by the NSF under Grant No.
+AST-1815461.
 

### 2019-08-08 paper/ms.tex
Add zenodo doi

@@
 \subsection*{Supplementary Material}
-All code used in this work is available at:
-\url{https://github.com/gusbeane/actions_systematic}.
+All code used in this work is available at
+\url{https://github.com/gusbeane/actions_systematic} and archived at Zenodo
+[doi:10.5281/zenodo.3364082].
 

### 2019-08-22 paper/ms.tex
Fix previous commit
Update KVJ's NSF grant
Remove bold from ref comments

@@
 
-\textbf{The true Galactic potential is not strictly axisymmetric, beyond even
+The true Galactic potential is not strictly axisymmetric, beyond even
 the typically-quoted gas clouds, spiral arms, and bar(s). The presence of disk
@@
 2018ApJ...854...47S, 2018Natur.555..334B}. See, e.g.,
-\citet{2018MNRAS.481..286L} for a discussion of these observations.}
-
-\textbf{With the vast improvement in the quality of phase-space measurements
+\citet{2018MNRAS.481..286L} for a discussion of these observations.
+
+With the vast improvement in the quality of phase-space measurements
 due to \textit{Gaia} the assumption of axisymmetry is increasingly inadequate
@@
 parameters used in axisymmetric models of the Galactic potential may be biased
-by the non-axisymmetries in the disk.}
+by the non-axisymmetries in the disk.
 
@@
 
-\textbf{In reality one only ever measures the position and velocity of a star
+In reality one only ever measures the position and velocity of a star
 at a single instance in time. However, there are at least two possibilities
@@
 observational challenges will likely limit this type of measurement to a small
-number of local stars for the near future.}
+number of local stars for the near future.
 
@@
 
-\textbf{Because the energies in the radial and vertical coordinates are
+Because the energies in the radial and vertical coordinates are
 related to the radial and vertical actions by a constant
@@
                      + \frac{\Delta J_z}{J_z}\text{.}
-\eeq}
+\eeq
 
@@
 induced in each action for an offset in the $z$-component. In the lower panels
-we consider offsets in the $x$ component (i.e. the solar radius). \textbf{The
+we consider offsets in the $x$ component (i.e. the solar radius). The
 left, center, and right columns show the fractional spread in the values
-computed for $J_R$, $J_{\phi}$, and $J_z$, respectively.}
+computed for $J_R$, $J_{\phi}$, and $J_z$, respectively.
 
@@
 the same deviation for \thick{}. We find that \halo{} is relatively
-\textbf{insensitive} to errors in the midplane, with only $\sim15\%$ error in
+insensitive to errors in the midplane, with only $\sim15\%$ error in
 $J_z$ out to an offset of $500\,\pc$.
@@
 \url{https://github.com/gusbeane/actions_systematic} and archived at Zenodo
-[doi:10.5281/zenodo.3364082].
+[doi:\href{https://doi.org/10.5281/zenodo.3364082}{10.5281/zenodo.3364082}].
 
@@
 K.V.J.'s contributions were supported in part by the NSF under Grant No.
-AST-1614743. M.-M.M.L. was partly supported by the NSF under Grant No.
+AST-1715582. M.-M.M.L. was partly supported by the NSF under Grant No.
 AST-1815461.
